### 2015-03-17 markdown/shade.Rmd
started markdown of WTC gm data

---
title: "Sun v Shade leaves (WTCIII)"
author: "Court Campany"
date: "March 17, 2015"
output: html_document
---

```
This set of figures attempts to show that sun and shade leaves were different on multiple levels.  
```

```{r, results='asis', echo=FALSE}
library(knitr)
opts_knit$set(root.dir = '../')
```

ACI curves were measured at saturating light for both sun and shade leaves.

```{r, results='asis', echo=FALSE, message=FALSE}
library(xtable)

aci_leaf <- read.csv("calculated_data/tdl_aci.csv")

# aci_table <- xtable(aci_leaf)
# digits(aci_table)[c(3,5:6)] <- 3
# digits(aci_table)[c(4)] <- 4
# print(aci_table, type='html', include.rownames=F,hline.after=NULL)
# print(aci_table,  type='html', include.rownames=F, comment=FALSE)

 kable(aci_leaf, row.names=NA, digits=2,format = "html")

```

```
Here are some predicted ACI curves for sun/shade with temperature treatements.  Both Vcmax and Jmax are reduced in shade leaves with little effect of treatment. Aci curves were only taken once towards the end of the experiment on well watered trees.
```
```{r,  echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}

source("functions and packages/functions.R")
source("master_scripts/plot_objects.R")
aci_leaf <- read.csv("calculated_data/tdl_aci.csv")

#simulated curves
  library(plantecophys)

  #simulate ACi curves for each leaf, ambient and elevated T
  sunAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[3,3], Jmax=aci_leaf[3,4],PPFD=1408)
  sunET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[4,3], Jmax=aci_leaf[4,4], PPFD=1408)
  shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[1,3], Jmax=aci_leaf[1,4], PPFD=375)
  shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[2,3], Jmax=aci_leaf[2,4],PPFD=375)

#plot
plot(sunAT_sim$Ci, sunAT_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab=cilab, ylab="", type="l", lwd=2)
  points(sunET_sim$Ci, sunET_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
  points(shaAT_sim$Ci, shaAT_sim$ALEAF, col="yellowgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2)
  points(shaET_sim$Ci, shaET_sim$ALEAF, col="yellowgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
  legend("bottomright", leglab2, lty=c(1,2,1,2), lwd=2,col=colaci, pt.bg=col4,inset = 0.03)
  title(ylab=satlab, mgp=ypos)

```

```
This figure shows the relationship between leaf N on an area basis and carbon isotope dicrimination between leaf types.
```
```{r, echo=FALSE,fig.align='center', message=FALSE}
source("master_scripts/plot_objects.R")
leaftraits <- read.csv("calculated_data/leaftraits_summary.csv")
##this has droughtleaves

palette(c("yellowgreen", "green4"))

plot(c13.mean ~ leafN_area.mean, data=leaftraits, pch=16, col=as.factor(leaf), ylim=c(-35,-25), 
                 xlim=c(1,3.5), cex=1.25, xlab=narealab, ylab=c13lab)
legend("bottomright", leaflab, pch=16, col=c("yellowgreen", "green4"),pt.cex=1.5,inset = 0.03)

```

Next I show the light environment in which gas exchange measurements for sun and shade leaves were measured.  This data was generated with a handheld par sensor for each individual leaf level and verified with ceptometer readings within at the height of the measured leaf. There was on average a ~75 reduction in PPFD from sun to shave leaves.

The second figure shows leaf mass area for each gas exchange campaign.  Here there

```{r,echo=FALSE,fig.align='center',message=FALSE, warning=FALSE}
source("functions and packages/functions.R")
source("master_scripts/plot_objects.R")
library(doBy)
####PPFD-----------------------------------------------------------------------------------------------
par <- read.csv("raw data/par.csv")
treatments <- read.csv("raw data/chamber_trt.csv")

#format function
par<- parformat(par)

par_leaf <- subset(par, ID !="shade-high")
Morder <- c("Oct", "Dec", "Jan", "Feb", "Mar", "Apr")

#data format for bar
parbar <- subset(par_leaf, select = c("par", "month", "leaf_type"))
parbar$month <- factor(parbar$month, levels = Morder)

bar(par, c(leaf_type, month), parbar, col=c("yellowgreen", "green4"), xlab="", ylab="", ylim=c(0, 2000), 
    half.errbar=FALSE)
title(ylab=parlab, mgp=ypos)

###LMA--------------------------------------------------------------------------------------------------
lma <- read.csv("calculated_data/leaf_chemistry.csv")

#data format for bar
lmabar <- subset(lma, select = c("lma", "Month", "leaf"))
  lmabar$Month <- factor(lmabar$Month, levels = Morder)

bar(lma, c(leaf, Month), lmabar, col=c("yellowgreen", "green4"), xlab="", ylab="", ylim=c(0, 185), half.errbar=FALSE)
title(ylab=lmalab, mgp=ypos)

##table summ by treatment
  lma_agg <- summaryBy(lma~ leaf+temp+drydown, data=lma, FUN=c(mean,se))
  lma_agg2 <- summaryBy(lma~ leaf+temp, data=lma, FUN=c(mean,se))

  lma_table <- xtable(lma_agg2)

```

## Mesophyll Conducatnce

```{r,echo=FALSE,fig.align='center',message=FALSE, warning=FALSE}
source("master_scripts/plot_objects.R")
source("functions and packages/functions.R")
library(doBy)

gmes <- read.csv("calculated_data/gmes_WTC.csv")

###for analysis first subset well watered and drought treatments
gm_drought <- gmes[gmes$drydown == "drought",]

gm_water <- gmes[gmes$drydown == "control",]
#write.csv(gm_water, "calculated_data/gmes_wellwatered.csv", row.names=FALSE)

palette(c("blue", "red"))

###drought----------------------------------------------------------------------------------------------------

##plots
plot(Photo~gm, data=gm_drought, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,25), xlim=c(0, .5))
  points(Photo~gm, data=gm_drought, subset=leaflight=="shade-low",pch=1, col=temp)
  legend("topright", templab, pch=16,inset = 0.03, col=palette())

###well watered------------------------------------------------------------------------------------------------

##sun-shade
plot(Photo~gm, data=gm_water, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,25), xlim=c(0,.5), 
     xlab=gmlab, ylab="")
points(Photo~gm, data=gm_water, subset=leaflight=="shade-low",pch=1, col=temp)
legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
title(ylab=satlab, mgp=ypos, cex=1.2)

##shade low-high
plot(Photo~gm, data=gm_water, subset=leaflight=="shade-high", pch=16, col=temp, ylim=c(0,25), xlim=c(0,.5), 
     xlab=gmlab, ylab="")
points(Photo~gm, data=gm_water, subset=leaflight=="shade-low",pch=1, col=temp)
legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
title(ylab=satlab, mgp=ypos, cex=1.2)

#leaf types
sun_gm <- gm_water[gm_water$leaf =="sun",]
shade_gm <- gm_water[gm_water$leaflight == "shade-low",]
lightson <- gm_water[gm_water$leaflight == "shade-high",]

#simple plots through time
plot(gm~campaign, data=sun_gm, pch=16, col=temp, ylim=c(0, .6), ylab="sun")
plot(gm~campaign, data=shade_gm, pch=16, col=temp, ylim=c(0, .6), ylab="shade")
plot(gm~campaign, data=lightson, pch=16, col=temp, ylim=c(0, .6), ylab="shade-high")

#means
gm_agg <- summaryBy(gm~ temp+leaf+light, data=gm_water,FUN=mean, keep.names=TRUE)
gm_drought_agg <- summaryBy(gm~ temp+leaf+light, data=gm_drought,FUN=mean, keep.names=TRUE)

##gm-temp
gm_ch<- summaryBy(gm+Photo+Cond+CTleaf ~id+chamber+temp+leaf+light+leaflight+Month, data=gm_water,FUN=mean,
                  keep.names=TRUE)

palette(c("blue", "red"))

##photosynthesis and temperature
plot(Photo~CTleaf, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,25), xlim=c(15, 40), 
     ylab="", xlab=leaftlab)
  points(Photo~CTleaf, data=gm_ch, subset=leaflight=="shade-low",pch=1, col=temp)
  legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
  title(ylab=satlab, mgp=ypos, cex=1.2)

plot(gm~CTleaf, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,.5), xlim=c(15, 40), 
     ylab="", xlab=leaftlab)
points(gm~CTleaf, data=gm_ch, subset=leaflight=="shade-low",pch=1, col=temp)
legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
title(ylab=gmlab, mgp=ypos, cex=1.2)

plot(Photo~gm, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,25), xlim=c(0,.5), 
     ylab="", xlab=gmlab)
points(Photo~gm, data=gm_ch, subset=leaflight=="shade-low",pch=1, col=temp)
legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
title(ylab=satlab, mgp=ypos, cex=1.2)

plot(gm~Cond, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,.4), xlim=c(0,.4), 
     ylab="", xlab=gmlab)
points(gm~Cond, data=gm_ch, subset=leaflight=="shade-low",pch=1, col=temp)
legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
title(ylab=condlab, mgp=ypos, cex=1.2)

###make a table of means and se of gm
tablenames <- c( "Leaf Type", "Light", "Temperature","Gm", "Gm(se)")

gm_dfr <- summaryBy(gm~ leaf+light+temp, data=gm_water,FUN=c(mean,se))
names(gm_dfr) <- tablenames

library(xtable)

gm_table <- xtable(gm_dfr)
digits(gm_table)[4:5] <- 4
print(gm_table, type='html')

```

### 2015-03-18 markdown/shade.Rmd
added more figures to gm markdown
fixed gam plot of gs and photo

changed some gm plots to show sun and shade not temp treatement

@@
 ---
-title: "Sun v Shade leaves (WTCIII)"
+## "Sun v Shade leaves (WTCIII)"
 author: "Court Campany"
@@
 ---
-
-```
-This set of figures attempts to show that sun and shade leaves were different on multiple levels.  
+```{r knitr_options , include=FALSE}
+knitr::opts_chunk$set(fig.width=6, fig.height=6,
+echo=FALSE, message=FALSE,
+fig.align='center', warning=FALSE)
 ```
@@
 
-ACI curves were measured at saturating light for both sun and shade leaves.
+### Shade versus Sun leaves
 
-```{r, results='asis', echo=FALSE, message=FALSE}
-library(xtable)
+This set of figures attempts to show that sun and shade leaves were different on multiple levels.
+
+* ACI curves were measured at saturating light for both sun and shade leaves.
+
+```{r, results='asis'}
+library(pander)
 
 aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
+names(aci_leaf) <-c("Leaf", "Temperature", "Vcmax", "Jmax", "Vcmax_se", "Jmax_se")
 
-# aci_table <- xtable(aci_leaf)
-# digits(aci_table)[c(3,5:6)] <- 3
-# digits(aci_table)[c(4)] <- 4
-# print(aci_table, type='html', include.rownames=F,hline.after=NULL)
-# print(aci_table,  type='html', include.rownames=F, comment=FALSE)
+aci_leaf2 <- aci_leaf[,c(1,2,3,5,4,6)]
 
- kable(aci_leaf, row.names=NA, digits=2,format = "html")
-
+panderOptions("digits", 2)
+pander(aci_leaf2, caption="Leaf Biochemistry")
 ```
 
-```
-Here are some predicted ACI curves for sun/shade with temperature treatements.
+* Here are some predicted ACI curves for sun/shade with temperature treatements.
 Both Vcmax and Jmax are reduced in shade leaves with little effect of treatment.
 Aci curves were only taken once towards the end of the experiment on well watered trees.
-```
-```{r,  echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
+
+```{r,results='asis'}
 
@@
 
-```
-This figure shows the relationship between leaf N on an area basis and carbon isotope dicrimination between leaf types.
-```
-```{r, echo=FALSE,fig.align='center', message=FALSE}
+* This figure shows the relationship between leaf N on an area basis and carbon isotope dicrimination between leaf types.
+
+```{r,results='asis'}
 source("master_scripts/plot_objects.R")
@@
 
-Next I show the light environment in which gas exchange measurements for sun and shade leaves were measured.
+* Next I show the light environment in which gas exchange measurements for sun and shade leaves were measured.
 This data was generated with a handheld par sensor for each individual leaf level and verified with ceptometer readings within at the height of the measured leaf.
@@
 
-The second figure shows leaf mass area for each gas exchange campaign.  Here there
+* The second figure shows leaf mass area for each gas exchange campaign.  Here there
 
-```{r,echo=FALSE,fig.align='center',message=FALSE, warning=FALSE}
+```{r,results='asis'}
  source("functions and packages/functions.R")
@@
  library(doBy)
+ library(xtable)
  ####PPFD-----------------------------------------------------------------------------------------------
@@
 
-```{r,echo=FALSE,fig.align='center',message=FALSE, warning=FALSE}
-source("master_scripts/plot_objects.R")
+* Below are means and standard errors of gm across all campaigns.
+
+* We took leaf measurements at their current temperature (ambient, +3) in order to do paired comparisons across leaves at a large range of temperatures.
+Im working on the paired comparisons now, but for this we will just focus on overall differences between leaf types.
+
+* Further analysis is thus needed to tease apart any effects of temperature treatment as these are just the overall means over the 6 campaigns
+
+####gm by leaf type and light environment
+
+* Gm was calculated with Rd and Gstar generated for E.globulus.
+Kristine Crous and I have compared tobacco with globulus calculations on FACE trees and they tend to lower the values of gm and generate less variance within a sample.
+I can easily alter the code to generate values with tobacco parameters if needed.
+
+* Here i seperated the well watered treatment from the drought treatment.
+Im much more confident in the well watered trees as I was often on the border with drawdown and xsi for droughted leave.
+They do so similar patterns
+
+* Shade gm is always lower than sun leaves but when the lights are turned on they have in instantenous response in gm that resembles sun leaves
+
+```{r,results='asis'}
  source("functions and packages/functions.R")
@@
  gm_drought <- gmes[gmes$drydown == "drought",]
+ gm_water <- gmes[gmes$drydown == "control",]
 
-gm_water <- gmes[gmes$drydown == "control",]
-#write.csv(gm_water, "calculated_data/gmes_wellwatered.csv", row.names=FALSE)
+###make a table of means and se of gm
 
-palette(c("blue", "red"))
+ gm_agg <- summaryBy(gm~ leaf+light+temp+leaflight, data=gm_water,FUN=c(mean,se))
+  Lorder <- c("shade-low", "sun-high", "shade-high")
+  gm_agg$leaflight <- factor(gm_agg$leaflight, levels = Lorder)
 
-###drought----------------------------------------------------------------------------------------------------
+   tablenames <- c( "Leaf Type", "Light", "Temperature","leaflight","Gm", "Gm(se)")
+   names(gm_agg) <- tablenames
 
-##plots
-plot(Photo~gm, data=gm_drought, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,25), xlim=c(0, .5))
-  points(Photo~gm, data=gm_drought, subset=leaflight=="shade-low",pch=1, col=temp)
-  legend("topright", templab, pch=16,inset = 0.03, col=palette())
+panderOptions("digits", 3)
+pander(gm_agg,caption="Well Watered")
 
-###well watered------------------------------------------------------------------------------------------------
+  gm_dr <- summaryBy(gm~ leaf+light+temp+leaflight, data=gm_drought,FUN=c(mean,se))
+  Lorder <- c("shade-low", "shade-low","sun-high", "sun-high","shade-high","shade-high")
+  gm_dr$leaflight <- factor(gm_dr$leaflight, levels = Lorder)
+  tablenames <- c( "Leaf Type", "Light", "Temperature","Leaf-Light","Gm", "Gm(se)")
+  names(gm_dr) <- tablenames
 
-##sun-shade
-plot(Photo~gm, data=gm_water, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,25), xlim=c(0,.5), 
-     xlab=gmlab, ylab="")
-points(Photo~gm, data=gm_water, subset=leaflight=="shade-low",pch=1, col=temp)
-legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
+pander(gm_dr,caption="Drought")
+
+```
+
+* Next, Ill just show means of sun and shade leaf physiology and keep it simple by ignoring the temperature treatment. We can address that later.
+
+```{r}
+source("functions and packages/functions.R")
+source("master_scripts/plot_objects.R")
+source("functions and packages/packages.R")
+
+#read in gm data set (no drought)
+
+gmes <- read.csv("calculated_data/gmes_wellwatered.csv")
+Ci_bar <- read.csv("calculated_data/Ci_bar.csv")
+
+##calculate CC
+gmes$Cc<- with(gmes, Ci-Photo/gm)
+gmes$cc_ci<- with(gmes, Ci/Cc)
+
+###get average by id
+gm_agg <- summaryBy(Photo+Cond+Ci+Cc+gm+VpdL+xsi+DELTA+cc_ci ~ chamber+id+leaf +light+temp+leaflight, 
+                    data=gmes, FUN=mean, keep.names=TRUE)
+
+##remove shade-high
+gm_sunsha <- gm_agg[gm_agg$leaflight != "shade-high",]
+  gm_sunsha <- droplevels(gm_sunsha)
+  gm_sunsha$id <- gsub("-high", "", gm_sunsha$id)
+  gm_sunsha$id <- gsub("-low", "", gm_sunsha$id)
+
+##merge
+gm_c13 <- merge(gm_sunsha, Ci_bar[, c(2,8)], by="id")
+  gm_c13$Cc_bar <- with(gm_c13, ci_bar-Photo/gm)
+
+###Photosynthesis vs gs (need to fit sun with something else)------------------------------------------------
+  library(mgcv)
+  library(lme4)
+
+  palette(c(shacol, suncol))
+
+  ####run gam models and then predict
+  #SUN leaves
+  sunmod <- gam(Photo ~ s(Cond, k=5), data=gm_c13, subset=leaflight=="sun-high")
+
+  #predict
+  #get apprpriate vector of gs from sun leaves
+  gsdat <- gm_c13[gm_c13$leaflight=="sun-high", "Cond"]
+
+  #generate sequence and then predict
+  gssun_seq <- seq(min(gsdat), max(gsdat), length=101)
+  gssun_pred <- predict(sunmod, newdata=data.frame(Cond=gssun_seq), se.fit=TRUE)
+
+  #ci and model fit
+  sunupr <- gssun_pred$fit + (2*gssun_pred$se.fit)
+  sunlwr <- gssun_pred$fit - (2*gssun_pred$se.fit)
+
+  #SHADE leaves
+  shamod <- gam(Photo ~ s(Cond, k=5), data=gm_c13, subset=leaflight=="shade-low")
+
+  #get apprpriate vector CC from sun leaves
+  gsdat2 <- gm_c13[gm_c13$leaflight=="shade-low", "Cond"]
+  #generate sequence and then predict
+  gssha_seq <- seq(min(gsdat2), max(gsdat2), length=101)
+  gssha_pred <- predict(shamod, newdata=data.frame(Cond=gssha_seq), type="link", se.fit=TRUE)
+
+  shaupr <- gssha_pred$fit + (2*gssha_pred$se.fit)
+  shalwr <- gssha_pred$fit - (2*gssha_pred$se.fit)
+
+  ###plot
+  plot(Photo~Cond, data=gm_c13, subset=leaflight=="sun-high", pch=16, col=suncol, ylim=c(0,25), 
+       xlim=c(0,.4), xlab=condlab, ylab="", cex=1.25)
+
+  lines(gssun_seq, sunupr, lty=2, lwd=2,col=suncol)
+  lines(gssun_seq, sunlwr, lty=2, lwd=2,col=suncol)
+  lines(gssun_seq, gssun_pred$fit, lty=1, lwd=2,col=suncol)
+
+  #shade
+  points(Photo~Cond, data=gm_c13, subset=leaflight=="shade-low", pch=16, col=shacol, cex=1.25)
+  lines(gssha_seq, shaupr, lty=2, lwd=2,col=shacol)
+  lines(gssha_seq, shalwr, lty=2, lwd=2,col=shacol)
+  lines(gssha_seq, gssha_pred$fit, lty=1, lwd=2,col=shacol)
+
   title(ylab=satlab, mgp=ypos, cex=1.2)
+  legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol)
 
-##shade low-high
-plot(Photo~gm, data=gm_water, subset=leaflight=="shade-high", pch=16, col=temp, ylim=c(0,25), xlim=c(0,.5), 
-     xlab=gmlab, ylab="")
-points(Photo~gm, data=gm_water, subset=leaflight=="shade-low",pch=1, col=temp)
-legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
+###Photosynthesis vs Ci--------------------------------------------------------------------------------
+
+#SUN leaves
+Aci_sun_lm <- lm(Photo~ Ci, data=gm_c13,subset=leaflight=="sun-high")
+
+  #predict
+  #get apprpriate vector CC from sun leaves
+  cidat <- gm_c13[gm_c13$leaflight=="sun-high", "Ci"]
+  #generate sequence and then predict
+  cisun_seq <- seq(min(cidat), max(cidat), length=101)
+  cisun_pred <- predict.lm(Aci_sun_lm, newdata=data.frame(Ci=cisun_seq), interval="confidence")
+
+#SHADE leaves
+Aci_shade_lm <- lm(Photo~ Ci, data=gm_c13, subset=leaflight=="shade-low")
+
+  #get apprpriate vector CC from sun leaves
+  cidat2 <- gm_c13[gm_c13$leaflight=="shade-low", "Ci"]
+  #generate sequence and then predict
+  cisha_seq <- seq(min(cidat2), max(cidat2), length=101)
+  cisha_pred <- predict.lm(Aci_shade_lm, newdata=data.frame(Ci=cisha_seq), interval="confidence")
+
+plot(Photo~Ci, data=gm_c13, subset=leaflight=="sun-high", pch=16, col=suncol, ylim=c(0,25), 
+     xlim=c(0,350), xlab=cilab, ylab="", cex=1.25)
+  ablineclip(Aci_sun_lm, lty=1, x1=min(gm_c13[gm_c13$leaf=="sun","Ci"]), x2=max(gm_c13[gm_c13$leaf=="sun","Ci"]), 
+             col="forestgreen", lwd=2)
+  lines(cisun_seq, cisun_pred[,2], lty=2, lwd=2, col="forestgreen")
+  lines(cisun_seq, cisun_pred[,3], lty=2, lwd=2, col="forestgreen")
+  #shade
+  points(Photo~Ci, data=gm_c13, subset=leaflight=="shade-low", pch=16, col=shacol, cex=1.25)
+
+  ablineclip(Aci_shade_lm, lty=1, x1=min(gm_c13[gm_c13$leaflight=="shade-low","Ci"]), 
+             x2=max(gm_c13[gm_c13$leaflight=="shade-low","Ci"]), col="yellow4", lwd=2)
+  lines(cisha_seq, cisha_pred[,2], lty=2, lwd=2,col="yellow4")
+  lines(cisha_seq, cisha_pred[,3], lty=2, lwd=2,col="yellow4")
+
   title(ylab=satlab, mgp=ypos, cex=1.2)
+  legend("topright", leaflab2, pch=16,inset = 0.03, col=leafcol)
 
-#leaf types
-sun_gm <- gm_water[gm_water$leaf =="sun",]
-shade_gm <- gm_water[gm_water$leaflight == "shade-low",]
-lightson <- gm_water[gm_water$leaflight == "shade-high",]
+###Photosynthesis vs Cc  
+  #SUN leaves
+  ##photo not different by temp treatment, so not used
+Acc_sun_lm <- lm(Photo~ Cc, data=gm_c13,subset=leaflight=="sun-high")
 
-#simple plots through time
-plot(gm~campaign, data=sun_gm, pch=16, col=temp, ylim=c(0, .6), ylab="sun")
-plot(gm~campaign, data=shade_gm, pch=16, col=temp, ylim=c(0, .6), ylab="shade")
-plot(gm~campaign, data=lightson, pch=16, col=temp, ylim=c(0, .6), ylab="shade-high")
+  ##instead predict Photo over a sequence of CC using the model fit
 
-#means
-gm_agg <- summaryBy(gm~ temp+leaf+light, data=gm_water,FUN=mean, keep.names=TRUE)
-gm_drought_agg <- summaryBy(gm~ temp+leaf+light, data=gm_drought,FUN=mean, keep.names=TRUE)
+  #get apprpriate vector CC from sun leaves
+  ccdat <- gm_c13[gm_c13$leaflight=="sun-high", "Cc"]
+  #generate sequence and then predict
+  ccsun_seq <- seq(min(ccdat), max(ccdat), length=101)
+  ccsun_pred <- predict.lm(Acc_sun_lm, newdata=data.frame(Cc=ccsun_seq), interval="confidence")
 
+#SHADE leaves
+Acc_shade_lm <- lm(Photo~ Cc, data=gm_c13,subset=leaflight=="shade-low")
+
+  #get apprpriate vector CC from sun leaves
+  ccdat2 <- gm_c13[gm_c13$leaflight=="shade-low", "Cc"]
+  #generate sequence and then predict
+  ccsha_seq <- seq(min(ccdat2), max(ccdat2), length=101)
+  ccsha_pred <- predict.lm(Acc_shade_lm, newdata=data.frame(Cc=ccsha_seq), interval="confidence")
+
+##plot
+plot(Photo~Cc, data=gm_c13, subset=leaflight=="sun-high", pch=16, col=suncol, ylim=c(0,25), 
+     xlim=c(0,350), xlab=cclab, ylab="", cex=1.25)
+  ablineclip(Acc_sun_lm, lty=1, x1=min(gm_c13[gm_c13$leaf=="sun","Cc"]), x2=max(gm_c13[gm_c13$leaf=="sun","Cc"]), 
+           col="forestgreen", lwd=2)
+  lines(ccsun_seq, ccsun_pred[,2], lty=2, lwd=2, col="forestgreen")
+  lines(ccsun_seq, ccsun_pred[,3], lty=2, lwd=2, col="forestgreen")
+  #shade
+  points(Photo~Cc, data=gm_c13, subset=leaflight=="shade-low", pch=16, col=shacol, cex=1.25)
+  ablineclip(Acc_shade_lm, lty=1, x1=min(gm_c13[gm_c13$leaflight=="shade-low","Cc"]), 
+             x2=max(gm_c13[gm_c13$leaflight=="shade-low","Cc"]), col="yellow4", lwd=2)
+  lines(ccsha_seq, ccsha_pred[,2], lty=2, lwd=2,col="yellow4")
+  lines(ccsha_seq, ccsha_pred[,3], lty=2, lwd=2,col="yellow4")
+
+  title(ylab=satlab, mgp=ypos, cex=1.2)
+  legend("topright", leaflab2, pch=16,inset = 0.03, col=leafcol)
+
+  ###Photosynthesis vs gm-------------------------------------------------------------------------------
+
+#SUN leaves
+Agm_sun_lm <- lm(Photo~ gm, data=gm_c13,subset=leaflight=="sun-high")
+
+#predict
+#get apprpriate vector CC from sun leaves
+gmdat <- gm_c13[gm_c13$leaflight=="sun-high", "gm"]
+#generate sequence and then predict
+gmsun_seq <- seq(min(gmdat), max(gmdat), length=101)
+gmsun_pred <- predict.lm(Agm_sun_lm, newdata=data.frame(gm=gmsun_seq), interval="confidence")
+
+#SHADE leaves
+Agm_sha_lm <- lm(Photo~ gm, data=gm_c13, subset=leaflight=="shade-low")
+
+#get apprpriate vector CC from sun leaves
+gmdat2 <- gm_c13[gm_c13$leaflight=="shade-low", "gm"]
+#generate sequence and then predict
+gmsha_seq <- seq(min(gmdat2), max(gmdat2), length=101)
+gmsha_pred <- predict.lm(Agm_sha_lm, newdata=data.frame(gm=gmsha_seq), interval="confidence")
+
+plot(Photo~gm, data=gm_c13, subset=leaflight=="sun-high", pch=16, col=suncol, ylim=c(0,25), 
+     xlim=c(0,.45), xlab=gmlab, ylab="", cex=1.25)
+ablineclip(Agm_sun_lm, lty=1, x1=min(gm_c13[gm_c13$leaf=="sun","gm"]), x2=max(gm_c13[gm_c13$leaf=="sun","gm"]), 
+           col="forestgreen", lwd=2)
+lines(gmsun_seq, gmsun_pred[,2], lty=2, lwd=2, col="forestgreen")
+lines(gmsun_seq, gmsun_pred[,3], lty=2, lwd=2, col="forestgreen")
+#shade
+points(Photo~gm, data=gm_c13, subset=leaflight=="shade-low", pch=16, col=shacol, cex=1.25)
+
+ablineclip(Agm_sha_lm, lty=1, x1=min(gm_c13[gm_c13$leaflight=="shade-low","gm"]), 
+           x2=max(gm_c13[gm_c13$leaflight=="shade-low","gm"]), col="yellow4", lwd=2)
+lines(gmsha_seq, gmsha_pred[,2], lty=2, lwd=2,col="yellow4")
+lines(gmsha_seq, gmsha_pred[,3], lty=2, lwd=2,col="yellow4")
+
+title(ylab=satlab, mgp=ypos, cex=1.2)
+legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol)
+
+```
+
+* Here we see that there is a very slight reduction in gm with increases in temperature, but only in sun leaves
+
+```{r}
 ##gm-temp
@@
 
-palette(c("blue", "red"))
+##photosynthesis and temperature
+gmt_sun_lm <- lm(gm~ CTleaf, data=gm_water,subset=leaflight=="sun-high")
+  sundat <- gm_water[gm_water$leaflight=="sun-high", "CTleaf"]
+  sun_seq <- seq(min(sundat), max(sundat), length=101)
+  sun_pred <- predict.lm(gmt_sun_lm, newdata=data.frame(CTleaf=sun_seq), interval="confidence")
 
-##photosynthesis and temperature
-plot(Photo~CTleaf, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,25), xlim=c(15, 40), 
+gmt_sha_lm <- lm(gm~ CTleaf, data=gm_water,subset=leaflight=="shade-low")
+  shadat <- gm_water[gm_water$leaflight=="shade-low", "CTleaf"]
+  sha_seq <- seq(min(shadat), max(shadat), length=101)
+  sha_pred <- predict.lm(gmt_sha_lm, newdata=data.frame(CTleaf=sha_seq), interval="confidence")
+
+plot(gm~CTleaf, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=suncol, ylim=c(0,.4), xlim=c(15, 40), 
      ylab="", xlab=leaftlab)
-  points(Photo~CTleaf, data=gm_ch, subset=leaflight=="shade-low",pch=1, col=temp)
-  legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
+
+  ablineclip(gmt_sun_lm, lty=1, x1=min(gm_water[gm_water$leaf=="sun","CTleaf"]), 
+             x2=max(gm_water[gm_water$leaf=="sun","CTleaf"]), col="forestgreen", lwd=2)
+  lines(sun_seq, sun_pred[,2], lty=2, lwd=2, col="forestgreen")
+  lines(sun_seq, sun_pred[,3], lty=2, lwd=2, col="forestgreen")
+  #shade
+  points(gm~CTleaf, data=gm_ch, subset=leaflight=="shade-low",pch=16, col=shacol)
+  ablineclip(gmt_sha_lm, lty=1, x1=min(gm_water[gm_water$leaf=="sun","CTleaf"]), 
+             x2=max(gm_water[gm_water$leaf=="sun","CTleaf"]), col="yellow4", lwd=2)
+  lines(sha_seq, sha_pred[,2], lty=2, lwd=2, col="yellow4")
+  lines(sha_seq, sha_pred[,3], lty=2, lwd=2, col="yellow4")
+
+legend("topright", c("sun", "shade"), pch=16,inset = 0.03, col=leafcol) 
+title(ylab=gmlab, mgp=ypos, cex=1.2)
+```
+
+* This figure shows the instantaneous response of gm in shade leaves when the lights are 'turned on'
+
+```{r}
+source("master_scripts/plot_objects.R")
+source("functions and packages/functions.R")
+library(doBy)
+
+gmes <- read.csv("calculated_data/gmes_WTC.csv")
+###for analysis first subset well watered and drought treatments
+gm_drought <- gmes[gmes$drydown == "drought",]
+gm_water <- gmes[gmes$drydown == "control",]
+
+lightscol <- alpha("darkorange2", .75)
+lightlab <- c(lightscol,shacol)
+lightleg <- c("High light", "Low light")
+##shade low-high
+plot(Photo~gm, data=gm_water, subset=leaflight=="shade-high", pch=16, col=lightscol, ylim=c(0,25), xlim=c(0,.4), 
+     xlab=gmlab, ylab="", cex=1.25)
+points(Photo~gm, data=gm_water, subset=leaflight=="shade-low",pch=16, col=shacol, cex=1.25)
+legend("topright", lightleg, pch=16,inset = 0.02, col=lightlab) 
+legend("topleft", "Shade Leaves", bty='n') 
 title(ylab=satlab, mgp=ypos, cex=1.2)
-
-plot(gm~CTleaf, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,.5), xlim=c(15, 40), 
-     ylab="", xlab=leaftlab)
-points(gm~CTleaf, data=gm_ch, subset=leaflight=="shade-low",pch=1, col=temp)
-legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
-title(ylab=gmlab, mgp=ypos, cex=1.2)
-
-plot(Photo~gm, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,25), xlim=c(0,.5), 
-     ylab="", xlab=gmlab)
-points(Photo~gm, data=gm_ch, subset=leaflight=="shade-low",pch=1, col=temp)
-legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
-title(ylab=satlab, mgp=ypos, cex=1.2)
-
-plot(gm~Cond, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,.4), xlim=c(0,.4), 
-     ylab="", xlab=gmlab)
-points(gm~Cond, data=gm_ch, subset=leaflight=="shade-low",pch=1, col=temp)
-legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
-title(ylab=condlab, mgp=ypos, cex=1.2)
-
-###make a table of means and se of gm
-tablenames <- c( "Leaf Type", "Light", "Temperature","Gm", "Gm(se)")
-
-gm_dfr <- summaryBy(gm~ leaf+light+temp, data=gm_water,FUN=c(mean,se))
-names(gm_dfr) <- tablenames
-
-library(xtable)
-
-gm_table <- xtable(gm_dfr)
-digits(gm_table)[4:5] <- 4
-print(gm_table, type='html')
-
 ```

### 2015-03-18 master_scripts/sunshade_phys.R

@@ -47,15 +47,14 @@ gm_c13 <- merge(gm_sunsha, Ci_bar[, c(2,8)], by="id")
   #predict
   #get apprpriate vector of gs from sun leaves
   gsdat <- gm_c13[gm_c13$leaflight=="sun-high", "Cond"]
+  
   #generate sequence and then predict
   gssun_seq <- seq(min(gsdat), max(gsdat), length=101)
   gssun_pred <- predict(sunmod, newdata=data.frame(Cond=gssun_seq), se.fit=TRUE)
   
-  sunupper <- gssun_pred$fit + (1.96*gssun_pred$se.fit)
-  sunlower <- gssun_pred$fit - (1.96*gssun_pred$se.fit)
-  
-  sunupr <- sunmod$family$linkinv(sunupper)
-  sunlwr <- sunmod$family$linkinv(sunlower)
+  #ci and model fit
+  sunupr <- gssun_pred$fit + (2*gssun_pred$se.fit)
+  sunlwr <- gssun_pred$fit - (2*gssun_pred$se.fit)
   
   #SHADE leaves
   shamod <- gam(Photo ~ s(Cond, k=5), data=gm_c13, subset=leaflight=="shade-low")
@@ -66,28 +65,30 @@ gm_c13 <- merge(gm_sunsha, Ci_bar[, c(2,8)], by="id")
   gssha_seq <- seq(min(gsdat2), max(gsdat2), length=101)
   gssha_pred <- predict(shamod, newdata=data.frame(Cond=gssha_seq), type="link", se.fit=TRUE)
   
-  shaupper <- gssha_pred$fit + (2*gssha_pred$se.fit)
-  shalower <- gssha_pred$fit - (2*gssha_pred$se.fit)
+  shaupr <- gssha_pred$fit + (2*gssha_pred$se.fit)
+  shalwr <- gssha_pred$fit - (2*gssha_pred$se.fit)
   
-  shaupr <- shamod$family$linkinv(shaupper)
-  shalwr <- shamod$family$linkinv(shalower)
-  
-  ###redo plot
-  
+  ###plot
   windows(10,8)
+  plot(Photo~Cond, data=gm_c13, subset=leaflight=="sun-high", pch=16, col=suncol, ylim=c(0,25), 
+       xlim=c(0,.4), xlab=condlab, ylab="", cex=1.25)
   
[…]

### 2015-03-20 markdown/shade.Rmd
finsihed bootstrapping lme of leaf phys
finished markdown of wtc data (1st draft)

@@
 ---
-## "Sun v Shade leaves (WTCIII)"
+title: "Sun v Shade leaves (WTCIII)"
 author: "Court Campany"
@@
 
-```{r, results='asis', echo=FALSE}
+```{r, results='asis'}
 library(knitr)
@@
 
-### Shade versus Sun leaves
-
-This set of figures attempts to show that sun and shade leaves were different on multiple levels.
-
-* ACI curves were measured at saturating light for both sun and shade leaves.
-
-```{r, results='asis'}
-library(pander)
-
-aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
-names(aci_leaf) <-c("Leaf", "Temperature", "Vcmax", "Jmax", "Vcmax_se", "Jmax_se")
-
-aci_leaf2 <- aci_leaf[,c(1,2,3,5,4,6)]
-
-panderOptions("digits", 2)
-pander(aci_leaf2, caption="Leaf Biochemistry")
-```
-
-* Here are some predicted ACI curves for sun/shade with temperature treatements.
-Both Vcmax and Jmax are reduced in shade leaves with little effect of treatment.
-Aci curves were only taken once towards the end of the experiment on well watered trees.
-
-```{r,results='asis'}
-
-source("functions and packages/functions.R")
-source("master_scripts/plot_objects.R")
-aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
-
-#simulated curves
-  library(plantecophys)
-
-  #simulate ACi curves for each leaf, ambient and elevated T
-  sunAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[3,3], Jmax=aci_leaf[3,4],PPFD=1408)
-  sunET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[4,3], Jmax=aci_leaf[4,4], PPFD=1408)
-  shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[1,3], Jmax=aci_leaf[1,4], PPFD=375)
-  shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[2,3], Jmax=aci_leaf[2,4],PPFD=375)
-
-#plot
-plot(sunAT_sim$Ci, sunAT_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab=cilab, ylab="", type="l", lwd=2)
-  points(sunET_sim$Ci, sunET_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
-  points(shaAT_sim$Ci, shaAT_sim$ALEAF, col="yellowgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2)
-  points(shaET_sim$Ci, shaET_sim$ALEAF, col="yellowgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
-  legend("bottomright", leglab2, lty=c(1,2,1,2), lwd=2,col=colaci, pt.bg=col4,inset = 0.03)
-  title(ylab=satlab, mgp=ypos)
-
-```
-
-* This figure shows the relationship between leaf N on an area basis and carbon isotope dicrimination between leaf types.
-
-```{r,results='asis'}
-source("master_scripts/plot_objects.R")
-leaftraits <- read.csv("calculated_data/leaftraits_summary.csv")
-##this has droughtleaves
-
-palette(c("yellowgreen", "green4"))
-
-plot(c13.mean ~ leafN_area.mean, data=leaftraits, pch=16, col=as.factor(leaf), ylim=c(-35,-25), 
-                 xlim=c(1,3.5), cex=1.25, xlab=narealab, ylab=c13lab)
-legend("bottomright", leaflab, pch=16, col=c("yellowgreen", "green4"),pt.cex=1.5,inset = 0.03)
-
-```
-
-* Next I show the light environment in which gas exchange measurements for sun and shade leaves were measured.
-This data was generated with a handheld par sensor for each individual leaf level and verified with ceptometer readings within at the height of the measured leaf.
-There was on average a ~75 reduction in PPFD from sun to shave leaves.
-
-* The second figure shows leaf mass area for each gas exchange campaign.  Here there
-
-```{r,results='asis'}
- source("functions and packages/functions.R")
- source("master_scripts/plot_objects.R")
- library(doBy)
- library(xtable)
- ####PPFD-----------------------------------------------------------------------------------------------
- par <- read.csv("raw data/par.csv")
- treatments <- read.csv("raw data/chamber_trt.csv")
-
- #format function
- par<- parformat(par)
-
- par_leaf <- subset(par, ID !="shade-high")
- Morder <- c("Oct", "Dec", "Jan", "Feb", "Mar", "Apr")
-
- #data format for bar
- parbar <- subset(par_leaf, select = c("par", "month", "leaf_type"))
- parbar$month <- factor(parbar$month, levels = Morder)
-
- bar(par, c(leaf_type, month), parbar, col=c("yellowgreen", "green4"), xlab="", ylab="", ylim=c(0, 2000), 
-     half.errbar=FALSE)
- title(ylab=parlab, mgp=ypos)
-
- ###LMA--------------------------------------------------------------------------------------------------
- lma <- read.csv("calculated_data/leaf_chemistry.csv")
-
- #data format for bar
- lmabar <- subset(lma, select = c("lma", "Month", "leaf"))
-   lmabar$Month <- factor(lmabar$Month, levels = Morder)
-
- bar(lma, c(leaf, Month), lmabar, col=c("yellowgreen", "green4"), xlab="", ylab="", ylim=c(0, 185), half.errbar=FALSE)
- title(ylab=lmalab, mgp=ypos)
-
- ##table summ by treatment
-   lma_agg <- summaryBy(lma~ leaf+temp+drydown, data=lma, FUN=c(mean,se))
-   lma_agg2 <- summaryBy(lma~ leaf+temp, data=lma, FUN=c(mean,se))
-
-   lma_table <- xtable(lma_agg2)
-
-```
-
-## Mesophyll Conducatnce
+### Mesophyll Conducatnce
 
@@
 
-* We took leaf measurements at their current temperature (ambient, +3) in order to do paired comparisons across leaves at a large range of temperatures.
-Im working on the paired comparisons now, but for this we will just focus on overall differences between leaf types.
+* We took leaf measurements at their current temperature (ambient, +3) for paired comparisons across leaves at a large range of temperatures.
 
-* Further analysis is thus needed to tease apart any effects of temperature treatment as these are just the overall means over the 6 campaigns
+* For simplicity I have mostly generalized the results to sun vs shade leaves below.
+
+* Further analysis is thus needed to tease apart any effects of temperature treatment as data below are overall means across the 6 campaigns
 
@@
 Im much more confident in the well watered trees as I was often on the border with drawdown and xsi for droughted leave.
-They do so similar patterns
+They do so similar patterns but gm seems to be higher in droughted trees
 
@@
  library(doBy)
+library(pander)
 
@@
    names(gm_agg) <- tablenames
+gm_agg2 <- gm_agg[c(3,4,5,6,1,2),]
+row.names(gm_agg2) <- NULL
 
 panderOptions("digits", 3)
-pander(gm_agg,caption="Well Watered")
+pander(gm_agg2,caption="Well Watered")
 
@@
 
-pander(gm_dr,caption="Drought")
+  gm_dr2 <- gm_dr[c(3,4,5,6,1,2),]
+  row.names(gm_dr2) <- NULL
+
+pander(gm_dr2,caption="Drought")
 
@@
 
-* Next, Ill just show means of sun and shade leaf physiology and keep it simple by ignoring the temperature treatment. We can address that later.
+* This figure shows the instantaneous response of gm in shade leaves when the lights are 'turned on'
+
+```{r}
+source("master_scripts/plot_objects.R")
+source("functions and packages/functions.R")
+library(doBy)
+
+gmes <- read.csv("calculated_data/gmes_WTC.csv")
+###for analysis first subset well watered and drought treatments
+gm_drought <- gmes[gmes$drydown == "drought",]
+gm_water <- gmes[gmes$drydown == "control",]
+
+lightscol <- alpha("darkorange2", .75)
+lightlab <- c(lightscol,shacol)
+lightleg <- c("High light", "Low light")
+##shade low-high
+plot(Photo~gm, data=gm_water, subset=leaflight=="shade-high", pch=16, col=lightscol, ylim=c(0,25), xlim=c(0,.4), 
+     xlab=gmlab, ylab="", cex=1.25)
+points(Photo~gm, data=gm_water, subset=leaflight=="shade-low",pch=16, col=shacol, cex=1.25)
+legend("topright", lightleg, pch=16,inset = 0.02, col=lightlab) 
+legend("topleft", "Shade Leaves", bty='n') 
+title(ylab=satlab, mgp=ypos, cex=1.2)
+```
+
+* Here is a simple version of the paired comparisons.  Only shade low light and sun leaves are used (watered and drought treatments together).
+
+```{r}
+source("functions and packages/functions.R")
+source("functions and packages/packages.R")
+#source("scripts/read_data.R")
+source("master_scripts/plot_objects_paired.R")
+
+###there is no ch10 shade in march, check for this in script and paired comparisons
+
+#read in gmsunsha
+#gm data, no drought, no shade high
+gm<- read.csv("calculated_data/gmes_wtc_pair.csv")
+gm$variable <- as.factor(paste(gm$leaf, gm$temp, sep="-"))
+
+gm2 <- gm[gm$leaflight != "shade-high",]
+
+#paired id dfr
+idDF <- data.frame(pair_id = unique(gm$pair_id))
+
+gm_id <- summaryBy(gm~ variable+pair_id, data=gm2, FUN=c(mean), keep.names=TRUE)
+gm_sp <- dlply(gm_id, .(variable), function(x) merge(x, idDF, all = TRUE))
+
+par(cex.axis=1.21, cex.lab=1.51,las=1,mgp=c(3.5,1,0),mfrow=c(3,1),  
+    omi=c(.5,0,0.1,0.1))   
+# First Panel
+par(mar=c(0,7,2,2))
+plot(gm_sp[[4]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[3], cex=2.5, ylab="", xlab="", xlim=c(0,.5), ylim=c(0,.5),
+     axes=FALSE) 
+abline(0,1, lty=2)
+axis(1, labels=FALSE) 
+axis(2, labels=TRUE) 
+box()
+legend(.4,.45, pairlab[1], pch=21, pt.bg=paircol[1], pt.cex=2,bg="white",inset = 0.06)
+
+# Second panel   
+par(mar=c(0,7,0,2))
+plot(gm_sp[[1]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[1], cex=2.5, ylab=gmlab, xlim=c(0,.5), ylim=c(0,.5),xlab="",
+     axes=FALSE)     
+abline(0,1, lty=2)
+legend(.4,.45, pairlab[2], pch=21, pt.bg=paircol[2], pt.cex=2,bg="white",inset = 0.06)
+axis(1, labels=FALSE) 
+axis(2, labels=TRUE) 
+box()
+
+#third panel
+par(mar=c(5,7,0,2))
+plot(gm_sp[[2]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[5], cex=2.5, ylab="", xlim=c(0,.5), ylim=c(0,.5),xlab=suatgm)   
+abline(0,1, lty=2)
+legend(.4,.45, pairlab[3], pch=21, pt.bg=paircol[3], pt.cex=2,bg="white",inset = 0.06) 
+```
+
+## Leaf Physiology
+* Below are means of sun and shade leaf physiology, keeping it simple by ignoring the temperature treatment. We can address that later.
 
@@
   legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol) 
+```
 
-###Photosynthesis vs Ci--------------------------------------------------------------------------------
+```{r}  
+source("functions and packages/functions.R")
+source("master_scripts/plot_objects.R")
+library(nlme)
+library(doBy)
 
+#read in gm data set (no drought)
+
+gmes <- read.csv("calculated_data/gmes_wellwatered.csv")
+Ci_bar <- read.csv("calculated_data/Ci_bar.csv")
+
+##calculate CC
+gmes$Cc<- with(gmes, Ci-Photo/gm)
+gmes$cc_ci<- with(gmes, Ci/Cc)
+
+###get average by id
+gm_agg <- summaryBy(Photo+Cond+Ci+Cc+gm+VpdL+xsi+DELTA+cc_ci ~ chamber+id+leaf +light+temp+leaflight, 
+                    data=gmes, FUN=mean, keep.names=TRUE)
+
+##remove shade-high
+gm_sunsha <- gm_agg[gm_agg$leaflight != "shade-high",]
+  gm_sunsha <- droplevels(gm_sunsha)
+  gm_sunsha$id <- gsub("-high", "", gm_sunsha$id)
+  gm_sunsha$id <- gsub("-low", "", gm_sunsha$id)
+
+##merge
+gm_c13 <- merge(gm_sunsha, Ci_bar[, c(2,8)], by="id")
+  gm_c13$Cc_bar <- with(gm_c13, ci_bar-Photo/gm)
+
+###Photosynthesis vs Ci
+#library(nlme)
 #SUN leaves
@@
   cisun_seq <- seq(min(cidat), max(cidat), length=101)
-  cisun_pred <- predict.lm(Aci_sun_lm, newdata=data.frame(Ci=cisun_seq), interval="confidence")
+  cisun_pred <- predict(Aci_sun_lm, newdata=data.frame(Ci=cisun_seq),interval="confidence")
 
@@
   cisha_seq <- seq(min(cidat2), max(cidat2), length=101)
-  cisha_pred <- predict.lm(Aci_shade_lm, newdata=data.frame(Ci=cisha_seq), interval="confidence")
+  cisha_pred <- predict(Aci_shade_lm, newdata=data.frame(Ci=cisha_seq), interval="confidence")
 
@@
 
-* This figure shows the instantaneous response of gm in shade leaves when the lights are 'turned on'
+### Shade versus Sun leaves
 
-```{r}
+This set of figures attempts to show that sun and shade leaves were different on multiple levels.
+
+* ACI curves were measured at saturating light for both sun and shade leaves.
+
+```{r, results='asis'}
+library(pander)
+
+aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
+names(aci_leaf) <-c("Leaf", "Temperature", "Vcmax", "Jmax", "Vcmax_se", "Jmax_se")
+
+aci_leaf2 <- aci_leaf[,c(1,2,3,5,4,6)]
+
+panderOptions("digits", 2)
+pander(aci_leaf2, caption="Leaf Biochemistry")
+```
+
+* Here are some predicted ACI curves for sun/shade with temperature treatements.
+Both Vcmax and Jmax are reduced in shade leaves with little effect of treatment.
+Aci curves were only taken once towards the end of the experiment on well watered trees.
+
+```{r,results='asis'}
+
+source("functions and packages/functions.R")
 source("master_scripts/plot_objects.R")
+aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
+
+#simulated curves
+  library(plantecophys)
+
+  #simulate ACi curves for each leaf, ambient and elevated T
+  sunAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[3,3], Jmax=aci_leaf[3,4],PPFD=1408)
+  sunET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[4,3], Jmax=aci_leaf[4,4], PPFD=1408)
+  shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[1,3], Jmax=aci_leaf[1,4], PPFD=375)
+  shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[2,3], Jmax=aci_leaf[2,4],PPFD=375)
+
+#plot
+plot(sunAT_sim$Ci, sunAT_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab=cilab, ylab="", type="l", lwd=2)
+  points(sunET_sim$Ci, sunET_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
+  points(shaAT_sim$Ci, shaAT_sim$ALEAF, col="yellowgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2)
+  points(shaET_sim$Ci, shaET_sim$ALEAF, col="yellowgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
+  legend("bottomright", leglab2, lty=c(1,2,1,2), lwd=2,col=colaci, pt.bg=col4,inset = 0.03)
+  title(ylab=satlab, mgp=ypos)
+
+```
+
+* This figure shows the relationship between leaf N on an area basis and carbon isotope dicrimination between leaf types.
+
+```{r,results='asis'}
+source("master_scripts/plot_objects.R")
+leaftraits <- read.csv("calculated_data/leaftraits_summary.csv")
+##this has droughtleaves
+
+palette(c("yellowgreen", "green4"))
+
+plot(c13.mean ~ leafN_area.mean, data=leaftraits, pch=16, col=as.factor(leaf), ylim=c(-35,-25), 
+                 xlim=c(1,3.5), cex=1.25, xlab=narealab, ylab=c13lab)
+legend("bottomright", leaflab, pch=16, col=c("yellowgreen", "green4"),pt.cex=1.5,inset = 0.03)
+
+```
+
+* Next I show the light environment in which gas exchange measurements for sun and shade leaves were measured.
+This data was generated with a handheld par sensor for each individual leaf level and verified with ceptometer readings within at the height of the measured leaf.
+There was on average a ~75 reduction in PPFD from sun to shave leaves.
+
+* The second figure shows leaf mass area for each gas exchange campaign.  Here there
+
+```{r,results='asis'}
  source("functions and packages/functions.R")
+ source("master_scripts/plot_objects.R")
  library(doBy)
+ library(xtable)
+ ####PPFD-----------------------------------------------------------------------------------------------
+ par <- read.csv("raw data/par.csv")
+ treatments <- read.csv("raw data/chamber_trt.csv")
 
-gmes <- read.csv("calculated_data/gmes_WTC.csv")
-###for analysis first subset well watered and drought treatments
-gm_drought <- gmes[gmes$drydown == "drought",]
-gm_water <- gmes[gmes$drydown == "control",]
+ #format function
+ par<- parformat(par)
 
-lightscol <- alpha("darkorange2", .75)
-lightlab <- c(lightscol,shacol)
-lightleg <- c("High light", "Low light")
-##shade low-high
-plot(Photo~gm, data=gm_water, subset=leaflight=="shade-high", pch=16, col=lightscol, ylim=c(0,25), xlim=c(0,.4), 
-     xlab=gmlab, ylab="", cex=1.25)
-points(Photo~gm, data=gm_water, subset=leaflight=="shade-low",pch=16, col=shacol, cex=1.25)
-legend("topright", lightleg, pch=16,inset = 0.02, col=lightlab) 
-legend("topleft", "Shade Leaves", bty='n') 
-title(ylab=satlab, mgp=ypos, cex=1.2)
+ par_leaf <- subset(par, ID !="shade-high")
+ Morder <- c("Oct", "Dec", "Jan", "Feb", "Mar", "Apr")
+
+ #data format for bar
+ parbar <- subset(par_leaf, select = c("par", "month", "leaf_type"))
+ parbar$month <- factor(parbar$month, levels = Morder)
+
+ bar(par, c(leaf_type, month), parbar, col=c("yellowgreen", "green4"), xlab="", ylab="", ylim=c(0, 2000), 
+     half.errbar=FALSE)
+ title(ylab=parlab, mgp=ypos)
+
+ ###LMA--------------------------------------------------------------------------------------------------
+ lma <- read.csv("calculated_data/leaf_chemistry.csv")
+
+ #data format for bar
+ lmabar <- subset(lma, select = c("lma", "Month", "leaf"))
+   lmabar$Month <- factor(lmabar$Month, levels = Morder)
+
+ bar(lma, c(leaf, Month), lmabar, col=c("yellowgreen", "green4"), xlab="", ylab="", ylim=c(0, 185), half.errbar=FALSE)
+ title(ylab=lmalab, mgp=ypos)
+
+ ##table summ by treatment
+   lma_agg <- summaryBy(lma~ leaf+temp+drydown, data=lma, FUN=c(mean,se))
+   lma_agg2 <- summaryBy(lma~ leaf+temp, data=lma, FUN=c(mean,se))
+
+   lma_table <- xtable(lma_agg2)
+
 ```

### 2015-03-20 markdown/wtc_sunshade.Rmd
finsihed bootstrapping lme of leaf phys
finished markdown of wtc data (1st draft)

---
title: "Sun v Shade leaves (WTCIII)"
author: "Court Campany"
date: "March 17, 2015"
output: html_document
---
```{r knitr_options , include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=6,
echo=FALSE, message=FALSE,
fig.align='center', warning=FALSE)
```

```{r, results='asis'}
library(knitr)
opts_knit$set(root.dir = '../')
```

```{r global data/sourcing , include=FALSE}
source("functions and packages/packages_md.R")
source("functions and packages/functions.R")
source("functions and packages/plot_objects_all.R")

gmes <- read.csv("calculated_data/gmes_WTC.csv")
gmpair<- read.csv("calculated_data/gmes_wtc_pair.csv")
gmwet <- read.csv("calculated_data/gmes_wellwatered.csv")
Ci_bar <- read.csv("calculated_data/Ci_bar.csv")
 ###for analysis first subset well watered and drought treatments
 gm_drought <- gmes[gmes$drydown == "drought",]
 gm_water <- gmes[gmes$drydown == "control",]

leaftraits <- read.csv("calculated_data/leaftraits_summary.csv") 
par <- read.csv("raw data/par.csv")
treatments <- read.csv("raw data/chamber_trt.csv")

```

### <u>Mesophyll Conducatnce</u>

* We took leaf measurements at their current temperature (ambient, +3C) for paired comparisons across leaf types at a large range of temperatures.

* Sun leaves were taken from upper canopy at their current light level.  Shade leaves were taken from lower canopy, first at their current light level and then again at sun leaf PPFD.

* g<sub>m</sub> was calculated with R<sub>d</sub> and G<sub>star</sub> generated for E.globulus.  Kristine Crous and I have compared tobacco with globulus calculations on FACE trees and they tend to lower the values of g<sub>m</sub> and generate less variance within a leaf sample. I can alter the code to generate values with tobacco parameters, if needed.

* For simplicity I have generalized most of the results below to just sun vs shade leaves (temperature not standardized).

* Further analysis is also needed to tease apart any effects of temperature treatment as data below are overall means across the 6 campaigns

####g<sub>m</sub> by leaf type and light environment

*  Below are means and standard errors of g<sub>m</sub> across all campaigns.

* Here I separated the well watered treatment from the drought treatment.  I'm more confident in the well watered values as I was often on the cutoff point with drawdown and xsi for drought leaves. They do show similar patterns but g<sub>m</sub> seems to be higher in drought trees

* Shade g<sub>m</sub> is always lower than sun leaves but when the lights are turned on they have in instantaneous response in g<sub>m</sub> that resembles sun leaves

```{r,results='asis'}
###make a table of means and se of gm

 gm_agg <- summaryBy(gm~ leaf+light+temp+leaflight, data=gm_water,FUN=c(mean,se))
   tablenames <- c( "Leaf Type", "Light", "Temperature","leaflight","Gm", "Gm(se)")
   names(gm_agg) <- tablenames
  gm_agg2 <- gm_agg[c(3,4,5,6,1,2),]
  row.names(gm_agg2) <- NULL

panderOptions("digits", 3)
pander(gm_agg2,caption="Well Watered")

  gm_dr <- summaryBy(gm~ leaf+light+temp+leaflight, data=gm_drought,FUN=c(mean,se))
    tablenames <- c( "Leaf Type", "Light", "Temperature","Leaf-Light","Gm", "Gm(se)")
    names(gm_dr) <- tablenames
  gm_dr2 <- gm_dr[c(3,4,5,6,1,2),]
  row.names(gm_dr2) <- NULL

pander(gm_dr2,caption="Drought")
```

* This figure shows the instantaneous response of g<sub>m</sub> in shade leaves when the lights are 'turned on'

```{r gm lights on}
#plot low vs high light shade leaves
plot(Photo~gm, data=gm_water, subset=leaflight=="shade-high", pch=16, col=lightscol, ylim=c(0,25), xlim=c(0,.4), 
     xlab=gmlab, ylab="", cex=1.25)
  points(Photo~gm, data=gm_water, subset=leaflight=="shade-low",pch=16, col=shacol, cex=1.25)
  legend("topright", lightleg, pch=16,inset = 0.02, col=lightlab) 
  legend("topleft", "Shade Leaves", bty='n') 
  title(ylab=satlab, mgp=ypos, cex=1.2)
```

* Here is a simple version of the paired comparisons.  Only shade low light and sun leaves are used (watered and drought treatments together). All leaf types are compared to sun leaves at ambient temperature treatment (x axis)

```{r gm paired, fig.height=10}
wa<- palette(wes_palette("Zissou",5))
paircol <- c(wa[3], wa[1],wa[5])

gmpair$variable <- as.factor(paste(gmpair$leaf, gmpair$temp, sep="-"))

gmpair2 <- gmpair[gmpair$leaflight != "shade-high",]

#paired id dfr
idDF <- data.frame(pair_id = unique(gmpair2$pair_id))

gm_id <- summaryBy(gm~ variable+pair_id, data=gmpair2, FUN=c(mean), keep.names=TRUE)
gm_sp <- dlply(gm_id, .(variable), function(x) merge(x, idDF, all = TRUE))

par(cex.axis=1.21, cex.lab=1.51,las=1,mgp=c(3.5,1,0),mfrow=c(3,1),  
    omi=c(.5,0,0.1,0.1))   
# First Panel
par(mar=c(0,7,2,2))
plot(gm_sp[[4]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[3], cex=2.5, ylab="", xlab="", xlim=c(0,.5), ylim=c(0,.5),
     axes=FALSE) 
abline(0,1, lty=2)
axis(1, labels=FALSE) 
axis(2, labels=TRUE) 
box()
legend(.4,.45, pairlab[1], pch=21, pt.bg=paircol[1], pt.cex=2,bg="white",inset = 0.06)

# Second panel   
par(mar=c(0,7,0,2))
plot(gm_sp[[1]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[1], cex=2.5, ylab=gmlab, xlim=c(0,.5), ylim=c(0,.5),xlab="",
     axes=FALSE)     
abline(0,1, lty=2)
legend(.4,.45, pairlab[2], pch=21, pt.bg=paircol[2], pt.cex=2,bg="white",inset = 0.06)
axis(1, labels=FALSE) 
axis(2, labels=TRUE) 
box()

#third panel
par(mar=c(5,7,0,2))
plot(gm_sp[[2]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[5], cex=2.5, ylab="", xlim=c(0,.5), ylim=c(0,.5),xlab=suatgm)   
abline(0,1, lty=2)
legend(.4,.45, pairlab[3], pch=21, pt.bg=paircol[3], pt.cex=2,bg="white",inset = 0.06) 
```

* Here we see a slight reduction in g<sub>m</sub> with increase in temperature, but only in sun leaves

```{r gm temp}
gm_ch<- summaryBy(gm+Photo+Cond+CTleaf ~id+chamber+temp+leaf+light+leaflight+Month, data=gm_water,FUN=mean,
                  keep.names=TRUE)

##photosynthesis and temperature
gmt_sun_lm <- lm(gm~ CTleaf, data=gm_water,subset=leaflight=="sun-high")
  sundat <- gm_water[gm_water$leaflight=="sun-high", "CTleaf"]
  sun_seq <- seq(min(sundat), max(sundat), length=101)
  sun_pred <- predict.lm(gmt_sun_lm, newdata=data.frame(CTleaf=sun_seq), interval="confidence")

gmt_sha_lm <- lm(gm~ CTleaf, data=gm_water,subset=leaflight=="shade-low")
  shadat <- gm_water[gm_water$leaflight=="shade-low", "CTleaf"]
  sha_seq <- seq(min(shadat), max(shadat), length=101)
  sha_pred <- predict.lm(gmt_sha_lm, newdata=data.frame(CTleaf=sha_seq), interval="confidence")

plot(gm~CTleaf, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=suncol, ylim=c(0,.4), xlim=c(15, 40), 
     ylab="", xlab=leaftlab)

  ablineclip(gmt_sun_lm, lty=1, x1=min(gm_water[gm_water$leaf=="sun","CTleaf"]), 
             x2=max(gm_water[gm_water$leaf=="sun","CTleaf"]), col="forestgreen", lwd=2)
  lines(sun_seq, sun_pred[,2], lty=2, lwd=2, col="forestgreen")
  lines(sun_seq, sun_pred[,3], lty=2, lwd=2, col="forestgreen")
  #shade
  points(gm~CTleaf, data=gm_ch, subset=leaflight=="shade-low",pch=16, col=shacol)
  ablineclip(gmt_sha_lm, lty=1, x1=min(gm_water[gm_water$leaf=="sun","CTleaf"]), 
             x2=max(gm_water[gm_water$leaf=="sun","CTleaf"]), col="yellow4", lwd=2)
  lines(sha_seq, sha_pred[,2], lty=2, lwd=2, col="yellow4")
  lines(sha_seq, sha_pred[,3], lty=2, lwd=2, col="yellow4")

legend("topleft", c("sun", "shade"), pch=16,inset = 0.03, col=leafcol) 
title(ylab=gmlab, mgp=ypos, cex=1.2)
```

### <u>Leaf Physiology</u>
* Below are figures of sun and shade leaf physiology, keeping it simple by ignoring the temperature treatment. We can address that later.

```{r A_gs}
##calculate CC
gmwet$Cc<- with(gmwet, Ci-Photo/gm)
gmwet$cc_ci<- with(gmwet, Ci/Cc)

###get average by id
gm_agg <- summaryBy(Photo+Cond+Ci+Cc+gm+VpdL+xsi+DELTA+cc_ci ~ chamber+id+leaf +light+temp+leaflight, 
                    data=gmwet, FUN=mean, keep.names=TRUE)

##remove shade-high
gm_sunsha <- gm_agg[gm_agg$leaflight != "shade-high",]
  gm_sunsha <- droplevels(gm_sunsha)
  gm_sunsha$id <- gsub("-high", "", gm_sunsha$id)
  gm_sunsha$id <- gsub("-low", "", gm_sunsha$id)

##merge
gm_c13 <- merge(gm_sunsha, Ci_bar[, c(2,8)], by="id")
  gm_c13$Cc_bar <- with(gm_c13, ci_bar-Photo/gm)

###Photosynthesis vs gs

  ####run gam models and then predict
  #SUN leaves
  sunmod <- gam(Photo ~ s(Cond, k=5), data=gm_c13, subset=leaflight=="sun-high")

  #predict
  #get apprpriate vector of gs from sun leaves
  gsdat <- gm_c13[gm_c13$leaflight=="sun-high", "Cond"]

  #generate sequence and then predict
  gssun_seq <- seq(min(gsdat), max(gsdat), length=101)
  gssun_pred <- predict(sunmod, newdata=data.frame(Cond=gssun_seq), se.fit=TRUE)

  #ci and model fit
  sunupr <- gssun_pred$fit + (2*gssun_pred$se.fit)
  sunlwr <- gssun_pred$fit - (2*gssun_pred$se.fit)

  #SHADE leaves
  shamod <- gam(Photo ~ s(Cond, k=5), data=gm_c13, subset=leaflight=="shade-low")

  #get apprpriate vector CC from sun leaves
  gsdat2 <- gm_c13[gm_c13$leaflight=="shade-low", "Cond"]
  #generate sequence and then predict
  gssha_seq <- seq(min(gsdat2), max(gsdat2), length=101)
  gssha_pred <- predict(shamod, newdata=data.frame(Cond=gssha_seq), type="link", se.fit=TRUE)

  shaupr <- gssha_pred$fit + (2*gssha_pred$se.fit)
  shalwr <- gssha_pred$fit - (2*gssha_pred$se.fit)

  ###plot
  plot(Photo~Cond, data=gm_c13, subset=leaflight=="sun-high", pch=16, col=suncol, ylim=c(0,25), 
       xlim=c(0,.4), xlab=condlab, ylab="", cex=1.25)

  lines(gssun_seq, sunupr, lty=2, lwd=2,col=suncol)
  lines(gssun_seq, sunlwr, lty=2, lwd=2,col=suncol)
  lines(gssun_seq, gssun_pred$fit, lty=1, lwd=2,col=suncol)

  #shade
  points(Photo~Cond, data=gm_c13, subset=leaflight=="shade-low", pch=16, col=shacol, cex=1.25)
  lines(gssha_seq, shaupr, lty=2, lwd=2,col=shacol)
  lines(gssha_seq, shalwr, lty=2, lwd=2,col=shacol)
  lines(gssha_seq, gssha_pred$fit, lty=1, lwd=2,col=shacol)

  title(ylab=satlab, mgp=ypos, cex=1.2)
  legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol) 
```

```{r leafphys}
source("calculated_data/bootstrap/readboot.R")

###Photosynthesis vs gm
plot(Photo~gm, data=sundat, pch=16, col=suncol, ylim=c(0,25), xlim=c(0,.4), xlab=gmlab, ylab="", cex=1.25)
  points(Photo~gm, data=shadat, pch=16, col=shacol, cex=1.25)
with(agm_sun, {
  lines(gm, lcl, lty=2, lwd=2,col="forestgreen")
  lines(gm, ucl, lty=2, lwd=2,col="forestgreen")
  lines(gm, pred, lty=1, lwd=2,col="forestgreen")
})
with(agm_sha, {
  lines(gm, lcl, lty=2, lwd=2,col="yellow4")
  lines(gm, ucl, lty=2, lwd=2,col="yellow4")
  lines(gm, pred, lty=1, lwd=2,col="yellow4")
})
title(ylab=satlab, mgp=ypos, cex=1.2)
legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol)

###Photosynthesis vs Ci
plot(Photo~Ci, data=sundat, pch=16, col=suncol, ylim=c(0,25), xlim=c(0,350), xlab=cilab, ylab="", cex=1.25)
points(Photo~Ci, data=shadat, pch=16, col=shacol, cex=1.25)
with(aci_sun, {
  lines(Ci, lcl, lty=2, lwd=2,col="forestgreen")
  lines(Ci, ucl, lty=2, lwd=2,col="forestgreen")
  lines(Ci, pred, lty=1, lwd=2,col="forestgreen")
})
with(aci_sha, {
  lines(Ci, lcl, lty=2, lwd=2,col="yellow4")
  lines(Ci, ucl, lty=2, lwd=2,col="yellow4")
  lines(Ci, pred, lty=1, lwd=2,col="yellow4")
})
title(ylab=satlab, mgp=ypos, cex=1.2)
legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol)

 ###Photosynthesis vs Cc 
  plot(Photo~Cc, data=sundat, pch=16, col=suncol, ylim=c(0,25), xlim=c(0,350), xlab=cclab, ylab="", cex=1.25)
  points(Photo~Cc, data=shadat, pch=16, col=shacol, cex=1.25)
  with(acc_sun, {
    lines(Cc, lcl, lty=2, lwd=2,col="forestgreen")
    lines(Cc, ucl, lty=2, lwd=2,col="forestgreen")
    lines(Cc, pred, lty=1, lwd=2,col="forestgreen")
  })
  with(acc_sha, {
    lines(Cc, lcl, lty=2, lwd=2,col="yellow4")
    lines(Cc, ucl, lty=2, lwd=2,col="yellow4")
    lines(Cc, pred, lty=1, lwd=2,col="yellow4")
  })
  title(ylab=satlab, mgp=ypos, cex=1.2)
  legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol) 
```

* Photosynthesis and Ci calculated from leaf ${\delta}$^13^C

```{r cibar}
###Photosynthesis vs cibar
plot(Photo~ci_bar, data=sundat, pch=16, col=suncol, ylim=c(0,25), 
     xlim=c(0,350), xlab=cibarlab2, ylab="", cex=1.25)
#shade
points(Photo~ci_bar, data=shadat, pch=16, col=shacol, cex=1.25)
  with(acib_sun, {
    lines(ci_bar, lcl, lty=2, lwd=2,col="forestgreen")
    lines(ci_bar, ucl, lty=2, lwd=2,col="forestgreen")
    lines(ci_bar, pred, lty=1, lwd=2,col="forestgreen")
  })
  with(acib_sha, {
    lines(ci_bar, lcl, lty=2, lwd=2,col="yellow4")
    lines(ci_bar, ucl, lty=2, lwd=2,col="yellow4")
    lines(ci_bar, pred, lty=1, lwd=2,col="yellow4")
  })
title(ylab=satlab, mgp=ypos, cex=1.2)
legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol) 
```

### <u>Shade versus Sun leaves</u>

This set of figures attempts to show how sun and shade leaves were fundamentally different.

* ACi curves were measured at saturating light for both sun and shade leaves.

```{r, results='asis'}
aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
names(aci_leaf) <-c("Leaf", "Temperature", "Vcmax", "Jmax", "Vcmax_se", "Jmax_se")

aci_leaf2 <- aci_leaf[,c(1,2,3,5,4,6)]

panderOptions("digits", 3)
pander(aci_leaf2, caption="Leaf Biochemistry")
```

* Here are predicted ACi curves for leaf types with temperature treatments. Both Vc<sub>max</sub> and J<sub>max</sub> are reduced in shade leaves with little effect of the temperature treatment. ACi curves were taken once towards the end of the experiment on well watered trees.

```{r,results='asis'}
  #simulate ACi curves for each leaf, ambient and elevated T
  sunAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[3,3], Jmax=aci_leaf[3,4],PPFD=1408)
  sunET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[4,3], Jmax=aci_leaf[4,4], PPFD=1408)
  shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[1,3], Jmax=aci_leaf[1,4], PPFD=375)
  shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[2,3], Jmax=aci_leaf[2,4],PPFD=375)

#plot
plot(sunAT_sim$Ci, sunAT_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab=cilab, ylab="", type="l", lwd=2)
  points(sunET_sim$Ci, sunET_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
  points(shaAT_sim$Ci, shaAT_sim$ALEAF, col="yellowgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2)
  points(shaET_sim$Ci, shaET_sim$ALEAF, col="yellowgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
  legend("bottomright", leglab2, lty=c(1,2,1,2), lwd=2,col=colaci, pt.bg=col4,inset = 0.03)
  title(ylab=satlab, mgp=ypos)
```

* This figure shows the relationship between leaf N on an area basis and ${\delta}$^13^C between leaf types (chamber means).

```{r,results='asis'}
##this has droughtleaves
palette(c("yellowgreen", "green4"))
par(mar=c(5,5,2,2))
plot(c13.mean ~ leafN_area.mean, data=leaftraits, pch=16, col=as.factor(leaf), ylim=c(-35,-25), 
                 xlim=c(1,3.5), cex=1.25, xlab=narealab, ylab=c13lab)
legend("bottomright", leaflab, pch=16, col=c("yellowgreen", "green4"),pt.cex=1.5,inset = 0.03)

```

* Next is the light environment in which gas exchange measurements for sun and shade leaves were measured.  This data was generated with a handheld par sensor for each individual leaf and verified with ceptometer readings within the canopy at the height of the measured leaf. There was a ~75% reduction in PPFD from sun to shave leaves.

* The second figure shows leaf mass area for each gas exchange campaign.  Surprisingly, not very different.

```{r,results='asis'}
 #format function
 par<- parformat(par)

 par_leaf <- subset(par, ID !="shade-high")
 Morder <- c("Oct", "Dec", "Jan", "Feb", "Mar", "Apr")

 #data format for bar
 parbar <- subset(par_leaf, select = c("par", "month", "leaf_type"))
 parbar$month <- factor(parbar$month, levels = Morder)

 bar(par, c(leaf_type, month), parbar, col=c("yellowgreen", "green4"), xlab="", ylab="", ylim=c(0, 2000), 
     half.errbar=FALSE)
 title(ylab=parlab, mgp=ypos)

 ###LMA--------------------------------------------------------------------------------------------------
 lma <- read.csv("calculated_data/leaf_chemistry.csv")

 #data format for bar
 lmabar <- subset(lma, select = c("lma", "Month", "leaf"))
   lmabar$Month <- factor(lmabar$Month, levels = Morder)

 bar(lma, c(leaf, Month), lmabar, col=c("yellowgreen", "green4"), xlab="", ylab="", ylim=c(0, 185), half.errbar=FALSE)
 title(ylab=lmalab, mgp=ypos)

```

### 2015-03-20 functions and packages/packages.R

@@ -1,7 +1,6 @@
 require(doBy)
 require (lattice)
 require(lubridate)
-require(doBy)
 require (lattice)
 library(plyr)
 library(plotBy)

### 2015-03-20 master_scripts/sunshade_phys.R

@@ -1,24 +1,21 @@
 source("functions and packages/functions.R")
 source("master_scripts/plot_objects.R")
 source("functions and packages/packages.R")
+library(mgcv)
+library(lme4)
 
-#read in gm data set (no drought)
+#read in gm data set (no drought) and Cibar(discrimination)
 
-gmes <- read.csv("calculated_data/gmes_wellwatered.csv")
 Ci_bar <- read.csv("calculated_data/Ci_bar.csv")
 
-
-##calculate CC
-gmes$Cc<- with(gmes, Ci-Photo/gm)
-gmes$cc_ci<- with(gmes, Ci/Cc)
-
-##need to combine Ci_bar with CC and gm
+gmes <- read.csv("calculated_data/gmes_wellwatered.csv")
+  ##calculate CC
+  gmes$Cc<- with(gmes, Ci-Photo/gm)
+  gmes$cc_ci<- with(gmes, Ci/Cc)
 
 ###get average by id
 gm_agg <- summaryBy(Photo+Cond+Ci+Cc+gm+VpdL+xsi+DELTA+cc_ci ~ chamber+id+leaf +light+temp+leaflight, 
                     data=gmes, FUN=mean, keep.names=TRUE)
-
-
 
 ##remove shade-high
 gm_sunsha <- gm_agg[gm_agg$leaflight != "shade-high",]
@@ -32,9 +29,6 @@ gm_c13 <- merge(gm_sunsha, Ci_bar[, c(2,8)], by="id")
 
   
 ###Photosynthesis vs gs (need to fit sun with something else)------------------------------------------------
-  library(mgcv)
-  library(lme4)
-  
   ###simple plot use smoothplot from RD and get CI polygon
   palette(c(shacol, suncol))
 
@@ -89,210 +83,194 @@ gm_c13 <- merge(gm_sunsha, Ci_bar[, c(2,8)], by="id")
   dev.copy2pdf(file="master_scripts/figures/photo_gs.pdf")
   dev.off()
 
+##bootstrap plots of confidence intervals with chamber as random effect---------------------------------
+  library(lmerTest)
+  library(plyr)
   
+  palette(c(shacol, suncol))
   
[…]

### 2015-03-20 old gm calc/paired.R

@@ -38,13 +38,16 @@ title(ylab=satlab, mgp=ypos, cex=1.2)
 legend("topright", sslab, pch=21, pt.bg=cl, pt.cex=2, bg="white",inset = 0.03) 
 
 
+#gm data----------------------------------------------------------------------------------------
+
 #paired id dfr
 idDF <- data.frame(id = unique(gm$id))
 
-#library(reshape)
-#gmCst <- cast(gm_id, id ~ variable, value= "gm")
+#summary of spot measurements for each id
+gm_id <- summaryBy(gm~ variable+id, data=gm, FUN=c(mean), keep.names=TRUE)
+gm_sp <- dlply(gm_id, .(variable), function(x) merge(x, idDF, all = TRUE))
 
-#gm data----------------------------------------------------------------------------------------
+
 
 #make an emptyr plot for presentation
 png(filename = "output/presentations/emptyplot.png", width = 11, height = 8.5, units = "in", res= 400)
@@ -59,22 +62,6 @@ title(ylab=cilab, mgp=ypos, cex=1.2)
 legend("topright", pairlab, pch=21, pt.bg=paircol, pt.cex=2,bg="white",inset = 0.03)  
 dev.off()
 
-
-x <- rnorm(100)
-y <- x + rnorm(100)
-lmfit <- lm(y~x)
-plot(x,y,xlim=c(-3.5,3.5))
-ablineclip(lmfit,x1=-2,x2=2,lty=2)
-ablineclip(h=0,x1=-2,x2=2,lty=3,col="red")
-ablineclip(v=0,y1=-2.5,y2=1.5,lty=4,col="green")
-
-
-
-
-
-#summary of spot measurements for each id
-gm_id <- summaryBy(gm~ variable+id, data=gm, FUN=c(mean), keep.names=TRUE)
-gm_sp <- dlply(gm_id, .(variable), function(x) merge(x, idDF, all = TRUE))
 
 plot(gm_sp[[4]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[3], cex=2.5, ylab="", xlab=suatgm, xlim=c(0,1), ylim=c(0,1)) 
 abline(0,1, lty=2)

### 2015-03-30 markdown/wtc_sunshade.Rmd
put captions below figures

@@
 
-* This figure shows the instantaneous response of g<sub>m</sub> in shade leaves when the lights are 'turned on'
-
 ```{r gm lights on}
@@
 ```
-
-* Here is a simple version of the paired comparisons.
-Only shade low light and sun leaves are used (watered and drought treatments together).
-All leaf types are compared to sun leaves at ambient temperature treatment (x axis)
+* This figure shows the instantaneous response of g<sub>m</sub> in shade leaves when the lights are 'turned on'
 
@@
 ```
-
-* Here we see a slight reduction in g<sub>m</sub> with increase in temperature, but only in sun leaves
+* Here is a simple version of the paired comparisons.
+Only shade low light and sun leaves are used (watered and drought treatments together).
+All leaf types are compared to sun leaves at ambient temperature treatment (x axis)
 
@@
 ```
+* Here we see a slight reduction in g<sub>m</sub> with increase in temperature, but only in sun leaves
 
@@
 
-* ACi curves were measured at saturating light for both sun and shade leaves.
-
 ```{r, results='asis'}
@@
 ```    
-
-* Here are predicted ACi curves for leaf types with temperature treatments.
-Both Vc<sub>max</sub> and J<sub>max</sub> are reduced in shade leaves with little effect of the temperature treatment.
-ACi curves were taken once towards the end of the experiment on well watered trees.
+* ACi curves were measured at saturating light for both sun and shade leaves.
 
@@
 ```
-
-* This figure shows the relationship between leaf N on an area basis and ${\delta}$^13^C between leaf types (chamber means).
+* Here are predicted ACi curves for leaf types with temperature treatments.
+Both Vc<sub>max</sub> and J<sub>max</sub> are reduced in shade leaves with little effect of the temperature treatment.
+ACi curves were taken once towards the end of the experiment on well watered trees.
 
@@
 ```
-
-* Next is the light environment in which gas exchange measurements for sun and shade leaves were measured.
-This data was generated with a handheld par sensor for each individual leaf and verified with ceptometer readings within the canopy at the height of the measured leaf.
-There was a ~75% reduction in PPFD from sun to shave leaves.
-
-* The second figure shows leaf mass area for each gas exchange campaign.  Surprisingly, not very different.
+* This figure shows the relationship between leaf N on an area basis and ${\delta}$^13^C between leaf types (chamber means).
 
@@
  title(ylab=parlab, mgp=ypos)
+```
 
- ###LMA--------------------------------------------------------------------------------------------------
+* Next is the light environment in which gas exchange measurements for sun and shade leaves were measured.
+This data was generated with a handheld par sensor for each individual leaf and verified with ceptometer readings within the canopy at the height of the measured leaf.
+There was a ~75% reduction in PPFD from sun to shave leaves.
+
+```{r}
+ ###LMA-
  lma <- read.csv("calculated_data/leaf_chemistry.csv")
@@
  title(ylab=lmalab, mgp=ypos)
+```
 
-```
+* The second figure shows leaf mass area for each gas exchange campaign.  Surprisingly, not very different.

### 2015-04-01 markdown/wtc_sunshade.Rmd
made two plotting scripts for markdown multi panel plots
more updates to markdown

--- 2015-03-20 markdown/wtc_sunshade.Rmd
@@
 ---
-title: "Sun v Shade leaves (WTCIII)"
+title: "Sun vs Shade leaves (WTCIII)"
 author: "Court Campany"
-date: "March 17, 2015"
+date: "April 1, 2015"
 output: html_document
@@
 
-### <u>Mesophyll Conducatnce</u>
+#### <u>Design</u>
+* WTCIII was a warming experiemnt with ambient and +3°C warming (n=6).  A drought manipulation was added in the last few months of the experiment.
+* Sun leaves were measured in the upper canopy at their current PPFD.  Shade leaves were measured in the lower canopy, first at their current PPFD and then again at sun leaf PPFD.
+* New fully expanded leaves were measured. Shade leaves were sampled into the middle canopy as the experiment progressed to avoid issues with leaf age.
+* Leaf physiological measurements were taken at their current temperature (ambient, +3°C) for paired comparisons across leaf types.
+Thus, data below represent a large range of temperatures from ~18-35°C.
+* For simplicity I have generalized most of the results below to just sun vs.
+shade leaves.
+So effects of the warming treatment are often not included and there is no standarizatiopn of leaf temperature.
+* Unless it is explicitly stated I have left off the drought treatment as well.  
+* Data below are mostly overall means across the 6 campaigns.  We can tease apart temperature/drought effects later.
 
-* We took leaf measurements at their current temperature (ambient, +3C) for paired comparisons across leaf types at a large range of temperatures.
+#### <u>Were Sun and Shade leaves actually different?</u>
 
-* Sun leaves were taken from upper canopy at their current light level.  Shade leaves were taken from lower canopy, first at their current light level and then again at sun leaf PPFD.
+```{r, results='asis'}
+aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
+names(aci_leaf) <-c("Leaf", "Temperature", "Vcmax", "Jmax", "Vcmax_se", "Jmax_se")
+
+aci_leaf2 <- aci_leaf[,c(1,2,3,5,4,6)]
+
+panderOptions("digits", 3)
+pander(aci_leaf2, caption="Leaf Biochemistry")
+```
+
+```{r,results='asis'}
+  #simulate ACi curves for each leaf, ambient and elevated T
+  sunAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[3,3], Jmax=aci_leaf[3,4],PPFD=1408)
+  sunET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[4,3], Jmax=aci_leaf[4,4], PPFD=1408)
+  shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[1,3], Jmax=aci_leaf[1,4], PPFD=375)
+  shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[2,3], Jmax=aci_leaf[2,4],PPFD=375)
+
+#plot
+plot(sunAT_sim$Ci, sunAT_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab=cilab, ylab="", type="l", lwd=2, ylim=c(0,30))
+  points(sunET_sim$Ci, sunET_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
+  points(shaAT_sim$Ci, shaAT_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2)
+  points(shaET_sim$Ci, shaET_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
+  legend("bottomright", leglab2, lty=c(1,2,1,2), lwd=2,col=colaci, pt.bg=col4,inset = 0.03)
+  title(ylab=satlab, mgp=ypos)
+```
+
+* ACi curves were measured at saturating light for both sun and shade leaves. 
+* Here are predicted ACi curves for leaf types with temperature treatments.
+Both Vc<sub>max</sub> and J<sub>max</sub> are reduced in shade leaves with little effect of the temperature treatment.
+ACi curves were taken once towards the end of the experiment on well watered trees
+
+#### <u>Mesophyll Conducatnce</u>
 
@@
 
-* For simplicity I have generalized most of the results below to just sun vs shade leaves (temperature not standardized).
+#### g<sub>m</sub> by leaf type and light environment
 
-* Further analysis is also needed to tease apart any effects of temperature treatment as data below are overall means across the 6 campaigns
+* The two tables below are means and standard errors of g<sub>m</sub> across all campaigns.
 
-####g<sub>m</sub> by leaf type and light environment
-
-*  Below are means and standard errors of g<sub>m</sub> across all campaigns.
-
-* Here I separated the well watered treatment from the drought treatment.
-I'm more confident in the well watered values as I was often on the cutoff point with drawdown and xsi for drought leaves.
+* The well-watered treatment from the drought treatment are seperated.
+I'm more confident in the well-watered values as I was often on the cutoff point with drawdown and xsi for drought leaves.
 They do show similar patterns but g<sub>m</sub> seems to be higher in drought trees
 
-* Shade g<sub>m</sub> is always lower than sun leaves but when the lights are turned on they have in instantaneous response in g<sub>m</sub> that resembles sun leaves
+* Shade leaf g<sub>m</sub> is always lower than sun leaves but when the lights are 'turned on' they have in instantaneous response in g<sub>m</sub> that resembles sun leaves.
 
@@
 
-* This figure shows the instantaneous response of g<sub>m</sub> in shade leaves when the lights are 'turned on'
-
 ```{r gm lights on}
 #plot low vs high light shade leaves
-plot(Photo~gm, data=gm_water, subset=leaflight=="shade-high", pch=16, col=lightscol, ylim=c(0,25), xlim=c(0,.4), 
+plot(Photo~gm, data=gm_water, subset=leaflight=="shade-high", pch=16, col=lightscol, ylim=c(0,30), xlim=c(0,.4), 
      xlab=gmlab, ylab="", cex=1.25)
@@
 
-* Here is a simple version of the paired comparisons.
-Only shade low light and sun leaves are used (watered and drought treatments together).
-All leaf types are compared to sun leaves at ambient temperature treatment (x axis)
+* This figure shows the instantaneous response of g<sub>m</sub> in shade leaves when the lights are 'turned on'.
 
@@
 par(mar=c(0,7,2,2))
-plot(gm_sp[[4]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[3], cex=2.5, ylab="", xlab="", xlim=c(0,.5), ylim=c(0,.5),
+plot(gm_sp[[4]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[3], cex=2.5, ylab="", xlab="", xlim=c(0,.4), ylim=c(0,.4),
      axes=FALSE) 
@@
 box()
-legend(.4,.45, pairlab[1], pch=21, pt.bg=paircol[1], pt.cex=2,bg="white",inset = 0.06)
+legend(.3,.35, pairlab[1], pch=21, pt.bg=paircol[1], pt.cex=2,bg="white",inset = 0.06)
 
@@
 par(mar=c(0,7,0,2))
-plot(gm_sp[[1]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[1], cex=2.5, ylab=gmlab, xlim=c(0,.5), ylim=c(0,.5),xlab="",
+plot(gm_sp[[1]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[1], cex=2.5, ylab=gmlab, xlim=c(0,.4), ylim=c(0,.4),xlab="",
      axes=FALSE)     
 abline(0,1, lty=2)
-legend(.4,.45, pairlab[2], pch=21, pt.bg=paircol[2], pt.cex=2,bg="white",inset = 0.06)
+legend(.3,.35, pairlab[2], pch=21, pt.bg=paircol[2], pt.cex=2,bg="white",inset = 0.06)
 axis(1, labels=FALSE) 
@@
 par(mar=c(5,7,0,2))
-plot(gm_sp[[2]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[5], cex=2.5, ylab="", xlim=c(0,.5), ylim=c(0,.5),xlab=suatgm)   
+plot(gm_sp[[2]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[5], cex=2.5, ylab="", xlim=c(0,.4), ylim=c(0,.4),xlab=suatgm)   
 abline(0,1, lty=2)
-legend(.4,.45, pairlab[3], pch=21, pt.bg=paircol[3], pt.cex=2,bg="white",inset = 0.06) 
+legend(.3,.35, pairlab[3], pch=21, pt.bg=paircol[3], pt.cex=2,bg="white",inset = 0.06) 
 ```
 
-* Here we see a slight reduction in g<sub>m</sub> with increase in temperature, but only in sun leaves
+* Here is a simple version of the paired leaf comparisons.
+Shade low light and sun leaves are used (watered and drought treatments are combined).
+Leaf types (with temperature treatment) are compared to sun leaves at ambient temperature as the control treatment (x axis).
 
@@
 
-### <u>Leaf Physiology</u>
-* Below are figures of sun and shade leaf physiology, keeping it simple by ignoring the temperature treatment. We can address that later.
+* This figure shows the relationship between g<sub>m</sub> and temperature with 95% confidence intervals.
+Here we see a slight reduction in g<sub>m</sub> with increases in temperature, but only significant in sun leaves.
 
-```{r A_gs}
-##calculate CC
-gmwet$Cc<- with(gmwet, Ci-Photo/gm)
-gmwet$cc_ci<- with(gmwet, Ci/Cc)
+### <u>Sun vs. Shade Leaf Physiology</u>
 
-###get average by id
-gm_agg <- summaryBy(Photo+Cond+Ci+Cc+gm+VpdL+xsi+DELTA+cc_ci ~ chamber+id+leaf +light+temp+leaflight, 
-                    data=gmwet, FUN=mean, keep.names=TRUE)
-
-##remove shade-high
-gm_sunsha <- gm_agg[gm_agg$leaflight != "shade-high",]
-  gm_sunsha <- droplevels(gm_sunsha)
-  gm_sunsha$id <- gsub("-high", "", gm_sunsha$id)
-  gm_sunsha$id <- gsub("-low", "", gm_sunsha$id)
-
-##merge
-gm_c13 <- merge(gm_sunsha, Ci_bar[, c(2,8)], by="id")
-  gm_c13$Cc_bar <- with(gm_c13, ci_bar-Photo/gm)
-
-###Photosynthesis vs gs
-
-  ####run gam models and then predict
-  #SUN leaves
-  sunmod <- gam(Photo ~ s(Cond, k=5), data=gm_c13, subset=leaflight=="sun-high")
-
-  #predict
-  #get apprpriate vector of gs from sun leaves
-  gsdat <- gm_c13[gm_c13$leaflight=="sun-high", "Cond"]
-
-  #generate sequence and then predict
-  gssun_seq <- seq(min(gsdat), max(gsdat), length=101)
-  gssun_pred <- predict(sunmod, newdata=data.frame(Cond=gssun_seq), se.fit=TRUE)
-
-  #ci and model fit
-  sunupr <- gssun_pred$fit + (2*gssun_pred$se.fit)
-  sunlwr <- gssun_pred$fit - (2*gssun_pred$se.fit)
-
-  #SHADE leaves
-  shamod <- gam(Photo ~ s(Cond, k=5), data=gm_c13, subset=leaflight=="shade-low")
-
-  #get apprpriate vector CC from sun leaves
-  gsdat2 <- gm_c13[gm_c13$leaflight=="shade-low", "Cond"]
-  #generate sequence and then predict
-  gssha_seq <- seq(min(gsdat2), max(gsdat2), length=101)
-  gssha_pred <- predict(shamod, newdata=data.frame(Cond=gssha_seq), type="link", se.fit=TRUE)
-
-  shaupr <- gssha_pred$fit + (2*gssha_pred$se.fit)
-  shalwr <- gssha_pred$fit - (2*gssha_pred$se.fit)
-
-  ###plot
-  plot(Photo~Cond, data=gm_c13, subset=leaflight=="sun-high", pch=16, col=suncol, ylim=c(0,25), 
-       xlim=c(0,.4), xlab=condlab, ylab="", cex=1.25)
-
-  lines(gssun_seq, sunupr, lty=2, lwd=2,col=suncol)
-  lines(gssun_seq, sunlwr, lty=2, lwd=2,col=suncol)
-  lines(gssun_seq, gssun_pred$fit, lty=1, lwd=2,col=suncol)
-
-  #shade
-  points(Photo~Cond, data=gm_c13, subset=leaflight=="shade-low", pch=16, col=shacol, cex=1.25)
-  lines(gssha_seq, shaupr, lty=2, lwd=2,col=shacol)
-  lines(gssha_seq, shalwr, lty=2, lwd=2,col=shacol)
-  lines(gssha_seq, gssha_pred$fit, lty=1, lwd=2,col=shacol)
-
-  title(ylab=satlab, mgp=ypos, cex=1.2)
-  legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol) 
+```{r A_gs, fig.width=12, fig.height=6}
+source("markdown/ags_plots.R")
 ```
 
-```{r leafphys}
+* Figures of photosynthesis and g<sub>s</sub>.
+Panel (a) is shown with 95% confidence intervals.
+Panel (b) shows predicted photosynthesis rates generated with g<sub>s</sub> values, mean leaf temperature and PPFD (sun/shade), and ACi parameters including ambient and elevated temperature treatments.
+Here we can see optimal stomatal behaviour in sun leaves but not in shade leaves.
+
+```{r a-gm, }
 source("calculated_data/bootstrap/readboot.R")
-
   ###Photosynthesis vs gm
-plot(Photo~gm, data=sundat, pch=16, col=suncol, ylim=c(0,25), xlim=c(0,.4), xlab=gmlab, ylab="", cex=1.25)
+plot(Photo~gm, data=sundat, pch=16, col=suncol, ylim=c(0,30), xlim=c(0,.4), xlab=gmlab, ylab="", cex=1.25)
   points(Photo~gm, data=shadat, pch=16, col=shacol, cex=1.25)
@@
 legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol) 
+```
 
+* This figure shows the relationship between photosynthesis and g<sub>m</sub> for sun and shade leaves with 95% confidence intervals.
+
+```{r leafphys,fig.width=12, fig.height=6}
 ###Photosynthesis vs Ci
-plot(Photo~Ci, data=sundat, pch=16, col=suncol, ylim=c(0,25), xlim=c(0,350), xlab=cilab, ylab="", cex=1.25)
+  shade <- gm_c13[gm_c13$leaf=="shade",]
+  sun <- gm_c13[gm_c13$leaf=="sun",]
+
+  #read in jmax vcmax
+  jvc<- read.csv("calculated_data/tdl_aci.csv")
+
+  #simulate ACi curves for each leaf, ambient and elevated T
+  sunAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[3,3], Jmax=jvc[3,4],PPFD=1407.959)
+  sunET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[4,3], Jmax=jvc[4,4], PPFD=1407.959)
+  shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[1,3], Jmax=jvc[1,4], PPFD=305)
+  shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[2,3], Jmax=jvc[2,4],PPFD=305)
+
+par(cex.axis=.96, cex.lab=1.2,mfrow=c(1,2),oma=c(0.1,0.1,0.1,0.1), las=1)
+
+par(mar=c(5,4,2,0))
+plot(Photo~Ci, data=sundat, pch=16, col=suncol, ylim=c(0,30), xlim=c(0,350), xlab=cilab, ylab="", cex=1.25)
 points(Photo~Ci, data=shadat, pch=16, col=shacol, cex=1.25)
@@
 title(ylab=satlab, mgp=ypos, cex=1.2)
-legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol)
+legend("topright", leaflab2, pch=16,inset = 0.03, col=leafcol) 
+text(x=.01, 29.5, "(a)", cex=1.2)
 
+par(mar=c(5,2.5,2,2)) 
+plot(sunAT_sim$Ci, sunAT_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab=cilab, ylab="", type="l",  lwd=2,ylim=c(0,30), xlim=c(0,1000))
+  points(sunET_sim$Ci, sunET_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
+  points(shaAT_sim$Ci, shaAT_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2)
+  points(shaET_sim$Ci, shaET_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
+  points(Photo~Ci, data=sundat, pch=16, col=suncol, cex=1.25)
+  points(Photo~Ci, data=shadat, pch=16, col=shacol, cex=1.25)
+
+  legend("bottomright", leglab2, lty=c(1,2,1,2), lwd=2,col=colaci, pt.bg=col4,inset = 0.03)
+  text(x=.01, 29.5, "(b)", cex=1.2)
+```
+
+* These figures show the relationship of photosynthesis to Ci.
+Panel (a) has 95% confidence intervals.
+Panel (b) has the predicted ACi curves using the biochemical parameters measured at saturating light but fitted with mean PPFD of sun and shade leaves.
+
+```{r}
  ###Photosynthesis vs Cc 
-  plot(Photo~Cc, data=sundat, pch=16, col=suncol, ylim=c(0,25), xlim=c(0,350), xlab=cclab, ylab="", cex=1.25)
+  plot(Photo~Cc, data=sundat, pch=16, col=suncol, ylim=c(0,30), xlim=c(0,350), xlab=cclab, ylab="", cex=1.25)
   points(Photo~Cc, data=shadat, pch=16, col=shacol, cex=1.25)
@@
 
-* Photosynthesis and Ci calculated from leaf ${\delta}$^13^C
+* This figures show the relationship of photosynthesis to Cc with 95% confidence intervals, calculated with g<sub>m</sub> data
 
-```{r cibar}
-###Photosynthesis vs cibar
-plot(Photo~ci_bar, data=sundat, pch=16, col=suncol, ylim=c(0,25), 
-     xlim=c(0,350), xlab=cibarlab2, ylab="", cex=1.25)
-#shade
-points(Photo~ci_bar, data=shadat, pch=16, col=shacol, cex=1.25)
-  with(acib_sun, {
-    lines(ci_bar, lcl, lty=2, lwd=2,col="forestgreen")
-    lines(ci_bar, ucl, lty=2, lwd=2,col="forestgreen")
-    lines(ci_bar, pred, lty=1, lwd=2,col="forestgreen")
-  })
-  with(acib_sha, {
-    lines(ci_bar, lcl, lty=2, lwd=2,col="yellow4")
-    lines(ci_bar, ucl, lty=2, lwd=2,col="yellow4")
-    lines(ci_bar, pred, lty=1, lwd=2,col="yellow4")
-  })
-title(ylab=satlab, mgp=ypos, cex=1.2)
-legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol) 
-```
+### <u>Shade versus Sun leaves, continued</u>
 
-### <u>Shade versus Sun leaves</u>
-
-This set of figures attempts to show how sun and shade leaves were fundamentally different.
-
-* ACi curves were measured at saturating light for both sun and shade leaves.
-
-```{r, results='asis'}
-aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
-names(aci_leaf) <-c("Leaf", "Temperature", "Vcmax", "Jmax", "Vcmax_se", "Jmax_se")
-
-aci_leaf2 <- aci_leaf[,c(1,2,3,5,4,6)]
-
-panderOptions("digits", 3)
-pander(aci_leaf2, caption="Leaf Biochemistry")
-```
-
-* Here are predicted ACi curves for leaf types with temperature treatments.
-Both Vc<sub>max</sub> and J<sub>max</sub> are reduced in shade leaves with little effect of the temperature treatment.
-ACi curves were taken once towards the end of the experiment on well watered trees.
+This next set of figures attempts to show how sun and shade leaves were fundamentally different.
 
 ```{r,results='asis'}
-  #simulate ACi curves for each leaf, ambient and elevated T
-  sunAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[3,3], Jmax=aci_leaf[3,4],PPFD=1408)
-  sunET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[4,3], Jmax=aci_leaf[4,4], PPFD=1408)
-  shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[1,3], Jmax=aci_leaf[1,4], PPFD=375)
-  shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[2,3], Jmax=aci_leaf[2,4],PPFD=375)
+##this has droughtleaves
+palette(c("yellow4", "forestgreen"))
+par(mar=c(5,5,2,2))
+plot(c13.mean ~ leafN_area.mean, data=leaftraits, pch=16, col=as.factor(leaf), ylim=c(-35,-25), 
+                 xlim=c(1,3.5), cex=1.25, xlab=narealab, ylab=c13lab)
+legend("bottomright", leaflab, pch=16, col=c("yellow4", "green4"),pt.cex=1.5,inset = 0.03)
 
-#plot
-plot(sunAT_sim$Ci, sunAT_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab=cilab, ylab="", type="l", lwd=2)
-  points(sunET_sim$Ci, sunET_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
-  points(shaAT_sim$Ci, shaAT_sim$ALEAF, col="yellowgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2)
-  points(shaET_sim$Ci, shaET_sim$ALEAF, col="yellowgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
-  legend("bottomright", leglab2, lty=c(1,2,1,2), lwd=2,col=colaci, pt.bg=col4,inset = 0.03)
-  title(ylab=satlab, mgp=ypos)
 ```
@@
 * This figure shows the relationship between leaf N on an area basis and ${\delta}$^13^C between leaf types (chamber means).
-
-```{r,results='asis'}
-##this has droughtleaves
-palette(c("yellowgreen", "green4"))
-par(mar=c(5,5,2,2))
-plot(c13.mean ~ leafN_area.mean, data=leaftraits, pch=16, col=as.factor(leaf), ylim=c(-35,-25), 
-                 xlim=c(1,3.5), cex=1.25, xlab=narealab, ylab=c13lab)
-legend("bottomright", leaflab, pch=16, col=c("yellowgreen", "green4"),pt.cex=1.5,inset = 0.03)
-
-```
-
-* Next is the light environment in which gas exchange measurements for sun and shade leaves were measured.
-This data was generated with a handheld par sensor for each individual leaf and verified with ceptometer readings within the canopy at the height of the measured leaf.
-There was a ~75% reduction in PPFD from sun to shave leaves.
-
-* The second figure shows leaf mass area for each gas exchange campaign.  Surprisingly, not very different.
 
@@
 
- bar(par, c(leaf_type, month), parbar, col=c("yellowgreen", "green4"), xlab="", ylab="", ylim=c(0, 2000), 
+ bar(par, c(leaf_type, month), parbar, col=c("yellow4", "forestgreen"), xlab="", ylab="", ylim=c(0, 2000), 
      half.errbar=FALSE)
  title(ylab=parlab, mgp=ypos)
+```
 
- ###LMA--------------------------------------------------------------------------------------------------
+* Next is the light environment in which gas exchange measurements for sun and shade leaves were measured.
+This data was generated with a handheld par sensor for each individual leaf and verified with ceptometer readings within the canopy at the height of the measured leaf.
+There was a ~75% reduction in PPFD from sun to shave leaves.
+
+```{r}
+ ###LMA-
  lma <- read.csv("calculated_data/leaf_chemistry.csv")
@@
 
- bar(lma, c(leaf, Month), lmabar, col=c("yellowgreen", "green4"), xlab="", ylab="", ylim=c(0, 185), half.errbar=FALSE)
+ bar(lma, c(leaf, Month), lmabar, col=c("yellow4", "forestgreen"), xlab="", ylab="", ylim=c(0, 185), half.errbar=FALSE)
  title(ylab=lmalab, mgp=ypos)
+```
 
-```
+* This figure shows leaf mass area for each gas exchange campaign.  Surprisingly, not very different between leaf types.

### 2015-04-01 master_scripts/aci_predict.R

@@ -69,8 +69,8 @@ Ci_bar <- read.csv("calculated_data/Ci_bar.csv")
   #simulate ACi curves for each leaf, ambient and elevated T
   sunAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[3,3], Jmax=jvc[3,4],PPFD=mean(sun$PARi))
   sunET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[4,3], Jmax=jvc[4,4], PPFD=mean(sun$PARi))
-  shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[1,3], Jmax=jvc[1,4], PPFD=375)
-  shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[2,3], Jmax=jvc[2,4],PPFD=375)
+  shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[1,3], Jmax=jvc[1,4], PPFD=mean(shade$PARi))
+  shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[2,3], Jmax=jvc[2,4],PPFD=mean(shade$PARi))
   
   
   windows(10,8)

### 2015-04-02 markdown/wtc_sunshade.Rmd
new file to compare leaf N with with vcmax
final touches to markdown

removed duplciates in aci sun (need to implement this elsewhere

@@
 author: "Court Campany"
-date: "April 1, 2015"
+date: "April 2, 2015"
 output: html_document
@@
 #### <u>Design</u>
-* WTCIII was a warming experiemnt with ambient and +3°C warming (n=6).  A drought manipulation was added in the last few months of the experiment.
-* Sun leaves were measured in the upper canopy at their current PPFD.  Shade leaves were measured in the lower canopy, first at their current PPFD and then again at sun leaf PPFD.
+* WTCIII was a warming experiment with ambient and +3°C warming (n=6).  A drought manipulation was added in the last few months of the experiment.
+* Sun leaves were measured in the upper canopy at their current PPFD.  Shade leaves were measured in the lower/middle canopy, first at their current PPFD and then again at sun leaf PPFD.
 * New fully expanded leaves were measured. Shade leaves were sampled into the middle canopy as the experiment progressed to avoid issues with leaf age.
@@
 shade leaves.
-So effects of the warming treatment are often not included and there is no standarizatiopn of leaf temperature.
+So effects of the warming treatment are often not included and there is no standardization of leaf temperature.
 * Unless it is explicitly stated I have left off the drought treatment as well.  
 * Data below are mostly overall means across the 6 campaigns.  We can tease apart temperature/drought effects later. 
+* The green/brown color scheme always refers to sun leaves and shade leaves at low light
 
@@
 
+```
+
+* ACi curves were measured at saturating light for both sun and shade leaves.
+
+```{r, aci plot}
 #plot
@@
 
-* ACi curves were measured at saturating light for both sun and shade leaves. 
 * Here are predicted ACi curves for leaf types with temperature treatments.
-Both Vc<sub>max</sub> and J<sub>max</sub> are reduced in shade leaves with little effect of the temperature treatment.
+Both Vc<sub>max</sub> and J<sub>max</sub> are reduced in shade leaves with little effect of the temperature treatment(AT/ET).
 ACi curves were taken once towards the end of the experiment on well watered trees
@@
 
-* The well-watered treatment from the drought treatment are seperated.
-I'm more confident in the well-watered values as I was often on the cutoff point with drawdown and xsi for drought leaves.
-They do show similar patterns but g<sub>m</sub> seems to be higher in drought trees
+* The well-watered treatment from the drought treatment are separated.
+I'm more confident in the well-watered values as I was often on the cutoff point with drawdown and thus xsi for drought leaves.
+They show similar patterns but g<sub>m</sub> seems to be higher in drought trees
 
-* Shade leaf g<sub>m</sub> is always lower than sun leaves but when the lights are 'turned on' they have in instantaneous response in g<sub>m</sub> that resembles sun leaves.
+* g<sub>m</sub> in shade is always lower than in sun but when the lights are 'turned on' shade leaves show an instantaneous response in g<sub>m</sub> that resembles sun leaves.
 
@@
 Panel (b) shows predicted photosynthesis rates generated with g<sub>s</sub> values, mean leaf temperature and PPFD (sun/shade), and ACi parameters including ambient and elevated temperature treatments.
-Here we can see optimal stomatal behaviour in sun leaves but not in shade leaves.
+Here we can see optimal stomatal behavior in sun leaves but not in shade leaves.
 
@@
                  xlim=c(1,3.5), cex=1.25, xlab=narealab, ylab=c13lab)
-legend("bottomright", leaflab, pch=16, col=c("yellow4", "green4"),pt.cex=1.5,inset = 0.03)
+legend("bottomright", leaflab2, pch=16, col=c("green4","yellow4"),inset = 0.03)
 
@@
 
-* Next is the light environment in which gas exchange measurements for sun and shade leaves were measured.
+* Here is the light environment in which gas exchange measurements for sun and shade leaves were measured.
 This data was generated with a handheld par sensor for each individual leaf and verified with ceptometer readings within the canopy at the height of the measured leaf.

### 2015-04-02 master_scripts/leafN_vcmax.R

@@ -1,0 +1,39 @@
+source("master_scripts/plot_objects.R")
+
+###need to correlate vcmax with leaf N
+
+###get vcmax value per chamber per leaf
+
+aciparam <- read.csv("calculated_data/aciparameters.csv")
+
+
+### get leaf N by chamber
+leafN <- read.csv("calculated_data/leaf_chemistry.csv")
+#no drought
+leafN2 <- leafN[leafN$drydown != "drought",]
+
+library(doBy)
+palette(c(shacol, suncol))
+
+Nagg <- summaryBy(leafN_area ~ chamber + leaf, data=leafN2)
+
+N_aci <- merge(aciparam, Nagg, by= c("chamber", "leaf"))
+
+sun <- N_aci[N_aci$leaf=="sun",]
+shade <- N_aci[N_aci$leaf=="shade",]
+
+
+###stats
+
+library(mgcv)
+library(lme4)
+library(lmerTest)
+library(plotrix)
+
+aciN_sun_lm <- lm(Vcmax~ leafN_area.mean, data=sun)
+aciN_sha_lm <- lm(Vcmax~ leafN_area.mean, data=shade)
+
+
+plot(Vcmax~leafN_area.mean, data=N_aci, col=as.factor(leaf), xlim=c(1,4), ylim=c(0, 130), pch=16)
+ablineclip(aciN_sun_lm, x1=min(sun$leafN_area.mean), x2=max(sun$leafN_area.mean), col=suncol, lwd=2)
+ablineclip(aciN_sha_lm, x1=min(shade$leafN_area.mean), x2=max(shade$leafN_area.mean), col=shacol, lwd=2)

### 2015-04-02 scripts/aci_wtc.R

@@ -60,6 +60,9 @@ sun_coefs <- sun_coefs[chamberorder,]
 #write a df with coefs for sun and shade leaves
 tdlaci <- rbind(sun_coefs, shade_coefs)
 tdlaci <- merge(tdlaci, plotsumm, by = "chamber")
+test <- tdlaci[-c(8, 13),]
+
+write.csv(test, "calculated_data/aciparameters.csv", row.names=FALSE)
 
 aci_means2 <- summaryBy(Vcmax+Jmax ~ leaf , data = tdlaci,  FUN=c(mean,se))
 

### 2015-06-01 WTC3/WTC3/index.Rmd
removed lock and added folders and files

---
title       : Are tree canopies optimized for carbon gain?
subtitle    : How wasteful water use in shade leaves constrain theoretical relationships of photosynthesis and resource distribution in Eucalyptus trees.
author      : Courtney Campany
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : standalone # {standalone, draft}
knit        : slidify::knit2slides

---

##slide1

---

##slide2

### 2015-06-01 presentation/example.Rmd
removed lock and added folders and files

---
title: "Untitled"
author: "Court Campany"
date: "May 26, 2015"
output: ioslides_presentation
---

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

# In the morning

## Getting up | What I like to do first thing

- Turn off alarm
- Get out of bed

## Breakfast

- Eat eggs
- Drink coffee

# In the evening

## Dinner

- Eat spaghetti
- Drink wine

### 2015-06-01 presentation/workflowtalk.Rmd
removed lock and added folders and files

---
title: "Project management and workflow"
author: Remko Duursma
date: 17 April 2015
output: ioslides_presentation
---

## Reproducible research

- Organize your work so that you have **everything** in a script
- The script reproduces all of your work, when run from a **clean workspace**
- Outputs (figures, processed datasets) are **disposable**, your scripts can always re-produce the output

![](chamber.jpg)

## A few key principles (1)

- Organize code and data in a single folder for each 'project'
- Do not mix code for separate projects
- Keep raw (original) data in a sub-folder, and **never modify raw data**

- Use *projects* in Rstudio to manage files and your workspace

## A few key principles (2)

- Write functions as much as possible
- Keep code that defines functions separate from other code
- Use a logical folder structure to organize files in a project (example later, and see Chapter 11)

![](../screenshots/folderstructure.png)

## Project size

- Try to avoid very large projects, instead split them into more **manageable chunks**.

- As a rule of thumb, a 'project' is about the size of the analysis for a single manuscript.

- Over time, projects always have the tendency to grow in size. If you archive old work, try to get used to: 
    + **deleting garbage** code and/or files that have accumulated
    + **completely starting over**. Sometimes it is best to reorganize your work completely.

## Folder structure

To keep raw data separate from scripts, functions, and outputs, a good folder structure is important. Below is an example, but this is of course flexible, and depends on the type of project.

![](../screenshots/folderstructuresmall.png)

In this simple example, we keep functions and scripts in the `R` folder, the raw data files (normally as CSV) in `rawdata`, and output is sent to the `output` folder.

## Working with folders

### Reading raw data

```{r, eval=FALSE}
allom <- read.csv("rawdata/allometry.csv")
```

### Saving a pdf

```{r, eval=FALSE}
pdf("output/Figure1.pdf")
plot(height ~ diameter, data=allom)
dev.off()
```

## Calling other scripts {.smaller}

I like to have a single 'master script', that loads other scripts that do particular bits, like read and clean the raw data, make figures, and fit linear models. This script may look like the following.

```{r eval=FALSE}
# Load packages
library(gplots)
library(car)

# Load functions
source("R/functions.R")

# Read raw data and add new variables
source("R/readdata.R")

# Make figures
source("R/figures.R")

# Do stats
source("R/linearmodels.R")
```

## Projects in Rstudio

Finally, I strongly recommend using **projects in Rstudio**.

- When you setup a project, a small file is added to your folder (e.g. `myproject.Rproj`)
- When you open the project, you start with a clean workspace, in the correct working directory
- So no more: `setwd()` and `rm(list=ls())` in your scripts
- This still works when you move the entire folder to a new location

### 2015-06-01 presentation/wtcIII_talk.Rmd
removed lock and added folders and files

---
title: "Are tree canopies optimized for carbon gain?"
subtitle: How wasteful water use in shade leaves constrain theoretical relationships of photosynthesis and resource distribution in Eucalyptus trees.
author: Courtney Campany, Mark Tjoelker and Remko Duursma
date: ''
output: ioslides_presentation
widescreen: yes
smaller: yes
transition: 0.4
---

```{r knitr_options , include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=6,
echo=FALSE, message=FALSE,
fig.align='center', warning=FALSE)
```

```{r, results='asis'}
library(knitr)
opts_knit$set(root.dir = '../')
```

```{r global data/sourcing , include=FALSE}
source("functions and packages/packages_md.R")
source("functions and packages/functions.R")
source("functions and packages/plot_objects_all.R")

gmes <- read.csv("calculated_data/gmes_WTC.csv")
gmpair<- read.csv("calculated_data/gmes_wtc_pair.csv")
gmwet <- read.csv("calculated_data/gmes_wellwatered.csv")
Ci_bar <- read.csv("calculated_data/Ci_bar.csv")
 ###for analysis first subset well watered and drought treatments
 gm_drought <- gmes[gmes$drydown == "drought",]
 gm_water <- gmes[gmes$drydown == "control",]

leaftraits <- read.csv("calculated_data/leaftraits_summary.csv") 
par <- read.csv("raw data/par.csv")
treatments <- read.csv("raw data/chamber_trt.csv")

aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
names(aci_leaf) <-c("Leaf", "Temperature", "Vcmax", "Jmax", "Vcmax_se", "Jmax_se")

  #simulate ACi curves for each leaf, ambient and elevated T
  sunAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[3,3], Jmax=aci_leaf[3,4],PPFD=1408)
  sunET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[4,3], Jmax=aci_leaf[4,4], PPFD=1408)
  shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[1,3], Jmax=aci_leaf[1,4], PPFD=375)
  shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[2,3], Jmax=aci_leaf[2,4],PPFD=375)
```

---
## When scaling to the canopy...Don't LEAF things to chance

- Trees can't put all leaves in the sun
- Drives changes in anatomy, morphology and physiology
- results in ↓ photosynthetic capacity in shade leaves

![](leaf_light.jpg)

- This should result in a gradient in resources available for photosynthetic with light....in theory

---

```{r, aci plot}
#plot
plot(sunAT_sim$Ci, sunAT_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab=cilab, ylab="", type="l", lwd=2, ylim=c(0,30))
  points(sunET_sim$Ci, sunET_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
  points(shaAT_sim$Ci, shaAT_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2)
  points(shaET_sim$Ci, shaET_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
  legend("bottomright", leglab2, lty=c(1,2,1,2), lwd=2,col=colaci, pt.bg=col4,inset = 0.03)
  title(ylab=satlab, mgp=ypos)
```
---

### 2015-06-01 WTC3/WTC3/WTC3/startslidify.R

@@ -1,0 +1,6 @@
+library(devtools)
+install_github('slidify', 'ramnathv')
+install_github('slidifyLibraries','ramnathv')
+
+library(slidify)
+author('WTC3')

### 2015-06-01 makepngs/sunshade_leaf.R

@@ -1,0 +1,67 @@
+
+source("functions and packages/packages_md.R")
+source("functions and packages/functions.R")
+source("functions and packages/plot_objects_all.R")
+
+
+leaftraits <- read.csv("calculated_data/leaftraits_summary.csv") 
+
+##this has droughtleaves
+
+
+png(filename = "makepngs/c13_N.png", width = 11, height = 8.5, units = "in", res= 400)
+
+par(mar=c(5,5,2,2), cex.lab=1.25)
+
+palette(c("yellow4", "forestgreen"))
+par(mar=c(5,5,2,2))
+plot(c13.mean ~ leafN_area.mean, data=leaftraits, pch=16, col=as.factor(leaf), ylim=c(-35,-25), 
+     xlim=c(1,3.5), cex=1.5, xlab=narealab, ylab=c13lab)
+
+legend("topright", leaflab2, pch=16,inset = 0.03, col=leafcol,bty='n', cex=1.25) 
+
+dev.off()
+
+
+###----aci
+
+aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
+
+#simulate ACi curves for each leaf, ambient and elevated T
+sunAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[3,3], Jmax=aci_leaf[3,4],PPFD=1408)
+sunET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[4,3], Jmax=aci_leaf[4,4], PPFD=1408)
+shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[1,3], Jmax=aci_leaf[1,4], PPFD=375)
+shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[2,3], Jmax=aci_leaf[2,4],PPFD=375)
+
+
+#plot
+
+png(filename = "makepngs/aci.png", width = 11, height = 8.5, units = "in", res= 400)
+
+par(mar=c(5,5,2,2), cex.lab=1.25)
+plot(sunAT_sim$Ci, sunAT_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab=cilab, ylab="", type="l", lwd=2, ylim=c(0,30))
[…]

### 2015-06-01 makepngs/sunshade_phys_png.R

@@ -1,0 +1,328 @@
+source("functions and packages/functions.R")
+source("master_scripts/plot_objects.R")
+source("functions and packages/packages.R")
+library(mgcv)
+library(lme4)
+
+#read in gm data set (no drought) and Cibar(discrimination)
+
+Ci_bar <- read.csv("calculated_data/Ci_bar.csv")
+
+gmes <- read.csv("calculated_data/gmes_wellwatered.csv")
+  ##calculate CC
+  gmes$Cc<- with(gmes, Ci-Photo/gm)
+  gmes$cc_ci<- with(gmes, Ci/Cc)
+
+###get average by id
+gm_agg <- summaryBy(Photo+Cond+Ci+Cc+gm+VpdL+xsi+DELTA+cc_ci ~ chamber+id+leaf +light+temp+leaflight, 
+                    data=gmes, FUN=mean, keep.names=TRUE)
+
+##remove shade-high
+gm_sunsha <- gm_agg[gm_agg$leaflight != "shade-high",]
+  gm_sunsha <- droplevels(gm_sunsha)
+  gm_sunsha$id <- gsub("-high", "", gm_sunsha$id)
+  gm_sunsha$id <- gsub("-low", "", gm_sunsha$id)
+
+##merge
+gm_c13 <- merge(gm_sunsha, Ci_bar[, c(2,8)], by="id")
+  gm_c13$Cc_bar <- with(gm_c13, ci_bar-Photo/gm)
+  #add total conductance to CO2
+  gm_c13$gmgs <- with(gm_c13, gm+Cond)
+
+  
+###Photosynthesis vs gs (need to fit sun with something else)------------------------------------------------
+  ###simple plot use smoothplot from RD and get CI polygon
+  palette(c(shacol, suncol))
+
+  ###try to add 95%ci as dotted line and then turn shading of addpoly lighter
+  
+  ####run gam models and then predict
+  #SUN leaves
+  sunmod <- gam(Photo ~ s(Cond, k=5), data=gm_c13, subset=leaflight=="sun-high")
+  
+  #predict
+  #get apprpriate vector of gs from sun leaves
+  gsdat <- gm_c13[gm_c13$leaflight=="sun-high", "Cond"]
+  
[…]

### 2015-06-11 markdown/wtc_gm_compare.Rmd
re ran gmes with tobacco parameters
created new gmes calc function with tob

have new RMD that just compares those values

tweaked c13 fig

--- 2015-04-02 markdown/wtc_sunshade.Rmd
@@
 
+#gm calc from E.globulus
 gmes <- read.csv("calculated_data/gmes_WTC.csv")
 gmpair<- read.csv("calculated_data/gmes_wtc_pair.csv")
-gmwet <- read.csv("calculated_data/gmes_wellwatered.csv")
-Ci_bar <- read.csv("calculated_data/Ci_bar.csv")
 ###for analysis first subset well watered and drought treatments
@@
 
-leaftraits <- read.csv("calculated_data/leaftraits_summary.csv") 
-par <- read.csv("raw data/par.csv")
-treatments <- read.csv("raw data/chamber_trt.csv")
+#gm calc from tobacco 
+gmes_tob <- read.csv("calculated_data/gmes_WTC_tobacco.csv")
+gmpair_tob<- read.csv("calculated_data/gmes_wtc_tobacco_pair.csv")
+
+###for analysis first subset well watered and drought treatments
+ gm_drought_tob <- gmes_tob[gmes_tob$drydown == "drought",]
+ gm_water_tob <- gmes_tob[gmes_tob$drydown == "control",]
 
@@
 
-#### <u>Design</u>
-* WTCIII was a warming experiment with ambient and +3°C warming (n=6).  A drought manipulation was added in the last few months of the experiment.
-* Sun leaves were measured in the upper canopy at their current PPFD.  Shade leaves were measured in the lower/middle canopy, first at their current PPFD and then again at sun leaf PPFD.
-* New fully expanded leaves were measured. Shade leaves were sampled into the middle canopy as the experiment progressed to avoid issues with leaf age.
-* Leaf physiological measurements were taken at their current temperature (ambient, +3°C) for paired comparisons across leaf types.
-Thus, data below represent a large range of temperatures from ~18-35°C.
-* For simplicity I have generalized most of the results below to just sun vs.
-shade leaves.
-So effects of the warming treatment are often not included and there is no standardization of leaf temperature.
-* Unless it is explicitly stated I have left off the drought treatment as well.  
-* Data below are mostly overall means across the 6 campaigns.  We can tease apart temperature/drought effects later. 
-* The green/brown color scheme always refers to sun leaves and shade leaves at low light
-
-#### <u>Were Sun and Shade leaves actually different?</u>
-
-```{r, results='asis'}
-aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
-names(aci_leaf) <-c("Leaf", "Temperature", "Vcmax", "Jmax", "Vcmax_se", "Jmax_se")
-
-aci_leaf2 <- aci_leaf[,c(1,2,3,5,4,6)]
-
-panderOptions("digits", 3)
-pander(aci_leaf2, caption="Leaf Biochemistry")
-```
-
-```{r,results='asis'}
-  #simulate ACi curves for each leaf, ambient and elevated T
-  sunAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[3,3], Jmax=aci_leaf[3,4],PPFD=1408)
-  sunET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[4,3], Jmax=aci_leaf[4,4], PPFD=1408)
-  shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[1,3], Jmax=aci_leaf[1,4], PPFD=375)
-  shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[2,3], Jmax=aci_leaf[2,4],PPFD=375)
-
-```
-
-* ACi curves were measured at saturating light for both sun and shade leaves.
-
-```{r, aci plot}
-#plot
-plot(sunAT_sim$Ci, sunAT_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab=cilab, ylab="", type="l", lwd=2, ylim=c(0,30))
-  points(sunET_sim$Ci, sunET_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
-  points(shaAT_sim$Ci, shaAT_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2)
-  points(shaET_sim$Ci, shaET_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
-  legend("bottomright", leglab2, lty=c(1,2,1,2), lwd=2,col=colaci, pt.bg=col4,inset = 0.03)
-  title(ylab=satlab, mgp=ypos)
-```
-
-* Here are predicted ACi curves for leaf types with temperature treatments.
-Both Vc<sub>max</sub> and J<sub>max</sub> are reduced in shade leaves with little effect of the temperature treatment(AT/ET).
-ACi curves were taken once towards the end of the experiment on well watered trees
-
-#### <u>Mesophyll Conducatnce</u>
-
-* g<sub>m</sub> was calculated with R<sub>d</sub> and G<sub>star</sub> generated for E.globulus.
-Kristine Crous and I have compared tobacco with globulus calculations on FACE trees and they tend to lower the values of g<sub>m</sub> and generate less variance within a leaf sample.
-I can alter the code to generate values with tobacco parameters, if needed.
-
-#### g<sub>m</sub> by leaf type and light environment
-
-* The two tables below are means and standard errors of g<sub>m</sub> across all campaigns.
-
-* The well-watered treatment from the drought treatment are separated.
-I'm more confident in the well-watered values as I was often on the cutoff point with drawdown and thus xsi for drought leaves.
-They show similar patterns but g<sub>m</sub> seems to be higher in drought trees
-
-* g<sub>m</sub> in shade is always lower than in sun but when the lights are 'turned on' shade leaves show an instantaneous response in g<sub>m</sub> that resembles sun leaves.
+#### <u>Mesophyll Conducatnce</u> from Euc parameters
 
@@
 
-```{r gm lights on}
-#plot low vs high light shade leaves
-plot(Photo~gm, data=gm_water, subset=leaflight=="shade-high", pch=16, col=lightscol, ylim=c(0,30), xlim=c(0,.4), 
-     xlab=gmlab, ylab="", cex=1.25)
-  points(Photo~gm, data=gm_water, subset=leaflight=="shade-low",pch=16, col=shacol, cex=1.25)
-  legend("topright", lightleg, pch=16,inset = 0.02, col=lightlab) 
-  legend("topleft", "Shade Leaves", bty='n') 
-  title(ylab=satlab, mgp=ypos, cex=1.2)
-```
-
-* This figure shows the instantaneous response of g<sub>m</sub> in shade leaves when the lights are 'turned on'.
-
-```{r gm paired, fig.height=10}
-wa<- palette(wes_palette("Zissou",5))
-paircol <- c(wa[3], wa[1],wa[5])
-
-gmpair$variable <- as.factor(paste(gmpair$leaf, gmpair$temp, sep="-"))
-
-gmpair2 <- gmpair[gmpair$leaflight != "shade-high",]
-
-#paired id dfr
-idDF <- data.frame(pair_id = unique(gmpair2$pair_id))
-
-gm_id <- summaryBy(gm~ variable+pair_id, data=gmpair2, FUN=c(mean), keep.names=TRUE)
-gm_sp <- dlply(gm_id, .(variable), function(x) merge(x, idDF, all = TRUE))
-
-par(cex.axis=1.21, cex.lab=1.51,las=1,mgp=c(3.5,1,0),mfrow=c(3,1),  
-    omi=c(.5,0,0.1,0.1))   
-# First Panel
-par(mar=c(0,7,2,2))
-plot(gm_sp[[4]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[3], cex=2.5, ylab="", xlab="", xlim=c(0,.4), ylim=c(0,.4),
-     axes=FALSE) 
-abline(0,1, lty=2)
-axis(1, labels=FALSE) 
-axis(2, labels=TRUE) 
-box()
-legend(.3,.35, pairlab[1], pch=21, pt.bg=paircol[1], pt.cex=2,bg="white",inset = 0.06)
-
-# Second panel   
-par(mar=c(0,7,0,2))
-plot(gm_sp[[1]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[1], cex=2.5, ylab=gmlab, xlim=c(0,.4), ylim=c(0,.4),xlab="",
-     axes=FALSE)     
-abline(0,1, lty=2)
-legend(.3,.35, pairlab[2], pch=21, pt.bg=paircol[2], pt.cex=2,bg="white",inset = 0.06)
-axis(1, labels=FALSE) 
-axis(2, labels=TRUE) 
-box()
-
-#third panel
-par(mar=c(5,7,0,2))
-plot(gm_sp[[2]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[5], cex=2.5, ylab="", xlim=c(0,.4), ylim=c(0,.4),xlab=suatgm)   
-abline(0,1, lty=2)
-legend(.3,.35, pairlab[3], pch=21, pt.bg=paircol[3], pt.cex=2,bg="white",inset = 0.06) 
-```
-
-* Here is a simple version of the paired leaf comparisons.
-Shade low light and sun leaves are used (watered and drought treatments are combined).
-Leaf types (with temperature treatment) are compared to sun leaves at ambient temperature as the control treatment (x axis).
-
-```{r gm temp}
-gm_ch<- summaryBy(gm+Photo+Cond+CTleaf ~id+chamber+temp+leaf+light+leaflight+Month, data=gm_water,FUN=mean,
-                  keep.names=TRUE)
-
-##photosynthesis and temperature
-gmt_sun_lm <- lm(gm~ CTleaf, data=gm_water,subset=leaflight=="sun-high")
-  sundat <- gm_water[gm_water$leaflight=="sun-high", "CTleaf"]
-  sun_seq <- seq(min(sundat), max(sundat), length=101)
-  sun_pred <- predict.lm(gmt_sun_lm, newdata=data.frame(CTleaf=sun_seq), interval="confidence")
-
-gmt_sha_lm <- lm(gm~ CTleaf, data=gm_water,subset=leaflight=="shade-low")
-  shadat <- gm_water[gm_water$leaflight=="shade-low", "CTleaf"]
-  sha_seq <- seq(min(shadat), max(shadat), length=101)
-  sha_pred <- predict.lm(gmt_sha_lm, newdata=data.frame(CTleaf=sha_seq), interval="confidence")
-
-plot(gm~CTleaf, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=suncol, ylim=c(0,.4), xlim=c(15, 40), 
-     ylab="", xlab=leaftlab)
-
-  ablineclip(gmt_sun_lm, lty=1, x1=min(gm_water[gm_water$leaf=="sun","CTleaf"]), 
-             x2=max(gm_water[gm_water$leaf=="sun","CTleaf"]), col="forestgreen", lwd=2)
-  lines(sun_seq, sun_pred[,2], lty=2, lwd=2, col="forestgreen")
-  lines(sun_seq, sun_pred[,3], lty=2, lwd=2, col="forestgreen")
-  #shade
-  points(gm~CTleaf, data=gm_ch, subset=leaflight=="shade-low",pch=16, col=shacol)
-  ablineclip(gmt_sha_lm, lty=1, x1=min(gm_water[gm_water$leaf=="sun","CTleaf"]), 
-             x2=max(gm_water[gm_water$leaf=="sun","CTleaf"]), col="yellow4", lwd=2)
-  lines(sha_seq, sha_pred[,2], lty=2, lwd=2, col="yellow4")
-  lines(sha_seq, sha_pred[,3], lty=2, lwd=2, col="yellow4")
-
-legend("topleft", c("sun", "shade"), pch=16,inset = 0.03, col=leafcol) 
-title(ylab=gmlab, mgp=ypos, cex=1.2)
-```
-
-* This figure shows the relationship between g<sub>m</sub> and temperature with 95% confidence intervals.
-Here we see a slight reduction in g<sub>m</sub> with increases in temperature, but only significant in sun leaves.
-
-### <u>Sun vs. Shade Leaf Physiology</u>
-
-```{r A_gs, fig.width=12, fig.height=6}
-source("markdown/ags_plots.R")
-```
-
-* Figures of photosynthesis and g<sub>s</sub>.
-Panel (a) is shown with 95% confidence intervals.
-Panel (b) shows predicted photosynthesis rates generated with g<sub>s</sub> values, mean leaf temperature and PPFD (sun/shade), and ACi parameters including ambient and elevated temperature treatments.
-Here we can see optimal stomatal behavior in sun leaves but not in shade leaves.
-
-```{r a-gm, }
-source("calculated_data/bootstrap/readboot.R")
-  ###Photosynthesis vs gm
-plot(Photo~gm, data=sundat, pch=16, col=suncol, ylim=c(0,30), xlim=c(0,.4), xlab=gmlab, ylab="", cex=1.25)
-  points(Photo~gm, data=shadat, pch=16, col=shacol, cex=1.25)
-with(agm_sun, {
-  lines(gm, lcl, lty=2, lwd=2,col="forestgreen")
-  lines(gm, ucl, lty=2, lwd=2,col="forestgreen")
-  lines(gm, pred, lty=1, lwd=2,col="forestgreen")
-})
-with(agm_sha, {
-  lines(gm, lcl, lty=2, lwd=2,col="yellow4")
-  lines(gm, ucl, lty=2, lwd=2,col="yellow4")
-  lines(gm, pred, lty=1, lwd=2,col="yellow4")
-})
-title(ylab=satlab, mgp=ypos, cex=1.2)
-legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol) 
-```
-
-* This figure shows the relationship between photosynthesis and g<sub>m</sub> for sun and shade leaves with 95% confidence intervals.
-
-```{r leafphys,fig.width=12, fig.height=6}
-###Photosynthesis vs Ci
-  shade <- gm_c13[gm_c13$leaf=="shade",]
-  sun <- gm_c13[gm_c13$leaf=="sun",]
-
-  #read in jmax vcmax
-  jvc<- read.csv("calculated_data/tdl_aci.csv")
-
-  #simulate ACi curves for each leaf, ambient and elevated T
-  sunAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[3,3], Jmax=jvc[3,4],PPFD=1407.959)
-  sunET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[4,3], Jmax=jvc[4,4], PPFD=1407.959)
-  shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[1,3], Jmax=jvc[1,4], PPFD=305)
-  shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[2,3], Jmax=jvc[2,4],PPFD=305)
-
-par(cex.axis=.96, cex.lab=1.2,mfrow=c(1,2),oma=c(0.1,0.1,0.1,0.1), las=1)
-
-par(mar=c(5,4,2,0))
-plot(Photo~Ci, data=sundat, pch=16, col=suncol, ylim=c(0,30), xlim=c(0,350), xlab=cilab, ylab="", cex=1.25)
-points(Photo~Ci, data=shadat, pch=16, col=shacol, cex=1.25)
-with(aci_sun, {
-  lines(Ci, lcl, lty=2, lwd=2,col="forestgreen")
-  lines(Ci, ucl, lty=2, lwd=2,col="forestgreen")
-  lines(Ci, pred, lty=1, lwd=2,col="forestgreen")
-})
-with(aci_sha, {
-  lines(Ci, lcl, lty=2, lwd=2,col="yellow4")
-  lines(Ci, ucl, lty=2, lwd=2,col="yellow4")
-  lines(Ci, pred, lty=1, lwd=2,col="yellow4")
-})
-title(ylab=satlab, mgp=ypos, cex=1.2)
-legend("topright", leaflab2, pch=16,inset = 0.03, col=leafcol) 
-text(x=.01, 29.5, "(a)", cex=1.2)
-
-par(mar=c(5,2.5,2,2)) 
-plot(sunAT_sim$Ci, sunAT_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab=cilab, ylab="", type="l",  lwd=2,ylim=c(0,30), xlim=c(0,1000))
-  points(sunET_sim$Ci, sunET_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
-  points(shaAT_sim$Ci, shaAT_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2)
-  points(shaET_sim$Ci, shaET_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
-  points(Photo~Ci, data=sundat, pch=16, col=suncol, cex=1.25)
-  points(Photo~Ci, data=shadat, pch=16, col=shacol, cex=1.25)
-
-  legend("bottomright", leglab2, lty=c(1,2,1,2), lwd=2,col=colaci, pt.bg=col4,inset = 0.03)
-  text(x=.01, 29.5, "(b)", cex=1.2)
-```
-
-* These figures show the relationship of photosynthesis to Ci.
-Panel (a) has 95% confidence intervals.
-Panel (b) has the predicted ACi curves using the biochemical parameters measured at saturating light but fitted with mean PPFD of sun and shade leaves.
-
-```{r}
- ###Photosynthesis vs Cc 
-  plot(Photo~Cc, data=sundat, pch=16, col=suncol, ylim=c(0,30), xlim=c(0,350), xlab=cclab, ylab="", cex=1.25)
-  points(Photo~Cc, data=shadat, pch=16, col=shacol, cex=1.25)
-  with(acc_sun, {
-    lines(Cc, lcl, lty=2, lwd=2,col="forestgreen")
-    lines(Cc, ucl, lty=2, lwd=2,col="forestgreen")
-    lines(Cc, pred, lty=1, lwd=2,col="forestgreen")
-  })
-  with(acc_sha, {
-    lines(Cc, lcl, lty=2, lwd=2,col="yellow4")
-    lines(Cc, ucl, lty=2, lwd=2,col="yellow4")
-    lines(Cc, pred, lty=1, lwd=2,col="yellow4")
-  })
-  title(ylab=satlab, mgp=ypos, cex=1.2)
-  legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol) 
-```
-
-* This figures show the relationship of photosynthesis to Cc with 95% confidence intervals, calculated with g<sub>m</sub> data
-
-### <u>Shade versus Sun leaves, continued</u>
-
-This next set of figures attempts to show how sun and shade leaves were fundamentally different.
+#### <u>Mesophyll Conducatnce</u> from tobacco parameters
 
 ```{r,results='asis'}
-##this has droughtleaves
-palette(c("yellow4", "forestgreen"))
-par(mar=c(5,5,2,2))
-plot(c13.mean ~ leafN_area.mean, data=leaftraits, pch=16, col=as.factor(leaf), ylim=c(-35,-25), 
-                 xlim=c(1,3.5), cex=1.25, xlab=narealab, ylab=c13lab)
-legend("bottomright", leaflab2, pch=16, col=c("green4","yellow4"),inset = 0.03)
+###make a table of means and se of gm
 
+ gm_agg_tob <- summaryBy(gm~ leaf+light+temp+leaflight, data=gm_water_tob,FUN=c(mean,se))
+   tablenames <- c( "Leaf Type", "Light", "Temperature","leaflight","Gm", "Gm(se)")
+   names(gm_agg_tob) <- tablenames
+  gm_agg3 <- gm_agg_tob[c(3,4,5,6,1,2),]
+  row.names(gm_agg3) <- NULL
+
+panderOptions("digits", 3)
+pander(gm_agg3,caption="Well Watered")
+
+  gm_dr_tob <- summaryBy(gm~ leaf+light+temp+leaflight, data=gm_drought_tob,FUN=c(mean,se))
+    tablenames <- c( "Leaf Type", "Light", "Temperature","Leaf-Light","Gm", "Gm(se)")
+    names(gm_dr) <- tablenames
+  gm_dr3 <- gm_dr_tob[c(3,4,5,6,1,2),]
+  row.names(gm_dr3) <- NULL
+
+pander(gm_dr3,caption="Drought")
 ```
-
-* This figure shows the relationship between leaf N on an area basis and ${\delta}$^13^C between leaf types (chamber means).
-
-```{r,results='asis'}
- #format function
- par<- parformat(par)
-
- par_leaf <- subset(par, ID !="shade-high")
- Morder <- c("Oct", "Dec", "Jan", "Feb", "Mar", "Apr")
-
- #data format for bar
- parbar <- subset(par_leaf, select = c("par", "month", "leaf_type"))
- parbar$month <- factor(parbar$month, levels = Morder)
-
- bar(par, c(leaf_type, month), parbar, col=c("yellow4", "forestgreen"), xlab="", ylab="", ylim=c(0, 2000), 
-     half.errbar=FALSE)
- title(ylab=parlab, mgp=ypos)
-```
-
-* Here is the light environment in which gas exchange measurements for sun and shade leaves were measured.
-This data was generated with a handheld par sensor for each individual leaf and verified with ceptometer readings within the canopy at the height of the measured leaf.
-There was a ~75% reduction in PPFD from sun to shave leaves.
-
-```{r}
- ###LMA-
- lma <- read.csv("calculated_data/leaf_chemistry.csv")
-
- #data format for bar
- lmabar <- subset(lma, select = c("lma", "Month", "leaf"))
-   lmabar$Month <- factor(lmabar$Month, levels = Morder)
-
- bar(lma, c(leaf, Month), lmabar, col=c("yellow4", "forestgreen"), xlab="", ylab="", ylim=c(0, 185), half.errbar=FALSE)
- title(ylab=lmalab, mgp=ypos)
-```
-
-* This figure shows leaf mass area for each gas exchange campaign.  Surprisingly, not very different between leaf types.

### 2015-06-11 functions and packages/gmes_calc_tobacco.R

@@ -1,0 +1,36 @@
+####gmescalculation----------------------------------------------------------------------------------------------
+gmcalc_tob_func <- function(x, a=4.4, ab= 2.9, e=30, b=29, f=16.2,del_growth = -8 , delR=-38, 
+                        k25r=0.718, k25g=38.89, Ea_r = 46.39, Ea_g = 24.46,Rgc=8.314472, c_r=18.72, c_g=13.49){
+  
+  x$CiCa <- x$Ci/x$CO2R
+  x$a_prime <- (ab*(x$CO2S-x$C2sfc)+a*(x$C2sfc-x$Ci))/(x$CO2S-x$Ci)
+  
+  x$Rd <- k25r * exp(c_r - Ea_r* 1000 / (Rgc*(273 + x$Tleaf)))
+  x$Gstar <- exp(c_g - Ea_g * 1000 / (Rgc*(273 + x$Tleaf))) * x$O2/21
+  
+#   x$Rd <- k25r * exp(Ea_r*((x$Tleaf+273.15)-298)/(298*Rgc*(x$Tleaf+273.15)))
+#   x$Gstar <- k25g * exp(Ea_g*((x$Tleaf+273.15)-298)/(298*Rgc*(x$Tleaf+273.15)))
+  
+  x$rd_term <- e*x$Rd*(x$Ci-x$Gstar)/((x$Photo+x$Rd)*x$CO2S)
+  x$f_term <- f*x$Gstar/x$CO2S
+  
+  x$TleafminusTair <- x$Tleaf - x$Tair
+  x$TblockminusTair <- x$TBlk - x$Tair
+  
+  x$CO2Rdry <- x$CO2R/(1-x$H2OR/1000)
+  x$CO2Sdry <- x$CO2S/(1-x$H2OS/1000)
+  
+  x$t <- (1+x$a_prime/1000)*x$Trmmol/x$CndCO2/1000/2
+  x$t2 <- 1/(1-x$t)
+  x$t3 <- (1+x$t)/(1-x$t)
+  
+  x$Di <- x$a_prime * x$t2+ (x$t3 * b-x$a_prime * x$t2) * x$CiCa
+  x$DiminusDo <- x$Di - x$DELTA
+  
+  x$rd_term2 <- x$t3- x$rd_term
+  x$f_term2 <- x$t3 - x$f_term
+  
[…]

### 2015-06-23 scripts/narea_ci.R
added some plots to markdown

@@ -47,12 +47,20 @@ palette(c(shacol, suncol))
 
 greyalpha <- alpha("grey95", alpha=.75)
 
+legcols <- c(shacol, suncol)
+legpchs <- c(1, 16)
+leaflabel <- c( "Sun", "Shade")
+leaflabel2 <- c( cibarlab2,cilab)
 
-windows(8.8)
+windows(8,8)
+png(filename = "makepngs/ci_narea.png", width = 11, height = 8.5, units = "in", res= 400)
 par(mar=c(5,5,2,2), cex=1.5)
-plot(Ci ~ leafN_area, data=ci_narea, col=as.factor(leaf), pch=16, ylim=c(0, 400), xlim=c(0,5), ylab=cilab, xlab=narealab)
+plot(Ci ~ leafN_area, data=ci_narea, col=as.factor(leaf), pch=16, ylim=c(0, 400), xlim=c(0,6), ylab=cilab, xlab=narealab)
 points(ci_bar ~ leafN_area, data=ci_narea,pch=21, col=as.factor(leaf), bg=greyalpha)
-dev.copy2pdf(file="master_scripts/figures/ci_narea.pdf")
+
+legend("topright", leaflabel, pch=16,inset = 0.03, col=leafcol, bty='n', cex=1) 
+legend("bottomright", leaflabel2, pch=legpchs,inset = 0.03,  bty='n', cex=1) 
+#dev.copy2pdf(file="master_scripts/figures/ci_narea.pdf")
 dev.off()
 
 

### 2015-06-23 markdown/WTC3_figures.Rmd
moved non master files to scripts folder
started water use script

--- 2015-04-02 markdown/wtc_sunshade.Rmd
@@
 ---
-title: "Sun vs Shade leaves (WTCIII)"
+title: "WTCIII-Figures"
 author: "Court Campany"
-date: "April 2, 2015"
+date: "June 23, 2015"
 output: html_document
 ---
-```{r knitr_options , include=FALSE}
-knitr::opts_chunk$set(fig.width=6, fig.height=6,
-echo=FALSE, message=FALSE,
-fig.align='center', warning=FALSE)
-```
 
-```{r, results='asis'}
-library(knitr)
-opts_knit$set(root.dir = '../')
-```
+![ciN](ppfd.png)
 
-```{r global data/sourcing , include=FALSE}
-source("functions and packages/packages_md.R")
-source("functions and packages/functions.R")
-source("functions and packages/plot_objects_all.R")
+* Figure 1. Here is the light environment in which gas exchange measurements for sun and shade leaves were measured.
 
-gmes <- read.csv("calculated_data/gmes_WTC.csv")
-gmpair<- read.csv("calculated_data/gmes_wtc_pair.csv")
-gmwet <- read.csv("calculated_data/gmes_wellwatered.csv")
-Ci_bar <- read.csv("calculated_data/Ci_bar.csv")
- ###for analysis first subset well watered and drought treatments
- gm_drought <- gmes[gmes$drydown == "drought",]
- gm_water <- gmes[gmes$drydown == "control",]
-
-leaftraits <- read.csv("calculated_data/leaftraits_summary.csv") 
-par <- read.csv("raw data/par.csv")
-treatments <- read.csv("raw data/chamber_trt.csv")
-
-```
-
-#### <u>Design</u>
-* WTCIII was a warming experiment with ambient and +3°C warming (n=6).  A drought manipulation was added in the last few months of the experiment.
-* Sun leaves were measured in the upper canopy at their current PPFD.  Shade leaves were measured in the lower/middle canopy, first at their current PPFD and then again at sun leaf PPFD.
-* New fully expanded leaves were measured. Shade leaves were sampled into the middle canopy as the experiment progressed to avoid issues with leaf age.
-* Leaf physiological measurements were taken at their current temperature (ambient, +3°C) for paired comparisons across leaf types.
-Thus, data below represent a large range of temperatures from ~18-35°C.
-* For simplicity I have generalized most of the results below to just sun vs.
-shade leaves.
-So effects of the warming treatment are often not included and there is no standardization of leaf temperature.
-* Unless it is explicitly stated I have left off the drought treatment as well.  
-* Data below are mostly overall means across the 6 campaigns.  We can tease apart temperature/drought effects later. 
-* The green/brown color scheme always refers to sun leaves and shade leaves at low light
-
-#### <u>Were Sun and Shade leaves actually different?</u>
-
-```{r, results='asis'}
-aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
-names(aci_leaf) <-c("Leaf", "Temperature", "Vcmax", "Jmax", "Vcmax_se", "Jmax_se")
-
-aci_leaf2 <- aci_leaf[,c(1,2,3,5,4,6)]
-
-panderOptions("digits", 3)
-pander(aci_leaf2, caption="Leaf Biochemistry")
-```
-
-```{r,results='asis'}
-  #simulate ACi curves for each leaf, ambient and elevated T
-  sunAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[3,3], Jmax=aci_leaf[3,4],PPFD=1408)
-  sunET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[4,3], Jmax=aci_leaf[4,4], PPFD=1408)
-  shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[1,3], Jmax=aci_leaf[1,4], PPFD=375)
-  shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=aci_leaf[2,3], Jmax=aci_leaf[2,4],PPFD=375)
-
-```
-
-* ACi curves were measured at saturating light for both sun and shade leaves.
-
-```{r, aci plot}
-#plot
-plot(sunAT_sim$Ci, sunAT_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab=cilab, ylab="", type="l", lwd=2, ylim=c(0,30))
-  points(sunET_sim$Ci, sunET_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
-  points(shaAT_sim$Ci, shaAT_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2)
-  points(shaET_sim$Ci, shaET_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
-  legend("bottomright", leglab2, lty=c(1,2,1,2), lwd=2,col=colaci, pt.bg=col4,inset = 0.03)
-  title(ylab=satlab, mgp=ypos)
-```
-
-* Here are predicted ACi curves for leaf types with temperature treatments.
-Both Vc<sub>max</sub> and J<sub>max</sub> are reduced in shade leaves with little effect of the temperature treatment(AT/ET).
-ACi curves were taken once towards the end of the experiment on well watered trees
-
-#### <u>Mesophyll Conducatnce</u>
-
-* g<sub>m</sub> was calculated with R<sub>d</sub> and G<sub>star</sub> generated for E.globulus.
-Kristine Crous and I have compared tobacco with globulus calculations on FACE trees and they tend to lower the values of g<sub>m</sub> and generate less variance within a leaf sample.
-I can alter the code to generate values with tobacco parameters, if needed.
-
-#### g<sub>m</sub> by leaf type and light environment
-
-* The two tables below are means and standard errors of g<sub>m</sub> across all campaigns.
-
-* The well-watered treatment from the drought treatment are separated.
-I'm more confident in the well-watered values as I was often on the cutoff point with drawdown and thus xsi for drought leaves.
-They show similar patterns but g<sub>m</sub> seems to be higher in drought trees
-
-* g<sub>m</sub> in shade is always lower than in sun but when the lights are 'turned on' shade leaves show an instantaneous response in g<sub>m</sub> that resembles sun leaves.
-
-```{r,results='asis'}
-###make a table of means and se of gm
-
- gm_agg <- summaryBy(gm~ leaf+light+temp+leaflight, data=gm_water,FUN=c(mean,se))
-   tablenames <- c( "Leaf Type", "Light", "Temperature","leaflight","Gm", "Gm(se)")
-   names(gm_agg) <- tablenames
-  gm_agg2 <- gm_agg[c(3,4,5,6,1,2),]
-  row.names(gm_agg2) <- NULL
-
-panderOptions("digits", 3)
-pander(gm_agg2,caption="Well Watered")
-
-  gm_dr <- summaryBy(gm~ leaf+light+temp+leaflight, data=gm_drought,FUN=c(mean,se))
-    tablenames <- c( "Leaf Type", "Light", "Temperature","Leaf-Light","Gm", "Gm(se)")
-    names(gm_dr) <- tablenames
-  gm_dr2 <- gm_dr[c(3,4,5,6,1,2),]
-  row.names(gm_dr2) <- NULL
-
-pander(gm_dr2,caption="Drought")
-```
-
-```{r gm lights on}
-#plot low vs high light shade leaves
-plot(Photo~gm, data=gm_water, subset=leaflight=="shade-high", pch=16, col=lightscol, ylim=c(0,30), xlim=c(0,.4), 
-     xlab=gmlab, ylab="", cex=1.25)
-  points(Photo~gm, data=gm_water, subset=leaflight=="shade-low",pch=16, col=shacol, cex=1.25)
-  legend("topright", lightleg, pch=16,inset = 0.02, col=lightlab) 
-  legend("topleft", "Shade Leaves", bty='n') 
-  title(ylab=satlab, mgp=ypos, cex=1.2)
-```
-
-* This figure shows the instantaneous response of g<sub>m</sub> in shade leaves when the lights are 'turned on'.
-
-```{r gm paired, fig.height=10}
-wa<- palette(wes_palette("Zissou",5))
-paircol <- c(wa[3], wa[1],wa[5])
-
-gmpair$variable <- as.factor(paste(gmpair$leaf, gmpair$temp, sep="-"))
-
-gmpair2 <- gmpair[gmpair$leaflight != "shade-high",]
-
-#paired id dfr
-idDF <- data.frame(pair_id = unique(gmpair2$pair_id))
-
-gm_id <- summaryBy(gm~ variable+pair_id, data=gmpair2, FUN=c(mean), keep.names=TRUE)
-gm_sp <- dlply(gm_id, .(variable), function(x) merge(x, idDF, all = TRUE))
-
-par(cex.axis=1.21, cex.lab=1.51,las=1,mgp=c(3.5,1,0),mfrow=c(3,1),  
-    omi=c(.5,0,0.1,0.1))   
-# First Panel
-par(mar=c(0,7,2,2))
-plot(gm_sp[[4]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[3], cex=2.5, ylab="", xlab="", xlim=c(0,.4), ylim=c(0,.4),
-     axes=FALSE) 
-abline(0,1, lty=2)
-axis(1, labels=FALSE) 
-axis(2, labels=TRUE) 
-box()
-legend(.3,.35, pairlab[1], pch=21, pt.bg=paircol[1], pt.cex=2,bg="white",inset = 0.06)
-
-# Second panel   
-par(mar=c(0,7,0,2))
-plot(gm_sp[[1]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[1], cex=2.5, ylab=gmlab, xlim=c(0,.4), ylim=c(0,.4),xlab="",
-     axes=FALSE)     
-abline(0,1, lty=2)
-legend(.3,.35, pairlab[2], pch=21, pt.bg=paircol[2], pt.cex=2,bg="white",inset = 0.06)
-axis(1, labels=FALSE) 
-axis(2, labels=TRUE) 
-box()
-
-#third panel
-par(mar=c(5,7,0,2))
-plot(gm_sp[[2]][,3]~gm_sp[[3]][,3], pch=21, bg=wa[5], cex=2.5, ylab="", xlim=c(0,.4), ylim=c(0,.4),xlab=suatgm)   
-abline(0,1, lty=2)
-legend(.3,.35, pairlab[3], pch=21, pt.bg=paircol[3], pt.cex=2,bg="white",inset = 0.06) 
-```
-
-* Here is a simple version of the paired leaf comparisons.
-Shade low light and sun leaves are used (watered and drought treatments are combined).
-Leaf types (with temperature treatment) are compared to sun leaves at ambient temperature as the control treatment (x axis).
-
-```{r gm temp}
-gm_ch<- summaryBy(gm+Photo+Cond+CTleaf ~id+chamber+temp+leaf+light+leaflight+Month, data=gm_water,FUN=mean,
-                  keep.names=TRUE)
-
-##photosynthesis and temperature
-gmt_sun_lm <- lm(gm~ CTleaf, data=gm_water,subset=leaflight=="sun-high")
-  sundat <- gm_water[gm_water$leaflight=="sun-high", "CTleaf"]
-  sun_seq <- seq(min(sundat), max(sundat), length=101)
-  sun_pred <- predict.lm(gmt_sun_lm, newdata=data.frame(CTleaf=sun_seq), interval="confidence")
-
-gmt_sha_lm <- lm(gm~ CTleaf, data=gm_water,subset=leaflight=="shade-low")
-  shadat <- gm_water[gm_water$leaflight=="shade-low", "CTleaf"]
-  sha_seq <- seq(min(shadat), max(shadat), length=101)
-  sha_pred <- predict.lm(gmt_sha_lm, newdata=data.frame(CTleaf=sha_seq), interval="confidence")
-
-plot(gm~CTleaf, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=suncol, ylim=c(0,.4), xlim=c(15, 40), 
-     ylab="", xlab=leaftlab)
-
-  ablineclip(gmt_sun_lm, lty=1, x1=min(gm_water[gm_water$leaf=="sun","CTleaf"]), 
-             x2=max(gm_water[gm_water$leaf=="sun","CTleaf"]), col="forestgreen", lwd=2)
-  lines(sun_seq, sun_pred[,2], lty=2, lwd=2, col="forestgreen")
-  lines(sun_seq, sun_pred[,3], lty=2, lwd=2, col="forestgreen")
-  #shade
-  points(gm~CTleaf, data=gm_ch, subset=leaflight=="shade-low",pch=16, col=shacol)
-  ablineclip(gmt_sha_lm, lty=1, x1=min(gm_water[gm_water$leaf=="sun","CTleaf"]), 
-             x2=max(gm_water[gm_water$leaf=="sun","CTleaf"]), col="yellow4", lwd=2)
-  lines(sha_seq, sha_pred[,2], lty=2, lwd=2, col="yellow4")
-  lines(sha_seq, sha_pred[,3], lty=2, lwd=2, col="yellow4")
-
-legend("topleft", c("sun", "shade"), pch=16,inset = 0.03, col=leafcol) 
-title(ylab=gmlab, mgp=ypos, cex=1.2)
-```
-
-* This figure shows the relationship between g<sub>m</sub> and temperature with 95% confidence intervals.
-Here we see a slight reduction in g<sub>m</sub> with increases in temperature, but only significant in sun leaves.
-
-### <u>Sun vs. Shade Leaf Physiology</u>
-
-```{r A_gs, fig.width=12, fig.height=6}
-source("markdown/ags_plots.R")
-```
-
-* Figures of photosynthesis and g<sub>s</sub>.
-Panel (a) is shown with 95% confidence intervals.
-Panel (b) shows predicted photosynthesis rates generated with g<sub>s</sub> values, mean leaf temperature and PPFD (sun/shade), and ACi parameters including ambient and elevated temperature treatments.
-Here we can see optimal stomatal behavior in sun leaves but not in shade leaves.
-
-```{r a-gm, }
-source("calculated_data/bootstrap/readboot.R")
-  ###Photosynthesis vs gm
-plot(Photo~gm, data=sundat, pch=16, col=suncol, ylim=c(0,30), xlim=c(0,.4), xlab=gmlab, ylab="", cex=1.25)
-  points(Photo~gm, data=shadat, pch=16, col=shacol, cex=1.25)
-with(agm_sun, {
-  lines(gm, lcl, lty=2, lwd=2,col="forestgreen")
-  lines(gm, ucl, lty=2, lwd=2,col="forestgreen")
-  lines(gm, pred, lty=1, lwd=2,col="forestgreen")
-})
-with(agm_sha, {
-  lines(gm, lcl, lty=2, lwd=2,col="yellow4")
-  lines(gm, ucl, lty=2, lwd=2,col="yellow4")
-  lines(gm, pred, lty=1, lwd=2,col="yellow4")
-})
-title(ylab=satlab, mgp=ypos, cex=1.2)
-legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol) 
-```
-
-* This figure shows the relationship between photosynthesis and g<sub>m</sub> for sun and shade leaves with 95% confidence intervals.
-
-```{r leafphys,fig.width=12, fig.height=6}
-###Photosynthesis vs Ci
-  shade <- gm_c13[gm_c13$leaf=="shade",]
-  sun <- gm_c13[gm_c13$leaf=="sun",]
-
-  #read in jmax vcmax
-  jvc<- read.csv("calculated_data/tdl_aci.csv")
-
-  #simulate ACi curves for each leaf, ambient and elevated T
-  sunAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[3,3], Jmax=jvc[3,4],PPFD=1407.959)
-  sunET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[4,3], Jmax=jvc[4,4], PPFD=1407.959)
-  shaAT_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[1,3], Jmax=jvc[1,4], PPFD=305)
-  shaET_sim <- Aci(Ci=seq(50,1000,length=101), Vcmax=jvc[2,3], Jmax=jvc[2,4],PPFD=305)
-
-par(cex.axis=.96, cex.lab=1.2,mfrow=c(1,2),oma=c(0.1,0.1,0.1,0.1), las=1)
-
-par(mar=c(5,4,2,0))
-plot(Photo~Ci, data=sundat, pch=16, col=suncol, ylim=c(0,30), xlim=c(0,350), xlab=cilab, ylab="", cex=1.25)
-points(Photo~Ci, data=shadat, pch=16, col=shacol, cex=1.25)
-with(aci_sun, {
-  lines(Ci, lcl, lty=2, lwd=2,col="forestgreen")
-  lines(Ci, ucl, lty=2, lwd=2,col="forestgreen")
-  lines(Ci, pred, lty=1, lwd=2,col="forestgreen")
-})
-with(aci_sha, {
-  lines(Ci, lcl, lty=2, lwd=2,col="yellow4")
-  lines(Ci, ucl, lty=2, lwd=2,col="yellow4")
-  lines(Ci, pred, lty=1, lwd=2,col="yellow4")
-})
-title(ylab=satlab, mgp=ypos, cex=1.2)
-legend("topright", leaflab2, pch=16,inset = 0.03, col=leafcol) 
-text(x=.01, 29.5, "(a)", cex=1.2)
-
-par(mar=c(5,2.5,2,2)) 
-plot(sunAT_sim$Ci, sunAT_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab=cilab, ylab="", type="l",  lwd=2,ylim=c(0,30), xlim=c(0,1000))
-  points(sunET_sim$Ci, sunET_sim$ALEAF, col="forestgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
-  points(shaAT_sim$Ci, shaAT_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2)
-  points(shaET_sim$Ci, shaET_sim$ALEAF, col="yellow4", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
-  points(Photo~Ci, data=sundat, pch=16, col=suncol, cex=1.25)
-  points(Photo~Ci, data=shadat, pch=16, col=shacol, cex=1.25)
-
-  legend("bottomright", leglab2, lty=c(1,2,1,2), lwd=2,col=colaci, pt.bg=col4,inset = 0.03)
-  text(x=.01, 29.5, "(b)", cex=1.2)
-```
-
-* These figures show the relationship of photosynthesis to Ci.
-Panel (a) has 95% confidence intervals.
-Panel (b) has the predicted ACi curves using the biochemical parameters measured at saturating light but fitted with mean PPFD of sun and shade leaves.
-
-```{r}
- ###Photosynthesis vs Cc 
-  plot(Photo~Cc, data=sundat, pch=16, col=suncol, ylim=c(0,30), xlim=c(0,350), xlab=cclab, ylab="", cex=1.25)
-  points(Photo~Cc, data=shadat, pch=16, col=shacol, cex=1.25)
-  with(acc_sun, {
-    lines(Cc, lcl, lty=2, lwd=2,col="forestgreen")
-    lines(Cc, ucl, lty=2, lwd=2,col="forestgreen")
-    lines(Cc, pred, lty=1, lwd=2,col="forestgreen")
-  })
-  with(acc_sha, {
-    lines(Cc, lcl, lty=2, lwd=2,col="yellow4")
-    lines(Cc, ucl, lty=2, lwd=2,col="yellow4")
-    lines(Cc, pred, lty=1, lwd=2,col="yellow4")
-  })
-  title(ylab=satlab, mgp=ypos, cex=1.2)
-  legend("topleft", leaflab2, pch=16,inset = 0.03, col=leafcol) 
-```
-
-* This figures show the relationship of photosynthesis to Cc with 95% confidence intervals, calculated with g<sub>m</sub> data
-
-### <u>Shade versus Sun leaves, continued</u>
-
-This next set of figures attempts to show how sun and shade leaves were fundamentally different.
-
-```{r,results='asis'}
-##this has droughtleaves
-palette(c("yellow4", "forestgreen"))
-par(mar=c(5,5,2,2))
-plot(c13.mean ~ leafN_area.mean, data=leaftraits, pch=16, col=as.factor(leaf), ylim=c(-35,-25), 
-                 xlim=c(1,3.5), cex=1.25, xlab=narealab, ylab=c13lab)
-legend("bottomright", leaflab2, pch=16, col=c("green4","yellow4"),inset = 0.03)
-
-```
-
-* This figure shows the relationship between leaf N on an area basis and ${\delta}$^13^C between leaf types (chamber means).
-
-```{r,results='asis'}
- #format function
- par<- parformat(par)
-
- par_leaf <- subset(par, ID !="shade-high")
- Morder <- c("Oct", "Dec", "Jan", "Feb", "Mar", "Apr")
-
- #data format for bar
- parbar <- subset(par_leaf, select = c("par", "month", "leaf_type"))
- parbar$month <- factor(parbar$month, levels = Morder)
-
- bar(par, c(leaf_type, month), parbar, col=c("yellow4", "forestgreen"), xlab="", ylab="", ylim=c(0, 2000), 
-     half.errbar=FALSE)
- title(ylab=parlab, mgp=ypos)
-```
-
-* Here is the light environment in which gas exchange measurements for sun and shade leaves were measured.
 This data was generated with a handheld par sensor for each individual leaf and verified with ceptometer readings within the canopy at the height of the measured leaf.
@@
 
-```{r}
- ###LMA-
- lma <- read.csv("calculated_data/leaf_chemistry.csv")
+![aci](aci_curves.png)
 
- #data format for bar
- lmabar <- subset(lma, select = c("lma", "Month", "leaf"))
-   lmabar$Month <- factor(lmabar$Month, levels = Morder)
-
- bar(lma, c(leaf, Month), lmabar, col=c("yellow4", "forestgreen"), xlab="", ylab="", ylim=c(0, 185), half.errbar=FALSE)
- title(ylab=lmalab, mgp=ypos)
-```
-
-* This figure shows leaf mass area for each gas exchange campaign.  Surprisingly, not very different between leaf types.
+* ACi curves were measured at saturating light for both sun and shade leaves.
+Both Vc<sub>max</sub> and J<sub>max</sub> are reduced in shade leaves with little effect of the temperature treatment(AT/ET).
+ACi curves were taken once towards the end of the experiment on well watered trees

### 2015-06-23 master_scripts/plot_objects.R

@@ -21,7 +21,6 @@ ksunlab <- expression(Leaf[sun]~Hyrdaulic~Conductance~~(units))
 kshadelab <- expression(Leaf[shade]~Hyrdaulic~Conductance~~(units))
 
 
-
 c13lab <-expression(paste(delta^{13}, "C (\u2030)"))
 ypos <- c(2.5,1,0)
 satlab <- expression(A~~(mu*mol~m^-2~s^-1))
@@ -30,6 +29,8 @@ condlab <- expression(g[s]~~(mol~m^-2~s^-1))
 totcondlab <- expression(Total~Conductance~to~CO[2]~~(mol~m^-2~s^-1))
 
 vpdlab <- "VPD  (kPa)"
+itelab <- expression(ITE~~(mu*mol~CO[2]~mmol~H[2]*O^-1))
+
 ratelab <- expression(mol~m^-2~s^-1)
 parlab <- expression(PPFD~~(mol~m^-2~s^-1))
 photolab <- expression(italic(A)~~(mu*mol~m^-2~s^-1))

### 2015-06-24 markdown/WTC3_figures.Rmd
lots of stuff done!!!!!!
made more manuscript figures in master folder

moved more scripts out of master folder

added figures to markdown figures rmd

updated packages script so it is universal for scripts and markdowns

added new labels to plot objects

@@
 
-![ciN](ppfd.png)
+```{r knitr_options , include=FALSE}
+knitr::opts_chunk$set(fig.width=6, fig.height=6,
+echo=FALSE, message=FALSE,
+fig.align='center', warning=FALSE)
+```
 
-* Figure 1. Here is the light environment in which gas exchange measurements for sun and shade leaves were measured.
+```{r, results='asis'}
+library(knitr)
+opts_knit$set(root.dir = '../')
+```
 
-This data was generated with a handheld par sensor for each individual leaf and verified with ceptometer readings within the canopy at the height of the measured leaf.
+```{r global data/sourcing , include=FALSE}
+source("functions and packages/functions.R")
+source("functions and packages/packages.R")
+source("master_scripts/plot_objects.R")
+library(pander)
+
+par <- read.csv("raw data/par.csv")
+treatments <- read.csv("raw data/chamber_trt.csv")
+
+```
+
+#### <u>Design</u>
+* WTCIII was a warming experiment with ambient and +3°C warming (n=6).  A drought manipulation was added in the last few months of the experiment but is not included here.
+* Sun leaves were measured in the upper canopy at their current PPFD.  Shade leaves were measured in the lower/middle canopy, first at their current PPFD and then again at sun leaf PPFD.
+* New fully expanded leaves were measured. Shade leaves were sampled into the middle canopy as the experiment progressed to avoid issues with leaf age.
+* Leaf physiological measurements were taken at their current temperature (ambient, +3°C). Thus, data below represent a large range of temperatures from ~18-35°C.
+
+```{r par, fig.width=12, fig.height=6}
+source("master_scripts/par.R")
+```
+
+* Figure 1.
+Local light environment for sun and shade leaves for each gas exchange campaign.
+Means and standard errors represent 1m intregrated PPFD (n=5), measured with a ceptometer, at the canopy height of each selected leaf.
+
 There was a ~75% reduction in PPFD from sun to shave leaves.
@@
 
-* ACi curves were measured at saturating light for both sun and shade leaves.
-Both Vc<sub>max</sub> and J<sub>max</sub> are reduced in shade leaves with little effect of the temperature treatment(AT/ET).
-ACi curves were taken once towards the end of the experiment on well watered trees
+* Figure2.
+Predicted photosynthetic CO2 response (ACi))curves for sun and shade leaves at elevated and ambient temperature treatments.
+Aci curves were developed at 25C and at saturating light (1800 umols m^-1 s^-1).
+ACi curves were taken once on well watered trees before the initation of the drought treatment.
+
+```{r, results='asis'}
+aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
+names(aci_leaf) <-c("Leaf", "Temperature", "Vcmax", "Jmax", "Vcmax_se", "Jmax_se")
+
+aci_leaf2 <- aci_leaf[,c(1,2,3,5,4,6)]
+
+panderOptions("digits", 3)
+pander(aci_leaf2, caption="Leaf Biochemistry")
+```
+
+![wateruse](wateruse.png)
+
+* Figure 3.
+Response of instantaneous water use efficiency (ITE) to leaf vapor pressure deficit (VPD) (a) and response of leaf photosynthesis rates to stomatal conductance (b) for sun and shade leaves at elevated and ambient temperature treatments.
+For (a)  lines represent predictions from the optimal ITE model with k=0.05 and a specific g1 value for each leaf type and treatment.
+For (b) lines represent 95% confidence intervals from a general additive model with k=5.
+
+![c13N](c13_nitro.png)
+
+* Figure 4. The relationship between leaf c13 and nitrogen[area] for sun and shade leaves at elevated and ambient temperature treatments.
+
+```{r gm, fig.width=8, fig.height=10}
+source("master_scripts/gm_figures.R")
+```
+
+* Figure 5.
+Response of mesophyll conductance to leaf temperature (a) and the response of leaf photosynthesis to mesophyll conductance (b) for sun and shade leaves at elevated and ambient temperature treatments.
+The mean difference in mesophyll conductance (c) and photosyntheis (d) between sun and shade leaves with standard errors.
+For (a) and (b) lines represent 95% confidence intervals generated with bootstrapping methods (n=999) from mixed effect linear model equations with chambers as a random effect.
+
+```{r lights, fig.width=8, fig.height=10}
+source("master_scripts/gm_lightson.R")
+```
+
+* Figure 6.
+Response of photosynthesis to mesophyll conductance following a maximum sunfleck response of shade leaves (a) and the mean difference in mesophyll conductance (c) and photosyntheis (d) to sun flecks in shade leaves with standard errors.
+The sunfleck was simulated by increasing the PPFD to the level of the sun leaf in the same chamber.
+The response of leaf physiology to sun flecks was recorded once both photosynthesis and stomatal conductance were stable in the gas exchange cuvette (ca.
+25m).

### 2015-06-24 master_scripts/gm_lightson.R

@@ -1,0 +1,53 @@
+# source("functions and packages/functions.R")
+# source("master_scripts/plot_objects.R")
+# source("functions and packages/packages.R")
+
+gasex <- read.csv("calculated_data/gmes_wellwatered.csv")
+lightson <- gasex[gasex$leaflight != "sun-high",]
+
+lightson_agg <- summaryBy(Photo+gm~ chamber+id+leaf +light+temp+leaflight, data=lightson, FUN=mean, keep.names=TRUE)
+
+  lightson_agg$leaf <- gsub("s", "S", lightson_agg$leaf)
+  lightson_agg$light <- gsub("high", "Sun-light", lightson_agg$light)
+  lightson_agg$light <- gsub("low", "Shade-light", lightson_agg$light)
+
+###plotting----------------------------------------------------------------------------------------------  
+  
+# windows(11,8.5)  
+
+#png(filename = "markdown/gm_lightson.png", width = 11, height = 8.5, units = "in", res= 400)
+
+par(fig=c(0,1,0.45,1), new=TRUE)
+
+par(mar=c(4,6,2,5), las=1) 
+plot(Photo~gm, data=lightson_agg, subset=leaflight=="shade-high", col=lightscol, ylim=c(5,25), xlim=c(0,0.3), 
+     xlab=gmlab, ylab="", pch=c(16, 17)[pch=lightson_agg$temp],cex=2, cex.axis=1.25, cex.lab=1.25)
+points(Photo~gm, data=lightson_agg, subset=leaflight=="shade-low", col=shacol, pch=c(16, 17)[pch=lightson_agg$temp], cex=2)
+
+legend("topleft", lightleg2, pch=c(16,17,16,17),inset = 0.02, col=collights, bty='n', cex=1.25) 
+text(x=.3, y=24.5, "(a)", cex=1.25)
+mtext(satlab, side=2, line=3.5, cex=1.25, las=3)
+
+###bar plots
+par(fig=c(0,0.5,0,.45), new=TRUE)
+bar(gm, light, lightson_agg, col=c(shacol,lightscol), xlab="", half.errbar=FALSE, ylim=c(0, 0.2),
+    mar=c(3,6,1,0), ylab="", cex.axis=1.25, cex.lab = 1.25, legend=F, cex.names=1.25)
+text(x=.25, y=.18, "(b)", cex=1.25)
[…]

### 2015-06-24 master_scripts/par.R

@@ -2,9 +2,9 @@ source("functions and packages/functions.R")
 source("functions and packages/packages.R")
 source("master_scripts/plot_objects.R")
 
-par <- read.csv("raw data/par.csv")
-
-treatments <- read.csv("raw data/temp_trt.csv")
+# par <- read.csv("raw data/par.csv")
+# 
+# treatments <- read.csv("raw data/temp_trt.csv")
 
 #format function
 par<- parformat(par)
@@ -21,29 +21,29 @@ parbar$leaf_type <- gsub("s", "S", parbar$leaf_type)
 par_agg <- summaryBy(par ~ leaf_type, data=parbar, FUN=mean)
 
 ###Stats---------------------------------------------------------------------------------------------------------
-library(visreg)
-library(multcomp)
-library(nlme)
+# library(visreg)
+# library(multcomp)
+# library(nlme)
 
 #examite data with boxplots, then remove any outliers
-boxplot(par~temp, data=par_leaf[par_leaf$leaf_type =="sun",])
-boxplot(par~temp, data=par_leaf[par_leaf$leaf_type =="shade",])
+# boxplot(par~temp, data=par_leaf[par_leaf$leaf_type =="sun",])
+# boxplot(par~temp, data=par_leaf[par_leaf$leaf_type =="shade",])
 
 
 ###lma not different between leaf types or temp treatment.
-par_leaf <- lme(par ~ leaf_type, random=~1|chamber, data=par_leaf)
-summary(par_leaf)
-anova(par_leaf)
+# par_leaf <- lme(par ~ leaf_type, random=~1|chamber, data=par_leaf)
+# summary(par_leaf)
+# anova(par_leaf)
 
 
 ###PLOTTING------------------------------------------------------------------------------------------------
 
-windows(7,5)
+#windows(7,5)
 bar(par, c(leaf_type, month), parbar, col=c("yellowgreen", "green4"), xlab="", ylab="", ylim=c(0, 2000), 
       half.errbar=FALSE)
 title(ylab=parlab, mgp=ypos)
-dev.copy2pdf(file="master_scripts/figures/ppfd.pdf")
-dev.off()
+# dev.copy2pdf(file="master_scripts/figures/ppfd.pdf")
+# dev.off()
 
 
 ###for png

### 2015-06-24 scripts/canopy_chem.R

@@ -136,12 +136,10 @@ dev.copy2pdf(file="master_scripts/figures/c13_nitro.pdf")
 dev.off()
 
 windows(8,8)
+par(mar=c(5,5,2,2), cex=1.25, cex=1.25, las=1)
 plot(c13 ~ leafN_area, data=canopy_chem, pch=16, col=as.factor(leaf), ylim=c(-35,-25), xlim=c(0,4), 
       cex=1.25, xlab=narealab, ylab=c13lab)
 
-
-plot(c13 ~ leafN_area, data=canopy_chem, pch=16, col=c(shacol, suncol), ylim=c(-35,-25), xlim=c(0,4), 
-     cex=1.25, xlab=narealab, ylab=c13lab)
 
 
 

### 2015-06-25 markdown/WTC3_figures.Rmd
added degree symbol
tweaks to rmd

@@
 treatments <- read.csv("raw data/chamber_trt.csv")
-
 ```
@@
 * WTCIII was a warming experiment with ambient and +3°C warming (n=6).  A drought manipulation was added in the last few months of the experiment but is not included here.
-* Sun leaves were measured in the upper canopy at their current PPFD.  Shade leaves were measured in the lower/middle canopy, first at their current PPFD and then again at sun leaf PPFD.
+* Sun leaves were measured in the upper canopy at their current PPFD.
+Shade leaves were measured in the lower/middle canopy, first at their current PPFD and then again at sun leaf PPFD as a simulated sunfleck.
 * New fully expanded leaves were measured. Shade leaves were sampled into the middle canopy as the experiment progressed to avoid issues with leaf age.
-* Leaf physiological measurements were taken at their current temperature (ambient, +3°C). Thus, data below represent a large range of temperatures from ~18-35°C.
+* Leaf physiological measurements were taken at their current temperature (ambient, +3°C). Thus, data below represent a large range of temperatures from 18-35°C.
 
-```{r par, fig.width=12, fig.height=6}
+```{r par, fig.width=8, fig.height=6}
 source("master_scripts/par.R")
 ```  
-
-* Figure 1.
+Figure 1.
 Local light environment for sun and shade leaves for each gas exchange campaign.
-Means and standard errors represent 1m intregrated PPFD (n=5), measured with a ceptometer, at the canopy height of each selected leaf.
-
-There was a ~75% reduction in PPFD from sun to shave leaves.
+Means and standard errors represent 1m integrated PPFD (n=5), measured with a ceptometer, at the canopy height of each selected leaf.
 
 ![aci](aci_curves.png)
-
-* Figure2.
-Predicted photosynthetic CO2 response (ACi))curves for sun and shade leaves at elevated and ambient temperature treatments.
-Aci curves were developed at 25C and at saturating light (1800 umols m^-1 s^-1).
-ACi curves were taken once on well watered trees before the initation of the drought treatment.
+Figure2.
+Predicted photosynthetic CO~2~ response (ACi) curves for sun and shade leaves at elevated and ambient temperature treatments.
+ACi curves were developed at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+ACi curves were taken once on well watered trees before the initiation of the drought treatment.
 
@@
 panderOptions("digits", 3)
-pander(aci_leaf2, caption="Leaf Biochemistry")
+pander(aci_leaf2, caption = "Table 1.  Biochemical parameters estimated from ACi curves above")
 ```
 
-![wateruse](wateruse.png)
-
-* Figure 3.
+```{r wateruse, fig.width=8, fig.height=10}
+source("master_scripts/wateruse.R")
+```  
+Figure 3.
 Response of instantaneous water use efficiency (ITE) to leaf vapor pressure deficit (VPD) (a) and response of leaf photosynthesis rates to stomatal conductance (b) for sun and shade leaves at elevated and ambient temperature treatments.
 For (a)  lines represent predictions from the optimal ITE model with k=0.05 and a specific g1 value for each leaf type and treatment.
-For (b) lines represent 95% confidence intervals from a general additive model with k=5.
+For (b) dashed lines represent 95% confidence intervals from a general additive model with k=5.
 
-![c13N](c13_nitro.png)
-
-* Figure 4. The relationship between leaf c13 and nitrogen[area] for sun and shade leaves at elevated and ambient temperature treatments.
+```{r c13n, fig.width=8, fig.height=6}
+source("master_scripts/c13_Narea.R")
+```  
+Figure 4. The relationship between leaf  $\delta$^13^C and nitrogen~area~ for sun and shade leaves at elevated and ambient temperature treatments.
 
@@
 ```  
-
-* Figure 5.
+Figure 5.
 Response of mesophyll conductance to leaf temperature (a) and the response of leaf photosynthesis to mesophyll conductance (b) for sun and shade leaves at elevated and ambient temperature treatments.
-The mean difference in mesophyll conductance (c) and photosyntheis (d) between sun and shade leaves with standard errors.
-For (a) and (b) lines represent 95% confidence intervals generated with bootstrapping methods (n=999) from mixed effect linear model equations with chambers as a random effect.
+The mean difference in mesophyll conductance (c) and photosynthesis (d) between sun and shade leaves with standard errors.
+For (a) and (b) lines represent 95% confidence intervals generated with bootstrapping methods (n=999) from mixed effect linear model equations with chamber as a random effect.
 
@@
 ```  
-
-* Figure 6.
-Response of photosynthesis to mesophyll conductance following a maximum sunfleck response of shade leaves (a) and the mean difference in mesophyll conductance (c) and photosyntheis (d) to sun flecks in shade leaves with standard errors.
+Figure 6.
+Response of photosynthesis to mesophyll conductance following a maximum sunfleck response of shade leaves (a) and overall mean values of mesophyll conductance (c) and photosynthesis (d) to sun flecks in shade leaves with standard errors.
 The sunfleck was simulated by increasing the PPFD to the level of the sun leaf in the same chamber.
-The response of leaf physiology to sun flecks was recorded once both photosynthesis and stomatal conductance were stable in the gas exchange cuvette (ca.
-25m).
+The response of leaf physiology to sun flecks was recorded once both photosynthesis and stomatal conductance were stable in the gas exchange leaf cuvette (ca.
+25min).

### 2015-06-25 master_scripts/aci_curves.R

@@ -1,8 +1,8 @@
-# source("functions and packages/packages.R")
-# source("functions and packages/functions.R")
-# source("master_scripts/plot_objects.R")
-# 
-# treatments <- read.csv("raw data/temp_trt.csv") 
+source("functions and packages/packages.R")
+source("functions and packages/functions.R")
+source("master_scripts/plot_objects.R")
+
+treatments <- read.csv("raw data/temp_trt.csv") 
 
 ###make png figure for talk that also plots points from licor------------------------------------------------  
   

### 2015-06-25 master_scripts/par.R

@@ -1,7 +1,7 @@
-source("functions and packages/functions.R")
-source("functions and packages/packages.R")
-source("master_scripts/plot_objects.R")
-
+# source("functions and packages/functions.R")
+# source("functions and packages/packages.R")
+# source("master_scripts/plot_objects.R")
+# 
 # par <- read.csv("raw data/par.csv")
 # 
 # treatments <- read.csv("raw data/temp_trt.csv")
@@ -39,9 +39,9 @@ par_agg <- summaryBy(par ~ leaf_type, data=parbar, FUN=mean)
 ###PLOTTING------------------------------------------------------------------------------------------------
 
 #windows(7,5)
-bar(par, c(leaf_type, month), parbar, col=c("yellowgreen", "green4"), xlab="", ylab="", ylim=c(0, 2000), 
-      half.errbar=FALSE)
-title(ylab=parlab, mgp=ypos)
+bar(par, c(leaf_type, month), parbar, col=c(shacol,suncol), xlab="", ylab=parlab, ylim=c(0, 2000), 
+      half.errbar=FALSE, mar=c(5,5,2,2))
+#title(ylab=parlab, mgp=ypos)
 # dev.copy2pdf(file="master_scripts/figures/ppfd.pdf")
 # dev.off()
 

### 2015-06-25 master_scripts/plot_objects.R

@@ -48,7 +48,8 @@ relparlab <- expression(l[shade]:l[sun])
 relnitrolab <- expression(N[shade]:N[sun])
 relklab <- expression(K[shade]:K[sun])
 
-leaftlab <- ("Leaf Temperature  (C)")
+leaftlab2 <- ("Leaf Temperature  (C)")
+leaftlab = expression(paste("Temperature  (",degree,"C)")) 
 
 templab <- c("AT", "ET")
 leaflab <- c("shade", "sun")
@@ -76,6 +77,9 @@ lightlab <- c(lightscol,shacol)
 lightleg <- c("High light", "Low light")
 lightleg2 <- c(expression(paste("High light", "-" ,"AT")), expression(paste("High light", "-" ,"ET")),
   expression(paste("Low light","-" ,"AT")), expression(paste("Low light", "-" ,"ET")))
+
+lightleg3 <- c(expression(paste("Sun light", "-" ,"AT")), expression(paste("Sun light", "-" ,"ET")),
+               expression(paste("Shade light","-" ,"AT")), expression(paste("Shade light", "-" ,"ET")))
 
 collights<-c(lightscol, lightscol, shacol, shacol)
 

### 2015-06-25 markdown/testmark.Rmd

--- 2015-03-17 markdown/shade.Rmd
@@
 ---
-title: "Sun v Shade leaves (WTCIII)"
-author: "Court Campany"
+title: "Shade leaves"
+author: "Court"
 date: "March 17, 2015"
@@
 
-```
-This set of figures attempts to show that sun and shade leaves were different on multiple levels.  
+```{r echo=FALSE}
+library(knitr)
+opts_knit$set(root.dir = '../')
+opts_chunk$set(message=FALSE, warnings=FALSE)
 ```
 
-```{r, results='asis', echo=FALSE}
-library(knitr)
-opts_knit$set(root.dir = '../')
+```{r echo=FALSE}
+source("functions and packages/functions.R")
+source("master_scripts/plot_objects.R")
+library(doBy)
+library(xtable)
+
+#wtc treatments
+treatments <- read.csv("raw data/temp_trt.csv")
+
+###aci graph and table---------------------------------------------------------------------------------
+aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
 ```
 
-ACI curves were measured at saturating light for both sun and shade leaves.
-
-```{r, results='asis', echo=FALSE, message=FALSE}
-library(xtable)
-
-aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
-
-# aci_table <- xtable(aci_leaf)
-# digits(aci_table)[c(3,5:6)] <- 3
-# digits(aci_table)[c(4)] <- 4
-# print(aci_table, type='html', include.rownames=F,hline.after=NULL)
-# print(aci_table,  type='html', include.rownames=F, comment=FALSE)
-
- kable(aci_leaf, row.names=NA, digits=2,format = "html")
-
+```{r results='asis'}
+aci_table <- xtable(aci_leaf)
+digits(aci_table)[c(3,5:6)] <- 3
+digits(aci_table)[c(4)] <- 4
+print(aci_table,floating=FALSE)
 ```
 
-```
-Here are some predicted ACI curves for sun/shade with temperature treatements.
-Both Vcmax and Jmax are reduced in shade leaves with little effect of treatment.
-Aci curves were only taken once towards the end of the experiment on well watered trees.
-```
-```{r,  echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
-
-source("functions and packages/functions.R")
-source("master_scripts/plot_objects.R")
-aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
-
+```{r echo=FALSE}
 #simulated curves
@@
   points(shaET_sim$Ci, shaET_sim$ALEAF, col="yellowgreen", pch=21,  cex=1.1,xlab="Ci", ylab="", type="l", lwd=2, lty=2)
-  legend("bottomright", leglab2, lty=c(1,2,1,2), lwd=2,col=colaci, pt.bg=col4,inset = 0.03)
+  legend("topleft", leglab2, lty=c(1,2,1,2), lwd=2,col=colaci, pt.bg=col4,inset = 0.03)
   title(ylab=satlab, mgp=ypos)
 
-```
-
-```
-This figure shows the relationship between leaf N on an area basis and carbon isotope dicrimination between leaf types.
-```
-```{r, echo=FALSE,fig.align='center', message=FALSE}
-source("master_scripts/plot_objects.R")
+#####leaf c13v N---------------------------------------------------------------------------------------
 leaftraits <- read.csv("calculated_data/leaftraits_summary.csv")
@@
 
-```
-
-Next I show the light environment in which gas exchange measurements for sun and shade leaves were measured.
-This data was generated with a handheld par sensor for each individual leaf level and verified with ceptometer readings within at the height of the measured leaf.
-There was on average a ~75 reduction in PPFD from sun to shave leaves.
-
-The second figure shows leaf mass area for each gas exchange campaign.  Here there
-
-```{r,echo=FALSE,fig.align='center',message=FALSE, warning=FALSE}
-source("functions and packages/functions.R")
-source("master_scripts/plot_objects.R")
-library(doBy)
 ####PPFD-----------------------------------------------------------------------------------------------
@@
 parbar <- subset(par_leaf, select = c("par", "month", "leaf_type"))
-parbar$month <- factor(parbar$month, levels = Morder)
+#parbar$month <- factor(parbar$month, levels = Morder)
 
@@
 ```
-
-## Mesophyll Conducatnce
-
-```{r,echo=FALSE,fig.align='center',message=FALSE, warning=FALSE}
-source("master_scripts/plot_objects.R")
-source("functions and packages/functions.R")
-library(doBy)
-
-gmes <- read.csv("calculated_data/gmes_WTC.csv")
-
-###for analysis first subset well watered and drought treatments
-gm_drought <- gmes[gmes$drydown == "drought",]
-
-gm_water <- gmes[gmes$drydown == "control",]
-#write.csv(gm_water, "calculated_data/gmes_wellwatered.csv", row.names=FALSE)
-
-palette(c("blue", "red"))
-
-###drought----------------------------------------------------------------------------------------------------
-
-##plots
-plot(Photo~gm, data=gm_drought, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,25), xlim=c(0, .5))
-  points(Photo~gm, data=gm_drought, subset=leaflight=="shade-low",pch=1, col=temp)
-  legend("topright", templab, pch=16,inset = 0.03, col=palette())
-
-###well watered------------------------------------------------------------------------------------------------
-
-##sun-shade
-plot(Photo~gm, data=gm_water, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,25), xlim=c(0,.5), 
-     xlab=gmlab, ylab="")
-points(Photo~gm, data=gm_water, subset=leaflight=="shade-low",pch=1, col=temp)
-legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
-title(ylab=satlab, mgp=ypos, cex=1.2)
-
-##shade low-high
-plot(Photo~gm, data=gm_water, subset=leaflight=="shade-high", pch=16, col=temp, ylim=c(0,25), xlim=c(0,.5), 
-     xlab=gmlab, ylab="")
-points(Photo~gm, data=gm_water, subset=leaflight=="shade-low",pch=1, col=temp)
-legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
-title(ylab=satlab, mgp=ypos, cex=1.2)
-
-#leaf types
-sun_gm <- gm_water[gm_water$leaf =="sun",]
-shade_gm <- gm_water[gm_water$leaflight == "shade-low",]
-lightson <- gm_water[gm_water$leaflight == "shade-high",]
-
-#simple plots through time
-plot(gm~campaign, data=sun_gm, pch=16, col=temp, ylim=c(0, .6), ylab="sun")
-plot(gm~campaign, data=shade_gm, pch=16, col=temp, ylim=c(0, .6), ylab="shade")
-plot(gm~campaign, data=lightson, pch=16, col=temp, ylim=c(0, .6), ylab="shade-high")
-
-#means
-gm_agg <- summaryBy(gm~ temp+leaf+light, data=gm_water,FUN=mean, keep.names=TRUE)
-gm_drought_agg <- summaryBy(gm~ temp+leaf+light, data=gm_drought,FUN=mean, keep.names=TRUE)
-
-##gm-temp
-gm_ch<- summaryBy(gm+Photo+Cond+CTleaf ~id+chamber+temp+leaf+light+leaflight+Month, data=gm_water,FUN=mean,
-                  keep.names=TRUE)
-
-palette(c("blue", "red"))
-
-##photosynthesis and temperature
-plot(Photo~CTleaf, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,25), xlim=c(15, 40), 
-     ylab="", xlab=leaftlab)
-  points(Photo~CTleaf, data=gm_ch, subset=leaflight=="shade-low",pch=1, col=temp)
-  legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
-  title(ylab=satlab, mgp=ypos, cex=1.2)
-
-plot(gm~CTleaf, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,.5), xlim=c(15, 40), 
-     ylab="", xlab=leaftlab)
-points(gm~CTleaf, data=gm_ch, subset=leaflight=="shade-low",pch=1, col=temp)
-legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
-title(ylab=gmlab, mgp=ypos, cex=1.2)
-
-plot(Photo~gm, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,25), xlim=c(0,.5), 
-     ylab="", xlab=gmlab)
-points(Photo~gm, data=gm_ch, subset=leaflight=="shade-low",pch=1, col=temp)
-legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
-title(ylab=satlab, mgp=ypos, cex=1.2)
-
-plot(gm~Cond, data=gm_ch, subset=leaflight=="sun-high", pch=16, col=temp, ylim=c(0,.4), xlim=c(0,.4), 
-     ylab="", xlab=gmlab)
-points(gm~Cond, data=gm_ch, subset=leaflight=="shade-low",pch=1, col=temp)
-legend("topright", templab, pch=16,inset = 0.03, col=palette()) 
-title(ylab=condlab, mgp=ypos, cex=1.2)
-
-###make a table of means and se of gm
-tablenames <- c( "Leaf Type", "Light", "Temperature","Gm", "Gm(se)")
-
-gm_dfr <- summaryBy(gm~ leaf+light+temp, data=gm_water,FUN=c(mean,se))
-names(gm_dfr) <- tablenames
-
-library(xtable)
-
-gm_table <- xtable(gm_dfr)
-digits(gm_table)[4:5] <- 4
-print(gm_table, type='html')
-
-```

### 2015-06-25 markdown/WTC3_figures.Rmd
more minor tweaks to figure markdown

@@
 
-![aci](aci_curves.png)
+```{r acicurves, fig.width=8, fig.height=6}
+source("master_scripts/aci_curves.R")
+```  
 Figure2.
@@
 Response of instantaneous water use efficiency (ITE) to leaf vapor pressure deficit (VPD) (a) and response of leaf photosynthesis rates to stomatal conductance (b) for sun and shade leaves at elevated and ambient temperature treatments.
-For (a)  lines represent predictions from the optimal ITE model with k=0.05 and a specific g1 value for each leaf type and treatment.
+For (a)  lines represent predictions from the optimal ITE model with a g1 value for each leaf type and treatment.
 For (b) dashed lines represent 95% confidence intervals from a general additive model with k=5.
@@
 25min).
+
+```{r table, results="asis"}
+ge_data <- read.csv("master_scripts/ge_table.csv")
+ge_tab <- ge_data[,1:8]
+
+colnames(ge_tab) <- c("Leaf", "Temperature", "Light Level", "Photosynthesis", "g~s~","g~m~",
+                      "Ci", "Transpiration")
+pandoc.table(ge_tab, caption = "Gas exchange data for sun and shade leaves", split.cells = c(1,1,1,1,1,1,1,1))
+```

### 2015-06-25 master_scripts/data_tables.R

@@ -21,6 +21,10 @@ ge_dry <- ge_overall[ge_overall$drydown == "drought",]
 ge_table <- ge_wet[ c(6,5,2,4,1,3),]
 
 ge_vars <- ge_table[, 1:3]
+  ge_vars$leaflight <- gsub("sun-", "", ge_vars$leaflight)
+  ge_vars$leaflight <- gsub("shade-", "", ge_vars$leaflight)
+  ge_vars$leaflight <- gsub("high", "High", ge_vars$leaflight)
+  ge_vars$leaflight <- gsub("l", "L", ge_vars$leaflight)
 ge_means <- ge_table[, c(5,6,9,7,8, 10)]
 ge_se <- ge_table[, c(11,12,15,13,14,16)]
 
@@ -42,7 +46,7 @@ ge_table3 <- cbind(ge_vars, ge1)
   
 ###add sigletters
 
-
+write.csv(ge_table3, "master_scripts/ge_table.csv", row.names=FALSE)
   
 #####vaiarbles with no shade high light
 ####leafk, PAR, ACI, 

### 2015-06-25 master_scripts/gm_lightson.R

@@ -25,14 +25,14 @@ plot(Photo~gm, data=lightson_agg, subset=leaflight=="shade-high", col=lightscol,
 points(Photo~gm, data=lightson_agg, subset=leaflight=="shade-low", col=shacol, pch=c(16, 17)[pch=lightson_agg$temp], cex=1.5)
 
 legend("topleft", lightleg3, pch=c(16,17,16,17),inset = 0.02, col=collights, bty='n', cex=1) 
-text(x=.3, y=24.5, "(a)", cex=1.25)
+text(x=.3, y=24.75, "(a)", cex=1.25)
 mtext(satlab, side=2, line=3.5, cex=1.25, las=3)
 
 ###bar plots
 par(fig=c(0,0.5,0,.45), new=TRUE)
 bar(gm, light, lightson_agg, col=c(shacol,lightscol), xlab="", half.errbar=FALSE, ylim=c(0, 0.2),
     mar=c(3,6,1,0), ylab="", cex.axis=1.25, cex.lab = 1.25, legend=F, cex.names=1.25)
-text(x=.25, y=.18, "(b)", cex=1.25)
+text(x=.25, y=.19, "(b)", cex=1.25)
 mtext(gmlab, side=2, line=3.5, cex=1.25, las=3)
 
 par(fig=c(.5,1,0,0.45), new=TRUE)
@@ -41,7 +41,7 @@ bar(Photo, light, lightson_agg, col=c(shacol,lightscol), xlab="", half.errbar=FA
 axis(side=4, labels=TRUE, cex.axis=1.21)
 mtext(photolab, side=4, line=3.5, cex=1.25, las=3)
 
-text(x=.25, y=18, "(c)", cex=1.25)
+text(x=.25, y=19, "(c)", cex=1.25)
 
 #dev.copy(png, file="markdown/gm_lightson.png", width=1100, height=800)
 #dev.copy2pdf(file="master_scripts/paper_figures/gm_lightson.pdf")

### 2015-06-30 markdown/WTC3_figures.Rmd
made changes to master scripts plots so they will run in rmd

@@
 
+```{r c13n, fig.width=8, fig.height=6}
+source("master_scripts/leafN_c13_vcmax.R")
+```  
+Figure 4. The relationship between leaf  $\delta$^13^C (a) and Vc~max~ (b) with nitrogen~area~ for sun and shade leaves at elevated and ambient temperature treatments.
+
 ```{r wateruse, fig.width=8, fig.height=10}
@@
 
-```{r c13n, fig.width=8, fig.height=6}
-source("master_scripts/c13_Narea.R")
-```  
-Figure 4. The relationship between leaf  $\delta$^13^C and nitrogen~area~ for sun and shade leaves at elevated and ambient temperature treatments.
-
 ```{r gm, fig.width=8, fig.height=10}
@@
 Response of mesophyll conductance to leaf temperature (a) and the response of leaf photosynthesis to mesophyll conductance (b) for sun and shade leaves at elevated and ambient temperature treatments.
-The mean difference in mesophyll conductance (c) and photosynthesis (d) between sun and shade leaves with standard errors.
 For (a) and (b) lines represent 95% confidence intervals generated with bootstrapping methods (n=999) from mixed effect linear model equations with chamber as a random effect.
 
-```{r lights, fig.width=8, fig.height=10}
+```{r lights, fig.width=8, fig.height=6}
 source("master_scripts/gm_lightson.R")
@@
 Figure 6.
-Response of photosynthesis to mesophyll conductance following a maximum sunfleck response of shade leaves (a) and overall mean values of mesophyll conductance (c) and photosynthesis (d) to sun flecks in shade leaves with standard errors.
+Response of photosynthesis to mesophyll conductance following a maximum sunfleck response of shade leaves.
 The sunfleck was simulated by increasing the PPFD to the level of the sun leaf in the same chamber.
@@
 25min).
+
+```{r barplots, fig.width=12, fig.height=6}
+source("master_scripts/physiology_barplots.R")
+```  
+The mean difference in stomatal conductance (a), mesophyll conductance (b) and photosynthesis (c) between sun, shade and shade leaves under high light with standard errors
+
+```{r gmparN, fig.width=8, fig.height=10}
+source("master_scripts/gm_par_nitro.R")
+``` 
+The relationship between mesophyll conductance, leaf PPFD and leaf nitrogen.
 

### 2015-06-30 master_scripts/gm_par_nitro.R

@@ -1,8 +1,8 @@
-source("functions and packages/functions.R")
-source("functions and packages/packages.R")
-source("master_scripts/plot_objects.R")
-
-treatments <- read.csv("raw data/temp_trt.csv")
+# source("functions and packages/functions.R")
+# source("functions and packages/packages.R")
+# source("master_scripts/plot_objects.R")
+# 
+# treatments <- read.csv("raw data/temp_trt.csv")
 
 ####read and format PAR data----------------------------------------------------------------------------------------------
 par <- read.csv("raw data/par.csv")
@@ -62,55 +62,54 @@ library(visreg)
 # library(lme4)
 # library(lmerTest)
 
-gmpar_sha <- lme(gm~ par, random=~1|chamber, data=leafdat2, subset=leaflight == "shade-low")
-summary(gmpar_sha)
-anova(gmpar_sha)
-visreg(gmpar_sha)  
-
-gmpar_sun <- lme(gm~ par, random=~1|chamber, data=leafdat2, subset=leaflight == "sun-high")
-  summary(gmpar_sun)
-  anova(gmpar_sun)
-  visreg(gmpar_sun)
+# gmpar_sha <- lme(gm~ par, random=~1|chamber, data=leafdat2, subset=leaflight == "shade-low")
+#   summary(gmpar_sha)
+#   anova(gmpar_sha)
+#   visreg(gmpar_sha)  
+# 
+# gmpar_sun <- lme(gm~ par, random=~1|chamber, data=leafdat2, subset=leaflight == "sun-high")
+#   summary(gmpar_sun)
+#   anova(gmpar_sun)
+#   visreg(gmpar_sun)
 gmpar_sun2 <- lm(gm~ par, data=leafdat2, subset=leaflight == "sun-high")
-anova(gmpar_sun, gmpar_sun2)
+# anova(gmpar_sun, gmpar_sun2)
 ###lower AIC with chamber as random but models not different
   
-gmpar_fleck <- lme(gm~ par, random=~1|chamber, data=fleck_dat)
-  summary(gmpar_fleck)
-  anova(gmpar_fleck)
-  visreg(gmpar_fleck)   
+# gmpar_fleck <- lme(gm~ par, random=~1|chamber, data=fleck_dat)
+#   summary(gmpar_fleck)
+#   anova(gmpar_fleck)
[…]

### 2015-06-30 master_scripts/leafN_c13_vcmax.R

@@ -31,8 +31,9 @@ N_aci <- merge(aciparam, Nagg, by= c("chamber", "leaf"))
   
 
 ###plot 13 vs Narea------------------------------------------------------------------------------------------------------
-
-windows(8,8)
+palette(c(shacol, suncol))
+ 
+#windows(8,8)
 par(mfrow=c(2,1))  
   
 #png(filename = "markdown/c13_nitro.png", width = 11, height = 8.5, units = "in", res= 400)
@@ -47,5 +48,5 @@ par(mar=c(5,5,0,2), cex=1.25, las=1,cex.axis=1.25, cex.axis=.8, cex.lab=.96)
 plot(Vcmax~leafN_area.mean, data=N_aci, col=as.factor(leaf),  ylim=c(0, 133),xlim=c(0,4.5),
      pch=c(16, 17)[pch=N_aci$temp],xlab=narealab, ylab=vclab)
 
-dev.copy2pdf(file="master_scripts/paper_figures/leafN_c13_vcmax.pdf")
-dev.off()
+# dev.copy2pdf(file="master_scripts/paper_figures/leafN_c13_vcmax.pdf")
+# dev.off()

### 2015-06-30 master_scripts/physiology_barplots.R

@@ -1,8 +1,8 @@
 ##physiology bar plots of gs, gm and A overall means with sun, shade and lights on
 
-source("functions and packages/functions.R")
-source("master_scripts/plot_objects.R")
-source("functions and packages/packages.R")
+# source("functions and packages/functions.R")
+# source("master_scripts/plot_objects.R")
+# source("functions and packages/packages.R")
 
 gasex <- read.csv("calculated_data/gmes_wellwatered.csv")
 
@@ -24,7 +24,7 @@ cols <- c(shacol, suncol, lightscol)
 
 #1: stomatal conductance
 
-windows(12,6)
+# windows(12,6)
 
 par(mfrow=c(1,3))
 
@@ -50,5 +50,5 @@ bar(Photo, leaflight, gasex_agg, col=cols, xlab="", half.errbar=FALSE, ylim=c(0,
 abline(v=2.5, lty=5)
 text(x=.25, y=19.5, "(c)", cex=1.5)
 
-dev.copy2pdf(file="master_scripts/paper_figures/physiology_bar.pdf")
-dev.off()
+# dev.copy2pdf(file="master_scripts/paper_figures/physiology_bar.pdf")
+# dev.off()

### 2015-07-01 markdown/WTC3_figures.Rmd
fixing plots for manuscript
major changes in legends

changed colors

started aci model plot

--- 2015-06-25 markdown/WTC3_figures.Rmd
@@
 
+#### <u>Figures</u>
 ```{r par, fig.width=8, fig.height=6}
@@
 
-![aci](aci_curves.png)
+```{r acicurves, fig.width=8, fig.height=6}
+source("master_scripts/aci_vcmaxnitro.R")
+```  
 Figure2.
 Predicted photosynthetic CO~2~ response (ACi) curves for sun and shade leaves at elevated and ambient temperature treatments.
-ACi curves were developed at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+ACi curves were developed at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^)(a).
+The relationship between Vc~max~  and leaf nitrogen, on an area basis, for sun and shade leaves at elevated and ambient temperature treatments (b).
 ACi curves were taken once on well watered trees before the initiation of the drought treatment.
-
-```{r, results='asis'}
-aci_leaf <- read.csv("calculated_data/tdl_aci.csv")
-names(aci_leaf) <-c("Leaf", "Temperature", "Vcmax", "Jmax", "Vcmax_se", "Jmax_se")
-
-aci_leaf2 <- aci_leaf[,c(1,2,3,5,4,6)]
-
-panderOptions("digits", 3)
-pander(aci_leaf2, caption = "Table 1.  Biochemical parameters estimated from ACi curves above")
-```
 
@@
 Figure 3.
-Response of instantaneous water use efficiency (ITE) to leaf vapor pressure deficit (VPD) (a) and response of leaf photosynthesis rates to stomatal conductance (b) for sun and shade leaves at elevated and ambient temperature treatments.
-For (a)  lines represent predictions from the optimal ITE model with k=0.05 and a specific g1 value for each leaf type and treatment.
-For (b) dashed lines represent 95% confidence intervals from a general additive model with k=5.
+Response of instantaneous water use efficiency (ITE) to leaf vapor pressure deficit (VPD) for sun leaves and shade leaves at both low and high light (a) and the relationship between leaf $\delta$^13^C and leaf nitrogen~area~ for sun and shade leaves at elevated and ambient temperature treatments.
+For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g1 value for each leaf type and treatment.
 
-```{r c13n, fig.width=8, fig.height=6}
-source("master_scripts/c13_Narea.R")
+```{r gmgs, fig.width=8, fig.height=10}
+source("master_scripts/gmgs")
 ```  
-Figure 4. The relationship between leaf  $\delta$^13^C and nitrogen~area~ for sun and shade leaves at elevated and ambient temperature treatments.
+Figure 4.
+The response of leaf photosynthesis rates to stomatal conductance (a) and mesophyll conductance (b) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
+Dashed lines represent 95% confidence intervals from a general additive model with a smoothing factor of 5.
+The response of shade leaf physiology to high light was recorded once both photosynthesis and stomatal conductance were stable in the gas exchange leaf cuvette (ca.
+25min).
 
-```{r gm, fig.width=8, fig.height=10}
-source("master_scripts/gm_figures.R")
+```{r cicc, fig.width=12, fig.height=8}
+source("master_scripts/cicc_bar.R")
 ```  
 Figure 5.
-Response of mesophyll conductance to leaf temperature (a) and the response of leaf photosynthesis to mesophyll conductance (b) for sun and shade leaves at elevated and ambient temperature treatments.
-The mean difference in mesophyll conductance (c) and photosynthesis (d) between sun and shade leaves with standard errors.
+The average intercellular CO~2~ concentraion (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and hight light with standard errors.
+
+```{r barplots, fig.width=12, fig.height=6}
+source("master_scripts/physiology_barplots.R")
+```  
+The average stomatal conductance (a), mesophyll conductance (b) and photosynthesis (c) between sun leavse and  shade leaves at both low and high light with standard errors.
+
+```{r acimod, fig.width=12, fig.height=6}
+source("master_scripts/aci_acc_model.R")
+```  
+The average stomatal conductance (a), mesophyll conductance (b) and photosynthesis (c) between sun leavse and  shade leaves at both low and high light with standard errors
+
+#### <u>Supplementary Figures</u>
+```{r gmparN, fig.width=8, fig.height=10}
+source("master_scripts/gm_par_nitro.R")
+``` 
+Figure S1. The relationship between mesophyll conductance, leaf PPFD and leaf nitrogen.
+
+```{r gmT, fig.width=8, fig.height=10}
+source("master_scripts/gm_temp.R")
+``` 
+Figure S2.
+Response of mesophyll conductance to leaf temperature  for sun and shade leaves at elevated and ambient temperature treatments.
 For (a) and (b) lines represent 95% confidence intervals generated with bootstrapping methods (n=999) from mixed effect linear model equations with chamber as a random effect.
-
-```{r lights, fig.width=8, fig.height=10}
-source("master_scripts/gm_lightson.R")
-```  
-Figure 6.
-Response of photosynthesis to mesophyll conductance following a maximum sunfleck response of shade leaves (a) and overall mean values of mesophyll conductance (c) and photosynthesis (d) to sun flecks in shade leaves with standard errors.
-The sunfleck was simulated by increasing the PPFD to the level of the sun leaf in the same chamber.
-The response of leaf physiology to sun flecks was recorded once both photosynthesis and stomatal conductance were stable in the gas exchange leaf cuvette (ca.
-25min).

### 2015-07-01 master_scripts/cicc_bar.R

@@ -1,8 +1,8 @@
 ##physiology bar plots of ci, cc, and drawdown
 
-source("functions and packages/functions.R")
-source("master_scripts/plot_objects.R")
-source("functions and packages/packages.R")
+# source("functions and packages/functions.R")
+# source("master_scripts/plot_objects.R")
+# source("functions and packages/packages.R")
 
 gasex <- read.csv("calculated_data/gmes_wellwatered.csv")
 
@@ -30,7 +30,7 @@ cols <- c(shacol2, suncol2, lightscol2)
 
 #1: CI
 
-windows(12,6)
+# windows(12,6)
 
 par(mfrow=c(1,3))
 
@@ -38,7 +38,7 @@ bar(Ci, leaflight, gasex_agg, col=cols, half.errbar=FALSE, ylim=c(0, 325),names.
     mar=c(5,5,2,1), ylab=cilab, cex.axis=1.25, cex.lab = 1.5, cex.names=1.5,legend=F)
 abline(v=2.5, lty=5)
 text(x=.25, y=315, "(a)", cex=1.5)
-mtext("Shade", side=1,at=.725, line=2.5 )
+mtext("Shade \n Low Light", side=1,at=.725, line=3.5 )
 mtext("Sun", side=1,at=1.9, line=2.5 )
 mtext("Shade \n High Light", side=1,at=3.15, line=3.5 )
 
@@ -47,7 +47,7 @@ bar(Cc, leaflight, gasex_agg, col=cols, xlab="", half.errbar=FALSE, ylim=c(0, 32
     mar=c(5,5,2,1), ylab=cclab,  cex.axis=1.25, cex.lab = 1.5, cex.names=1.5,legend=F)
 abline(v=2.5, lty=5)
 text(x=.25, y=315, "(b)", cex=1.5)
-mtext("Shade", side=1,at=.725, line=2.5 )
+mtext("Shade \n Low Light", side=1,at=.725, line=3.5 )
 mtext("Sun", side=1,at=1.9, line=2.5 )
 mtext("Shade \n High Light", side=1,at=3.15, line=3.5 )
 
@@ -56,9 +56,9 @@ bar(drawdown, leaflight, gasex_agg, col=cols, xlab="", half.errbar=FALSE, ylim=c
     mar=c(5,5,2,2), ylab=drawdownlab,  cex.axis=1.25, cex.lab = 1.5, cex.names=1.5,legend=F)
[…]

### 2015-07-01 master_scripts/gmgs.R

@@ -1,6 +1,6 @@
-source("functions and packages/functions.R")
-source("master_scripts/plot_objects.R")
-source("functions and packages/packages.R")
+# source("functions and packages/functions.R")
+# source("master_scripts/plot_objects.R")
+# source("functions and packages/packages.R")
 
 #read in gm data set (no drought) and Cibar(discrimination)-------------------------------------------------------
 gmes <- read.csv("calculated_data/gmes_wellwatered.csv")
@@ -74,7 +74,7 @@ agm_fleck <-  read.csv( "master_scripts/bootstrap_results/agm_fleck.csv")
 ####PLOTTING: 2 panel figure with gm, gs, A---------------------------------------------------------------------------------
 
 #1: panel Photosynthesis vs gs (gam plots)
-windows(10, 12)
+# windows(10, 12)
 
 par(mfrow=c(2,1))
 
@@ -97,14 +97,15 @@ plot(Photo~Cond, data=gm_sunsha, subset=leaflight=="sun-high",  col=suncol, ylim
   lines(gsfleck_seq, fleckupr, lty=2, lwd=2,col=lightscol)
   lines(gsfleck_seq, flecklwr, lty=2, lwd=2,col=lightscol)
   lines(gsfleck_seq, gsfleck_pred$fit, lty=1, lwd=2,col=lightscol)
-  legend("topleft", leglab3, pch=rep(c(16,17),3), col=leafcols,inset = 0.01, bty='n',cex=.8)
+  legend("topleft", alllab, pch=c(16,16,16,16,17), col=allcols,inset = 0.01, bty='n',cex=.8)
+
 
   text(x=.5, y=24.8, "(a)", cex=1)
 
 ####panel2: gm vs A
 
 par(mar=c(4,4,1,1), cex=1.25, las=1, cex.axis=.8, cex.lab=1, mgp=c(2.5,1,0))
-plot(Photo~gm, data=sundat,  col=suncol, ylim=c(5,25), xlim=c(0,.4), xlab=gmlab, ylab=satlab, 
+plot(Photo~gm, data=sundat,  col=suncol, ylim=c(5,25), xlim=c(0,.5), xlab=gmlab, ylab=satlab, 
      pch=c(16, 17)[pch=gm_sunsha$temp])
[…]

### 2015-07-01 master_scripts/physiology_barplots.R

@@ -1,8 +1,8 @@
 ##physiology bar plots of gs, gm and A overall means with sun, shade and lights on
 
-source("functions and packages/functions.R")
-source("master_scripts/plot_objects.R")
-source("functions and packages/packages.R")
+# source("functions and packages/functions.R")
+# source("master_scripts/plot_objects.R")
+# source("functions and packages/packages.R")
 
 gasex <- read.csv("calculated_data/gmes_wellwatered.csv")
 
@@ -24,7 +24,7 @@ cols <- c(shacol2, suncol2, lightscol2)
 
 #1: stomatal conductance
 
-windows(12,6)
+# windows(12,6)
 
 par(mfrow=c(1,3))
 
@@ -32,7 +32,7 @@ bar(Cond, leaflight, gasex_agg, col=cols, half.errbar=FALSE, ylim=c(0, 0.3),name
     mar=c(5,5,2,1), ylab=condlab, cex.axis=1.25, cex.lab = 1.5, cex.names=1.5,legend=F)
 abline(v=2.5, lty=5)
 text(x=.25, y=.2925, "(a)", cex=1.5)
-mtext("Shade", side=1,at=.725, line=2.5 )
+mtext("Shade \n Low Light", side=1,at=.725, line=3.5 )
 mtext("Sun", side=1,at=1.9, line=2.5 )
 mtext("Shade \n High Light", side=1,at=3.15, line=3.5 )
 
@@ -41,7 +41,7 @@ bar(gm, leaflight, gasex_agg, col=cols, xlab="", half.errbar=FALSE, ylim=c(0, 0.
     mar=c(5,5,2,1), ylab=gmlab,  cex.axis=1.25, cex.lab = 1.5, cex.names=1.5,legend=F)
 abline(v=2.5, lty=5)
 text(x=.25, y=.22, "(b)", cex=1.5)
-mtext("Shade", side=1,at=.725, line=2.5 )
+mtext("Shade \n Low Light", side=1,at=.725, line=3.5 )
 mtext("Sun", side=1,at=1.9, line=2.5 )
 mtext("Shade \n High Light", side=1,at=3.15, line=3.5 )
 
@@ -50,9 +50,9 @@ bar(Photo, leaflight, gasex_agg, col=cols, xlab="", half.errbar=FALSE, ylim=c(0,
[…]

### 2015-07-01 markdown/WTC3_figures.Rmd
plot changes

@@
 
-par <- read.csv("raw data/par.csv")
-treatments <- read.csv("raw data/chamber_trt.csv")
+#treatments <- read.csv("raw data/chamber_trt.csv")
 ```
@@
 #### <u>Figures</u>
-```{r par, fig.width=8, fig.height=6}
+```{r par, fig.width=6, fig.height=6}
 source("master_scripts/par.R")
@@
 
-```{r acicurves, fig.width=8, fig.height=6}
+```{r acicurves, fig.width=8, fig.height=10}
 source("master_scripts/aci_vcmaxnitro.R")
@@
 ACi curves were developed at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^)(a).
-The relationship between Vc~max~  and leaf nitrogen, on an area basis, for sun and shade leaves at elevated and ambient temperature treatments (b).
+The relationship between Vc~max~ and leaf nitrogen~area~ for sun and shade leaves at elevated and ambient temperature treatments (b).
 ACi curves were taken once on well watered trees before the initiation of the drought treatment.
@@
 ```{r gmgs, fig.width=8, fig.height=10}
-source("master_scripts/gmgs")
+source("master_scripts/gmgs.R")
 ```  
@@
 
-```{r cicc, fig.width=12, fig.height=8}
+```{r cicc, fig.width=10, fig.height=6}
 source("master_scripts/cicc_bar.R")
@@
 
-```{r barplots, fig.width=12, fig.height=6}
+```{r barplots, fig.width=10, fig.height=6}
 source("master_scripts/physiology_barplots.R")
@@
 The average stomatal conductance (a), mesophyll conductance (b) and photosynthesis (c) between sun leavse and  shade leaves at both low and high light with standard errors.
-
-```{r acimod, fig.width=12, fig.height=6}
-source("master_scripts/aci_acc_model.R")
-```  
-The average stomatal conductance (a), mesophyll conductance (b) and photosynthesis (c) between sun leavse and  shade leaves at both low and high light with standard errors
 
@@
 
-```{r gmT, fig.width=8, fig.height=10}
+```{r gmT, fig.width=6, fig.height=6}
 source("master_scripts/gm_temp.R")

### 2015-07-01 master_scripts/aci_vcmaxnitro.R

@@ -1,6 +1,6 @@
-source("functions and packages/packages.R")
-source("functions and packages/functions.R")
-source("master_scripts/plot_objects.R")
+# source("functions and packages/packages.R")
+# source("functions and packages/functions.R")
+# source("master_scripts/plot_objects.R")
 
 treatments <- read.csv("raw data/temp_trt.csv") 
 
@@ -66,7 +66,7 @@ nitrovc_mod <- lm(Vcmax~leafN_area.mean, data=N_aci)
 
 palette(c(shacol2, suncol2))
   
-windows(8,12)
+#windows(8,12)
 
 par(fig=c(0, 1, .5, 1),mar=c(4,4,1,1), cex=1.25, las=1, cex.axis=.8, cex.lab=1, mgp=c(2.5,1,0))
   
@@ -85,23 +85,6 @@ plot(Photo~Ci ,data= acishade_clean, col=lights50col, ylim=c(0, 42), xlim=c(0,20
   legend("topleft", c("Sun", "Shade-High Light", "AT", "ET"), pch=c(-1,-1, 16,17), lty=c(1,1, 1,2),
          col=c(suncol2, lightscol2, "black", "black"),inset = 0.01, bty='n',cex=.8)
   
-  ###inset figure  
-  par(fig=c(0.525, 0.95, 0.6,0.75), mar=c(2,2,0,0),new=T, cex=.7, las=1,  cex.axis=.7, cex.lab=.7, mgp=c(2.5,1,0), tcl=-.25)
-    
-  plot(Photo~Ci ,data= acishade_clean, ylim=c(0, 15.5), xlim=c(45,300), xlab="", ylab="",xaxt="n", yaxt="n", pch="")
-  axis(2, mgp=c(3, .5, 0))
-  axis(1, mgp=c(3, .5, 0))
-
-  points(Photo~Ci ,data= acishade_clean, col=lights50col,pch=c(16, 17)[pch=acishade_clean$temp])
-  points(Photo~Ci ,data= shade_redo,col=lights50col, pch=c(16, 17)[pch=shade_redo$temp])
-  points(Photo~Ci ,data= sunaci_clean2, col=suncol50, pch=c(16, 17)[pch=sunaci_clean2$temp])
[…]

### 2015-07-01 master_scripts/par.R

@@ -2,9 +2,9 @@
 # source("functions and packages/packages.R")
 # source("master_scripts/plot_objects.R")
 # 
-# par <- read.csv("raw data/par.csv")
-# 
-# treatments <- read.csv("raw data/temp_trt.csv")
+ par <- read.csv("raw data/par.csv")
+ 
+ treatments <- read.csv("raw data/temp_trt.csv")
 
 #format function
 par<- parformat(par)
@@ -42,7 +42,7 @@ par_agg <- summaryBy(par ~ leaf_type, data=parbar, FUN=mean)
 ###PLOTTING------------------------------------------------------------------------------------------------
 # windows(8,6)
 bar(par, c(leaf_type, Month), parbar, col=c(shacol,suncol), xlab="", ylab=parlab, ylim=c(0, 2000), 
-#       half.errbar=FALSE, mar=c(5,5,2,2))
+      half.errbar=FALSE, mar=c(5,5,2,2))
 #   dev.copy2pdf(file="master_scripts/paper_figures/ppfd.pdf")
 #   dev.off()
 

### 2015-07-02 markdown/WTC3_figures.Rmd
finished smooth plot CIs of several figures.  Calculated r2 for main figure linear fits.  minor adjsutments to plots
made plots fit on a smaller window for later ease

@@
 Local light environment for sun and shade leaves for each gas exchange campaign.
-Means and standard errors represent 1m integrated PPFD (n=5), measured with a ceptometer, at the canopy height of each selected leaf.
+Means with standard errors represent 1 m integrated PPFD (n=5), measured with a ceptometer, at the canopy height of each selected leaf.
 
-```{r acicurves, fig.width=8, fig.height=10}
+```{r acicurves, fig.width=6, fig.height=8}
 source("master_scripts/aci_vcmaxnitro.R")
@@
 Figure2.
-Predicted photosynthetic CO~2~ response (ACi) curves for sun and shade leaves at elevated and ambient temperature treatments.
-ACi curves were developed at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^)(a).
-The relationship between Vc~max~ and leaf nitrogen~area~ for sun and shade leaves at elevated and ambient temperature treatments (b).
-ACi curves were taken once on well watered trees before the initiation of the drought treatment.
+(a) Photosynthetic CO~2~ response (ACi) curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.
+ACi curves were developed once for all trees, before the initiation of the drought treatment, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+(b) The relationship between Vc~max~ and mean leaf nitrogen~area~ for each chamber, including sun leaves and shade leaves at low light.
+For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional r^2^ of 0.28 and 0.35, respectively.
 
-```{r wateruse, fig.width=8, fig.height=10}
+```{r wateruse, fig.width=6, fig.height=8}
 source("master_scripts/wateruse.R")
@@
 Figure 3.
-Response of instantaneous water use efficiency (ITE) to leaf vapor pressure deficit (VPD) for sun leaves and shade leaves at both low and high light (a) and the relationship between leaf $\delta$^13^C and leaf nitrogen~area~ for sun and shade leaves at elevated and ambient temperature treatments.
+(a) Response of instantaneous water use efficiency (ITE) to leaf vapor pressure deficit (VPD) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
+(b) The relationship between leaf $\delta$^13^C and leaf nitrogen~area~ for sun leaves and shade leaves at low light.
 For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g1 value for each leaf type and treatment.
+For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional r^2^ of 0.41 and 0.45, respectively.
 
-```{r gmgs, fig.width=8, fig.height=10}
-source("master_scripts/gmgs.R")
+```{r Agmgs, fig.width=6, fig.height=8}
+source("master_scripts/Agmgs.R")
 ```  
@@
 The response of leaf photosynthesis rates to stomatal conductance (a) and mesophyll conductance (b) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
-Dashed lines represent 95% confidence intervals from a general additive model with a smoothing factor of 5.
 The response of shade leaf physiology to high light was recorded once both photosynthesis and stomatal conductance were stable in the gas exchange leaf cuvette (ca.
 25min).
+Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
+Grey areas are approximately 95% confidence intervals from the mean.
 
-```{r cicc, fig.width=10, fig.height=6}
+```{r barplots, fig.width=10, fig.height=5}
+source("master_scripts/physiology_barplots.R")
+```  
+The mean stomatal conductance (a), mesophyll conductance (b) and photosynthesis rate (c) of sun leaves and shade leaves at both low and high light with standard errors.
+
+```{r cicc, fig.width=10, fig.height=5}
 source("master_scripts/cicc_bar.R")
@@
 Figure 5.
-The average intercellular CO~2~ concentraion (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and hight light with standard errors.
-
-```{r barplots, fig.width=10, fig.height=6}
-source("master_scripts/physiology_barplots.R")
-```  
-The average stomatal conductance (a), mesophyll conductance (b) and photosynthesis (c) between sun leavse and  shade leaves at both low and high light with standard errors.
+The mean intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light with standard errors.
 
 #### <u>Supplementary Figures</u>
-```{r gmparN, fig.width=8, fig.height=10}
-source("master_scripts/gm_par_nitro.R")
+```{r gmgs, fig.width=6, fig.height=6}
+source("master_scripts/gmgs.R")
 ``` 
-Figure S1. The relationship between mesophyll conductance, leaf PPFD and leaf nitrogen.
+Figure S1.
+Mesophyll conductance as a function of stomatal conductance for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
+The dashed line is the 1:1 relationship.
 
@@
 Response of mesophyll conductance to leaf temperature for sun and shade leaves at elevated and ambient temperature treatments.
-For (a) and (b) lines represent 95% confidence intervals generated with bootstrapping methods (n=999) from mixed effect linear model equations with chamber as a random effect.
+Solid lines represent linear model fits and grey areas are approximately 95% confidence intervals from the mean.
+
+```{r gmparN, fig.width=6, fig.height=10}
+source("master_scripts/gm_par_nitro.R")
+``` 
+Figure S3.
+(a) Mesophyll conductance as a function of leaf nitrogen~area~ for sun leaves and shade leaves at low light with elevated and ambient temperature treatments.
+(b) Mesophyll conductance has a function of PPFD for all leaf types.
+(c) Leaf nitrogen~area~ as a function of PPFD for all leaf types.
+Dashed line represents linear model fits.

### 2015-07-02 master_scripts/gmgs.R

@@ -7,185 +7,18 @@ gmes <- read.csv("calculated_data/gmes_wellwatered.csv")
 
 ###get average by id
 gm_agg <- summaryBy(Photo+ Cond + gm ~ chamber+id+leaf +light+temp+leaflight+Month, data=gmes, FUN=mean, keep.names=TRUE)
-  gm_agg$leaf <- gsub("s", "S", gm_agg$leaf)
-  gm_agg$light <- gsub("high", "Sun-light", gm_agg$light)
-  gm_agg$light <- gsub("low", "Shade-light", gm_agg$light)
+gm_agg$leaf <- gsub("s", "S", gm_agg$leaf)
+gm_agg$light <- gsub("high", "Sun-light", gm_agg$light)
+gm_agg$light <- gsub("low", "Shade-light", gm_agg$light)
 
-##remove shade-high
-gm_sunsha <- gm_agg[gm_agg$leaflight != "shade-high",]
-  gm_sunsha <- droplevels(gm_sunsha)
-
-##sun-shade dataframes
-sundat <- gm_sunsha[gm_sunsha$leaflight =="sun-high",]
-shadat <- gm_sunsha[gm_sunsha$leaflight =="shade-low",]
-##dfr with lights on
-fleckdat <- gm_agg[gm_agg$leaflight == "shade-high",]
-  fleckdat <- droplevels(fleckdat)
-
-#### GS vs A data: use gam for CI of non-linear relationship between A and gs----------------------------------------
-
-#SUN leaves
-sunmod <- gam(Photo ~ s(Cond, k=5), data=gm_sunsha, subset=leaflight=="sun-high")
-
-#predict
-#get apprpriate vector of gs from sun leaves
-gsdat <- gm_sunsha[gm_sunsha$leaflight=="sun-high", "Cond"]
-
-#generate sequence and then predict
-gssun_seq <- seq(min(gsdat), max(gsdat), length=101)
-gssun_pred <- predict(sunmod, newdata=data.frame(Cond=gssun_seq), se.fit=TRUE)
-
-#ci and model fit
-sunupr <- gssun_pred$fit + (2*gssun_pred$se.fit)
-sunlwr <- gssun_pred$fit - (2*gssun_pred$se.fit)
-
-#SHADE leaves
[…]

### 2015-07-02 master_scripts/par.R

@@ -1,7 +1,7 @@
 # source("functions and packages/functions.R")
 # source("functions and packages/packages.R")
 # source("master_scripts/plot_objects.R")
-# 
+
  par <- read.csv("raw data/par.csv")
  
  treatments <- read.csv("raw data/temp_trt.csv")
@@ -40,11 +40,12 @@ par_agg <- summaryBy(par ~ leaf_type, data=parbar, FUN=mean)
 # anova(par_mod2)
 
 ###PLOTTING------------------------------------------------------------------------------------------------
-# windows(8,6)
+ #windows(6,6)
+
 bar(par, c(leaf_type, Month), parbar, col=c(shacol,suncol), xlab="", ylab=parlab, ylim=c(0, 2000), 
-      half.errbar=FALSE, mar=c(5,5,2,2))
-#   dev.copy2pdf(file="master_scripts/paper_figures/ppfd.pdf")
-#   dev.off()
+      half.errbar=FALSE, mar=c(4,4,1,1),las=1,cex.axis=.8, cex.lab=1, cex.names=1,mgp=c(2.5,1,0))
+#    dev.copy2pdf(file="master_scripts/paper_figures/ppfd.pdf")
+#    dev.off()
 
 
 ###for png

### 2015-07-02 stats/c13_nitro_stats.R

@@ -1,0 +1,37 @@
+source("functions and packages/functions.R")
+source("master_scripts/plot_objects.R")
+source("functions and packages/packages.R")
+require(lme4)
+require(lmerTest)
+library(LMERConvenienceFunctions)
+
+##Stats on c13 and leaf N-------------------------------
+
+#read data
+treatments <- read.csv("raw data/temp_trt.csv")
+
+leaf_chem <- read.csv("raw data/leaf_chem.csv")
+leaf_mass <- read.csv("raw data/leaf_data.csv")
+
+#get vcmax per chamber 
+aciparam <- read.csv("calculated_data/aciparameters.csv")
+
+#calculate nitrogen content in leaves
+canopy_chem <- merge(leaf_mass[leaf_mass$wp_type=="mid",c(1, 3:4, 7:8)], leaf_chem[, 4:8], by=c("Month", "chamber", "leaf"))
+#add treatments  
+canopy_chem <- addtrt_func(canopy_chem)
+canopy_chem$leafN <- with(canopy_chem, leaf_mass *(n_perc/100))
+canopy_chem$lma <- with(canopy_chem, (leaf_mass/leaf_area)*10000) #g/m2
+canopy_chem$leafN_area <- with(canopy_chem, lma *(n_perc/100))
+canopy_chem<- add_campaign(canopy_chem)
+
+chem_mod <- lm(c13 ~ leafN_area, data=canopy_chem)
+summary(chem_mod)
+
+chem_mod2 <- lmer(c13 ~ leafN_area + (1|chamber),data=canopy_chem)
+anova(chem_mod2)
+summary(chem_mod2)
+mcp.fnc(chem_mod2)
+source("functions and packages/r2glmm.R")
+rsquared.glmm(chem_mod2)
+

### 2015-07-06 markdown/WTC3_figures.Rmd
added tobacco dataframes to testing script

@@
 * Sun leaves were measured in the upper canopy at their current PPFD.
-Shade leaves were measured in the lower/middle canopy, first at their current PPFD and then again at sun leaf PPFD as a simulated sunfleck.
+Shade leaves were measured in the lower/middle canopy, first at their current PPFD and then again at sun leaf PPFD as a simulated sunfleck (Shade-Low light & Shade-High light).
 * New fully expanded leaves were measured. Shade leaves were sampled into the middle canopy as the experiment progressed to avoid issues with leaf age.

### 2015-07-06 master_scripts/gmes_WTC_tobacco.R

@@ -251,16 +251,14 @@ write.csv(gm_wtc_pair, "calculated_data/gmes_wtc_tobacco_pair.csv", row.names=FA
 
 
 
-####useful code to save
-# xsi_dfr2 <- setNames(xsi_dfr, names2)
-# list2env(lapply(xsi_dfr2, as.data.frame), .GlobalEnv)
+###for analysis first subset well watered and drought treatments--------------------------------------------------------
+# gm_drought <- gm_WTC2[gm_WTC2$drydown == "drought",]
+# write.csv(gm_drought, "calculated_data/gmes_drought.csv", row.names=FALSE)
+
+gm_tob_water <- gm_WTC2[gm_WTC2$drydown == "control",]
+write.csv(gm_tob_water, "calculated_data/gmes_tob_wellwatered.csv", row.names=FALSE)
 
 
-# zzz <- list()
-# for(i in 1:length(xsi_feb_h4)){
-#   zzz[[i]] <- gmesdata_func(xsi_feb_h4[[i]], licor_gmes, licor_times, licorrows=5,whichlicor="H4")
-#   message(i)
-# }
 
 
 

### 2015-07-12 manuscript/manuscript.Rmd
added manuscript folder with rmd from other project
started sigletters scripts and folder

added first para of results (resouces)

removed old scripts

---
title: Super catchy title of the manuscript goes here
author: ''
date: ''
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: global-change-biology.csl
bibliography: references.bib
---

\
Remko A. Duursma^1^, Richard Feynman^2^.

\
^1^ Hawkesbury Institute for the Environment, University of Western Sydney, Locked Bag 1797, Penrith, NSW, Australia

\
^2^ INRA, UMR 1391 ISPA, 33140 Villenave d'Ornon, France

\
*Corresponding author*: 
Remko Duursma
E: remkoduursma@gmail.com
T: +61(0)422096908

```{r echo=FALSE, message=FALSE, warnings=FALSE}
# Two useful packages for printing numbers, accessing content of objects.
library(broom)
library(reporttools)

# Do calculations here, create objects that we can use throughout the paper.

# Or, place them all in a separate script:
# source("manuscript_calcs.R")

# Example data
x <- rnorm(100)
mydata <- data.frame(x=x, y=x+rnorm(100))

# Linear regression
lmobj <- lm(y ~ x, data=mydata)

# Shorthand function to print numbers with two decimal points (and retain trailing zeroes)
f <- function(x, digits=2)sprintf(x, fmt=paste0("%.",digits,"f"))

# Nice function to format P values.
pval <- function(...)formatPval(..., includeEquality=TRUE)

# Set Flag to include figures in text or not.
includeFigs <- TRUE

# This document can be compiled with the 'Knit Word' button in Rstudio, or:
# library(rmarkdown)
# render("manuscript.Rmd", "word_document", "manuscript.docx")
```

# Abstract {.unnumbered}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris vehicula, quam nec tempus consequat, magna enim sodales ex, vitae condimentum urna enim nec libero. Vivamus sit amet lacus quam. Donec imperdiet sodales enim nec convallis. Praesent enim dolor, sagittis quis ex et, dapibus fringilla neque. Ut sed ultricies tortor. Duis ullamcorper dui nibh, ac ultricies arcu consectetur eget. Etiam blandit odio eu nulla sollicitudin, finibus elementum ligula luctus. Nulla lacus justo, bibendum eu justo elementum, varius maximus elit. Proin aliquam velit aliquam tellus ullamcorper, eget elementum magna imperdiet. Vestibulum id nulla erat. Quisque ante nulla, vestibulum ut magna sit amet, lacinia tempus leo. Proin sit amet lorem est.

# Introduction

**Notice how every major section starts on a new page, this is set in the manuscriptstyle.docx template file.** Some things were said [@Xiao2011], and also by Zhu et al. [-@Zhu2012]. Other remarks were made like that [see for example, @Zhang2005]. In conclusion, many things were said [@Zhu2012; @Xiao2011]. We can do subscripts, for example CO~2~, as well as superscripts (e.g., R^2^), and *italics*.

\
Proin fringilla magna in lobortis finibus. Sed in malesuada lectus. Nulla et diam tincidunt, scelerisque velit in, luctus nisl. Fusce fringilla tellus sit amet ante ullamcorper mattis. Curabitur metus sapien, efficitur elementum finibus sit amet, volutpat eget neque. Morbi id vulputate augue. In vel arcu et augue ullamcorper gravida. Nunc faucibus blandit felis, vitae feugiat mauris posuere sed. Nam quis quam non risus auctor placerat vitae non ex. Fusce at nunc diam. Nam dui libero, tristique in vulputate at, consequat et ante. Aenean consequat metus vitae velit dapibus pretium. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Mauris varius pellentesque ex vel ultricies. Fusce suscipit dapibus lacinia.

\
Aliquam ut ligula at nulla mattis rutrum. Duis dictum tempus diam ac tincidunt. Vivamus bibendum, augue non bibendum porta, metus orci lacinia justo, vitae consectetur ipsum nisl at libero. In convallis sem id dui blandit, id placerat libero molestie. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Sed malesuada aliquam risus, pharetra pharetra lorem tincidunt vitae. Ut lobortis lorem eu tortor sollicitudin, et consectetur lorem tincidunt. Morbi ut ligula vitae felis convallis cursus. Fusce placerat nisl et mi imperdiet viverra. In semper nibh sollicitudin tortor commodo pretium. Vestibulum ultricies dignissim sollicitudin. Vivamus tristique pellentesque ante vitae tempor. Morbi facilisis urna vitae sapien blandit feugiat. Integer dapibus ut ex id dignissim.

# Materials and Methods

## Subheader text
Curabitur suscipit, orci ac luctus pellentesque, lacus quam lobortis leo, sit amet rhoncus arcu dolor in felis. Morbi suscipit ullamcorper velit, vitae facilisis ligula tincidunt eu. Mauris molestie metus metus, sit amet eleifend mi volutpat sit amet. Nam dignissim ipsum eu tellus malesuada dignissim. Integer eleifend id metus vitae eleifend. Aliquam ultricies mi a eros vulputate, sit amet venenatis ante suscipit. Morbi vehicula sed nisi vitae semper. In et libero ac magna tincidunt mattis. Duis et enim eu dui scelerisque ullamcorper. Praesent pulvinar erat vitae bibendum iaculis. Maecenas dignissim ante nec dui vehicula dictum. Donec ac sodales arcu. Morbi nec lacus luctus, ultrices lacus vitae, rutrum justo.

\
Curabitur suscipit, orci ac luctus pellentesque, lacus quam lobortis leo, sit amet rhoncus arcu dolor in felis. Morbi suscipit ullamcorper velit, vitae facilisis ligula tincidunt eu. Mauris molestie metus metus, sit amet eleifend mi volutpat sit amet. Nam dignissim ipsum eu tellus malesuada dignissim. Integer eleifend id metus vitae eleifend. Aliquam ultricies mi a eros vulputate, sit amet venenatis ante suscipit. Morbi vehicula sed nisi vitae semper. In et libero ac magna tincidunt mattis. Duis et enim eu dui scelerisque ullamcorper. Praesent pulvinar erat vitae bibendum iaculis. Maecenas dignissim ante nec dui vehicula dictum. Donec ac sodales arcu. Morbi nec lacus luctus, ultrices lacus vitae, rutrum justo.

## More methods subheader
Aliquam ut ligula at nulla mattis rutrum. Duis dictum tempus diam ac tincidunt. Vivamus bibendum, augue non bibendum porta, metus orci lacinia justo, vitae consectetur ipsum nisl at libero. In convallis sem id dui blandit, id placerat libero molestie. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Sed malesuada aliquam risus, pharetra pharetra lorem tincidunt vitae. Ut lobortis lorem eu tortor sollicitudin, et consectetur lorem tincidunt. Morbi ut ligula vitae felis convallis cursus. Fusce placerat nisl et mi imperdiet viverra. In semper nibh sollicitudin tortor commodo pretium. Vestibulum ultricies dignissim sollicitudin. Vivamus tristique pellentesque ante vitae tempor. Morbi facilisis urna vitae sapien blandit feugiat. Integer dapibus ut ex id dignissim,

\
(1) 
$$\tau_d = \bar{Q}_{below} / Q_{above}$$

that was an equation in Latex.

# Results

Our data contains `r nrow(mydata)` rows and X has a standard deviation of `r sd(mydata$x)`. The regression of y on x gave an R^2^ of `r f(summary(lmobj)$r.squared)`. The relationship was significant (P `r pval(glance(lmobj)$p.value)`).

\
Curabitur semper nisl eu turpis tempus molestie. Suspendisse potenti. Proin laoreet nibh ac nisi iaculis, at efficitur augue tincidunt. Cras mi lectus, maximus a venenatis eget, pellentesque vel erat. Cras tincidunt, quam ultricies feugiat condimentum, velit ligula tincidunt velit, id venenatis lacus risus id sapien. Ut pulvinar quam ut faucibus ultricies. Ut quis enim et eros hendrerit dictum. Nunc tempus magna vel ante convallis pretium. Praesent sit amet risus vestibulum justo ultrices ultricies vel eu metus. In aliquam ultrices diam, ut tempus mi ultrices ut. Sed id lacus ac sem ullamcorper maximus ut eu leo.

\
Aliquam ut ligula at nulla mattis rutrum. Duis dictum tempus diam ac tincidunt. Vivamus bibendum, augue non bibendum porta, metus orci lacinia justo, vitae consectetur ipsum nisl at libero. In convallis sem id dui blandit, id placerat libero molestie. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Sed malesuada aliquam risus, pharetra pharetra lorem tincidunt vitae. Ut lobortis lorem eu tortor sollicitudin, et consectetur lorem tincidunt. Morbi ut ligula vitae felis convallis cursus. Fusce placerat nisl et mi imperdiet viverra. In semper nibh sollicitudin tortor commodo pretium. Vestibulum ultricies dignissim sollicitudin. Vivamus tristique pellentesque ante vitae tempor. Morbi facilisis urna vitae sapien blandit feugiat. Integer dapibus ut ex id dignissim.

# Discussion

Curabitur semper nisl eu turpis tempus molestie. Suspendisse potenti. Proin laoreet nibh ac nisi iaculis, at efficitur augue tincidunt. Cras mi lectus, maximus a venenatis eget, pellentesque vel erat. Cras tincidunt, quam ultricies feugiat condimentum, velit ligula tincidunt velit, id venenatis lacus risus id sapien. Ut pulvinar quam ut faucibus ultricies. Ut quis enim et eros hendrerit dictum. Nunc tempus magna vel ante convallis pretium. Praesent sit amet risus vestibulum justo ultrices ultricies vel eu metus. In aliquam ultrices diam, ut tempus mi ultrices ut. Sed id lacus ac sem ullamcorper maximus ut eu leo.

\
Aliquam ut ligula at nulla mattis rutrum. Duis dictum tempus diam ac tincidunt. Vivamus bibendum, augue non bibendum porta, metus orci lacinia justo, vitae consectetur ipsum nisl at libero. In convallis sem id dui blandit, id placerat libero molestie. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Sed malesuada aliquam risus, pharetra pharetra lorem tincidunt vitae. Ut lobortis lorem eu tortor sollicitudin, et consectetur lorem tincidunt. Morbi ut ligula vitae felis convallis cursus. Fusce placerat nisl et mi imperdiet viverra. In semper nibh sollicitudin tortor commodo pretium. Vestibulum ultricies dignissim sollicitudin. Vivamus tristique pellentesque ante vitae tempor. Morbi facilisis urna vitae sapien blandit feugiat. Integer dapibus ut ex id dignissim.

# Acknowledgements {.unnumbered}
We thank lots of people.

# Figures {.unnumbered}

```{r fig.width=8, fig.height=6, echo=FALSE, message=FALSE, eval=includeFigs}
# Ideally have your figures in functions (figure1() etc.). Just an example here.
with(mydata, plot(x,y, pch=19))
```

**Figure 1**. Pretty awesome figure caption.

# References {.unnumbered}

### 2015-07-12 manuscript/manuscript_wtc3.Rmd
added manuscript folder with rmd from other project
started sigletters scripts and folder

added first para of results (resouces)

removed old scripts

---
title: "Are shade leaves lying in wait? How instantaneous responses of mesophyll conductance to light availability in shade leaves affect theories of optimal canopy carbon gain"
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: new-phytologist.csl
bibliography: references.bib
---

\
Courtney E. Campany^1^, Belinda Medlyn^1^, Mark Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^.

\
^1^ Hawkesbury Institute for the Environment, University of Western Sydney, Locked Bag 1797, Penrith, NSW, Australia

\
^2^ Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, Australia

\
*Corresponding author*: 
Courtney Campany
E: courtneycampany@gmail.com

```{r, results='asis', echo=FALSE}
library(knitr)
opts_knit$set(root.dir = '../')
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Two useful packages for printing numbers, accessing content of objects.
library(broom)
library(reporttools)

# Shorthand function to print numbers with two decimal points (and retain trailing zeroes)
f <- function(x, digits=2)sprintf(x, fmt=paste0("%.",digits,"f"))

# Nice function to format P values.
pval <- function(...)formatPval(..., includeEquality=TRUE)

# Set Flag to include figures in text or not.
includeFigs <- TRUE

# This document can be compiled with the 'Knit Word' button in Rstudio, or:
# library(rmarkdown)
# render("manuscript.Rmd", "word_document", "manuscript.docx")
```

```{r global data/sourcing , include=FALSE}
source("functions and packages/functions.R")
source("functions and packages/packages.R")
source("master_scripts/plot_objects.R")
library(pander)

#treatments <- read.csv("raw data/chamber_trt.csv")
```

# Summary {.unnumbered}

# Introduction
The need to understand and predict carbon uptake in forest ecosystems is crucially important for assessing the impacts of a changing climate. Specifically, this involves accurate upscaling of leaf level measurements to predict whole canopy function. Due to the costs and limitations of efficient light harvesting plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling problematic. Incident PPFD declines exponentially with increasing leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor], and photosynthesis (A) increases with irradiance from shade to sun until biochemical limitation occurs [@evans1995carbon]. Consequently, leaves modify structural and functional properties to efficiently intercept variable light and enhance photosynthetic capacity [@vogelman1996leaves; @valladares2004photosynthetic]. To estimate whole canopy carbon gain it is thus necessary to account for the non-linear response of photosynthesis to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].

\
The distribution of resources required for A, including nitrogen (N) and water, are also partially defined by canopy light gradients. As A has a saturating response with light and a dependence on N, leaf N should be distributed in the upper canopy to maximize canopy carbon gain [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co]. This can lead to changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity across leaf types [@evans2001photosynthetic]. Sun leaves also experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization]. Higher rates of A and stomatal conductance (g~s~) can only be sustained if the hydraulic conductance (K) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal]. Optimal photosynthetic N investment in the upper canopy will be wasted if photosynthetic capacity exceeds hydraulic supply [@niinemets2012optimization], thus K should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].

\
These theoretical distributions of N and water regulate leaf physiological traits that constrain A differently for sun and shade leaves. The balance between the costs and energetic benefits of these leaf traits affecting photosynthetic capacity create trade-offs impacting canopy carbon gain [@givnish1988adaptation]. During A, CO~2~ diffuses from the atmosphere through stomata (g~s~), intercellular air space and into the chloroplast for fixation by Rubisco (mesophyll conductance, g~m~) [@tazoe2011using] and trade-offs exists that constrain each of these diffusion pathways. For optimal leaf carbon gain these pathways should act to minimize the resistance to CO~2~ diffusion while also limiting the intrinsic energy, water and construction costs which diminish the carbon return from A. As these processes should not be considered independent of each other and must be integrated with hydraulic constraints [@griffiths2013you] predicting their behaviour within canopies remains difficult.

\
First, g~s~ should be efficiently distributed within a canopy to utilize supplies of light, N and water to maximize A [@peltoniemi2012co]. •	This is because stomata are hypothesized to exhibit an optimal behaviour to maximize carbon gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal]. Second, variation in photosynthetic capacity between sun and shade leaves has also been hypothesized to be due to differences in g~m~ [@duursma2006vertical]. Mesophyll conductance can impose limitations on A as large as those associated with g~s~ [@warren2008stand; @ubierna2011estimation]and these limitations can reduce the efficiency of N use in A [@niinemets2007photosynthesis]. Interpreting the coupled responses of both g~s~ and g~m~ to A across sun and shade leaves has major implications for predicting canopy carbon gain, but empirical measurements across tree canopies are still lacking.

\
Additionally, assessing shade leaf behaviour is made difficult with accounting of short term light fluctuations within a canopy, via sunflecks. How shade leaves utilize sunflecks for short term carbon gain depends on the combined response time of gs and gm and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks]. For example, at short timescales the utilization of sunflecks is limited by the slow response of stomatal opening, effectively limiting the maximum assimilation rate that can be achieved [@way2012sunflecks]. Mesophyll conductance has been shown to respond to environmental factors across both long and rapid time scales, possibly faster than gs [@flexas2008mesophyll]. However, anatomical parameters which regulate gm with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are not likely adaptable during short light fluctuations. The physiological behaviour of shade leaves to maximize carbon gain must be assessed as both a degree of acclimation to local irradiance and as a potential response to transitory light availability.

\
Climate warming may also potentially affect the physiological behaviour of leaves within a canopy. This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) scale positively with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis]. How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~ will have implications for upscaling A for sun and shade leaves. Additionally, light saturated rates of A are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature]. How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients. The impacts of warming on plant physiological processes are obviously vast, yet differentiating their impacts on leaf physiology within a canopy will be essential for evaluating whole tree responses to a changing climate.

\
In this study we use *Eucalyptus tereticornis* Sm. trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (+3&deg;C) treatments, to empirically evaluate the distribution of resources and leaf physiological behaviour of sun and shade leaves.
1.	If whole tree canopies are optimized for carbon gain, then leaf N, K and photosynthetic capacity were predicted to be higher in full sun leaves.
2.	Under comparable leaf VPD, g~s~ and A should exhibit theoretical optimal behaviour across both sun and shade leaves, while g~m~ was predicted to scale positively with photosynthetic capacity.
3.	As shade leaves are constrained by their underlying biochemistry and slow physiological responses, increases in A following sunfleck simulations were not expected to reach levels of full sun leaves.
4.	The effects of climate warming was predicted to be greater in sun than shade leaves, seen as increases in stomatal closure and reductions in leaf carbon gain during summer months.

# Materials and Methods

## Whole tree chamber experimental design
Twelve *Eucalyptus tereticornis* seedlings, chosen from a single local Cumberland plain cohort, were planted in March 2013 into 12 whole-tree chambers (WTC) at the Hawkesbury Forest Experiment site near Richmond, NSW, Australia. Each chamber has a maximum height of 10 m and seedlings were grown for 15 months.  A detailed description of the WTC operation and design is available in [@barton2010whole]. Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced a constant elevated air temperature treatment (ET, 3&deg;C). The CO~2~ concentration inside the chamber was set to match outside air, tracking diurnal changes. Due to the temperature treatment air humidity was not explicitly controlled. Trees were watered with 70 l weekly from March 2013 to November 2013.  From December 2013 to final harvest trees were watered fortnightly with mean monthly (100 yr) rainfall amount.  In February 2013 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering. Due to a limited range of data for the drought treatment only well-watered trees are reported here.

\
Before seedlings were planted into each chamber they were maintained under well watered conditions in 35 l pots and kept inside each chamber. This allowed for seedlings to gain sufficient biomass before planting while also allowing them to acclimate to chamber temperature treatments. Seedlings were planted into each chamber after mean seedling height reached 100 cm.  After 2 months, floors were installed 45 cm from the soil surface.  This enabled chamber fluxes of CO~2~ and H~2~O from the whole tree canopy to be monitored. Gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.  At this point, trees had a mean height of 34&pm;15.1 and 418&pm;323.1 cm and an estimated leaf area of 3.9&pm;0.1 and 6.2&pm;0.2 m^2^ for ambient and elevated temperature treatments, respectively.

##Leaf gas exchange, online carbon isotope discrimination and mesophyll conductance
Leaf gas exchange measurements were performed monthly through to the final harvest.  Measurements were taken on a representative sun and shade leaf for each tree during each measurement campaign.  The youngest fully expanded leaf from the stem apex was chosen for gas exchange measurements and sun leaves were measured in the upper third of the canopy. In order to minimize any confounding effects of leaf age, shade leaves were measured first in the lower canopy then gradually higher during each campaign.  As shade leaves most likely developed slower this assured that older leaves that exist in the lower canopy were avoided. The nearest leaf on each branch was sampled for measurement of predawn leaf water potential. All leaves were selected and flagged 24 hours prior to initiation of measurements.

\
Prior to gas exchange measurements photosynthetic photon flux density (PPFD) was recorded at the individual leaf level and at the canopy position for each selected leaf.  A hand-held photosynthetically available radiation meter was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either full sun or full shade. A ceptometer (AccuPAR LP-80, Decagon Devices, Pullman, USA) was then used to measure 1 m integrated PPFD at the canopy height of each chosen leaf type. Five ceptometer readings were recorded within the canopy at random locations of the height of each selected leaf.  The mean of these readings was assumed to represent the overall leaf light environment of representative full sun and shade leaves for each tree. All measurements of PPFD and gas exchange were performed on full sun days between 10:00-14:30 h.

\
Leaf level gas exchange was measured with a standard 6 cm^2^ leaf chamber using a portable gas exchange system (LI-6400, LI-COR, Lincoln, NE, USA).  This system was coupled with a tunable diode laser (TDL;TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of online carbon isotope discrimination. The CO~2~ in the leaf chamber was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mols s^-1^.  Two identical gas exchanges systems were run simultaneously, one in each of a randomly chosen WTC for each temperature treatment. This paired design allowed for direct leaf comparisons to be made with simultaneous measurements of similar leaf types. Leaf temperatures were controlled at the current ambient or +3&deg;C chamber air temperature. PPFD in the chamber was set to match the individual light environment of each leaf type (explained above). Sun flecks were then simulated for shade leaves by increasing the chamber PPFD to the light environment of the full sun leaf in the same tree. When shade leaves were exposed to high light they were allowed enough time to equilibrate with chamber conditions until stable (ca. 25 min).

\
Once the chamber environment for leaves was stable the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet. These gases were dried by passing through napion gas dryers in the respective gas lines, and then ^12^CO~2~ and ^13^CO~2~ concentrations were measured for each gas by the TDL. Reference, sample and 2 calibration gases were run on alternating 80 s loops (20 s each), one for each paired leaf, for a total of 12 min. This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference and samples gases for calculations. The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDLAS on each measurement cycle. Net CO~2~ assimilation rate, g~s~, transpiration, and intercellular [CO~2~] were auto logged every 15 s for each gas exchange system over the same 12 min interval.

\
Using online C isotope discrimination measurements, the difference  between the observed discrimination and what is predicted for light saturated gas exchange is proportional to g~m~ [@griffiths2013mesophyll]. First, leaf discrimination was calculated by comparing the isotopic composition of the reference gas entering the chamber ($\delta$^13^C~e~) with the sample gas ($\delta$^13^C~o~) such that:

$$\delta = \left(\frac{R_s}{R_{stnd}} -1\right)1000$$

where R~s~ is the isotopic ratio of the sample and R~stnd~ is the isotopic ratio of the standard Vienna Pee Dee Belemnite. Next, the observed discrimination ($\Delta$~obs~) is calculated from [@evans1986carbon]:

$$\Delta_{obs} = \frac{\xi(\delta^13C_o-\delta^13C_e)}{1000+\delta^13C_e-\xi(\delta^13C_o-\delta^13C_e)}$$

where:

$$\xi = \frac{C_{e}}{C_{e}-C_{o}}$$

$\xi$ is the ratio of the CO~2~ entering the well mixed leaf chamber to the CO~2~ draw down by the leaf.

\
Second, leaf ^13^CO~2~ during C~3~ photosynthesis ($\delta$^13^C) is the resultant discrimination from CO~2~ diffusion from the atmosphere to the site of carboxylation, consisting or a series of fractionation steps described in [@evans1986carbon]. In this experiment, a modified form of this equation presented in [@tazoe2011using] with ternary effect corrections by [@farquhar2012ternary] was used such that:

$$
\begin{aligned}
\Delta^13 = &\frac{1}{1-t}a'+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}\\
            &\frac{1+t}{1-t}\left(b-a'-\frac{eR_d}{A+R_d}\right)\frac{A}{g_mC_a}-\\
            &\frac{1+t}{1-t}\left(\frac{eR_D}{(A+R_d)C_a}(C_i-\Gamma_*)\right)- \frac{1+t}{1-t}\left(f\frac{\Gamma_*}{C_a}\right)\\
\end{aligned}
$$

where:

$$t = \frac{(1+a')E}{2g^t_{ac}}$$

E denotes the transpiration rate and $\text{g}^t_{ac}$ is the total conductance to CO~2~ diffusion to both the boundary layer and stomatal conductance (von Caemmerer 1981). C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_*$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~). In this experiment both $\Gamma_*$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulis* from [@crous2012light]. The different fractionation factors include; diffusion through water (a~i~, 1.8&permil;), Rubisco carboxylation (b, 29&permil;), the photorespiratory fractionation (f, 16.2&permil;) and the combined fractionation through the boundary layer and the stomata (a'). a' is defined by:

$$a' = \frac{a_b(C_a-C_i)+a(C_s-C_i)}{C_a-C_i}$$

where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b (2.9&permil;) is the fractionation from boundary layer diffusion and a is the fractionation due to diffusion in air~(4.4&permil;) [@evans1986carbon]. The CO~2~ diffusion from the intercellular airspace to the chloroplast, g~m~, is given by its relationship to the CO~2~ assimilation rate (A) by:

$$g_m = \frac{A}{C_i-C_c}$$

where C~c~ is the chloroplast CO~2~ partial pressure. Examples of this approach to measure gas exchange and carbon isotope discrimination are presented in [@evans2013temperature].

\
Photosynthetic CO~2~ response (AC~i~) curves were developed at 25&deg;C for one sun and shade leaf for each chamber prior to the initiation of the drought treatment. Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25&deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^). From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical].

##Leaf chemistry and hydraulic parameters
Following gas exchange measurements each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, dried to constant mass and then weighed.  These leaves were then milled and analyzed for leaf N content and $\delta$^13^C. Leaf samples were analysed on a Delta V Advantage coupled to a Flash HT and Conflo IV (Thermo Fisher Scientific, Bremen, Germany) in dual-reactor setup. Samples were flash combusted at 1000&deg;C and converted to CO~2~ and N~2~ and then subjected to stable isotope ratio mass spectrometry. Isotopic signatures are reported relative to the VPDP scale.

\
Predawn ($\Psi_{pd}$) and midday ($\Psi_l$) leaf water potentials were measured for sun and shade leaves during each gas exchange campaign using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA). Predawn leaf water potential on selected leaves for each chamber was measured before sunrise on the same day as gas exchange measurements. Leaves used for gas exchange were immediately sampled for $\Psi_l$ once measurements were completed. All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed. Leaf water potential and transpiration (E) from gas exchange were then used to calculate leaf specific hydraulic conductance (K) through the equation:

$$K = \frac{E}{\Psi_{pd}~-\Psi_l-hp_wg}$$

where the term Hp~w~g represents the gravitational potential assumed to be minimal in this experiment due to small tree size [@whitehead1998regulation]. Leaf level instantaneous transpiration efficiency (ITE) was calculated as leaf photosynthesis divided by leaf transpiration. The g~1~ parameter, an index of water cost of carbon assimilation, was generated by fitting observed values of g~s~ into the optimal stomatal conductance model from [@medlyn2012reconciling].

##Data analysis
Differences in experimental parameters to either the warming treatment or leaf type were analysed by one-way analysis of variance in R [@RDevelopmentCoreTeam2011] with chambers as random effects.  Mixed model ANOVAs of A max and leaf chemistry were performed using the nlme package [@Pinheiro2015] in R  and r~2~ values of mixed models were computed as in [@nakagawa2013general]. Confidence intervals (95 %) of mixed effect linear models of leaf photosynthesis as a functions of different physiological parameters were generated using bootstrapping methods with 999 simulations in the lme4 package [@bates2013lme4] in R. For non-linear mixed models models confidence intervals were estimated by fitting a generalized additive model to the data with the mgcv package in R.  Results were considered significant at P&leq;0.05.

# Results

## Sun and Shade leaf resource distribution

Across all measurement campaigns 1m canopy integrated PPFD was significantly different between sun and shade leaves (P<0.001).  PPFD was reduced by 78% in the shade (Figure 1). The distribution of H~2~O, via K, was not different across sun and shade leaves (Table 1).  This was a result of neither $\Psi_{pd}$, $\Psi_l$ (Table 1) nor E (Table 2) differing between leaf types. Leaf nitrogen, on an area basis, was significantly higher in sun leaves than in shade leaves (p<0.001) by ca. 20% (Table1). No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K, E or leaf N either within or across leaf types.

## Leaf photosynthesis rates and photosynthetic capacity

Leaf photosynthesis across time, leaf types, lightson and temp treatment
Leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environoment, across the wide range of leaf temperatures measured.

classic aci (acc later after gmes)
vcmax and jmax
realtionship of vcmax and photosynthesis with N

## water use
ITE with VPD for all leaf types and temp treatments
leaf 13C with leaf N and what this stands for

## stomatal and mesophyll conductance

# Discussion

# Figures {.unnumbered}

```{r par, fig.width=6, fig.height=6, echo=FALSE}
source("master_scripts/par.R")
```

**Figure 1**. Local light environment for sun and shade leaves for each gas exchange campaign. Means with standard errors represent 1 m integrated PPFD (n=5), measured with a ceptometer, at the canopy height of each selected leaf.

```{r acicurves, fig.width=6, fig.height=8, echo=FALSE}
source("master_scripts/aci_vcmaxnitro.R")
```

**Figure2**. (a) Photosynthetic CO~2~ response (ACi) curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments. ACi curves were developed once for all trees, before the initiation of the drought treatment, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^). (b) The relationship between Vc~max~ and mean leaf nitrogen~area~ for each chamber, including sun leaves and shade leaves at low light. For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional r^2^ of 0.28 and 0.35, respectively.

```{r wateruse, fig.width=6, fig.height=8, echo=FALSE}
source("master_scripts/wateruse.R")
```

**Figure 3**. (a) Response of instantaneous water use efficiency (ITE) to leaf vapor pressure deficit (VPD) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments. (b) The relationship between leaf $\delta$^13^C and leaf nitrogen~area~ for sun leaves and shade leaves at low light. For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g1 value for each leaf type and treatment. For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional r^2^ of 0.41 and 0.45, respectively.

```{r Agmgs, fig.width=6, fig.height=8, echo=FALSE}
source("master_scripts/Agmgs.R")
```

**Figure 4**. The response of leaf photosynthesis rates to stomatal conductance (a) and mesophyll conductance (b) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.  The response of shade leaf physiology to high light was recorded once both photosynthesis and stomatal conductance were stable in the gas exchange leaf cuvette (ca. 25min). Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b). Grey areas are approximately 95% confidence intervals from the mean.

```{r barplots, fig.width=10, fig.height=5, echo=FALSE}
source("master_scripts/physiology_barplots.R")
```

**Figure 5**. The mean stomatal conductance (a), mesophyll conductance (b) and photosynthesis rate (c) of sun leaves and shade leaves at both low and high light with standard errors.

```{r cicc, fig.width=10, fig.height=5, echo=FALSE}
source("master_scripts/cicc_bar.R")
```

**Figure 6**. The mean intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light with standard errors.

# References {.unnumbered}

### 2015-07-12 gasexchange_testing.R

@@ -1,35 +1,0 @@
-source("functions and packages/functions.R")
-source("master_scripts/plot_objects.R")
-source("functions and packages/packages.R")
-
-#read in gm data set (no drought) and Cibar(discrimination)-------------------------------------------------------
-gmes <- read.csv("calculated_data/gmes_wellwatered.csv")
-
-###get average by id
-gm_agg <- summaryBy(Photo+ Cond +Ci+ gm ~ chamber+id+leaf +light+temp+leaflight+Month, data=gmes, FUN=mean, keep.names=TRUE)
-gm2 <- summaryBy(Photo+ Cond +Ci+ gm ~ leaflight, data=gm_agg, FUN=c(mean,se))
-
-
-### we can test gas exchange data here
-
-boxplot(Ci~leaflight, data=gm_agg)
-boxplot(Photo~leaflight, data=gm_agg)
-boxplot(gm~leaflight, data=gm_agg)
-boxplot(Cond~leaflight, data=gm_agg)
-
-stats_gm <- boxplot(gm~leaflight, data=gm_agg)
-
-
-
-###read in tobacco dataset
-gmes_tob <- read.csv("calculated_data/gmes_tob_wellwatered.csv")
-
-###get average by id
-gm_tob_agg <- summaryBy(Photo+ Cond +Ci+ gm ~ chamber+id+leaf +light+temp+leaflight+Month, data=gmes_tob, FUN=mean, keep.names=TRUE)
-
-gm_tob2 <- summaryBy(Photo+ Cond +Ci+ gm ~ leaflight, data=gm_tob_agg, FUN=c(mean,se))
-
-
-###aci
-#read in jmax vcmax
-jvc<- read.csv("calculated_data/tdl_aci.csv")

### 2015-07-12 gm_testingbymonth.R

@@ -1,66 +1,0 @@
-
-
-source("functions and packages/gmes_functions.R")
-source("functions and packages/functions.R")
-
-###required packakges
-library(data.table)
-library(plyr)
-library(doBy)
-
-##read in treatments
-treatments <- read.csv("raw data/temp_trt.csv")
-
-####read licor data and run licor formatting functions------------------------------------------------------------------
-licor_master <- read.csv("tdl_files/april/licor_april.csv")
-licor_master <- chlab_func(licor_master)
-
-###remove pair ids for now (?) bind them back after formatting (should beed id, chamber)
-pairs <- licor_master[, 2:7]
-licor2 <- licor_master[, -7]
-
-####format master licor file, will have two dataframes one with all data, and one with times for matching
-licor_gmes <- chooseidfunc(licor2, c("campaign" , "chamber",  "leaf",  "light"))
-
-licor_gmes <- licorformat_func(licor_gmes)
-
-###now run time range function if get samples id and time range
-licor_times <- timerange_func(licor_gmes)
-
-
-
-####april
-apr_names<- list.files(path="tdl_files/april/",pattern="tdl",full.names=TRUE)
-apr_names2 <- gsub("tdl_files/april/", "", apr_names)
-apr_names2 <- gsub(".csv", "", apr_names2)
-
-apr_files <- llply(list.files(path="tdl_files/april/",pattern="tdl",full.names=TRUE),function(filename) {
-  dat=read.csv(filename)})
-
-apr_formatted <- llply(apr_files, tdlformat_func)
-
-xsi_apr <- llply(apr_formatted, function(x)  xsicalc_func(x))
-xsi_apr_dfr <- llply(xsi_apr, function(x) data.frame(x))
-xsi_apr_dfr2 <- setNames(xsi_apr_dfr, apr_names2)
-
-##for now run gmes functions for each licor #
-xsi_apr_h1 <- xsi_apr_dfr2[c(5,8,11)] 
-xsi_apr_h3 <- xsi_apr_dfr2[c(1,2,4,6,7,9)] 
-xsi_apr_h4 <- xsi_apr_dfr2[c(3,10)] 
-
[…]

### 2015-07-12 master_scripts/lightson_sigletters.R

@@ -1,0 +1,17 @@
+source("functions and packages/functions.R")
+library(doBy)
+
+gasexchange  <- read.csv("calculated_data/gmes_wellwatered.csv")
+
+###leaf data
+ge_agg <- summaryBy(Photo+Cond+Ci+Trmmol+gm+VpdL ~ chamber+id+leaf +light+temp+leaflight+Month+drydown, 
+                    data=gasexchange, FUN=mean, keep.names=TRUE)
+ge_agg$tukeyid <- as.factor(paste(ge_agg$leaflight, ge_agg$temp, sep="-"))
+
+###linear models of gas exchange by treatments
+library(visreg)
+library(multcomp)
+library(nlme)
+
+#Photosynthesis
+photo_agg <- summaryBy(Photo+Cond+gm ~ leaf, data=ge_agg[ge_agg$leaflight == "shade-high",] , FUN=c(mean,se))

### 2015-07-12 scripts/leaf_K.R

@@ -129,5 +129,63 @@ shajan <- subset(leafcond, Month == "Jan" & leaf == "shade")
 #sun- jan = ch2 midday wp really high, ch8/5 transpiration low
 #sha - jan = ch 9/2 transpiration too high
 
+###stats on wp and leaf K
+leafK_stats <- droplevels(leafK_nodrought)
 
 
+library(visreg)
+library(multcomp)
+library(nlme)
+
+#examite data with boxplots, then remove any outliers
+boxplot(leafK~temp, data=leafK_stats[leafK_stats$leaf =="sun",])
+boxplot(leafK~temp, data=leafK_stats[leafK_stats$leaf =="shade",])
+
+
+##PAR significantly different between leaves not by temperature treatment
+pre_mod <- lme(pre_mp ~ leaf, random=~1|chamber, data=leafK_stats)
+summary(pre_mod)
+anova(pre_mod)
+
+pre_mod2 <- lme(pre_mp ~ leaf*temp, random=~1|chamber, data=leafK_stats)
+anova(pre_mod2)
+summary(pre_mod2)
+  ##predawn wp not different between at/et or sun/shade
+mid_mod <- lme(mid_mp ~ leaf, random=~1|chamber, data=leafK_stats)
+summary(mid_mod)
+anova(mid_mod)
+
+mid_mod2 <- lme(mid_mp ~ leaf*temp, random=~1|chamber, data=leafK_stats)
+anova(mid_mod2)
+summary(mid_mod2)
+ ##mid wp not different between at/et or sun/shade
+
+e_mod <- lme(Trmmol ~ leaf, random=~1|chamber, data=leafK_stats)
+summary(e_mod)
+anova(e_mod)
+
+e_mod2 <- lme(Trmmol ~ leaf*temp, random=~1|chamber, data=leafK_stats)
+anova(e_mod2)
+summary(e_mod2)
+
+k_mod <- lme(leafK ~ leaf, random=~1|chamber, data=leafK_stats)
+summary(k_mod)
+anova(k_mod)
+
+k_mod2 <- lme(leafK ~ leaf*temp, random=~1|chamber, data=leafK_stats)
+anova(k_mod2)
+summary(k_mod2)
+
+
+Edat <- read.csv("calculated_data/gmes_wellwatered.csv")
+
+###get average by id
[…]

### 2015-07-12 stats/gasexchange_sigletters.R

@@ -1,16 +1,12 @@
 source("functions and packages/functions.R")
 library(doBy)
 
-gasexchange  <- read.csv("calculated_data/gmes_wtc.csv")
+gasexchange  <- read.csv("calculated_data/gmes_wellwatered.csv")
 
 ###leaf data
 ge_agg <- summaryBy(Photo+Cond+Ci+Trmmol+gm+VpdL ~ chamber+id+leaf +light+temp+leaflight+Month+drydown, 
                     data=gasexchange, FUN=mean, keep.names=TRUE)
-
-###split by drought and then rearrange
-ge_wet <- ge_agg[ge_agg$drydown == "control",]
-ge_dry <- ge_agg[ge_agg$drydown == "drought",]
-
+ge_agg$tukeyid <- as.factor(paste(ge_agg$leaflight, ge_agg$temp, sep="-"))
 
 ###linear models of gas exchange by treatments
 library(visreg)
@@ -18,61 +14,76 @@ library(multcomp)
 library(nlme)
 
 #Photosynthesis
-photo_agg <- summaryBy(Photo+Cond+gm ~ leaf, data=ge_wet[ge_wet$leaflight != "shade-high",] , FUN=c(mean,se))
+ge_sunsha <- ge_agg[ge_agg$leaflight != "shade-high",]
+ge_sunsha <- droplevels(ge_sunsha)
 
-A_leaf <- lme(Photo~ leaf, random=~1|chamber, data=ge_wet, subset=leaflight != "shade-high")
+A_sun_temp <- lme(Photo ~ temp ,random=~1|chamber, data=ge_sunsha, subset=leaflight=="sun-high")
+summary(A_sun_temp)
+anova(A_sun_temp)
+visreg(A_sun_temp)
+
+A_sha_temp <- lme(Photo ~ temp ,random=~1|chamber, data=ge_sunsha, subset=leaflight=="shade-low")
+summary(A_sha_temp)
+anova(A_sha_temp)
+visreg(A_sha_temp)
+##first no difference between temp treatments with sun shad leaves, 
+##look at model without for sun vs shade
+
+A_leaf <- lme(Photo~ leaf, random=~1|chamber, data=ge_sunsha, subset=leaflight != "shade-high")
 summary(A_leaf)
 anova(A_leaf)
 visreg(A_leaf)
+##A higher in sun leaves (23%)
 
[…]

### 2015-07-13 manuscript/manuscript_wtc3.Rmd
added results, figures, and new cites

@@
 
-```{r, results='asis', echo=FALSE}
+```{r, results='asis', echo=FALSE, warning=FALSE}
 library(knitr)
@@
 
-where C~c~ is the chloroplast CO~2~ partial pressure. Examples of this approach to measure gas exchange and carbon isotope discrimination are presented in [@evans2013temperature].
+where C~c~ is the chloroplast CO~2~ partial pressure.
+Examples of this approach to measure gas exchange and carbon isotope discrimination are presented in [@evans2013temperature].
+Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular arispaces to the site of carboxylation were then estimated using eq.
+7.
 
@@
 Leaf level instantaneous transpiration efficiency (ITE) was calculated as leaf photosynthesis divided by leaf transpiration.
-The g~1~ parameter, an index of water cost of carbon assimilation, was generated by fitting observed values of g~s~ into the optimal stomatal conductance model from [@medlyn2012reconciling].
+The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2012reconciling] using non-linear resgression, where K=0.5 [see @duursma2013near].
 
@@
 ## Sun and Shade leaf resource distribution
-
-Across all measurement campaigns 1m canopy integrated PPFD was significantly different between sun and shade leaves (P<0.001).
-PPFD was reduced by 78% in the shade (Figure 1).
+Across all measurement campaigns 1m canopy integrated PPFD was significantly different between sun and shade leaves (P<0.001) and PPFD was reduced by 78% in the shade (Figure 1).
 The distribution of H~2~O, via K, was not different across sun and shade leaves (Table 1).
-This was a result of neither $\Psi_{pd}$, $\Psi_l$ (Table 1) nor E (Table 2) differing between leaf types.
-Leaf nitrogen, on an area basis, was significantly higher in sun leaves than in shade leaves (p<0.001) by ca.
+This was the result of neither $\Psi_{pd}$, $\Psi_l$ (Table 1) nor E (Table 2) differing between leaf types.
+Leaf N, on an area basis, was significantly higher in sun leaves than in shade leaves (p<0.001) by ca.
 20% (Table1).
-No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K, E or leaf N either within or across leaf types.
+No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K, E or N~area~ either within or across leaf types.
 
 ## Leaf photosynthesis rates and photosynthetic capacity
+Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environoment (P<0.001, Table 2).
+Following an increase in available light, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P<0.001).
+No effect of the warming treatment was detected on rates of A of sun leaves or shade leaves at low or high light.
+Photosynthesis across all leaf types and treatments did not vary according across the leaf temperatures measured in this study or time of year (Figure S1a).
 
-Leaf photosynthesis across time, leaf types, lightson and temp treatment
-Leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environoment, across the wide range of leaf temperatures measured.
+\
+The photsynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (p<0.001 & p=0.02, respectively), as evident from AC~i~ curves from each chamber (Figure 2a).
+No effect of the warming treatment was detected on either parameter (Table 1).
+Within individual chambers Vc~max~ was postively related to leaf N~area~ across leaf types and temperature treatments (P =0.01, Figure 2b).
+Additionally, leaf N~area~ was postively related to A across gas exchange campaigns and leaf types (P<0.001, Figure 2c).
+As Vc~max~ is condisdered implicit with g~m~ (results below) with conventional AC~i~ curves, a AC~c~ curves were also simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.
+When accounting for the effect of g~m~, photosynthetic capacity was still greater in sun leaves than shade leaves (Figure S2)
 
-classic aci (acc later after gmes)
-vcmax and jmax
-realtionship of vcmax and photosynthesis with N
+## Leaf water use efficiency
+Leaf ITE was significantly greater in sun leaves than in shade leaves (P=0.001, Table 2).
+Following an increase in available light, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P<0.001).
+ITE in sun leaves was reduced in the warming treatment (P=0.021) but no effect was noted in shade leaves with low or high light.
+For all leaf types and light treatments there was a strong response of ITE to VPD, and individual data points broadly corresponded to response curves from the optimal ITE model (Figure 3a).
+The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
+Across leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure S2a)
 
-## water use
-ITE with VPD for all leaf types and temp treatments
+Leaf $\delta$^13^C significantly decreased from sun leaves to shade leaves by ca.
+2&permil; (p<0.001, Table 1).
+Within leaf types no affects of the warming treatment on leaf $\delta$^13^C were detected.
+Leaf $\delta$^13^C and N~area~ were postively correlated for all leaves (P<0.001, Figure 3b).
+
 leaf 13C with leaf N and what this stands for
 
-## stomatal and mesophyll conductance
+## Stomatal conductance, mesophyll conductance and CO~2~ drawdown
+Mean g~s~ was significantly higher in shade compared to sun leaves (+18%), under their local light environment (P=0.004, Table2).
+Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P<0.001, Figure 4a).
+In sun leaves, g~s~ and A showed optimal behavior with diminishing rates of A with highest values of g~s~ (Figure 5a).
+However, in shade leaves this relationship was clearly uncoupled.
+With increased light availability, both increases in A and g~s~ in shade leaves extened the realtionship of A and g~s~ seen in sun leaves.
+
+\
+Mean g~m~ was significantly higher in sun compared to shade leaves (+27%), under their local light environment (P=<0.001, Table2).
+Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P<0.001, Figure 4b).
+Relationships between g~m~ and A were nearly proportional within leaf types and light treatments (Figure 4b,c).
+Photosynthesis scaled postively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves in high light resulted in the highest rates of A (Figure 5b).
+No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types or light treatments.
+Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figures S1b and S2b, respectively) and only a slight negative relationship was detected with g~m~ and increasing leaf temperature.
+
+\
+The combination of a non-limiting supply of leaf K and more open stomata lead to a significant increase in C~i~ in shade compared to sun leaves (Figure 6a).
+As increases in A in shade leaves were associated with higher rates of g~s~ the C~i~ was similar than under low light but still greater than sun leaves.
+As the drop in CO~2~ for intecellular spaces to the cholorplast, C~i~-C~c~, measures the coordiation between gm and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure 6c).
+This was a direct result of the increase in g~m~ and its subsequent direct effect on A, when compared to both sun and shade leaves that are acclimated to their local light environment.
+CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable acorss the range of temperatures measured and gas exchange campaigns (Figure S2c and S1c, respectively).
+
+## delta contrib, deltav cica
+
+The difference between $\delta$^13^Ce and $\delta$^13^Co was mostly accounted for by discrimination associated with mesophyll conductance (85%) compared to contributions from respiration and photorespiration.
 
 # Discussion
+
+photosynthetic capacity: Mesophyll conductance asymmetrically affects key photosynthetic parameters and an assumption of infinite mesophyll conductance leads to underestimation of the maximum carboxylation rate Vcmax, maximum electron transport rate Jmax [@sun2014asymmetrical].
 
@@
 
-**Figure2**.
+**Figure 2**.
 (a) Photosynthetic CO~2~ response (ACi) curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.
@@
 (b) The relationship between Vc~max~ and mean leaf nitrogen~area~ for each chamber, including sun leaves and shade leaves at low light.
-For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional r^2^ of 0.28 and 0.35, respectively.
+(c) the relationship between For (b,c) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional r^2^ of 0.28 and 0.35 for (b), and 0.24 and 0.33 for (c).
 
-```{r wateruse, fig.width=6, fig.height=8, echo=FALSE}
+```{r wateruse, fig.width=6, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE}
 source("master_scripts/wateruse.R")
@@
 
+```{r barplots, fig.width=10, fig.height=5, echo=FALSE}
+source("master_scripts/physiology_barplots.R")
+```
+
+**Figure 4**. The mean stomatal conductance (a), mesophyll conductance (b) and photosynthesis rate (c) of sun leaves and shade leaves at both low and high light with standard errors.
+
 ```{r Agmgs, fig.width=6, fig.height=8, echo=FALSE}
@@
 ```  
-
-**Figure 4**.
+**Figure 5**.
 The response of leaf photosynthesis rates to stomatal conductance (a) and mesophyll conductance (b) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
@@
 Grey areas are approximately 95% confidence intervals from the mean.
-
-```{r barplots, fig.width=10, fig.height=5, echo=FALSE}
-source("master_scripts/physiology_barplots.R")
-```
-
-**Figure 5**. The mean stomatal conductance (a), mesophyll conductance (b) and photosynthesis rate (c) of sun leaves and shade leaves at both low and high light with standard errors.
 
@@
 
+# Supporting Information {.unnumbered}
+```{r gasex_temp1, fig.width=6, fig.height=10, echo=FALSE}
+source("master_scripts/gasex_temp.R")
+```  
+**Figure S1**.
+Reponse of photosynthesis (a), mesophyll conductance (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
+Shaded symbols represents each monthyl measurement campaign.
+
+```{r gasex_temp2, fig.width=6, fig.height=10, echo=FALSE}
+source("master_scripts/gasex_temp2.R")
+```  
+**Figure S2**.
+Reponse of vapor pressure deficit (a), stomatal conductance (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
+Shaded symbols represents each monthyl measurement campaign.
+
+```{r acc, fig.width=6, fig.height=6, echo=FALSE}
+source("master_scripts/ACc_figure.R")
+```
+
+**Figure S3**.
+Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
+C~c~ values were predicited with mesophyll conductance, thus curves represent chloroplastic photosynthetic parameters at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+
+```{r deltacica, fig.width=6, fig.height=6, echo=FALSE}
+source("master_scripts/delta_cica.R")
+```
+
+**Figure S4**.
+Relationship between the observed discrimination of ^13^CO~2~ and measured C~i~/C~a~ for sun leaves and shade leaves at both low and high light.
+The solid line represents the the theortical line for C3 plants from [@evans1986carbon].
+
 # References {.unnumbered}

### 2015-07-14 manuscript/manuscript_wtc3.Rmd
added data table for resouces
started 2 data tables in manuscrupts

@@
 ---
-title: "Are shade leaves lying in wait? How instantaneous responses of mesophyll conductance to light availability in shade leaves affect theories of optimal canopy carbon gain"
+title: "Are shade leaves lying in wait? How instantaneous responses of mesophyll conductance to light availability in shade leaves impacts canopy carbon gain"
 author: ""
@@
 # Summary {.unnumbered}
+
+* overview
+* method
+* big result
+* implication
 
@@
 
+# Tables {.unnumbered}
+
+```{r tablex, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
+library(ascii)
+resource_table <- read.csv("master_scripts/resource_table.csv")
+print(ascii(resource_table,  include.rownames=FALSE,
+          colnames =c("Leaf", "Temperature", "LMA", "Narea", "Vcmax", "Jmax", 
+                       "Leaf K", "WP_pre", "WP_mid", "13C")),type="pandoc")
+
+```
+**Table 1.** Canopy resouce distribution among sun and shade leaves under elevated and ambient termperatures
+
+```{r table1, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
+library(ascii)
+phystable <- read.csv("master_scripts/ge_table.csv")
+print(ascii(phystable,
+            include.rownames=FALSE,
+            colnames=c("Leaf", "Temperature", "Light", "Photosynthesis", "gs", "gm", 
+                       "Ci", "Transpiration", "VPD")),
+            type='pandoc')
+
+```
+**Table 2.** Mean gas exchange parameters across leaf types with light and temperature treatments
+
 # Figures {.unnumbered}

### 2015-07-23 manuscript/manuscript_wtc3.Rmd
added equations in methods
air vars script added with met data download script

@@
 Second, leaf ^13^CO~2~ during C~3~ photosynthesis ($\delta$^13^C) is the resultant discrimination from CO~2~ diffusion from the atmosphere to the site of carboxylation, consisting or a series of fractionation steps described in [@evans1986carbon].
-In this experiment, a modified form of this equation presented in [@tazoe2011using] with ternary effect corrections by [@farquhar2012ternary] was used such that:
+In this experiment, a modified form of this equation presented in [@evans2013temperature] with ternary effect corrections by [@farquhar2012ternary] was used such that:
+
+from Evans & Von Caemmerer
+$\Delta$~o~ = $\Delta$~i~ - $\Delta$~gm~ - $\Delta$~e~ - $\Delta$~f~
+where $\Delta$~o~ is the observed discrimation and  $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contritbutions of fractionataion if C~i~=C~c~, g~m~, respiration and photorespiration, respectively.
+The equations for each are as follows:
 
@@
 \begin{aligned}
-\Delta^13 = &\frac{1}{1-t}a'+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}\\
+\Delta = &\frac{1}{1-t}a'+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}\\
             &\frac{1+t}{1-t}\left(b-a'-\frac{eR_d}{A+R_d}\right)\frac{A}{g_mC_a}-\\
             &\frac{1+t}{1-t}\left(\frac{eR_D}{(A+R_d)C_a}(C_i-\Gamma_*)\right)- \frac{1+t}{1-t}\left(f\frac{\Gamma_*}{C_a}\right)\\
+\end{aligned}
+$$
+
+$$
+\begin{aligned}
+\Delta_i = &\frac{1}{1-t}a'+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}\\
+\end{aligned}
+$$
+
+$$
+\begin{aligned}
+\Delta_gm = &\frac{1+t}{1-t}\left(b-a'-\frac{eR_d}{A+R_d}\right)\frac{A}{g_mC_a}-\\
+
+\end{aligned}
+$$
+
+$$
+\begin{aligned}
+\Delta_e = &\frac{1+t}{1-t}\left(\frac{eR_D}{(A+R_d)C_a}(C_i-\Gamma_*)\right)
+\end{aligned}
+$$
+
+$$
+\begin{aligned}
+\Delta = &\frac{1}{1-t}a'+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}\\
+            &\frac{1+t}{1-t}\left(b-a'-\frac{eR_d}{A+R_d}\right)\frac{A}{g_mC_a}-\\
+            &\frac{1+t}{1-t}\left(\frac{eR_D}{(A+R_d)C_a}(C_i-\Gamma_*)\right)- \frac{1+t}{1-t}\left(f\frac{\Gamma_*}{C_a}\right)\\
+\end{aligned}
+$$
+
+$$
+\begin{aligned}
+\Delta_f = & \frac{1+t}{1-t}\left(f\frac{\Gamma_*}{C_a}\right)\\
 \end{aligned}
@@
 With increased light availability, both increases in A and g~s~ in shade leaves extened the realtionship of A and g~s~ seen in sun leaves.
+
+discrimination contributions and \delta vs ci/ca go here
 

### 2015-07-27 manuscript/manuscript.Rmd
added sigletters to both data tables
made equation edits in methods

minor edits in methods and results of manuscript

@@
     number_sections: yes
-csl: global-change-biology.csl
+csl: new-phytologist.csl
 bibliography: references.bib

### 2015-07-27 manuscript/manuscript_wtc3.Rmd
added sigletters to both data tables
made equation edits in methods

minor edits in methods and results of manuscript

@@
 ---
-title: "Are shade leaves lying in wait? How instantaneous responses of mesophyll conductance to light availability in shade leaves impacts canopy carbon gain"
+title: "Near instantaneous response of mesophyll conductance to light availability allows shade leaves to lie in wait"
 author: ""
@@
 # Introduction
-The need to understand and predict carbon uptake in forest ecosystems is crucially important for assessing the impacts of a changing climate.
+The need to understand and predict carbon (C) uptake in forest ecosystems is crucially important for assessing the impacts of a changing climate.
 Specifically, this involves accurate upscaling of leaf level measurements to predict whole canopy function.
 Due to the costs and limitations of efficient light harvesting plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling problematic.
-Incident PPFD declines exponentially with increasing leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor], and photosynthesis (A) increases with irradiance from shade to sun until biochemical limitation occurs [@evans1995carbon].
+Incident PPFD declines exponentially with increasing leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor], and photosynthesis increases with irradiance from shade to sun until biochemical limitation occurs [@evans1995carbon].
 Consequently, leaves modify structural and functional properties to efficiently intercept variable light and enhance photosynthetic capacity [@vogelman1996leaves; @valladares2004photosynthetic].
-To estimate whole canopy carbon gain it is thus necessary to account for the non-linear response of photosynthesis to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
+To estimate whole canopy C gain it is thus necessary to account for the non-linear response of photosynthesis to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
 
@@
 The distribution of resources required for A, including nitrogen (N) and water, are also partially defined by canopy light gradients.
-As A has a saturating response with light and a dependence on N, leaf N should be distributed in the upper canopy to maximize canopy carbon gain [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
+As A has a saturating response with light and a dependence on N, leaf N should be distributed in the upper canopy to maximize canopy C gain [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
 This can lead to changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity across leaf types [@evans2001photosynthetic].
@@
 These theoretical distributions of N and water regulate leaf physiological traits that constrain A differently for sun and shade leaves.
-The balance between the costs and energetic benefits of these leaf traits affecting photosynthetic capacity create trade-offs impacting canopy carbon gain [@givnish1988adaptation].
+The balance between the costs and energetic benefits of these leaf traits affecting photosynthetic capacity create trade-offs impacting canopy C gain [@givnish1988adaptation].
 During A, CO~2~ diffuses from the atmosphere through stomata (g~s~), intercellular air space and into the chloroplast for fixation by Rubisco (mesophyll conductance, g~m~) [@tazoe2011using] and trade-offs exists that constrain each of these diffusion pathways.
-For optimal leaf carbon gain these pathways should act to minimize the resistance to CO~2~ diffusion while also limiting the intrinsic energy, water and construction costs which diminish the carbon return from A.
+For optimal leaf C gain these pathways should act to minimize the resistance to CO~2~ diffusion while also limiting the intrinsic energy, water and construction costs which diminish the C return from A.
 As these processes should not be considered independent of each other and must be integrated with hydraulic constraints [@griffiths2013you] predicting their behaviour within canopies remains difficult.
@@
 First, g~s~ should be efficiently distributed within a canopy to utilize supplies of light, N and water to maximize A [@peltoniemi2012co].
-•	This is because stomata are hypothesized to exhibit an optimal behaviour to maximize carbon gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
+•	This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
 Second, variation in photosynthetic capacity between sun and shade leaves has also been hypothesized to be due to differences in g~m~ [@duursma2006vertical].
 Mesophyll conductance can impose limitations on A as large as those associated with g~s~ [@warren2008stand; @ubierna2011estimation]and these limitations can reduce the efficiency of N use in A [@niinemets2007photosynthesis].
-Interpreting the coupled responses of both g~s~ and g~m~ to A across sun and shade leaves has major implications for predicting canopy carbon gain, but empirical measurements across tree canopies are still lacking.
+Interpreting the coupled responses of both g~s~ and g~m~ to A across sun and shade leaves has major implications for predicting canopy C gain, but empirical measurements across tree canopies are still lacking.
 
@@
 Additionally, assessing shade leaf behaviour is made difficult with accounting of short term light fluctuations within a canopy, via sunflecks.
-How shade leaves utilize sunflecks for short term carbon gain depends on the combined response time of gs and gm and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks].
+How shade leaves utilize sunflecks for short term C gain depends on the combined response time of gs and gm and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks].
 For example, at short timescales the utilization of sunflecks is limited by the slow response of stomatal opening, effectively limiting the maximum assimilation rate that can be achieved [@way2012sunflecks].
@@
 However, anatomical parameters which regulate gm with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are not likely adaptable during short light fluctuations.
-The physiological behaviour of shade leaves to maximize carbon gain must be assessed as both a degree of acclimation to local irradiance and as a potential response to transitory light availability.
+The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance and as a potential response to transitory light availability.
 
@@
 trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (+3&deg;C) treatments, to empirically evaluate the distribution of resources and leaf physiological behaviour of sun and shade leaves.
-1.	If whole tree canopies are optimized for carbon gain, then leaf N, K and photosynthetic capacity were predicted to be higher in full sun leaves.
+1.	If whole tree canopies are optimized for C gain, then leaf N, K and photosynthetic capacity were predicted to be higher in full sun leaves.
 2.	Under comparable leaf VPD, g~s~ and A should exhibit theoretical optimal behaviour across both sun and shade leaves, while g~m~ was predicted to scale positively with photosynthetic capacity.
 3.	As shade leaves are constrained by their underlying biochemistry and slow physiological responses, increases in A following sunfleck simulations were not expected to reach levels of full sun leaves.
-4.	The effects of climate warming was predicted to be greater in sun than shade leaves, seen as increases in stomatal closure and reductions in leaf carbon gain during summer months.
+4.	The effects of climate warming was predicted to be greater in sun than shade leaves, seen as increases in stomatal closure and reductions in leaf C gain during summer months.
 
@@
 A detailed description of the WTC operation and design is available in [@barton2010whole].
-Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced a constant elevated air temperature treatment (ET, 3&deg;C).
+Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced a constant elevated air temperature treatment of +3&deg;C (ET, ).
 The CO~2~ concentration inside the chamber was set to match outside air, tracking diurnal changes.
 Due to the temperature treatment air humidity was not explicitly controlled.
-Trees were watered with 70 l weekly from March 2013 to November 2013.
-From December 2013 to final harvest trees were watered fortnightly with mean monthly (100 yr) rainfall amount.
+Trees were watered weekly with 70 L from March 2013 to November 2013.
+From December 2013 to final harvest trees were watered fortnightly with the mean monthly (100 yr) rainfall amount.
 In February 2013 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering.
@@
 \
-Before seedlings were planted into each chamber they were maintained under well watered conditions in 35 l pots and kept inside each chamber.
+Before seedlings were planted into each chamber they were maintained under well watered conditions in 35 L pots and kept inside each chamber.
 This allowed for seedlings to gain sufficient biomass before planting while also allowing them to acclimate to chamber temperature treatments.
@@
 Gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.
-At this point, trees had a mean height of 34&pm;15.1 and 418&pm;323.1 cm and an estimated leaf area of 3.9&pm;0.1 and 6.2&pm;0.2 m^2^ for ambient and elevated temperature treatments, respectively.
+At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm, height of 348&pm;15.1 cm and an estimated leaf area of 3.9&pm;0.1 m^2^.
+For ET treatments trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an estimated leaf area of 6.2&pm;0.2 m^2^.
 
@@
 Measurements were taken on a representative sun and shade leaf for each tree during each measurement campaign.
-The youngest fully expanded leaf from the stem apex was chosen for gas exchange measurements and sun leaves were measured in the upper third of the canopy.
+The newest fully expanded leaf from the stem apex was chosen for gas exchange measurements and sun leaves were measured in the upper third of the canopy.
 In order to minimize any confounding effects of leaf age, shade leaves were measured first in the lower canopy then gradually higher during each campaign.
-As shade leaves most likely developed slower this assured that older leaves that exist in the lower canopy were avoided.
+As shade leaves most likely developed slower this assured that older leaves in the lower canopy were avoided.
 The nearest leaf on each branch was sampled for measurement of predawn leaf water potential.
@@
 Five ceptometer readings were recorded within the canopy at random locations of the height of each selected leaf.
-The mean of these readings was assumed to represent the overall leaf light environment of representative full sun and shade leaves for each tree.
+The mean of these readings was assumed to represent the local light environment of full sun and shade leaves for each tree.
 All measurements of PPFD and gas exchange were performed on full sun days between 10:00-14:30 h.
@@
 Leaf level gas exchange was measured with a standard 6 cm^2^ leaf chamber using a portable gas exchange system (LI-6400, LI-COR, Lincoln, NE, USA).
-This system was coupled with a tunable diode laser (TDL;TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of online carbon isotope discrimination.
+This system was coupled with a tunable diode laser (TDL;TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of online C isotope discrimination.
 The CO~2~ in the leaf chamber was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mols s^-1^.
 Two identical gas exchanges systems were run simultaneously, one in each of a randomly chosen WTC for each temperature treatment.
-This paired design allowed for direct leaf comparisons to be made with simultaneous measurements of similar leaf types.
-Leaf temperatures were controlled at the current ambient or +3&deg;C chamber air temperature.
+Leaf temperatures were controlled at the current AT or ET chamber air temperature.
 PPFD in the chamber was set to match the individual light environment of each leaf type (explained above).
 Sun flecks were then simulated for shade leaves by increasing the chamber PPFD to the light environment of the full sun leaf in the same tree.
-When shade leaves were exposed to high light they were allowed enough time to equilibrate with chamber conditions until stable (ca.
+The maximum sunfleck response was then recorded once shade leaves re-stablized in the leaf cuvette (ca.
 25 min).
@@
 \
-Once the chamber environment for leaves was stable the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet.
+Once the chamber environment for leaves was stable, for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet.
 These gases were dried by passing through napion gas dryers in the respective gas lines, and then ^12^CO~2~ and ^13^CO~2~ concentrations were measured for each gas by the TDL.
-Reference, sample and 2 calibration gases were run on alternating 80 s loops (20 s each), one for each paired leaf, for a total of 12 min.
+Reference, sample and 2 calibration gases were run on alternating 80 s loops (20 s each), one for each paired WTC leaf, for a total of 12 min.
 This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference and samples gases for calculations.
 The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDLAS on each measurement cycle.
-Net CO~2~ assimilation rate, g~s~, transpiration, and intercellular [CO~2~] were auto logged every 15 s for each gas exchange system over the same 12 min interval.
+Photosynthesis, g~s~, transpiration, and intercellular [CO~2~] were auto logged every 15 s for each gas exchange system over the same 12 min interval.
 
@@
 
+\
+(1)
 $$\delta = \left(\frac{R_s}{R_{stnd}} -1\right)1000$$
@@
 
+\
+(2)
 $$\Delta_{obs} = \frac{\xi(\delta^13C_o-\delta^13C_e)}{1000+\delta^13C_e-\xi(\delta^13C_o-\delta^13C_e)}$$
@@
 
+\
+(3)
 $$\xi = \frac{C_{e}}{C_{e}-C_{o}}$$
 
-$\xi$ is the ratio of the CO~2~ entering the well mixed leaf chamber to the CO~2~ draw down by the leaf.
+and $\xi$ is the ratio of the CO~2~ entering the well mixed leaf cuvette to the CO~2~ draw down by the leaf.
+
+Second, C isotope discrimination during C~3~ photosynthesis ($\Delta$) is the resultant discrimination from CO~2~ diffusion from the atmosphere to the site of carboxylation, consisting or a series of fractionation steps described in [@evans1986carbon].
+In this experiment, a modified form of this equation presented in [@evans2013temperature] with ternary effect corrections by [@farquhar2012ternary] was used such that:
 
 \
-Second, leaf ^13^CO~2~ during C~3~ photosynthesis ($\delta$^13^C) is the resultant discrimination from CO~2~ diffusion from the atmosphere to the site of carboxylation, consisting or a series of fractionation steps described in [@evans1986carbon].
-In this experiment, a modified form of this equation presented in [@evans2013temperature] with ternary effect corrections by [@farquhar2012ternary] was used such that:
+(4)
+$$\Delta_o = \Delta_i - \Delta_{gm} - \Delta_e - \Delta_f$$
 
-from Evans & Von Caemmerer
-$\Delta$~o~ = $\Delta$~i~ - $\Delta$~gm~ - $\Delta$~e~ - $\Delta$~f~
-where $\Delta$~o~ is the observed discrimation and  $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contritbutions of fractionataion if C~i~=C~c~, g~m~, respiration and photorespiration, respectively.
+where $\Delta$~o~ is the observed discrimation and  $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionataion if C~i~ = C~c~, g~m~, respiration and photorespiration, respectively.
 The equations for each are as follows:
 
+\
+(5)
 $$
 \begin{aligned}
-\Delta = &\frac{1}{1-t}a'+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}\\
-            &\frac{1+t}{1-t}\left(b-a'-\frac{eR_d}{A+R_d}\right)\frac{A}{g_mC_a}-\\
-            &\frac{1+t}{1-t}\left(\frac{eR_D}{(A+R_d)C_a}(C_i-\Gamma_*)\right)- \frac{1+t}{1-t}\left(f\frac{\Gamma_*}{C_a}\right)\\
+\Delta_i = &\frac{1}{1-t}a'+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}
 \end{aligned}
@@
 
+\
+(6)
 $$
 \begin{aligned}
-\Delta_i = &\frac{1}{1-t}a'+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}\\
+\Delta_{gm} = &\frac{1+t}{1-t}\left(b-a'-\frac{eR_d}{A+R_d}\right)\frac{A}{g_mC_a}
 \end{aligned}
@@
 
-$$
-\begin{aligned}
-\Delta_gm = &\frac{1+t}{1-t}\left(b-a'-\frac{eR_d}{A+R_d}\right)\frac{A}{g_mC_a}-\\
-
-\end{aligned}
-$$
-
+\
+(7)
 $$
@@
 
-$$
-\begin{aligned}
-\Delta = &\frac{1}{1-t}a'+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}\\
-            &\frac{1+t}{1-t}\left(b-a'-\frac{eR_d}{A+R_d}\right)\frac{A}{g_mC_a}-\\
-            &\frac{1+t}{1-t}\left(\frac{eR_D}{(A+R_d)C_a}(C_i-\Gamma_*)\right)- \frac{1+t}{1-t}\left(f\frac{\Gamma_*}{C_a}\right)\\
-\end{aligned}
-$$
-
+\
+(8)
 $$
@@
 
-where:
-
-$$t = \frac{(1+a')E}{2g^t_{ac}}$$
-
-E denotes the transpiration rate and $\text{g}^t_{ac}$ is the total conductance to CO~2~ diffusion to both the boundary layer and stomatal conductance (von Caemmerer 1981).
-C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_*$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~).
-In this experiment both $\Gamma_*$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulis* from [@crous2012light].
-The different fractionation factors include; diffusion through water (a~i~, 1.8&permil;), Rubisco carboxylation (b, 29&permil;), the photorespiratory fractionation (f, 16.2&permil;) and the combined fractionation through the boundary layer and the stomata (a').
+where the different fractionation factors include; diffusion through water (a~i~, 1.8&permil;), Rubisco carboxylation (b, 29&permil;), the photorespiratory fractionation (f, 16.2&permil;) and the combined fractionation through the boundary layer and the stomata (a').
 a' is defined by:
 
+\
+(9)
 $$a' = \frac{a_b(C_a-C_i)+a(C_s-C_i)}{C_a-C_i}$$
@@
 where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b (2.9&permil;) is the fractionation from boundary layer diffusion and a is the fractionation due to diffusion in air~(4.4&permil;) [@evans1986carbon].
+C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$  is the compensation point in the absence of mitochondrial respiration in the light (R~d~).
+In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulis* from [@crous2012light].
+E denotes the transpiration rate and $\text{g}^t_{ac}$ is the total conductance to CO~2~ diffusion to both the boundary layer and stomatal conductance **(von Caemmerer 1981).
+The ternary effect corrections (t) are desribed by:
+
+\
+(10)
+$$t = \frac{(1+a')E}{2g^t_{ac}}$$
+
 The CO~2~ diffusion from the intercellular airspace to the chloroplast, g~m~, is given by its relationship to the CO~2~ assimilation rate (A) by:
 
+\
+(11)
 $$g_m = \frac{A}{C_i-C_c}$$
@@
 where C~c~ is the chloroplast CO~2~ partial pressure.
-Examples of this approach to measure gas exchange and carbon isotope discrimination are presented in [@evans2013temperature].
-Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular arispaces to the site of carboxylation were then estimated using eq.
-7.
+Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular arispaces to the site of carboxylation were then estimated using Equation 11.
+Examples of this approach to measure gas exchange and C isotope discrimination are presented in [@evans2013temperature].
+The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck where then compared as a function of C~i~:C~a~.
 
 \
-Photosynthetic CO~2~ response (AC~i~) curves were developed at 25&deg;C for one sun and shade leaf for each chamber prior to the initiation of the drought treatment.
-Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25&deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical].
+Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each chamber prior to the initiation of the drought treatment.
+Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@Duursma2014] in R [@RDevelopmentCoreTeam2011].
 
@@
 Predawn ($\Psi_{pd}$) and midday ($\Psi_l$) leaf water potentials were measured for sun and shade leaves during each gas exchange campaign using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA).
-Predawn leaf water potential on selected leaves for each chamber was measured before sunrise on the same day as gas exchange measurements.
+Predawn leaf water potential was measured before sunrise on the same day as gas exchange measurements.
 Leaves used for gas exchange were immediately sampled for $\Psi_l$ once measurements were completed.
@@
 
+\
+(12)
 $$K = \frac{E}{\Psi_{pd}~-\Psi_l-hp_wg}$$
@@
 where the term Hp~w~g represents the gravitational potential assumed to be minimal in this experiment due to small tree size [@whitehead1998regulation].
-Leaf level instantaneous transpiration efficiency (ITE) was calculated as leaf photosynthesis divided by leaf transpiration.
+Leaf level instantaneous transpiration efficiency (ITE) was calculated as leaf photosynthesis divided by transpiration.
 The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2012reconciling] using non-linear resgression, where K=0.5 [see @duursma2013near].
@@
 ##Data analysis
-Differences in experimental parameters to either the warming treatment or leaf type were analysed by one-way analysis of variance in R [@RDevelopmentCoreTeam2011] with chambers as random effects.
-Mixed model ANOVAs of A max and leaf chemistry were performed using the nlme package [@Pinheiro2015] in R  and r~2~ values of mixed models were computed as in [@nakagawa2013general].
-Confidence intervals (95 %) of mixed effect linear models of leaf photosynthesis as a functions of different physiological parameters were generated using bootstrapping methods with 999 simulations in the lme4 package [@bates2013lme4] in R.
-For non-linear mixed models models confidence intervals were estimated by fitting a generalized additive model to the data with the mgcv package in R.
+Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R with WTC as a random effect.
+Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general].
+Confidence intervals (95 %) of mixed effect linear models of leaf photosynthesis as a functions of different physiological parameters were generated using bootstrapping methods with 999 simulations in the 'lme4' package [@bates2013lme4] in R.
+For non-linear mixed models models confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package in R.
 Results were considered significant at P&leq;0.05.
@@
 
-## Sun and Shade leaf resource distribution
-Across all measurement campaigns 1m canopy integrated PPFD was significantly different between sun and shade leaves (P<0.001) and PPFD was reduced by 78% in the shade (Figure 1).
+## Leaf resource distribution
+Across all measurement campaigns 1 m canopy integrated PPFD was significantly different between sun and shade leaves (P<0.001) and PPFD was reduced by 78% in the shade (Figure 1).
 The distribution of H~2~O, via K, was not different across sun and shade leaves (Table 1).
@@
 Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P<0.001, Figure 4a).
-In sun leaves, g~s~ and A showed optimal behavior with diminishing rates of A with highest values of g~s~ (Figure 5a).
+In sun leaves, g~s~ and A showed optimal behavior **as in (cite)** with diminishing rates of A with highest values of g~s~ (Figure 5a).
 However, in shade leaves this relationship was clearly uncoupled.
-With increased light availability, both increases in A and g~s~ in shade leaves extened the realtionship of A and g~s~ seen in sun leaves.
+With increased light availability, both increases in A and g~s~ in shade leaves extened the relationship of A and g~s~ seen in sun leaves.
 

### 2015-07-27 master_scripts/lightson_sigletters.R

@@ -93,4 +93,37 @@ gm_lightson_siglets2 <- gm_lightson_siglets$mcletters$Letters
 
 visreg(gm_leaf)
 write.csv(gm_lightson_siglets2, "master_scripts/sigletters/sl_gm_lightson.csv", row.names=FALSE)
+
+###vpd-----------------------------------------------------------------------------------------
+vpd_temp <- lme(VpdL ~ temp ,random=~1|chamber, data=lightson)
+summary(vpd_temp)
+anova(vpd_temp)
+visreg(vpd_temp)
+
+vpd_leaf <- lme(VpdL~ tukeyid, random=~1|chamber, data=ge_agg)
+  summary(vpd_leaf)
+  anova(vpd_leaf)
+  visreg(vpd_leaf)
+
+tukey_vpd<- glht(vpd_leaf, linfct = mcp(tukeyid = "Tukey"))
+vpd_lightson_siglets<- cld(tukey_vpd)
+vpd_lightson_siglets2 <- vpd_lightson_siglets$mcletters$Letters
+
+write.csv(vpd_lightson_siglets2, "master_scripts/sigletters/sl_vpd_lightson.csv", row.names=FALSE)
+
+###E-----------------------------------------------------------------------------------------
+e_temp <- lme(Trmmol ~ temp ,random=~1|chamber, data=lightson)
+summary(e_temp)
+anova(e_temp)
+
+e_leaf <- lme(Trmmol~ tukeyid, random=~1|chamber, data=ge_agg)
+summary(e_leaf)
+anova(e_leaf)
+visreg(e_leaf)
+
+tukey_e<- glht(e_leaf, linfct = mcp(tukeyid = "Tukey"))
+e_lightson_siglets<- cld(tukey_e)
+e_lightson_siglets2 <- e_lightson_siglets$mcletters$Letters
+
+write.csv(e_lightson_siglets2, "master_scripts/sigletters/sl_e_lightson.csv", row.names=FALSE)
   

### 2015-07-27 master_scripts/resouces_sigletters.R

@@ -1,0 +1,187 @@
+source("functions and packages/functions.R")
+library(doBy)
+library(visreg)
+library(multcomp)
+library(nlme)
+
+
+treatments <- read.csv("raw data/temp_trt.csv") 
+
+####generate sig letters for data tables for wtc3 manuscript
+
+##variables include:  WP*2, K, Narea, 13C, Vcmax, Jmax
+
+####aci parameters--------------------------------------------------------------------------------------------------
+aciparam <- read.csv("calculated_data/aciparameters.csv")
+  aciparam$tukeyid <- as.factor(paste(aciparam$leaf, aciparam$temp, sep="-"))
+
+#1: jmax : no effect of ET
+J_temp <- lme(Jmax ~ temp ,random=~1|chamber, data=aciparam)
+  summary(J_temp)
+  anova(J_temp)
+  visreg(J_temp)
+
+##full model
+
+J_leaf <- lme(Jmax~ tukeyid, random=~1|chamber, data=aciparam)
+  summary(J_leaf)
+  anova(J_leaf)
+  visreg(J_leaf)
+
+tukey_Jmax<- glht(J_leaf, linfct = mcp(tukeyid = "Tukey"))
+Jmax_siglets<- cld(tukey_Jmax)
+Jmax_siglets2 <- Jmax_siglets$mcletters$Letters
+
+write.csv(Jmax_siglets2, "master_scripts/sigletters/slr_jmax.csv", row.names=FALSE)
+
+#2: Vcmax: no effect of ET
+
+vc_temp <- lme(Vcmax ~ leaf ,random=~1|chamber, data=aciparam)
+  summary(vc_temp)
+  anova(vc_temp)
+  visreg(vc_temp)
+
+##full model
+
+vc_leaf <- lme(Vcmax~ tukeyid, random=~1|chamber, data=aciparam)
+  summary(vc_leaf)
+  anova(vc_leaf)
+  visreg(vc_leaf)
+
+tukey_vcmax<- glht(vc_leaf, linfct = mcp(tukeyid = "Tukey"))
+vcmax_siglets<- cld(tukey_vcmax)
+vcmax_siglets2 <- vcmax_siglets$mcletters$Letters
+
+write.csv(vcmax_siglets2, "master_scripts/sigletters/slr_vcmax.csv", row.names=FALSE)
+
+
[…]

### 2015-07-27 scripts/fulldelta_eq.R

@@ -1,0 +1,7 @@
+$$
+  \begin{aligned}
+\Delta = &\frac{1}{1-t}a'+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}\\
+&\frac{1+t}{1-t}\left(b-a'-\frac{eR_d}{A+R_d}\right)\frac{A}{g_mC_a}-\\
+                      &\frac{1+t}{1-t}\left(\frac{eR_D}{(A+R_d)C_a}(C_i-\Gamma_*)\right)- \frac{1+t}{1-t}\left(f\frac{\Gamma_*}{C_a}\right)\\
+                      \end{aligned}
+                      $$

### 2015-07-28 manuscript/manuscript_wtc3.Rmd
edits to figures, methods and results
added contribution of gm to delt obs

@@
 
+```{r, echo=FALSE}
+airvarsfig <- "1"
+parfig <- "2"
+acifig <- "3"
+waterusefig <- "4"
+gebarfig <- "5"
+agmgsfig <- "6"
+deltafig <- "7"
+ciccfig <- "8"
+
+get1fig <- "S1"
+get2fig <- "S2"
+accfig <- "S3"
+```
+
 # Summary {.unnumbered}
@@
 First, g~s~ should be efficiently distributed within a canopy to utilize supplies of light, N and water to maximize A [@peltoniemi2012co].
-•	This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
+This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
 Second, variation in photosynthetic capacity between sun and shade leaves has also been hypothesized to be due to differences in g~m~ [@duursma2006vertical].
@@
 Additionally, assessing shade leaf behaviour is made difficult with accounting of short term light fluctuations within a canopy, via sunflecks.
-How shade leaves utilize sunflecks for short term C gain depends on the combined response time of gs and gm and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks].
+How shade leaves utilize sunflecks for short term C gain depends on the combined response time of g~s~ and g~m~ and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks].
 For example, at short timescales the utilization of sunflecks is limited by the slow response of stomatal opening, effectively limiting the maximum assimilation rate that can be achieved [@way2012sunflecks].
-Mesophyll conductance has been shown to respond to environmental factors across both long and rapid time scales, possibly faster than gs [@flexas2008mesophyll].
-However, anatomical parameters which regulate gm with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are not likely adaptable during short light fluctuations.
+Mesophyll conductance has been shown to respond to environmental factors across both long and rapid time scales, possibly faster than g~s~ [@flexas2008mesophyll].
+However, anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are not likely adaptable during short light fluctuations.
 The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance and as a potential response to transitory light availability.
@@
 trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (+3&deg;C) treatments, to empirically evaluate the distribution of resources and leaf physiological behaviour of sun and shade leaves.
+Our hypothesese are as follows:
+
+\
 1.	If whole tree canopies are optimized for C gain, then leaf N, K and photosynthetic capacity were predicted to be higher in full sun leaves.
@@
 The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDLAS on each measurement cycle.
-Photosynthesis, g~s~, transpiration, and intercellular [CO~2~] were auto logged every 15 s for each gas exchange system over the same 12 min interval.
+Photosynthesis, g~s~, transpiration, vapor pressure deficit and intercellular [CO~2~] were auto logged every 15 s for each gas exchange system over the same 12 min interval.
 
@@
 \
-Predawn ($\Psi_{pd}$) and midday ($\Psi_l$) leaf water potentials were measured for sun and shade leaves during each gas exchange campaign using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA).
+Predawn ($\Psi_{pd}$) and midday ($\Psi_l$) leaf water potentials (MPa) were measured for sun and shade leaves during each gas exchange campaign using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA).
 Predawn leaf water potential was measured before sunrise on the same day as gas exchange measurements.
@@
 All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
-Leaf water potential and transpiration (E) from gas exchange were then used to calculate leaf specific hydraulic conductance (K) through the equation:
+Leaf water potential and transpiration (E, mmol^-2^ s^-1^) from gas exchange were then used to calculate leaf specific hydraulic conductance (K) through the equation:
 
@@
 ## Leaf resource distribution
-Across all measurement campaigns 1 m canopy integrated PPFD was significantly different between sun and shade leaves (P<0.001) and PPFD was reduced by 78% in the shade (Figure 1).
+Across all measurement campaigns 1 m canopy integrated PPFD was significantly different between sun and shade leaves (P<0.001) and PPFD was reduced by 78% in the shade (Figure `r parfig`).
 The distribution of H~2~O, via K, was not different across sun and shade leaves (Table 1).
 This was the result of neither $\Psi_{pd}$, $\Psi_l$ (Table 1) nor E (Table 2) differing between leaf types.
-Leaf N, on an area basis, was significantly higher in sun leaves than in shade leaves (p<0.001) by ca.
+Leaf N, on an area basis, was significantly higher in sun leaves than in shade leaves (p < 0.001) by ca.
 20% (Table1).
-No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K, E or N~area~ either within or across leaf types.
+Leaf mass per area was not different between leaf types.
+No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K, E, N~area~ or LMA either within or across leaf types.
 
 ## Leaf photosynthesis rates and photosynthetic capacity
-Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environoment (P<0.001, Table 2).
-Following an increase in available light, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P<0.001).
+Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environoment (P < 0.001, Table 2).
+Following an increase in available light, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001).
 No effect of the warming treatment was detected on rates of A of sun leaves or shade leaves at low or high light.
-Photosynthesis across all leaf types and treatments did not vary according across the leaf temperatures measured in this study or time of year (Figure S1a).
+Photosynthesis across all leaf types and treatments did not vary according across the leaf temperatures measured in this study or time of year (Figure `r get1fig`a).
 
 \
-The photsynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (p<0.001 & p=0.02, respectively), as evident from AC~i~ curves from each chamber (Figure 2a).
+The photsynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (P < 0.001 & P = 0.02, respectively), as evident from AC~i~ curves from each chamber (Figure `r acifig`a).
 No effect of the warming treatment was detected on either parameter (Table 1).
-Within individual chambers Vc~max~ was postively related to leaf N~area~ across leaf types and temperature treatments (P =0.01, Figure 2b).
-Additionally, leaf N~area~ was postively related to A across gas exchange campaigns and leaf types (P<0.001, Figure 2c).
+Within individual chambers Vc~max~ was postively related to leaf N~area~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
+Additionally, leaf N~area~ was postively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).
 As Vc~max~ is condisdered implicit with g~m~ (results below) with conventional AC~i~ curves, a AC~c~ curves were also simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.
-When accounting for the effect of g~m~, photosynthetic capacity was still greater in sun leaves than shade leaves (Figure S2)
+When accounting for the effect of g~m~, photosynthetic capacity was still greater in sun leaves than shade leaves (Figure `r accfig`)
 
 ## Leaf water use efficiency
-Leaf ITE was significantly greater in sun leaves than in shade leaves (P=0.001, Table 2).
-Following an increase in available light, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P<0.001).
-ITE in sun leaves was reduced in the warming treatment (P=0.021) but no effect was noted in shade leaves with low or high light.
-For all leaf types and light treatments there was a strong response of ITE to VPD, and individual data points broadly corresponded to response curves from the optimal ITE model (Figure 3a).
+Leaf ITE was significantly greater in sun leaves than in shade leaves (P = 0.001, Table 2).
+Following an increase in available light, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
+ITE in sun leaves was reduced in the warming treatment (P = 0.021) but no effect was noted in shade leaves with low or high light.
+For all leaf types and light treatments there was a strong response of ITE to VPD, and individual data points broadly corresponded to response curves from the optimal ITE model (Figure `r waterusefig`a).
 The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
-Across leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure S2a)
+Across leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a)
 
-Leaf $\delta$^13^C significantly decreased from sun leaves to shade leaves by ca.
-2&permil; (p<0.001, Table 1).
-Within leaf types no affects of the warming treatment on leaf $\delta$^13^C were detected.
-Leaf $\delta$^13^C and N~area~ were postively correlated for all leaves (P<0.001, Figure 3b).
+Leaf $\delta$^13^C, as an index of water use efficiencey [@marshall2007sources], significantly decreased from sun leaves to shade leaves by ca.
+2&permil; (p < 0.001, Table 1).
+No effects of the warming treatment on leaf $\delta$^13^C were detected.
+Leaf $\delta$^13^C and N~area~ were postively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C and higher N investment in sun leaves.
 
-leaf 13C with leaf N and what this stands for
+*leaf 13C with leaf N and what this stands for*
 
 ## Stomatal conductance, mesophyll conductance and CO~2~ drawdown
-Mean g~s~ was significantly higher in shade compared to sun leaves (+18%), under their local light environment (P=0.004, Table2).
-Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P<0.001, Figure 4a).
-In sun leaves, g~s~ and A showed optimal behavior **as in (cite)** with diminishing rates of A with highest values of g~s~ (Figure 5a).
+Mean g~s~ was significantly higher in shade compared to sun leaves (+18%), under their local light environment (P = 0.004, Table2).
+Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001, Figure `r gebarfig`a).
+In sun leaves, g~s~ and A showed optimal behavior *as in (cite)* with diminishing rates of A with highest values of g~s~ (Figure `r agmgsfig`a).
 However, in shade leaves this relationship was clearly uncoupled.
 With increased light availability, both increases in A and g~s~ in shade leaves extened the relationship of A and g~s~ seen in sun leaves.
-
-discrimination contributions and \delta vs ci/ca go here
+\
+The observed carbob isotope discrimination was postively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`).
+Here, the carbon isotope discrimination associtaed with g~m~ accounted for the majority of $\Delta$~o~ (69.7%pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments.
+Mean g~m~ was significantly higher in sun compared to shade leaves (+27%), under their local light environment (P < 0.001, Table2).
+Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001).
+Relationships between g~m~ and A were nearly proportional within leaf types and light treatments (Figure `r gebarfig`b,c).
+Photosynthesis scaled postively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves in high light resulted in the highest rates of A (Figure `r agmgsfig`b).
+No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types or light treatments.
+Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figures `r get1fig`b and `r get2fig`b, respectively) and a slightly negative relationship with g~m~ and increasing leaf temperature.
 
 \
-Mean g~m~ was significantly higher in sun compared to shade leaves (+27%), under their local light environment (P=<0.001, Table2).
-Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P<0.001, Figure 4b).
-Relationships between g~m~ and A were nearly proportional within leaf types and light treatments (Figure 4b,c).
-Photosynthesis scaled postively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves in high light resulted in the highest rates of A (Figure 5b).
-No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types or light treatments.
-Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figures S1b and S2b, respectively) and only a slight negative relationship was detected with g~m~ and increasing leaf temperature.
-
-\
-The combination of a non-limiting supply of leaf K and more open stomata lead to a significant increase in C~i~ in shade compared to sun leaves (Figure 6a).
+The combination of a non-limiting supply of leaf K and more open stomata lead to a significant increase in C~i~ in shade compared to sun leaves (Figure `r ciccfig`a).
 As increases in A in shade leaves were associated with higher rates of g~s~ the C~i~ was similar than under low light but still greater than sun leaves.
-As the drop in CO~2~ for intecellular spaces to the cholorplast, C~i~-C~c~, measures the coordiation between gm and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure 6c).
+As the drop in CO~2~ for intecellular spaces to the cholorplast, C~i~-C~c~, measures the coordiation between gm and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure `r ciccfig`b,c).
 This was a direct result of the increase in g~m~ and its subsequent direct effect on A, when compared to both sun and shade leaves that are acclimated to their local light environment.
-CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable acorss the range of temperatures measured and gas exchange campaigns (Figure S2c and S1c, respectively).
-
-## delta contrib, deltav cica
-
-The difference between $\delta$^13^Ce and $\delta$^13^Co was mostly accounted for by discrimination associated with mesophyll conductance (85%) compared to contributions from respiration and photorespiration.
+CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable acorss the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
 
@@
 
-# Tables {.unnumbered}
+# List of Tables
 
-```{r tablex, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-library(ascii)
-resource_table <- read.csv("master_scripts/resource_table.csv")
-print(ascii(resource_table,  include.rownames=FALSE,
-          colnames =c("Leaf", "Temperature", "LMA", "Narea", "Vcmax", "Jmax", 
-                       "Leaf K", "WP_pre", "WP_mid", "13C")),type="pandoc")
+**Table 1**.
+*Eucalyptus tereticornis* leaf traits and the distribution of resources between full sun and shade leaves under ambient and elevated temperature treatments.
+Leaf mass per area, leaf Nitrogen~area~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
+Values of Vc~max~ and J~max~ are treament mean(standard error) from AC~i~ curves measured in each chamber before the drought treatment.
+Different letters represent significant differences between treatments.
+The P value represents the overall difference between leaf types and warming treatments.
 
-```
-**Table 1.** Canopy resouce distribution among sun and shade leaves under elevated and ambient termperatures
+\
+**Table 2**.
+Responses of leaf level gas exchange parameters of *Eucalyptus tereticornis* trees between full sun and shade leaves under ambient and elevated temperature treatments.
+Each value reflects the mean (&pm; 1 standard error) for each treatment across all gas exchange campaigns (n=6).
+Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD~leaf~ are kPa.
+Different letters represent significant differences between treatments.
+The P value represents the overall difference between leaf types and both light and warming treatments.
 
-```{r table1, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
-library(ascii)
-phystable <- read.csv("master_scripts/ge_table.csv")
-print(ascii(phystable,
-            include.rownames=FALSE,
-            colnames=c("Leaf", "Temperature", "Light", "Photosynthesis", "gs", "gm", 
-                       "Ci", "Transpiration", "VPD")),
-            type='pandoc')
+# List of Figures
 
-```
-**Table 2.** Mean gas exchange parameters across leaf types with light and temperature treatments
+**Figure 1**. Daily maximum and minimum temperature (a) and total daily PPFD (b) for each chamber across the experiment duration.
 
-# Figures {.unnumbered}
-
-```{r par, fig.width=6, fig.height=6, echo=FALSE}
-source("master_scripts/par.R")
-```
-
-**Figure 1**.
+\
+**Figure 2**.
 Local light environment for sun and shade leaves for each gas exchange campaign.
@@
 
-```{r acicurves, fig.width=6, fig.height=8, echo=FALSE}
-source("master_scripts/aci_vcmaxnitro.R")
-```
-
-**Figure 2**.
+\
+**Figure 3**.
 (a) Photosynthetic CO~2~ response (ACi) curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.
@@
 
-```{r wateruse, fig.width=6, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE}
-source("master_scripts/wateruse.R")
-```
-
-**Figure 3**.
+\
+**Figure 4**.
 (a) Response of instantaneous water use efficiency (ITE) to leaf vapor pressure deficit (VPD) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
@@
 
-```{r barplots, fig.width=10, fig.height=5, echo=FALSE}
-source("master_scripts/physiology_barplots.R")
-```
+\
+**Figure 5**.
+The mean &pm; 1 standard error of stomatal conductance (a), mesophyll conductance (b) and photosynthesis rate (c) of sun leaves and shade leaves at both low and high light with standard errors.
 
-**Figure 4**. The mean stomatal conductance (a), mesophyll conductance (b) and photosynthesis rate (c) of sun leaves and shade leaves at both low and high light with standard errors.
-
-```{r Agmgs, fig.width=6, fig.height=8, echo=FALSE}
-source("master_scripts/Agmgs.R")
-```  
-**Figure 5**.
+\
+**Figure 6**.
 The response of leaf photosynthesis rates to stomatal conductance (a) and mesophyll conductance (b) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
@@
 
-```{r cicc, fig.width=10, fig.height=5, echo=FALSE}
-source("master_scripts/cicc_bar.R")
-```
+\
+**Figure 7**.
+Relationship between the observed discrimination of ^13^CO~2~ and measured C~i~/C~a~ for sun leaves and shade leaves at both low and high light.
+The solid line represents the the theortical line for C3 plants from [@evans1986carbon].
 
-**Figure 6**.
-The mean intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light with standard errors.
+\
+**Figure 8**.
+The mean &pm; 1 standard error of intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light with standard error
 
-# Supporting Information {.unnumbered}
-```{r gasex_temp1, fig.width=6, fig.height=10, echo=FALSE}
-source("master_scripts/gasex_temp.R")
-```  
+\
 **Figure S1**.
@@
 
-```{r gasex_temp2, fig.width=6, fig.height=10, echo=FALSE}
-source("master_scripts/gasex_temp2.R")
-```  
+\
 **Figure S2**.
@@
 
-```{r acc, fig.width=6, fig.height=6, echo=FALSE}
+\
+**Figure S3**.
+Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
+C~c~ values were predicited with mesophyll conductance, thus curves represent chloroplastic photosynthetic parameters at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+
+# Tables {.unnumbered}
+
+**Table 1**.
+
+\
+**Table 2**.
+
+# Figures {.unnumbered}
+
+**airvars goes here**
+**Figure 1**
+
+```{r par, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/par.R")
+```
+
+**Figure 2**.
+
+```{r acicurves, fig.width=6, fig.height=8, warning=FALSE,echo=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/aci_vcmaxnitro.R")
+```
+
+**Figure 3**.
+
+```{r wateruse, fig.width=6, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
+source("master_scripts/wateruse.R")
+```
+
+**Figure 4**.
+
+```{r barplots, fig.width=10, fig.height=5, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/physiology_barplots.R")
+```
+
+**Figure 5**.
+
+```{r Agmgs, fig.width=6, fig.height=8, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/Agmgs.R")
+```
+
+**Figure 6**.
+
+```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/delta_cica.R")
+```
+
+**Figure 7**.
+
+```{r cicc, fig.width=10, fig.height=5, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/cicc_bar.R")
+```
+
+**Figure 8**.
+
+# Supporting Information {.unnumbered}
+```{r gasex_temp1, fig.width=6, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/gasex_temp.R")
+```  
+**Figure S1**.
+
+```{r gasex_temp2, fig.width=6, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/gasex_temp2.R")
+```  
+**Figure S2**.
+
+```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
 source("master_scripts/ACc_figure.R")
@@
 **Figure S3**.
-Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
-C~c~ values were predicited with mesophyll conductance, thus curves represent chloroplastic photosynthetic parameters at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-
-```{r deltacica, fig.width=6, fig.height=6, echo=FALSE}
-source("master_scripts/delta_cica.R")
-```
-
-**Figure S4**.
-Relationship between the observed discrimination of ^13^CO~2~ and measured C~i~/C~a~ for sun leaves and shade leaves at both low and high light.
-The solid line represents the the theortical line for C3 plants from [@evans1986carbon].
 

### 2015-07-29 manuscript/manuscript_wtc3.Rmd
what i said b4

@@
 A detailed description of the WTC operation and design is available in [@barton2010whole].
-Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced a constant elevated air temperature treatment of +3&deg;C (ET, ).
+Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced a constant elevated air temperature treatment of +3&deg;C (ET, Figure `r airvars`).
 The CO~2~ concentration inside the chamber was set to match outside air, tracking diurnal changes.
@@
 # Figures {.unnumbered}
+```{r airvars, fig.width=6, fig.height=8, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/air_vars.R")
+```
 
-**airvars goes here**
 **Figure 1**

### 2015-07-30 manuscript/manuscript_wtc3.Rmd
added ecophys cite

@@
 A detailed description of the WTC operation and design is available in [@barton2010whole].
-Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced a constant elevated air temperature treatment of +3&deg;C (ET, Figure `r airvars`).
+Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced a constant elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarsfig`).
 The CO~2~ concentration inside the chamber was set to match outside air, tracking diurnal changes.
@@
 Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@Duursma2014] in R [@RDevelopmentCoreTeam2011].
+From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011].
 

### 2015-07-31 manuscript/manuscript_wtc3.Rmd
fixed a space

@@
 Second, variation in photosynthetic capacity between sun and shade leaves has also been hypothesized to be due to differences in g~m~ [@duursma2006vertical].
-Mesophyll conductance can impose limitations on A as large as those associated with g~s~ [@warren2008stand; @ubierna2011estimation]and these limitations can reduce the efficiency of N use in A [@niinemets2007photosynthesis].
+Mesophyll conductance can impose limitations on A as large as those associated with g~s~ [@warren2008stand; @ubierna2011estimation] and these limitations can reduce the efficiency of N use in A [@niinemets2007photosynthesis].
 Interpreting the coupled responses of both g~s~ and g~m~ to A across sun and shade leaves has major implications for predicting canopy C gain, but empirical measurements across tree canopies are still lacking.

### 2015-08-02 manuscript/manuscript_wtc3.Rmd
added discussion to manuscript
made pngs scripts and figures for esa talk

@@
 Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001, Figure `r gebarfig`a).
-In sun leaves, g~s~ and A showed optimal behavior *as in (cite)* with diminishing rates of A with highest values of g~s~ (Figure `r agmgsfig`a).
+In sun leaves, g~s~ and A exhibited optimal behavior with diminishing rates of A with highest values of g~s~ (Figure `r agmgsfig`a).
 However, in shade leaves this relationship was clearly uncoupled.
@@
 As increases in A in shade leaves were associated with higher rates of g~s~ the C~i~ was similar than under low light but still greater than sun leaves.
-As the drop in CO~2~ for intecellular spaces to the cholorplast, C~i~-C~c~, measures the coordiation between gm and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure `r ciccfig`b,c).
+As the drop in CO~2~ for intecellular spaces to the cholorplast, C~i~-C~c~, measures the coordiation between g~m~ and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure `r ciccfig`b,c).
 This was a direct result of the increase in g~m~ and its subsequent direct effect on A, when compared to both sun and shade leaves that are acclimated to their local light environment.
@@
 # Discussion
+Here we show that dynamic physiological responses in shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy.
+Tree canopies may maintain the ability to upregulate A beyond the constraints of their local light environment, even at the cost of valuable resources.
+Empirical data accounting for the total conductance of CO2 (both g~s~ and g~m~) reveal that these responses can lead to leaf C gain equivalent to sun leaves and may happen quicker than previously thought.
+Thus, failure to integrate the effects of g~m~ into process based tree growth models may underestimate the ability of canopies to alter their photosynthetic capacity when environmental conditions change.
+Additionally, with measurements recorded across a large natural range of temperatures we were unable to detect any effects of a +3 &deg;C warming treatment on leaf physiology.
 
-photosynthetic capacity: Mesophyll conductance asymmetrically affects key photosynthetic parameters and an assumption of infinite mesophyll conductance leads to underestimation of the maximum carboxylation rate Vcmax, maximum electron transport rate Jmax [@sun2014asymmetrical].
+## sub heading
+The allocation of N constrains photosynthesis rates and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
+Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979).
+As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vcmax and Jmax [@farquhar1980biochemical].
+Our data agree with these conventional conclusions as the distribution of N, both biochemical parameters and A were all reduced in shade leaves.
+
+\
+Photosynthesis in trees is also limited by the ability to supply water to the upper canopy.
+Using a two-leaf model Peltoniemi et al.
+[-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if leaf K is also optimally distributed.
+In this study, variation in leaf N distribution and A rates were not associated with subsequent changes in leaf K between sun and shade leaves.
+This is meaningful as foliage tends to operate at greater water limitations in the upper canopy, despite the ability of sun leaves to develop more effective vascular systems [@sellin2008effects;@niinemets2012optimization].
+As both sun and shade leaves were acclimated to their local light environment this sub-optimal distribution of water supply has the potential to diminish the C return for the whole canopy.
+
+\
+The inefficient distribution of resources led to decreased instantaneous WUE in shade leaves throughout the experiment.
+Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was constant across leaf lifespans.
+From a canopy perspective this pattern in WUE appears to be detrimental to C gain as A in sun leaves was characterized by low rates of g~s~ and low C~i~ compared to similar Eucalyptus species [see @drake2014capacity].
+As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower WUE while the photosynthetic efficiency of sun leaves appears at least partially constrained by water supply.
+
+\
+The pattern of wasteful water use in shade leaves is important as we hypothesized that gs and A would be proportional across sun and shade leaves.
+In sun leaves variation in A and g~s~ were strongly correlated, exhibiting behaviour agreeing with optimal stomatal theory.
+However, lower rates of A in shade leaves were not coupled with decreases in g~s~ which resulted in the observed decreases in WUE.
+This is significant as optimal stomatal behaviour has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on sun leaves [see @prentice2014balancing; @lin2015optimal].
+As a result, the often used economic framework of balancing costs of using water versus N allocation to predict photosynthesis [@wright2003least] may break down when considering all leaves within a tree canopy.
+
+\
+This behaviour may be explained by the response of shade leaves to increases in light availability.
+Contrary to hypothesis, shade leaves were able to increase A to rates equivalent or even outperforming sun leaves under identical light intensity.
+It is possible that sustaining higher g~s~, at the cost of larger transpiration losses, is a strategy to increase the utilization efficiency of sunflecks.
+These results should not be entirely unexpected as shade leaves experience interchanging periods of sun and shade, despite canopy light extinction [@pearcy1990sunflecks].
+Additionally, empirical studies that assess photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@leakey2002relative] and understory plants [@chazdon1991importance; @brantley2009contribution].
+Our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy.
+
+\
+Increasing the drawdown of CO~2~ into the intercellular air spaces via g~s~, however, does not necessarily infer concomitant increases in A.
+This is because CO~2~ supply to the site of carboxylation, via g~m~, may be anatomically constrained.
+In this study A and g~m~ scaled positively across leaf types and increased proportionately with increased light intensity in shade leaves.
+Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of mesophyll conductance [@hanba2004overexpression; @heinen2009role; @li2014plant].
+This provides a potential explanation for our observed rapid increases in g~m~ but is untested in leaves of woody species.
+Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
+Here we provide empirical data showing g~m~ not only varies within a canopy but the up-regulation of g~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.
+
+\
+Similar to Laisk [-@laisk2005adjustment] we show evidence that optimal acclimation, in this case physiological adjustment of the drawdown of CO~2~ in shade leaves, is possibly directed towards occasional maximums of light availability over adaptations towards a low light environment.
+If shade leaves ‘lie in wait’ for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximise C gain, beyond conventional trade-offs associated with canopy resource distribution.
+If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour in shade leaves could be offset over the leaf lifespan.
+The potential benefits of C gain in leaves constrained by low light to levels equivalent to sun leaves may be large enough to tolerate excess losses from transpiration.
+However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.
+Thus, models which predict leaf photosynthesis from N distribution within a canopy will be flawed unless inclusion of canopy light extinction and the integrations of sunflecks on shade leaves are included [@de1997simple].
+
+\
+Furthermore, our findings have important implications for process based models that predict canopy C gain from rates of leaf photosynthesis.
+Although the distributions of resources, such as N and water, define leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
+The observed physiological behaviour leaves in the shade, at both low and high irradiance, suggests that current theories of leaf optimal behaviour are incomplete.
+Furthermore, gm must now be considered as a dynamic process in modelling whole canopy C gain.
+Additional empirical data are needed to determine both the mechanisms responsible for and the capacity of gm to rapidly increase CO2 drawdown.
+To improve our ability to predict whole canopy C gain future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.
 

### 2015-08-02 makepngs/wue.R

@@ -1,0 +1,68 @@
+source("functions and packages/functions.R")
+source("master_scripts/plot_objects.R")
+source("functions and packages/packages.R")
+library(mgcv)
+library(lme4)
+
+####panel a = ITE and VPD----------------------------------------------------------------------------------------------------
+
+ite <- read.csv("calculated_data/gmes_wellwatered.csv")
+g1_ite <- read.csv("calculated_data/g1_ite.csv")
+
+###get average by id
+ite_agg <- summaryBy(Photo+Cond+ Trmmol+VpdL ~ chamber+id+leaf +light+temp+leaflight+Month, 
+                     data=ite, FUN=mean, keep.names=TRUE)
+
+#add iWUE
+ite_agg$ite <- with(ite_agg, Photo/Trmmol)
+
+##remove shade-high
+ite_sunsha <- ite_agg[ite_agg$leaflight != "shade-high",]
+ite_sunsha <- droplevels(ite_sunsha)
+
+##dfr with lights on
+ite_lightson <- ite_agg[ite_agg$leaflight == "shade-high",]
+ite_lightson <- droplevels(ite_lightson)
+
+####panel B: 13C and N area--------------------------------------------------------------------------------------------------
+#read data
+treatments <- read.csv("raw data/temp_trt.csv")
+
+leaf_chem <- read.csv("raw data/leaf_chem.csv")
+leaf_mass <- read.csv("raw data/leaf_data.csv")
+
+#get vcmax per chamber 
+aciparam <- read.csv("calculated_data/aciparameters.csv")
+
+#calculate nitrogen content in leaves
+canopy_chem <- merge(leaf_mass[leaf_mass$wp_type=="mid",c(1, 3:4, 7:8)], leaf_chem[, 4:8], by=c("Month", "chamber", "leaf"))
+  #add treatments  
+  canopy_chem <- addtrt_func(canopy_chem)
+  canopy_chem$leafN <- with(canopy_chem, leaf_mass *(n_perc/100))
+  canopy_chem$lma <- with(canopy_chem, (leaf_mass/leaf_area)*10000) #g/m2
+  canopy_chem$leafN_area <- with(canopy_chem, lma *(n_perc/100))
[…]

### 2015-08-02 master_scripts/data_tables.R

@@ -2,12 +2,15 @@ source("functions and packages/functions.R")
 library(doBy)
 
 gasexchange  <- read.csv("calculated_data/gmes_wtc.csv")
+##calculate CC
+gasexchange$Cc<- with(gasexchange, Ci-Photo/gm)
 
 ###leaf data
-ge_agg <- summaryBy(Photo+Cond+Ci+Trmmol+gm+VpdL ~ chamber+id+leaf +light+temp+leaflight+Month+drydown, 
+ge_agg <- summaryBy(Photo+Cond+Ci+Cc+Trmmol+gm+VpdL ~ chamber+id+leaf +light+temp+leaflight+Month+drydown, 
                     data=gasexchange, FUN=mean, keep.names=TRUE)
 
-ge_overall <- summaryBy(Photo+Cond+Ci+Trmmol+gm+VpdL ~ leaf +temp+leaflight+drydown, 
+
+ge_overall <- summaryBy(Photo+Cond+Ci+Cc+Trmmol+gm+VpdL ~ leaf +temp+leaflight+drydown, 
                         data=gasexchange, FUN=c(mean, se))
 
 
@@ -36,10 +39,10 @@ ge_vars <- ge_table[, 1:3]
   ge_vars$temp <- gsub("ambient", "Ambient", ge_vars$temp)
   ge_vars$temp <- gsub("elevated", "Elevated", ge_vars$temp)
   
-##no Ci/Cc for now
-##order: A, gs, gm, ite, e, vpd  
-ge_means <- ge_table[, c(5,6,9,17,8,10)]
-ge_se <- ge_table[, c(11,12,15,18,14,16)]
+
+##order: A, gs, gm, ite, e, vpd, Ci, CC  
+ge_means <- ge_table[, c(5,6,10,19,9,11,7,8)]
+ge_se <- ge_table[, c(12,13,17,20,16,18,14,15)]
 
 
 ###paste and round means and se together
@@ -49,7 +52,9 @@ ge3 <- data.frame(paste0(sprintf("%1.3f", round(ge_means[,3], 3)), " (", sprintf
 ge4 <- data.frame(paste0(sprintf("%1.2f", round(ge_means[,4], 2)), " (", sprintf("%1.2f", round(ge_se[,4],2)),")"))
 ge5 <- data.frame(paste0(sprintf("%1.2f",round(ge_means[,5], 2)), " (", round(ge_se[,5],2),")"))
[…]

### 2015-08-02 master_scripts/lightson_sigletters.R

@@ -2,12 +2,14 @@ source("functions and packages/functions.R")
 library(doBy)
 
 gasexchange  <- read.csv("calculated_data/gmes_wellwatered.csv")
+  ##calculate CC
+  gasexchange$Cc<- with(gasexchange, Ci-Photo/gm)
 
 ITE <- read.csv("calculated_data/ITE.csv")  
 ITE$tukeyid <- as.factor(paste(ITE$leaflight, ITE$temp, sep="-"))
 
 ###leaf data
-ge_agg <- summaryBy(Photo+Cond+Ci+Trmmol+gm+VpdL ~ chamber+id+leaf +light+temp+leaflight+Month+drydown, 
+ge_agg <- summaryBy(Photo+Cond+Ci+Cc+Trmmol+gm+VpdL ~ chamber+id+leaf +light+temp+leaflight+Month+drydown, 
                     data=gasexchange, FUN=mean, keep.names=TRUE)
 ge_agg$tukeyid <- as.factor(paste(ge_agg$leaflight, ge_agg$temp, sep="-"))
 
@@ -126,4 +128,50 @@ e_lightson_siglets<- cld(tukey_e)
 e_lightson_siglets2 <- e_lightson_siglets$mcletters$Letters
 
 write.csv(e_lightson_siglets2, "master_scripts/sigletters/sl_e_lightson.csv", row.names=FALSE)
+
+
+###ci-----------------------------------------------------------------------------------------
+ci_temp <- lme(Ci ~ temp ,random=~1|chamber, data=lightson)
+summary(ci_temp)
+anova(ci_temp)
+
+ci_leaf <- lme(Ci~ tukeyid, random=~1|chamber, data=ge_agg)
+summary(ci_leaf)
+anova(ci_leaf)
+visreg(ci_leaf)
+
+tukey_ci<- glht(ci_leaf, linfct = mcp(tukeyid = "Tukey"))
+ci_lightson_siglets<- cld(tukey_ci)
+ci_lightson_siglets2 <- ci_lightson_siglets$mcletters$Letters
+
+write.csv(ci_lightson_siglets2, "master_scripts/sigletters/sl_ci_lightson.csv", row.names=FALSE)
+
+###cc-----------------------------------------------------------------------------------------
+cc_temp <- lme(Cc ~ temp ,random=~1|chamber, data=lightson)
+summary(cc_temp)
[…]

### 2015-08-02 master_scripts/par.R

@@ -50,14 +50,9 @@ bar(par, c(leaf_type, Month), parbar, col=c(shacol,suncol), xlab="", ylab=parlab
 
 
 ###for png
-# png(filename = "markdown/ppfd.png", width = 11, height = 8.5, units = "in", res= 400)
-# 
-# bar(par, c(leaf_type, month), parbar, col=c("yellowgreen", "green4"), xlab="",  ylim=c(0, 2000), 
-#     half.errbar=FALSE, mar=c(5,7,2,2), ylab=parlab, cex.axis=1.75, cex.lab = 2, cex.names=2,
-#     legend = FALSE)
-# legend("topleft", c("Shade", "Sun"), pch=22,inset = 0.01, pt.bg=c("yellowgreen", "green4"),
-#        bty='n', cex=2)
-# 
+# png(filename = "makepngs/ppfd.png", width = 11, height = 8.5, units = "in", res= 400)
+# bar(par, c(leaf_type, Month), parbar, col=c(shacol,suncol), xlab="", ylab=parlab, ylim=c(0, 2000), 
+#     half.errbar=FALSE, mar=c(5,6,1,1),las=1,cex.axis=1.5, cex.lab=2, cex.names=2,mgp=c(3.5,1,0))
 # dev.off()
 
 

### 2015-08-03 manuscript/manuscript_wtc3.Rmd
changes to methods

@@
 # Introduction
-The need to understand and predict carbon (C) uptake in forest ecosystems is crucially important for assessing the impacts of a changing climate.
+The need to understand and predict carbon (C) uptake in forest ecosystems is crucially important for assessing the impacts of environmental change.
 Specifically, this involves accurate upscaling of leaf level measurements to predict whole canopy function.
 Due to the costs and limitations of efficient light harvesting plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling problematic.
-Incident PPFD declines exponentially with increasing leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor], and photosynthesis increases with irradiance from shade to sun until biochemical limitation occurs [@evans1995carbon].
+Incident PPFD declines exponentially with increasing leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor], and photosynthesis (A) increases with irradiance from shade to sun until biochemical limitation occurs [@evans1995carbon].
 Consequently, leaves modify structural and functional properties to efficiently intercept variable light and enhance photosynthetic capacity [@vogelman1996leaves; @valladares2004photosynthetic].
-To estimate whole canopy C gain it is thus necessary to account for the non-linear response of photosynthesis to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
+To estimate whole canopy C gain it is thus necessary to account for the non-linear response of A to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
 
@@
 Mesophyll conductance has been shown to respond to environmental factors across both long and rapid time scales, possibly faster than g~s~ [@flexas2008mesophyll].
-However, anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are not likely adaptable during short light fluctuations.
+However, anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are unlikely to adapt during short light fluctuations.
 The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance and as a potential response to transitory light availability.
@@
 How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients.
-The impacts of warming on plant physiological processes are obviously vast, yet differentiating their impacts on leaf physiology within a canopy will be essential for evaluating whole tree responses to a changing climate.
+The impacts of warming on plant physiological processes are obviously vast, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.
 
@@
 1.	If whole tree canopies are optimized for C gain, then leaf N, K and photosynthetic capacity were predicted to be higher in full sun leaves.
-2.	Under comparable leaf VPD, g~s~ and A should exhibit theoretical optimal behaviour across both sun and shade leaves, while g~m~ was predicted to scale positively with photosynthetic capacity.
+
+\
+2.	Under comparable leaf VPD, g~s~ and A should be proportional across both sun and shade leaves, while g~m~ was predicted to scale positively with photosynthetic capacity.
+
+\
 3.	As shade leaves are constrained by their underlying biochemistry and slow physiological responses, increases in A following sunfleck simulations were not expected to reach levels of full sun leaves.
+
+\
 4.	The effects of climate warming was predicted to be greater in sun than shade leaves, seen as increases in stomatal closure and reductions in leaf C gain during summer months.
@@
 Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced a constant elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarsfig`).
+Due to the temperature treatment air humidity was not explicitly controlled.
 The CO~2~ concentration inside the chamber was set to match outside air, tracking diurnal changes.
-Due to the temperature treatment air humidity was not explicitly controlled.
 Trees were watered weekly with 70 L from March 2013 to November 2013.
@@
 Seedlings were planted into each chamber after mean seedling height reached 100 cm.
+The top soils at this site, used in both pots and chambers, are an alluvial formation of low-fertility sandy loam soils with low organic matter and low water holding capacity.
 After 2 months, floors were installed 45 cm from the soil surface.
@@
 The nearest leaf on each branch was sampled for measurement of predawn leaf water potential.
-All leaves were selected and flagged 24 hours prior to initiation of measurements.
-
 \
 Prior to gas exchange measurements photosynthetic photon flux density (PPFD) was recorded at the individual leaf level and at the canopy position for each selected leaf.
-A hand-held photosynthetically available radiation meter was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either full sun or full shade.
+A hand-held quantum sensor () was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either full sun or full shade.
 A ceptometer (AccuPAR LP-80, Decagon Devices, Pullman, USA) was then used to measure 1 m integrated PPFD at the canopy height of each chosen leaf type.
@@
 Tree canopies may maintain the ability to upregulate A beyond the constraints of their local light environment, even at the cost of valuable resources.
-Empirical data accounting for the total conductance of CO2 (both g~s~ and g~m~) reveal that these responses can lead to leaf C gain equivalent to sun leaves and may happen quicker than previously thought.
+Empirical data accounting for the total conductance of CO~2~ reveal that shifts in g~s~ and g~m~ in shade leaves can lead to leaf C gain equivalent to sun leaves and may happen quicker than previously thought.
 Thus, failure to integrate the effects of g~m~ into process based tree growth models may underestimate the ability of canopies to alter their photosynthetic capacity when environmental conditions change.
@@
 Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979).
-As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vcmax and Jmax [@farquhar1980biochemical].
+As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical].
 Our data agree with these conventional conclusions as the distribution of N, both biochemical parameters and A were all reduced in shade leaves.
@@
 In this study, variation in leaf N distribution and A rates were not associated with subsequent changes in leaf K between sun and shade leaves.
-This is meaningful as foliage tends to operate at greater water limitations in the upper canopy, despite the ability of sun leaves to develop more effective vascular systems [@sellin2008effects;@niinemets2012optimization].
+This is especially meaningful in trees as foliage generally operates at greater water limitations in the upper canopy.
 As both sun and shade leaves were acclimated to their local light environment this sub-optimal distribution of water supply has the potential to diminish the C return for the whole canopy.
@@
 \
-The pattern of wasteful water use in shade leaves is important as we hypothesized that gs and A would be proportional across sun and shade leaves.
+The pattern of wasteful water use in shade leaves is important as we hypothesized that g~s~ and A would be proportional across sun and shade leaves.
 In sun leaves variation in A and g~s~ were strongly correlated, exhibiting behaviour agreeing with optimal stomatal theory.
-However, lower rates of A in shade leaves were not coupled with decreases in g~s~ which resulted in the observed decreases in WUE.
+However, lower rates of A in shade leaves were not coupled with decreases in g~s~, explaining the observed decreases in WUE.
 This is significant as optimal stomatal behaviour has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on sun leaves [see @prentice2014balancing; @lin2015optimal].
-As a result, the often used economic framework of balancing costs of using water versus N allocation to predict photosynthesis [@wright2003least] may break down when considering all leaves within a tree canopy.
+As a result, the often used economic framework of balancing costs of using water versus N allocation to predict A [@wright2003least] may break down when considering all leaves within a tree canopy.
 
@@
 These results should not be entirely unexpected as shade leaves experience interchanging periods of sun and shade, despite canopy light extinction [@pearcy1990sunflecks].
-Additionally, empirical studies that assess photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@leakey2002relative] and understory plants [@chazdon1991importance; @brantley2009contribution].
+Unfortunately, empirical studies that assess photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@leakey2002relative] and understory plants [@chazdon1991importance; @brantley2009contribution].
 Our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy.
@@
 Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of mesophyll conductance [@hanba2004overexpression; @heinen2009role; @li2014plant].
-This provides a potential explanation for our observed rapid increases in g~m~ but is untested in leaves of woody species.
+This provides a potential explanation for the observed rapid increases in g~m~ but is laregly untested in leaves of tree species.
 Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
@@
 \
-Similar to Laisk [-@laisk2005adjustment] we show evidence that optimal acclimation, in this case physiological adjustment of the drawdown of CO~2~ in shade leaves, is possibly directed towards occasional maximums of light availability over adaptations towards a low light environment.
-If shade leaves ‘lie in wait’ for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximise C gain, beyond conventional trade-offs associated with canopy resource distribution.
-If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour in shade leaves could be offset over the leaf lifespan.
-The potential benefits of C gain in leaves constrained by low light to levels equivalent to sun leaves may be large enough to tolerate excess losses from transpiration.
+Similar to Laisk [-@laisk2005adjustment] we provide evidence that optimal acclimation, in this case physiological adjustment of the drawdown of CO~2~ in shade leaves, is possibly directed towards occasional maximums of light availability over adaptations towards a low light environment.
+If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximise C gain, beyond conventional trade-offs associated with canopy resource distribution.
+If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan.
+For example, the potential benefits of C gain in leaves constrained by low light to levels equivalent to sun leaves may be large enough to tolerate excess losses from transpiration.
 However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.
@@
 The observed physiological behaviour leaves in the shade, at both low and high irradiance, suggests that current theories of leaf optimal behaviour are incomplete.
-Furthermore, gm must now be considered as a dynamic process in modelling whole canopy C gain.
-Additional empirical data are needed to determine both the mechanisms responsible for and the capacity of gm to rapidly increase CO2 drawdown.
+Furthermore, g~m~ must now be considered as a dynamic process in modelling whole canopy C gain.
+Additional empirical data are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown.
 To improve our ability to predict whole canopy C gain future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.

### 2015-08-03 manuscript/manuscript_wtc3.Rmd
more edits to results

@@
 Prior to gas exchange measurements photosynthetic photon flux density (PPFD) was recorded at the individual leaf level and at the canopy position for each selected leaf.
-A hand-held quantum sensor () was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either full sun or full shade.
+A hand-held quantum sensor (LI-COR, Lincoln, NE, USA) was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either full sun or full shade.
 A ceptometer (AccuPAR LP-80, Decagon Devices, Pullman, USA) was then used to measure 1 m integrated PPFD at the canopy height of each chosen leaf type.
@@
 where R~s~ is the isotopic ratio of the sample and R~stnd~ is the isotopic ratio of the standard Vienna Pee Dee Belemnite.
-Next, the observed discrimination ($\Delta$~obs~) is calculated from [@evans1986carbon]:
+Next, the observed discrimination ($\Delta$~obs~) is calculated from Evans et al.
+[-@evans1986carbon]:
 
@@
 
-where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b (2.9&permil;) is the fractionation from boundary layer diffusion and a is the fractionation due to diffusion in air~(4.4&permil;) [@evans1986carbon].
+where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ (2.9&permil;) is the fractionation from boundary layer diffusion and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon].
 C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~).
 In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulis* from [@crous2012light].
-E denotes the transpiration rate and $\text{g}^t_{ac}$ is the total conductance to CO~2~ diffusion to both the boundary layer and stomatal conductance **(von Caemmerer 1981).
 The ternary effect corrections (t) are desribed by:
@@
 $$t = \frac{(1+a')E}{2g^t_{ac}}$$
+
+where E denotes the transpiration rate and $\text{g}^t_{ac}$ is the total conductance to CO~2~ diffusion to both the boundary layer and stomatal conductance.
 
@@
 Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general].
-Confidence intervals (95 %) of mixed effect linear models of leaf photosynthesis as a functions of different physiological parameters were generated using bootstrapping methods with 999 simulations in the 'lme4' package [@bates2013lme4] in R.
+Confidence intervals (95 %) of mixed effect linear models of A as a function of different physiological parameters were generated using bootstrapping methods with 999 simulations in the 'lme4' package [@bates2013lme4] in R.
 For non-linear mixed models models confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package in R.
-Results were considered significant at P&leq;0.05.
+Results were considered significant at P &leq; 0.05.
 
@@
 This was the result of neither $\Psi_{pd}$, $\Psi_l$ (Table 1) nor E (Table 2) differing between leaf types.
-Leaf N, on an area basis, was significantly higher in sun leaves than in shade leaves (p < 0.001) by ca.
-20% (Table1).
+Leaf N, on an area basis, was approximately 20% higher in sun leaves than in shade leaves (Table1).
 Leaf mass per area was not different between leaf types.
@@
 ## Leaf photosynthesis rates and photosynthetic capacity
-Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environoment (P < 0.001, Table 2).
+Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environoment (Table 2).
 Following an increase in available light, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001).
 No effect of the warming treatment was detected on rates of A of sun leaves or shade leaves at low or high light.
-Photosynthesis across all leaf types and treatments did not vary according across the leaf temperatures measured in this study or time of year (Figure `r get1fig`a).
+Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
 
 \
-The photsynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (P < 0.001 & P = 0.02, respectively), as evident from AC~i~ curves from each chamber (Figure `r acifig`a).
+The photsynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (Table 1), as evident from AC~i~ curves from each chamber (Figure `r acifig`a).
 No effect of the warming treatment was detected on either parameter (Table 1).
@@
 The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
-Across leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a)
+Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a)
 
+\
 Leaf $\delta$^13^C, as an index of water use efficiencey [@marshall2007sources], significantly decreased from sun leaves to shade leaves by ca.
-2&permil; (p < 0.001, Table 1).
+2&permil; (Table 1).
 No effects of the warming treatment on leaf $\delta$^13^C were detected.
 Leaf $\delta$^13^C and N~area~ were postively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C and higher N investment in sun leaves.
-
-*leaf 13C with leaf N and what this stands for*
 
@@
 With increased light availability, both increases in A and g~s~ in shade leaves extened the relationship of A and g~s~ seen in sun leaves.
+
 \
-The observed carbob isotope discrimination was postively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`).
-Here, the carbon isotope discrimination associtaed with g~m~ accounted for the majority of $\Delta$~o~ (69.7%pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments.
-Mean g~m~ was significantly higher in sun compared to shade leaves (+27%), under their local light environment (P < 0.001, Table2).
+The observed carbon isotope discrimination was postively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`).
+Here, the carbon isotope discrimination associtaed with g~m~ accounted for the majority of $\Delta$~o~ (69.7&pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments.
+
+\
+Mean g~m~ was higher in sun compared to shade leaves (+27%), under their local light environment ( Table2).
 Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001).
 Relationships between g~m~ and A were nearly proportional within leaf types and light treatments (Figure `r gebarfig`b,c).
-Photosynthesis scaled postively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves in high light resulted in the highest rates of A (Figure `r agmgsfig`b).
-No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types or light treatments.
-Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figures `r get1fig`b and `r get2fig`b, respectively) and a slightly negative relationship with g~m~ and increasing leaf temperature.
+Photosynthesis scaled postively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b).
+No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types.
+Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figures `r get1fig`b and `r get2fig`b, respectively) and a slightly negative relationship with g~m~ and increasing leaf temperature was detected.
 

### 2015-08-03 manuscript/manuscript_wtc3.Rmd
finalized first draft

@@
 ---
-title: "Near instantaneous response of mesophyll conductance to light availability allows shade leaves to lie in wait"
+title: "Response of mesophyll conductance to light availability allows shade leaves to lie in wait"
 author: ""
@@
 trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (+3&deg;C) treatments, to empirically evaluate the distribution of resources and leaf physiological behaviour of sun and shade leaves.
-Our hypothesese are as follows:
+Our hypotheses are as follows:
 
@@
 Sun flecks were then simulated for shade leaves by increasing the chamber PPFD to the light environment of the full sun leaf in the same tree.
-The maximum sunfleck response was then recorded once shade leaves re-stablized in the leaf cuvette (ca.
+The maximum sunfleck response was then recorded once shade leaves re-stabilized in the leaf cuvette (ca.
 25 min).
@@
 
-where $\Delta$~o~ is the observed discrimation and  $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionataion if C~i~ = C~c~, g~m~, respiration and photorespiration, respectively.
+where $\Delta$~o~ is the observed discrimination and  $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, g~m~, respiration and photorespiration, respectively.
 The equations for each are as follows:
@@
 \begin{aligned}
-\Delta_i = &\frac{1}{1-t}a'+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}
+\Delta_i = &\frac{1}{1-t}a_i+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}
 \end{aligned}
@@
 In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulis* from [@crous2012light].
-The ternary effect corrections (t) are desribed by:
+The ternary effect corrections (t) are described by:
 
@@
 
+\
 The CO~2~ diffusion from the intercellular airspace to the chloroplast, g~m~, is given by its relationship to the CO~2~ assimilation rate (A) by:
@@
 where C~c~ is the chloroplast CO~2~ partial pressure.
-Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular arispaces to the site of carboxylation were then estimated using Equation 11.
+Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were then estimated using Equation 11.
 Examples of this approach to measure gas exchange and C isotope discrimination are presented in [@evans2013temperature].
@@
 Leaf level instantaneous transpiration efficiency (ITE) was calculated as leaf photosynthesis divided by transpiration.
-The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2012reconciling] using non-linear resgression, where K=0.5 [see @duursma2013near].
+The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2012reconciling] using non-linear regression, where K=0.5 [see @duursma2013near].
 
@@
 ## Leaf photosynthesis rates and photosynthetic capacity
-Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environoment (Table 2).
+Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environment (Table 2).
 Following an increase in available light, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001).
@@
 \
-The photsynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (Table 1), as evident from AC~i~ curves from each chamber (Figure `r acifig`a).
-No effect of the warming treatment was detected on either parameter (Table 1).
-Within individual chambers Vc~max~ was postively related to leaf N~area~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
-Additionally, leaf N~area~ was postively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).
-As Vc~max~ is condisdered implicit with g~m~ (results below) with conventional AC~i~ curves, a AC~c~ curves were also simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.
+The photosynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (Table 1), as evident from AC~i~ curves from each chamber (Figure `r acifig`a).
+Within leaf types, no effect of the warming treatment was detected on either parameter.
+Within individual chambers Vc~max~ was positively related to leaf N~area~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
+Additionally, leaf N~area~ was positively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).
+As Vc~max~ is considered implicit with g~m~ (results below) with conventional AC~i~ curves, a AC~c~ curves were also simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.
 When accounting for the effect of g~m~, photosynthetic capacity was still greater in sun leaves than shade leaves (Figure `r accfig`)
@@
 ## Leaf water use efficiency
-Leaf ITE was significantly greater in sun leaves than in shade leaves (P = 0.001, Table 2).
+Leaf ITE was significantly greater in sun leaves than in shade leaves at low light (P = 0.001).
 Following an increase in available light, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
-ITE in sun leaves was reduced in the warming treatment (P = 0.021) but no effect was noted in shade leaves with low or high light.
+ITE in sun leaves was reduced in the warming treatment but no effect was noted in shade leaves with low or high light (Table 2).
 For all leaf types and light treatments there was a strong response of ITE to VPD, and individual data points broadly corresponded to response curves from the optimal ITE model (Figure `r waterusefig`a).
@@
 \
-Leaf $\delta$^13^C, as an index of water use efficiencey [@marshall2007sources], significantly decreased from sun leaves to shade leaves by ca.
+Leaf $\delta$^13^C, as an index of water use efficiency [@marshall2007sources], significantly decreased from sun leaves to shade leaves by ca.
 2&permil; (Table 1).
 No effects of the warming treatment on leaf $\delta$^13^C were detected.
-Leaf $\delta$^13^C and N~area~ were postively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C and higher N investment in sun leaves.
+Leaf $\delta$^13^C and N~area~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C and higher N investment in sun leaves.
 
-## Stomatal conductance, mesophyll conductance and CO~2~ drawdown
-Mean g~s~ was significantly higher in shade compared to sun leaves (+18%), under their local light environment (P = 0.004, Table2).
-Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001, Figure `r gebarfig`a).
-In sun leaves, g~s~ and A exhibited optimal behavior with diminishing rates of A with highest values of g~s~ (Figure `r agmgsfig`a).
-However, in shade leaves this relationship was clearly uncoupled.
-With increased light availability, both increases in A and g~s~ in shade leaves extened the relationship of A and g~s~ seen in sun leaves.
+## Leaf physiology
+Mean g~s~ was significantly higher in shade compared to sun leaves (+18%) under their local light environment (P = 0.004).
+Increases in g~s~ led to increases in A in sun leaves, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
+Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves ( Figure `r gebarfig`a).
+With increased light availability, both increases in A and g~s~ in shade leaves exhibited similar behavior as observed in sun leaves.
 
 \
-The observed carbon isotope discrimination was postively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`).
-Here, the carbon isotope discrimination associtaed with g~m~ accounted for the majority of $\Delta$~o~ (69.7&pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments.
+The observed carbon isotope discrimination was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`).
+Here, the carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta$~o~ (69.7&pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments.
 
 \
-Mean g~m~ was higher in sun compared to shade leaves (+27%), under their local light environment ( Table2).
-Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001).
+Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
+Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (Table 2).
 Relationships between g~m~ and A were nearly proportional within leaf types and light treatments (Figure `r gebarfig`b,c).
-Photosynthesis scaled postively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b).
+Photosynthesis scaled positively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b).
 No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types.
@@
 As increases in A in shade leaves were associated with higher rates of g~s~ the C~i~ was similar than under low light but still greater than sun leaves.
-As the drop in CO~2~ for intecellular spaces to the cholorplast, C~i~-C~c~, measures the coordiation between g~m~ and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure `r ciccfig`b,c).
+As the drop in CO~2~ for intecellular spaces to the cholorplast, C~i~-C~c~, measures the coordination between g~m~ and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure `r ciccfig`b,c).
 This was a direct result of the increase in g~m~ and its subsequent direct effect on A, when compared to both sun and shade leaves that are acclimated to their local light environment.
-CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable acorss the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
+CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
 
@@
 
-## sub heading
+## Resource distribution and photosynthetic capacity
 The allocation of N constrains photosynthesis rates and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
@@
 
+## Leaf physiology of sun and shade leaves
 \
@@
 Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of mesophyll conductance [@hanba2004overexpression; @heinen2009role; @li2014plant].
-This provides a potential explanation for the observed rapid increases in g~m~ but is laregly untested in leaves of tree species.
+This provides a potential explanation for the observed rapid increases in g~m~ but is largely untested in leaves of tree species.
 Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
@@
 
+## Conclusions
 \
 Similar to Laisk [-@laisk2005adjustment] we provide evidence that optimal acclimation, in this case physiological adjustment of the drawdown of CO~2~ in shade leaves, is possibly directed towards occasional maximums of light availability over adaptations towards a low light environment.
-If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximise C gain, beyond conventional trade-offs associated with canopy resource distribution.
+If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution.
 If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan.
-For example, the potential benefits of C gain in leaves constrained by low light to levels equivalent to sun leaves may be large enough to tolerate excess losses from transpiration.
+For example, the potential C gain in leaves which experience high amounts of sunflecks may be large enough to tolerate excess losses from transpiration.
 However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.
-Thus, models which predict leaf photosynthesis from N distribution within a canopy will be flawed unless inclusion of canopy light extinction and the integrations of sunflecks on shade leaves are included [@de1997simple].
+Thus, models which predict leaf photosynthesis from N distribution within a canopy will be flawed unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].
 
@@
 The observed physiological behaviour leaves in the shade, at both low and high irradiance, suggests that current theories of leaf optimal behaviour are incomplete.
-Furthermore, g~m~ must now be considered as a dynamic process in modelling whole canopy C gain.
+Furthermore, g~m~ must now be considered as a dynamic process in tree growth models that cannot be simply parameterized.
 Additional empirical data are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown.
@@
 Leaf mass per area, leaf Nitrogen~area~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
-Values of Vc~max~ and J~max~ are treament mean(standard error) from AC~i~ curves measured in each chamber before the drought treatment.
+Values of Vc~max~ and J~max~ are treatment mean(standard error) from AC~i~ curves measured in each chamber at saturating light.
 Different letters represent significant differences between treatments.
@@
 Relationship between the observed discrimination of ^13^CO~2~ and measured C~i~/C~a~ for sun leaves and shade leaves at both low and high light.
-The solid line represents the the theortical line for C3 plants from [@evans1986carbon].
+The solid line represents the the theoretical line for C3 plants from [@evans1986carbon].
 
@@
 **Figure S1**.
-Reponse of photosynthesis (a), mesophyll conductance (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
-Shaded symbols represents each monthyl measurement campaign.
+Response of photosynthesis (a), mesophyll conductance (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
+Shaded symbols represents each monthly measurement campaign.
 
@@
 **Figure S2**.
-Reponse of vapor pressure deficit (a), stomatal conductance (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
-Shaded symbols represents each monthyl measurement campaign.
+Response of vapor pressure deficit (a), stomatal conductance (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
+Shaded symbols represents each monthly measurement campaign.
 
@@
 Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
-C~c~ values were predicited with mesophyll conductance, thus curves represent chloroplastic photosynthetic parameters at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+C~c~ values were predicted with mesophyll conductance, thus curves represent chloroplastic photosynthetic parameters at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
 

### 2015-08-03 master_scripts/resouces_sigletters.R

@@ -36,7 +36,7 @@ write.csv(Jmax_siglets2, "master_scripts/sigletters/slr_jmax.csv", row.names=FAL
 
 #2: Vcmax: no effect of ET
 
-vc_temp <- lme(Vcmax ~ leaf ,random=~1|chamber, data=aciparam)
+vc_temp <- lme(Vcmax ~ temp ,random=~1|chamber, data=aciparam)
   summary(vc_temp)
   anova(vc_temp)
   visreg(vc_temp)

### 2015-08-18 manuscript/manuscript_wtc3.Rmd
started corrections for gm manuscript
did some examples for sachin

@@
 ---
-title: "Response of mesophyll conductance to light availability allows shade leaves to lie in wait"
+title: "Fast response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks."
 author: ""
@@
 # Introduction
-The need to understand and predict carbon (C) uptake in forest ecosystems is crucially important for assessing the impacts of environmental change.
+Understanding and predict carbon (C) uptake in forest ecosystems is crucially important for assessing the impacts of environmental change.
 Specifically, this involves accurate upscaling of leaf level measurements to predict whole canopy function.
-Due to the costs and limitations of efficient light harvesting plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling problematic.
-Incident PPFD declines exponentially with increasing leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor], and photosynthesis (A) increases with irradiance from shade to sun until biochemical limitation occurs [@evans1995carbon].
+Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling problematic.
+Incident PPFD declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor] and photosynthesis (A) responds to this changing irradiance until until biochemical limitation occurs [@evans1995carbon].
 Consequently, leaves modify structural and functional properties to efficiently intercept variable light and enhance photosynthetic capacity [@vogelman1996leaves; @valladares2004photosynthetic].
@@
 The distribution of resources required for A, including nitrogen (N) and water, are also partially defined by canopy light gradients.
-As A has a saturating response with light and a dependence on N, leaf N should be distributed in the upper canopy to maximize canopy C gain [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
+As A has a saturating response with light and a dependence on N, it has been argued that leaf N should be proportional to PPFD along the canopy gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
 This can lead to changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity across leaf types [@evans2001photosynthetic].
-Sun leaves also experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
+Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
 Higher rates of A and stomatal conductance (g~s~) can only be sustained if the hydraulic conductance (K) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal].
-Optimal photosynthetic N investment in the upper canopy will be wasted if photosynthetic capacity exceeds hydraulic supply [@niinemets2012optimization], thus K should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
+Optimal photosynthetic N investment in the upper canopy will be wasted if photosynthetic capacity exceeds hydraulic supply [@niinemets2012optimization; @peltoniemi2012co], thus K should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
 
@@
 These theoretical distributions of N and water regulate leaf physiological traits that constrain A differently for sun and shade leaves.
-The balance between the costs and energetic benefits of these leaf traits affecting photosynthetic capacity create trade-offs impacting canopy C gain [@givnish1988adaptation].
-During A, CO~2~ diffuses from the atmosphere through stomata (g~s~), intercellular air space and into the chloroplast for fixation by Rubisco (mesophyll conductance, g~m~) [@tazoe2011using] and trade-offs exists that constrain each of these diffusion pathways.
+The balance between the costs and benefits of these leaf traits affecting photosynthetic capacity create trade-offs impacting canopy C gain [@givnish1988adaptation].
+During A, CO~2~ diffuses from the atmosphere through stomata (g~s~), intercellular air space and into the chloroplast for fixation by Rubisco (mesophyll conductance, g~m~) [@tazoe2011using].
+Trade-offs exists that constrain each of these diffusion pathways, although these are yet to be explictily quantified for g~m~.
 For optimal leaf C gain these pathways should act to minimize the resistance to CO~2~ diffusion while also limiting the intrinsic energy, water and construction costs which diminish the C return from A.
@@
 This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
-Second, variation in photosynthetic capacity between sun and shade leaves has also been hypothesized to be due to differences in g~m~ [@duursma2006vertical].
+Second, variation in photosynthetic capacity between sun and shade leaves has also been hypothesized to be due to differences in g~m~ [@piel2002effect; @duursma2006vertical; @warren2007internal].
 Mesophyll conductance can impose limitations on A as large as those associated with g~s~ [@warren2008stand; @ubierna2011estimation] and these limitations can reduce the efficiency of N use in A [@niinemets2007photosynthesis].
@@
 How shade leaves utilize sunflecks for short term C gain depends on the combined response time of g~s~ and g~m~ and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks].
-For example, at short timescales the utilization of sunflecks is limited by the slow response of stomatal opening, effectively limiting the maximum assimilation rate that can be achieved [@way2012sunflecks].
-Mesophyll conductance has been shown to respond to environmental factors across both long and rapid time scales, possibly faster than g~s~ [@flexas2008mesophyll].
+For example, at short timescales the utilization of sunflecks is limited by delayed responses of stomata, effectively limiting the maximum assimilation rate that can be achieved [@vico2011effects; @way2012sunflecks].
+Mesophyll conductance has been shown to respond to environmental factors across both short and long time scales, possibly faster than g~s~ [@flexas2008mesophyll].
 However, anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are unlikely to adapt during short light fluctuations.
@@
 \
-1.	If whole tree canopies are optimized for C gain, then leaf N, K and photosynthetic capacity were predicted to be higher in full sun leaves.
+1.	If whole tree canopies are optimized for C gain, then leaf N, K and photosynthetic capacity were predicted to be higher in sun leaves compared to shade leaves.
 
 \
-2.	Under comparable leaf VPD, g~s~ and A should be proportional across both sun and shade leaves, while g~m~ was predicted to scale positively with photosynthetic capacity.
+2.	Under comparable leaf VPD, g~s~ should be proportional to A across both sun and shade leaves and g~m~ should scale positively with photosynthetic capacity.
 
@@
 \
-4.	The effects of climate warming was predicted to be greater in sun than shade leaves, seen as increases in stomatal closure and reductions in leaf C gain during summer months.
+4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months.
 
@@
 Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced a constant elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarsfig`).
-Due to the temperature treatment air humidity was not explicitly controlled.
-The CO~2~ concentration inside the chamber was set to match outside air, tracking diurnal changes.
 Trees were watered weekly with 70 L from March 2013 to November 2013.
@@
 In February 2013 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering.
-Due to a limited range of data for the drought treatment only well-watered trees are reported here.
+Due to a limited range of data for the drought treatment only well-watered trees are reported, which reduces the sample size of WTC (n=6) after the initiation of this treatment.
 
@@
 The top soils at this site, used in both pots and chambers, are an alluvial formation of low-fertility sandy loam soils with low organic matter and low water holding capacity.
+A root exclusion barrier extended from chamber walls to the hard layer (ca.
+1 m) and roots were allowed to grow freely below the barrier.
 After 2 months, floors were installed 45 cm from the soil surface.
-This enabled chamber fluxes of CO~2~ and H~2~O from the whole tree canopy to be monitored.
-Gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.
+This enabled chamber fluxes of CO~2~ and H~2~O from the whole tree canopy to be monitored (for methods see Barton et al.
+[-@barton2010whole]).
+Leaf gas exchange measurements were then initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.
 At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm, height of 348&pm;15.1 cm and an estimated leaf area of 3.9&pm;0.1 m^2^.
@@
 The nearest leaf on each branch was sampled for measurement of predawn leaf water potential.
+
 \
@@
 All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
-Leaf water potential and transpiration (E, mmol^-2^ s^-1^) from gas exchange were then used to calculate leaf specific hydraulic conductance (K) through the equation:
+Leaf water potential and transpiration (E, mmol^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K) through the equation:
 
@@
 where the term Hp~w~g represents the gravitational potential assumed to be minimal in this experiment due to small tree size [@whitehead1998regulation].
-Leaf level instantaneous transpiration efficiency (ITE) was calculated as leaf photosynthesis divided by transpiration.
+Leaf level instantaneous water use efficiency (WUE~i~) was calculated as leaf photosynthesis divided by transpiration.
 The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2012reconciling] using non-linear regression, where K=0.5 [see @duursma2013near].
@@
 Confidence intervals (95 %) of mixed effect linear models of A as a function of different physiological parameters were generated using bootstrapping methods with 999 simulations in the 'lme4' package [@bates2013lme4] in R.
-For non-linear mixed models models confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package in R.
+For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package in R, using WTC as a random effect.
 Results were considered significant at P &leq; 0.05.
@@
 Across all measurement campaigns 1 m canopy integrated PPFD was significantly different between sun and shade leaves (P<0.001) and PPFD was reduced by 78% in the shade (Figure `r parfig`).
-The distribution of H~2~O, via K, was not different across sun and shade leaves (Table 1).
+Leaf-specific hydraulic conductance was not different across sun and shade leaves (Table 1).
 This was the result of neither $\Psi_{pd}$, $\Psi_l$ (Table 1) nor E (Table 2) differing between leaf types.
@@
 
-## Leaf photosynthesis rates and photosynthetic capacity
-Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environment (Table 2).
+## Photosynthetic capacity and leaf photosynthesis rates 
+The photosynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (Table 1), as evident from AC~i~ curves from each chamber (Figure `r acifig`a).
+Within leaf types, no effect of the warming treatment was detected on either parameter.
+Within individual chambers Vc~max~ was positively related to leaf N~area~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
+Additionally, leaf N~area~ was positively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).
+As g~m~ (results below) is included in the estimates of J~max~ and Vc~max~ with conventional AC~i~, AC~c~ curves were also simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.
+When accounting for the effect of g~m~, chloroplastic Vc~max~ and J~max~ were still higher in sun leaves (Figure `r accfig`).
+
+\
+Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table 2).
 Following an increase in available light, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001).
@@
 
-\
-The photosynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (Table 1), as evident from AC~i~ curves from each chamber (Figure `r acifig`a).
-Within leaf types, no effect of the warming treatment was detected on either parameter.
-Within individual chambers Vc~max~ was positively related to leaf N~area~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
-Additionally, leaf N~area~ was positively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).
-As Vc~max~ is considered implicit with g~m~ (results below) with conventional AC~i~ curves, a AC~c~ curves were also simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.
-When accounting for the effect of g~m~, photosynthetic capacity was still greater in sun leaves than shade leaves (Figure `r accfig`)
-
 ## Leaf water use efficiency
-Leaf ITE was significantly greater in sun leaves than in shade leaves at low light (P = 0.001).
-Following an increase in available light, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
-ITE in sun leaves was reduced in the warming treatment but no effect was noted in shade leaves with low or high light (Table 2).
-For all leaf types and light treatments there was a strong response of ITE to VPD, and individual data points broadly corresponded to response curves from the optimal ITE model (Figure `r waterusefig`a).
+Leaf WUE~i~ was significantly greater in sun leaves than in shade leaves at low light (P = 0.001).
+Following an increase in available light, WUE~i~ of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
+WUE~i~ in sun leaves was reduced in the warming treatment but no effect was noted in shade leaves with low or high light (Table 2).
+For all leaf types and light treatments there was a strong response of WUE~i~ to VPD, and individual data points broadly corresponded to response curves from the optimal WUE~i~ model (Figure `r waterusefig`a).
 The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
@@
 \
-Leaf $\delta$^13^C, as an index of water use efficiency [@marshall2007sources], significantly decreased from sun leaves to shade leaves by ca.
+Leaf $\delta$^13^C, as an index of integrated water use efficiency [@marshall2007sources], significantly decreased from sun leaves to shade leaves by ca.
 2&permil; (Table 1).
@@
 Mean g~s~ was significantly higher in shade compared to sun leaves (+18%) under their local light environment (P = 0.004).
-Increases in g~s~ led to increases in A in sun leaves, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
+Increases in g~s~ were postively correlated to increases in A in sun leaves, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
 Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves ( Figure `r gebarfig`a).
@@
 Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (Table 2).
-Relationships between g~m~ and A were nearly proportional within leaf types and light treatments (Figure `r gebarfig`b,c).
+Mesophyll conductance was nearly proportional to A within leaf types and light treatments (Figure `r gebarfig`b,c).
 Photosynthesis scaled positively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b).
 No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types.
-Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figures `r get1fig`b and `r get2fig`b, respectively) and a slightly negative relationship with g~m~ and increasing leaf temperature was detected.
+Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figure `r get1fig`b and `r get2fig`b, respectively) and only a weak negative relationship with g~m~ and increasing leaf temperature was detected (Figure `r get1fig`b).
 
@@
 As increases in A in shade leaves were associated with higher rates of g~s~ the C~i~ was similar than under low light but still greater than sun leaves.
-As the drop in CO~2~ for intecellular spaces to the cholorplast, C~i~-C~c~, measures the coordination between g~m~ and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure `r ciccfig`b,c).
+As the drop in CO~2~ for intercellular spaces to the cholorplast, C~i~-C~c~, measures the coordination between g~m~ and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure `r ciccfig`b,c).
 This was a direct result of the increase in g~m~ and its subsequent direct effect on A, when compared to both sun and shade leaves that are acclimated to their local light environment.
@@
 ## Resource distribution and photosynthetic capacity
-The allocation of N constrains photosynthesis rates and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
+The allocation of N~a~ constrains photosynthesis rates and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
 Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979).
 As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical].
-Our data agree with these conventional conclusions as the distribution of N, both biochemical parameters and A were all reduced in shade leaves.
+Our data agree with these conventional conclusions as the distribution of N~a~ , both biochemical parameters and A were all reduced in shade leaves.
 
@@
 It is possible that sustaining higher g~s~, at the cost of larger transpiration losses, is a strategy to increase the utilization efficiency of sunflecks.
-These results should not be entirely unexpected as shade leaves experience interchanging periods of sun and shade, despite canopy light extinction [@pearcy1990sunflecks].
+These results should not be entirely unexpected as shade leaves experience interchanging periods of sun and shade, despite lower mean PPFD [@pearcy1990sunflecks].
 Unfortunately, empirical studies that assess photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@leakey2002relative] and understory plants [@chazdon1991importance; @brantley2009contribution].

### 2015-08-19 manuscript/manuscript_wtc3.Rmd
made change
manuscript edits
made changes to sup figure alpha of points

added inset to acc plot

add more cites

@@
 acifig <- "3"
-waterusefig <- "4"
+agmgsfig <- "4"
 gebarfig <- "5"
-agmgsfig <- "6"
+waterusefig <- "6"
 deltafig <- "7"
@@
 
-get1fig <- "S1"
-get2fig <- "S2"
-accfig <- "S3"
+accfig <- "S1"
+get1fig <- "S2"
+get2fig <- "S3"
+
 ```
@@
 where the term Hp~w~g represents the gravitational potential assumed to be minimal in this experiment due to small tree size [@whitehead1998regulation].
-Leaf level instantaneous water use efficiency (WUE~i~) was calculated as leaf photosynthesis divided by transpiration.
+Leaf level instantaneous water use efficiency (~i~WUE) was calculated as leaf photosynthesis divided by transpiration.
 The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2012reconciling] using non-linear regression, where K=0.5 [see @duursma2013near].
@@
 ## Leaf resource distribution
-Across all measurement campaigns 1 m canopy integrated PPFD was significantly different between sun and shade leaves (P<0.001) and PPFD was reduced by 78% in the shade (Figure `r parfig`).
-Leaf-specific hydraulic conductance was not different across sun and shade leaves (Table 1).
-This was the result of neither $\Psi_{pd}$, $\Psi_l$ (Table 1) nor E (Table 2) differing between leaf types.
-Leaf N, on an area basis, was approximately 20% higher in sun leaves than in shade leaves (Table1).
+Across all measurement campaigns PPFD was significantly different between sun and shade leaves (P<0.001) and PPFD was reduced by 78% in the shade (Figure `r parfig`).
+Leaf-specific hydraulic conductance was similar across sun and shade leaves (Table 1).
+This was because $\Psi_{pd}$, $\Psi_l$ (Table 1) and E (Table 2) did not differ between leaf types.
+Leaf N on an area basis (N~a~) was approximately 20% higher in sun leaves than in shade leaves (Table1).
 Leaf mass per area was not different between leaf types.
@@
 Within leaf types, no effect of the warming treatment was detected on either parameter.
-Within individual chambers Vc~max~ was positively related to leaf N~area~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
-Additionally, leaf N~area~ was positively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).
+Within individual chambers Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
 As g~m~ (results below) is included in the estimates of J~max~ and Vc~max~ with conventional AC~i~, AC~c~ curves were also simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.
@@
 Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table 2).
+Additionally, leaf N~area~ was positively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).
 Following an increase in available light, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001).
@@
 
-## Leaf water use efficiency
-Leaf WUE~i~ was significantly greater in sun leaves than in shade leaves at low light (P = 0.001).
-Following an increase in available light, WUE~i~ of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
-WUE~i~ in sun leaves was reduced in the warming treatment but no effect was noted in shade leaves with low or high light (Table 2).
-For all leaf types and light treatments there was a strong response of WUE~i~ to VPD, and individual data points broadly corresponded to response curves from the optimal WUE~i~ model (Figure `r waterusefig`a).
+## Stomatal conductance and leaf water use efficiency
+Mean g~s~ was significantly higher in shade compared to sun leaves (+18%) under their local light environment (P = 0.004).
+Increases in g~s~ were postively correlated to increases in A in sun leaves, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
+Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves ( Figure `r gebarfig`a).
+With increased light availability, both increases in A and g~s~ in shade leaves exhibited similar behavior as observed in sun leaves.
+
+\
+Leaf ~i~WUE was significantly greater in sun leaves than in shade leaves at low light (P = 0.001).
+Following an increase in available light, ~i~WUE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
+~i~WUE in sun leaves was reduced in the warming treatment but no effect was noted in shade leaves with low or high light (Table 2).
+For all leaf types and light treatments there was a strong response of ~i~WUE to VPD, and individual data points broadly corresponded to response curves from the optimal ~i~WUE model (Figure `r waterusefig`a).
 The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
@@
 
-## Leaf physiology
-Mean g~s~ was significantly higher in shade compared to sun leaves (+18%) under their local light environment (P = 0.004).
-Increases in g~s~ were postively correlated to increases in A in sun leaves, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
-Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves ( Figure `r gebarfig`a).
-With increased light availability, both increases in A and g~s~ in shade leaves exhibited similar behavior as observed in sun leaves.
-
-\
+## Leaf carbon isotope discrimination and mesophyll conductance
 The observed carbon isotope discrimination was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`).
@@
 
-\
-The combination of a non-limiting supply of leaf K and more open stomata lead to a significant increase in C~i~ in shade compared to sun leaves (Figure `r ciccfig`a).
+## Variation in CO~2~ drawdown
+More open stomata in shade leaves, possible as a result of a non-limiting supply of leaf K, lead to a significant increase in C~i~ compared to sun leaves (Figure `r ciccfig`a).
 As increases in A in shade leaves were associated with higher rates of g~s~ the C~i~ was similar than under low light but still greater than sun leaves.
@@
 Photosynthesis in trees is also limited by the ability to supply water to the upper canopy.
+Ultimately, the ability of a trees hydraulic architecture to supply water to foliage across increasing pathlengths affects productivity and survival [@sellin2008effects].
 Using a two-leaf model Peltoniemi et al.
@@
 In this study, variation in leaf N distribution and A rates were not associated with subsequent changes in leaf K between sun and shade leaves.
-This is especially meaningful in trees as foliage generally operates at greater water limitations in the upper canopy.
 As both sun and shade leaves were acclimated to their local light environment this sub-optimal distribution of water supply has the potential to diminish the C return for the whole canopy.
@@
 \
-The inefficient distribution of resources led to decreased instantaneous WUE in shade leaves throughout the experiment.
+Unexpected higher rates of g~s~ led to decreased instantaneous WUE in shade leaves throughout the experiment.
 Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was constant across leaf lifespans.
-From a canopy perspective this pattern in WUE appears to be detrimental to C gain as A in sun leaves was characterized by low rates of g~s~ and low C~i~ compared to similar Eucalyptus species [see @drake2014capacity].
+From a canopy perspective this pattern in WUE initially appears to be detrimental to C gain as A in sun leaves was characterized by low rates of g~s~ and low C~i~.
 As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower WUE while the photosynthetic efficiency of sun leaves appears at least partially constrained by water supply.
@@
 \
-This behaviour may be explained by the response of shade leaves to increases in light availability.
-Contrary to hypothesis, shade leaves were able to increase A to rates equivalent or even outperforming sun leaves under identical light intensity.
 It is possible that sustaining higher g~s~, at the cost of larger transpiration losses, is a strategy to increase the utilization efficiency of sunflecks.
-These results should not be entirely unexpected as shade leaves experience interchanging periods of sun and shade, despite lower mean PPFD [@pearcy1990sunflecks].
+Rapid increases in leaf temperature with sunflecks has the potential to inhibit C gain [@leakey2003high].
+It has been suggested that transpiration-induced cooling is effective in avoiding sunfleck induced heat damage, however leaves need to keep stomata open in the shade for this to occur [@schymanski2013stomatal].
+Here, shade leaves were able to increase A to rates equivalent or even outperforming sun leaves under identical light intensity.
+Whether more open stomata is a strategy to reduce stomatal response time or to decrease thermal load on shade leaves in trees is still unknown.
 Unfortunately, empirical studies that assess photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@leakey2002relative] and understory plants [@chazdon1991importance; @brantley2009contribution].
-Our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy.
+Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy (see De Pury & Farquhar, [-@de1997simple]).
 
@@
 Furthermore, our findings have important implications for process based models that predict canopy C gain from rates of leaf photosynthesis.
-Although the distributions of resources, such as N and water, define leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
-The observed physiological behaviour leaves in the shade, at both low and high irradiance, suggests that current theories of leaf optimal behaviour are incomplete.
+Although the distribution of resources, such as N and water, define leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
+
+The observed physiological behaviour of leaves in the shade, at both low and high irradiance, suggests that current theories of leaf optimal behaviour are incomplete.
 Furthermore, g~m~ must now be considered as a dynamic process in tree growth models that cannot be simply parameterized.
-Additional empirical data are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown.
+Additional empirical data across multiple tree species are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown.
 To improve our ability to predict whole canopy C gain future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.
+
+These results should not be entirely unexpected as shade leaves experience interchanging periods of sun and shade, despite lower mean PPFD [@pearcy1990sunflecks].
 
@@
 *Eucalyptus tereticornis* leaf traits and the distribution of resources between full sun and shade leaves under ambient and elevated temperature treatments.
-Leaf mass per area, leaf Nitrogen~area~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
+Leaf mass per area, leaf N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
 Values of Vc~max~ and J~max~ are treatment mean(standard error) from AC~i~ curves measured in each chamber at saturating light.
@@
 **Figure 4**.
-(a) Response of instantaneous water use efficiency (ITE) to leaf vapor pressure deficit (VPD) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
-(b) The relationship between leaf $\delta$^13^C and leaf nitrogen~area~ for sun leaves and shade leaves at low light.
-For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g1 value for each leaf type and treatment.
-For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional r^2^ of 0.41 and 0.45, respectively.
+The response of leaf photosynthesis rates to stomatal conductance (a) and mesophyll conductance (b) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
+The response of shade leaf physiology to high light was recorded once both photosynthesis and stomatal conductance were stable in the gas exchange leaf cuvette (ca.
+25min).
+Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
+Grey areas are approximately 95% confidence intervals from the mean.
 
@@
 **Figure 6**.
-The response of leaf photosynthesis rates to stomatal conductance (a) and mesophyll conductance (b) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
-The response of shade leaf physiology to high light was recorded once both photosynthesis and stomatal conductance were stable in the gas exchange leaf cuvette (ca.
-25min).
-Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
-Grey areas are approximately 95% confidence intervals from the mean.
+(a) Response of instantaneous water use efficiency (ITE) to leaf vapor pressure deficit (VPD) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
+(b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves and shade leaves at low light.
+For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g1 value for each leaf type and treatment.
+For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional r^2^ of 0.41 and 0.45, respectively.
 
@@
 **Figure S1**.
+Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
+C~c~ values were predicted with mesophyll conductance, thus curves represent chloroplastic photosynthetic parameters at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+
+\
+**Figure S2**.
 Response of photosynthesis (a), mesophyll conductance (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
@@
 \
-**Figure S2**.
+**Figure S3**.
 Response of vapor pressure deficit (a), stomatal conductance (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
 Shaded symbols represents each monthly measurement campaign.
-
-\
-**Figure S3**.
-Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
-C~c~ values were predicted with mesophyll conductance, thus curves represent chloroplastic photosynthetic parameters at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
 
@@
 
-```{r wateruse, fig.width=6, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
-source("master_scripts/wateruse.R")
+```{r Agmgs, fig.width=6, fig.height=8, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/Agmgs.R")
 ```
@@
 
-```{r Agmgs, fig.width=6, fig.height=8, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/Agmgs.R")
+```{r wateruse, fig.width=6, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
+source("master_scripts/wateruse.R")
 ```
@@
 # Supporting Information {.unnumbered}
-```{r gasex_temp1, fig.width=6, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/gasex_temp.R")
-```  
-**Figure S1**.
-
-```{r gasex_temp2, fig.width=6, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/gasex_temp2.R")
-```  
-**Figure S2**.
 
@@
 
+**Figure S1**.
+
+```{r gasex_temp1, fig.width=6, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/gasex_temp.R")
+```
+
+**Figure S2**.
+
+```{r gasex_temp2, fig.width=6, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/gasex_temp2.R")
+```
+
 **Figure S3**.

### 2015-08-20 manuscript/manuscript_wtc3.Rmd
edits to manuscript
deleted some test files on repo

small edits to plots

@@
 
-# Summary {.unnumbered}
+# Summary
+* Light gradients within tree canopies play a major role in the distribution of resources that define photosynthetic capacity of individual leaves.
+The transient nature of light availability within a canopy, however, complicates our ability to assess the contribution of shade leaves to canopy carbon gain.
+* Leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis, stomatal conductance and mesophyll conductance of sun and shade leaves in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
+The physiological behaviour of individual leaves was then correlated with the distribution of nitrogen and water within the canopy to evaluate how trees are optimized for carbon gain.
+* More open stomata and rapid increases in mesophyll conductance allowed shade leaves to "lie in wait" for sunflecks and lead to leaf carbon gain exceeding sun leaves.
+* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
+Evidence that mesophyll conductance not only varies within a canopy but can be up-regulated over short time intervals must now be considered in process based tree growth models.
 
-* overview
-* method
-* big result
-* implication
+# Key words
+photosynthesis, stomatal conductance, mesophyll conductance,  shade, leaf optimal behavior
 
@@
 Mesophyll conductance can impose limitations on A as large as those associated with g~s~ [@warren2008stand; @ubierna2011estimation] and these limitations can reduce the efficiency of N use in A [@niinemets2007photosynthesis].
-Interpreting the coupled responses of both g~s~ and g~m~ to A across sun and shade leaves has major implications for predicting canopy C gain, but empirical measurements across tree canopies are still lacking.
+Interpreting the coupled responses of both g~s~ and g~m~ to A across sun and shade leaves is critical for predicting canopy C gain, but empirical measurements across tree canopies are still lacking.
 
@@
 \
-2.	Under comparable leaf VPD, g~s~ should be proportional to A across both sun and shade leaves and g~m~ should scale positively with photosynthetic capacity.
+2.	Stomatal conductance should be proportional to A across sun and shade leaves under similar leaf VPD and g~m~ should scale positively with photosynthetic capacity.
 
@@
 [-@barton2010whole]).
-Leaf gas exchange measurements were then initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.
+Leaf gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.
 At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm, height of 348&pm;15.1 cm and an estimated leaf area of 3.9&pm;0.1 m^2^.
@@
 The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDLAS on each measurement cycle.
-Photosynthesis, g~s~, transpiration, vapor pressure deficit and intercellular [CO~2~] were auto logged every 15 s for each gas exchange system over the same 12 min interval.
+Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] were auto logged every 15 s for each gas exchange system over the same 12 min interval.
 
@@
 
-where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ (2.9&permil;) is the fractionation from boundary layer diffusion and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon].
+where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ is the fractionation from boundary layer diffusion (2.9&permil;) and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon].
 C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~).
@@
 Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011].
+From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@Duursma2014] in R [@RDevelopmentCoreTeam2011].
 
@@
 where the term Hp~w~g represents the gravitational potential assumed to be minimal in this experiment due to small tree size [@whitehead1998regulation].
-Leaf level instantaneous water use efficiency (~i~WUE) was calculated as leaf photosynthesis divided by transpiration.
+Leaf level instantaneous water use efficiency (WUE~i~) was calculated as leaf photosynthesis divided by transpiration.
 The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2012reconciling] using non-linear regression, where K=0.5 [see @duursma2013near].
@@
 Within individual chambers Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
-As g~m~ (results below) is included in the estimates of J~max~ and Vc~max~ with conventional AC~i~, AC~c~ curves were also simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.
+As g~m~ (results below) is included in the estimates of J~max~ and Vc~max~ with conventional AC~i~ curves, AC~c~ curves were simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.
 When accounting for the effect of g~m~, chloroplastic Vc~max~ and J~max~ were still higher in sun leaves (Figure `r accfig`).
@@
 Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves ( Figure `r gebarfig`a).
-With increased light availability, both increases in A and g~s~ in shade leaves exhibited similar behavior as observed in sun leaves.
+With increased light availability, increases in A and g~s~ in shade leaves exhibited similar behavior as observed in sun leaves.
 
 \
-Leaf ~i~WUE was significantly greater in sun leaves than in shade leaves at low light (P = 0.001).
-Following an increase in available light, ~i~WUE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
-~i~WUE in sun leaves was reduced in the warming treatment but no effect was noted in shade leaves with low or high light (Table 2).
-For all leaf types and light treatments there was a strong response of ~i~WUE to VPD, and individual data points broadly corresponded to response curves from the optimal ~i~WUE model (Figure `r waterusefig`a).
+Leaf WUE~i~ was significantly greater in sun leaves than in shade leaves at low light (P = 0.001).
+Following an increase in available light, WUE~i~ of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
+WUE~i~ in sun leaves was reduced in the warming treatment but no effect was noted in shade leaves with low or high light (Table 2).
+For all leaf types and light treatments there was a strong response of WUE~i~ to VPD, and individual data points broadly corresponded to response curves from the optimal WUE model (Figure `r waterusefig`a).
 The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
@@
 The observed carbon isotope discrimination was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`).
-Here, the carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta$~o~ (69.7&pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments.
+Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta$~o~ (69.7&pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments.
 
@@
 
-## Variation in CO~2~ drawdown
-More open stomata in shade leaves, possible as a result of a non-limiting supply of leaf K, lead to a significant increase in C~i~ compared to sun leaves (Figure `r ciccfig`a).
-As increases in A in shade leaves were associated with higher rates of g~s~ the C~i~ was similar than under low light but still greater than sun leaves.
-As the drop in CO~2~ for intercellular spaces to the cholorplast, C~i~-C~c~, measures the coordination between g~m~ and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure `r ciccfig`b,c).
-This was a direct result of the increase in g~m~ and its subsequent direct effect on A, when compared to both sun and shade leaves that are acclimated to their local light environment.
-CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
+## Variation in intercellular and chloroplastic CO~2~ concentrations
+More open stomata in shade leaves under low and high light lead to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
+The drawdown of CO~2~ from intercellular spaces to the cholorplast, C~i~-C~c~, measures the coordination between g~m~ and A [@caemmerer2014temperature].
+This drawdown was similar between sun and shade leaves at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
+This was the result of the proportional relationship between g~m~ and A across all leaves.
+The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
 
 # Discussion
-Here we show that dynamic physiological responses in shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy.
-Tree canopies may maintain the ability to upregulate A beyond the constraints of their local light environment, even at the cost of valuable resources.
-Empirical data accounting for the total conductance of CO~2~ reveal that shifts in g~s~ and g~m~ in shade leaves can lead to leaf C gain equivalent to sun leaves and may happen quicker than previously thought.
-Thus, failure to integrate the effects of g~m~ into process based tree growth models may underestimate the ability of canopies to alter their photosynthetic capacity when environmental conditions change.
+Here we show that leaves within canopies of *Eucalyptus tereticornis* trees maintain the ability to upregulate A beyond the limitations of their local light environment.
+More open stomata and rapid increases in g~m~ allowed shade leaves to readily utilize increases in light availability, leading to leaf C gain exceeding sun leaves.
+Although these results should not be entirely unexpected as shade leaves experience interchanging periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies still impedes our ability to predict whole canopy C gain.
+Failure to integrate this dynamic physiology into process based tree growth models may underestimate the ability of canopies to alter leaf C gain when environmental conditions change.
 Additionally, with measurements recorded across a large natural range of temperatures we were unable to detect any effects of a +3 &deg;C warming treatment on leaf physiology.
@@
 \
-Unexpected higher rates of g~s~ led to decreased instantaneous WUE in shade leaves throughout the experiment.
-Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was constant across leaf lifespans.
+Unexpected higher rates of g~s~ led to decreased WUE~i~ in shade leaves throughout the experiment.
+Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was constant across the leaf lifespan.
 From a canopy perspective this pattern in WUE initially appears to be detrimental to C gain as A in sun leaves was characterized by low rates of g~s~ and low C~i~.
-As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower WUE while the photosynthetic efficiency of sun leaves appears at least partially constrained by water supply.
+As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower WUE when the photosynthetic efficiency of sun leaves appears at least partially constrained by water supply.
 
-## Leaf physiology of sun and shade leaves
-\
+## Physiological behaviour of sun and shade leaves
 The pattern of wasteful water use in shade leaves is important as we hypothesized that g~s~ and A would be proportional across sun and shade leaves.
@@
 It has been suggested that transpiration-induced cooling is effective in avoiding sunfleck induced heat damage, however leaves need to keep stomata open in the shade for this to occur [@schymanski2013stomatal].
-Here, shade leaves were able to increase A to rates equivalent or even outperforming sun leaves under identical light intensity.
+Here, shade leaves were able to increase A equivalent or even outperforming sun leaves under identical light intensity.
 Whether more open stomata is a strategy to reduce stomatal response time or to decrease thermal load on shade leaves in trees is still unknown.
 Unfortunately, empirical studies that assess photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@leakey2002relative] and understory plants [@chazdon1991importance; @brantley2009contribution].
-Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy (see De Pury & Farquhar, [-@de1997simple]).
+Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy [see @de1997simple].
 
@@
 In this study A and g~m~ scaled positively across leaf types and increased proportionately with increased light intensity in shade leaves.
-Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of mesophyll conductance [@hanba2004overexpression; @heinen2009role; @li2014plant].
+Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
 This provides a potential explanation for the observed rapid increases in g~m~ but is largely untested in leaves of tree species.
@@
 
-## Conclusions
 \
-Similar to Laisk [-@laisk2005adjustment] we provide evidence that optimal acclimation, in this case physiological adjustment of the drawdown of CO~2~ in shade leaves, is possibly directed towards occasional maximums of light availability over adaptations towards a low light environment.
+Similar to Laisk [-@laisk2005adjustment] we provide evidence that optimal acclimation, in this case the physiological adjustment of CO~2~ drawdown into shade leaves, is possibly directed towards occasional maximums of light availability over adaptation to a low light environment.
 If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution.
@@
 
-\
-Furthermore, our findings have important implications for process based models that predict canopy C gain from rates of leaf photosynthesis.
+## Conclusions
+Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy.
 Although the distribution of resources, such as N and water, define leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
-
-The observed physiological behaviour of leaves in the shade, at both low and high irradiance, suggests that current theories of leaf optimal behaviour are incomplete.
+These findings suggest that current theories of leaf optimal behaviour may be incomplete and have important implications for process based models that predict canopy C gain from rates of leaf photosynthesis.
 Furthermore, g~m~ must now be considered as a dynamic process in tree growth models that cannot be simply parameterized.
@@
 
-These results should not be entirely unexpected as shade leaves experience interchanging periods of sun and shade, despite lower mean PPFD [@pearcy1990sunflecks].
-
 # List of Tables
-
 **Table 1**.
@@
 Leaf mass per area, leaf N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
-Values of Vc~max~ and J~max~ are treatment mean(standard error) from AC~i~ curves measured in each chamber at saturating light.
+Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
 Different letters represent significant differences between treatments.
@@
 Each value reflects the mean (&pm; 1 standard error) for each treatment across all gas exchange campaigns (n=6).
-Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD~leaf~ are kPa.
+Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa.
 Different letters represent significant differences between treatments.
@@
 # List of Figures
-
 **Figure 1**. Daily maximum and minimum temperature (a) and total daily PPFD (b) for each chamber across the experiment duration.
@@
 Local light environment for sun and shade leaves for each gas exchange campaign.
-Means with standard errors represent 1 m integrated PPFD (n=5), measured with a ceptometer, at the canopy height of each selected leaf.
+Means &pm; 1 standard error represent 1 m integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf.
 
@@
 **Figure 8**.
-The mean &pm; 1 standard error of intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light with standard error
+The mean &pm; 1 standard error of intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light.
 
@@
 # Tables {.unnumbered}
-
 **Table 1**.
@@
 # Figures {.unnumbered}
-```{r airvars, fig.width=6, fig.height=8, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r airvars, fig.width=5, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
 source("master_scripts/air_vars.R")
@@
 
-```{r acicurves, fig.width=6, fig.height=8, warning=FALSE,echo=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r acicurves, fig.width=8, fig.height=8, warning=FALSE,echo=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
 source("master_scripts/aci_vcmaxnitro.R")
@@
 
-```{r Agmgs, fig.width=6, fig.height=8, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r Agmgs, fig.width=5, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
 source("master_scripts/Agmgs.R")
@@
 
-```{r wateruse, fig.width=6, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
+```{r wateruse, fig.width=5, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
 source("master_scripts/wateruse.R")

### 2015-08-20 master_scripts/cicc_bar.R

@@ -11,6 +11,7 @@ gasex <- read.csv("calculated_data/gmes_wellwatered.csv")
   gasex$cc_ci<- with(gasex, Ci/Cc)
   ##calculate drawdown
   gasex$drawdown <- with(gasex, Ci-Cc)
+  
 
 ###get average by id
 gasex_agg <- summaryBy(Ci+Cc+drawdown~ chamber+id+leaf +light+temp+leaflight+Month, data=gasex, FUN=mean, keep.names=TRUE)
@@ -22,6 +23,15 @@ gasex_agg <- summaryBy(Ci+Cc+drawdown~ chamber+id+leaf +light+temp+leaflight+Mon
 leaforder <- c("Shade", "Sun", "Sunfleck")  
 gasex_agg$leaflight <- factor(gasex_agg$leaflight, levels = leaforder)
 
+
+# library(visreg)
+# library(multcomp)
+# library(nlme)
+# drawdown_mod <- lme(drawdown~ leaflight, random=~1|chamber, data=gasex_agg)
+# summary(drawdown_mod)
+# anova(drawdown_mod)  
+# visreg(drawdown_mod)
+
 #lightson <- gasex_agg[gasex_agg$leaflight != "sun-high",]
 
 cols <- c(shacol2, suncol2, lightscol2)
@@ -30,7 +40,7 @@ cols <- c(shacol2, suncol2, lightscol2)
 
 #1: CI
 
-# windows(12,6)
+#windows(12,6)
 
 par(mfrow=c(1,3))
 

### 2015-08-20 master_scripts/gasex_temp2.R

@@ -1,4 +1,3 @@
-# 
 # source("functions and packages/functions.R")
 # source("master_scripts/plot_objects.R")
 # source("functions and packages/packages.R")
@@ -65,7 +64,7 @@ legalpha <- c(oct4, dec4, jan4, feb4, mar4, apr4)
 
 
 ####second set of figures----------------------------------------------------------------------------------------------------
-#windows(8,12)
+#windows(6,10)
 
 par(mfrow=c(3,1))
 par(mar=c(0,4,1,1), cex=1.25, las=1, cex.axis=.8, cex.lab=1, mgp=c(2.5,1,0))

### 2015-08-20 manuscript/manuscript_wtc3.Rmd
spelling fixes

@@
 Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling problematic.
-Incident PPFD declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor] and photosynthesis (A) responds to this changing irradiance until until biochemical limitation occurs [@evans1995carbon].
+Incident PPFD declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor] and photosynthesis (A) responds to this changing irradiance until biochemical limitation occurs [@evans1995carbon].
 Consequently, leaves modify structural and functional properties to efficiently intercept variable light and enhance photosynthetic capacity [@vogelman1996leaves; @valladares2004photosynthetic].
@@
 During A, CO~2~ diffuses from the atmosphere through stomata (g~s~), intercellular air space and into the chloroplast for fixation by Rubisco (mesophyll conductance, g~m~) [@tazoe2011using].
-Trade-offs exists that constrain each of these diffusion pathways, although these are yet to be explictily quantified for g~m~.
+Trade-offs exists that constrain each of these diffusion pathways, although these are yet to be explicitly quantified for g~m~.
 For optimal leaf C gain these pathways should act to minimize the resistance to CO~2~ diffusion while also limiting the intrinsic energy, water and construction costs which diminish the C return from A.
@@
 Mean g~s~ was significantly higher in shade compared to sun leaves (+18%) under their local light environment (P = 0.004).
-Increases in g~s~ were postively correlated to increases in A in sun leaves, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
+Increases in g~s~ were positively correlated to increases in A in sun leaves, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
 Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves ( Figure `r gebarfig`a).
@@
 Photosynthesis scaled positively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b).
-No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types.
+No differences in g~s~ or g~m~ were detected with the warming treatment within leaf types.
 Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figure `r get1fig`b and `r get2fig`b, respectively) and only a weak negative relationship with g~m~ and increasing leaf temperature was detected (Figure `r get1fig`b).
@@
 More open stomata in shade leaves under low and high light lead to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
-The drawdown of CO~2~ from intercellular spaces to the cholorplast, C~i~-C~c~, measures the coordination between g~m~ and A [@caemmerer2014temperature].
+The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and A [@caemmerer2014temperature].
 This drawdown was similar between sun and shade leaves at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
@@
 Relationship between the observed discrimination of ^13^CO~2~ and measured C~i~/C~a~ for sun leaves and shade leaves at both low and high light.
-The solid line represents the the theoretical line for C3 plants from [@evans1986carbon].
+The solid line represents the theoretical line for C3 plants from [@evans1986carbon].
 

### 2015-09-20 manuscript/manuscript_wtc3.Rmd
edits to manuscrpt from RD

@@
 ---
-title: "Fast response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks."
+title: "Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks"
 author: ""
@@
 \
-^1^ Hawkesbury Institute for the Environment, University of Western Sydney, Locked Bag 1797, Penrith, NSW, Australia
+^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia
 
@@
 The transient nature of light availability within a canopy, however, complicates our ability to assess the contribution of shade leaves to canopy carbon gain.
-* Leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis, stomatal conductance and mesophyll conductance of sun and shade leaves in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
-The physiological behaviour of individual leaves was then correlated with the distribution of nitrogen and water within the canopy to evaluate how trees are optimized for carbon gain.
-* More open stomata and rapid increases in mesophyll conductance allowed shade leaves to "lie in wait" for sunflecks and lead to leaf carbon gain exceeding sun leaves.
+* Leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis (A), stomatal conductance (g~s~) and mesophyll conductance (g~m~) of sun and shade leaves in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
+* More open stomata and rapid increases in mesophyll conductance
+
 * Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
-Evidence that mesophyll conductance not only varies within a canopy but can be up-regulated over short time intervals must now be considered in process based tree growth models.
+We argue that the rapid response of g~m~ with light allows shade leaves to "lie in wait" for sunflecks.
+Evidence that g~m~ not only complicates including gm~m~ in process based models of canopy carbon gain.
 
-# Key words
+## Key words
 photosynthesis, stomatal conductance, mesophyll conductance,  shade, leaf optimal behavior
@@
 # Introduction
-Understanding and predict carbon (C) uptake in forest ecosystems is crucially important for assessing the impacts of environmental change.
+Understanding and predicting carbon (C) uptake in forest ecosystems is crucially important for assessing the impacts of environmental change.
 Specifically, this involves accurate upscaling of leaf level measurements to predict whole canopy function.
@@
 Incident PPFD declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor] and photosynthesis (A) responds to this changing irradiance until biochemical limitation occurs [@evans1995carbon].
-Consequently, leaves modify structural and functional properties to efficiently intercept variable light and enhance photosynthetic capacity [@vogelman1996leaves; @valladares2004photosynthetic].
+Consequently, leaves modify structural and functional properties to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic].
 To estimate whole canopy C gain it is thus necessary to account for the non-linear response of A to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
@@
 \
-The distribution of resources required for A, including nitrogen (N) and water, are also partially defined by canopy light gradients.
-As A has a saturating response with light and a dependence on N, it has been argued that leaf N should be proportional to PPFD along the canopy gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
+The distribution of resources required for A, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients.
+As A has a saturating response with light and a the maximum rates depends on N, it has been argued that leaf N should be proportional to PPFD along the canopy gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
 This can lead to changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity across leaf types [@evans2001photosynthetic].
@@
 Higher rates of A and stomatal conductance (g~s~) can only be sustained if the hydraulic conductance (K) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal].
-Optimal photosynthetic N investment in the upper canopy will be wasted if photosynthetic capacity exceeds hydraulic supply [@niinemets2012optimization; @peltoniemi2012co], thus K should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
+Optimal photosynthetic N investment in the upper canopy will be wasted if hydraulic supply is insufficient [@niinemets2012optimization; @peltoniemi2012co], thus K should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
 
@@
 This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
-Second, variation in photosynthetic capacity between sun and shade leaves has also been hypothesized to be due to differences in g~m~ [@piel2002effect; @duursma2006vertical; @warren2007internal].
+Second, variation in photosynthetic capacity between sun and shade leaves has also been hypothesized to be due to differences in g~m~ [@piel2002effect; @warren2007internal].
 Mesophyll conductance can impose limitations on A as large as those associated with g~s~ [@warren2008stand; @ubierna2011estimation] and these limitations can reduce the efficiency of N use in A [@niinemets2007photosynthesis].
@@
 \
-1.	If whole tree canopies are optimized for C gain, then leaf N, K and photosynthetic capacity were predicted to be higher in sun leaves compared to shade leaves.
+1.	If whole tree canopies are optimized for C gain, then leaf N, K and photosynthetic capacity are predicted to be higher in sun leaves compared to shade leaves.
 
 \
-2.	Stomatal conductance should be proportional to A across sun and shade leaves under similar leaf VPD and g~m~ should scale positively with photosynthetic capacity.
+2.	Stomatal conductance is proportional to A across sun and shade leaves under similar leaf VPD and g~m~ scales positively with photosynthetic capacity.
 
@@
 Twelve *Eucalyptus tereticornis* seedlings, chosen from a single local Cumberland plain cohort, were planted in March 2013 into 12 whole-tree chambers (WTC) at the Hawkesbury Forest Experiment site near Richmond, NSW, Australia.
-Each chamber has a maximum height of 10 m and seedlings were grown for 15 months.
+Each chamber has a height of 9 m and seedlings were grown for 15 months.
 A detailed description of the WTC operation and design is available in [@barton2010whole].
-Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced a constant elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarsfig`).
+Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarsfig`).
 Trees were watered weekly with 70 L from March 2013 to November 2013.
@@
 Leaf level gas exchange was measured with a standard 6 cm^2^ leaf chamber using a portable gas exchange system (LI-6400, LI-COR, Lincoln, NE, USA).
-This system was coupled with a tunable diode laser (TDL;TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of online C isotope discrimination.
+This system was coupled with a tunable diode laser (TDL; TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of online C isotope discrimination.
 The CO~2~ in the leaf chamber was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mols s^-1^.
@@
 
+\
 Second, C isotope discrimination during C~3~ photosynthesis ($\Delta$) is the resultant discrimination from CO~2~ diffusion from the atmosphere to the site of carboxylation, consisting or a series of fractionation steps described in [@evans1986carbon].
@@
 
-\
+##Biochemical parameters of photosynthesis
 Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each chamber prior to the initiation of the drought treatment.
@@
 (12)
-$$K = \frac{E}{\Psi_{pd}~-\Psi_l-hp_wg}$$
+$$K = \frac{E}{\Psi_{pd}~-\Psi_l}$$
 
-where the term Hp~w~g represents the gravitational potential assumed to be minimal in this experiment due to small tree size [@whitehead1998regulation].
 Leaf level instantaneous water use efficiency (WUE~i~) was calculated as leaf photosynthesis divided by transpiration.
-The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2012reconciling] using non-linear regression, where K=0.5 [see @duursma2013near].
+The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2012reconciling] using non-linear regression, where k=0.5 [see @duursma2013near].
 
 ##Data analysis
-Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R with WTC as a random effect.
+Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R [@RDevelopmentCoreTeam2011] with WTC as a random effect.
 Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general].
-Confidence intervals (95 %) of mixed effect linear models of A as a function of different physiological parameters were generated using bootstrapping methods with 999 simulations in the 'lme4' package [@bates2013lme4] in R.
-For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package in R, using WTC as a random effect.
-Results were considered significant at P &leq; 0.05.
+Confidence intervals (95 %) of mixed effect linear models of A as a function of different physiological parameters were generated using bootstrapping methods with 999 simulations using the bootMer function in the 'lme4' package [@bates2013lme4].
+For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package, using WTC as a random effect.
+Results were considered significant when P &leq; 0.05.
 
@@
 ## Photosynthetic capacity and leaf photosynthesis rates 
-The photosynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (Table 1), as evident from AC~i~ curves from each chamber (Figure `r acifig`a).
+The photosynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (Table 1), as esimated from AC~i~ curves from each chamber (Figure `r acifig`a).
 Within leaf types, no effect of the warming treatment was detected on either parameter.
 Within individual chambers Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
-As g~m~ (results below) is included in the estimates of J~max~ and Vc~max~ with conventional AC~i~ curves, AC~c~ curves were simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.
+As g~m~ (results below) is implicit in the estimates of J~max~ and Vc~max~ with conventional AC~i~ curves, we also simulated AC~c~ curves to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.
 When accounting for the effect of g~m~, chloroplastic Vc~max~ and J~max~ were still higher in sun leaves (Figure `r accfig`).
@@
 Mean g~s~ was significantly higher in shade compared to sun leaves (+18%) under their local light environment (P = 0.004).
-Increases in g~s~ were positively correlated to increases in A in sun leaves, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
-Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves ( Figure `r gebarfig`a).
+In sun leaves, A was postively correlated with g~s~, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
+Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (Figure `r gebarfig`a).
 With increased light availability, increases in A and g~s~ in shade leaves exhibited similar behavior as observed in sun leaves.
@@
 \
-Leaf $\delta$^13^C, as an index of integrated water use efficiency [@marshall2007sources], significantly decreased from sun leaves to shade leaves by ca.
+Bulk-leaf $\delta$^13^C, as an index of integrated water use efficiency [@marshall2007sources], was significantly lower in shade leaves compared to sun leaves by ca.
 2&permil; (Table 1).
@@
 Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
-Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (Table 2).
+Following an increase in available light, g~m~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (Table 2).
 Mesophyll conductance was nearly proportional to A within leaf types and light treatments (Figure `r gebarfig`b,c).
@@
 Here we show that leaves within canopies of *Eucalyptus tereticornis* trees maintain the ability to upregulate A beyond the limitations of their local light environment.
-More open stomata and rapid increases in g~m~ allowed shade leaves to readily utilize increases in light availability, leading to leaf C gain exceeding sun leaves.
+More open stomata and rapid increases in g~m~ allowed shade leaves to readily utilize increases in light availability, leading to rates of leaf C gain exceeding sun leaves.
 Although these results should not be entirely unexpected as shade leaves experience interchanging periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies still impedes our ability to predict whole canopy C gain.
@@
 As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical].
-Our data agree with these conventional conclusions as the distribution of N~a~ , both biochemical parameters and A were all reduced in shade leaves.
+Our data agree with these conventional conclusions as the distribution of N~a~ , both measures of photosynthetic capacity and A were all reduced in shade leaves.
 
 \
-Photosynthesis in trees is also limited by the ability to supply water to the upper canopy.
+Photosynthesis is also limited by the ability to supply water to the upper canopy.
 Ultimately, the ability of a trees hydraulic architecture to supply water to foliage across increasing pathlengths affects productivity and survival [@sellin2008effects].
-Using a two-leaf model Peltoniemi et al.
+Using a two-leaf model, Peltoniemi et al.
 [-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if leaf K is also optimally distributed.
@@
 \
-Unexpected higher rates of g~s~ led to decreased WUE~i~ in shade leaves throughout the experiment.
-Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was constant across the leaf lifespan.
+Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased WUE~i~ throughout the experiment.
+Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was prevalent across the leaf lifespan.
 From a canopy perspective this pattern in WUE initially appears to be detrimental to C gain as A in sun leaves was characterized by low rates of g~s~ and low C~i~.
@@
 The pattern of wasteful water use in shade leaves is important as we hypothesized that g~s~ and A would be proportional across sun and shade leaves.
-In sun leaves variation in A and g~s~ were strongly correlated, exhibiting behaviour agreeing with optimal stomatal theory.
+In sun leaves, A and g~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
 However, lower rates of A in shade leaves were not coupled with decreases in g~s~, explaining the observed decreases in WUE.
@@
 \
-It is possible that sustaining higher g~s~, at the cost of larger transpiration losses, is a strategy to increase the utilization efficiency of sunflecks.
+It is possible that sustaining higher g~s~ in the shade, at the cost of larger transpiration losses, is a strategy to increase the utilization efficiency of sunflecks.
+Here, shade leaves were able to increase A equivalent or even outperforming sun leaves under identical light intensity.
 Rapid increases in leaf temperature with sunflecks has the potential to inhibit C gain [@leakey2003high].
 It has been suggested that transpiration-induced cooling is effective in avoiding sunfleck induced heat damage, however leaves need to keep stomata open in the shade for this to occur [@schymanski2013stomatal].
-Here, shade leaves were able to increase A equivalent or even outperforming sun leaves under identical light intensity.
 Whether more open stomata is a strategy to reduce stomatal response time or to decrease thermal load on shade leaves in trees is still unknown.
@@
 \
-Increasing the drawdown of CO~2~ into the intercellular air spaces via g~s~, however, does not necessarily infer concomitant increases in A.
-This is because CO~2~ supply to the site of carboxylation, via g~m~, may be anatomically constrained.
-In this study A and g~m~ scaled positively across leaf types and increased proportionately with increased light intensity in shade leaves.
+We found that A and g~m~ scaled positively across leaf types and, suprisingly, increased proportionately with increased light intensity in shade leaves.
 Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
@@
 \
-Similar to Laisk [-@laisk2005adjustment] we provide evidence that optimal acclimation, in this case the physiological adjustment of CO~2~ drawdown into shade leaves, is possibly directed towards occasional maximums of light availability over adaptation to a low light environment.
 If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution.

### 2015-09-21 manuscript/manuscript_wtc3.Rmd
manuscript edits RD

@@
 Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling problematic.
-Incident PPFD declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor] and photosynthesis (A) responds to this changing irradiance until biochemical limitation occurs [@evans1995carbon].
-Consequently, leaves modify structural and functional properties to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic].
+Incident PPFD declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor].
+Consequently, leaves modify structural and functional properties to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic] as changing irradiance strongly affects rates of photosynthesis (A) [@evans1995carbon].
 To estimate whole canopy C gain it is thus necessary to account for the non-linear response of A to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
@@
 As A has a saturating response with light and a the maximum rates depends on N, it has been argued that leaf N should be proportional to PPFD along the canopy gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
-This can lead to changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity across leaf types [@evans2001photosynthetic].
+Acclimiation of chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic].
 Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
@@
 \
-These theoretical distributions of N and water regulate leaf physiological traits that constrain A differently for sun and shade leaves.
-The balance between the costs and benefits of these leaf traits affecting photosynthetic capacity create trade-offs impacting canopy C gain [@givnish1988adaptation].
 During A, CO~2~ diffuses from the atmosphere through stomata (g~s~), intercellular air space and into the chloroplast for fixation by Rubisco (mesophyll conductance, g~m~) [@tazoe2011using].
-Trade-offs exists that constrain each of these diffusion pathways, although these are yet to be explicitly quantified for g~m~.
-For optimal leaf C gain these pathways should act to minimize the resistance to CO~2~ diffusion while also limiting the intrinsic energy, water and construction costs which diminish the C return from A.
-As these processes should not be considered independent of each other and must be integrated with hydraulic constraints [@griffiths2013you] predicting their behaviour within canopies remains difficult.
-
-\
-First, g~s~ should be efficiently distributed within a canopy to utilize supplies of light, N and water to maximize A [@peltoniemi2012co].
+Stomatal conductance should be efficiently distributed within a canopy to utilize supplies of light, N and water to maximize A [@peltoniemi2012co].
 This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
-Second, variation in photosynthetic capacity between sun and shade leaves has also been hypothesized to be due to differences in g~m~ [@piel2002effect; @warren2007internal].
-Mesophyll conductance can impose limitations on A as large as those associated with g~s~ [@warren2008stand; @ubierna2011estimation] and these limitations can reduce the efficiency of N use in A [@niinemets2007photosynthesis].
-Interpreting the coupled responses of both g~s~ and g~m~ to A across sun and shade leaves is critical for predicting canopy C gain, but empirical measurements across tree canopies are still lacking.
+Mesophyll conductance can also impose limitations on A as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in A [@niinemets2007photosynthesis].
+Variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
+Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack empirical data currently hinders are ability to interpreting their coupled responses to A across sun and shade leaves.
 
@@
 For example, at short timescales the utilization of sunflecks is limited by delayed responses of stomata, effectively limiting the maximum assimilation rate that can be achieved [@vico2011effects; @way2012sunflecks].
-Mesophyll conductance has been shown to respond to environmental factors across both short and long time scales, possibly faster than g~s~ [@flexas2008mesophyll].
-However, anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are unlikely to adapt during short light fluctuations.
+Mesophyll conductance has been shown to respond to environmental factors (e.g.
+CO~2~, temperature or VPD) across short time scales, possibly faster than g~s~ [@flexas2008mesophyll and references therin], yet short term response to light availabiity are unclear.
+For example, g~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid].
+Anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are also unlikely to adapt during short light fluctuations.
 The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance and as a potential response to transitory light availability.
@@
 A hand-held quantum sensor (LI-COR, Lincoln, NE, USA) was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either full sun or full shade.
-A ceptometer (AccuPAR LP-80, Decagon Devices, Pullman, USA) was then used to measure 1 m integrated PPFD at the canopy height of each chosen leaf type.
-Five ceptometer readings were recorded within the canopy at random locations of the height of each selected leaf.
+A ceptometer (AccuPAR LP-80, Decagon Devices, Pullman, USA) was then used to measure integrated PPFD at the canopy height of each chosen leaf type.
+Each ceptometer reading integrated an array of 80 sensors over a total length of 84 cm.
+Five ceptometer readings were recorded within the canopy at the same height and clsoe to each selected leaf.
 The mean of these readings was assumed to represent the local light environment of full sun and shade leaves for each tree.
-All measurements of PPFD and gas exchange were performed on full sun days between 10:00-14:30 h.
+All measurements of PPFD and gas exchange were performed on sunny days between 10:00-14:30 h.
 
@@
 Samples were flash combusted at 1000&deg;C and converted to CO~2~ and N~2~ and then subjected to stable isotope ratio mass spectrometry.
-Isotopic signatures are reported relative to the VPDP scale.
+Leaf N is reported on an area basis (N~a~) and isotopic signatures are reported relative to the VPDP scale.
 
@@
 
-Leaf level instantaneous water use efficiency (WUE~i~) was calculated as leaf photosynthesis divided by transpiration.
-The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2012reconciling] using non-linear regression, where k=0.5 [see @duursma2013near].
+Leaf level instantaneous transpiration efficiency (ITE) was calculated as leaf photosynthesis divided by transpiration.
+The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2011reconciling] using non-linear regression, where k=0.5 [see @duursma2013near].
 
@@
 This was because $\Psi_{pd}$, $\Psi_l$ (Table 1) and E (Table 2) did not differ between leaf types.
-Leaf N on an area basis (N~a~) was approximately 20% higher in sun leaves than in shade leaves (Table1).
+Leaf N~a~ was approximately 20% higher in sun leaves than in shade leaves (Table1).
 Leaf mass per area was not different between leaf types.
@@
 Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table 2).
-Additionally, leaf N~area~ was positively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).
+Additionally, leaf N~a~ was positively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).
 Following an increase in available light, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001).
@@
 \
-Leaf WUE~i~ was significantly greater in sun leaves than in shade leaves at low light (P = 0.001).
-Following an increase in available light, WUE~i~ of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
-WUE~i~ in sun leaves was reduced in the warming treatment but no effect was noted in shade leaves with low or high light (Table 2).
-For all leaf types and light treatments there was a strong response of WUE~i~ to VPD, and individual data points broadly corresponded to response curves from the optimal WUE model (Figure `r waterusefig`a).
+Leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001).
+Following an increase in available light, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
+Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment but no effect was noted in shade leaves with low or high light (Table 2).
+For all leaf types and light treatments there was a strong response of ITE to VPD, and individual data points broadly corresponded to response curves from the optimal WUE model (Figure `r waterusefig`a).
 The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
@@
 The observed carbon isotope discrimination was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`).
-Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta$~o~ (69.7&pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments.
+Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta$~o~ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
+No differences in g~s~ were detected with the warming treatment within leaf types and g~s~ did not vary significantly across measurement campaigns (Figure `r get1fig`b).
 
@@
 Photosynthesis scaled positively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b).
-No differences in g~s~ or g~m~ were detected with the warming treatment within leaf types.
-Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figure `r get1fig`b and `r get2fig`b, respectively) and only a weak negative relationship with g~m~ and increasing leaf temperature was detected (Figure `r get1fig`b).
+No differences in g~m~ were detected with the warming treatment within leaf types.
+Additionally, g~m~ did not vary significantly across measurements campaigns (Figure `r get2fig`b) and only a weak negative relationship with increasing leaf temperature was detected (Figure `r get1fig`b).
 
@@
 \
-Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased WUE~i~ throughout the experiment.
+Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased ITE throughout the experiment.
 Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was prevalent across the leaf lifespan.
@@
 \
-It is possible that sustaining higher g~s~ in the shade, at the cost of larger transpiration losses, is a strategy to increase the utilization efficiency of sunflecks.
-Here, shade leaves were able to increase A equivalent or even outperforming sun leaves under identical light intensity.
-Rapid increases in leaf temperature with sunflecks has the potential to inhibit C gain [@leakey2003high].
-It has been suggested that transpiration-induced cooling is effective in avoiding sunfleck induced heat damage, however leaves need to keep stomata open in the shade for this to occur [@schymanski2013stomatal].
-Whether more open stomata is a strategy to reduce stomatal response time or to decrease thermal load on shade leaves in trees is still unknown.
-Unfortunately, empirical studies that assess photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@leakey2002relative] and understory plants [@chazdon1991importance; @brantley2009contribution].
+It is possible that reducing stomatal response time, by sustaining higher g~s~, is a strategy to take advantage of high light quickly in shade leaves.
+Evidence from this study supports this hypothesis, as shade leaves increased A equivalent or even outperforming sun leaves under identical light intensity.
+Transipiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck induced thermal load [@schymanski2013stomatal].
+This is because rapid increases in leaf temperature with sunflecks have been shown to inhibit C gain [@leakey2003high].
+However, this response likely occurs at already high air temperatures and may not explain the observed rates of g~s~ in shade leaves across the large range of temperatures included in this study.
+How prevalent each of these strategies are within tree canopies is still unknown, yet empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@leakey2002relative] and understory plants [@chazdon1991importance; @brantley2009contribution].
 Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy [see @de1997simple].
@@
 \
-We found that A and g~m~ scaled positively across leaf types and, suprisingly, increased proportionately with increased light intensity in shade leaves.
+We found that A and g~m~ scaled positively across leaf types and, suprisingly, increased proportionately when light intensity was increased in shade leaves.
 Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
-This provides a potential explanation for the observed rapid increases in g~m~ but is largely untested in leaves of tree species.
+This provides a potential explanation for the observed rapid increases in g~m~ but the impacts of aquaporins on g~m~ are untested in leaves of tree species.
 Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
@@
 Although the distribution of resources, such as N and water, define leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
-These findings suggest that current theories of leaf optimal behaviour may be incomplete and have important implications for process based models that predict canopy C gain from rates of leaf photosynthesis.
-Furthermore, g~m~ must now be considered as a dynamic process in tree growth models that cannot be simply parameterized.
+These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have important implications for process based models that predict canopy C gain from rates of leaf photosynthesis.
+Furthermore, the dynamic nature of g~m~ cannot be simply parameterized in tree growth models and possibly should be excluded until it can be represented properly.
 Additional empirical data across multiple tree species are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown.
@@
 **Figure 3**.
-(a) Photosynthetic CO~2~ response (ACi) curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.
-ACi curves were developed once for all trees, before the initiation of the drought treatment, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-(b) The relationship between Vc~max~ and mean leaf nitrogen~area~ for each chamber, including sun leaves and shade leaves at low light.
-(c) the relationship between For (b,c) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional r^2^ of 0.28 and 0.35 for (b), and 0.24 and 0.33 for (c).
+(a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.
+ACi curves were measured once on all trees, before the initiation of the drought treatment, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+(b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light.
+(c) The relationship between A and leaf N~a~ for all each sun and shade leaf.
+For (b,c) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional R^2^ of 0.28 and 0.35 for (b), and 0.24 and 0.33 for (c).
 
@@
 **Figure 4**.
-The response of leaf photosynthesis rates to stomatal conductance (a) and mesophyll conductance (b) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
-The response of shade leaf physiology to high light was recorded once both photosynthesis and stomatal conductance were stable in the gas exchange leaf cuvette (ca.
+The response of A to g~s~ (a) and g~m~ (b) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
+The response of shade leaf physiology to high light was recorded once both A and g~s~ were stable in the gas exchange leaf cuvette (ca.
 25min).
@@
 \
-**Figure 5**.
-The mean &pm; 1 standard error of stomatal conductance (a), mesophyll conductance (b) and photosynthesis rate (c) of sun leaves and shade leaves at both low and high light with standard errors.
+**Figure 5**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light with standard errors.
 
@@
 **Figure 6**.
-(a) Response of instantaneous water use efficiency (ITE) to leaf vapor pressure deficit (VPD) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
+(a) Response of ITE to leaf vapor pressure deficit (VPD) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
 (b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves and shade leaves at low light.
-For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g1 value for each leaf type and treatment.
-For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional r^2^ of 0.41 and 0.45, respectively.
+For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment.
+For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
 
@@
 Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
-C~c~ values were predicted with mesophyll conductance, thus curves represent chloroplastic photosynthetic parameters at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+C~c~ values were predicted with g~m~, thus curves represent chloroplastic photosynthetic parameters at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
 
 \
-**Figure S2**.
-Response of photosynthesis (a), mesophyll conductance (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
-Shaded symbols represents each monthly measurement campaign.
+**Figure S2**. Response of A (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign.
 
 \
-**Figure S3**.
-Response of vapor pressure deficit (a), stomatal conductance (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
-Shaded symbols represents each monthly measurement campaign.
+**Figure S3**. Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign.
 

### 2015-09-21 master_scripts/wateruse.R

@@ -20,6 +20,11 @@ ite_agg$ite <- with(ite_agg, Photo/Trmmol)
 ite_sunsha <- ite_agg[ite_agg$leaflight != "shade-high",]
 ite_sunsha <- droplevels(ite_sunsha)
 
+# ite_mod <- lme(ite ~ leaf ,random=~1|chamber, data=ite_sunsha)
+# summary(ite_mod)
+# ite_sun <- mean(ite_sunsha[ite_sunsha$leaf == "sun", "ite"])
+# ite_sha<- mean(ite_sunsha[ite_sunsha$leaf == "shade", "ite"])
+
 ##dfr with lights on
 ite_lightson <- ite_agg[ite_agg$leaflight == "shade-high",]
 ite_lightson <- droplevels(ite_lightson)

### 2015-09-22 manuscript/manuscript_wtc3.Rmd
manuscript edits RD

@@
 * Leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis (A), stomatal conductance (g~s~) and mesophyll conductance (g~m~) of sun and shade leaves in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
-* More open stomata and rapid increases in mesophyll conductance
+* Reduction in A, Vc~max, J~max and leaf nitrogen were detected in shade leaves, however, they maintained consistently higher g~s~ than sun leaves.
+When light intensity was increased for shade leaves g~s~ was higher in shade than in sun leaves, g~m~ increased compared to the low light environment and A in shade leaves was greater than sun leaves at the same PPFD.
 
@@
 A detailed description of the WTC operation and design is available in [@barton2010whole].
-Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarsfig`).
-Trees were watered weekly with 70 L from March 2013 to November 2013.
-From December 2013 to final harvest trees were watered fortnightly with the mean monthly (100 yr) rainfall amount.
+Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature October 2013 to the end of the experiment trees were watered every 15 days with the mean monthly rainfall amount for Richmond, NSW.
 In February 2013 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering.
@@
 More open stomata and rapid increases in g~m~ allowed shade leaves to readily utilize increases in light availability, leading to rates of leaf C gain exceeding sun leaves.
-Although these results should not be entirely unexpected as shade leaves experience interchanging periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies still impedes our ability to predict whole canopy C gain.
-Failure to integrate this dynamic physiology into process based tree growth models may underestimate the ability of canopies to alter leaf C gain when environmental conditions change.
+This finding is unexpected because although we know shade leaves experience interchanging periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies still impedes our ability to predict whole canopy C gain.
+This dynamic behaviour of shade leaves complicates optimizations of canopy C uptake and accounting for the response of leaf physiology to transient light dynamics is an essential need in future studies.
 Additionally, with measurements recorded across a large natural range of temperatures we were unable to detect any effects of a +3 &deg;C warming treatment on leaf physiology.
@@
 Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was prevalent across the leaf lifespan.
-From a canopy perspective this pattern in WUE initially appears to be detrimental to C gain as A in sun leaves was characterized by low rates of g~s~ and low C~i~.
-As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower WUE when the photosynthetic efficiency of sun leaves appears at least partially constrained by water supply.
+From a canopy perspective this pattern in water use efficiency initially appears to be detrimental to C gain as A in sun leaves was characterized by low rates of g~s~ and low C~i~.
+Relative to the differences in photosynthesis rate between leaf types the higher rates of g~s~ in shade leaves appear to exhibit waste water loss.
+As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower ITE.
 

### 2015-09-23 manuscript/manuscript_wtc3.Rmd
adding citations from SP suggestions
added definition of shade leaf

@@
 * Leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis (A), stomatal conductance (g~s~) and mesophyll conductance (g~m~) of sun and shade leaves in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
-* Reduction in A, Vc~max, J~max and leaf nitrogen were detected in shade leaves, however, they maintained consistently higher g~s~ than sun leaves.
+* Reduction in A, Vc~max~, J~max~ and leaf nitrogen were detected in shade leaves, however, they maintained consistently higher g~s~ than sun leaves.
 When light intensity was increased for shade leaves g~s~ was higher in shade than in sun leaves, g~m~ increased compared to the low light environment and A in shade leaves was greater than sun leaves at the same PPFD.
-
 * Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
@@
 Additionally, assessing shade leaf behaviour is made difficult with accounting of short term light fluctuations within a canopy, via sunflecks.
-How shade leaves utilize sunflecks for short term C gain depends on the combined response time of g~s~ and g~m~ and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks].
+How shade leaves utilize sunflecks for short term C gain depends on the combined response time of g~s~ and g~m~ and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks;@tausz2005dynamic].
 For example, at short timescales the utilization of sunflecks is limited by delayed responses of stomata, effectively limiting the maximum assimilation rate that can be achieved [@vico2011effects; @way2012sunflecks].
@@
 The newest fully expanded leaf from the stem apex was chosen for gas exchange measurements and sun leaves were measured in the upper third of the canopy.
+Here, shade leaves are defined as inner-canopy leaves developing in a low light enivironment.
 In order to minimize any confounding effects of leaf age, shade leaves were measured first in the lower canopy then gradually higher during each campaign.
@@
 Relative to the differences in photosynthesis rate between leaf types the higher rates of g~s~ in shade leaves appear to exhibit waste water loss.
-As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower ITE.
+As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower ITE compared to sun leaves.
 
@@
 \
-It is possible that reducing stomatal response time, by sustaining higher g~s~, is a strategy to take advantage of high light quickly in shade leaves.
+It is possible that reducing stomatal response time, by sustaining higher g~s~, is a strategy to take advantage of high light quickly in shade leaves [@tausz2005dynamic].
 Evidence from this study supports this hypothesis, as shade leaves increased A equivalent or even outperforming sun leaves under identical light intensity.
@@
 However, this response likely occurs at already high air temperatures and may not explain the observed rates of g~s~ in shade leaves across the large range of temperatures included in this study.
-How prevalent each of these strategies are within tree canopies is still unknown, yet empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@leakey2002relative] and understory plants [@chazdon1991importance; @brantley2009contribution].
+How prevalent each of these strategies are within tree canopies is still unknown, yet empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution].
 Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy [see @de1997simple].

### 2015-09-23 manuscript/manuscript_wtc3.Rmd
attempted tp add pixie dust to table
started energy balance script

@@
 source("master_scripts/plot_objects.R")
-library(pander)
+library(pixiedust)
+
+##read table ready dataframes
+# resourcetab <- read.csv("master_scripts/resource_table.csv")
+# getab <- read.csv("master_scripts/ge_table.csv")
 

### 2015-09-23 EB_shade.R

@@ -1,0 +1,36 @@
+##run energy balance model on E and gs
+
+source("functions and packages/functions.R")
+source("master_scripts/plot_objects.R")
+source("functions and packages/packages.R")
+
+#read in gm data set (no drought) and Cibar(discrimination)-------------------------------------------------------
+gmes <- read.csv("calculated_data/gmes_wellwatered.csv")
+
+###get average by id
+gm_agg <- summaryBy(Photo+Cond+VpdL+Trmmol+PARi+gm+CO2R+Tair+Ci+CTleaf ~ 
+                      chamber+id+leaf +light+temp+leaflight+Month, data=gmes, FUN=mean, keep.names=TRUE)
+gm_agg$leaf <- gsub("s", "S", gm_agg$leaf)
+gm_agg$light <- gsub("high", "Sun-light", gm_agg$light)
+gm_agg$light <- gsub("low", "Shade-light", gm_agg$light)
+
+names(gm_agg)[10] <- "VPD"
+
+##remove shade-high
+gm_sunsha <- gm_agg[gm_agg$leaflight != "shade-high",]
+gm_sunsha <- droplevels(gm_sunsha)
+
+##sun-shade dataframes
+sundat <- gm_sunsha[gm_sunsha$leaflight =="sun-high",]
+shadat <- gm_sunsha[gm_sunsha$leaflight =="shade-low",]
+##dfr with lights on
+fleckdat <- gm_agg[gm_agg$leaflight == "shade-high",]
+fleckdat <- droplevels(fleckdat)
+
+
+shade_agg <- summaryBy(. ~ leaf, data=shadat, keep.names = TRUE)
+####run plantecophys on leaf-type dataframes
+shade_EB <- PhotosynEB(GS=shade_agg$Cond, Ca = shade_agg$CO2R,Tair=shade_agg$Tair)
+                     
+shade_EB <- PhotosynEB(GS=shadat$Cond, Ca = shadat$CO2R,Tair=shadat$Tair)                     
+

### 2015-09-25 manuscript/manuscript_wtc3.Rmd
pixie dust tables into manuscript

@@
 ##read table ready dataframes
-# resourcetab <- read.csv("master_scripts/resource_table.csv")
+resourcetab <- read.csv("master_scripts/resource_table.csv", stringsAsFactors = FALSE)
+getab <- read.csv("master_scripts/ge_table.csv", stringsAsFactors = FALSE)
+pval <- c(0.785, 0.001, 0.028, 0.002, 0.001, 0.3486, 0.6385, 0.973)
+pval2 <- c(0.001,	0.001,	0.001,	0.001,	0.001,	0.005,	0.001,	0.001)
 # getab <- read.csv("master_scripts/ge_table.csv")
@@
 Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
+Units of LMA and leaf N~area~ are g m^2^, K is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;.
 Different letters represent significant differences between treatments.
@@
 **Table 1**.
+*Eucalyptus tereticornis* leaf traits and the distribution of resources between full sun and shade leaves under ambient and elevated temperature treatments.
+Leaf mass per area, leaf N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
+Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
+Units of LMA and Leaf Nitrogen~area~ are g m^2^, K is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;.
+Different letters represent significant differences between treatments.
+The P value represents the overall difference between leaf types and warming treatments.
+
+```{r, echo=FALSE}
+resourcetab2 <- rbind(resourcetab, c("P value","", as.character(pval)))
+greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
+
+dust(resourcetab2)+
+  sprinkle_colnames("Leaf", "Temperature", "LMA", "Leaf N~area~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0(greeks['delta'],"^13^c")) +
+  sprinkle(rows= 1, part="head",bold=TRUE)+
+  sprinkle_print_method("markdown")
+```
 
@@
 **Table 2**.
+Responses of leaf level gas exchange parameters of *Eucalyptus tereticornis* trees between full sun and shade leaves under ambient and elevated temperature treatments.
+Each value reflects the mean (&pm; 1 standard error) for each treatment across all gas exchange campaigns (n=6).
+Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa.
+Different letters represent significant differences between treatments.
+The P value represents the overall difference between leaf types and both light and warming treatments.
+
+```{r, echo=FALSE}
+getab2 <- rbind(getab, c("P value","","", as.character(pval2)))
+greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
+
+dust(getab2)+
+  sprinkle_colnames("Leaf", "Light","Temperature", "A", "g~s~", "g~m~", "ITE","E","VPD","C~i~", "C~c~") +
+  sprinkle(rows= 1, part="head",bold=TRUE)+
+  sprinkle_print_method("markdown")
+```
 

### 2015-09-25 master_scripts/resource_table.R

@@ -48,6 +48,10 @@ leaf_vars <- leaf_table[, 1:2]
   leaf_vars$leaf <- gsub("s", "S", leaf_vars$leaf)
   leaf_vars$temp <- gsub("ambient", "Ambient", leaf_vars$temp)
   leaf_vars$temp <- gsub("elevated", "Elevated", leaf_vars$temp)
+
+leaf_vars2 <-data.frame(Leaf = c("Sun", "", "Shade", ""), Temperature = c("AT", "ET", "AT","ET"))  
+  
+  
 ##split means and SE
 leaf_means <- leaf_table[, 3:10]
 leaf_se <- leaf_table[, 11:18]
@@ -63,7 +67,7 @@ v7 <- data.frame(paste0(sprintf("%3.2f",round(leaf_means[,7], 1)), " (", sprintf
 v8 <- data.frame(paste0(sprintf("%3.1f",round(leaf_means[,8], 1)), " (", sprintf("%1.2f",round(leaf_se[,8],2)),")"))
 
 
-leaf_table2 <- cbind(leaf_vars, v1)
+leaf_table2 <- cbind(leaf_vars2, v1)
 leaf_table2 <- cbind(leaf_table2, v2)
 leaf_table2 <- cbind(leaf_table2, v3)
 leaf_table2 <- cbind(leaf_table2, v4)
@@ -107,12 +111,11 @@ leaf_table2[[10]] <- paste(leaf_table2[[10]], siglet3[[1]][,1])
 
 ##add pval at bottom
 
-pval <- as.vector(c("Container Effect (P)",0.785, 0.001, 0.028, 0.002, 0.001, 0.3486, 0.6385, 0.973))
-
+# pval <- as.vector(c("P value","Container Effect",0.785, 0.001, 0.028, 0.002, 0.001, 0.3486, 0.6385, 0.973))
+# 
+# test <- rbind(leaf_table2, pval)
 
 write.csv(leaf_table2, "master_scripts/resource_table.csv", row.names=FALSE)
-
-
 
 
 

### 2015-09-28 manuscript/manuscript_wtc3.Rmd
erase RD's functions in RMD
edits to manuscript.  this version sent to MT and BM

@@
 library(reporttools)
-
-# Shorthand function to print numbers with two decimal points (and retain trailing zeroes)
-f <- function(x, digits=2)sprintf(x, fmt=paste0("%.",digits,"f"))
-
-# Nice function to format P values.
-pval <- function(...)formatPval(..., includeEquality=TRUE)
 
@@
 * Leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis (A), stomatal conductance (g~s~) and mesophyll conductance (g~m~) of sun and shade leaves in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
-* Reduction in A, Vc~max~, J~max~ and leaf nitrogen were detected in shade leaves, however, they maintained consistently higher g~s~ than sun leaves.
-When light intensity was increased for shade leaves g~s~ was higher in shade than in sun leaves, g~m~ increased compared to the low light environment and A in shade leaves was greater than sun leaves at the same PPFD.
+* Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
+When light intensity was increased for shade leaves both g~s~ and g~m~ were higher compared to the low light environment, leading to increases in A greater than sun leaves at the same PPFD.
 * Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
 We argue that the rapid response of g~m~ with light allows shade leaves to "lie in wait" for sunflecks.
-Evidence that g~m~ not only complicates including gm~m~ in process based models of canopy carbon gain.
+Evidence that g~m~ not only varies within a canopy but can be up-regulated over short time intervals complicates including g~m~ in process-based models of canopy carbon gain
 
@@
 Incident PPFD declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor].
-Consequently, leaves modify structural and functional properties to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic] as changing irradiance strongly affects rates of photosynthesis (A) [@evans1995carbon].
+Consequently, leaves modify structural and functional properties to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of photosynthesis (A) [@evans1995carbon].
 To estimate whole canopy C gain it is thus necessary to account for the non-linear response of A to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
@@
 As A has a saturating response with light and a the maximum rates depends on N, it has been argued that leaf N should be proportional to PPFD along the canopy gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
-Acclimiation of chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic].
+Acclimation of chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic].
 Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
@@
 Mesophyll conductance has been shown to respond to environmental factors (e.g.
-CO~2~, temperature or VPD) across short time scales, possibly faster than g~s~ [@flexas2008mesophyll and references therin], yet short term response to light availabiity are unclear.
+CO~2~, temperature or VPD) across short time scales, possibly faster than g~s~ [@flexas2008mesophyll and references therein], yet short term response to light availability are unclear.
 For example, g~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid].
@@
 This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) scale positively with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis].
-How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~ will have implications for upscaling A for sun and shade leaves.
+How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for upscaling A for sun and shade leaves.
 Additionally, light saturated rates of A are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
@@
 \
-3.	As shade leaves are constrained by their underlying biochemistry and slow physiological responses, increases in A following sunfleck simulations were not expected to reach levels of full sun leaves.
+3.	As shade leaves are constrained by their underlying biochemistry and slow physiological responses, increases in A following sunfleck simulations were not expected to reach rates of full sun leaves.
 
@@
 A detailed description of the WTC operation and design is available in [@barton2010whole].
-Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature October 2013 to the end of the experiment trees were watered every 15 days with the mean monthly rainfall amount for Richmond, NSW.
+Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarsfig`).
+Trees were watered weekly with 70 L from March 2013 to November 2013.
+From October 2013 to the end of the experiment trees were watered every 15 days with the mean monthly rainfall amount for Richmond, NSW.
 In February 2013 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering.
@@
 \
-Before seedlings were planted into each chamber they were maintained under well watered conditions in 35 L pots and kept inside each chamber.
+Before seedlings were planted into each chamber they were maintained under well-watered conditions in 35 L pots and kept inside each chamber.
 This allowed for seedlings to gain sufficient biomass before planting while also allowing them to acclimate to chamber temperature treatments.
@@
 1 m) and roots were allowed to grow freely below the barrier.
-After 2 months, floors were installed 45 cm from the soil surface.
-This enabled chamber fluxes of CO~2~ and H~2~O from the whole tree canopy to be monitored (for methods see Barton et al.
-[-@barton2010whole]).
+After 2 months, floors were installed 45 cm from the soil surface, enabling chamber fluxes of CO~2~ and H~2~O from the whole tree canopy to be monitored.
 Leaf gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.
 At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm, height of 348&pm;15.1 cm and an estimated leaf area of 3.9&pm;0.1 m^2^.
-For ET treatments trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an estimated leaf area of 6.2&pm;0.2 m^2^.
+For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an estimated leaf area of 6.2&pm;0.2 m^2^.
 
 ##Leaf gas exchange, online carbon isotope discrimination and mesophyll conductance
-Leaf gas exchange measurements were performed monthly through to the final harvest.
+Leaf gas exchange measurements were performed in October 2013 and monthly from December 2013 to April 2014.
 Measurements were taken on a representative sun and shade leaf for each tree during each measurement campaign.
 The newest fully expanded leaf from the stem apex was chosen for gas exchange measurements and sun leaves were measured in the upper third of the canopy.
-Here, shade leaves are defined as inner-canopy leaves developing in a low light enivironment.
+Here, shade leaves are defined as inner-canopy leaves developing in a low light environment.
 In order to minimize any confounding effects of leaf age, shade leaves were measured first in the lower canopy then gradually higher during each campaign.
 As shade leaves most likely developed slower this assured that older leaves in the lower canopy were avoided.
-The nearest leaf on each branch was sampled for measurement of predawn leaf water potential.
 
@@
 Each ceptometer reading integrated an array of 80 sensors over a total length of 84 cm.
-Five ceptometer readings were recorded within the canopy at the same height and clsoe to each selected leaf.
-The mean of these readings was assumed to represent the local light environment of full sun and shade leaves for each tree.
+Five ceptometer readings were recorded within the canopy at the same height and close to each selected leaf.
+The mean of these readings was assumed to represent the local light environment of sun and shade leaves for each tree.
 All measurements of PPFD and gas exchange were performed on sunny days between 10:00-14:30 h.
@@
 \
-Leaf level gas exchange was measured with a standard 6 cm^2^ leaf chamber using a portable gas exchange system (LI-6400, LI-COR, Lincoln, NE, USA).
+Leaf level gas exchange was measured with a standard (2 x 3 cm) leaf chamber using a portable gas exchange system (LI-6400, LI-COR, Lincoln, NE, USA).
 This system was coupled with a tunable diode laser (TDL; TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of online C isotope discrimination.
@@
 PPFD in the chamber was set to match the individual light environment of each leaf type (explained above).
-Sun flecks were then simulated for shade leaves by increasing the chamber PPFD to the light environment of the full sun leaf in the same tree.
-The maximum sunfleck response was then recorded once shade leaves re-stabilized in the leaf cuvette (ca.
+Sun flecks were simulated for shade leaves by increasing the chamber PPFD to the light environment of the full sun leaf in the same tree.
+The maximum sunfleck response was recorded once CO~2~ and water vapor fluxes re-stabilized in the leaf cuvette (ca.
 25 min).
@@
 \
-Once the chamber environment for leaves was stable, for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet.
+Once CO~2~ and water vapor flux values were stable for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet.
 These gases were dried by passing through napion gas dryers in the respective gas lines, and then ^12^CO~2~ and ^13^CO~2~ concentrations were measured for each gas by the TDL.
@@
 The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDLAS on each measurement cycle.
-Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] were auto logged every 15 s for each gas exchange system over the same 12 min interval.
+Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the same 12 min interval.
 
@@
 \
-Second, C isotope discrimination during C~3~ photosynthesis ($\Delta$) is the resultant discrimination from CO~2~ diffusion from the atmosphere to the site of carboxylation, consisting or a series of fractionation steps described in [@evans1986carbon].
-In this experiment, a modified form of this equation presented in [@evans2013temperature] with ternary effect corrections by [@farquhar2012ternary] was used such that:
+Second, C isotope discrimination during C~3~ photosynthesis ($\Delta$) is the resultant discrimination from CO~2~ diffusion from the atmosphere to the site of carboxylation, consisting or a series of fractionation steps described in Evans et al.
+[-@evans1986carbon].
+In this experiment, a modified form of this equation presented in Evans & Von Caemmerer [-@evans2013temperature] with ternary effect corrections by Farquhar & Cernusak [-@farquhar2012ternary] was used such that:
 
@@
 ##Leaf chemistry and hydraulic parameters
-Following gas exchange measurements each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, dried to constant mass and then weighed.
+Following gas exchange measurements each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, dried to constant mass and weighed.
 These leaves were then milled and analyzed for leaf N content and $\delta$^13^C.
@@
 Predawn ($\Psi_{pd}$) and midday ($\Psi_l$) leaf water potentials (MPa) were measured for sun and shade leaves during each gas exchange campaign using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA).
+The leaf closest to the leaf used for gas exchange was sampled for measurement of $\Psi_{pd}$.
 Predawn leaf water potential was measured before sunrise on the same day as gas exchange measurements.
@@
 
-Leaf level instantaneous transpiration efficiency (ITE) was calculated as leaf photosynthesis divided by transpiration.
-The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2011reconciling] using non-linear regression, where k=0.5 [see @duursma2013near].
+Leaf level instantaneous transpiration efficiency (ITE) was calculated as A divided by E.
+The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression, where k=0.5 [see @duursma2013near].
 
@@
 Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general].
-Confidence intervals (95 %) of mixed effect linear models of A as a function of different physiological parameters were generated using bootstrapping methods with 999 simulations using the bootMer function in the 'lme4' package [@bates2013lme4].
+Confidence intervals (95 %) of mixed effect linear models were generated using bootstrapping methods with 999 simulations, using the bootMer function in the 'lme4' package [@bates2013lme4].
 For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package, using WTC as a random effect.
@@
 ## Leaf resource distribution
-Across all measurement campaigns PPFD was significantly different between sun and shade leaves (P<0.001) and PPFD was reduced by 78% in the shade (Figure `r parfig`).
+Across all monthly measurement campaigns PPFD was significantly different between sun and shade leaves (P<0.001) and PPFD was reduced by >75% in the shade (Figure `r parfig`).
 Leaf-specific hydraulic conductance was similar across sun and shade leaves (Table 1).
 This was because $\Psi_{pd}$, $\Psi_l$ (Table 1) and E (Table 2) did not differ between leaf types.
-Leaf N~a~ was approximately 20% higher in sun leaves than in shade leaves (Table1).
+Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table1).
 Leaf mass per area was not different between leaf types.
@@
 ## Photosynthetic capacity and leaf photosynthesis rates 
-The photosynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (Table 1), as esimated from AC~i~ curves from each chamber (Figure `r acifig`a).
+The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table 1), as estimated from AC~i~ curves from each chamber (Figure `r acifig`a).
 Within leaf types, no effect of the warming treatment was detected on either parameter.
@@
 \
-Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table 2).
+Mean A was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table 2).
 Additionally, leaf N~a~ was positively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).
-Following an increase in available light, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001).
-No effect of the warming treatment was detected on rates of A of sun leaves or shade leaves at low or high light.
+Following an increase light intensity, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001).
+No effect of the warming treatment was detected on A of sun leaves or shade leaves at low or high light.
 Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
@@
 ## Stomatal conductance and leaf water use efficiency
-Mean g~s~ was significantly higher in shade compared to sun leaves (+18%) under their local light environment (P = 0.004).
-In sun leaves, A was postively correlated with g~s~, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
-Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (Figure `r gebarfig`a).
-With increased light availability, increases in A and g~s~ in shade leaves exhibited similar behavior as observed in sun leaves.
+On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table 2).
+In sun leaves, A was positively correlated with g~s~, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
+Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (Figure `r gebarfig`a).
+With increased PPFD, increases in A and g~s~ in shade leaves were positively correlated  as observed in sun leaves.
 
@@
 Leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001).
-Following an increase in available light, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
+Following an increase in  PPFD, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
 Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment but no effect was noted in shade leaves with low or high light (Table 2).
-For all leaf types and light treatments there was a strong response of ITE to VPD, and individual data points broadly corresponded to response curves from the optimal WUE model (Figure `r waterusefig`a).
+For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal WUE model (Figure `r waterusefig`a).
 The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
@@
 Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
-Following an increase in available light, g~m~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (Table 2).
+With increased PPFD, g~m~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (Table 2).
 Mesophyll conductance was nearly proportional to A within leaf types and light treatments (Figure `r gebarfig`b,c).
@@
 More open stomata and rapid increases in g~m~ allowed shade leaves to readily utilize increases in light availability, leading to rates of leaf C gain exceeding sun leaves.
-This finding is unexpected because although we know shade leaves experience interchanging periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies still impedes our ability to predict whole canopy C gain.
+Although we know shade leaves experience interchanging periods of sun and shade [@pearcy1990sunflecks], these findings are unexpected because a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
 This dynamic behaviour of shade leaves complicates optimizations of canopy C uptake and accounting for the response of leaf physiology to transient light dynamics is an essential need in future studies.
-Additionally, with measurements recorded across a large natural range of temperatures we were unable to detect any effects of a +3 &deg;C warming treatment on leaf physiology.
+Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology.
 
 ## Resource distribution and photosynthetic capacity
-The allocation of N~a~ constrains photosynthesis rates and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
+The allocation of N~a~ constrains A and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
 Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979).
 As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical].
-Our data agree with these conventional conclusions as the distribution of N~a~ , both measures of photosynthetic capacity and A were all reduced in shade leaves.
+Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and A were all reduced in shade leaves.
 
@@
 From a canopy perspective this pattern in water use efficiency initially appears to be detrimental to C gain as A in sun leaves was characterized by low rates of g~s~ and low C~i~.
-Relative to the differences in photosynthesis rate between leaf types the higher rates of g~s~ in shade leaves appear to exhibit waste water loss.
+Relative to the differences in A between leaf types, higher rates of g~s~ in shade leaves appear to exhibit wasteful water loss.
 As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower ITE compared to sun leaves.
@@
 Evidence from this study supports this hypothesis, as shade leaves increased A equivalent or even outperforming sun leaves under identical light intensity.
-Transipiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck induced thermal load [@schymanski2013stomatal].
+Transpiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck induced thermal load [@schymanski2013stomatal].
 This is because rapid increases in leaf temperature with sunflecks have been shown to inhibit C gain [@leakey2003high].
-However, this response likely occurs at already high air temperatures and may not explain the observed rates of g~s~ in shade leaves across the large range of temperatures included in this study.
-How prevalent each of these strategies are within tree canopies is still unknown, yet empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution].
+However, this response likely occurs at already high air temperatures and may not explain the observed g~s~ in shade leaves across the large range of temperatures included in this study.
+How prevalent each of these strategies are within tree canopies is still unknown, as empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution].
 Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy [see @de1997simple].
@@
 \
-We found that A and g~m~ scaled positively across leaf types and, suprisingly, increased proportionately when light intensity was increased in shade leaves.
+We found that A and g~m~ scaled positively across leaf types and, surprisingly, increased proportionately when light intensity was increased in shade leaves.
 Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
@@
 Furthermore, the dynamic nature of g~m~ cannot be simply parameterized in tree growth models and possibly should be excluded until it can be represented properly.
-Additional empirical data across multiple tree species are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown.
+Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown.
 To improve our ability to predict whole canopy C gain future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.
@@
 *Eucalyptus tereticornis* leaf traits and the distribution of resources between full sun and shade leaves under ambient and elevated temperature treatments.
-Leaf mass per area, leaf N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
+Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
 Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
-Units of LMA and leaf N~area~ are g m^2^, K is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;.
-Different letters represent significant differences between treatments.
-The P value represents the overall difference between leaf types and warming treatments.
+Units of LMA and Leaf Nitrogen~area~ are g m^2^, K is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;.
+Different letters represent significant differences between leaf type and temperature treatments.
+The P value represents the overall effect of leaf type and temperature treatment on each trait.
 
@@
 **Table 2**.
-Responses of leaf level gas exchange parameters of *Eucalyptus tereticornis* trees between full sun and shade leaves under ambient and elevated temperature treatments.
-Each value reflects the mean (&pm; 1 standard error) for each treatment across all gas exchange campaigns (n=6).
+Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
+Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).
 Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa.
-Different letters represent significant differences between treatments.
-The P value represents the overall difference between leaf types and both light and warming treatments.
+Different letters represent significant differences between leaf type, light environment and temperature treatments.
+The P value represents the overall effect of leaf type, light environment and temperature treatment on each parameter.
 
@@
 **Figure 2**.
-Local light environment for sun and shade leaves for each gas exchange campaign.
-Means &pm; 1 standard error represent 1 m integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf.
+Local light environment for sun and shade leaves for each monthly gas exchange campaign.
+Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf.
 
@@
 (b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light.
-(c) The relationship between A and leaf N~a~ for all each sun and shade leaf.
-For (b,c) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional R^2^ of 0.28 and 0.35 for (b), and 0.24 and 0.33 for (c).
+(c) The relationship between A and leaf N~a~ for each sun and shade leaf.
+For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
 
@@
 The response of A to g~s~ (a) and g~m~ (b) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
-The response of shade leaf physiology to high light was recorded once both A and g~s~ were stable in the gas exchange leaf cuvette (ca.
-25min).
 Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
-Grey areas are approximately 95% confidence intervals from the mean.
+Grey areas are 95% confidence intervals from the mean.
 
 \
-**Figure 5**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light with standard errors.
+**Figure 5**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light.
 
@@
 **Figure 6**.
-(a) Response of ITE to leaf vapor pressure deficit (VPD) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
+(a) Response of ITE to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
 (b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves and shade leaves at low light.
@@
 Relationship between the observed discrimination of ^13^CO~2~ and measured C~i~/C~a~ for sun leaves and shade leaves at both low and high light.
-The solid line represents the theoretical line for C3 plants from [@evans1986carbon].
+The solid line represents the theoretical line for C3 plants from Evans et al.
+[-@evans1986carbon].
 
@@
 Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
-C~c~ values were predicted with g~m~, thus curves represent chloroplastic photosynthetic parameters at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+C~c~ values were predicted with g~m~, thus curves represent chloroplastic photosynthetic parameters at 25°C and saturating light (1800 $\mu$mols m^-1^ s^-1^).
 
@@
 *Eucalyptus tereticornis* leaf traits and the distribution of resources between full sun and shade leaves under ambient and elevated temperature treatments.
-Leaf mass per area, leaf N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
+Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
 Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
 Units of LMA and Leaf Nitrogen~area~ are g m^2^, K is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;.
-Different letters represent significant differences between treatments.
-The P value represents the overall difference between leaf types and warming treatments.
+Different letters represent significant differences between leaf type and temperature treatments.
+The P value represents the overall effect of leaf type and temperature treatment on each trait.
 
@@
 dust(resourcetab2)+
-  sprinkle_colnames("Leaf", "Temperature", "LMA", "Leaf N~area~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0(greeks['delta'],"^13^c")) +
+  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0(greeks['delta'],"^13^c")) +
   sprinkle(rows= 1, part="head",bold=TRUE)+
@@
 # Figures {.unnumbered}
-```{r airvars, fig.width=5, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r airvars, fig.width=7, fig.height=9, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
 source("master_scripts/air_vars.R")
@@
 
-```{r acicurves, fig.width=8, fig.height=8, warning=FALSE,echo=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
 source("master_scripts/aci_vcmaxnitro.R")
@@
 
-```{r Agmgs, fig.width=5, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
 source("master_scripts/Agmgs.R")
@@
 
-```{r barplots, fig.width=10, fig.height=5, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
 source("master_scripts/physiology_barplots.R")
@@
 
-```{r wateruse, fig.width=5, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
+```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
 source("master_scripts/wateruse.R")
@@
 
-```{r cicc, fig.width=10, fig.height=5, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
 source("master_scripts/cicc_bar.R")
@@
 
-```{r gasex_temp1, fig.width=6, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r gasex_temp1, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
 source("master_scripts/gasex_temp.R")
@@
 
-```{r gasex_temp2, fig.width=6, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r gasex_temp2, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
 source("master_scripts/gasex_temp2.R")

### 2015-10-08 manuscript/manuscript_wtc3.Rmd
manuscrpt edits...MT

@@
 * Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
-When light intensity was increased for shade leaves both g~s~ and g~m~ were higher compared to the low light environment, leading to increases in A greater than sun leaves at the same PPFD.
+When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves at the same PPFD.
 * Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
-We argue that the rapid response of g~m~ with light allows shade leaves to "lie in wait" for sunflecks.
+We argue that the rapid response of g~m~ with light enables shade leaves to respond quickly to sunflecks.
 Evidence that g~m~ not only varies within a canopy but can be up-regulated over short time intervals complicates including g~m~ in process-based models of canopy carbon gain
@@
 Specifically, this involves accurate upscaling of leaf level measurements to predict whole canopy function.
-Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling problematic.
+Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling based on solar irradiance problematic.
 Incident PPFD declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor].
-Consequently, leaves modify structural and functional properties to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of photosynthesis (A) [@evans1995carbon].
+Consequently, structural and functional properties of leaves are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of photosynthesis (A) [@evans1995carbon].
 To estimate whole canopy C gain it is thus necessary to account for the non-linear response of A to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
@@
 The distribution of resources required for A, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients.
-As A has a saturating response with light and a the maximum rates depends on N, it has been argued that leaf N should be proportional to PPFD along the canopy gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
+As A has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
 Acclimation of chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic].
@@
 Higher rates of A and stomatal conductance (g~s~) can only be sustained if the hydraulic conductance (K) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal].
-Optimal photosynthetic N investment in the upper canopy will be wasted if hydraulic supply is insufficient [@niinemets2012optimization; @peltoniemi2012co], thus K should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
+Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing A if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co], thus K should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
 
@@
 During A, CO~2~ diffuses from the atmosphere through stomata (g~s~), intercellular air space and into the chloroplast for fixation by Rubisco (mesophyll conductance, g~m~) [@tazoe2011using].
-Stomatal conductance should be efficiently distributed within a canopy to utilize supplies of light, N and water to maximize A [@peltoniemi2012co].
+Based on optimality theory, stomatal conductance should be efficiently distributed within a canopy to utilize supplies of light, N and water to maximize A [@peltoniemi2012co].
 This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].

### 2015-10-21 manuscript/manuscript_wtc3.Rmd
manuscript edits from MT

@@
 This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
-Mesophyll conductance can also impose limitations on A as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in A [@niinemets2007photosynthesis].
-Variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
+Mesophyll conductance can also impose limitations on A as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in A [@niinemets2007photosynthesis] if g~m~ constrains CO~2~ supply to the chloroplast.
+Part of the variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
 Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack empirical data currently hinders are ability to interpreting their coupled responses to A across sun and shade leaves.
@@
 Mesophyll conductance has been shown to respond to environmental factors (e.g.
-CO~2~, temperature or VPD) across short time scales, possibly faster than g~s~ [@flexas2008mesophyll and references therein], yet short term response to light availability are unclear.
+CO~2~, temperature or vapor pressure deficit) across short time scales, possibly faster than g~s~ [@flexas2008mesophyll and references therein], yet short term response to light availability are unclear.
 For example, g~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid].
-Anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are also unlikely to adapt during short light fluctuations.
+Anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are also unlikely to adjust during short light fluctuations.
 The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance and as a potential response to transitory light availability.
@@
 In this study we use *Eucalyptus tereticornis* Sm.
-trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (+3&deg;C) treatments, to empirically evaluate the distribution of resources and leaf physiological behaviour of sun and shade leaves.
+trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to empirically evaluate the distribution of resources and leaf physiological behaviour of sun and shade leaves.
 Our hypotheses are as follows:
@@
 \
-Before seedlings were planted into each chamber they were maintained under well-watered conditions in 35 L pots and kept inside each chamber.
-This allowed for seedlings to gain sufficient biomass before planting while also allowing them to acclimate to chamber temperature treatments.
+Before seedlings were planted into each chamber they were maintained under well-watered conditions in 25 L pots and kept inside each chamber.
+This allowed for seedlings to gain sufficient size before planting while also allowing them to acclimate to chamber temperature treatments.
 Seedlings were planted into each chamber after mean seedling height reached 100 cm.
@@
 1 m) and roots were allowed to grow freely below the barrier.
-After 2 months, floors were installed 45 cm from the soil surface, enabling chamber fluxes of CO~2~ and H~2~O from the whole tree canopy to be monitored.
 Leaf gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.
-At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm, height of 348&pm;15.1 cm and an estimated leaf area of 3.9&pm;0.1 m^2^.
-For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an estimated leaf area of 6.2&pm;0.2 m^2^.
+At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm, height of 348&pm;15.1 cm and an leaf area of 3.9&pm;0.1 m^2^.
+For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an leaf area of 6.2&pm;0.2 m^2^.
 
 ##Leaf gas exchange, online carbon isotope discrimination and mesophyll conductance
-Leaf gas exchange measurements were performed in October 2013 and monthly from December 2013 to April 2014.
+Leaf gas exchange measurements were performed six times, beginning in October 2013 and monthly from December 2013 to April 2014.
 Measurements were taken on a representative sun and shade leaf for each tree during each measurement campaign.
-The newest fully expanded leaf from the stem apex was chosen for gas exchange measurements and sun leaves were measured in the upper third of the canopy.
+The newest fully expanded leaf from the branch apex was chosen for gas exchange measurements and sun leaves were measured in the upper third of the canopy.
 Here, shade leaves are defined as inner-canopy leaves developing in a low light environment.
@@
 \
-Prior to gas exchange measurements photosynthetic photon flux density (PPFD) was recorded at the individual leaf level and at the canopy position for each selected leaf.
+Prior to gas exchange measurements photosynthetic photon flux density (PPFD) was recorded as a point measurement at the individual leaf level and a spatially averaged measurement at the canopy position for each selected leaf.
 A hand-held quantum sensor (LI-COR, Lincoln, NE, USA) was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either full sun or full shade.
@@
 \
-Leaf level gas exchange was measured with a standard (2 x 3 cm) leaf chamber using a portable gas exchange system (LI-6400, LI-COR, Lincoln, NE, USA).
+Leaf level gas exchange was measured with a standard (2 x 3 cm) leaf cuvette using a portable gas exchange system (LI-6400XT, LI-COR, Lincoln, NE, USA).
 This system was coupled with a tunable diode laser (TDL; TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of online C isotope discrimination.
-The CO~2~ in the leaf chamber was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mols s^-1^.
+The CO~2~ in the leaf cuvette was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mol s^-1^.
 Two identical gas exchanges systems were run simultaneously, one in each of a randomly chosen WTC for each temperature treatment.
-Leaf temperatures were controlled at the current AT or ET chamber air temperature.
-PPFD in the chamber was set to match the individual light environment of each leaf type (explained above).
-Sun flecks were simulated for shade leaves by increasing the chamber PPFD to the light environment of the full sun leaf in the same tree.
+Leaf temperatures were controlled at the current AT or ET WTC air temperature.
+PPFD in the cuvette was set to match the individual light environment of each leaf type (explained above).
+Sun flecks were simulated for shade leaves by increasing the leaf cuvette PPFD (LI-COR red/blue light source) to match the light environment of the full sun leaf in the same tree.
 The maximum sunfleck response was recorded once CO~2~ and water vapor fluxes re-stabilized in the leaf cuvette (ca.
@@
 \
-Once CO~2~ and water vapor flux values were stable for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet.
-These gases were dried by passing through napion gas dryers in the respective gas lines, and then ^12^CO~2~ and ^13^CO~2~ concentrations were measured for each gas by the TDL.
-Reference, sample and 2 calibration gases were run on alternating 80 s loops (20 s each), one for each paired WTC leaf, for a total of 12 min.
-This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference and samples gases for calculations.
-The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDLAS on each measurement cycle.
+Once CO~2~ and water vapor flux values were stable for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet of the LI-6400.
+The gas streams were dried by passing through napion gas dryers in the respective gas lines, and then ^12^CO~2~ and ^13^CO~2~ concentrations were measured for each gas stream by the TDL.
+Reference, sample and two calibration gases were run on alternating 80 s loops (20 s each), one for each AT and ET leaf at a matched canopy position, for a total of 12 min.
+This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference line and samples line gas streams for calculations.
+The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDL on each measurement cycle.
 Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the same 12 min interval.
@@
 Using online C isotope discrimination measurements, the difference  between the observed discrimination and what is predicted for light saturated gas exchange is proportional to g~m~ [@griffiths2013mesophyll].
-First, leaf discrimination was calculated by comparing the isotopic composition of the reference gas entering the chamber ($\delta$^13^C~e~) with the sample gas ($\delta$^13^C~o~) such that:
+First, leaf C isotope discrimination was calculated by comparing the isotopic composition of the reference gas entering the leaf cuvette ($\delta$^13^C~e~) with the sample gas ($\delta$^13^C~o~) such that:
 
@@
 
-and $\xi$ is the ratio of the CO~2~ entering the well mixed leaf cuvette to the CO~2~ draw down by the leaf.
+and $\xi$ is the ratio of the CO~2~ entering the well mixed leaf cuvette to the CO~2~ draw down in the gas stream by the leaf.
 
@@
 \begin{aligned}
-\Delta_e = &\frac{1+t}{1-t}\left(\frac{eR_D}{(A+R_d)C_a}(C_i-\Gamma_*)\right)
+\Delta_e = &\frac{1+t}{1-t}\left(\frac{eR_d}{(A+R_d)C_a}(C_i-\Gamma_*)\right)
 \end{aligned}
@@
 C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~).
-In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulis* from [@crous2012light].
+In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulus* from [@crous2012light].
 The ternary effect corrections (t) are described by:
@@
 Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were then estimated using Equation 11.
-Examples of this approach to measure gas exchange and C isotope discrimination are presented in [@evans2013temperature].
+Examples of this approach to measure gas exchange and C isotope discrimination are presented in Evans and Von Caemmerer [-@evans2013temperature].
 The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck where then compared as a function of C~i~:C~a~.
@@
 ##Biochemical parameters of photosynthesis
-Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each chamber prior to the initiation of the drought treatment.
+Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each WTC prior to the initiation of the drought treatment.
 Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
@@
 ##Leaf chemistry and hydraulic parameters
-Following gas exchange measurements each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, dried to constant mass and weighed.
+Following gas exchange measurements each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, dried and weighed.
 These leaves were then milled and analyzed for leaf N content and $\delta$^13^C.
@@
 All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
-Leaf water potential and transpiration (E, mmol^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K) through the equation:
+Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K) through the equation:
 
@@
 ## Photosynthetic capacity and leaf photosynthesis rates 
-The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table 1), as estimated from AC~i~ curves from each chamber (Figure `r acifig`a).
+The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table 1), as estimated from AC~i~ curves (Figure `r acifig`a).
 Within leaf types, no effect of the warming treatment was detected on either parameter.
-Within individual chambers Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
+Among the sampled leaves, Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
 As g~m~ (results below) is implicit in the estimates of J~max~ and Vc~max~ with conventional AC~i~ curves, we also simulated AC~c~ curves to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.
@@
 Additionally, leaf N~a~ was positively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).
-Following an increase light intensity, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001).
-No effect of the warming treatment was detected on A of sun leaves or shade leaves at low or high light.
+Following an increase light intensity to match high-light conditions, A of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001).
+No effect of the warming treatment was detected on A of sun leaves measured at high light or shade leaves at either low or high light.
 Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
@@
 In sun leaves, A was positively correlated with g~s~, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
-Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (Figure `r gebarfig`a).
+Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
 With increased PPFD, increases in A and g~s~ in shade leaves were positively correlated  as observed in sun leaves.
@@
 \
-Leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001).
-Following an increase in  PPFD, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
-Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment but no effect was noted in shade leaves with low or high light (Table 2).
+Measured under ambient light and temperature conditions leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001).
+Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
+Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment but no warmding effect was detected in shade leaves measured at low or high light (Table 2).
 For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal WUE model (Figure `r waterusefig`a).
 The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
-Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a)
+Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).
 
@@
 No effects of the warming treatment on leaf $\delta$^13^C were detected.
-Leaf $\delta$^13^C and N~area~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C and higher N investment in sun leaves.
+Leaf $\delta$^13^C and N~area~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C (higher water-use efficiency) and higher N investment in sun leaves.
 
@@
 However, this response likely occurs at already high air temperatures and may not explain the observed g~s~ in shade leaves across the large range of temperatures included in this study.
-How prevalent each of these strategies are within tree canopies is still unknown, as empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution].
+How prevalent each of these strategies are within tree canopies is still unknown, as empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants, often in deep shade [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution].
 Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy [see @de1997simple].
@@
 \
-We found that A and g~m~ scaled positively across leaf types and, surprisingly, increased proportionately when light intensity was increased in shade leaves.
+We found that A and g~m~ scaled positively across leaf types and, surprisingly, increased rapidly (within minutes) and proportionately when light intensity was increased in shade leaves.
 Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
@@
 **Figure 4**.
-The response of A to g~s~ (a) and g~m~ (b) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
+The response of A to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
 Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
@@
 \
-**Figure 5**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light.
+**Figure 5**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
 
@@
 **Figure 6**.
-(a) Response of ITE to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
-(b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves and shade leaves at low light.
+(a) Response of instantaneous transpiration efficiency (ITE) to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
+(b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light.
 For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment.
@@
 **Figure 7**.
-Relationship between the observed discrimination of ^13^CO~2~ and measured C~i~/C~a~ for sun leaves and shade leaves at both low and high light.
+Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis and measured C~i~/C~a~ for sun leaves measured at hight light and shade leaves measured at both low and high light.
 The solid line represents the theoretical line for C3 plants from Evans et al.

### 2015-10-21 manuscript/manuscript_wtc3.Rmd
pixiedust change

@@
 
-dust(resourcetab2)+
-  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0(greeks['delta'],"^13^c")) +
-  sprinkle(rows= 1, part="head",bold=TRUE)+
+dust(resourcetab2)%>%
+  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0(greeks['delta'],"^13^c")) %>%
+  sprinkle(rows= 1, part="head",bold=TRUE)%>%
   sprinkle_print_method("markdown")
@@
 
-dust(getab2)+
-  sprinkle_colnames("Leaf", "Light","Temperature", "A", "g~s~", "g~m~", "ITE","E","VPD","C~i~", "C~c~") +
-  sprinkle(rows= 1, part="head",bold=TRUE)+
+dust(getab2)%>%
+  sprinkle_colnames("Leaf", "Light","Temperature", "A", "g~s~", "g~m~", "ITE","E","VPD","C~i~", "C~c~") %>%
+  sprinkle(rows= 1, part="head",bold=TRUE)%>%
   sprinkle_print_method("markdown")

### 2015-11-10 manuscript/manuscript_wtc3.Rmd
manuscript edits mt

@@
 # Summary
-* Light gradients within tree canopies play a major role in the distribution of resources that define photosynthetic capacity of individual leaves.
+* Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves.
 The transient nature of light availability within a canopy, however, complicates our ability to assess the contribution of shade leaves to canopy carbon gain.
@@
 In February 2013 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering.
-Due to a limited range of data for the drought treatment only well-watered trees are reported, which reduces the sample size of WTC (n=6) after the initiation of this treatment.
+Due to a limited range of data for the drought treatment only well-watered trees are reported, which reduces the sample size of WTC (n=6 to n=3) for the final 3 months of the experiment.
 
@@
 Seedlings were planted into each chamber after mean seedling height reached 100 cm.
-The top soils at this site, used in both pots and chambers, are an alluvial formation of low-fertility sandy loam soils with low organic matter and low water holding capacity.
+The top soils at this site, used in both pots and chambers, are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg-1 total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity.
 A root exclusion barrier extended from chamber walls to the hard layer (ca.

### 2015-11-16 manuscript/manuscript_wtc3.Rmd
figure and text edits from MT

@@
 ```{r, echo=FALSE}
-airvarsfig <- "1"
-parfig <- "2"
-acifig <- "3"
-agmgsfig <- "4"
-gebarfig <- "5"
-waterusefig <- "6"
-deltafig <- "7"
-ciccfig <- "8"
 
-accfig <- "S1"
-get1fig <- "S2"
-get2fig <- "S3"
+parfig <- "1"
+acifig <- "2"
+agmgsfig <- "3"
+gebarfig <- "4"
+waterusefig <- "5"
+deltafig <- "6"
+ciccfig <- "7"
+
+airvarsfig <- "S1"
+accfig <- "S2"
+get1fig <- "S3"
+get2fig <- "S4"
 
@@
 Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling based on solar irradiance problematic.
-Incident PPFD declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor].
+Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor].
 Consequently, structural and functional properties of leaves are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of photosynthesis (A) [@evans1995carbon].
@@
 As A has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
-Acclimation of chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic].
+Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic].
 Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
@@
 In this study we use *Eucalyptus tereticornis* Sm.
-trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to empirically evaluate the distribution of resources and leaf physiological behaviour of sun and shade leaves.
+trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to empirically evaluate the distribution of nitrogen and water supply and leaf physiological behaviour of sun and shade leaves.
 Our hypotheses are as follows:
@@
 Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarsfig`).
+Trees grew quickly and developed large canopies, with height growth reaching the top of the WTCs over the experiment duration.
 Trees were watered weekly with 70 L from March 2013 to November 2013.
@@
 For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an leaf area of 6.2&pm;0.2 m^2^.
+Leaf area was calculated based on leaf counts and mean leaf size.
 
@@
 The newest fully expanded leaf from the branch apex was chosen for gas exchange measurements and sun leaves were measured in the upper third of the canopy.
-Here, shade leaves are defined as inner-canopy leaves developing in a low light environment.
-In order to minimize any confounding effects of leaf age, shade leaves were measured first in the lower canopy then gradually higher during each campaign.
-As shade leaves most likely developed slower this assured that older leaves in the lower canopy were avoided.
+Here, shade leaves are defined as inner-canopy leaves developing on secondary branches in a low light environment.
+Shade leaves were always measured in the lower canopy, but leaves were sampled on subsequent higher branches across measurement campaigns to minimize confounding effects of leaf age.
+As shade leaves most likely developed slower this assured that older leaves in the bottom canopy were avoided.
 
@@
 A hand-held quantum sensor (LI-COR, Lincoln, NE, USA) was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either full sun or full shade.
-A ceptometer (AccuPAR LP-80, Decagon Devices, Pullman, USA) was then used to measure integrated PPFD at the canopy height of each chosen leaf type.
+A ceptometer (AccuPAR LP-80, Decagon Devices, Pullman, USA) was then used to measure a spatially averaged PPFD at the canopy height of each chosen leaf type.
 Each ceptometer reading integrated an array of 80 sensors over a total length of 84 cm.
-Five ceptometer readings were recorded within the canopy at the same height and close to each selected leaf.
+Five ceptometer readings were recorded at different locations within the canopy, but at the same height and close to each selected leaf.
 The mean of these readings was assumed to represent the local light environment of sun and shade leaves for each tree.
@@
 \
-Second, C isotope discrimination during C~3~ photosynthesis ($\Delta$) is the resultant discrimination from CO~2~ diffusion from the atmosphere to the site of carboxylation, consisting or a series of fractionation steps described in Evans et al.
+Second, C isotope discrimination during C~3~ photosynthesis ($\Delta$) is the resultant discrimination from CO~2~ diffusion from the atmosphere to the site of carboxylation, consisting of a series of fractionation steps described in Evans et al.
 [-@evans1986carbon].
@@
 Samples were flash combusted at 1000&deg;C and converted to CO~2~ and N~2~ and then subjected to stable isotope ratio mass spectrometry.
-Leaf N is reported on an area basis (N~a~) and isotopic signatures are reported relative to the VPDP scale.
+Leaf N is reported on an area basis (N~a~, g m^-2^) and isotopic signatures are reported relative to the VPDP scale.
 
@@
 All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
-Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K) through the equation:
+Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K, mmol m^-2^ s^-1^ MPa) through the equation:
 
@@
 For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package, using WTC as a random effect.
-Results were considered significant when P &leq; 0.05.
-
 # Results
@@
 ## Leaf resource distribution
-Across all monthly measurement campaigns PPFD was significantly different between sun and shade leaves (P<0.001) and PPFD was reduced by >75% in the shade (Figure `r parfig`).
-Leaf-specific hydraulic conductance was similar across sun and shade leaves (Table 1).
+Across six measurement campaigns over the 7 month period PPFD was  reduced on average by >75% in the shade (Figure `r parfig`).
+Leaf-specific hydraulic conductance (K) was similar across sun and shade leaves (Table 1).
 This was because $\Psi_{pd}$, $\Psi_l$ (Table 1) and E (Table 2) did not differ between leaf types.
@@
 Mean A was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table 2).
-Additionally, leaf N~a~ was positively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).
+Additionally, leaf N~a~ was positively related to A across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
 Following an increase light intensity to match high-light conditions, A of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001).
@@
 On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table 2).
-In sun leaves, A was positively correlated with g~s~, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
+Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
 Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
-With increased PPFD, increases in A and g~s~ in shade leaves were positively correlated  as observed in sun leaves.
 
@@
 Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
-Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment but no warmding effect was detected in shade leaves measured at low or high light (Table 2).
+Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment, but no warmding effect was detected in shade leaves measured at low or high light (Table 2).
 For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal WUE model (Figure `r waterusefig`a).
@@
 ## Leaf carbon isotope discrimination and mesophyll conductance
-The observed carbon isotope discrimination was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`).
-Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta$~o~ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
+The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at liw light (Figure `r deltafig`).
+Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
+The remainder $\Delta_o$ consists of the contributions of stomatal conductance, respirations and photosrespiration to discrimination.
 No differences in g~s~ were detected with the warming treatment within leaf types and g~s~ did not vary significantly across measurement campaigns (Figure `r get1fig`b).
@@
 Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
-With increased PPFD, g~m~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (Table 2).
-Mesophyll conductance was nearly proportional to A within leaf types and light treatments (Figure `r gebarfig`b,c).
+Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table 2).
+Proportional increases in g~m~ were matched by proportional increases in A from low to high light in shade leaves (Figure `r gebarfig`b,c).
 Photosynthesis scaled positively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b).
@@
 ## Variation in intercellular and chloroplastic CO~2~ concentrations
-More open stomata in shade leaves under low and high light lead to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
+Higher rates of g~s~ in shade leaves under low and high light lead to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
 The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and A [@caemmerer2014temperature].
@@
 More open stomata and rapid increases in g~m~ allowed shade leaves to readily utilize increases in light availability, leading to rates of leaf C gain exceeding sun leaves.
-Although we know shade leaves experience interchanging periods of sun and shade [@pearcy1990sunflecks], these findings are unexpected because a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
+Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], these findings are unexpected because a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
 This dynamic behaviour of shade leaves complicates optimizations of canopy C uptake and accounting for the response of leaf physiology to transient light dynamics is an essential need in future studies.
@@
 # List of Figures
-**Figure 1**. Daily maximum and minimum temperature (a) and total daily PPFD (b) for each chamber across the experiment duration.
 
-\
-**Figure 2**.
+**Figure 1**.
 Local light environment for sun and shade leaves for each monthly gas exchange campaign.
@@
 \
-**Figure 3**.
+**Figure 2**.
 (a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.
-ACi curves were measured once on all trees, before the initiation of the drought treatment, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
 (b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light.
-(c) The relationship between A and leaf N~a~ for each sun and shade leaf.
+(c) The relationship between A and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
 For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
@@
 \
-**Figure 4**.
+**Figure 3**.
 The response of A to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
@@
 \
-**Figure 5**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
+**Figure 4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
 
 \
-**Figure 6**.
+**Figure 5**.
 (a) Response of instantaneous transpiration efficiency (ITE) to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
@@
 \
-**Figure 7**.
+**Figure 6**.
 Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis and measured C~i~/C~a~ for sun leaves measured at hight light and shade leaves measured at both low and high light.
@@
 \
-**Figure 8**.
+**Figure 7**.
 The mean &pm; 1 standard error of intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light.
@@
 \
-**Figure S1**.
+**Figure S1**. Daily maximum and minimum temperature (a) and total daily PPFD (b) for each chamber across the experiment duration.
+
+\
+**Figure S2**.
 Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
@@
 \
-**Figure S2**. Response of A (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign.
+**Figure S3**. Response of A (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign.
 
 \
-**Figure S3**. Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign.
+**Figure S4**. Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign.
 
@@
 # Figures {.unnumbered}
-```{r airvars, fig.width=7, fig.height=9, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/air_vars.R")
-```
-
-**Figure 1**
-
 ```{r par, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
@@
 
+```{r airvars, fig.width=7, fig.height=9, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/air_vars.R")
+```
+
+**Figure S1**
+
 ```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
@@
 
-**Figure S1**.
+**Figure S2**.
 
@@
 
-**Figure S2**.
+**Figure S3**.
 
@@
 
-**Figure S3**.
+**Figure S4**.
 

### 2015-11-16 master_scripts/plot_objects.R

@@ -28,7 +28,7 @@ condlab <- expression(g[s]~~(mol~m^-2~s^-1))
 totcondlab <- expression(Total~Conductance~to~CO[2]~~(mol~m^-2~s^-1))
 
 vpdlab <- "VPD  (kPa)"
-itelab <- expression(ITE~~(mu*mol~CO[2]~mmol~H[2]*O^-1))
+itelab <- expression(ITE~~(mmol~CO[2]~mol~H[2]*O^-1))
 
 ratelab <- expression(mol~m^-2~s^-1)
 parlab <- expression(PPFD~~(mu*mol~m^-2~s^-1))

### 2015-11-20 manuscript/manuscript_wtc3.Rmd
MT comments

@@
 We argue that the rapid response of g~m~ with light enables shade leaves to respond quickly to sunflecks.
-Evidence that g~m~ not only varies within a canopy but can be up-regulated over short time intervals complicates including g~m~ in process-based models of canopy carbon gain
+Evidence that g~m~ not only varies within a canopy but can be up-regulated over short time intervals possibly represents a new mechanism underpinning leaf gas exchange responses to light.
 
@@
 # Introduction
-Understanding and predicting carbon (C) uptake in forest ecosystems is crucially important for assessing the impacts of environmental change.
-Specifically, this involves accurate upscaling of leaf level measurements to predict whole canopy function.
+Light availability is one of the most important environmental drivers of leaf carbon (C) uptake in trees.
+Predicting C uptake if forests usually involves upscaling leaf level measurements to assess whole canopy function.
 Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling based on solar irradiance problematic.
@@
 \
-During A, CO~2~ diffuses from the atmosphere through stomata (g~s~), intercellular air space and into the chloroplast for fixation by Rubisco (mesophyll conductance, g~m~) [@tazoe2011using].
-Based on optimality theory, stomatal conductance should be efficiently distributed within a canopy to utilize supplies of light, N and water to maximize A [@peltoniemi2012co].
+Rates of photosynthesis in C~3~ plants are limited by the [CO~2~] available for fixation by Rubisco within the chlorplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmoshpehre to the site of carboxylation [@warren2008stand].
+This drawdown consists of multiple resistance pathways to CO~2~ diffusion which include CO~2~ diffusion from the atmosphere through stomata (stomatal conductance, g~s~) and then from these intercellular air spaces into the chloroplast (mesophyll conductance, g~m~).
+Based on optimality theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize A [@peltoniemi2012co].
 This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
@@
 How shade leaves utilize sunflecks for short term C gain depends on the combined response time of g~s~ and g~m~ and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks;@tausz2005dynamic].
-For example, at short timescales the utilization of sunflecks is limited by delayed responses of stomata, effectively limiting the maximum assimilation rate that can be achieved [@vico2011effects; @way2012sunflecks].
+The utilization of sunflecks is first limited by delayed responses of stomata opening, which may take minutes, effectively limiting the maximum assimilation rate that can be achieved [@pearcy1990sunflecks; @vico2011effects; @way2012sunflecks].
 Mesophyll conductance has been shown to respond to environmental factors (e.g.
-CO~2~, temperature or vapor pressure deficit) across short time scales, possibly faster than g~s~ [@flexas2008mesophyll and references therein], yet short term response to light availability are unclear.
+CO~2~, temperature or vapor pressure deficit) at timescales of minutes, possibly faster than g~s~ [@flexas2008mesophyll and references therein], yet short term response to light availability are unclear.
 For example, g~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid].

### 2015-11-23 manuscript/manuscript_wtc3.Rmd
comments MT

@@
 \
-Climate warming may also potentially affect the physiological behaviour of leaves within a canopy.
-This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) scale positively with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis].
+Climate warming may also affect the physiological behaviour of leaves within a canopy.
+This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis].
 How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for upscaling A for sun and shade leaves.
@@
 \
-1.	If whole tree canopies are optimized for C gain, then leaf N, K and photosynthetic capacity are predicted to be higher in sun leaves compared to shade leaves.
+1.	If whole tree canopies are optimized for C gain, then leaf N, hydraulic conductance and biochemical photosynthetic capacity are predicted to be higher in sun leaves compared to shade leaves.
 
@@
 \
-3.	As shade leaves are constrained by their underlying biochemistry and slow physiological responses, increases in A following sunfleck simulations were not expected to reach rates of full sun leaves.
+3.
+As shade leaves were expected to develop lower biochemical photosynthetic capacity and leaf physiolgoical responses to light are typically slow (minutes), increases in A following sunfleck simulations were not expected to reach rates of full sun leaves.
 
 \
-4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months.
+4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporative demand from higher temperatures will lead to stomatal closure.
 
@@
 
-where R~s~ is the isotopic ratio of the sample and R~stnd~ is the isotopic ratio of the standard Vienna Pee Dee Belemnite.
+where R~s~ is the isotopic ratio of the sample and R~stnd~ is the isotopic ratio of the standard Vienna Pee Dee Belemnite (VPDP).
 Next, the observed discrimination ($\Delta$~obs~) is calculated from Evans et al.
@@
 The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression, where k=0.5 [see @duursma2013near].
+Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use [@lin2015optimal].
 
@@
 For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package, using WTC as a random effect.
+All tests of statistical significance were conducted at an alpha of 0.05.
+
 # Results
@@
 **Table 1**.
-*Eucalyptus tereticornis* leaf traits and the distribution of resources between full sun and shade leaves under ambient and elevated temperature treatments.
+*Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments.
 Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
@@
 Different letters represent significant differences between leaf type and temperature treatments.
-The P value represents the overall effect of leaf type and temperature treatment on each trait.
+The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.
 
@@
 Different letters represent significant differences between leaf type, light environment and temperature treatments.
-The P value represents the overall effect of leaf type, light environment and temperature treatment on each parameter.
+The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.
 

### 2015-11-23 manuscript/manuscript_wtc3.Rmd
comments MT

@@
 In this study, variation in leaf N distribution and A rates were not associated with subsequent changes in leaf K between sun and shade leaves.
-As both sun and shade leaves were acclimated to their local light environment this sub-optimal distribution of water supply has the potential to diminish the C return for the whole canopy.
+As these trees were kept well watered, it appears that water supply was adequate and likely non-limiting to both sun and shade leaves.
 

### 2015-11-24 manuscript/manuscript_wtc3.Rmd
edits MT...need to look at stats for supp figures

@@
 Among the sampled leaves, Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
-As g~m~ (results below) is implicit in the estimates of J~max~ and Vc~max~ with conventional AC~i~ curves, we also simulated AC~c~ curves to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.
-When accounting for the effect of g~m~, chloroplastic Vc~max~ and J~max~ were still higher in sun leaves (Figure `r accfig`).
 
@@
 Additionally, leaf N~a~ was positively related to A across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
-Following an increase light intensity to match high-light conditions, A of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001).
+Following an increase in light intensity to match high-light conditions, A of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table 2).
 No effect of the warming treatment was detected on A of sun leaves measured at high light or shade leaves at either low or high light.
@@
 Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
+No effect of the warming treatment was detected on g~s~ within or across leaf types.
+Stomatal conductance within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get2fig`b).
 
@@
 Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
-Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment, but no warmding effect was detected in shade leaves measured at low or high light (Table 2).
-For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal WUE model (Figure `r waterusefig`a).
+Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment, but no warming effect was detected in shade leaves measured at low or high light (Table 2).
 The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
+For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal ITE model with a g1 value for each leaf type and treatment (Figure `r waterusefig`a).
 Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).
@@
 Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
-The remainder $\Delta_o$ consists of the contributions of stomatal conductance, respirations and photosrespiration to discrimination.
-No differences in g~s~ were detected with the warming treatment within leaf types and g~s~ did not vary significantly across measurement campaigns (Figure `r get1fig`b).
+The remainder $\Delta_o$ consists of the contributions of g~s~, respiration and photorespiration to discrimination.
 
@@
 No differences in g~m~ were detected with the warming treatment within leaf types.
-Additionally, g~m~ did not vary significantly across measurements campaigns (Figure `r get2fig`b) and only a weak negative relationship with increasing leaf temperature was detected (Figure `r get1fig`b).
+Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments, but a weak negative relationship with increasing leaf temperature was detected with leaf types under local light conditions (Figure `r get1fig`b).
+We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~.
+Comparison of A-Cc curves (Figure `r accfig`) and A-Ci curves revealed similar differences between sun and shade leaves.
 
@@
 Higher rates of g~s~ in shade leaves under low and high light lead to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
+The chloroplast CO~2~ partial pressure was comparable between shade leaves when measured at both low and high light conditions (Figure `r ciccfig`c).
+In sun leaves C~c~ was significantly lower than shade leaves as a likely consequence of a lower C~i~.
 The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and A [@caemmerer2014temperature].
-This drawdown was similar between sun and shade leaves at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
+This drawdown was similar between sun and shade leaves measured at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
 This was the result of the proportional relationship between g~m~ and A across all leaves.
@@
 Here we show that leaves within canopies of *Eucalyptus tereticornis* trees maintain the ability to upregulate A beyond the limitations of their local light environment.
+
 More open stomata and rapid increases in g~m~ allowed shade leaves to readily utilize increases in light availability, leading to rates of leaf C gain exceeding sun leaves.
+
 Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], these findings are unexpected because a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
+
 This dynamic behaviour of shade leaves complicates optimizations of canopy C uptake and accounting for the response of leaf physiology to transient light dynamics is an essential need in future studies.
+
 Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology.
@@
 In this study, variation in leaf N distribution and A rates were not associated with subsequent changes in leaf K between sun and shade leaves.
-As these trees were kept well watered, it appears that water supply was adequate and likely non-limiting to both sun and shade leaves.
+Thus, no direct relationship between water supply and N distribution or A within the canopy were detected.
 
@@
 Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased ITE throughout the experiment.
-Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was prevalent across the leaf lifespan.
+Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was likely prevalent long term.
 From a canopy perspective this pattern in water use efficiency initially appears to be detrimental to C gain as A in sun leaves was characterized by low rates of g~s~ and low C~i~.
-Relative to the differences in A between leaf types, higher rates of g~s~ in shade leaves appear to exhibit wasteful water loss.
+Relative to the differences in A between leaf types, higher rates of g~s~ in shade leaves appear to exhibit inefficient water use.
 As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower ITE compared to sun leaves.
@@
 ## Physiological behaviour of sun and shade leaves
-The pattern of wasteful water use in shade leaves is important as we hypothesized that g~s~ and A would be proportional across sun and shade leaves.
+The pattern of inefficient water use in shade leaves is important as we hypothesized that g~s~ and A would be proportional across sun and shade leaves.
 In sun leaves, A and g~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
 However, lower rates of A in shade leaves were not coupled with decreases in g~s~, explaining the observed decreases in WUE.
-This is significant as optimal stomatal behaviour has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on sun leaves [see @prentice2014balancing; @lin2015optimal].
+This is significant as optimal stomatal regulation to balance C gain with water loss has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on full sun leaves [see @prentice2014balancing; @lin2015optimal].
 As a result, the often used economic framework of balancing costs of using water versus N allocation to predict A [@wright2003least] may break down when considering all leaves within a tree canopy.
@@
 This is because rapid increases in leaf temperature with sunflecks have been shown to inhibit C gain [@leakey2003high].
-However, this response likely occurs at already high air temperatures and may not explain the observed g~s~ in shade leaves across the large range of temperatures included in this study.
+However, this response likely occurs at very high air temperatures and may not explain the observed g~s~ in shade leaves across the large natural range of temperatures included in this study.
 How prevalent each of these strategies are within tree canopies is still unknown, as empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants, often in deep shade [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution].
@@
 If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan.
-For example, the potential C gain in leaves which experience high amounts of sunflecks may be large enough to tolerate excess losses from transpiration.
+For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to tolerate decreases in WUE.
 However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.
-Thus, models which predict leaf photosynthesis from N distribution within a canopy will be flawed unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].
+Thus, models which predict leaf photosynthesis from N distribution within a canopy will be incomplete unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].
 
@@
 Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy.
-Although the distribution of resources, such as N and water, define leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
+Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
 These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have important implications for process based models that predict canopy C gain from rates of leaf photosynthesis.

### 2015-11-30 manuscript/manuscript_wtc3.Rmd
comments from MT

@@
 * Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves.
-The transient nature of light availability within a canopy, however, complicates our ability to assess the contribution of shade leaves to canopy carbon gain.
-* Leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis (A), stomatal conductance (g~s~) and mesophyll conductance (g~m~) of sun and shade leaves in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
+A lack of empirical data relating photosynthesis to leaf physiolgoical behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
+* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis (A), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
 * Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
@@
 PPFD in the cuvette was set to match the individual light environment of each leaf type (explained above).
-Sun flecks were simulated for shade leaves by increasing the leaf cuvette PPFD (LI-COR red/blue light source) to match the light environment of the full sun leaf in the same tree.
-The maximum sunfleck response was recorded once CO~2~ and water vapor fluxes re-stabilized in the leaf cuvette (ca.
+Periods of high irradiance were simulated for shade leaves by increasing the leaf cuvette PPFD (LI-COR red/blue light source) to match the light environment of the full sun leaf in the same tree.
+The maximum sunfleck response of shade leaves was then recorded once CO~2~ and water vapor fluxes re-stabilized in the leaf cuvette (ca.
 25 min).
@@
 Proportional increases in g~m~ were matched by proportional increases in A from low to high light in shade leaves (Figure `r gebarfig`b,c).
-Photosynthesis scaled positively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b).
+Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186).
+The large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b).
 No differences in g~m~ were detected with the warming treatment within leaf types.
-Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments, but a weak negative relationship with increasing leaf temperature was detected with leaf types under local light conditions (Figure `r get1fig`b).
+Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively).
 We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~.
@@
 # Discussion
-Here we show that leaves within canopies of *Eucalyptus tereticornis* trees maintain the ability to upregulate A beyond the limitations of their local light environment.
-
-More open stomata and rapid increases in g~m~ allowed shade leaves to readily utilize increases in light availability, leading to rates of leaf C gain exceeding sun leaves.
-
-Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], these findings are unexpected because a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
-
-This dynamic behaviour of shade leaves complicates optimizations of canopy C uptake and accounting for the response of leaf physiology to transient light dynamics is an essential need in future studies.
-
+Here we show that A in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
+Although shade leaves in lower light environments exhibited relatively high g~s~ and c~i~, it was rapid increases in g~m~ under periods of high light availability that led to this upregulation of A.
+Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
+These findings offer new insights into how aspects of leaf physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake in future studies.
 Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology.
@@
 If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution.
+This is because the role of g~s~ in regulating photosynthetic induction impacts the capacity of a leaf to utilize sunflecks [@way2012sunflecks].
 If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan.

### 2015-11-30 master_scripts/gasex_temp3.R

@@ -73,6 +73,11 @@ meandraw <-summaryBy(drawdown ~ leaflight, data=gasex_agg, FUN=mean)
 gmt_sun_mod <- lm(gm~ CTleaf , data=gasex_agg, subset=leaflight=="sun-high")
 gmt_sha_mod <- lm(gm~ CTleaf , data=gasex_agg,subset=leaflight=="shade-low")
 gmt_fleck_mod <- lm(gm~ CTleaf , data=gasex_agg,subset=leaflight=="shade-high")
+
+summary(gmt_sun_mod)
+summary(gmt_sha_mod)
+summary(gmt_fleck_mod)
+
 coefsun <- coef(gmt_sun_mod)
 coefsha <- coef(gmt_sha_mod)
 coeffleck <- coef(gmt_fleck_mod)

### 2015-12-01 manuscript/manuscript_wtc3.Rmd
lma in discussion
new par figure

regression on gases appendix fig

@@
 * Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves.
-A lack of empirical data relating photosynthesis to leaf physiolgoical behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
+A lack of empirical data relating photosynthesis to leaf physiological behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
 * To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis (A), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
@@
 \
-Rates of photosynthesis in C~3~ plants are limited by the [CO~2~] available for fixation by Rubisco within the chlorplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmoshpehre to the site of carboxylation [@warren2008stand].
+Rates of photosynthesis in C~3~ plants are limited by the [CO~2~] available for fixation by Rubisco within the chloroplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmosphere to the site of carboxylation [@warren2008stand].
 This drawdown consists of multiple resistance pathways to CO~2~ diffusion which include CO~2~ diffusion from the atmosphere through stomata (stomatal conductance, g~s~) and then from these intercellular air spaces into the chloroplast (mesophyll conductance, g~m~).
-Based on optimality theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize A [@peltoniemi2012co].
+Based on optimal theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize A [@peltoniemi2012co].
 This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
@@
 3.
-As shade leaves were expected to develop lower biochemical photosynthetic capacity and leaf physiolgoical responses to light are typically slow (minutes), increases in A following sunfleck simulations were not expected to reach rates of full sun leaves.
+As shade leaves were expected to develop lower biochemical photosynthetic capacity and leaf physiological responses to light are typically slow (minutes), increases in A following sunfleck simulations were not expected to reach rates of full sun leaves.
 
 \
-4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporative demand from higher temperatures will lead to stomatal closure.
+4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporation demand from higher temperatures will lead to stomatal closure.
 
@@
 ## Leaf carbon isotope discrimination and mesophyll conductance
-The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at liw light (Figure `r deltafig`).
+The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
 Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
@@
 Here we show that A in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
-Although shade leaves in lower light environments exhibited relatively high g~s~ and c~i~, it was rapid increases in g~m~ under periods of high light availability that led to this upregulation of A.
+Although shade leaves in lower light environments exhibited relatively high g~s~ and c~i~, it was rapid increases in g~m~ under periods of high light availability that led to this up-regulation of A.
 Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
@@
 Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and A were all reduced in shade leaves.
+Leaf mass area, however, was not different between sun and shade leaves.
+This could be due to leaf formation under comparative light conditions or possible differences in total non-structural carbohydrates contents between leaf types.
 
@@
 **Figure 1**.
-Local light environment for sun and shade leaves for each monthly gas exchange campaign.
+Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014.
 Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf.
+Each date represents the starting date for each measurement campaign.
+Points represent the maximum daily air temperature during each campaign period.
 
@@
 **Figure 6**.
-Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis and measured C~i~/C~a~ for sun leaves measured at hight light and shade leaves measured at both low and high light.
+Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
 The solid line represents the theoretical line for C3 plants from Evans et al.
@@
 \
-**Figure S3**. Response of A (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign.
+**Figure S3**.
+Response of A (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
+Shaded symbols represents each monthly measurement campaign.
+Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
+All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
+Weak negative relationships with g~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (R^2^ = 0.16 and 0.08, respectively).
 
 \
-**Figure S4**. Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign.
+**Figure S4**.
+Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
+Shaded symbols represents each monthly measurement campaign.
+Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
+All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
+Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (R^2^ = 0.73, 0.58 and 0.72, respectively)
 
@@
 ```{r par, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/par.R")
+source("master_scripts/par2.R")
 ```
@@
 
-```{r gasex_temp1, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/gasex_temp.R")
+```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/gasex_temp3.R")
 ```
@@
 
-```{r gasex_temp2, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/gasex_temp2.R")
+```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/gasex_temp4.R")
 ```

### 2015-12-01 functions and packages/functions.R

@@ -427,3 +427,32 @@ add_trend_line <- function(x,y,d,col, ...){
              x1=min(d[[x]]), x2=max(d[[x]]), ...)
 }
 #------------------------------------------------------------------------
+
+###mtexti for reading direction arrangement
+mtexti <- function(text, side, off = 0.25,
+                   srt = if(side == 2) 90  else
+                     if(side == 4) 270 else 0, 
+                   outer=TRUE,cex=1,
+                   ...) {
+  # dimensions of plotting region in user units
+  usr <- par('usr')
+  # dimensions of plotting region in inches
+  pin <- par('pin')
+  # user units per inch
+  upi <- c(usr[2]-usr[1],
+           usr[4]-usr[3]) / pin
+  # default x and y positions
+  xpos <- (usr[1] + usr[2])/2
+  ypos <- (usr[3] + usr[4])/2
+  if(1 == side)
+    ypos <- usr[3] - upi[2] * off
+  if(2 == side)
+    xpos <- usr[1] - upi[1] * off
+  if(3 == side)
+    ypos <- usr[4] + upi[2] * off
+  if(4 == side)
+    xpos <- usr[2] + upi[1] * off
+  text(x=xpos, y=ypos, text, xpd=NA, srt=srt, cex=cex,...)
+}
+
+

### 2015-12-01 master_scripts/gasex_temp4.R

@@ -1,0 +1,167 @@
+# source("functions and packages/functions.R")
+# source("master_scripts/plot_objects.R")
+# source("functions and packages/packages.R")
+
+#read in gm data set (no drought) and Cibar(discrimination)-------------------------------------------------------
+gasex<- read.csv("calculated_data/gmes_wellwatered.csv")
+##calculate CC
+gasex$Cc<- with(gasex, Ci-Photo/gm)
+##calculate drawdown
+gasex$drawdown <- with(gasex, Ci-Cc)
+gasex$drawdowngs <- with(gasex, CO2R-Ci)
+
+###get average by id
+gasex_agg <- summaryBy(Photo+ gm + drawdown +CTleaf+VpdL+Cond+drawdowngs ~ chamber+id+leaf +light+temp+leaflight+Month, data=gasex, 
+                       FUN=mean, keep.names=TRUE)
+
+####show data as a function of campaign (lets shade col by campaign)
+
+##lets order by month
+Morder <- c("Oct", "Dec", "Jan", "Feb", "Mar", "Apr")
+gasex_agg$Month <- factor(gasex_agg$Month, levels = Morder)
+
+###shade black colors for each month to test
+
+##sun col
+octcol <- alpha(suncol2, alpha=.4)
+deccol <- alpha(suncol2, alpha=.55)
+jancol <- alpha(suncol2, alpha=.65)
+febcol <- alpha(suncol2, alpha=.75)
+marcol <- alpha(suncol2, alpha=.9)
+aprcol <- suncol2
+##shad cols
+octcol2 <- alpha(shacol2, alpha=.4)
+deccol2<- alpha(shacol2, alpha=.55)
+jancol2 <- alpha(shacol2, alpha=.65)
+febcol2 <- alpha(shacol2, alpha=.75)
+marcol2 <- alpha(shacol2, alpha=.9)
+aprcol2 <- shacol2
+#fleck col
+octcol3 <- alpha(lightscol2, alpha=.4)
+deccol3<- alpha(lightscol2, alpha=.55)
+jancol3 <- alpha(lightscol2, alpha=.65)
+febcol3 <- alpha(lightscol2, alpha=.75)
+marcol3 <- alpha(lightscol2, alpha=.9)
+aprcol3 <- lightscol2
+
[…]

### 2015-12-01 manuscript/manuscript_wtc3.Rmd
abbreviation updates to manuscript

@@
 A lack of empirical data relating photosynthesis to leaf physiological behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
-* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis (A), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
+* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis (*A*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
 * Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
@@
 Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor].
-Consequently, structural and functional properties of leaves are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of photosynthesis (A) [@evans1995carbon].
+Consequently, structural and functional properties of leaves are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of photosynthesis (*A*) [@evans1995carbon].
 To estimate whole canopy C gain it is thus necessary to account for the non-linear response of A to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
@@
 \
-The distribution of resources required for A, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients.
-As A has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
+The distribution of resources required for *A*, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients.
+As *A* has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
 Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic].
 Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
-Higher rates of A and stomatal conductance (g~s~) can only be sustained if the hydraulic conductance (K) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal].
-Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing A if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co], thus K should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
+Higher rates of *A* and stomatal conductance (g~s~) can only be sustained if the hydraulic conductance (K) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal].
+Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A* if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co], thus K should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
 
@@
 This drawdown consists of multiple resistance pathways to CO~2~ diffusion which include CO~2~ diffusion from the atmosphere through stomata (stomatal conductance, g~s~) and then from these intercellular air spaces into the chloroplast (mesophyll conductance, g~m~).
-Based on optimal theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize A [@peltoniemi2012co].
+Based on optimal theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize *A* [@peltoniemi2012co].
 This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
-Mesophyll conductance can also impose limitations on A as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in A [@niinemets2007photosynthesis] if g~m~ constrains CO~2~ supply to the chloroplast.
+Mesophyll conductance can also impose limitations on *A* as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A* [@niinemets2007photosynthesis] if g~m~ constrains CO~2~ supply to the chloroplast.
 Part of the variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
-Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack empirical data currently hinders are ability to interpreting their coupled responses to A across sun and shade leaves.
+Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack empirical data currently hinders are ability to interpreting their coupled responses to *A* across sun and shade leaves.
 
@@
 This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis].
-How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for upscaling A for sun and shade leaves.
+How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for upscaling *A* for sun and shade leaves.
 Additionally, light saturated rates of A are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
@@
 \
-2.	Stomatal conductance is proportional to A across sun and shade leaves under similar leaf VPD and g~m~ scales positively with photosynthetic capacity.
+2.	Stomatal conductance is proportional to *A* across sun and shade leaves under similar leaf VPD and g~m~ scales positively with photosynthetic capacity.
 
@@
 3.
-As shade leaves were expected to develop lower biochemical photosynthetic capacity and leaf physiological responses to light are typically slow (minutes), increases in A following sunfleck simulations were not expected to reach rates of full sun leaves.
+As shade leaves were expected to develop lower biochemical photosynthetic capacity and leaf physiological responses to light are typically slow (minutes), increases in *A* following sunfleck simulations were not expected to reach rates of full sun leaves.
 
@@
 \
-The CO~2~ diffusion from the intercellular airspace to the chloroplast, g~m~, is given by its relationship to the CO~2~ assimilation rate (A) by:
+The CO~2~ diffusion from the intercellular airspace to the chloroplast, g~m~, is given by its relationship to the CO~2~ assimilation rate (*A*) by:
 
@@
 (11)
-$$g_m = \frac{A}{C_i-C_c}$$
+$$g_m = \frac{*A*}{C_i-C_c}$$
 
@@
 All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
-Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K, mmol m^-2^ s^-1^ MPa) through the equation:
+Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) through the equation:
 
@@
 Across six measurement campaigns over the 7 month period PPFD was  reduced on average by >75% in the shade (Figure `r parfig`).
-Leaf-specific hydraulic conductance (K) was similar across sun and shade leaves (Table 1).
+Leaf-specific hydraulic conductance (K~l~) was similar across sun and shade leaves (Table 1).
 This was because $\Psi_{pd}$, $\Psi_l$ (Table 1) and E (Table 2) did not differ between leaf types.
@@
 Leaf mass per area was not different between leaf types.
-No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K, E, N~area~ or LMA either within or across leaf types.
+No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~area~ or LMA either within or across leaf types.
 
@@
 \
-Mean A was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table 2).
-Additionally, leaf N~a~ was positively related to A across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
+Mean *A* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table 2).
+Additionally, leaf N~a~ was positively related to *A* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
 Following an increase in light intensity to match high-light conditions, A of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table 2).
-No effect of the warming treatment was detected on A of sun leaves measured at high light or shade leaves at either low or high light.
+No effect of the warming treatment was detected on *A* of sun leaves measured at high light or shade leaves at either low or high light.
 Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
@@
 On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table 2).
-Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
+Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and *A* were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
 Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
@@
 The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
-For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal ITE model with a g1 value for each leaf type and treatment (Figure `r waterusefig`a).
+For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal ITE model with a g~1~ value for each leaf type and treatment (Figure `r waterusefig`a).
 Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).
@@
 Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table 2).
-Proportional increases in g~m~ were matched by proportional increases in A from low to high light in shade leaves (Figure `r gebarfig`b,c).
+Proportional increases in g~m~ were matched by proportional increases in *A* from low to high light in shade leaves (Figure `r gebarfig`b,c).
 Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186).
-The large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b).
+The large increases in g~m~ in shade leaves under high light resulted in the highest rates of *A* (Figure `r agmgsfig`b).
 No differences in g~m~ were detected with the warming treatment within leaf types.
@@
 In sun leaves C~c~ was significantly lower than shade leaves as a likely consequence of a lower C~i~.
-The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and A [@caemmerer2014temperature].
+The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and *A* [@caemmerer2014temperature].
 This drawdown was similar between sun and shade leaves measured at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
-This was the result of the proportional relationship between g~m~ and A across all leaves.
+This was the result of the proportional relationship between g~m~ and *A* across all leaves.
 The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
@@
 # Discussion
-Here we show that A in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
-Although shade leaves in lower light environments exhibited relatively high g~s~ and c~i~, it was rapid increases in g~m~ under periods of high light availability that led to this up-regulation of A.
+Here we show that *A* in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
+Although shade leaves in lower light environments exhibited relatively high g~s~ and c~i~, it was rapid increases in g~m~ under periods of high light availability that led to this up-regulation of *A*.
 Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
@@
 ## Resource distribution and photosynthetic capacity
-The allocation of N~a~ constrains A and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
+The allocation of N~a~ constrains *A* and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
 Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979).
 As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical].
-Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and A were all reduced in shade leaves.
+Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and *A* were all reduced in shade leaves.
 Leaf mass area, however, was not different between sun and shade leaves.
@@
 Using a two-leaf model, Peltoniemi et al.
-[-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if leaf K is also optimally distributed.
-In this study, variation in leaf N distribution and A rates were not associated with subsequent changes in leaf K between sun and shade leaves.
-Thus, no direct relationship between water supply and N distribution or A within the canopy were detected.
+[-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if K~l~ is also optimally distributed.
+In this study, variation in leaf N distribution and *A* rates were not associated with subsequent changes in K~l~ between sun and shade leaves.
+Thus, no direct relationship between water supply and N distribution or *A* within the canopy were detected.
 
@@
 Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was likely prevalent long term.
-From a canopy perspective this pattern in water use efficiency initially appears to be detrimental to C gain as A in sun leaves was characterized by low rates of g~s~ and low C~i~.
-Relative to the differences in A between leaf types, higher rates of g~s~ in shade leaves appear to exhibit inefficient water use.
+From a canopy perspective this pattern in water use efficiency initially appears to be detrimental to C gain as *A* in sun leaves was characterized by low rates of g~s~ and low C~i~.
+Relative to the differences in *A* between leaf types, higher rates of g~s~ in shade leaves appear to exhibit inefficient water use.
 As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower ITE compared to sun leaves.
@@
 ## Physiological behaviour of sun and shade leaves
-The pattern of inefficient water use in shade leaves is important as we hypothesized that g~s~ and A would be proportional across sun and shade leaves.
-In sun leaves, A and g~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
-However, lower rates of A in shade leaves were not coupled with decreases in g~s~, explaining the observed decreases in WUE.
+The pattern of inefficient water use in shade leaves is important as we hypothesized that g~s~ and *A* would be proportional across sun and shade leaves.
+In sun leaves, *A* and g~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
+However, lower rates of *A* in shade leaves were not coupled with decreases in g~s~, explaining the observed decreases in ITE.
 This is significant as optimal stomatal regulation to balance C gain with water loss has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on full sun leaves [see @prentice2014balancing; @lin2015optimal].
-As a result, the often used economic framework of balancing costs of using water versus N allocation to predict A [@wright2003least] may break down when considering all leaves within a tree canopy.
+As a result, the often used economic framework of balancing costs of using water versus N allocation to predict *A* [@wright2003least] may break down when considering all leaves within a tree canopy.
 
@@
 \
-We found that A and g~m~ scaled positively across leaf types and, surprisingly, increased rapidly (within minutes) and proportionately when light intensity was increased in shade leaves.
+We found that *A* and g~m~ scaled positively across leaf types and, surprisingly, increased rapidly (within minutes) and proportionately when light intensity was increased in shade leaves.
 Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
@@
 If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan.
-For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to tolerate decreases in WUE.
-However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.
+For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to tolerate decreases in ITE However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.
 Thus, models which predict leaf photosynthesis from N distribution within a canopy will be incomplete unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].
@@
 *Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments.
-Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
+Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
 Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
-Units of LMA and Leaf Nitrogen~area~ are g m^2^, K is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;.
+Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;.
 Different letters represent significant differences between leaf type and temperature treatments.

### 2015-12-03 manuscript/manuscript_wtc3.Rmd
A abbr, K abbreviation, Water-use and model language double checked

@@
 A lack of empirical data relating photosynthesis to leaf physiological behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
-* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis (*A*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
+* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
 * Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
@@
 Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor].
-Consequently, structural and functional properties of leaves are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of photosynthesis (*A*) [@evans1995carbon].
+Consequently, structural and functional properties of leaves are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of leaf photosynthesis (*A~n~*) [@evans1995carbon].
 To estimate whole canopy C gain it is thus necessary to account for the non-linear response of A to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
@@
 \
-The distribution of resources required for *A*, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients.
-As *A* has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
+The distribution of resources required for *A~n~*, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients.
+As *A~n~* has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
 Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic].
 Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
-Higher rates of *A* and stomatal conductance (g~s~) can only be sustained if the hydraulic conductance (K) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal].
-Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A* if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co], thus K should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
+Higher rates of *A~n~* and stomatal conductance (g~s~) can only be sustained if the leaf specific hydraulic conductance (K~l~) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal].
+Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A~n~* if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co], thus K~l~ should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
 
@@
 This drawdown consists of multiple resistance pathways to CO~2~ diffusion which include CO~2~ diffusion from the atmosphere through stomata (stomatal conductance, g~s~) and then from these intercellular air spaces into the chloroplast (mesophyll conductance, g~m~).
-Based on optimal theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize *A* [@peltoniemi2012co].
+Based on optimal theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize *A~n~* [@peltoniemi2012co].
 This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
-Mesophyll conductance can also impose limitations on *A* as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A* [@niinemets2007photosynthesis] if g~m~ constrains CO~2~ supply to the chloroplast.
+Mesophyll conductance can also impose limitations on *A~n~* as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A~n~* [@niinemets2007photosynthesis] if g~m~ constrains CO~2~ supply to the chloroplast.
 Part of the variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
-Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack empirical data currently hinders are ability to interpreting their coupled responses to *A* across sun and shade leaves.
+Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack empirical data currently hinders are ability to interpreting their coupled responses to *A~n~* across sun and shade leaves.
 
@@
 This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis].
-How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for upscaling *A* for sun and shade leaves.
+How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for upscaling *A~n~* for sun and shade leaves.
 Additionally, light saturated rates of A are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
@@
 \
-2.	Stomatal conductance is proportional to *A* across sun and shade leaves under similar leaf VPD and g~m~ scales positively with photosynthetic capacity.
+2.	Stomatal conductance is proportional to *A~n~* across sun and shade leaves under similar leaf VPD and g~m~ scales positively with photosynthetic capacity.
 
@@
 3.
-As shade leaves were expected to develop lower biochemical photosynthetic capacity and leaf physiological responses to light are typically slow (minutes), increases in *A* following sunfleck simulations were not expected to reach rates of full sun leaves.
+As shade leaves were expected to develop lower biochemical photosynthetic capacity and leaf physiological responses to light are typically slow (minutes), increases in *A~n~* following sunfleck simulations were not expected to reach rates of full sun leaves.
 
@@
 \
-The CO~2~ diffusion from the intercellular airspace to the chloroplast, g~m~, is given by its relationship to the CO~2~ assimilation rate (*A*) by:
+The CO~2~ diffusion from the intercellular airspace to the chloroplast, g~m~, is given by its relationship to the leaf photosynthesis rate (*A~n~*) by:
 
@@
 (11)
-$$g_m = \frac{*A*}{C_i-C_c}$$
+$$g_m = \frac{*A~n~*}{C_i-C_c}$$
 
@@
 \
-Mean *A* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table 2).
-Additionally, leaf N~a~ was positively related to *A* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
+Mean *A~n~* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table 2).
+Additionally, leaf N~a~ was positively related to *A~n~* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
 Following an increase in light intensity to match high-light conditions, A of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table 2).
-No effect of the warming treatment was detected on *A* of sun leaves measured at high light or shade leaves at either low or high light.
+No effect of the warming treatment was detected on *A~n~* of sun leaves measured at high light or shade leaves at either low or high light.
 Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
 
-## Stomatal conductance and leaf water use efficiency
+## Stomatal conductance and leaf water-use efficiency
 On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table 2).
-Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and *A* were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
+Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and *A~n~* were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
 Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
@@
 \
-Bulk-leaf $\delta$^13^C, as an index of integrated water use efficiency [@marshall2007sources], was significantly lower in shade leaves compared to sun leaves by ca.
+Bulk-leaf $\delta$^13^C, as an index of integrated water-use efficiency [@marshall2007sources], was significantly lower in shade leaves compared to sun leaves by ca.
 2&permil; (Table 1).
@@
 Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table 2).
-Proportional increases in g~m~ were matched by proportional increases in *A* from low to high light in shade leaves (Figure `r gebarfig`b,c).
+Proportional increases in g~m~ were matched by proportional increases in *A~n~* from low to high light in shade leaves (Figure `r gebarfig`b,c).
 Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186).
-The large increases in g~m~ in shade leaves under high light resulted in the highest rates of *A* (Figure `r agmgsfig`b).
+The large increases in g~m~ in shade leaves under high light resulted in the highest rates of *A~n~* (Figure `r agmgsfig`b).
 No differences in g~m~ were detected with the warming treatment within leaf types.
@@
 In sun leaves C~c~ was significantly lower than shade leaves as a likely consequence of a lower C~i~.
-The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and *A* [@caemmerer2014temperature].
+The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and *A~n~* [@caemmerer2014temperature].
 This drawdown was similar between sun and shade leaves measured at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
-This was the result of the proportional relationship between g~m~ and *A* across all leaves.
+This was the result of the proportional relationship between g~m~ and *A~n~* across all leaves.
 The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
@@
 # Discussion
-Here we show that *A* in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
-Although shade leaves in lower light environments exhibited relatively high g~s~ and c~i~, it was rapid increases in g~m~ under periods of high light availability that led to this up-regulation of *A*.
+Here we show that *A~n~* in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
+Although shade leaves in lower light environments exhibited relatively high g~s~ and c~i~, it was rapid increases in g~m~ under periods of high light availability that led to this up-regulation of *A~n~*.
 Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
@@
 ## Resource distribution and photosynthetic capacity
-The allocation of N~a~ constrains *A* and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
+The allocation of N~a~ constrains *A~n~* and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
 Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979).
 As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical].
-Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and *A* were all reduced in shade leaves.
+Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and *A~n~* were all reduced in shade leaves.
 Leaf mass area, however, was not different between sun and shade leaves.
@@
 [-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if K~l~ is also optimally distributed.
-In this study, variation in leaf N distribution and *A* rates were not associated with subsequent changes in K~l~ between sun and shade leaves.
-Thus, no direct relationship between water supply and N distribution or *A* within the canopy were detected.
+In this study, variation in leaf N distribution and *A~n~* rates were not associated with subsequent changes in K~l~ between sun and shade leaves.
+Thus, no direct relationship between water supply and N distribution or *A~n~* within the canopy were detected.
 
@@
 Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was likely prevalent long term.
-From a canopy perspective this pattern in water use efficiency initially appears to be detrimental to C gain as *A* in sun leaves was characterized by low rates of g~s~ and low C~i~.
-Relative to the differences in *A* between leaf types, higher rates of g~s~ in shade leaves appear to exhibit inefficient water use.
+From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A~n~* in sun leaves was characterized by low rates of g~s~ and low C~i~.
+Relative to the differences in *A~n~* between leaf types, higher rates of g~s~ in shade leaves appear to exhibit inefficient water use.
 As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower ITE compared to sun leaves.
@@
 ## Physiological behaviour of sun and shade leaves
-The pattern of inefficient water use in shade leaves is important as we hypothesized that g~s~ and *A* would be proportional across sun and shade leaves.
-In sun leaves, *A* and g~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
-However, lower rates of *A* in shade leaves were not coupled with decreases in g~s~, explaining the observed decreases in ITE.
+The pattern of inefficient water use in shade leaves is important as we hypothesized that g~s~ and *A~n~* would be proportional across sun and shade leaves.
+In sun leaves, *A~n~* and g~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
+However, lower rates of *A~n~* in shade leaves were not coupled with decreases in g~s~, explaining the observed decreases in ITE.
 This is significant as optimal stomatal regulation to balance C gain with water loss has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on full sun leaves [see @prentice2014balancing; @lin2015optimal].
-As a result, the often used economic framework of balancing costs of using water versus N allocation to predict *A* [@wright2003least] may break down when considering all leaves within a tree canopy.
+As a result, the often used economic framework of balancing costs of using water versus N allocation to predict *A~n~* [@wright2003least] may break down when considering all leaves within a tree canopy.
 
@@
 \
-We found that *A* and g~m~ scaled positively across leaf types and, surprisingly, increased rapidly (within minutes) and proportionately when light intensity was increased in shade leaves.
+We found that *A~n~* and g~m~ scaled positively across leaf types and, surprisingly, increased rapidly (within minutes) and proportionately when light intensity was increased in shade leaves.
 Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
@@
 Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
-These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have important implications for process based models that predict canopy C gain from rates of leaf photosynthesis.
+These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have important implications for process-based models that predict canopy C gain from rates of leaf photosynthesis.
 Furthermore, the dynamic nature of g~m~ cannot be simply parameterized in tree growth models and possibly should be excluded until it can be represented properly.

### 2015-12-03 manuscript/manuscript_wtc3.Rmd
finsihed par figure

@@
 # Figures {.unnumbered}
-```{r par, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
 source("master_scripts/par2.R")

### 2015-12-03 master_scripts/par2.R

@@ -1,19 +1,11 @@
 # source("functions and packages/functions.R")
 # source("functions and packages/packages.R")
 # source("master_scripts/plot_objects.R")
+
 treatments <- read.csv("raw data/temp_trt.csv")
 
-###need campaign dates in order to remake PAR graph with temp
-
-octdate <- as.Date("2013-10-25")
-decdate <- as.Date("2013-12-10")
-jandate <- as.Date("2014-01-28")
-febdate <- as.Date("2014-02-24")
-mardate <- as.Date("2014-03-19")
-aprdate <- as.Date("2014-04-21")
-
-campaigns <- c(octdate, decdate, jandate, febdate, mardate, aprdate)
-campaigns2 <- data.frame(Month = c("Oct", "Dec", "Jan", "Feb", "Mar", "Apr"), Date = campaigns)
+###need campaign dates in order to remake PAR graph with temp  (campaigns, campaigns2, campaindates/simple)are
+##stored in plot objects
 
 ###add temp to par figure so read met data
 
@@ -21,13 +13,25 @@ met <- read.csv("calculated_data/met_chamber.csv")
   met$Date <- as.Date(met$Date)
   met$chamber <- gsub("C", "ch", met$chamber)
 
-metcampaigns <- met[met$Date %in% campaigns,]
+##use vector of all measurement dates from plot objects script
+metcampaigns <- met[met$Date %in% campaigndates_simple,]
   ###i want ambient air temp so need to exclude ET treatments
   metcampaigns <- merge(metcampaigns, treatments)
 
 met_amb <- metcampaigns[metcampaigns$temp == "ambient",]
+  ##extract month and create a mean for each campaign
+  library(lubridate)
+  met_amb$Month <- as.character(month(met_amb$Date, label=TRUE))
+  ##keep only variables i need (for ease of NA removal later)
+  met_amb2 <- met_amb[, c(1:2, 4,6, 8:10)]
+  met_amb2 <- met_amb2[complete.cases(met_amb2),]
 
-met_amb_agg <- summaryBy(Tair_al.mean + Tair_al.max ~ Date, data= met_amb, FUN=c(mean, se))
[…]

### 2015-12-06 manuscript/manuscript_wtc3.Rmd
ppfd2 in caption
changed arivars fig tag so not same as eucpve tag

@@
 
-airvarsfig <- "S1"
+airvarfig <- "S1"
 accfig <- "S2"
@@
 A detailed description of the WTC operation and design is available in [@barton2010whole].
-Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarsfig`).
+Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarfig`).
 Trees grew quickly and developed large canopies, with height growth reaching the top of the WTCs over the experiment duration.
@@
 Each date represents the starting date for each measurement campaign.
-Points represent the maximum daily air temperature during each campaign period.
+Points represent the mean (&pm; 1 standard error) daily maximum air temperature during each campaign period.
 

### 2015-12-06 manuscript/manuscript_wtc3.Rmd
references updates and tags changed in text for co2 compatibility

@@
 How shade leaves utilize sunflecks for short term C gain depends on the combined response time of g~s~ and g~m~ and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks;@tausz2005dynamic].
-The utilization of sunflecks is first limited by delayed responses of stomata opening, which may take minutes, effectively limiting the maximum assimilation rate that can be achieved [@pearcy1990sunflecks; @vico2011effects; @way2012sunflecks].
+The utilization of sunflecks is first limited by delayed responses of stomata opening, which may take minutes, effectively limiting the maximum assimilation rate that can be achieved [@pearcy1990sunflecks; @vicomanz2011; @way2012sunflecks].
 Mesophyll conductance has been shown to respond to environmental factors (e.g.

### 2015-12-07 manuscript/manuscript_wtc3.Rmd
figure caption edit

@@
 All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
-Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (R^2^ = 0.73, 0.58 and 0.72, respectively)
+Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (R^2^ = 0.73, 0.58 and 0.72, respectively).
 

### 2015-12-07 manuscript/manuscript_wtc3.Rmd
figures and tables moved in text
have backup figs and tables and new phyt summary in seperate rmd

started notes script

@@
 
-parfig <- "1"
-acifig <- "2"
-agmgsfig <- "3"
-gebarfig <- "4"
-waterusefig <- "5"
-deltafig <- "6"
-ciccfig <- "7"
+parfig <- "3.1"
+acifig <- "3.2"
+agmgsfig <- "3.3"
+gebarfig <- "3.4"
+waterusefig <- "3.5"
+deltafig <- "3.6"
+ciccfig <- "3.7"
 
-airvarfig <- "S1"
-accfig <- "S2"
-get1fig <- "S3"
-get2fig <- "S4"
+airvarfig <- "3.S1"
+accfig <- "3.S2"
+get1fig <- "3.S3"
+get2fig <- "3.S4"
 
@@
 
-# Summary
-* Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves.
+# Abstract
+Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves.
 A lack of empirical data relating photosynthesis to leaf physiological behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
-* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
-* Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
+To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
+Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
 When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves at the same PPFD.
-* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
+Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
 We argue that the rapid response of g~m~ with light enables shade leaves to respond quickly to sunflecks.
@@
 
+\
+```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/par2.R")
+```  
+\
+**Figure 1**.
+Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014.
+Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf.
+Each date represents the starting date for each measurement campaign.
+Points represent the mean (&pm; 1 standard error) daily maximum air temperature during each campaign period.
+
+\
+**Table 1**.
+*Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments.
+Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
+Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
+Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;.
+Different letters represent significant differences between leaf type and temperature treatments.
+The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.
+
+```{r, echo=FALSE}
+resourcetab2 <- rbind(resourcetab, c("P value","", as.character(pval)))
+greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
+
+dust(resourcetab2)%>%
+  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0(greeks['delta'],"^13^c")) %>%
+  sprinkle(rows= 1, part="head",bold=TRUE)%>%
+  sprinkle_print_method("markdown")
+```
+
 ## Photosynthetic capacity and leaf photosynthesis rates 
@@
 Among the sampled leaves, Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
+
+\
+```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/aci_vcmaxnitro.R")
+```  
+\
+**Figure 2**.
+(a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.
+ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+(b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light.
+(c) The relationship between A and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
+For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
 
@@
 Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
+
+\
+**Table 2**.
+Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
+Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).
+Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa.
+Different letters represent significant differences between leaf type, light environment and temperature treatments.
+The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.
+
+```{r, echo=FALSE}
+getab2 <- rbind(getab, c("P value","","", as.character(pval2)))
+greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
+
+dust(getab2)%>%
+  sprinkle_colnames("Leaf", "Light","Temperature", "A", "g~s~", "g~m~", "ITE","E","VPD","C~i~", "C~c~") %>%
+  sprinkle(rows= 1, part="head",bold=TRUE)%>%
+  sprinkle_print_method("markdown")
+```
 
@@
 \
+```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/Agmgs.R")
+```  
+\
+**Figure 3**.
+The response of A to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
+Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
+Grey areas are 95% confidence intervals from the mean.
+
+\
+```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/physiology_barplots.R")
+```  
+\
+**Figure 4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
+
+\
 Measured under ambient light and temperature conditions leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001).
@@
 Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).
+
+\
+```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
+source("master_scripts/wateruse.R")
+```  
+\
+**Figure 5**.
+(a) Response of instantaneous transpiration efficiency (ITE) to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
+(b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light.
+For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment.
+For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
 
@@
 The remainder $\Delta_o$ consists of the contributions of g~s~, respiration and photorespiration to discrimination.
+
+\
+```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/delta_cica.R")
+```  
+\
+**Figure 6**.
+Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
+The solid line represents the theoretical line for C3 plants from Evans et al.
+[-@evans1986carbon].
 
@@
 The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
+
+\
+```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/cicc_bar.R")
+```  
+\
+**Figure 7**.
+The mean &pm; 1 standard error of intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light.
 
@@
 
-# List of Tables
-**Table 1**.
-*Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments.
-Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
-Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
-Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;.
-Different letters represent significant differences between leaf type and temperature treatments.
-The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.
+# Supporting Information {.unnumbered}
+
+```{r airvars, fig.width=7, fig.height=9, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/air_vars.R")
+```
+\
+**Figure 3.S1**. Daily maximum and minimum temperature (a) and total daily PPFD (b) for each chamber across the experiment duration.
 
 \
-**Table 2**.
-Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
-Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).
-Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa.
-Different letters represent significant differences between leaf type, light environment and temperature treatments.
-The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.
-
-# List of Figures
-
-**Figure 1**.
-Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014.
-Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf.
-Each date represents the starting date for each measurement campaign.
-Points represent the mean (&pm; 1 standard error) daily maximum air temperature during each campaign period.
-
+```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/ACc_figure.R")
+```  
 \
-**Figure 2**.
-(a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.
-ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-(b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light.
-(c) The relationship between A and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
-For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
-
-\
-**Figure 3**.
-The response of A to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
-Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
-Grey areas are 95% confidence intervals from the mean.
-
-\
-**Figure 4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
-
-\
-**Figure 5**.
-(a) Response of instantaneous transpiration efficiency (ITE) to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
-(b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light.
-For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment.
-For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
-
-\
-**Figure 6**.
-Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
-The solid line represents the theoretical line for C3 plants from Evans et al.
-[-@evans1986carbon].
-
-\
-**Figure 7**.
-The mean &pm; 1 standard error of intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light.
-
-\
-**Figure S1**. Daily maximum and minimum temperature (a) and total daily PPFD (b) for each chamber across the experiment duration.
-
-\
-**Figure S2**.
+**Figure 3.S2**.
 Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
@@
 \
-**Figure S3**.
+```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/gasex_temp3.R")
+```  
+\
+**Figure 3.S3**.
 Response of A (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
@@
 \
-**Figure S4**.
+```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+source("master_scripts/gasex_temp4.R")
+```  
+\
+**Figure 3.S4**.
 Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
@@
 
-# Tables {.unnumbered}
-**Table 1**.
-*Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments.
-Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
-Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
-Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;.
-Different letters represent significant differences between leaf type and temperature treatments.
-The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.
-
-```{r, echo=FALSE}
-resourcetab2 <- rbind(resourcetab, c("P value","", as.character(pval)))
-greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
-
-dust(resourcetab2)%>%
-  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0(greeks['delta'],"^13^c")) %>%
-  sprinkle(rows= 1, part="head",bold=TRUE)%>%
-  sprinkle_print_method("markdown")
-```
-
-\
-**Table 2**.
-Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
-Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).
-Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa.
-Different letters represent significant differences between leaf type, light environment and temperature treatments.
-The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.
-
-```{r, echo=FALSE}
-getab2 <- rbind(getab, c("P value","","", as.character(pval2)))
-greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
-
-dust(getab2)%>%
-  sprinkle_colnames("Leaf", "Light","Temperature", "A", "g~s~", "g~m~", "ITE","E","VPD","C~i~", "C~c~") %>%
-  sprinkle(rows= 1, part="head",bold=TRUE)%>%
-  sprinkle_print_method("markdown")
-```
-
-# Figures {.unnumbered}
-```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/par2.R")
-```
-
-**Figure 2**.
-
-```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/aci_vcmaxnitro.R")
-```
-
-**Figure 3**.
-
-```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/Agmgs.R")
-```
-
-**Figure 4**.
-
-```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/physiology_barplots.R")
-```
-
-**Figure 5**.
-
-```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
-source("master_scripts/wateruse.R")
-```
-
-**Figure 6**.
-
-```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/delta_cica.R")
-```
-
-**Figure 7**.
-
-```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/cicc_bar.R")
-```
-
-**Figure 8**.
-
-# Supporting Information {.unnumbered}
-
-```{r airvars, fig.width=7, fig.height=9, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/air_vars.R")
-```
-
-**Figure S1**
-
-```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/ACc_figure.R")
-```
-
-**Figure S2**.
-
-```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/gasex_temp3.R")
-```
-
-**Figure S3**.
-
-```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
-source("master_scripts/gasex_temp4.R")
-```
-
-**Figure S4**.
-
 # References {.unnumbered}

### 2015-12-07 manuscript/tablesfigures.Rmd
figures and tables moved in text
have backup figs and tables and new phyt summary in seperate rmd

started notes script

--- 2015-12-07 manuscript/manuscript_wtc3.Rmd
@@
 ---
-title: "Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks"
-author: ""
-date: ""
-output:
-  word_document:
-    reference_docx: manuscriptstyle.docx
-  pdf_document: default
-  html_document:
-    number_sections: yes
-csl: new-phytologist.csl
-bibliography: references.bib
+title: "tablesfigures"
+author: "Court Campany"
+date: "December 7, 2015"
+output: word_document
 ---
-
-\
-Courtney E. Campany^1^, Belinda Medlyn^1^, Mark Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^.
-
-\
-^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia
-
-\
-^2^ Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, Australia
-
-\
-*Corresponding author*: 
-Courtney Campany
-E: courtneycampany@gmail.com
-
-```{r, results='asis', echo=FALSE, warning=FALSE}
-library(knitr)
-opts_knit$set(root.dir = '../')
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE}
-# Two useful packages for printing numbers, accessing content of objects.
-library(broom)
-library(reporttools)
-
-# Set Flag to include figures in text or not.
-includeFigs <- TRUE
-
-# This document can be compiled with the 'Knit Word' button in Rstudio, or:
-# library(rmarkdown)
-# render("manuscript.Rmd", "word_document", "manuscript.docx")
-```
 
@@
 ```
-
-# Summary
-* Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves.
-A lack of empirical data relating photosynthesis to leaf physiological behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
-* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
-* Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
-When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves at the same PPFD.
-* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
-We argue that the rapid response of g~m~ with light enables shade leaves to respond quickly to sunflecks.
-Evidence that g~m~ not only varies within a canopy but can be up-regulated over short time intervals possibly represents a new mechanism underpinning leaf gas exchange responses to light.
-
-## Key words
-photosynthesis, stomatal conductance, mesophyll conductance,  shade, leaf optimal behavior
-
-# Introduction
-Light availability is one of the most important environmental drivers of leaf carbon (C) uptake in trees.
-Predicting C uptake if forests usually involves upscaling leaf level measurements to assess whole canopy function.
-Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling based on solar irradiance problematic.
-Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor].
-Consequently, structural and functional properties of leaves are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of leaf photosynthesis (*A~n~*) [@evans1995carbon].
-To estimate whole canopy C gain it is thus necessary to account for the non-linear response of A to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
-
-\
-The distribution of resources required for *A~n~*, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients.
-As *A~n~* has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
-Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic].
-Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
-Higher rates of *A~n~* and stomatal conductance (g~s~) can only be sustained if the leaf specific hydraulic conductance (K~l~) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal].
-Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A~n~* if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co], thus K~l~ should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
-
-\
-Rates of photosynthesis in C~3~ plants are limited by the [CO~2~] available for fixation by Rubisco within the chloroplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmosphere to the site of carboxylation [@warren2008stand].
-This drawdown consists of multiple resistance pathways to CO~2~ diffusion which include CO~2~ diffusion from the atmosphere through stomata (stomatal conductance, g~s~) and then from these intercellular air spaces into the chloroplast (mesophyll conductance, g~m~).
-Based on optimal theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize *A~n~* [@peltoniemi2012co].
-This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
-Mesophyll conductance can also impose limitations on *A~n~* as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A~n~* [@niinemets2007photosynthesis] if g~m~ constrains CO~2~ supply to the chloroplast.
-Part of the variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
-Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack empirical data currently hinders are ability to interpreting their coupled responses to *A~n~* across sun and shade leaves.
-
-\
-Additionally, assessing shade leaf behaviour is made difficult with accounting of short term light fluctuations within a canopy, via sunflecks.
-How shade leaves utilize sunflecks for short term C gain depends on the combined response time of g~s~ and g~m~ and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks;@tausz2005dynamic].
-The utilization of sunflecks is first limited by delayed responses of stomata opening, which may take minutes, effectively limiting the maximum assimilation rate that can be achieved [@pearcy1990sunflecks; @vicomanz2011; @way2012sunflecks].
-Mesophyll conductance has been shown to respond to environmental factors (e.g.
-CO~2~, temperature or vapor pressure deficit) at timescales of minutes, possibly faster than g~s~ [@flexas2008mesophyll and references therein], yet short term response to light availability are unclear.
-For example, g~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid].
-Anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are also unlikely to adjust during short light fluctuations.
-The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance and as a potential response to transitory light availability.
-
-\
-Climate warming may also affect the physiological behaviour of leaves within a canopy.
-This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis].
-How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for upscaling *A~n~* for sun and shade leaves.
-Additionally, light saturated rates of A are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
-How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients.
-The impacts of warming on plant physiological processes are obviously vast, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.
-
-\
-In this study we use *Eucalyptus tereticornis* Sm.
-trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to empirically evaluate the distribution of nitrogen and water supply and leaf physiological behaviour of sun and shade leaves.
-Our hypotheses are as follows:
-
-\
-1.	If whole tree canopies are optimized for C gain, then leaf N, hydraulic conductance and biochemical photosynthetic capacity are predicted to be higher in sun leaves compared to shade leaves.
-
-\
-2.	Stomatal conductance is proportional to *A~n~* across sun and shade leaves under similar leaf VPD and g~m~ scales positively with photosynthetic capacity.
-
-\
-3.
-As shade leaves were expected to develop lower biochemical photosynthetic capacity and leaf physiological responses to light are typically slow (minutes), increases in *A~n~* following sunfleck simulations were not expected to reach rates of full sun leaves.
-
-\
-4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporation demand from higher temperatures will lead to stomatal closure.
-
-# Materials and Methods
-
-## Whole tree chamber experimental design
-Twelve *Eucalyptus tereticornis* seedlings, chosen from a single local Cumberland plain cohort, were planted in March 2013 into 12 whole-tree chambers (WTC) at the Hawkesbury Forest Experiment site near Richmond, NSW, Australia.
-Each chamber has a height of 9 m and seedlings were grown for 15 months.
-A detailed description of the WTC operation and design is available in [@barton2010whole].
-Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarfig`).
-Trees grew quickly and developed large canopies, with height growth reaching the top of the WTCs over the experiment duration.
-Trees were watered weekly with 70 L from March 2013 to November 2013.
-From October 2013 to the end of the experiment trees were watered every 15 days with the mean monthly rainfall amount for Richmond, NSW.
-In February 2013 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering.
-Due to a limited range of data for the drought treatment only well-watered trees are reported, which reduces the sample size of WTC (n=6 to n=3) for the final 3 months of the experiment.
-
-\
-Before seedlings were planted into each chamber they were maintained under well-watered conditions in 25 L pots and kept inside each chamber.
-This allowed for seedlings to gain sufficient size before planting while also allowing them to acclimate to chamber temperature treatments.
-Seedlings were planted into each chamber after mean seedling height reached 100 cm.
-The top soils at this site, used in both pots and chambers, are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg-1 total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity.
-A root exclusion barrier extended from chamber walls to the hard layer (ca.
-1 m) and roots were allowed to grow freely below the barrier.
-Leaf gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.
-At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm, height of 348&pm;15.1 cm and an leaf area of 3.9&pm;0.1 m^2^.
-For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an leaf area of 6.2&pm;0.2 m^2^.
-Leaf area was calculated based on leaf counts and mean leaf size.
-
-##Leaf gas exchange, online carbon isotope discrimination and mesophyll conductance
-Leaf gas exchange measurements were performed six times, beginning in October 2013 and monthly from December 2013 to April 2014.
-Measurements were taken on a representative sun and shade leaf for each tree during each measurement campaign.
-The newest fully expanded leaf from the branch apex was chosen for gas exchange measurements and sun leaves were measured in the upper third of the canopy.
-Here, shade leaves are defined as inner-canopy leaves developing on secondary branches in a low light environment.
-Shade leaves were always measured in the lower canopy, but leaves were sampled on subsequent higher branches across measurement campaigns to minimize confounding effects of leaf age.
-As shade leaves most likely developed slower this assured that older leaves in the bottom canopy were avoided.
-
-\
-Prior to gas exchange measurements photosynthetic photon flux density (PPFD) was recorded as a point measurement at the individual leaf level and a spatially averaged measurement at the canopy position for each selected leaf.
-A hand-held quantum sensor (LI-COR, Lincoln, NE, USA) was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either full sun or full shade.
-A ceptometer (AccuPAR LP-80, Decagon Devices, Pullman, USA) was then used to measure a spatially averaged PPFD at the canopy height of each chosen leaf type.
-Each ceptometer reading integrated an array of 80 sensors over a total length of 84 cm.
-Five ceptometer readings were recorded at different locations within the canopy, but at the same height and close to each selected leaf.
-The mean of these readings was assumed to represent the local light environment of sun and shade leaves for each tree.
-All measurements of PPFD and gas exchange were performed on sunny days between 10:00-14:30 h.
-
-\
-Leaf level gas exchange was measured with a standard (2 x 3 cm) leaf cuvette using a portable gas exchange system (LI-6400XT, LI-COR, Lincoln, NE, USA).
-This system was coupled with a tunable diode laser (TDL; TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of online C isotope discrimination.
-The CO~2~ in the leaf cuvette was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mol s^-1^.
-Two identical gas exchanges systems were run simultaneously, one in each of a randomly chosen WTC for each temperature treatment.
-Leaf temperatures were controlled at the current AT or ET WTC air temperature.
-PPFD in the cuvette was set to match the individual light environment of each leaf type (explained above).
-Periods of high irradiance were simulated for shade leaves by increasing the leaf cuvette PPFD (LI-COR red/blue light source) to match the light environment of the full sun leaf in the same tree.
-The maximum sunfleck response of shade leaves was then recorded once CO~2~ and water vapor fluxes re-stabilized in the leaf cuvette (ca.
-25 min).
-
-\
-Once CO~2~ and water vapor flux values were stable for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet of the LI-6400.
-The gas streams were dried by passing through napion gas dryers in the respective gas lines, and then ^12^CO~2~ and ^13^CO~2~ concentrations were measured for each gas stream by the TDL.
-Reference, sample and two calibration gases were run on alternating 80 s loops (20 s each), one for each AT and ET leaf at a matched canopy position, for a total of 12 min.
-This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference line and samples line gas streams for calculations.
-The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDL on each measurement cycle.
-Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the same 12 min interval.
-
-\
-Using online C isotope discrimination measurements, the difference  between the observed discrimination and what is predicted for light saturated gas exchange is proportional to g~m~ [@griffiths2013mesophyll].
-First, leaf C isotope discrimination was calculated by comparing the isotopic composition of the reference gas entering the leaf cuvette ($\delta$^13^C~e~) with the sample gas ($\delta$^13^C~o~) such that:
-
-\
-(1)
-$$\delta = \left(\frac{R_s}{R_{stnd}} -1\right)1000$$
-
-where R~s~ is the isotopic ratio of the sample and R~stnd~ is the isotopic ratio of the standard Vienna Pee Dee Belemnite (VPDP).
-Next, the observed discrimination ($\Delta$~obs~) is calculated from Evans et al.
-[-@evans1986carbon]:
-
-\
-(2)
-$$\Delta_{obs} = \frac{\xi(\delta^13C_o-\delta^13C_e)}{1000+\delta^13C_e-\xi(\delta^13C_o-\delta^13C_e)}$$
-
-where:
-
-\
-(3)
-$$\xi = \frac{C_{e}}{C_{e}-C_{o}}$$
-
-and $\xi$ is the ratio of the CO~2~ entering the well mixed leaf cuvette to the CO~2~ draw down in the gas stream by the leaf.
-
-\
-Second, C isotope discrimination during C~3~ photosynthesis ($\Delta$) is the resultant discrimination from CO~2~ diffusion from the atmosphere to the site of carboxylation, consisting of a series of fractionation steps described in Evans et al.
-[-@evans1986carbon].
-In this experiment, a modified form of this equation presented in Evans & Von Caemmerer [-@evans2013temperature] with ternary effect corrections by Farquhar & Cernusak [-@farquhar2012ternary] was used such that:
-
-\
-(4)
-$$\Delta_o = \Delta_i - \Delta_{gm} - \Delta_e - \Delta_f$$
-
-where $\Delta$~o~ is the observed discrimination and  $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, g~m~, respiration and photorespiration, respectively.
-The equations for each are as follows:
-
-\
-(5)
-$$
-\begin{aligned}
-\Delta_i = &\frac{1}{1-t}a_i+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}
-\end{aligned}
-$$
-
-\
-(6)
-$$
-\begin{aligned}
-\Delta_{gm} = &\frac{1+t}{1-t}\left(b-a'-\frac{eR_d}{A+R_d}\right)\frac{A}{g_mC_a}
-\end{aligned}
-$$
-
-\
-(7)
-$$
-\begin{aligned}
-\Delta_e = &\frac{1+t}{1-t}\left(\frac{eR_d}{(A+R_d)C_a}(C_i-\Gamma_*)\right)
-\end{aligned}
-$$
-
-\
-(8)
-$$
-\begin{aligned}
-\Delta_f = & \frac{1+t}{1-t}\left(f\frac{\Gamma_*}{C_a}\right)\\
-\end{aligned}
-$$
-
-where the different fractionation factors include; diffusion through water (a~i~, 1.8&permil;), Rubisco carboxylation (b, 29&permil;), the photorespiratory fractionation (f, 16.2&permil;) and the combined fractionation through the boundary layer and the stomata (a').
-a' is defined by:
-
-\
-(9)
-$$a' = \frac{a_b(C_a-C_i)+a(C_s-C_i)}{C_a-C_i}$$
-
-where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ is the fractionation from boundary layer diffusion (2.9&permil;) and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon].
-C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~).
-In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulus* from [@crous2012light].
-The ternary effect corrections (t) are described by:
-
-\
-(10)
-$$t = \frac{(1+a')E}{2g^t_{ac}}$$
-
-where E denotes the transpiration rate and $\text{g}^t_{ac}$ is the total conductance to CO~2~ diffusion to both the boundary layer and stomatal conductance.
-
-\
-The CO~2~ diffusion from the intercellular airspace to the chloroplast, g~m~, is given by its relationship to the leaf photosynthesis rate (*A~n~*) by:
-
-\
-(11)
-$$g_m = \frac{*A~n~*}{C_i-C_c}$$
-
-where C~c~ is the chloroplast CO~2~ partial pressure.
-Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were then estimated using Equation 11.
-Examples of this approach to measure gas exchange and C isotope discrimination are presented in Evans and Von Caemmerer [-@evans2013temperature].
-The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck where then compared as a function of C~i~:C~a~.
-
-##Biochemical parameters of photosynthesis
-Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each WTC prior to the initiation of the drought treatment.
-Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011].
-
-##Leaf chemistry and hydraulic parameters
-Following gas exchange measurements each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, dried and weighed.
-These leaves were then milled and analyzed for leaf N content and $\delta$^13^C.
-Leaf samples were analysed on a Delta V Advantage coupled to a Flash HT and Conflo IV (Thermo Fisher Scientific, Bremen, Germany) in dual-reactor setup.
-Samples were flash combusted at 1000&deg;C and converted to CO~2~ and N~2~ and then subjected to stable isotope ratio mass spectrometry.
-Leaf N is reported on an area basis (N~a~, g m^-2^) and isotopic signatures are reported relative to the VPDP scale.
-
-\
-Predawn ($\Psi_{pd}$) and midday ($\Psi_l$) leaf water potentials (MPa) were measured for sun and shade leaves during each gas exchange campaign using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA).
-The leaf closest to the leaf used for gas exchange was sampled for measurement of $\Psi_{pd}$.
-Predawn leaf water potential was measured before sunrise on the same day as gas exchange measurements.
-Leaves used for gas exchange were immediately sampled for $\Psi_l$ once measurements were completed.
-All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
-Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) through the equation:
-
-\
-(12)
-$$K = \frac{E}{\Psi_{pd}~-\Psi_l}$$
-
-Leaf level instantaneous transpiration efficiency (ITE) was calculated as A divided by E.
-The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression, where k=0.5 [see @duursma2013near].
-Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use [@lin2015optimal].
-
-##Data analysis
-Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R [@RDevelopmentCoreTeam2011] with WTC as a random effect.
-Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general].
-Confidence intervals (95 %) of mixed effect linear models were generated using bootstrapping methods with 999 simulations, using the bootMer function in the 'lme4' package [@bates2013lme4].
-For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package, using WTC as a random effect.
-All tests of statistical significance were conducted at an alpha of 0.05.
-
-# Results
-
-## Leaf resource distribution
-Across six measurement campaigns over the 7 month period PPFD was  reduced on average by >75% in the shade (Figure `r parfig`).
-Leaf-specific hydraulic conductance (K~l~) was similar across sun and shade leaves (Table 1).
-This was because $\Psi_{pd}$, $\Psi_l$ (Table 1) and E (Table 2) did not differ between leaf types.
-Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table1).
-Leaf mass per area was not different between leaf types.
-No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~area~ or LMA either within or across leaf types.
-
-## Photosynthetic capacity and leaf photosynthesis rates 
-The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table 1), as estimated from AC~i~ curves (Figure `r acifig`a).
-Within leaf types, no effect of the warming treatment was detected on either parameter.
-Among the sampled leaves, Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
-
-\
-Mean *A~n~* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table 2).
-Additionally, leaf N~a~ was positively related to *A~n~* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
-Following an increase in light intensity to match high-light conditions, A of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table 2).
-No effect of the warming treatment was detected on *A~n~* of sun leaves measured at high light or shade leaves at either low or high light.
-Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
-
-## Stomatal conductance and leaf water-use efficiency
-On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table 2).
-Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and *A~n~* were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
-Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
-No effect of the warming treatment was detected on g~s~ within or across leaf types.
-Stomatal conductance within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get2fig`b).
-
-\
-Measured under ambient light and temperature conditions leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001).
-Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
-Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment, but no warming effect was detected in shade leaves measured at low or high light (Table 2).
-The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
-For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal ITE model with a g~1~ value for each leaf type and treatment (Figure `r waterusefig`a).
-Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).
-
-\
-Bulk-leaf $\delta$^13^C, as an index of integrated water-use efficiency [@marshall2007sources], was significantly lower in shade leaves compared to sun leaves by ca.
-2&permil; (Table 1).
-No effects of the warming treatment on leaf $\delta$^13^C were detected.
-Leaf $\delta$^13^C and N~area~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C (higher water-use efficiency) and higher N investment in sun leaves.
-
-## Leaf carbon isotope discrimination and mesophyll conductance
-The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
-Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
-The remainder $\Delta_o$ consists of the contributions of g~s~, respiration and photorespiration to discrimination.
-
-\
-Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
-Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table 2).
-Proportional increases in g~m~ were matched by proportional increases in *A~n~* from low to high light in shade leaves (Figure `r gebarfig`b,c).
-Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186).
-The large increases in g~m~ in shade leaves under high light resulted in the highest rates of *A~n~* (Figure `r agmgsfig`b).
-No differences in g~m~ were detected with the warming treatment within leaf types.
-Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively).
-We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~.
-Comparison of A-Cc curves (Figure `r accfig`) and A-Ci curves revealed similar differences between sun and shade leaves.
-
-## Variation in intercellular and chloroplastic CO~2~ concentrations
-Higher rates of g~s~ in shade leaves under low and high light lead to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
-The chloroplast CO~2~ partial pressure was comparable between shade leaves when measured at both low and high light conditions (Figure `r ciccfig`c).
-In sun leaves C~c~ was significantly lower than shade leaves as a likely consequence of a lower C~i~.
-The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and *A~n~* [@caemmerer2014temperature].
-This drawdown was similar between sun and shade leaves measured at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
-This was the result of the proportional relationship between g~m~ and *A~n~* across all leaves.
-The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
-
-# Discussion
-Here we show that *A~n~* in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
-Although shade leaves in lower light environments exhibited relatively high g~s~ and c~i~, it was rapid increases in g~m~ under periods of high light availability that led to this up-regulation of *A~n~*.
-Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
-These findings offer new insights into how aspects of leaf physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake in future studies.
-Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology.
-
-## Resource distribution and photosynthetic capacity
-The allocation of N~a~ constrains *A~n~* and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
-Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979).
-As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical].
-Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and *A~n~* were all reduced in shade leaves.
-Leaf mass area, however, was not different between sun and shade leaves.
-This could be due to leaf formation under comparative light conditions or possible differences in total non-structural carbohydrates contents between leaf types.
-
-\
-Photosynthesis is also limited by the ability to supply water to the upper canopy.
-Ultimately, the ability of a trees hydraulic architecture to supply water to foliage across increasing pathlengths affects productivity and survival [@sellin2008effects].
-Using a two-leaf model, Peltoniemi et al.
-[-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if K~l~ is also optimally distributed.
-In this study, variation in leaf N distribution and *A~n~* rates were not associated with subsequent changes in K~l~ between sun and shade leaves.
-Thus, no direct relationship between water supply and N distribution or *A~n~* within the canopy were detected.
-
-\
-Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased ITE throughout the experiment.
-Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was likely prevalent long term.
-From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A~n~* in sun leaves was characterized by low rates of g~s~ and low C~i~.
-Relative to the differences in *A~n~* between leaf types, higher rates of g~s~ in shade leaves appear to exhibit inefficient water use.
-As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower ITE compared to sun leaves.
-
-## Physiological behaviour of sun and shade leaves
-The pattern of inefficient water use in shade leaves is important as we hypothesized that g~s~ and *A~n~* would be proportional across sun and shade leaves.
-In sun leaves, *A~n~* and g~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
-However, lower rates of *A~n~* in shade leaves were not coupled with decreases in g~s~, explaining the observed decreases in ITE.
-This is significant as optimal stomatal regulation to balance C gain with water loss has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on full sun leaves [see @prentice2014balancing; @lin2015optimal].
-As a result, the often used economic framework of balancing costs of using water versus N allocation to predict *A~n~* [@wright2003least] may break down when considering all leaves within a tree canopy.
-
-\
-It is possible that reducing stomatal response time, by sustaining higher g~s~, is a strategy to take advantage of high light quickly in shade leaves [@tausz2005dynamic].
-Evidence from this study supports this hypothesis, as shade leaves increased A equivalent or even outperforming sun leaves under identical light intensity.
-Transpiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck induced thermal load [@schymanski2013stomatal].
-This is because rapid increases in leaf temperature with sunflecks have been shown to inhibit C gain [@leakey2003high].
-However, this response likely occurs at very high air temperatures and may not explain the observed g~s~ in shade leaves across the large natural range of temperatures included in this study.
-How prevalent each of these strategies are within tree canopies is still unknown, as empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants, often in deep shade [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution].
-Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy [see @de1997simple].
-
-\
-We found that *A~n~* and g~m~ scaled positively across leaf types and, surprisingly, increased rapidly (within minutes) and proportionately when light intensity was increased in shade leaves.
-Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
-This provides a potential explanation for the observed rapid increases in g~m~ but the impacts of aquaporins on g~m~ are untested in leaves of tree species.
-Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
-Here we provide empirical data showing g~m~ not only varies within a canopy but the up-regulation of g~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.
-
-\
-If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution.
-This is because the role of g~s~ in regulating photosynthetic induction impacts the capacity of a leaf to utilize sunflecks [@way2012sunflecks].
-If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan.
-For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to tolerate decreases in ITE However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.
-Thus, models which predict leaf photosynthesis from N distribution within a canopy will be incomplete unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].
-
-## Conclusions
-Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy.
-Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
-These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have important implications for process-based models that predict canopy C gain from rates of leaf photosynthesis.
-Furthermore, the dynamic nature of g~m~ cannot be simply parameterized in tree growth models and possibly should be excluded until it can be represented properly.
-Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown.
-To improve our ability to predict whole canopy C gain future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.
 
@@
 
-**Figure 2**.
+**Figure 1**.
 
@@
 
-**Figure 3**.
+**Figure 2**.
 
@@
 
-**Figure 4**.
+**Figure 3**.
 
@@
 
-**Figure 5**.
+**Figure 4**.
 
@@
 
-**Figure 6**.
+**Figure 5**.
 
@@
 
-**Figure 7**.
+**Figure 6**.
 
@@
 
-**Figure 8**.
+**Figure 7**.
 
@@
 
-# References {.unnumbered}
+###new phytologist summary
+# Summary
+* Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves.
+A lack of empirical data relating photosynthesis to leaf physiological behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
+* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
+* Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
+When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves at the same PPFD.
+* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
+We argue that the rapid response of g~m~ with light enables shade leaves to respond quickly to sunflecks.
+Evidence that g~m~ not only varies within a canopy but can be up-regulated over short time intervals possibly represents a new mechanism underpinning leaf gas exchange responses to light.

### 2015-12-07 manuscript/manuscript_wtc3.Rmd
tables #ing are now automated

@@
 get2fig <- "3.S4"
+
+wtc3tab1 <- "3.1"
+wtc3tab2 <- "3.2"
 
@@
 Across six measurement campaigns over the 7 month period PPFD was  reduced on average by >75% in the shade (Figure `r parfig`).
-Leaf-specific hydraulic conductance (K~l~) was similar across sun and shade leaves (Table 1).
-This was because $\Psi_{pd}$, $\Psi_l$ (Table 1) and E (Table 2) did not differ between leaf types.
+Leaf-specific hydraulic conductance (K~l~) was similar across sun and shade leaves (Table `r wtc3tab1`).
+This was because $\Psi_{pd}$, $\Psi_l$ (Table `r wtc3tab1`) and E (Table `r wtc3tab2`) did not differ between leaf types.
 Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table1).
@@
 \
-**Figure 1**.
+**Figure 3.1**.
 Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014.
@@
 \
-**Table 1**.
+**Table 3.1**.
 *Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments.
@@
 ## Photosynthetic capacity and leaf photosynthesis rates 
-The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table 1), as estimated from AC~i~ curves (Figure `r acifig`a).
+The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from AC~i~ curves (Figure `r acifig`a).
 Within leaf types, no effect of the warming treatment was detected on either parameter.
@@
 \
-**Figure 2**.
+**Figure 3.2**.
 (a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.
@@
 \
-Mean *A~n~* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table 2).
+Mean *A~n~* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table `r wtc3tab2`).
 Additionally, leaf N~a~ was positively related to *A~n~* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
-Following an increase in light intensity to match high-light conditions, A of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table 2).
+Following an increase in light intensity to match high-light conditions, A of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table `r wtc3tab2`).
 No effect of the warming treatment was detected on *A~n~* of sun leaves measured at high light or shade leaves at either low or high light.
@@
 \
-**Table 2**.
+**Table 3.2**.
 Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
@@
 ## Stomatal conductance and leaf water-use efficiency
-On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table 2).
+On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table `r wtc3tab2`).
 Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and *A~n~* were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
@@
 \
-**Figure 3**.
+**Figure 3.3**.
 The response of A to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
@@
 \
-**Figure 4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
+**Figure 3.4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
 
@@
 Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
-Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment, but no warming effect was detected in shade leaves measured at low or high light (Table 2).
+Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment, but no warming effect was detected in shade leaves measured at low or high light (Table `r wtc3tab2`).
 The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
@@
 \
-**Figure 5**.
+**Figure 3.5**.
 (a) Response of instantaneous transpiration efficiency (ITE) to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
@@
 Bulk-leaf $\delta$^13^C, as an index of integrated water-use efficiency [@marshall2007sources], was significantly lower in shade leaves compared to sun leaves by ca.
-2&permil; (Table 1).
+2&permil; (Table `r wtc3tab1`).
 No effects of the warming treatment on leaf $\delta$^13^C were detected.
@@
 \
-**Figure 6**.
+**Figure 3.6**.
 Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
@@
 Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
-Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table 2).
+Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table `r wtc3tab2`).
 Proportional increases in g~m~ were matched by proportional increases in *A~n~* from low to high light in shade leaves (Figure `r gebarfig`b,c).
@@
 \
-**Figure 7**.
+**Figure 3.7**.
 The mean &pm; 1 standard error of intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light.

### 2015-12-08 manuscript/manuscript_wtc3.Rmd
shade leaves color change, figs now 600dpi

@@
 \
-```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/par2.R")
@@
 \
-```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/aci_vcmaxnitro.R")
@@
 \
-```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/Agmgs.R")
@@
 \
-```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/physiology_barplots.R")
@@
 \
-```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
+```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=600}
 source("master_scripts/wateruse.R")
@@
 \
-```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/delta_cica.R")
@@
 \
-```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/cicc_bar.R")
@@
 
-```{r airvars, fig.width=7, fig.height=9, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r airvars, fig.width=7, fig.height=9, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/air_vars.R")
@@
 \
-```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/ACc_figure.R")
@@
 \
-```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/gasex_temp3.R")
@@
 \
-```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/gasex_temp4.R")

### 2015-12-08 master_scripts/plot_objects.R

@@ -79,7 +79,12 @@ ltys <- c(1,2, 1,2)
 
 ###sun shade colors-------------------------------------------------------------------------------------------
 suncol2 <- "#005300"
-shacol2 <- "#61726C"
+
+
+# shacol2 <- "#61726C"
+shacol2 <- "#9d2f16"
+
+
 lightscol2 <- "#E68A00"
 newshade <- ("#613E1C")
 

### 2015-12-08 manuscript/tablesfigures.Rmd
dpi change on backup rmd

@@
 # Figures {.unnumbered}
-```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/par2.R")
@@
 
-```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/aci_vcmaxnitro.R")
@@
 
-```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/Agmgs.R")
@@
 
-```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/physiology_barplots.R")
@@
 
-```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
+```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=600}
 source("master_scripts/wateruse.R")
@@
 
-```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/delta_cica.R")
@@
 
-```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/cicc_bar.R")
@@
 
-```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/ACc_figure.R")
@@
 
-```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/gasex_temp3.R")
@@
 
-```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
+```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/gasex_temp4.R")

### 2015-12-08 manuscript/manuscript_wtc3.Rmd
swicthed to thesis docx for margins

@@
   word_document:
-    reference_docx: manuscriptstyle.docx
+    reference_docx: manuscriptstyle_thesis.docx
   pdf_document: default

### 2015-12-08 manuscript/manuscript_wtc3.Rmd
figure align and added chapter # designations to sections

@@
 
-# Abstract
+#3.1  Abstract
 Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves.
@@
 
-## Key words
-photosynthesis, stomatal conductance, mesophyll conductance,  shade, leaf optimal behavior
+<!-- ## Key words -->
+<!-- photosynthesis, stomatal conductance, mesophyll conductance,  shade, leaf optimal behavior -->
 
-# Introduction
+#3.2  Introduction
 Light availability is one of the most important environmental drivers of leaf carbon (C) uptake in trees.
@@
 
-# Materials and Methods
+#3.3  Materials and Methods
 
-## Whole tree chamber experimental design
+##3.3.1 Whole tree chamber experimental design
 Twelve *Eucalyptus tereticornis* seedlings, chosen from a single local Cumberland plain cohort, were planted in March 2013 into 12 whole-tree chambers (WTC) at the Hawkesbury Forest Experiment site near Richmond, NSW, Australia.
@@
 
-##Leaf gas exchange, online carbon isotope discrimination and mesophyll conductance
+##3.3.2 Leaf gas exchange, online carbon isotope discrimination and mesophyll conductance
 Leaf gas exchange measurements were performed six times, beginning in October 2013 and monthly from December 2013 to April 2014.
@@
 
-##Biochemical parameters of photosynthesis
+##3.3.3 Biochemical parameters of photosynthesis
 Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each WTC prior to the initiation of the drought treatment.
@@
 
-##Leaf chemistry and hydraulic parameters
+##3.3.4 Leaf chemistry and hydraulic parameters
 Following gas exchange measurements each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, dried and weighed.
@@
 
-##Data analysis
+##3.3.5 Data analysis
 Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R [@RDevelopmentCoreTeam2011] with WTC as a random effect.
@@
 
-# Results
+#3.4  Results
 
-## Leaf resource distribution
+##3.4.1 Leaf resource distribution
 Across six measurement campaigns over the 7 month period PPFD was  reduced on average by >75% in the shade (Figure `r parfig`).
@@
 No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~area~ or LMA either within or across leaf types.
-
 \
@@
 
-## Photosynthetic capacity and leaf photosynthesis rates 
+##3.4.2 Photosynthetic capacity and leaf photosynthesis rates 
 The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from AC~i~ curves (Figure `r acifig`a).
@@
 Among the sampled leaves, Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
-
 \
@@
 
-## Stomatal conductance and leaf water-use efficiency
+##3.4.3 Stomatal conductance and leaf water-use efficiency
 On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table `r wtc3tab2`).
@@
 Stomatal conductance within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get2fig`b).
-
 \
@@
 Grey areas are 95% confidence intervals from the mean.
-
 \
@@
 Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).
-
 \
@@
 
-## Leaf carbon isotope discrimination and mesophyll conductance
+##3.4.4 Leaf carbon isotope discrimination and mesophyll conductance
 The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
@@
 
-## Variation in intercellular and chloroplastic CO~2~ concentrations
+##3.4.5 Variation in intercellular and chloroplastic CO~2~ concentrations
 Higher rates of g~s~ in shade leaves under low and high light lead to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
@@
 The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
-
 \
@@
 
-# Discussion
+#3.5  Discussion
 Here we show that *A~n~* in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
@@
 
-## Resource distribution and photosynthetic capacity
+##3.5.1 Resource distribution and photosynthetic capacity
 The allocation of N~a~ constrains *A~n~* and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
@@
 
-## Physiological behaviour of sun and shade leaves
+##3.5.2 Physiological behaviour of sun and shade leaves
 The pattern of inefficient water use in shade leaves is important as we hypothesized that g~s~ and *A~n~* would be proportional across sun and shade leaves.
@@
 
-## Conclusions
+##3.5.3 Conclusions
 Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy.
@@
 
-# Supporting Information {.unnumbered}
+#3.6  Supporting Information
 
@@
 **Figure 3.S1**. Daily maximum and minimum temperature (a) and total daily PPFD (b) for each chamber across the experiment duration.
-
 \
@@
 C~c~ values were predicted with g~m~, thus curves represent chloroplastic photosynthetic parameters at 25°C and saturating light (1800 $\mu$mols m^-1^ s^-1^).
-
 \
@@
 Weak negative relationships with g~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (R^2^ = 0.16 and 0.08, respectively).
-
 \

### 2015-12-08 manuscript/manuscript_wtc3.Rmd
tried to add chapters to title:fail
hid corresponding author

@@
 ---
-title: "Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks"
+title: "Chapter 3 Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks"
 author: ""
@@
 
-\
-*Corresponding author*: 
-Courtney Campany
-E: courtneycampany@gmail.com
+<!-- \ -->
+<!-- *Corresponding author*: -->
+<!-- Courtney Campany -->
+<!-- E: courtneycampany@gmail.com -->
 

### 2015-12-08 reproducibility_notes.R

@@ -10,3 +10,5 @@
 #4. for RMD will use two dummy docx files for thesis and pub, thesis will need to reflect binding requirements.
 
 #comment out keywords for thesis
+#comment out author correspondence for thesis
+#leaving co-authors in thesis for now

### 2015-12-08 manuscript/manuscript_wtc3.Rmd
fixed chapter and title

@@
 ---
-title: "Chapter 3 Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks"
+title: |
+  | Chapter 3 
+  | Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks
 author: ""

### 2015-12-09 manuscript/manuscript_wtc3.Rmd
small edits
authors commented out

this version used in thesis draft1

@@
 
-\
-Courtney E. Campany^1^, Belinda Medlyn^1^, Mark Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^.
+<!-- \ -->
+<!-- Courtney E. Campany^1^, Belinda Medlyn^1^, Mark Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^. -->
 
-\
-^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia
+<!-- \ -->
+<!-- ^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia -->
 
-\
-^2^ Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, Australia
+<!-- \ -->
+<!-- ^2^ Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, Australia -->
 
@@
 getab <- read.csv("master_scripts/ge_table.csv", stringsAsFactors = FALSE)
-pval <- c(0.785, 0.001, 0.028, 0.002, 0.001, 0.3486, 0.6385, 0.973)
+pval <- c(0.781, 0.001, 0.028, 0.002, 0.973, 0.3486, 0.6385, 0.001)
 pval2 <- c(0.001,	0.001,	0.001,	0.001,	0.001,	0.005,	0.001,	0.001)
@@
 If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan.
-For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to tolerate decreases in ITE However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.
+For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to tolerate decreases in ITE.
+However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.
 Thus, models which predict leaf photosynthesis from N distribution within a canopy will be incomplete unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].
@@
 
-```{r airvars, fig.width=7, fig.height=9, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+```{r airvars, fig.width=7, fig.height=8.8, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/air_vars.R")

### 2015-12-09 manuscript/manuscript_wtc3.Rmd
changed thesis draft to tree phys refernces

@@
     number_sections: yes
-csl: new-phytologist.csl
+csl: tree-physiology.csl
 bibliography: references.bib

### 2015-12-11 manuscript/manuscript_wtc3.Rmd
new methods

@@
 
-<!-- \ -->
-<!-- Courtney E. Campany^1^, Belinda Medlyn^1^, Mark Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^. -->
+\
+Courtney E. Campany^1^, Belinda Medlyn^1^, Mark Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^.
 
-<!-- \ -->
-<!-- ^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia -->
+\
+^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia
 
-<!-- \ -->
-<!-- ^2^ Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, Australia -->
+\
+^2^ Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, Australia
 
-<!-- \ -->
-<!-- *Corresponding author*: -->
-<!-- Courtney Campany -->
-<!-- E: courtneycampany@gmail.com -->
+\
+*Corresponding author*:
+Courtney Campany
+E: courtneycampany@gmail.com
 

### 2015-12-11 markdown/methods_condensed.Rmd
new methods

--- 2015-12-11 manuscript/manuscript_wtc3.Rmd
@@
 ---
-title: |
-  | Chapter 3 
-  | Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks
-author: ""
-date: ""
-output:
-  word_document:
-    reference_docx: manuscriptstyle_thesis.docx
-  pdf_document: default
-  html_document:
-    number_sections: yes
-csl: tree-physiology.csl
-bibliography: references.bib
+title: "gm methods_revised"
+author: "Court Campany"
+date: "December 11, 2015"
+output: word_document
 ---
-
-\
-Courtney E. Campany^1^, Belinda Medlyn^1^, Mark Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^.
-
-\
-^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia
-
-\
-^2^ Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, Australia
-
-\
-*Corresponding author*:
-Courtney Campany
-E: courtneycampany@gmail.com
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE}
-# Two useful packages for printing numbers, accessing content of objects.
-library(broom)
-library(reporttools)
+#from SVCEVANS
+Gas exchange and carbon isotope discrimination measurements were made as described by Tazoe et al.
+(2011) and Evans & von Caemmerer (2013).
+Plants were placed in a temperature-controlled growth cabinet together with the measuring heads (6 cm2 chamber with a red blue lightemitting diode light source) of two LI-6400XT gas exchange systems (Li-Cor, Lincoln, NE, USA).
+These chambers were coupled to a tunable diode laser (TDL;TGA100,Campbell Scientific, Inc., Logan, UT, USA) with a T junction in the match valve tubing.
+This allowed leaves of two plants to be measured in sequence, with each LI-6400XT sampled by the TDL at 4 min intervals as outlined in Fig.
+1 of Tazoe et al.
+(2011).
+The LI-6400XT consoles were placed outside the growth chambers.
+The CO2 in the leaf chamber was set at 380 mol mol−1 in the sample cell, flow rate at 200 mol s−1 and irradiance of 1500 or 1000 mol quantam−2 s−1.N2 andO2 were mixed by mass flow controllers (Omega Engineering Inc., Stamford, CT, USA) to generate 2% O2, which was supplied to the consoles and specified for the LI-6400 calculations.
+Leaves were stabilized first at 25 °C, then temperature was dropped to 15 °C and then increased in 5 °C steps.
+Mesophyll conductance was calculated from carbon isotope discrimination with equations and fractionation factors presented in Evans & von Caemmerer (2013) including the ternary corrections proposed by Farquhar & Cernusak (2012).
+Low O2 partial pressure and high light minimized the contribution of respiratory and photorespiratory fractionation to the carbon isotope signal.
 
-# Set Flag to include figures in text or not.
-includeFigs <- TRUE
-
-# This document can be compiled with the 'Knit Word' button in Rstudio, or:
-# library(rmarkdown)
-# render("manuscript.Rmd", "word_document", "manuscript.docx")
-```
-
-```{r global data/sourcing , include=FALSE}
-source("functions and packages/functions.R")
-source("functions and packages/packages.R")
-source("master_scripts/plot_objects.R")
-library(pixiedust)
-
-##read table ready dataframes
-resourcetab <- read.csv("master_scripts/resource_table.csv", stringsAsFactors = FALSE)
-getab <- read.csv("master_scripts/ge_table.csv", stringsAsFactors = FALSE)
-pval <- c(0.781, 0.001, 0.028, 0.002, 0.973, 0.3486, 0.6385, 0.001)
-pval2 <- c(0.001,	0.001,	0.001,	0.001,	0.001,	0.005,	0.001,	0.001)
-# getab <- read.csv("master_scripts/ge_table.csv")
-
-#treatments <- read.csv("raw data/chamber_trt.csv")
-```
-
-```{r, echo=FALSE}
-
-parfig <- "3.1"
-acifig <- "3.2"
-agmgsfig <- "3.3"
-gebarfig <- "3.4"
-waterusefig <- "3.5"
-deltafig <- "3.6"
-ciccfig <- "3.7"
-
-airvarfig <- "3.S1"
-accfig <- "3.S2"
-get1fig <- "3.S3"
-get2fig <- "3.S4"
-
-wtc3tab1 <- "3.1"
-wtc3tab2 <- "3.2"
-
-```
-
-#3.1  Abstract
-Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves.
-A lack of empirical data relating photosynthesis to leaf physiological behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
-To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
-Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
-When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves at the same PPFD.
-Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
-We argue that the rapid response of g~m~ with light enables shade leaves to respond quickly to sunflecks.
-Evidence that g~m~ not only varies within a canopy but can be up-regulated over short time intervals possibly represents a new mechanism underpinning leaf gas exchange responses to light.
-
-<!-- ## Key words -->
-<!-- photosynthesis, stomatal conductance, mesophyll conductance,  shade, leaf optimal behavior -->
-
-#3.2  Introduction
-Light availability is one of the most important environmental drivers of leaf carbon (C) uptake in trees.
-Predicting C uptake if forests usually involves upscaling leaf level measurements to assess whole canopy function.
-Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling based on solar irradiance problematic.
-Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor].
-Consequently, structural and functional properties of leaves are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of leaf photosynthesis (*A~n~*) [@evans1995carbon].
-To estimate whole canopy C gain it is thus necessary to account for the non-linear response of A to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
-
-\
-The distribution of resources required for *A~n~*, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients.
-As *A~n~* has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
-Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic].
-Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
-Higher rates of *A~n~* and stomatal conductance (g~s~) can only be sustained if the leaf specific hydraulic conductance (K~l~) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal].
-Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A~n~* if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co], thus K~l~ should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
-
-\
-Rates of photosynthesis in C~3~ plants are limited by the [CO~2~] available for fixation by Rubisco within the chloroplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmosphere to the site of carboxylation [@warren2008stand].
-This drawdown consists of multiple resistance pathways to CO~2~ diffusion which include CO~2~ diffusion from the atmosphere through stomata (stomatal conductance, g~s~) and then from these intercellular air spaces into the chloroplast (mesophyll conductance, g~m~).
-Based on optimal theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize *A~n~* [@peltoniemi2012co].
-This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
-Mesophyll conductance can also impose limitations on *A~n~* as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A~n~* [@niinemets2007photosynthesis] if g~m~ constrains CO~2~ supply to the chloroplast.
-Part of the variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
-Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack empirical data currently hinders are ability to interpreting their coupled responses to *A~n~* across sun and shade leaves.
-
-\
-Additionally, assessing shade leaf behaviour is made difficult with accounting of short term light fluctuations within a canopy, via sunflecks.
-How shade leaves utilize sunflecks for short term C gain depends on the combined response time of g~s~ and g~m~ and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks;@tausz2005dynamic].
-The utilization of sunflecks is first limited by delayed responses of stomata opening, which may take minutes, effectively limiting the maximum assimilation rate that can be achieved [@pearcy1990sunflecks; @vicomanz2011; @way2012sunflecks].
-Mesophyll conductance has been shown to respond to environmental factors (e.g.
-CO~2~, temperature or vapor pressure deficit) at timescales of minutes, possibly faster than g~s~ [@flexas2008mesophyll and references therein], yet short term response to light availability are unclear.
-For example, g~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid].
-Anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are also unlikely to adjust during short light fluctuations.
-The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance and as a potential response to transitory light availability.
-
-\
-Climate warming may also affect the physiological behaviour of leaves within a canopy.
-This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis].
-How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for upscaling *A~n~* for sun and shade leaves.
-Additionally, light saturated rates of A are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
-How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients.
-The impacts of warming on plant physiological processes are obviously vast, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.
-
-\
-In this study we use *Eucalyptus tereticornis* Sm.
-trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to empirically evaluate the distribution of nitrogen and water supply and leaf physiological behaviour of sun and shade leaves.
-Our hypotheses are as follows:
-
-\
-1.	If whole tree canopies are optimized for C gain, then leaf N, hydraulic conductance and biochemical photosynthetic capacity are predicted to be higher in sun leaves compared to shade leaves.
-
-\
-2.	Stomatal conductance is proportional to *A~n~* across sun and shade leaves under similar leaf VPD and g~m~ scales positively with photosynthetic capacity.
-
-\
-3.
-As shade leaves were expected to develop lower biochemical photosynthetic capacity and leaf physiological responses to light are typically slow (minutes), increases in *A~n~* following sunfleck simulations were not expected to reach rates of full sun leaves.
-
-\
-4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporation demand from higher temperatures will lead to stomatal closure.
-
-#3.3  Materials and Methods
-
-##3.3.1 Whole tree chamber experimental design
-Twelve *Eucalyptus tereticornis* seedlings, chosen from a single local Cumberland plain cohort, were planted in March 2013 into 12 whole-tree chambers (WTC) at the Hawkesbury Forest Experiment site near Richmond, NSW, Australia.
-Each chamber has a height of 9 m and seedlings were grown for 15 months.
-A detailed description of the WTC operation and design is available in [@barton2010whole].
-Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarfig`).
-Trees grew quickly and developed large canopies, with height growth reaching the top of the WTCs over the experiment duration.
-Trees were watered weekly with 70 L from March 2013 to November 2013.
-From October 2013 to the end of the experiment trees were watered every 15 days with the mean monthly rainfall amount for Richmond, NSW.
-In February 2013 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering.
-Due to a limited range of data for the drought treatment only well-watered trees are reported, which reduces the sample size of WTC (n=6 to n=3) for the final 3 months of the experiment.
-
-\
-Before seedlings were planted into each chamber they were maintained under well-watered conditions in 25 L pots and kept inside each chamber.
-This allowed for seedlings to gain sufficient size before planting while also allowing them to acclimate to chamber temperature treatments.
-Seedlings were planted into each chamber after mean seedling height reached 100 cm.
-The top soils at this site, used in both pots and chambers, are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg-1 total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity.
-A root exclusion barrier extended from chamber walls to the hard layer (ca.
-1 m) and roots were allowed to grow freely below the barrier.
-Leaf gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.
-At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm, height of 348&pm;15.1 cm and an leaf area of 3.9&pm;0.1 m^2^.
-For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an leaf area of 6.2&pm;0.2 m^2^.
-Leaf area was calculated based on leaf counts and mean leaf size.
-
-##3.3.2 Leaf gas exchange, online carbon isotope discrimination and mesophyll conductance
-Leaf gas exchange measurements were performed six times, beginning in October 2013 and monthly from December 2013 to April 2014.
-Measurements were taken on a representative sun and shade leaf for each tree during each measurement campaign.
-The newest fully expanded leaf from the branch apex was chosen for gas exchange measurements and sun leaves were measured in the upper third of the canopy.
-Here, shade leaves are defined as inner-canopy leaves developing on secondary branches in a low light environment.
-Shade leaves were always measured in the lower canopy, but leaves were sampled on subsequent higher branches across measurement campaigns to minimize confounding effects of leaf age.
-As shade leaves most likely developed slower this assured that older leaves in the bottom canopy were avoided.
-
-\
-Prior to gas exchange measurements photosynthetic photon flux density (PPFD) was recorded as a point measurement at the individual leaf level and a spatially averaged measurement at the canopy position for each selected leaf.
-A hand-held quantum sensor (LI-COR, Lincoln, NE, USA) was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either full sun or full shade.
-A ceptometer (AccuPAR LP-80, Decagon Devices, Pullman, USA) was then used to measure a spatially averaged PPFD at the canopy height of each chosen leaf type.
-Each ceptometer reading integrated an array of 80 sensors over a total length of 84 cm.
-Five ceptometer readings were recorded at different locations within the canopy, but at the same height and close to each selected leaf.
-The mean of these readings was assumed to represent the local light environment of sun and shade leaves for each tree.
-All measurements of PPFD and gas exchange were performed on sunny days between 10:00-14:30 h.
+Synchronized gas exchange and C isopte discrimination measurements were similarily as described in Evans & von Caemmerer (2013)
 
@@
 The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck where then compared as a function of C~i~:C~a~.
-
-##3.3.3 Biochemical parameters of photosynthesis
-Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each WTC prior to the initiation of the drought treatment.
-Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011].
-
-##3.3.4 Leaf chemistry and hydraulic parameters
-Following gas exchange measurements each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, dried and weighed.
-These leaves were then milled and analyzed for leaf N content and $\delta$^13^C.
-Leaf samples were analysed on a Delta V Advantage coupled to a Flash HT and Conflo IV (Thermo Fisher Scientific, Bremen, Germany) in dual-reactor setup.
-Samples were flash combusted at 1000&deg;C and converted to CO~2~ and N~2~ and then subjected to stable isotope ratio mass spectrometry.
-Leaf N is reported on an area basis (N~a~, g m^-2^) and isotopic signatures are reported relative to the VPDP scale.
-
-\
-Predawn ($\Psi_{pd}$) and midday ($\Psi_l$) leaf water potentials (MPa) were measured for sun and shade leaves during each gas exchange campaign using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA).
-The leaf closest to the leaf used for gas exchange was sampled for measurement of $\Psi_{pd}$.
-Predawn leaf water potential was measured before sunrise on the same day as gas exchange measurements.
-Leaves used for gas exchange were immediately sampled for $\Psi_l$ once measurements were completed.
-All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
-Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) through the equation:
-
-\
-(12)
-$$K = \frac{E}{\Psi_{pd}~-\Psi_l}$$
-
-Leaf level instantaneous transpiration efficiency (ITE) was calculated as A divided by E.
-The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression, where k=0.5 [see @duursma2013near].
-Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use [@lin2015optimal].
-
-##3.3.5 Data analysis
-Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R [@RDevelopmentCoreTeam2011] with WTC as a random effect.
-Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general].
-Confidence intervals (95 %) of mixed effect linear models were generated using bootstrapping methods with 999 simulations, using the bootMer function in the 'lme4' package [@bates2013lme4].
-For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package, using WTC as a random effect.
-All tests of statistical significance were conducted at an alpha of 0.05.
-
-#3.4  Results
-
-##3.4.1 Leaf resource distribution
-Across six measurement campaigns over the 7 month period PPFD was  reduced on average by >75% in the shade (Figure `r parfig`).
-Leaf-specific hydraulic conductance (K~l~) was similar across sun and shade leaves (Table `r wtc3tab1`).
-This was because $\Psi_{pd}$, $\Psi_l$ (Table `r wtc3tab1`) and E (Table `r wtc3tab2`) did not differ between leaf types.
-Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table1).
-Leaf mass per area was not different between leaf types.
-No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~area~ or LMA either within or across leaf types.
-\
-```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/par2.R")
-```  
-\
-**Figure 3.1**.
-Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014.
-Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf.
-Each date represents the starting date for each measurement campaign.
-Points represent the mean (&pm; 1 standard error) daily maximum air temperature during each campaign period.
-
-\
-**Table 3.1**.
-*Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments.
-Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
-Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
-Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;.
-Different letters represent significant differences between leaf type and temperature treatments.
-The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.
-
-```{r, echo=FALSE}
-resourcetab2 <- rbind(resourcetab, c("P value","", as.character(pval)))
-greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
-
-dust(resourcetab2)%>%
-  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0(greeks['delta'],"^13^c")) %>%
-  sprinkle(rows= 1, part="head",bold=TRUE)%>%
-  sprinkle_print_method("markdown")
-```
-
-##3.4.2 Photosynthetic capacity and leaf photosynthesis rates 
-The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from AC~i~ curves (Figure `r acifig`a).
-Within leaf types, no effect of the warming treatment was detected on either parameter.
-Among the sampled leaves, Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
-\
-```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/aci_vcmaxnitro.R")
-```  
-\
-**Figure 3.2**.
-(a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.
-ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-(b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light.
-(c) The relationship between A and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
-For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
-
-\
-Mean *A~n~* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table `r wtc3tab2`).
-Additionally, leaf N~a~ was positively related to *A~n~* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
-Following an increase in light intensity to match high-light conditions, A of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table `r wtc3tab2`).
-No effect of the warming treatment was detected on *A~n~* of sun leaves measured at high light or shade leaves at either low or high light.
-Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
-
-\
-**Table 3.2**.
-Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
-Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).
-Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa.
-Different letters represent significant differences between leaf type, light environment and temperature treatments.
-The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.
-
-```{r, echo=FALSE}
-getab2 <- rbind(getab, c("P value","","", as.character(pval2)))
-greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
-
-dust(getab2)%>%
-  sprinkle_colnames("Leaf", "Light","Temperature", "A", "g~s~", "g~m~", "ITE","E","VPD","C~i~", "C~c~") %>%
-  sprinkle(rows= 1, part="head",bold=TRUE)%>%
-  sprinkle_print_method("markdown")
-```
-
-##3.4.3 Stomatal conductance and leaf water-use efficiency
-On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table `r wtc3tab2`).
-Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and *A~n~* were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
-Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
-No effect of the warming treatment was detected on g~s~ within or across leaf types.
-Stomatal conductance within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get2fig`b).
-\
-```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/Agmgs.R")
-```  
-\
-**Figure 3.3**.
-The response of A to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
-Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
-Grey areas are 95% confidence intervals from the mean.
-\
-```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/physiology_barplots.R")
-```  
-\
-**Figure 3.4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
-
-\
-Measured under ambient light and temperature conditions leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001).
-Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
-Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment, but no warming effect was detected in shade leaves measured at low or high light (Table `r wtc3tab2`).
-The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
-For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal ITE model with a g~1~ value for each leaf type and treatment (Figure `r waterusefig`a).
-Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).
-\
-```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=600}
-source("master_scripts/wateruse.R")
-```  
-\
-**Figure 3.5**.
-(a) Response of instantaneous transpiration efficiency (ITE) to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
-(b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light.
-For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment.
-For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
-
-\
-Bulk-leaf $\delta$^13^C, as an index of integrated water-use efficiency [@marshall2007sources], was significantly lower in shade leaves compared to sun leaves by ca.
-2&permil; (Table `r wtc3tab1`).
-No effects of the warming treatment on leaf $\delta$^13^C were detected.
-Leaf $\delta$^13^C and N~area~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C (higher water-use efficiency) and higher N investment in sun leaves.
-
-##3.4.4 Leaf carbon isotope discrimination and mesophyll conductance
-The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
-Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
-The remainder $\Delta_o$ consists of the contributions of g~s~, respiration and photorespiration to discrimination.
-
-\
-```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/delta_cica.R")
-```  
-\
-**Figure 3.6**.
-Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
-The solid line represents the theoretical line for C3 plants from Evans et al.
-[-@evans1986carbon].
-
-\
-Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
-Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table `r wtc3tab2`).
-Proportional increases in g~m~ were matched by proportional increases in *A~n~* from low to high light in shade leaves (Figure `r gebarfig`b,c).
-Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186).
-The large increases in g~m~ in shade leaves under high light resulted in the highest rates of *A~n~* (Figure `r agmgsfig`b).
-No differences in g~m~ were detected with the warming treatment within leaf types.
-Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively).
-We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~.
-Comparison of A-Cc curves (Figure `r accfig`) and A-Ci curves revealed similar differences between sun and shade leaves.
-
-##3.4.5 Variation in intercellular and chloroplastic CO~2~ concentrations
-Higher rates of g~s~ in shade leaves under low and high light lead to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
-The chloroplast CO~2~ partial pressure was comparable between shade leaves when measured at both low and high light conditions (Figure `r ciccfig`c).
-In sun leaves C~c~ was significantly lower than shade leaves as a likely consequence of a lower C~i~.
-The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and *A~n~* [@caemmerer2014temperature].
-This drawdown was similar between sun and shade leaves measured at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
-This was the result of the proportional relationship between g~m~ and *A~n~* across all leaves.
-The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
-\
-```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/cicc_bar.R")
-```  
-\
-**Figure 3.7**.
-The mean &pm; 1 standard error of intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light.
-
-#3.5  Discussion
-Here we show that *A~n~* in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
-Although shade leaves in lower light environments exhibited relatively high g~s~ and c~i~, it was rapid increases in g~m~ under periods of high light availability that led to this up-regulation of *A~n~*.
-Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
-These findings offer new insights into how aspects of leaf physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake in future studies.
-Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology.
-
-##3.5.1 Resource distribution and photosynthetic capacity
-The allocation of N~a~ constrains *A~n~* and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
-Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979).
-As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical].
-Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and *A~n~* were all reduced in shade leaves.
-Leaf mass area, however, was not different between sun and shade leaves.
-This could be due to leaf formation under comparative light conditions or possible differences in total non-structural carbohydrates contents between leaf types.
-
-\
-Photosynthesis is also limited by the ability to supply water to the upper canopy.
-Ultimately, the ability of a trees hydraulic architecture to supply water to foliage across increasing pathlengths affects productivity and survival [@sellin2008effects].
-Using a two-leaf model, Peltoniemi et al.
-[-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if K~l~ is also optimally distributed.
-In this study, variation in leaf N distribution and *A~n~* rates were not associated with subsequent changes in K~l~ between sun and shade leaves.
-Thus, no direct relationship between water supply and N distribution or *A~n~* within the canopy were detected.
-
-\
-Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased ITE throughout the experiment.
-Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was likely prevalent long term.
-From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A~n~* in sun leaves was characterized by low rates of g~s~ and low C~i~.
-Relative to the differences in *A~n~* between leaf types, higher rates of g~s~ in shade leaves appear to exhibit inefficient water use.
-As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower ITE compared to sun leaves.
-
-##3.5.2 Physiological behaviour of sun and shade leaves
-The pattern of inefficient water use in shade leaves is important as we hypothesized that g~s~ and *A~n~* would be proportional across sun and shade leaves.
-In sun leaves, *A~n~* and g~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
-However, lower rates of *A~n~* in shade leaves were not coupled with decreases in g~s~, explaining the observed decreases in ITE.
-This is significant as optimal stomatal regulation to balance C gain with water loss has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on full sun leaves [see @prentice2014balancing; @lin2015optimal].
-As a result, the often used economic framework of balancing costs of using water versus N allocation to predict *A~n~* [@wright2003least] may break down when considering all leaves within a tree canopy.
-
-\
-It is possible that reducing stomatal response time, by sustaining higher g~s~, is a strategy to take advantage of high light quickly in shade leaves [@tausz2005dynamic].
-Evidence from this study supports this hypothesis, as shade leaves increased A equivalent or even outperforming sun leaves under identical light intensity.
-Transpiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck induced thermal load [@schymanski2013stomatal].
-This is because rapid increases in leaf temperature with sunflecks have been shown to inhibit C gain [@leakey2003high].
-However, this response likely occurs at very high air temperatures and may not explain the observed g~s~ in shade leaves across the large natural range of temperatures included in this study.
-How prevalent each of these strategies are within tree canopies is still unknown, as empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants, often in deep shade [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution].
-Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy [see @de1997simple].
-
-\
-We found that *A~n~* and g~m~ scaled positively across leaf types and, surprisingly, increased rapidly (within minutes) and proportionately when light intensity was increased in shade leaves.
-Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
-This provides a potential explanation for the observed rapid increases in g~m~ but the impacts of aquaporins on g~m~ are untested in leaves of tree species.
-Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
-Here we provide empirical data showing g~m~ not only varies within a canopy but the up-regulation of g~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.
-
-\
-If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution.
-This is because the role of g~s~ in regulating photosynthetic induction impacts the capacity of a leaf to utilize sunflecks [@way2012sunflecks].
-If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan.
-For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to tolerate decreases in ITE.
-However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.
-Thus, models which predict leaf photosynthesis from N distribution within a canopy will be incomplete unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].
-
-##3.5.3 Conclusions
-Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy.
-Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
-These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have important implications for process-based models that predict canopy C gain from rates of leaf photosynthesis.
-Furthermore, the dynamic nature of g~m~ cannot be simply parameterized in tree growth models and possibly should be excluded until it can be represented properly.
-Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown.
-To improve our ability to predict whole canopy C gain future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.
-
-#3.6  Supporting Information
-
-```{r airvars, fig.width=7, fig.height=8.8, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/air_vars.R")
-```
-\
-**Figure 3.S1**. Daily maximum and minimum temperature (a) and total daily PPFD (b) for each chamber across the experiment duration.
-\
-```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/ACc_figure.R")
-```  
-\
-**Figure 3.S2**.
-Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
-C~c~ values were predicted with g~m~, thus curves represent chloroplastic photosynthetic parameters at 25°C and saturating light (1800 $\mu$mols m^-1^ s^-1^).
-\
-```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/gasex_temp3.R")
-```  
-\
-**Figure 3.S3**.
-Response of A (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
-Shaded symbols represents each monthly measurement campaign.
-Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
-All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
-Weak negative relationships with g~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (R^2^ = 0.16 and 0.08, respectively).
-\
-```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/gasex_temp4.R")
-```  
-\
-**Figure 3.S4**.
-Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
-Shaded symbols represents each monthly measurement campaign.
-Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
-All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
-Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (R^2^ = 0.73, 0.58 and 0.72, respectively).
-
-# References {.unnumbered}

### 2015-12-15 markdown/methods_condensed.Rmd
started new methods

@@
 
-Synchronized gas exchange and C isopte discrimination measurements were similarily as described in Evans & von Caemmerer (2013)
+Synchronized gas exchange and C isopte discrimination measurements were made similarily as described in Tazoe et al. [-@tazoe2011using] and Evans & von Caemmerer [-@evans2013temperature].
 
@@
 Using online C isotope discrimination measurements, the difference between the observed discrimination and what is predicted for light saturated gas exchange is proportional to g~m~ [@griffiths2013mesophyll].
-First, leaf C isotope discrimination was calculated by comparing the isotopic composition of the reference gas entering the leaf cuvette ($\delta$^13^C~e~) with the sample gas ($\delta$^13^C~o~) such that:
 
-\
-(1)
-$$\delta = \left(\frac{R_s}{R_{stnd}} -1\right)1000$$
+In this experiment, a modified form of this equation presented in Evans & Von Caemmerer [-@evans2013temperature] with ternary effect corrections by Farquhar & Cernusak [-@farquhar2012ternary].
 
-where R~s~ is the isotopic ratio of the sample and R~stnd~ is the isotopic ratio of the standard Vienna Pee Dee Belemnite (VPDP).
-Next, the observed discrimination ($\Delta$~obs~) is calculated from Evans et al.
-[-@evans1986carbon]:
+For this study, the light compenstation point ($\Gamma_{*}$) and dark respiration parameters dervived for tobaco plants in [@evans2013temperature, tazoe?] where replaced with parameters derived for *Eucalyptus*.
+A standard Arrhenius function was used to calculate \Gamma_{*}$ and R~d~ *Eucalyptus globulus* from [@crous2012light].
 
-\
-(2)
-$$\Delta_{obs} = \frac{\xi(\delta^13C_o-\delta^13C_e)}{1000+\delta^13C_e-\xi(\delta^13C_o-\delta^13C_e)}$$
-
-where:
-
-\
-(3)
-$$\xi = \frac{C_{e}}{C_{e}-C_{o}}$$
-
-and $\xi$ is the ratio of the CO~2~ entering the well mixed leaf cuvette to the CO~2~ draw down in the gas stream by the leaf.
-
-\
-Second, C isotope discrimination during C~3~ photosynthesis ($\Delta$) is the resultant discrimination from CO~2~ diffusion from the atmosphere to the site of carboxylation, consisting of a series of fractionation steps described in Evans et al.
-[-@evans1986carbon].
-In this experiment, a modified form of this equation presented in Evans & Von Caemmerer [-@evans2013temperature] with ternary effect corrections by Farquhar & Cernusak [-@farquhar2012ternary] was used such that:
-
-\
-(4)
-$$\Delta_o = \Delta_i - \Delta_{gm} - \Delta_e - \Delta_f$$
-
-where $\Delta$~o~ is the observed discrimination and  $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, g~m~, respiration and photorespiration, respectively.
-The equations for each are as follows:
-
-\
-(5)
-$$
-\begin{aligned}
-\Delta_i = &\frac{1}{1-t}a_i+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}
-\end{aligned}
-$$
-
-\
-(6)
-$$
-\begin{aligned}
-\Delta_{gm} = &\frac{1+t}{1-t}\left(b-a'-\frac{eR_d}{A+R_d}\right)\frac{A}{g_mC_a}
-\end{aligned}
-$$
-
-\
-(7)
-$$
-\begin{aligned}
-\Delta_e = &\frac{1+t}{1-t}\left(\frac{eR_d}{(A+R_d)C_a}(C_i-\Gamma_*)\right)
-\end{aligned}
-$$
-
-\
-(8)
-$$
-\begin{aligned}
-\Delta_f = & \frac{1+t}{1-t}\left(f\frac{\Gamma_*}{C_a}\right)\\
-\end{aligned}
-$$
-
-where the different fractionation factors include; diffusion through water (a~i~, 1.8&permil;), Rubisco carboxylation (b, 29&permil;), the photorespiratory fractionation (f, 16.2&permil;) and the combined fractionation through the boundary layer and the stomata (a').
-a' is defined by:
-
-\
-(9)
-$$a' = \frac{a_b(C_a-C_i)+a(C_s-C_i)}{C_a-C_i}$$
-
-where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ is the fractionation from boundary layer diffusion (2.9&permil;) and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon].
-C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~).
  In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulus* from [@crous2012light].
@@
 
-\
-(10)
-$$t = \frac{(1+a')E}{2g^t_{ac}}$$
+ Examples of this approach to measure gas exchange and C isotope discrimination are presented in Evans and Von Caemmerer [-@evans2013temperature].
 
-where E denotes the transpiration rate and $\text{g}^t_{ac}$ is the total conductance to CO~2~ diffusion to both the boundary layer and stomatal conductance.
-
-\
-The CO~2~ diffusion from the intercellular airspace to the chloroplast, g~m~, is given by its relationship to the leaf photosynthesis rate (*A~n~*) by:
-
-\
-(11)
-$$g_m = \frac{*A~n~*}{C_i-C_c}$$
-
-where C~c~ is the chloroplast CO~2~ partial pressure.
-Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were then estimated using Equation 11.
-Examples of this approach to measure gas exchange and C isotope discrimination are presented in Evans and Von Caemmerer [-@evans2013temperature].
  The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck where then compared as a function of C~i~:C~a~.

### 2015-12-15 markdown/methods_condensed.Rmd
simple methods draft

@@
 
-#from SVCEVANS
-Gas exchange and carbon isotope discrimination measurements were made as described by Tazoe et al.
-(2011) and Evans & von Caemmerer (2013).
-Plants were placed in a temperature-controlled growth cabinet together with the measuring heads (6 cm2 chamber with a red blue lightemitting diode light source) of two LI-6400XT gas exchange systems (Li-Cor, Lincoln, NE, USA).
-These chambers were coupled to a tunable diode laser (TDL;TGA100,Campbell Scientific, Inc., Logan, UT, USA) with a T junction in the match valve tubing.
-This allowed leaves of two plants to be measured in sequence, with each LI-6400XT sampled by the TDL at 4 min intervals as outlined in Fig.
-1 of Tazoe et al.
-(2011).
-The LI-6400XT consoles were placed outside the growth chambers.
-The CO2 in the leaf chamber was set at 380 mol mol−1 in the sample cell, flow rate at 200 mol s−1 and irradiance of 1500 or 1000 mol quantam−2 s−1.N2 andO2 were mixed by mass flow controllers (Omega Engineering Inc., Stamford, CT, USA) to generate 2% O2, which was supplied to the consoles and specified for the LI-6400 calculations.
-Leaves were stabilized first at 25 °C, then temperature was dropped to 15 °C and then increased in 5 °C steps.
-Mesophyll conductance was calculated from carbon isotope discrimination with equations and fractionation factors presented in Evans & von Caemmerer (2013) including the ternary corrections proposed by Farquhar & Cernusak (2012).
-Low O2 partial pressure and high light minimized the contribution of respiratory and photorespiratory fractionation to the carbon isotope signal.
-
-Synchronized gas exchange and C isopte discrimination measurements were made similarily as described in Tazoe et al. [-@tazoe2011using] and Evans & von Caemmerer [-@evans2013temperature].
-
-\
+Synchronized gas exchange and C isotope discrimination measurements were made similarly as described in Tazoe et al.
+[-@tazoe2011using] and Evans & von Caemmerer [-@evans2013temperature].
 Leaf level gas exchange was measured with a standard (2 x 3 cm) leaf cuvette using a portable gas exchange system (LI-6400XT, LI-COR, Lincoln, NE, USA).
@@
 Periods of high irradiance were simulated for shade leaves by increasing the leaf cuvette PPFD (LI-COR red/blue light source) to match the light environment of the full sun leaf in the same tree.
-The maximum sunfleck response of shade leaves was then recorded once CO~2~ and water vapor fluxes re-stabilized in the leaf cuvette (ca.
+The maximum sunfleck response of shade leaves was then recorded once CO~2~ and water vapour fluxes re-stabilized in the leaf cuvette (ca.
 25 min).
@@
 \
-Once CO~2~ and water vapor flux values were stable for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet of the LI-6400.
+Once CO~2~ and water vapour flux values were stable for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet of the LI-6400XT.
 The gas streams were dried by passing through napion gas dryers in the respective gas lines, and then ^12^CO~2~ and ^13^CO~2~ concentrations were measured for each gas stream by the TDL.
@@
 The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDL on each measurement cycle.
-Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the same 12 min interval.
-
-\
-Using online C isotope discrimination measurements, the difference between the observed discrimination and what is predicted for light saturated gas exchange is proportional to g~m~ [@griffiths2013mesophyll].
-
-In this experiment, a modified form of this equation presented in Evans & Von Caemmerer [-@evans2013temperature] with ternary effect corrections by Farquhar & Cernusak [-@farquhar2012ternary].
-
-For this study, the light compenstation point ($\Gamma_{*}$) and dark respiration parameters dervived for tobaco plants in [@evans2013temperature, tazoe?] where replaced with parameters derived for *Eucalyptus*.
-A standard Arrhenius function was used to calculate \Gamma_{*}$ and R~d~ *Eucalyptus globulus* from [@crous2012light].
-
- In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulus* from [@crous2012light].
-The ternary effect corrections (t) are described by:
-
- Examples of this approach to measure gas exchange and C isotope discrimination are presented in Evans and Von Caemmerer [-@evans2013temperature].
-
- The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck where then compared as a function of C~i~:C~a~.
+Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the 12 min interval.
+Mesophyll conductance was calculated from carbon isotope discrimination with equations and fractionation factors as presented in Evans & von Caemmerer (2013), including the ternary corrections proposed by Farquhar & Cernusak (2012).
+For this study, the CO~2~ compensation point ($\Gamma_{*}$) and respiration during the day (R~d~) equation parameters originally derived for tobacco plants [@von1994kinetics] were replaced with parameters derived for *Eucalyptus globulus* from Crous et al.
+[-@crous2012light].

### 2015-12-15 markdown/methods_condensed.Rmd
finished gm methods condensed

@@
 Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the 12 min interval.
-Mesophyll conductance was calculated from carbon isotope discrimination with equations and fractionation factors as presented in Evans & von Caemmerer (2013), including the ternary corrections proposed by Farquhar & Cernusak (2012).
-For this study, the CO~2~ compensation point ($\Gamma_{*}$) and respiration during the day (R~d~) equation parameters originally derived for tobacco plants [@von1994kinetics] were replaced with parameters derived for *Eucalyptus globulus* from Crous et al.
-[-@crous2012light].
+
+\
+Mesophyll conductance was calculated from carbon isotope discrimination with equations and fractionation factors as presented in Evans & von Caemmerer (2013), including the ternary corrections proposed by Farquhar & Cernusak (2012), such that:
+
+\
+(1)
+$$\Delta_o = \Delta_i - \Delta_{gm} - \Delta_e - \Delta_f$$
+
+where $\Delta$~o~ is the observed discrimination and $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, g~m~, respiration and photorespiration, respectively.
+For this study, the CO~2~ compensation point ($\Gamma_{*}$) and respiration during the day (R~d~) parameters originally derived for tobacco plants [@von1994kinetics] were replaced with parameters derived for *Eucalyptus globulus* from Crous et al.
+[-@crous2012light] when calculating g~m~.
+The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck where compared as a function of C~i~:C~a~.
+Once g~m~ was caluclated the chloroplast CO~2~ partial pressure (C~c~) and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were able to be estimated from the relationship between g~m~ to the leaf photosynthesis rate (*A~n~*) by:
+
+\
+(2)
+$$g_m = \frac{*A~n~*}{C_i-C_c}$$

### 2015-12-16 manuscript/manuscript_wtc3.Rmd
started new methods
photosynthesis abbr changes

@@
 (b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light.
-(c) The relationship between A and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
+(c) The relationship between *A~n~* and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
 For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
@@
 Additionally, leaf N~a~ was positively related to *A~n~* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
-Following an increase in light intensity to match high-light conditions, A of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table `r wtc3tab2`).
+Following an increase in light intensity to match high-light conditions, *A~n~* of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table `r wtc3tab2`).
 No effect of the warming treatment was detected on *A~n~* of sun leaves measured at high light or shade leaves at either low or high light.
@@
 Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).
-Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa.
+Units for *A~n~* and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa.
 Different letters represent significant differences between leaf type, light environment and temperature treatments.
@@
 dust(getab2)%>%
-  sprinkle_colnames("Leaf", "Light","Temperature", "A", "g~s~", "g~m~", "ITE","E","VPD","C~i~", "C~c~") %>%
+  sprinkle_colnames("Leaf", "Light","Temperature", "*A~n~*", "g~s~", "g~m~", "ITE","E","VPD","C~i~", "C~c~") %>%
   sprinkle(rows= 1, part="head",bold=TRUE)%>%
@@
 **Figure 3.S3**.
-Response of A (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
+Response of *A~n~* (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
 Shaded symbols represents each monthly measurement campaign.

### 2015-12-16 manuscript/tablesfigures.Rmd
photosynthesis abbr changes

@@
 **Figure 3**.
-The response of A to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
+The response of *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
 Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
@@
 \
-**Figure 4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
+**Figure 4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
 
@@
 **Figure S3**.
-Response of A (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
+Response of *A~n~* (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
 Shaded symbols represents each monthly measurement campaign.

### 2015-12-18 manuscript/manuscript_wtc3.Rmd
started RD comments on last version

@@
 To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
-Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
-When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves at the same PPFD.
+Compared to sun leaves, shade leaves had lower *A~n~*, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
+When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in *A~n~* greater than sun leaves at the same photosynthetic photon flux density.
 Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
@@
 Light availability is one of the most important environmental drivers of leaf carbon (C) uptake in trees.
-Predicting C uptake if forests usually involves upscaling leaf level measurements to assess whole canopy function.
+Predicting C uptake of forests usually involves upscaling leaf level measurements to assess whole canopy function.
 Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling based on solar irradiance problematic.
@@
 Consequently, structural and functional properties of leaves are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of leaf photosynthesis (*A~n~*) [@evans1995carbon].
-To estimate whole canopy C gain it is thus necessary to account for the non-linear response of A to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
+To estimate whole canopy C gain it is thus necessary to account for the non-linear response of *A~n~* to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
 
@@
 Part of the variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
-Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack empirical data currently hinders are ability to interpreting their coupled responses to *A~n~* across sun and shade leaves.
+Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack of empirical data currently hinders are ability to interpreting their coupled responses to *A~n~* across sun and shade leaves.
 
 \
-Additionally, assessing shade leaf behaviour is made difficult with accounting of short term light fluctuations within a canopy, via sunflecks.
+Additionally, accounting for short term light fluctuations within a canopy, via sunflecks, makes assessing shade leaf behaviour difficult.
 How shade leaves utilize sunflecks for short term C gain depends on the combined response time of g~s~ and g~m~ and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks;@tausz2005dynamic].
@@
 How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients.
-The impacts of warming on plant physiological processes are obviously vast, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.
+The impacts of warming on plant physiological processes are vast, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.
 
@@
 \
-3.
-As shade leaves were expected to develop lower biochemical photosynthetic capacity and leaf physiological responses to light are typically slow (minutes), increases in *A~n~* following sunfleck simulations were not expected to reach rates of full sun leaves.
+3. As shade leaves were expected to develop lower biochemical photosynthetic capacity, increases in *A~n~* following sunfleck simulations were not expected to reach rates of full sun leaves.
 
 \
-4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporation demand from higher temperatures will lead to stomatal closure.
+4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporation demand from higher temperatures may lead to stomatal closure.
 
@@
 This allowed for seedlings to gain sufficient size before planting while also allowing them to acclimate to chamber temperature treatments.
-Seedlings were planted into each chamber after mean seedling height reached 100 cm.
+Seedlings were planted into each chamber after mean seedling height reached ca.
+100 cm.
 The top soils at this site, used in both pots and chambers, are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg-1 total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity.
@@
 For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an leaf area of 6.2&pm;0.2 m^2^.
-Leaf area was calculated based on leaf counts and mean leaf size.
+Leaf area was calculated based on complete leaf counts and mean leaf size from a subsample.
 
@@
 The CO~2~ in the leaf cuvette was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mol s^-1^.
-Two identical gas exchanges systems were run simultaneously, one in each of a randomly chosen WTC for each temperature treatment.
+Two identical gas exchanges systems were run simultaneously, one in each of a AT or ET WTC.
 Leaf temperatures were controlled at the current AT or ET WTC air temperature.
@@
 C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~).
-In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulus* from [@crous2012light].
+In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulus* from Crous et al.
+[-@crous2012light].
 The ternary effect corrections (t) are described by:
@@
 (11)
-$$g_m = \frac{*A~n~*}{C_i-C_c}$$
+$$g_m = \frac{A_n}{C_i-C_c}$$
 
@@
 Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011].
+From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011] using default parameters.
 
@@
 All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
-Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) through the equation:
+Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) with the equation:
 
@@
 
-Leaf level instantaneous transpiration efficiency (ITE) was calculated as A divided by E.
+Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A~n~* divided by E.
 The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression, where k=0.5 [see @duursma2013near].
@@
 For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package, using WTC as a random effect.
-All tests of statistical significance were conducted at an alpha of 0.05.
+All tests of statistical significance were conducted at an $\alpha$ of 0.05.
 
@@
 Leaf mass per area was not different between leaf types.
-No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~area~ or LMA either within or across leaf types.
+No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~area~ or LMA either within or across leaf types (P > 0.05).
 \
@@
 dust(resourcetab2)%>%
-  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0(greeks['delta'],"^13^c")) %>%
+  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0("$\\delta$","^13^c")) %>%
   sprinkle(rows= 1, part="head",bold=TRUE)%>%

### 2015-12-18 reproducibility_notes.R

@@ -14,3 +14,7 @@
 #leaving co-authors in thesis for now
 
 # airvars fig orginal height=9
+
+
+#tables
+#1. delta in column names does not bold properly

### 2015-12-30 manuscript/manuscript_wtc3.Rmd
finished the majority of comments from RD on chapter 4

@@
 where R~s~ is the isotopic ratio of the sample and R~stnd~ is the isotopic ratio of the standard Vienna Pee Dee Belemnite (VPDP).
-Next, the observed discrimination ($\Delta$~obs~) is calculated from Evans et al.
+Next, the observed discrimination ($\Delta$~o~) is calculated from Evans et al.
 [-@evans1986carbon]:
@@
 (2)
-$$\Delta_{obs} = \frac{\xi(\delta^13C_o-\delta^13C_e)}{1000+\delta^13C_e-\xi(\delta^13C_o-\delta^13C_e)}$$
+$$\Delta_{o} = \frac{\xi(\delta^13C_o-\delta^13C_e)}{1000+\delta^13C_e-\xi(\delta^13C_o-\delta^13C_e)}$$
 
@@
 **Table 3.2**.
-Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
+*Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
 Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).
@@
 **Figure 3.3**.
-The response of A to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
+The response of *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
 Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
-Grey areas are 95% confidence intervals from the mean.
+Grey areas are 95% confidence intervals for the mean.
 \
@@
 \
-**Figure 3.4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and A (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
+**Figure 3.4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
 
 \
-Measured under ambient light and temperature conditions leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001).
-Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).
+Measured under ambient light and temperature, leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001).
+Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was therefore significantly lower than sun leaves (P < 0.001).
 Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment, but no warming effect was detected in shade leaves measured at low or high light (Table `r wtc3tab2`).
@@
 For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment.
-For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
+For (b) the dashed line represents the significant linear model fit across all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
 
@@
 No effects of the warming treatment on leaf $\delta$^13^C were detected.
-Leaf $\delta$^13^C and N~area~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C (higher water-use efficiency) and higher N investment in sun leaves.
+Leaf $\delta$^13^C and N~a~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C (higher water-use efficiency) and higher N investment in sun leaves.
 
@@
 **Figure 3.6**.
-Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
+Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis ($\Delta_o$) and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
 The solid line represents the theoretical line for C3 plants from Evans et al.
@@
 Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186).
-The large increases in g~m~ in shade leaves under high light resulted in the highest rates of *A~n~* (Figure `r agmgsfig`b).
+The large increases in g~m~ in shade leaves under high light likely resulted in the highest rates of *A~n~* (Figure `r agmgsfig`b).
 No differences in g~m~ were detected with the warming treatment within leaf types.
@@
 We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~.
-Comparison of A-Cc curves (Figure `r accfig`) and A-Ci curves revealed similar differences between sun and shade leaves.
+Comparison of AC~c~ curves (Figure `r accfig`) and AC~i~ curves revealed similar differences between sun and shade leaves.
 
 ##3.4.5 Variation in intercellular and chloroplastic CO~2~ concentrations
-Higher rates of g~s~ in shade leaves under low and high light lead to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
+Higher rates of g~s~ in shade leaves under low and high light led to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
 The chloroplast CO~2~ partial pressure was comparable between shade leaves when measured at both low and high light conditions (Figure `r ciccfig`c).
-In sun leaves C~c~ was significantly lower than shade leaves as a likely consequence of a lower C~i~.
+In sun leaves C~c~ was significantly lower than shade leaves, consistent with a lower C~i~.
 The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and *A~n~* [@caemmerer2014temperature].
@@
 **Figure 3.7**.
-The mean &pm; 1 standard error of intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light.
+The mean &pm; 1 standard error of (a) intercellular CO~2~ concentration (C~i~), (b) CO~2~ concentration in the chloroplasts (C~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (C~i~-C~c~).
 
@@
 Here we show that *A~n~* in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
-Although shade leaves in lower light environments exhibited relatively high g~s~ and c~i~, it was rapid increases in g~m~ under periods of high light availability that led to this up-regulation of *A~n~*.
+Although shade leaves in lower light environments exhibited relatively high g~s~ and C~i~, it was rapid increases in g~m~ under periods of high light availability that allowed for this up-regulation of *A~n~*.
 Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
@@
 Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and *A~n~* were all reduced in shade leaves.
-Leaf mass area, however, was not different between sun and shade leaves.
-This could be due to leaf formation under comparative light conditions or possible differences in total non-structural carbohydrates contents between leaf types.
+Leaf mass per area, however, was not different between sun and shade leaves.
+This could be due to leaf formation under comparative light conditions or possible unmeasured differences in total non-structural carbohydrates contents between leaf types.
 
@@
 In sun leaves, *A~n~* and g~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
-However, lower rates of *A~n~* in shade leaves were not coupled with decreases in g~s~, explaining the observed decreases in ITE.
-This is significant as optimal stomatal regulation to balance C gain with water loss has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on full sun leaves [see @prentice2014balancing; @lin2015optimal].
+However, lower rates of *A~n~* in shade leaves were not coupled with decreases in g~s~, leading to the observed decreases in ITE.
+This is significant as optimal stomatal regulation to balance C gain with water loss has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on sun leaves [e.g.
+@prentice2014balancing; @lin2015optimal].
 As a result, the often used economic framework of balancing costs of using water versus N allocation to predict *A~n~* [@wright2003least] may break down when considering all leaves within a tree canopy.
@@
 It is possible that reducing stomatal response time, by sustaining higher g~s~, is a strategy to take advantage of high light quickly in shade leaves [@tausz2005dynamic].
-Evidence from this study supports this hypothesis, as shade leaves increased A equivalent or even outperforming sun leaves under identical light intensity.
+Evidence from this study supports this hypothesis, as shade leaves increased *A~n~* equivalent or even outperforming sun leaves under identical light intensity.
 Transpiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck induced thermal load [@schymanski2013stomatal].
@@
 Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
-This provides a potential explanation for the observed rapid increases in g~m~ but the impacts of aquaporins on g~m~ are untested in leaves of tree species.
+This provides a potential explanation for the observed rapid increases in g~m~ but the impacts of aquaporins on g~m~ are yet to be tested in leaves of tree species.
 Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
@@
 \
-**Figure 3.S1**. Daily maximum and minimum temperature (a) and total daily PPFD (b) for each chamber across the experiment duration.
+**Figure 3.S1**. Daily maximum and minimum temperature (a), daily maximum VPD (b) and total daily PPFD (c) for each chamber across the experiment duration.
 \
@@
 Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
-C~c~ values were predicted with g~m~, thus curves represent chloroplastic photosynthetic parameters at 25°C and saturating light (1800 $\mu$mols m^-1^ s^-1^).
+C~c~ values were predicted with g~m~ and curves represent chloroplastic photosynthetic parameters at 25°C and saturating light (1800 $\mu$mols m^-1^ s^-1^).
 \

### 2015-12-30 manuscript/tablesfigures.Rmd
finished the majority of comments from RD on chapter 4

@@
 Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
-Grey areas are 95% confidence intervals from the mean.
+Grey areas are 95% confidence intervals for the mean.
 
@@
 **Figure 6**.
-Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
+Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis ($\Delta_o$) and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
 The solid line represents the theoretical line for C3 plants from Evans et al.
@@
 **Figure 7**.
-The mean &pm; 1 standard error of intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light.
+The mean &pm; 1 standard error of (a) intercellular CO~2~ concentration (C~i~), (b) CO~2~ concentration in the chloroplasts (C~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (C~i~-C~c~).
 
 \
-**Figure S1**. Daily maximum and minimum temperature (a) and total daily PPFD (b) for each chamber across the experiment duration.
+**Figure S1**. Daily maximum and minimum temperature (a), daily maximum VPD (b) and total daily PPFD (c) for each chamber across the experiment duration.
 
@@
 Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
-C~c~ values were predicted with g~m~, thus curves represent chloroplastic photosynthetic parameters at 25°C and saturating light (1800 $\mu$mols m^-1^ s^-1^).
+C~c~ values were predicted with g~m~ and curves represent chloroplastic photosynthetic parameters at 25°C and saturating light (1800 $\mu$mols m^-1^ s^-1^).
 
@@
 **Table 2**.
-Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
+*Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
 Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).

### 2015-12-30 master_scripts/aci_vcmaxnitro.R

@@ -137,7 +137,7 @@ plot(Vcmax~leafN_area.mean, data=N_aci, col=as.factor(leaf),  ylim=c(40, 133),xl
 ####panel bottom right 
 par(fig=c(.5, 1, 0, .5),new=T,mar=c(4,4,1,1), cex=1.25, las=1, cex.axis=.8, cex.lab=1, mgp=c(2.5,1,0))
 plot(Photo ~ leafN_area, data=Anitro[Anitro$leaflight=="sun-high",], col=suncol,  xlim=c(0,4), ylim=c(0,25),
-       pch=c(16, 17)[pch=Anitro$temp], ylab=satlab, xlab=narealab)
+       pch=c(16, 17)[pch=Anitro$temp], ylab=photolab, xlab=narealab)
   points(Photo ~ leafN_area, data=Anitro[Anitro$leaflight=="shade-low",], col=shacol,  pch=c(16, 17)[pch=Anitro$temp])
   ablineclip(photoN_mod, x1=min(Anitro$leafN_area), x2=max(Anitro$leafN_area), lwd=2, lty=3)
   #legend("topleft", leglab2, pch=c(16,17,16,17), col=trtcols,inset = 0.01, bty='n',cex=.8)  

### 2016-01-02 manuscript/manuscript_wtc3.Rmd
thesis edits from RD

@@
 The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression, where k=0.5 [see @duursma2013near].
-Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use [@lin2015optimal].
+Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use, whereas large values imply a lower C construction cost and decreased water-use efficiency [@lin2015optimal].
 
@@
 Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R [@RDevelopmentCoreTeam2011] with WTC as a random effect.
-Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general].
+Explained variance (R^2^) of mixed models were computed as in [@nakagawa2013general], in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
 Confidence intervals (95 %) of mixed effect linear models were generated using bootstrapping methods with 999 simulations, using the bootMer function in the 'lme4' package [@bates2013lme4].

### 2016-01-03 manuscript/manuscript_wtc3.Rmd
grammar edit
A abbr

@@
 How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for upscaling *A~n~* for sun and shade leaves.
-Additionally, light saturated rates of A are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
+Additionally, light saturated rates of *A~n~* are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
 How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients.
@@
 The CO~2~ in the leaf cuvette was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mol s^-1^.
-Two identical gas exchanges systems were run simultaneously, one in each of a AT or ET WTC.
+Two identical gas exchanges systems were run simultaneously, one in each of an AT or ET WTC.
 Leaf temperatures were controlled at the current AT or ET WTC air temperature.

### 2016-01-05 manuscript/manuscript_wtc3.Rmd
table numbering
moved table 2
added gm equation, now in thesis as well

@@
 \begin{aligned}
-\Delta_i = &\frac{1}{1-t}a_i+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}
+\Delta_i = &\frac{1}{1-t}a'+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}
 \end{aligned}
@@
 \begin{aligned}
-\Delta_{gm} = &\frac{1+t}{1-t}\left(b-a'-\frac{eR_d}{A+R_d}\right)\frac{A}{g_mC_a}
+\Delta_{gm} = &\frac{1+t}{1-t}\left(b-a_i-\frac{eR_d}{A+R_d}\right)\frac{A}{g_mC_a}
 \end{aligned}
@@
 \
+The g~m~ can then be calculated as:
+
+\
+(11)
+$$
+\begin{aligned}
+g_m = \left(\frac{1+t}{1-t}\left(b-a_i-\frac{eR_d}{A+R_d}\right)\frac{A}{C_a}\right) / \left(\Delta_i-\Delta_o-\Delta_e-\Delta_f\right)
+\end{aligned}
+$$
+
+\
 The CO~2~ diffusion from the intercellular airspace to the chloroplast, g~m~, is given by its relationship to the leaf photosynthesis rate (*A~n~*) by:
@@
 \
-(11)
+(12)
 $$g_m = \frac{A_n}{C_i-C_c}$$
@@
 where C~c~ is the chloroplast CO~2~ partial pressure.
-Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were then estimated using Equation 11.
+Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were then estimated using Equation 12.
 Examples of this approach to measure gas exchange and C isotope discrimination are presented in Evans and Von Caemmerer [-@evans2013temperature].
@@
 \
-(12)
+(13)
 $$K = \frac{E}{\Psi_{pd}~-\Psi_l}$$
@@
 This was because $\Psi_{pd}$, $\Psi_l$ (Table `r wtc3tab1`) and E (Table `r wtc3tab2`) did not differ between leaf types.
-Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table1).
+Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table `r wtc3tab1`).
 Leaf mass per area was not different between leaf types.
@@
 
+\
+**Table 3.2**.
+*Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
+Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).
+Units for *A~n~* and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa.
+Different letters represent significant differences between leaf type, light environment and temperature treatments.
+The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.
+
+```{r, echo=FALSE}
+getab2 <- rbind(getab, c("P value","","", as.character(pval2)))
+greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
+
+dust(getab2)%>%
+  sprinkle_colnames("Leaf", "Light","Temperature", "*A~n~*", "g~s~", "g~m~", "ITE","E","VPD","C~i~", "C~c~") %>%
+  sprinkle(rows= 1, part="head",bold=TRUE)%>%
+  sprinkle_print_method("markdown")
+```
+
 ##3.4.2 Photosynthetic capacity and leaf photosynthesis rates 
@@
 Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
-
-\
-**Table 3.2**.
-*Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
-Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).
-Units for *A~n~* and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa.
-Different letters represent significant differences between leaf type, light environment and temperature treatments.
-The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.
-
-```{r, echo=FALSE}
-getab2 <- rbind(getab, c("P value","","", as.character(pval2)))
-greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
-
-dust(getab2)%>%
-  sprinkle_colnames("Leaf", "Light","Temperature", "*A~n~*", "g~s~", "g~m~", "ITE","E","VPD","C~i~", "C~c~") %>%
-  sprinkle(rows= 1, part="head",bold=TRUE)%>%
-  sprinkle_print_method("markdown")
-```
 

### 2016-01-05 gm_newphyt/manuscript_gmes.Rmd
started new folder with gmes manuscript for new phytologist.  moved discrimination gm equations to supp.  moved delta fig to supp.

--- 2016-01-05 manuscript/manuscript_wtc3.Rmd
@@
   word_document:
-    reference_docx: manuscriptstyle_thesis.docx
+    reference_docx: manuscriptstyle.docx
   pdf_document: default
@@
     number_sections: yes
-csl: tree-physiology.csl
+csl: new-phytologist.csl
 bibliography: references.bib
@@
 
-parfig <- "3.1"
-acifig <- "3.2"
-agmgsfig <- "3.3"
-gebarfig <- "3.4"
-waterusefig <- "3.5"
-deltafig <- "3.6"
-ciccfig <- "3.7"
+parfig <- "1"
+acifig <- "2"
+agmgsfig <- "3"
+gebarfig <- "4"
+waterusefig <- "5"
+ciccfig <- "6"
 
-airvarfig <- "3.S1"
-accfig <- "3.S2"
-get1fig <- "3.S3"
-get2fig <- "3.S4"
+airvarfig <- "S1"
+deltafig <- "S2"
+accfig <- "S3"
+get1fig <- "S3"
+get2fig <- "S4"
 
-wtc3tab1 <- "3.1"
-wtc3tab2 <- "3.2"
+wtc3tab1 <- "1"
+wtc3tab2 <- "2"
 
@@
 
-#3.1  Abstract
+#Abstract
 Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves.
@@
 
-<!-- ## Key words -->
-<!-- photosynthesis, stomatal conductance, mesophyll conductance,  shade, leaf optimal behavior -->
+##Key words
+photosynthesis, stomatal conductance, mesophyll conductance,  shade, leaf optimal behavior
 
-#3.2  Introduction
+#Introduction
 Light availability is one of the most important environmental drivers of leaf carbon (C) uptake in trees.
@@
 
-#3.3  Materials and Methods
+#Materials and Methods
 
-##3.3.1 Whole tree chamber experimental design
+##Whole tree chamber experimental design
 Twelve *Eucalyptus tereticornis* seedlings, chosen from a single local Cumberland plain cohort, were planted in March 2013 into 12 whole-tree chambers (WTC) at the Hawkesbury Forest Experiment site near Richmond, NSW, Australia.
@@
 
-##3.3.2 Leaf gas exchange, online carbon isotope discrimination and mesophyll conductance
+##Leaf gas exchange, online carbon isotope discrimination and mesophyll conductance
 Leaf gas exchange measurements were performed six times, beginning in October 2013 and monthly from December 2013 to April 2014.
@@
 
-\
+###condensed methods (cite in supplem)
+Synchronized gas exchange and C isotope discrimination measurements were made similarly as described in Tazoe et al.
+[-@tazoe2011using] and Evans & von Caemmerer [-@evans2013temperature].
 Leaf level gas exchange was measured with a standard (2 x 3 cm) leaf cuvette using a portable gas exchange system (LI-6400XT, LI-COR, Lincoln, NE, USA).
@@
 The CO~2~ in the leaf cuvette was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mol s^-1^.
-Two identical gas exchanges systems were run simultaneously, one in each of an AT or ET WTC.
+Two identical gas exchanges systems were run simultaneously, one in each of a randomly chosen WTC for each temperature treatment.
 Leaf temperatures were controlled at the current AT or ET WTC air temperature.
@@
 The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDL on each measurement cycle.
-Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the same 12 min interval.
+Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the 12 min interval.
 
+\
+Mesophyll conductance was calculated from carbon isotope discrimination with equations and fractionation factors as presented in Evans & Von Caemmerer [-@evans2013temperature], including the ternary corrections proposed by Farquhar & Cernusak [-@farquhar2012ternary], such that:
+
+\
+(1)
+$$\Delta_o = \Delta_i - \Delta_{gm} - \Delta_e - \Delta_f$$
+
+where $\Delta$~o~ is the observed discrimination and $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, g~m~, respiration and photorespiration, respectively.
+For this study, the CO~2~ compensation point ($\Gamma_{*}$) and respiration during the day (R~d~) parameters originally derived for tobacco plants [@von1994kinetics] were replaced with parameters derived for *Eucalyptus globulus* from Crous et al.
+[-@crous2012light] when calculating g~m~.
+Full descriptions of the carbon isotope discrimination equations, with ternary corrections, are available in the Supporting Information.
+The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck where compared as a function of C~i~:C~a~.
+Once g~m~ was caluclated the chloroplast CO~2~ partial pressure (C~c~) and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were estimated from the relationship between g~m~ to the leaf photosynthesis rate (*A~n~*) by:
+
+\
+(2)
+$$g_m = \frac{*A~n~*}{C_i-C_c}$$
+
+##Biochemical parameters of photosynthesis
+Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each WTC prior to the initiation of the drought treatment.
+Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011] using default parameters.
+
+##Leaf chemistry and hydraulic parameters
+Following gas exchange measurements each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, dried and weighed.
+These leaves were then milled and analyzed for leaf N content and $\delta$^13^C.
+Leaf samples were analysed on a Delta V Advantage coupled to a Flash HT and Conflo IV (Thermo Fisher Scientific, Bremen, Germany) in dual-reactor setup.
+Samples were flash combusted at 1000&deg;C and converted to CO~2~ and N~2~ and then subjected to stable isotope ratio mass spectrometry.
+Leaf N is reported on an area basis (N~a~, g m^-2^) and isotopic signatures are reported relative to the VPDP scale.
+
+\
+Predawn ($\Psi_{pd}$) and midday ($\Psi_l$) leaf water potentials (MPa) were measured for sun and shade leaves during each gas exchange campaign using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA).
+The leaf closest to the leaf used for gas exchange was sampled for measurement of $\Psi_{pd}$.
+Predawn leaf water potential was measured before sunrise on the same day as gas exchange measurements.
+Leaves used for gas exchange were immediately sampled for $\Psi_l$ once measurements were completed.
+All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
+Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) with the equation:
+
+\
+(3)
+$$K = \frac{E}{\Psi_{pd}~-\Psi_l}$$
+
+Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A~n~* divided by E.
+The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression, where k=0.5 [see @duursma2013near].
+Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use, whereas large values imply a lower C construction cost and decreased water-use efficiency [@lin2015optimal].
+
+##Data analysis
+Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R [@RDevelopmentCoreTeam2011] with WTC as a random effect.
+Explained variance (R^2^) of mixed models were computed as in [@nakagawa2013general], in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
+Confidence intervals (95 %) of mixed effect linear models were generated using bootstrapping methods with 999 simulations, using the bootMer function in the 'lme4' package [@bates2013lme4].
+For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package, using WTC as a random effect.
+All tests of statistical significance were conducted at an $\alpha$ of 0.05.
+
+#Results
+
+##Leaf resource distribution
+Across six measurement campaigns over the 7 month period PPFD was  reduced on average by >75% in the shade (Figure `r parfig`).
+Leaf-specific hydraulic conductance (K~l~) was similar across sun and shade leaves (Table `r wtc3tab1`).
+This was because $\Psi_{pd}$, $\Psi_l$ (Table `r wtc3tab1`) and E (Table `r wtc3tab2`) did not differ between leaf types.
+Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table `r wtc3tab1`).
+Leaf mass per area was not different between leaf types.
+No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~area~ or LMA either within or across leaf types (P > 0.05).
+
+\
+##Photosynthetic capacity and leaf photosynthesis rates 
+The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from AC~i~ curves (Figure `r acifig`a).
+Within leaf types, no effect of the warming treatment was detected on either parameter.
+Among the sampled leaves, Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
+
+\
+Mean *A~n~* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table `r wtc3tab2`).
+Additionally, leaf N~a~ was positively related to *A~n~* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
+Following an increase in light intensity to match high-light conditions, *A~n~* of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table `r wtc3tab2`).
+No effect of the warming treatment was detected on *A~n~* of sun leaves measured at high light or shade leaves at either low or high light.
+Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
+
+##Stomatal conductance and leaf water-use efficiency
+On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table `r wtc3tab2`).
+Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and *A~n~* were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
+Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
+No effect of the warming treatment was detected on g~s~ within or across leaf types.
+Stomatal conductance within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get2fig`b).
+
+\
+Measured under ambient light and temperature, leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001).
+Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was therefore significantly lower than sun leaves (P < 0.001).
+Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment, but no warming effect was detected in shade leaves measured at low or high light (Table `r wtc3tab2`).
+The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
+For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal ITE model with a g~1~ value for each leaf type and treatment (Figure `r waterusefig`a).
+Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).
+
+\
+Bulk-leaf $\delta$^13^C, as an index of integrated water-use efficiency [@marshall2007sources], was significantly lower in shade leaves compared to sun leaves by ca.
+2&permil; (Table `r wtc3tab1`).
+No effects of the warming treatment on leaf $\delta$^13^C were detected.
+Leaf $\delta$^13^C and N~a~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C (higher water-use efficiency) and higher N investment in sun leaves.
+
+##Leaf carbon isotope discrimination and mesophyll conductance
+The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
+Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
+The remainder $\Delta_o$ consists of the contributions of g~s~, respiration and photorespiration to discrimination.
+
+\
+Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
+Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table `r wtc3tab2`).
+Proportional increases in g~m~ were matched by proportional increases in *A~n~* from low to high light in shade leaves (Figure `r gebarfig`b,c).
+Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186).
+The large increases in g~m~ in shade leaves under high light likely resulted in the highest rates of *A~n~* (Figure `r agmgsfig`b).
+No differences in g~m~ were detected with the warming treatment within leaf types.
+Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively).
+We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~.
+Comparison of AC~c~ curves (Figure `r accfig`) and AC~i~ curves revealed similar differences between sun and shade leaves.
+
+##Variation in intercellular and chloroplastic CO~2~ concentrations
+Higher rates of g~s~ in shade leaves under low and high light led to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
+The chloroplast CO~2~ partial pressure was comparable between shade leaves when measured at both low and high light conditions (Figure `r ciccfig`c).
+In sun leaves C~c~ was significantly lower than shade leaves, consistent with a lower C~i~.
+The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and *A~n~* [@caemmerer2014temperature].
+This drawdown was similar between sun and shade leaves measured at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
+This was the result of the proportional relationship between g~m~ and *A~n~* across all leaves.
+The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
+
+#Discussion
+Here we show that *A~n~* in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
+Although shade leaves in lower light environments exhibited relatively high g~s~ and C~i~, it was rapid increases in g~m~ under periods of high light availability that allowed for this up-regulation of *A~n~*.
+Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
+These findings offer new insights into how aspects of leaf physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake in future studies.
+Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology.
+
+##Resource distribution and photosynthetic capacity
+The allocation of N~a~ constrains *A~n~* and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
+Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979).
+As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical].
+Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and *A~n~* were all reduced in shade leaves.
+Leaf mass per area, however, was not different between sun and shade leaves.
+This could be due to leaf formation under comparative light conditions or possible unmeasured differences in total non-structural carbohydrates contents between leaf types.
+
+\
+Photosynthesis is also limited by the ability to supply water to the upper canopy.
+Ultimately, the ability of a trees hydraulic architecture to supply water to foliage across increasing pathlengths affects productivity and survival [@sellin2008effects].
+Using a two-leaf model, Peltoniemi et al.
+[-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if K~l~ is also optimally distributed.
+In this study, variation in leaf N distribution and *A~n~* rates were not associated with subsequent changes in K~l~ between sun and shade leaves.
+Thus, no direct relationship between water supply and N distribution or *A~n~* within the canopy were detected.
+
+\
+Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased ITE throughout the experiment.
+Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was likely prevalent long term.
+From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A~n~* in sun leaves was characterized by low rates of g~s~ and low C~i~.
+Relative to the differences in *A~n~* between leaf types, higher rates of g~s~ in shade leaves appear to exhibit inefficient water use.
+As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower ITE compared to sun leaves.
+
+##Physiological behaviour of sun and shade leaves
+The pattern of inefficient water use in shade leaves is important as we hypothesized that g~s~ and *A~n~* would be proportional across sun and shade leaves.
+In sun leaves, *A~n~* and g~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
+However, lower rates of *A~n~* in shade leaves were not coupled with decreases in g~s~, leading to the observed decreases in ITE.
+This is significant as optimal stomatal regulation to balance C gain with water loss has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on sun leaves [e.g.
+@prentice2014balancing; @lin2015optimal].
+As a result, the often used economic framework of balancing costs of using water versus N allocation to predict *A~n~* [@wright2003least] may break down when considering all leaves within a tree canopy.
+
+\
+It is possible that reducing stomatal response time, by sustaining higher g~s~, is a strategy to take advantage of high light quickly in shade leaves [@tausz2005dynamic].
+Evidence from this study supports this hypothesis, as shade leaves increased *A~n~* equivalent or even outperforming sun leaves under identical light intensity.
+Transpiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck induced thermal load [@schymanski2013stomatal].
+This is because rapid increases in leaf temperature with sunflecks have been shown to inhibit C gain [@leakey2003high].
+However, this response likely occurs at very high air temperatures and may not explain the observed g~s~ in shade leaves across the large natural range of temperatures included in this study.
+How prevalent each of these strategies are within tree canopies is still unknown, as empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants, often in deep shade [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution].
+Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy [see @de1997simple].
+
+\
+We found that *A~n~* and g~m~ scaled positively across leaf types and, surprisingly, increased rapidly (within minutes) and proportionately when light intensity was increased in shade leaves.
+Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
+This provides a potential explanation for the observed rapid increases in g~m~ but the impacts of aquaporins on g~m~ are yet to be tested in leaves of tree species.
+Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
+Here we provide empirical data showing g~m~ not only varies within a canopy but the up-regulation of g~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.
+
+\
+If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution.
+This is because the role of g~s~ in regulating photosynthetic induction impacts the capacity of a leaf to utilize sunflecks [@way2012sunflecks].
+If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan.
+For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to tolerate decreases in ITE.
+However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.
+Thus, models which predict leaf photosynthesis from N distribution within a canopy will be incomplete unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].
+
+##Conclusions
+Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy.
+Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
+These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have important implications for process-based models that predict canopy C gain from rates of leaf photosynthesis.
+Furthermore, the dynamic nature of g~m~ cannot be simply parameterized in tree growth models and possibly should be excluded until it can be represented properly.
+Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown.
+To improve our ability to predict whole canopy C gain future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.
+
+#Tables
+
+**Table 3.1**.
+*Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments.
+Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
+Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
+Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;.
+Different letters represent significant differences between leaf type and temperature treatments.
+The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.
+
+```{r, echo=FALSE}
+resourcetab2 <- rbind(resourcetab, c("P value","", as.character(pval)))
+greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
+
+dust(resourcetab2)%>%
+  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0("$\\delta$","^13^c")) %>%
+  sprinkle(rows= 1, part="head",bold=TRUE)%>%
+  sprinkle_print_method("markdown")
+```
+
+\
+**Table 3.2**.
+*Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
+Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).
+Units for *A~n~* and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa.
+Different letters represent significant differences between leaf type, light environment and temperature treatments.
+The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.
+
+```{r, echo=FALSE}
+getab2 <- rbind(getab, c("P value","","", as.character(pval2)))
+greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
+
+dust(getab2)%>%
+  sprinkle_colnames("Leaf", "Light","Temperature", "*A~n~*", "g~s~", "g~m~", "ITE","E","VPD","C~i~", "C~c~") %>%
+  sprinkle(rows= 1, part="head",bold=TRUE)%>%
+  sprinkle_print_method("markdown")
+```
+
+#Figures
+
+```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+source("master_scripts/par2.R")
+```  
+\
+**Figure 3.1**.
+Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014.
+Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf.
+Each date represents the starting date for each measurement campaign.
+Points represent the mean (&pm; 1 standard error) daily maximum air temperature during each campaign period.
+
+\
+```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+source("master_scripts/aci_vcmaxnitro.R")
+```  
+\
+**Figure 3.2**.
+(a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.
+ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+(b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light.
+(c) The relationship between *A~n~* and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
+For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
+\
+```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+source("master_scripts/Agmgs.R")
+```  
+\
+**Figure 3.3**.
+The response of *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
+Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
+Grey areas are 95% confidence intervals for the mean.
+\
+```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+source("master_scripts/physiology_barplots.R")
+```  
+\
+**Figure 3.4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
+\
+```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=600}
+source("master_scripts/wateruse.R")
+```  
+\
+**Figure 3.5**.
+(a) Response of instantaneous transpiration efficiency (ITE) to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
+(b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light.
+For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment.
+For (b) the dashed line represents the significant linear model fit across all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
+\
+```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+source("master_scripts/cicc_bar.R")
+```  
+\
+**Figure 3.6**.
+The mean &pm; 1 standard error of (a) intercellular CO~2~ concentration (C~i~), (b) CO~2~ concentration in the chloroplasts (C~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (C~i~-C~c~).
+
+#Supporting Information
+
+##Figures
+
+```{r airvars, fig.width=7, fig.height=8.8, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+source("master_scripts/air_vars.R")
+```
+\
+**Figure 3.S1**. Daily maximum and minimum temperature (a), daily maximum VPD (b) and total daily PPFD (c) for each chamber across the experiment duration.
+\
+```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+source("master_scripts/delta_cica.R")
+```  
+\
+**Figure 3.S2**.
+Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis ($\Delta_o$) and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
+The solid line represents the theoretical line for C3 plants from Evans et al.
+[-@evans1986carbon].
+\
+```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+source("master_scripts/ACc_figure.R")
+```  
+\
+**Figure 3.S3**.
+Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
+C~c~ values were predicted with g~m~ and curves represent chloroplastic photosynthetic parameters at 25°C and saturating light (1800 $\mu$mols m^-1^ s^-1^).
+\
+```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+source("master_scripts/gasex_temp3.R")
+```  
+\
+**Figure 3.S4**.
+Response of *A~n~* (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
+Shaded symbols represents each monthly measurement campaign.
+Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
+All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
+Weak negative relationships with g~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (R^2^ = 0.16 and 0.08, respectively).
+\
+```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+source("master_scripts/gasex_temp4.R")
+```  
+\
+**Figure 3.S5**.
+Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
+Shaded symbols represents each monthly measurement campaign.
+Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
+All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
+Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (R^2^ = 0.73, 0.58 and 0.72, respectively).
+
+##Methods
+
+###Description of the calculation of g~m~ from carbon isotope discrimination during C3 photosynthesis
 \
@@
 
-\
-The CO~2~ diffusion from the intercellular airspace to the chloroplast, g~m~, is given by its relationship to the leaf photosynthesis rate (*A~n~*) by:
-
-\
-(12)
-$$g_m = \frac{A_n}{C_i-C_c}$$
-
-where C~c~ is the chloroplast CO~2~ partial pressure.
-Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were then estimated using Equation 12.
-Examples of this approach to measure gas exchange and C isotope discrimination are presented in Evans and Von Caemmerer [-@evans2013temperature].
-The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck where then compared as a function of C~i~:C~a~.
-
-##3.3.3 Biochemical parameters of photosynthesis
-Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each WTC prior to the initiation of the drought treatment.
-Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011] using default parameters.
-
-##3.3.4 Leaf chemistry and hydraulic parameters
-Following gas exchange measurements each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, dried and weighed.
-These leaves were then milled and analyzed for leaf N content and $\delta$^13^C.
-Leaf samples were analysed on a Delta V Advantage coupled to a Flash HT and Conflo IV (Thermo Fisher Scientific, Bremen, Germany) in dual-reactor setup.
-Samples were flash combusted at 1000&deg;C and converted to CO~2~ and N~2~ and then subjected to stable isotope ratio mass spectrometry.
-Leaf N is reported on an area basis (N~a~, g m^-2^) and isotopic signatures are reported relative to the VPDP scale.
-
-\
-Predawn ($\Psi_{pd}$) and midday ($\Psi_l$) leaf water potentials (MPa) were measured for sun and shade leaves during each gas exchange campaign using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA).
-The leaf closest to the leaf used for gas exchange was sampled for measurement of $\Psi_{pd}$.
-Predawn leaf water potential was measured before sunrise on the same day as gas exchange measurements.
-Leaves used for gas exchange were immediately sampled for $\Psi_l$ once measurements were completed.
-All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
-Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) with the equation:
-
-\
-(13)
-$$K = \frac{E}{\Psi_{pd}~-\Psi_l}$$
-
-Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A~n~* divided by E.
-The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression, where k=0.5 [see @duursma2013near].
-Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use, whereas large values imply a lower C construction cost and decreased water-use efficiency [@lin2015optimal].
-
-##3.3.5 Data analysis
-Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R [@RDevelopmentCoreTeam2011] with WTC as a random effect.
-Explained variance (R^2^) of mixed models were computed as in [@nakagawa2013general], in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
-Confidence intervals (95 %) of mixed effect linear models were generated using bootstrapping methods with 999 simulations, using the bootMer function in the 'lme4' package [@bates2013lme4].
-For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package, using WTC as a random effect.
-All tests of statistical significance were conducted at an $\alpha$ of 0.05.
-
-#3.4  Results
-
-##3.4.1 Leaf resource distribution
-Across six measurement campaigns over the 7 month period PPFD was  reduced on average by >75% in the shade (Figure `r parfig`).
-Leaf-specific hydraulic conductance (K~l~) was similar across sun and shade leaves (Table `r wtc3tab1`).
-This was because $\Psi_{pd}$, $\Psi_l$ (Table `r wtc3tab1`) and E (Table `r wtc3tab2`) did not differ between leaf types.
-Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table `r wtc3tab1`).
-Leaf mass per area was not different between leaf types.
-No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~area~ or LMA either within or across leaf types (P > 0.05).
-\
-```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/par2.R")
-```  
-\
-**Figure 3.1**.
-Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014.
-Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf.
-Each date represents the starting date for each measurement campaign.
-Points represent the mean (&pm; 1 standard error) daily maximum air temperature during each campaign period.
-
-\
-**Table 3.1**.
-*Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments.
-Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
-Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
-Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;.
-Different letters represent significant differences between leaf type and temperature treatments.
-The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.
-
-```{r, echo=FALSE}
-resourcetab2 <- rbind(resourcetab, c("P value","", as.character(pval)))
-greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
-
-dust(resourcetab2)%>%
-  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0("$\\delta$","^13^c")) %>%
-  sprinkle(rows= 1, part="head",bold=TRUE)%>%
-  sprinkle_print_method("markdown")
-```
-
-\
-**Table 3.2**.
-*Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
-Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).
-Units for *A~n~* and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa.
-Different letters represent significant differences between leaf type, light environment and temperature treatments.
-The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.
-
-```{r, echo=FALSE}
-getab2 <- rbind(getab, c("P value","","", as.character(pval2)))
-greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
-
-dust(getab2)%>%
-  sprinkle_colnames("Leaf", "Light","Temperature", "*A~n~*", "g~s~", "g~m~", "ITE","E","VPD","C~i~", "C~c~") %>%
-  sprinkle(rows= 1, part="head",bold=TRUE)%>%
-  sprinkle_print_method("markdown")
-```
-
-##3.4.2 Photosynthetic capacity and leaf photosynthesis rates 
-The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from AC~i~ curves (Figure `r acifig`a).
-Within leaf types, no effect of the warming treatment was detected on either parameter.
-Among the sampled leaves, Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
-\
-```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/aci_vcmaxnitro.R")
-```  
-\
-**Figure 3.2**.
-(a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.
-ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-(b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light.
-(c) The relationship between *A~n~* and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
-For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
-
-\
-Mean *A~n~* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table `r wtc3tab2`).
-Additionally, leaf N~a~ was positively related to *A~n~* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
-Following an increase in light intensity to match high-light conditions, *A~n~* of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table `r wtc3tab2`).
-No effect of the warming treatment was detected on *A~n~* of sun leaves measured at high light or shade leaves at either low or high light.
-Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
-
-##3.4.3 Stomatal conductance and leaf water-use efficiency
-On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table `r wtc3tab2`).
-Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and *A~n~* were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
-Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
-No effect of the warming treatment was detected on g~s~ within or across leaf types.
-Stomatal conductance within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get2fig`b).
-\
-```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/Agmgs.R")
-```  
-\
-**Figure 3.3**.
-The response of *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
-Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
-Grey areas are 95% confidence intervals for the mean.
-\
-```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/physiology_barplots.R")
-```  
-\
-**Figure 3.4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
-
-\
-Measured under ambient light and temperature, leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001).
-Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was therefore significantly lower than sun leaves (P < 0.001).
-Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment, but no warming effect was detected in shade leaves measured at low or high light (Table `r wtc3tab2`).
-The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
-For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal ITE model with a g~1~ value for each leaf type and treatment (Figure `r waterusefig`a).
-Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).
-\
-```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=600}
-source("master_scripts/wateruse.R")
-```  
-\
-**Figure 3.5**.
-(a) Response of instantaneous transpiration efficiency (ITE) to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
-(b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light.
-For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment.
-For (b) the dashed line represents the significant linear model fit across all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
-
-\
-Bulk-leaf $\delta$^13^C, as an index of integrated water-use efficiency [@marshall2007sources], was significantly lower in shade leaves compared to sun leaves by ca.
-2&permil; (Table `r wtc3tab1`).
-No effects of the warming treatment on leaf $\delta$^13^C were detected.
-Leaf $\delta$^13^C and N~a~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C (higher water-use efficiency) and higher N investment in sun leaves.
-
-##3.4.4 Leaf carbon isotope discrimination and mesophyll conductance
-The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
-Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
-The remainder $\Delta_o$ consists of the contributions of g~s~, respiration and photorespiration to discrimination.
-
-\
-```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/delta_cica.R")
-```  
-\
-**Figure 3.6**.
-Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis ($\Delta_o$) and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
-The solid line represents the theoretical line for C3 plants from Evans et al.
-[-@evans1986carbon].
-
-\
-Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
-Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table `r wtc3tab2`).
-Proportional increases in g~m~ were matched by proportional increases in *A~n~* from low to high light in shade leaves (Figure `r gebarfig`b,c).
-Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186).
-The large increases in g~m~ in shade leaves under high light likely resulted in the highest rates of *A~n~* (Figure `r agmgsfig`b).
-No differences in g~m~ were detected with the warming treatment within leaf types.
-Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively).
-We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~.
-Comparison of AC~c~ curves (Figure `r accfig`) and AC~i~ curves revealed similar differences between sun and shade leaves.
-
-##3.4.5 Variation in intercellular and chloroplastic CO~2~ concentrations
-Higher rates of g~s~ in shade leaves under low and high light led to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
-The chloroplast CO~2~ partial pressure was comparable between shade leaves when measured at both low and high light conditions (Figure `r ciccfig`c).
-In sun leaves C~c~ was significantly lower than shade leaves, consistent with a lower C~i~.
-The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and *A~n~* [@caemmerer2014temperature].
-This drawdown was similar between sun and shade leaves measured at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
-This was the result of the proportional relationship between g~m~ and *A~n~* across all leaves.
-The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
-\
-```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/cicc_bar.R")
-```  
-\
-**Figure 3.7**.
-The mean &pm; 1 standard error of (a) intercellular CO~2~ concentration (C~i~), (b) CO~2~ concentration in the chloroplasts (C~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (C~i~-C~c~).
-
-#3.5  Discussion
-Here we show that *A~n~* in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
-Although shade leaves in lower light environments exhibited relatively high g~s~ and C~i~, it was rapid increases in g~m~ under periods of high light availability that allowed for this up-regulation of *A~n~*.
-Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
-These findings offer new insights into how aspects of leaf physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake in future studies.
-Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology.
-
-##3.5.1 Resource distribution and photosynthetic capacity
-The allocation of N~a~ constrains *A~n~* and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
-Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979).
-As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical].
-Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and *A~n~* were all reduced in shade leaves.
-Leaf mass per area, however, was not different between sun and shade leaves.
-This could be due to leaf formation under comparative light conditions or possible unmeasured differences in total non-structural carbohydrates contents between leaf types.
-
-\
-Photosynthesis is also limited by the ability to supply water to the upper canopy.
-Ultimately, the ability of a trees hydraulic architecture to supply water to foliage across increasing pathlengths affects productivity and survival [@sellin2008effects].
-Using a two-leaf model, Peltoniemi et al.
-[-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if K~l~ is also optimally distributed.
-In this study, variation in leaf N distribution and *A~n~* rates were not associated with subsequent changes in K~l~ between sun and shade leaves.
-Thus, no direct relationship between water supply and N distribution or *A~n~* within the canopy were detected.
-
-\
-Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased ITE throughout the experiment.
-Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was likely prevalent long term.
-From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A~n~* in sun leaves was characterized by low rates of g~s~ and low C~i~.
-Relative to the differences in *A~n~* between leaf types, higher rates of g~s~ in shade leaves appear to exhibit inefficient water use.
-As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower ITE compared to sun leaves.
-
-##3.5.2 Physiological behaviour of sun and shade leaves
-The pattern of inefficient water use in shade leaves is important as we hypothesized that g~s~ and *A~n~* would be proportional across sun and shade leaves.
-In sun leaves, *A~n~* and g~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
-However, lower rates of *A~n~* in shade leaves were not coupled with decreases in g~s~, leading to the observed decreases in ITE.
-This is significant as optimal stomatal regulation to balance C gain with water loss has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on sun leaves [e.g.
-@prentice2014balancing; @lin2015optimal].
-As a result, the often used economic framework of balancing costs of using water versus N allocation to predict *A~n~* [@wright2003least] may break down when considering all leaves within a tree canopy.
-
-\
-It is possible that reducing stomatal response time, by sustaining higher g~s~, is a strategy to take advantage of high light quickly in shade leaves [@tausz2005dynamic].
-Evidence from this study supports this hypothesis, as shade leaves increased *A~n~* equivalent or even outperforming sun leaves under identical light intensity.
-Transpiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck induced thermal load [@schymanski2013stomatal].
-This is because rapid increases in leaf temperature with sunflecks have been shown to inhibit C gain [@leakey2003high].
-However, this response likely occurs at very high air temperatures and may not explain the observed g~s~ in shade leaves across the large natural range of temperatures included in this study.
-How prevalent each of these strategies are within tree canopies is still unknown, as empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants, often in deep shade [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution].
-Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy [see @de1997simple].
-
-\
-We found that *A~n~* and g~m~ scaled positively across leaf types and, surprisingly, increased rapidly (within minutes) and proportionately when light intensity was increased in shade leaves.
-Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
-This provides a potential explanation for the observed rapid increases in g~m~ but the impacts of aquaporins on g~m~ are yet to be tested in leaves of tree species.
-Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
-Here we provide empirical data showing g~m~ not only varies within a canopy but the up-regulation of g~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.
-
-\
-If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution.
-This is because the role of g~s~ in regulating photosynthetic induction impacts the capacity of a leaf to utilize sunflecks [@way2012sunflecks].
-If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan.
-For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to tolerate decreases in ITE.
-However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.
-Thus, models which predict leaf photosynthesis from N distribution within a canopy will be incomplete unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].
-
-##3.5.3 Conclusions
-Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy.
-Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
-These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have important implications for process-based models that predict canopy C gain from rates of leaf photosynthesis.
-Furthermore, the dynamic nature of g~m~ cannot be simply parameterized in tree growth models and possibly should be excluded until it can be represented properly.
-Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown.
-To improve our ability to predict whole canopy C gain future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.
-
-#3.6  Supporting Information
-
-```{r airvars, fig.width=7, fig.height=8.8, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/air_vars.R")
-```
-\
-**Figure 3.S1**. Daily maximum and minimum temperature (a), daily maximum VPD (b) and total daily PPFD (c) for each chamber across the experiment duration.
-\
-```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/ACc_figure.R")
-```  
-\
-**Figure 3.S2**.
-Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
-C~c~ values were predicted with g~m~ and curves represent chloroplastic photosynthetic parameters at 25°C and saturating light (1800 $\mu$mols m^-1^ s^-1^).
-\
-```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/gasex_temp3.R")
-```  
-\
-**Figure 3.S3**.
-Response of *A~n~* (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
-Shaded symbols represents each monthly measurement campaign.
-Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
-All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
-Weak negative relationships with g~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (R^2^ = 0.16 and 0.08, respectively).
-\
-```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/gasex_temp4.R")
-```  
-\
-**Figure 3.S4**.
-Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
-Shaded symbols represents each monthly measurement campaign.
-Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
-All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
-Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (R^2^ = 0.73, 0.58 and 0.72, respectively).
-
 # References {.unnumbered}

### 2016-01-05 gm_newphyt/manuscript_gmes.Rmd
removed thesis chapter title

@@
 ---
-title: |
-  | Chapter 3 
-  | Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks
+title: Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks
 author: ""

### 2016-01-06 gm_newphyt/manuscript_gmes.Rmd
added word count script from RD to use in latex
finished a new phytologist draft of the gmes manuscript which was sent to SVC

@@
 \
-Courtney E. Campany^1^, Belinda Medlyn^1^, Mark Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^.
+Courtney E. Campany^1^, Mark Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^.
 
@@
 ^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia
-
 \
@@
 \
-*Corresponding author*:
-Courtney Campany
-E: courtneycampany@gmail.com
+Author for Correspondence:
+\
+*Courtney Campany*
+\
+*Tel: +61(0)432391114*
+\
+*Email: courtneycampany@gmail.com*
+
+\
+```{r wc, echo=FALSE, warning=FALSE}
+library(pixiedust)
+wc_tab <- data.frame(count = c("Total word count (main body):", "Summary:", "Introduction:", "Materials and Methods:", "Results:",                               "Discussion:", "Acknowledgements:"), 
+                     answer = c(5535, 219, 1207, 1847, 1102, 1359,20),
+                     count2 = c("No. of Figures:", "No of Tables:", "No. of Supporting Information files:"
+                                ,"","","",""),
+                     answer2 = c("6 (all in color)",2,"6 (5 Figures, 1 Methods)", "","","",""))
+#colnames(wc_tab) <- NULL 
+dust(wc_tab)%>%
+sprinkle_print_method("markdown")
+```
 
@@
 accfig <- "S3"
-get1fig <- "S3"
-get2fig <- "S4"
+get1fig <- "S4"
+get2fig <- "S5"
 
@@
 
-#Abstract
-Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves.
+# Summary
+* Light gradients within tree canopies play a major role in the distribution of plant resources that define leaf photosynthetic capacity.
 A lack of empirical data relating photosynthesis to leaf physiological behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
-To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
-Compared to sun leaves, shade leaves had lower *A~n~*, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
-When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in *A~n~* greater than sun leaves at the same photosynthetic photon flux density.
-Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
+* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
+* Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
+When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves.
+* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
 We argue that the rapid response of g~m~ with light enables shade leaves to respond quickly to sunflecks.
@@
 ##Key words
-photosynthesis, stomatal conductance, mesophyll conductance,  shade, leaf optimal behavior
+leaf optimal behavior, mesophyll conductance, photosynthesis, shade, stomatal conductance,
 
@@
 As *A~n~* has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
-Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll, and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic].
+Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic].
 Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
@@
 \
-In this study we use *Eucalyptus tereticornis* Sm.
-trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to empirically evaluate the distribution of nitrogen and water supply and leaf physiological behaviour of sun and shade leaves.
+In this study we use *Eucalyptus tereticornis* trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to empirically evaluate the distribution of nitrogen and water supply and leaf physiological behaviour of sun and shade leaves.
 Our hypotheses are as follows:
@@
 ##Whole tree chamber experimental design
-Twelve *Eucalyptus tereticornis* seedlings, chosen from a single local Cumberland plain cohort, were planted in March 2013 into 12 whole-tree chambers (WTC) at the Hawkesbury Forest Experiment site near Richmond, NSW, Australia.
+Twelve *Eucalyptus tereticornis* Sm.
+seedlings, chosen from a single local Cumberland plain cohort, were planted in March 2013 into 12 whole-tree chambers (WTC) at the Hawkesbury Forest Experiment site near Richmond, NSW, Australia.
 Each chamber has a height of 9 m and seedlings were grown for 15 months.
@@
 
-###condensed methods (cite in supplem)
+\
 Synchronized gas exchange and C isotope discrimination measurements were made similarly as described in Tazoe et al.
@@
 Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the 12 min interval.
-
 \
@@
 [-@crous2012light] when calculating g~m~.
-Full descriptions of the carbon isotope discrimination equations, with ternary corrections, are available in the Supporting Information.
+Full descriptions of the carbon isotope discrimination equations, with ternary corrections, are presented in Supporting Information Methods S1.
 The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck where compared as a function of C~i~:C~a~.
-Once g~m~ was caluclated the chloroplast CO~2~ partial pressure (C~c~) and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were estimated from the relationship between g~m~ to the leaf photosynthesis rate (*A~n~*) by:
+Once g~m~ was calculated the chloroplast CO~2~ partial pressure (C~c~) and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were estimated from the relationship between g~m~ to the leaf photosynthesis rate (*A~n~*) by:
 
@@
 (2)
-$$g_m = \frac{*A~n~*}{C_i-C_c}$$
+$$g_m = \frac{A_n}{C_i-C_c}$$
 
@@
 Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A~n~* divided by E.
-The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression, where k=0.5 [see @duursma2013near].
+The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near].
 Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use, whereas large values imply a lower C construction cost and decreased water-use efficiency [@lin2015optimal].
@@
 
-\
 ##Photosynthetic capacity and leaf photosynthesis rates 
@@
 
+### Acknowledgements
+We thank Craig Barton and Burhan Amiji for maintaining the Whole Tree Chambers experiment and for their outstanding technical assitance.
+
+### Author Contributions
+CC contributed to the design of the research, data analysis, collection, interpretation and writing the manuscript.
+MT contributed to the design of the research, interpretation  and writing the manuscript.
+SC contributed to the design of the research, performance of the research and interpretation.
+RD contributed to the design of the research, data analysis, interpretation and writing the manuscript.
+
 #Supporting Information
@@
 
-##Methods
+##Methods S1
 

### 2016-01-06 manuscript/manuscript_wtc3.Rmd
added word count script from RD to use in latex
finished a new phytologist draft of the gmes manuscript which was sent to SVC

@@
 
-\
-Courtney E. Campany^1^, Belinda Medlyn^1^, Mark Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^.
+<!-- \ -->
+<!-- Courtney E. Campany^1^, Belinda Medlyn^1^, Mark Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^. -->
 
-\
-^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia
+<!-- \ -->
+<!-- ^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia -->
 
-\
-^2^ Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, Australia
+<!-- \ -->
+<!-- ^2^ Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, Australia -->
 
-\
-*Corresponding author*:
-Courtney Campany
-E: courtneycampany@gmail.com
+<!-- \ -->
+<!-- *Corresponding author*: -->
+<!-- Courtney Campany -->
+<!-- E: courtneycampany@gmail.com -->
 

### 2016-01-08 gm_newphyt/manuscript_gmes.Rmd
MT thesis edits

@@
 \
-4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporation demand from higher temperatures may lead to stomatal closure.
+4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporation demand from higher temperatures lead to stomatal closure.
 
@@
 Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).
-Units for *A~n~* and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa.
+Units for *A~n~* and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD is kPa.
 Different letters represent significant differences between leaf type, light environment and temperature treatments.

### 2016-01-10 gm_newphyt/manuscript_gmes.Rmd
final thesis edits from MT.  Need to re-run manuscript rmd for SVC.

@@
 Each chamber has a height of 9 m and seedlings were grown for 15 months.
-A detailed description of the WTC operation and design is available in [@barton2010whole].
+A detailed description of the WTC operation and design is available in [-@barton2010whole].
 Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarfig`).
@@
 Photosynthesis is also limited by the ability to supply water to the upper canopy.
-Ultimately, the ability of a trees hydraulic architecture to supply water to foliage across increasing pathlengths affects productivity and survival [@sellin2008effects].
+Ultimately, the ability of tree hydraulic architecture to supply water to foliage across increasing pathlengths affects productivity and survival [@sellin2008effects].
 Using a two-leaf model, Peltoniemi et al.
@@
 \
-Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased ITE throughout the experiment.
+Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased ITE in shade leaves throughout the experiment.
 Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was likely prevalent long term.
@@
 Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
-This provides a potential explanation for the observed rapid increases in g~m~ but the impacts of aquaporins on g~m~ are yet to be tested in leaves of tree species.
+This provides a potential explanation for the observed rapid increases in g~m~, but the impacts of aquaporins on g~m~ are yet to be tested in leaves of tree species.
 Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
@@
 \
-If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution.
+If shade leaves "lie in wait" for sunflecks, then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution.
 This is because the role of g~s~ in regulating photosynthetic induction impacts the capacity of a leaf to utilize sunflecks [@way2012sunflecks].
-If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan.
+If the valuation of sunflecks as a C resource is large enough, then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan or across the entire canopy when considering both sun and shade leaf types.
 For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to tolerate decreases in ITE.
@@
 Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown.
-To improve our ability to predict whole canopy C gain future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.
+To improve our ability to predict whole canopy C gain, future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.
 
@@
 Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
-Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;.
+Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, WP is MPa and $\delta$^13^C is &permil;.
 Different letters represent significant differences between leaf type and temperature treatments.

### 2016-01-24 gm_newphyt/manuscript_gmes.Rmd
correct mooney 1979 reference
thesis edits

@@
 Each chamber has a height of 9 m and seedlings were grown for 15 months.
-A detailed description of the WTC operation and design is available in [-@barton2010whole].
+A detailed description of the WTC operation and design is available in Barton et al.
+[-@barton2010whole].
 Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarfig`).
@@
 The allocation of N~a~ constrains *A~n~* and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
-Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979).
+Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy [@mooney1979environmental].
 As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical].

### 2016-02-02 gm_newphyt/manuscript_gmes.Rmd
manuscripts edits from SVC, these are made in thesis too

@@
 \
-^2^ Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, Australia
+^2^ ARC Centre of Excellence for Translational Photosynthesis, Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, Australia
 
@@
 * To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
-* Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
+* Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but g~s~ was similar.
 When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves.
@@
 Part of the variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
-Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack of empirical data currently hinders are ability to interpreting their coupled responses to *A~n~* across sun and shade leaves.
+Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack of empirical data currently hinders our ability to interpreting their coupled responses to *A~n~* across sun and shade leaves.
 
@@
 Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the 12 min interval.
+
 \
-Mesophyll conductance was calculated from carbon isotope discrimination with equations and fractionation factors as presented in Evans & Von Caemmerer [-@evans2013temperature], including the ternary corrections proposed by Farquhar & Cernusak [-@farquhar2012ternary], such that:
+Mesophyll conductance was calculated from carbon isotope discrimination with equations and fractionation factors as presented in Evans & von Caemmerer [-@evans2013temperature], including the ternary corrections proposed by Farquhar & Cernusak [-@farquhar2012ternary], such that:
 
@@
 Samples were flash combusted at 1000&deg;C and converted to CO~2~ and N~2~ and then subjected to stable isotope ratio mass spectrometry.
-Leaf N is reported on an area basis (N~a~, g m^-2^) and isotopic signatures are reported relative to the VPDP scale.
+Leaf N is reported on an area basis (N~a~, g m^-2^) and isotopic signatures of dry matter are reported relative to standard Vienna Pee Dee Belemnite.
 
@@
 
-**Table 3.1**.
+**Table 1**.
 *Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments.
@@
 \
-**Table 3.2**.
+**Table 2**.
 *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
@@
 \
-**Figure 3.1**.
+**Figure 1**.
 Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014.
@@
 \
-**Figure 3.2**.
+**Figure 2**.
 (a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.
@@
 \
-**Figure 3.3**.
+**Figure 3**.
 The response of *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
@@
 \
-**Figure 3.4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
+**Figure 4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
 \
@@
 \
-**Figure 3.5**.
+**Figure 5**.
 (a) Response of instantaneous transpiration efficiency (ITE) to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
@@
 \
-**Figure 3.6**.
+**Figure 6**.
 The mean &pm; 1 standard error of (a) intercellular CO~2~ concentration (C~i~), (b) CO~2~ concentration in the chloroplasts (C~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (C~i~-C~c~).
@@
 \
-**Figure 3.S1**. Daily maximum and minimum temperature (a), daily maximum VPD (b) and total daily PPFD (c) for each chamber across the experiment duration.
+**Figure S1**. Daily maximum and minimum temperature (a), daily maximum VPD (b) and total daily PPFD (c) for each chamber across the experiment duration.
 \
@@
 \
-**Figure 3.S2**.
+**Figure S2**.
 Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis ($\Delta_o$) and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
@@
 \
-**Figure 3.S3**.
+**Figure S3**.
 Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
@@
 \
-**Figure 3.S4**.
+**Figure S4**.
 Response of *A~n~* (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
@@
 \
-**Figure 3.S5**.
+**Figure S5**.
 Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.

### 2016-02-02 manuscript/manuscript_wtc3.Rmd
correct mooney 1979 reference
manuscripts edits from SVC, these are made in thesis too

@@
 To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
-Compared to sun leaves, shade leaves had lower *A~n~*, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~.
+Compared to sun leaves, shade leaves had lower *A~n~*, g~m~, Vc~max~, J~max~ and leaf nitrogen but g~s~ was similar.
 When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in *A~n~* greater than sun leaves at the same photosynthetic photon flux density.
@@
 [-@evans1986carbon].
-In this experiment, a modified form of this equation presented in Evans & Von Caemmerer [-@evans2013temperature] with ternary effect corrections by Farquhar & Cernusak [-@farquhar2012ternary] was used such that:
+In this experiment, a modified form of this equation presented in Evans & von Caemmerer [-@evans2013temperature] with ternary effect corrections by Farquhar & Cernusak [-@farquhar2012ternary] was used such that:
 
@@
 Samples were flash combusted at 1000&deg;C and converted to CO~2~ and N~2~ and then subjected to stable isotope ratio mass spectrometry.
-Leaf N is reported on an area basis (N~a~, g m^-2^) and isotopic signatures are reported relative to the VPDP scale.
+Leaf N is reported on an area basis (N~a~, g m^-2^) and isotopic signatures of dry leaf matter are reported relative to VPDP.
 

### 2016-02-22 gm_newphyt/manuscript_gmes.Rmd
added hanba cite

@@
 For example, g~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid].
-Anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are also unlikely to adjust during short light fluctuations.
+Anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative; @hanba2002effect] are also unlikely to adjust during short light fluctuations.
 The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance and as a potential response to transitory light availability.

### 2016-02-23 gm_newphyt/manuscript_gmes.Rmd
updated new box plots and shade scatter plot
started new ACC script
updates to manuscrupt rmd

--- 2016-02-02 gm_newphyt/manuscript_gmes.Rmd
@@
 waterusefig <- "5"
-ciccfig <- "6"
+gmshadefig <- "6"
+ciccfig <- "7"
 
@@
 * To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
-* Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but g~s~ was similar.
-When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves.
+* Compared to sun leaves, shade leaves had lower A, g~m~, leaf nitrogen and photosynthetic biochemical parameters (Vc~max~ and J~max~), but g~s~ was similar.
+When light intensity was increased for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves.
 * Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
-We argue that the rapid response of g~m~ with light enables shade leaves to respond quickly to sunflecks.
-Evidence that g~m~ not only varies within a canopy but can be up-regulated over short time intervals possibly represents a new mechanism underpinning leaf gas exchange responses to light.
+We argue that the up-regulation of g~m~ over short time intervals with light enables shade leaves to respond quickly to sunflecks, possibly representing a new mechanism underpinning leaf gas exchange responses to light.
 
@@
 For example, g~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid].
-Anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are also unlikely to adjust during short light fluctuations.
+Anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative; terashima2001sun; @hanba2002effect] are also unlikely to adjust during short light fluctuations.
 The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance and as a potential response to transitory light availability.
@@
 \
-In this study we use *Eucalyptus tereticornis* trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to empirically evaluate the distribution of nitrogen and water supply and leaf physiological behaviour of sun and shade leaves.
-Our hypotheses are as follows:
-
-\
-1.	If whole tree canopies are optimized for C gain, then leaf N, hydraulic conductance and biochemical photosynthetic capacity are predicted to be higher in sun leaves compared to shade leaves.
-
-\
-2.	Stomatal conductance is proportional to *A~n~* across sun and shade leaves under similar leaf VPD and g~m~ scales positively with photosynthetic capacity.
-
-\
-3. As shade leaves were expected to develop lower biochemical photosynthetic capacity, increases in *A~n~* following sunfleck simulations were not expected to reach rates of full sun leaves.
-
-\
-4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporation demand from higher temperatures lead to stomatal closure.
+In this study we use *Eucalyptus tereticornis* trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to empirically test whether the distribution of nitrogen and water supply and leaf physiological behaviour result in  higher photosynthetic capacity in sun leaves compared to shade leaves.
+As leaves which developed in the shade were expected to have lower biochemical photosynthetic capacity, we then predicted that sunfleck simulations would not increase *A~n~* to rates similar to sun leaves.
+We further predicted that climate warming would decrease g~s~ and leaf C gain in sun leaves during summer months, as increased evaporation demand from higher temperatures lead to stomatal closure.
 
@@
 Leaf-specific hydraulic conductance (K~l~) was similar across sun and shade leaves (Table `r wtc3tab1`).
-This was because $\Psi_{pd}$, $\Psi_l$ (Table `r wtc3tab1`) and E (Table `r wtc3tab2`) did not differ between leaf types.
+This was because neither leaf water potentials ($\Psi_{pd}$ and $\Psi_l$, Table `r wtc3tab1`) nor transpiration rates (E, Table `r wtc3tab2`) differed between leaf types.
 Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table `r wtc3tab1`).
-Leaf mass per area was not different between leaf types.
-No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~area~ or LMA either within or across leaf types (P > 0.05).
+Leaf mass per area (LMA) was not different between leaf types (Table `r wtc3tab1`).
+No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~a~ or LMA either within or across leaf types (P > 0.05).
 
@@
 Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
+
 Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table `r wtc3tab2`).
 Proportional increases in g~m~ were matched by proportional increases in *A~n~* from low to high light in shade leaves (Figure `r gebarfig`b,c).
+
 Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186).
@@
 Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased ITE in shade leaves throughout the experiment.
-Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was likely prevalent long term.
+Additionally, consistently lower leaf $\delta$^13^C in shade leaves suggests that this pattern was likely prevalent long term.
 From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A~n~* in sun leaves was characterized by low rates of g~s~ and low C~i~.
@@
 **Figure 3**.
-The response of *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
+The relationship between *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
 Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
@@
 ```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/physiology_barplots.R")
+source("master_scripts/physiology_boxplots.R")
 ```  
 \
-**Figure 4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
+**Figure 4**. Box plots of measured g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
 \
@@
 **Figure 5**.
-(a) Response of instantaneous transpiration efficiency (ITE) to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
+(a) The relationship between instantaneous transpiration efficiency (ITE) and VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
 (b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light.
@@
 \
-```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/cicc_bar.R")
+```{r gmshade, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+source("master_scripts/gmes_shadeleaves_figure.R")
 ```  
@@
 **Figure 6**.
-The mean &pm; 1 standard error of (a) intercellular CO~2~ concentration (C~i~), (b) CO~2~ concentration in the chloroplasts (C~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (C~i~-C~c~).
+The change in measured g~m for individual shade leaves following an increase in PPFD to match the light environment of the full sun leaf in the same tree.
+Each measurement of g~m~ on the same leaf, for both low and high environments, was recorded once CO~2~ and water vapour fluxes were stable in the leaf cuvette.
+The dashed line is the 1:1 relationship.
+
+\
+```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+source("master_scripts/cicc_boxplots.R")
+```  
+\
+**Figure 7**.
+Boxplots of (a) intercellular CO~2~ concentration (C~i~), (b) CO~2~ concentration in the chloroplasts (C~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (C~i~-C~c~).
 
 ### Acknowledgements
-We thank Craig Barton and Burhan Amiji for maintaining the Whole Tree Chambers experiment and for their outstanding technical assitance.
+We thank Craig Barton and Burhan Amiji for maintaining the Whole Tree Chambers experiment and for their outstanding technical assistance.
 

### 2016-02-23 master_scripts/gmes_shadeleaves_figure.R

@@ -6,7 +6,7 @@ source("functions and packages/packages.R")
 gmes <- read.csv("calculated_data/gmes_wellwatered.csv")
 
 ###get average by id
-gm_agg <- summaryBy(Photo+ Cond + gm ~ chamber+id+leaf +light+temp+leaflight+Month, data=gmes, FUN=mean, keep.names=TRUE)
+gm_agg <- summaryBy(Photo+ Cond + gm_bar ~ chamber+id+leaf +light+temp+leaflight+Month, data=gmes, FUN=mean, keep.names=TRUE)
 gm_agg$leaf <- gsub("s", "S", gm_agg$leaf)
 gm_agg$light <- gsub("high", "Sun-light", gm_agg$light)
 gm_agg$light <- gsub("low", "Shade-light", gm_agg$light)
@@ -29,15 +29,13 @@ names(fleckdat2)[4] <- "fleck_gm"
 shadeleaf <- merge(shadat2, fleckdat2, by=c("chamber", "temp","Month"))
 
 ##plot
-shagmlab <- expression(g[m]~Low~Light~~(mol~m^-2~s^-1))
-fleckgmlab <- expression(g[m]~High~Light~~(mol~m^-2~s^-1))
+# windows(7,7)
 
-windows(7,7)
 par(mar=c(4,4,1,1), cex=1.25, las=1, cex.axis=.8, cex.lab=1, mgp=c(2.5,1,0))
 plot(fleck_gm~shade_gm, data=shadeleaf, ylim=c(0,.3), xlim=c(0,.3), xlab=shagmlab, ylab=fleckgmlab, 
      pch=c(16, 17)[shadeleaf$temp])
 abline(0,1, lty=2)
 legend("topleft", c("AT", "ET"), pch=c(16,17), inset = 0.01, bty='n', title="Shade Leaves", cex=.8)
 
-dev.copy2pdf(file="master_scripts/paper_figures/gmshade.pdf")
-dev.off()
+# dev.copy2pdf(file="master_scripts/paper_figures/gmshade.pdf")
+# dev.off()

### 2016-02-23 master_scripts/plot_objects.R

@@ -8,6 +8,9 @@ pchtri <- c(17,17,2,2)
 
 gmlab <-expression(g[m]~~(mol~m^-2~s^-1))
 gmlab2 <-expression(g[m]~~(mol~m^-2~s^-1~bar^-1))
+shagmlab <- expression(g[m]~Low~Light~~(mol~m^-2~s^-1~bar^-1))
+fleckgmlab <- expression(g[m]~High~Light~~(mol~m^-2~s^-1~bar^-1))
+
 lmalab <- expression(LMA~~(g~m^-2))
 masslab <- "Leaf Mass  (g)"
 

### 2016-02-24 gm_newphyt/manuscript_gmes.Rmd
updates to gmes manuscript from SVC

@@
 The distribution of resources required for *A~n~*, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients.
-As *A~n~* has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
+As *A~n~* has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co; @buckley2013does].
 Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic].
@@
 Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
-
-Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table `r wtc3tab2`).
+Following the increase in PPFD from low to high light on the same leaf, g~m~ values of shade leaves increased on average 55% after approximately 25 min (Figure `r gmshadefig`).
+These measured values of g~m~ for shade leaves at high light were also equivalent or greater than those of sun leaves (Table `r wtc3tab2`).
 Proportional increases in g~m~ were matched by proportional increases in *A~n~* from low to high light in shade leaves (Figure `r gebarfig`b,c).
-
 Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186).
@@
 \
-```{r gmshade, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+```{r gmshade, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/gmes_shadeleaves_figure.R")
@@
 **Figure 6**.
-The change in measured g~m for individual shade leaves following an increase in PPFD to match the light environment of the full sun leaf in the same tree.
-Each measurement of g~m~ on the same leaf, for both low and high environments, was recorded once CO~2~ and water vapour fluxes were stable in the leaf cuvette.
+The change in measured g~m~ for individual shade leaves following an increase in PPFD to match the light environment of full sun leaves in the same tree.
+Measurements of g~m~ were recorded once CO~2~ and water vapour fluxes were stable in the leaf cuvette, which took approximately 25 minutes after light intensity was increased.
 The dashed line is the 1:1 relationship.

### 2016-02-24 thesis2manuscript_notes.R

@@ -21,3 +21,4 @@
 ##introduction-------------------------------------------------------------------------------------------------------------
 #1. added hanba and terashima to Boardman cite, sunfleck paragraph
 #2. deleted hyptotheses and rewrote last paragraph
+#3. added Buckely cite to N to light sentence

### 2016-03-03 gm_newphyt/manuscript_gmes.Rmd
updating manuscrupt with SVC comments (mostly ACC and gm_bar)
rapid gm no longer written

@@
 * Compared to sun leaves, shade leaves had lower A, g~m~, leaf nitrogen and photosynthetic biochemical parameters (Vc~max~ and J~max~), but g~s~ was similar.
-When light intensity was increased for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves.
+When light intensity was increased for shade leaves both g~s~ and g~m~ increased, leading to increases in A greater than sun leaves.
 * Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
@@
 Here we show that *A~n~* in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
-Although shade leaves in lower light environments exhibited relatively high g~s~ and C~i~, it was rapid increases in g~m~ under periods of high light availability that allowed for this up-regulation of *A~n~*.
+Although shade leaves in lower light environments exhibited relatively high g~s~ and C~i~, it was concomitant increases in g~m~ under periods of high light availability that allowed for this up-regulation of *A~n~*.
 Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
@@
 Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased ITE in shade leaves throughout the experiment.
+
 Additionally, consistently lower leaf $\delta$^13^C in shade leaves suggests that this pattern was likely prevalent long term.
 From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A~n~* in sun leaves was characterized by low rates of g~s~ and low C~i~.
+
 Relative to the differences in *A~n~* between leaf types, higher rates of g~s~ in shade leaves appear to exhibit inefficient water use.
@@
 \
-We found that *A~n~* and g~m~ scaled positively across leaf types and, surprisingly, increased rapidly (within minutes) and proportionately when light intensity was increased in shade leaves.
+We found that *A~n~* and g~m~ scaled positively across leaf types and, unexpectedly, increased proportionately after leaf stability (~25 min) when light intensity was increased in shade leaves.
 Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
-This provides a potential explanation for the observed rapid increases in g~m~, but the impacts of aquaporins on g~m~ are yet to be tested in leaves of tree species.
+This provides a potential explanation for the observed increases in g~m~, but the impacts of aquaporins on g~m~ are yet to be tested in leaves of tree species.
 Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
@@
 Furthermore, the dynamic nature of g~m~ cannot be simply parameterized in tree growth models and possibly should be excluded until it can be represented properly.
-Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown.
+Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of g~m~ to opportunistically increase CO~2~ drawdown.
 To improve our ability to predict whole canopy C gain, future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.
@@
 Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).
-Units for *A~n~* and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD is kPa.
+Units for *A~n~* and E are $\mu$mol m^-2^ s^-1^, for g~s~ are mol m^-2^ s^-1^, for g~m~ are mol m^-2^ s^-1^ bar^-1^ and for VPD is kPa.
 Different letters represent significant differences between leaf type, light environment and temperature treatments.
@@
 
+Here, g~m~ is expressed in units of mol m^-2^ s^-1^ bar^-1^ to be consistent with Evans & Von Caemmerer [-@evans2013temperature].
+As the pressure term is nearly equal to unity, suitable comparisons between g~m~ and g~s~ can be made.
+
 # References {.unnumbered}

### 2016-03-04 gm_newphyt/manuscript_gmes.Rmd
jmax vcmax probably wrong
fixed bulk leaf 13C test in discussion

fixed references (can refer to these for working latex tags for other papers)

@@
 Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+
 From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011] using default parameters.
+
+As g~m~ was known
+
+An alternate non-rectangular hyperbola version of the biochemical model of [@farquhar1980biochemical], accounting for g~m~, was used to quantify chloroplastic J~max~ and Vc~max~ [@ethier2004need].
+All AC~i~ curves were fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011] using default parameters and treatment specific measured g~m~.
+Conventional AC~i~ curves, fitted without g~m~, are provided for reference in Supporting Information (Figure `r accfig`).
+
+We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~.
+Comparison of AC~c~ curves (Figure `r accfig`) and AC~i~ curves revealed similar differences between sun and shade leaves.
 
@@
 Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively).
-We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~.
-Comparison of AC~c~ curves (Figure `r accfig`) and AC~i~ curves revealed similar differences between sun and shade leaves.
 
@@
 Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased ITE in shade leaves throughout the experiment.
-
-Additionally, consistently lower leaf $\delta$^13^C in shade leaves suggests that this pattern was likely prevalent long term.
 From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A~n~* in sun leaves was characterized by low rates of g~s~ and low C~i~.
-
+Consistently lower leaf $\delta$^13^C in shade leaves also suggests that observed higher C~i~ and C~c~ in shade leaves was likely prevalent long term.
 Relative to the differences in *A~n~* between leaf types, higher rates of g~s~ in shade leaves appear to exhibit inefficient water use.
@@
 \
-```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/aci_vcmaxnitro.R")
-```  
+# ```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+# source("master_scripts/aci_vcmaxnitro.R")
+# ```  
  \

### 2016-05-11 gm_newphyt/manuscript_gmes.Rmd
added clean aci files and restarted ACc script for manuscript
waiting for gmes curve fits
thesis reviewer comments on manuscript and thesis versions

--- 2016-03-03 gm_newphyt/manuscript_gmes.Rmd
@@
 ---
-title: Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks
+title: Shade leaves of *Eucalpytus* trees lie in wait to maximize whole canopy carbon gain
 author: ""
@@
 sprinkle_print_method("markdown")
+#Dynamic respones of shade leaf physiology to light availability increases whole tree carbon gain
 ```
@@
 
+```{r install packages, echo=FALSE, message=FALSE, warning=FALSE,results="hide"}
+#install needed packages if not already installed
+.packages = c("lubridate", "doBy", "sciplot", "plyr","plotrix", "wesanderson", "plantecophys", "scales","reshape", "pixiedust")
+
+# Install CRAN packages (if not already installed)
+.inst <- .packages %in% installed.packages()
+if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst], repos = "http://cran.at.r-project.org")
+
+# Load packages into session
+lapply(.packages, library, character.only=TRUE)
+```
+
 ```{r global data/sourcing , include=FALSE}
 source("functions and packages/functions.R")
-source("functions and packages/packages.R")
 source("master_scripts/plot_objects.R")
-library(pixiedust)
 
@@
 deltafig <- "S2"
-accfig <- "S3"
-get1fig <- "S4"
-get2fig <- "S5"
+get1fig <- "S3"
+get2fig <- "S4"
 
@@
 Part of the variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
+It is likely that leaf anatomical costs associated with minimizing the length, diameter and tortuosity of the g~m~ diffusion pathway are necessary to maintain a high g~m~ [@hassiotou2009influence].
 Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack of empirical data currently hinders our ability to interpreting their coupled responses to *A~n~* across sun and shade leaves.
@@
 All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
-Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) with the equation:
+Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange in the leaf cuvette were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) with the equation:
 
@@
 (3)
-$$K = \frac{E}{\Psi_{pd}~-\Psi_l}$$
+$$K_l = \frac{E}{\Psi_{pd}~-\Psi_l}$$
 
 Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A~n~* divided by E.
-The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near].
+The g~1~ parameter, describing the plants water-use strategy, was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near].
 Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use, whereas large values imply a lower C construction cost and decreased water-use efficiency [@lin2015optimal].
@@
 Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
-Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, WP is MPa and $\delta$^13^C is &permil;.
+Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, $\Psi$ is MPa and $\delta$^13^C is &permil;.
 Different letters represent significant differences between leaf type and temperature treatments.
@@
 dust(resourcetab2)%>%
-  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0("$\\delta$","^13^c")) %>%
+  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K",paste0("$\\Psi$","~pre~"),paste0("$\\Psi$","~l~"), paste0("$\\delta$","^13^C")) %>%
   sprinkle(rows= 1, part="head",bold=TRUE)%>%
@@
 ```
+
 \
@@
 The dashed line is the 1:1 relationship.
-
 \
@@
 We thank Craig Barton and Burhan Amiji for maintaining the Whole Tree Chambers experiment and for their outstanding technical assistance.
+We thank Sune Linder and the Swedish University for Agricultural Science for providing the whole-tree chambers.
 
@@
 \
-```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/ACc_figure.R")
-```  
-\
-**Figure S3**.
-Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
-C~c~ values were predicted with g~m~ and curves represent chloroplastic photosynthetic parameters at 25°C and saturating light (1800 $\mu$mols m^-1^ s^-1^).
-\
 ```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
@@
 \
-**Figure S4**.
+**Figure S3**.
 Response of *A~n~* (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
@@
 \
-**Figure S5**.
+**Figure S4**.
 Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.

### 2016-05-11 manuscript/manuscript_wtc3.Rmd
thesis reviewer comments on manuscript and thesis versions

@@
 (13)
-$$K = \frac{E}{\Psi_{pd}~-\Psi_l}$$
+$$K_l = \frac{E}{\Psi_{pd}~-\Psi_l}$$
 
 Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A~n~* divided by E.
-The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near].
+The g~1~ parameter, describing the plants water-use strategy, was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near].
 Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use, whereas large values imply a lower C construction cost and decreased water-use efficiency [@lin2015optimal].
@@
 dust(resourcetab2)%>%
-  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K","WP~pre~","WP~mid~", paste0("$\\delta$","^13^c")) %>%
+  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K",paste0("$\\Psi$","~pre~"),paste0("$\\Psi$","~l~"), paste0("$\\delta$","^13^c")) %>%
   sprinkle(rows= 1, part="head",bold=TRUE)%>%
@@
 ```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/physiology_barplots.R")
+source("scripts/physiology_barplots.R")
 ```  
@@
 ```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/cicc_bar.R")
+source("scripts/cicc_bar.R")
 ```  
@@
 ```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/ACc_figure.R")
+source("scripts/ACc_figure.R")
 ```  

### 2016-05-12 gm_newphyt/manuscript_gmes.Rmd
new manuscrupt draft completed and sent to co-authors

@@
 ---
-title: Shade leaves of *Eucalpytus* trees lie in wait to maximize whole canopy carbon gain
+title: Dynamic physiological behavior allows shade leaves to outperform sun leaves in high light
 author: ""
@@
 *Email: courtneycampany@gmail.com*
-
 \
-```{r wc, echo=FALSE, warning=FALSE}
-library(pixiedust)
-wc_tab <- data.frame(count = c("Total word count (main body):", "Summary:", "Introduction:", "Materials and Methods:", "Results:",                               "Discussion:", "Acknowledgements:"), 
-                     answer = c(5535, 219, 1207, 1847, 1102, 1359,20),
-                     count2 = c("No. of Figures:", "No of Tables:", "No. of Supporting Information files:"
-                                ,"","","",""),
-                     answer2 = c("6 (all in color)",2,"6 (5 Figures, 1 Methods)", "","","",""))
-#colnames(wc_tab) <- NULL 
-dust(wc_tab)%>%
-sprinkle_print_method("markdown")
-#Dynamic respones of shade leaf physiology to light availability increases whole tree carbon gain
-```
+\
+Total word count (excluding summary, references and legends): 5563
+\
+Summary: 200
+\
+Introduction: 1186
+\
+Materials and Methods: 1863
+\
+Results: 1098
+\
+Discussion: 1380
+\
+Acknowledgements: 36
+\
+Number of figures: 7 (6 in color)
+\
+Number of tables: 2
+\
+Number of supporting information files: 4 figures, 1 methods
 
@@
 When light intensity was increased for shade leaves both g~s~ and g~m~ increased, leading to increases in A greater than sun leaves.
-* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
+* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for up-scaling leaf level measurements and predicting whole canopy carbon gain.
 We argue that the up-regulation of g~m~ over short time intervals with light enables shade leaves to respond quickly to sunflecks, possibly representing a new mechanism underpinning leaf gas exchange responses to light.
@@
 Light availability is one of the most important environmental drivers of leaf carbon (C) uptake in trees.
-Predicting C uptake of forests usually involves upscaling leaf level measurements to assess whole canopy function.
-Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling based on solar irradiance problematic.
+Predicting C uptake of forests usually involves up-scaling leaf level measurements to assess whole canopy function.
+Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple up-scaling based on solar irradiance problematic.
 Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor].
@@
 This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis].
-How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for upscaling *A~n~* for sun and shade leaves.
+How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for up-scaling *A~n~* for sun and shade leaves.
 Additionally, light saturated rates of *A~n~* are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
@@
 ##Conclusions
-Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy.
+Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for up-scaling leaf level measurements to the canopy.
 Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
@@
 Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
-Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, $\Psi$ is MPa and $\delta$^13^C is &permil;.
+Units of LMA and leaf N~a~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, $\Psi$ is MPa and $\delta$^13^C is &permil;.
 Different letters represent significant differences between leaf type and temperature treatments.
@@
 
-```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png",message=FALSE, dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/par2.R")
@@
 \
-```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600,message=FALSE, dev.args=list(type="cairo")}
  source("master_scripts/aci_vcmaxnitro.R")
@@
 \
-```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600,message=FALSE, dev.args=list(type="cairo")}
 source("master_scripts/physiology_boxplots.R")
@@
 \
-```{r gmshade, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+```{r gmshade, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600,message=FALSE, dev.args=list(type="cairo")}
 source("master_scripts/gmes_shadeleaves_figure.R")
@@
 \
-```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/cicc_boxplots.R")
@@
 
-Here, g~m~ is expressed in units of mol m^-2^ s^-1^ bar^-1^ to be consistent with Evans & Von Caemmerer [-@evans2013temperature].
+Here, g~m~ is expressed in units of mol m^-2^ s^-1^ bar^-1^ to be consistent with Evans & von Caemmerer [-@evans2013temperature].
 As the pressure term is nearly equal to unity, suitable comparisons between g~m~ and g~s~ can be made.

### 2016-05-20 gm_newphyt/manuscript_gmes.Rmd
deleted unused markdowns
added thesis review edits to only thesis version

@@
 #install needed packages if not already installed
-.packages = c("lubridate", "doBy", "sciplot", "plyr","plotrix", "wesanderson", "plantecophys", "scales","reshape", "pixiedust")
+.packages = c("lubridate", "doBy", "plyr","plotrix", "wesanderson", "plantecophys", "scales","reshape", "pixiedust")
 

### 2016-05-20 manuscript/manuscript_wtc3.Rmd
deleted unused markdowns
added thesis review edits to only thesis version

@@
 Due to a limited range of data for the drought treatment only well-watered trees are reported, which reduces the sample size of WTC (n=6 to n=3) for the final 3 months of the experiment.
+This limited amount of data is attributed to small sample sizes as well as a difficulty in generating a sufficient CO~2~ drawdown inside the leaf cuvette, due to drought conditions, needed to accurately measure g~m~.
 
@@
 The CO~2~ in the leaf cuvette was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mol s^-1^.
+We did not explicitly control relative humidity within the leaf cuvette.
 Two identical gas exchanges systems were run simultaneously, one in each of an AT or ET WTC.
@@
 All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
+The degree to which leaf cuvette conditions, especially for shade leaves, modified $\Psi_l$ is unknown.
 Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange in the leaf cuvette were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) with the equation:

### 2016-05-20 markdown/ags_plots.R

@@ -1,162 +1,0 @@
-# source("functions and packages/functions.R")
-# source("master_scripts/plot_objects.R")
-# source("functions and packages/packages.R")
-library(mgcv)
-library(lme4)
-
-library(plantecophys)
-
-#read in gm data set (no drought) and Cibar(discrimination)
-
-Ci_bar <- read.csv("calculated_data/Ci_bar.csv")
-
-gmes <- read.csv("calculated_data/gmes_wellwatered.csv")
-##calculate CC
-gmes$Cc<- with(gmes, Ci-Photo/gm)
-gmes$cc_ci<- with(gmes, Ci/Cc)
-
-###get average by id
-gm_agg <- summaryBy(Photo+Cond+CTleaf+Ci+Cc+gm+VpdL+xsi+DELTA+cc_ci ~ chamber+id+leaf +light+temp+leaflight, 
-                    data=gmes, FUN=mean, keep.names=TRUE)
-
-##remove shade-high
-gm_sunsha <- gm_agg[gm_agg$leaflight != "shade-high",]
-gm_sunsha <- droplevels(gm_sunsha)
-gm_sunsha$id <- gsub("-high", "", gm_sunsha$id)
-gm_sunsha$id <- gsub("-low", "", gm_sunsha$id)
-
-##merge
-gm_c13 <- merge(gm_sunsha, Ci_bar[, c(2,8)], by="id")
-gm_c13$Cc_bar <- with(gm_c13, ci_bar-Photo/gm)
-
-
-###Photosynthesis vs gs (need to fit sun with something else)------------------------------------------------
-###simple plot use smoothplot from RD and get CI polygon
-palette(c(shacol, suncol))
-
-###try to add 95%ci as dotted line and then turn shading of addpoly lighter
-
-####run gam models and then predict
-#SUN leaves
-sunmod <- gam(Photo ~ s(Cond, k=5), data=gm_c13, subset=leaflight=="sun-high")
-
-#predict
-#get apprpriate vector of gs from sun leaves
-gsdat <- gm_c13[gm_c13$leaflight=="sun-high", "Cond"]
-
-#generate sequence and then predict
[…]

### 2016-05-30 manuscript/manuscript_wtc3.Rmd
changed Acc figure in thesis rmd to include points

@@
 
+as higher temperatures
+
 #3.3  Materials and Methods
@@
 Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarfig`).
+Mixing fans in each chamber maintained air movement to minimize leaf boundary layer, which allows leaf temperatures in each WTC close to those of outside control trees [@barton2010whole].
+Overall, minimal WTC effects on tree performance were previously detected across a range of structural characteristics and physiological parameters compared to outside trees [@medhurst2006whole].
+
+\
 Trees grew quickly and developed large canopies, with height growth reaching the top of the WTCs over the experiment duration.
@@
 Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
-Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
+Values of Vc~max~ and J~max~ are treatment means (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light iwth units $\mu$mol m^-2^ s^-1^.
 Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, $\Psi$ is MPa and $\delta$^13^C is &permil;.
@@
 This could be due to leaf formation under comparative light conditions or possible unmeasured differences in total non-structural carbohydrates contents between leaf types.
+Alternatively, plasticity in leaf structure and anatomy of this species may have been necessary to achieve the observed dynamic physiological responses to light.
+As gradients in LMA within tree canopies are commonly observed, this finding warrants further study.
 
@@
 ```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("scripts/ACc_figure.R")
+source("scripts/ACc_figure_points.R")
 ```  

### 2016-06-01 manuscript/manuscript_wtc3.Rmd
change hypo

@@
 \
-4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporation demand from higher temperatures lead to stomatal closure.
-
-as higher temperatures
+4.
+We predicted that climate warming would decrease g~s~ and leaf C gain in sun leaves more so than shade leaves during summer months, as increased evaporation demand from higher temperatures and irradiance lead to stomatal closure.
 
@@
 Two identical gas exchanges systems were run simultaneously, one in each of an AT or ET WTC.
-Leaf temperatures were controlled at the current AT or ET WTC air temperature.
+Leaf temperatures were controlled at the current AT or ET WTC air temperature, thus observed VPD is based on leaf temperatures in the cuvette.
 PPFD in the cuvette was set to match the individual light environment of each leaf type (explained above).
@@
 Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased ITE in shade leaves throughout the experiment.
-Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was likely prevalent long term.
+As the VPD used to calculate ITE was recorded from controlled cuvette leaf temperatures, it is unknown if actual leaf temperatures varied within the canopy and how this could alter patterns in ITE.
+However, consistently higher leaf $\delta$^13^C in shade leaves suggests that the observed patterns in instantanesous ITE were likely prevalent long term.
 From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A~n~* in sun leaves was characterized by low rates of g~s~ and low C~i~.

### 2016-06-17 manuscript/manuscript_wtc3.Rmd
cleaning repo

@@
 Leaf gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.
-At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm, height of 348&pm;15.1 cm and an leaf area of 3.9&pm;0.1 m^2^.
-For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an leaf area of 6.2&pm;0.2 m^2^.
+At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm, height of 348&pm;15.1 cm and a leaf area of 3.9&pm;0.1 m^2^.
+For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and a leaf area of 6.2&pm;0.2 m^2^.
 Leaf area was calculated based on complete leaf counts and mean leaf size from a subsample.

### 2016-06-17 old gm calc/gmes.R

@@ -1,74 +1,0 @@
-#gmes from leaf level (shade and sun leaves) and either high or low light
-#light levels are set on a per tree basis
-#run paired ttests for each chamber comparison, should be 3-4 measurements per leaf per tree
-
-#read data
-gmes_oct <- read.csv("raw data/gmes_oct.csv")
-gmes_oct$chamber <- as.factor(gmes_oct$chamber)
-
-gmes_jan <- read.csv("raw data/gmes_jan.csv")
-gmes_jan$chamber <- as.factor(gmes_jan$chamber)
-
-
-#combine leaf and light variables to make a unique ID
-gmes_oct$leafID <- as.factor(paste(gmes_oct$leaf, gmes_oct$par_type, sep = "-"))
-gmes_jan$leafID <- as.factor(paste(gmes_jan$leaf, gmes_jan$type, sep = "-"))
-
-
-
-#quick plots
-
-#Jan-------------------------------------------------------------------------------------------
-
-jan_plot <- subset(gmes_jan, select = c("chamber", "photo", "cond", "leafID", "gm", "leaf_temp", "temp"))
-
-
-with(subset(jan_plot,leafID=="sun-high"), plot(gm~leaf_temp, col=temp, ylim=c(0.0, .50), xlim=c(18, 35)))
-points(gm~leaf_temp, data=subset(jan_plot,leafID=="shade-high"),  pch=16, col=temp)
-points(gm~leaf_temp, data=subset(jan_plot,leafID=="shade-low"),  pch=15, col=temp)
-
-
-
-#linear model to show fit on graph
-sunmodel <- lm(gm~ leaf_temp, data = subset(jan_plot,leafID=="sun-high"))
-summary(sunmodel)
-
-shadehighmodel <- lm(gm~ leaf_temp, data = subset(jan_plot,leafID=="shade-high"))
-summary(shadehighmodel)
-
-shadelowmodel <- lm(gm~ leaf_temp, data = subset(jan_plot,leafID=="shade-low"))
-summary(shadelowmodel)
-
-
-with(subset(jan_plot,leafID=="sun-high"), plot(gm~leaf_temp, pch=16,col="red",ylim=c(0.0, .50), xlim=c(18, 35)))
-points(gm~leaf_temp, data=subset(jan_plot,leafID=="shade-high"),  pch=16, col="blue")
[…]

### 2016-06-17 old gm calc/paired.R

@@ -1,316 +1,0 @@
-source("functions and packages/functions.R")
-source("functions and packages/packages.R")
-#source("scripts/read_data.R")
-source("master_scripts/plot_objects_paired.R")
-
-
-###there is no ch10 shade in march, check for this in script and paired comparisons
-
-#read in gmsunsha
-#gm data, no drought, no shade high
-gm<- read.csv("calculated_data/gm_sunsha.csv")
-gm$variable <- paste(gm$leaf, gm$temp, sep="-")
-
-cl <- (c("yellowgreen", "green4"))
-sslab <- c("Shade", "Sun")
-
-png(filename = "output/presentations/ciA.png", width = 11, height = 8.5, units = "in", res= 400)
-plot(Photo~Ci, data=gm, pch=21, bg=cl[leaf], xlim=c(0,400), ylim=c(0,30), xlab=cilab, ylab="", cex=1.3)
-title(ylab=satlab, mgp=ypos, cex=1.2)
-legend("topright", sslab, pch=21, pt.bg=cl, pt.cex=2, bg="white",inset = 0.03) 
-dev.off()
-
-png(filename = "output/presentations/gmA.png", width = 11, height = 8.5, units = "in", res= 400)
-plot(Photo~gm, data=gm, pch=21, bg=cl[leaf], xlim=c(0,1), ylim=c(0,30), xlab=gmlab, ylab="", cex=1.3)
-title(ylab=satlab, mgp=ypos, cex=1.2)
-legend("topright", sslab, pch=21, pt.bg=cl, pt.cex=2, bg="white",inset = 0.03) 
-dev.off()
-
-png(filename = "output/presentations/ccA.png", width = 11, height = 8.5, units = "in", res= 400)
-plot(Photo~Cc, data=gm, pch=21, bg=cl[leaf], xlim=c(0,400), ylim=c(0,30), xlab=cclab, ylab="", cex=1.3)
-title(ylab=satlab, mgp=ypos, cex=1.2)
-legend("topright", sslab, pch=21, pt.bg=cl, pt.cex=2, bg="white",inset = 0.03) 
-dev.off()
-
-gm$ratio<- with(gm, gm/Cond)
[…]

### 2016-06-17 thesis2manuscript_notes.R

@@ -1,24 +1,0 @@
-###notes on manuscrupt to thesis document
-
-###figures----------------------------------------------------------------------------------------------------------------
-
-#1. changed physiology bar plots to box plots (includes changes to caption)
-#2. changed drawdown bar plots to box plots (includes changes to caption)
-#3. replaced "respone to" to "relationship between" for several figures
-#4. added gmshade figure (as fig 6), renumbered figure 6to7
-
-
-
-###universal---------------------------------------------------------------------------------------------------------------
-
-#1. units on ppm changed to ubar
-#2. units on gm changed to included bar
-
-
-###reworked abstract summary (edits from SVC)
-
-
-##introduction-------------------------------------------------------------------------------------------------------------
-#1. added hanba and terashima to Boardman cite, sunfleck paragraph
-#2. deleted hyptotheses and rewrote last paragraph
-#3. added Buckely cite to N to light sentence

### 2016-06-17 gm_newphyt/manuscript_gmes.Rmd
MT and SVC manuscript edits, this version sent to RD

@@
 ---
-title: Dynamic physiological behavior allows shade leaves to outperform sun leaves in high light
+title: Coupled response of stomatal and mesophyll conductance to light enhances photosynthesis of shade leaves under sunflecks.
 author: ""
@@
 \
-Courtney E. Campany^1^, Mark Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^.
+Courtney E. Campany^1^, Mark G. Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^.
 
 \
-^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia
+^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, 2751 Australia
 \
-^2^ ARC Centre of Excellence for Translational Photosynthesis, Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, Australia
+^2^ ARC Centre of Excellence for Translational Photosynthesis, Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, 2601 Australia
 
@@
 includeFigs <- TRUE
-
-# This document can be compiled with the 'Knit Word' button in Rstudio, or:
-# library(rmarkdown)
-# render("manuscript.Rmd", "word_document", "manuscript.docx")
 ```
@@
 pval2 <- c(0.001,	0.001,	0.001,	0.001,	0.001,	0.005,	0.001,	0.001)
-# getab <- read.csv("master_scripts/ge_table.csv")
-
-#treatments <- read.csv("raw data/chamber_trt.csv")
 ```
 
-```{r, echo=FALSE}
+```{r figtabs, echo=FALSE}
 
@@
 # Summary
-* Light gradients within tree canopies play a major role in the distribution of plant resources that define leaf photosynthetic capacity.
-A lack of empirical data relating photosynthesis to leaf physiological behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
-* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
-* Compared to sun leaves, shade leaves had lower A, g~m~, leaf nitrogen and photosynthetic biochemical parameters (Vc~max~ and J~max~), but g~s~ was similar.
-When light intensity was increased for shade leaves both g~s~ and g~m~ increased, leading to increases in A greater than sun leaves.
-* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for up-scaling leaf level measurements and predicting whole canopy carbon gain.
-We argue that the up-regulation of g~m~ over short time intervals with light enables shade leaves to respond quickly to sunflecks, possibly representing a new mechanism underpinning leaf gas exchange responses to light.
+* Light gradients within tree canopies play a major role in the distribution of plant resources that define the photosynthetic capacity of sun and shade leaves.
+However, biochemical and diffusional constraints on gas exchange in sun and shade leaves in response to light remain poorly quantified, but critical for predicting carbon and water exchange in canopies.
+* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in the field in climate controlled whole tree chambers.
+* Compared to sun leaves, shade leaves had lower *A~n~*, g~m~, leaf nitrogen and photosynthetic biochemical parameters (Vc~max~ and J~max~), but g~s~ was similar.
+When light intensity was temporarily increased for shade leaves to match that of sun leaves, both g~s~ and g~m~ increased, leading to increases in *A~n~* greater than sun leaves.
+* Here we show that dynamic physiological responses of shade leaves to altered light environments have important implications for up-scaling leaf level measurements and predicting whole canopy carbon gain.
+Despite exhibiting reduced photosynthetic biochemical capacity, the rapid up-regulation of g~m~ with increased light enables shade leaves to respond quickly to sunflecks, possibly representing a new mechanism underpinning leaf gas exchange responses to light.
 
 ##Key words
-leaf optimal behavior, mesophyll conductance, photosynthesis, shade, stomatal conductance,
+*Eucalyptus tereticornis*, mesophyll conductance, photosynthesis, stomatal conductance, sunflecks
 
@@
 Predicting C uptake of forests usually involves up-scaling leaf level measurements to assess whole canopy function.
-Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple up-scaling based on solar irradiance problematic.
-Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor].
-Consequently, structural and functional properties of leaves are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of leaf photosynthesis (*A~n~*) [@evans1995carbon].
+Due to the costs and limitations of efficient light harvesting within plant canopies, not all leaves are exposed to full sun [@niinemets2010review], making simple up-scaling based on solar irradiance problematic.
+Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index (ratio of leaf area to ground area), creating a steep light gradient from the canopy top to bottom [@monsi2005factor].
+Consequently, structural and functional properties of leaves within canopies are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance with canopy depth strongly affects rates of leaf photosynthesis (*A~n~*) [@evans1995carbon].
 To estimate whole canopy C gain it is thus necessary to account for the non-linear response of *A~n~* to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
@@
 The distribution of resources required for *A~n~*, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients.
-As *A~n~* has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co; @buckley2013does].
-Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic].
+As *A~n~* has a saturating response with light and maximum rates depend, in part, on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co; @buckley2013does].
+Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance of the growth environment [@evans2001photosynthetic].
 Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
-Higher rates of *A~n~* and stomatal conductance (g~s~) can only be sustained if the leaf specific hydraulic conductance (K~l~) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal].
-Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A~n~* if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co], thus K~l~ should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
+Higher rates of *A~n~* and stomatal conductance (g~s~) in sun leaves can only be sustained if the leaf specific hydraulic conductance (K~l~) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal].
+Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A~n~* if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co]; thus K~l~ should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
 
@@
 Rates of photosynthesis in C~3~ plants are limited by the [CO~2~] available for fixation by Rubisco within the chloroplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmosphere to the site of carboxylation [@warren2008stand].
-This drawdown consists of multiple resistance pathways to CO~2~ diffusion which include CO~2~ diffusion from the atmosphere through stomata (stomatal conductance, g~s~) and then from these intercellular air spaces into the chloroplast (mesophyll conductance, g~m~).
-Based on optimal theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize *A~n~* [@peltoniemi2012co].
+This drawdown consists of two major resistance pathways to CO~2~ diffusion, namely, the CO~2~ diffusion from the atmosphere through stomata into intercellular air spaces (stomatal conductance, g~s~) and the pathway from the intercellular air spaces into the chloroplast of the mesophyll cells (mesophyll conductance, g~m~).
+Based on optimality theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize *A~n~* [@peltoniemi2012co].
 This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
-Mesophyll conductance can also impose limitations on *A~n~* as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A~n~* [@niinemets2007photosynthesis] if g~m~ constrains CO~2~ supply to the chloroplast.
-Part of the variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
-It is likely that leaf anatomical costs associated with minimizing the length, diameter and tortuosity of the g~m~ diffusion pathway are necessary to maintain a high g~m~ [@hassiotou2009influence].
+Mesophyll conductance can also impose limitations on *A~n~* that can be as large as g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A~n~* [@niinemets2007photosynthesis], if g~m~ constrains CO~2~ supply to the chloroplast.
+Part of the variation in photosynthetic capacity between sun and shade leaves may arise from differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
+It is likely that leaf anatomical costs associated with minimizing the length and tortuosity of the g~m~ diffusion pathway are necessary to maintain the benefits of a high g~m~ [@hassiotou2009influence].
 Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack of empirical data currently hinders our ability to interpreting their coupled responses to *A~n~* across sun and shade leaves.
@@
 How shade leaves utilize sunflecks for short term C gain depends on the combined response time of g~s~ and g~m~ and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks;@tausz2005dynamic].
-The utilization of sunflecks is first limited by delayed responses of stomata opening, which may take minutes, effectively limiting the maximum assimilation rate that can be achieved [@pearcy1990sunflecks; @vicomanz2011; @way2012sunflecks].
+The utilization of sunflecks in shade leaves is first limited by delayed responses of stomata opening, which may take minutes, effectively limiting the maximum assimilation rate that can be achieved [@pearcy1990sunflecks; @vicomanz2011; @way2012sunflecks].
 Mesophyll conductance has been shown to respond to environmental factors (e.g.
-CO~2~, temperature or vapor pressure deficit) at timescales of minutes, possibly faster than g~s~ [@flexas2008mesophyll and references therein], yet short term response to light availability are unclear.
+CO~2~, temperature or vapor pressure deficit) at timescales of minutes, possibly faster than g~s~ [@flexas2008mesophyll and references therein], yet the short term response to light availability is unclear.
 For example, g~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid].
-Anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative; terashima2001sun; @hanba2002effect] are also unlikely to adjust during short light fluctuations.
-The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance and as a potential response to transitory light availability.
+Anatomical parameters which regulate g~m~ with changing growth irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative; @terashima2001sun; @hanba2002effect] are also unlikely to adjust during transient fluctuations in light.
+The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance of the growth environment and as a potential response to transitory light availability.
 
@@
 How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for up-scaling *A~n~* for sun and shade leaves.
-Additionally, light saturated rates of *A~n~* are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
+Additionally, light-saturated rates of *A~n~* are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
 How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients.
@@
 \
-In this study we use *Eucalyptus tereticornis* trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to empirically test whether the distribution of nitrogen and water supply and leaf physiological behaviour result in  higher photosynthetic capacity in sun leaves compared to shade leaves.
-As leaves which developed in the shade were expected to have lower biochemical photosynthetic capacity, we then predicted that sunfleck simulations would not increase *A~n~* to rates similar to sun leaves.
-We further predicted that climate warming would decrease g~s~ and leaf C gain in sun leaves during summer months, as increased evaporation demand from higher temperatures lead to stomatal closure.
+In this study we use *Eucalyptus tereticornis* trees, planted outdoors in naturally sunlit climate controlled whole tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to test whether the distribution of N and water supply and leaf physiological behaviour result in  higher photosynthetic capacity in sun leaves compared to shade leaves.
+We further aimed to quantify the contraints on An in sun and shade leaves arising from biochemical photosynthetic capacity and the CO~2~ diffusion pathway via stomatal and mesophyll conductance.
+As leaves which developed in the shade were expected to have lower biochemical photosynthetic capacity and correspondingly lower diffusive conductance than sun leaves, we predicted that sunfleck simulations would increase *A~n~* and g~s~ to values lower than that of sun leaves.
+We further predicted that climate warming would decrease g~s~ and leaf C gain in sun leaves more so thatn shade leaves during summer months, as increased evaporation demand from higher temperatures and irradiance lead to stomatal closure.
 
@@
 Twelve *Eucalyptus tereticornis* Sm.
-seedlings, chosen from a single local Cumberland plain cohort, were planted in March 2013 into 12 whole-tree chambers (WTC) at the Hawkesbury Forest Experiment site near Richmond, NSW, Australia.
-Each chamber has a height of 9 m and seedlings were grown for 15 months.
+seedlings, chosen from a single local Cumberland plain cohort, were planted in March 2013 into 12 whole-tree chambers (WTC) at the Hawkesbury Forest Experiment site near Richmond, New South Wales, Australia.
+Each chamber was 9 m tall, which accommodated growth of trees for 15 months.
 A detailed description of the WTC operation and design is available in Barton et al.
-[-@barton2010whole].
-Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarfig`).
-Trees grew quickly and developed large canopies, with height growth reaching the top of the WTCs over the experiment duration.
-Trees were watered weekly with 70 L from March 2013 to November 2013.
+[-@barton2010whole] and methods for this experiment in [@drake2016does].
+Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of ambient +3&deg;C (ET, Figure `r airvarfig`).
+The trees grew quickly and developed large canopies, with height growth reaching the top of the WTCs at the end of the experiment.
+Trees were watered weekly with 70 L from March 2013 to September 2013.
 From October 2013 to the end of the experiment trees were watered every 15 days with the mean monthly rainfall amount for Richmond, NSW.
-In February 2013 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering.
-Due to a limited range of data for the drought treatment only well-watered trees are reported, which reduces the sample size of WTC (n=6 to n=3) for the final 3 months of the experiment.
+In February 2014 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering.
+Due to a limited range of data for the drought treatment only well-watered trees are reported, which reduces the sample size of WTC (*n*=6 to *n*=3) for the final 3 months of the experiment.
+This limited amount of data is attributed to small sample sizes as well as the difficulty in generating a sufficient CO~2~ drawdown inside the leaf cuvette, due to drought conditions, needed to accurately measure g~m~.
 
 \
-Before seedlings were planted into each chamber they were maintained under well-watered conditions in 25 L pots and kept inside each chamber.
-This allowed for seedlings to gain sufficient size before planting while also allowing them to acclimate to chamber temperature treatments.
-Seedlings were planted into each chamber after mean seedling height reached ca.
-100 cm.
-The top soils at this site, used in both pots and chambers, are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg-1 total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity.
+The top soils at this site, used in the chambers, are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg^-1^ total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity.
 A root exclusion barrier extended from chamber walls to the hard layer (ca.
 1 m) and roots were allowed to grow freely below the barrier.
+
+\
 Leaf gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.
@@
 For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an leaf area of 6.2&pm;0.2 m^2^.
-Leaf area was calculated based on complete leaf counts and mean leaf size from a subsample.
+Leaf area was calculated based on complete leaf counts and mean leaf size from a subsample [@drake2016does].
 
-##Leaf gas exchange, online carbon isotope discrimination and mesophyll conductance
+##Leaf gas exchange, couple with concurrent measurements of carbon isotope discrimination to estimate mesophyll conductance
 Leaf gas exchange measurements were performed six times, beginning in October 2013 and monthly from December 2013 to April 2014.
@@
 Shade leaves were always measured in the lower canopy, but leaves were sampled on subsequent higher branches across measurement campaigns to minimize confounding effects of leaf age.
-As shade leaves most likely developed slower this assured that older leaves in the bottom canopy were avoided.
+As shade leaves most likely developed more slowly this assured that older leaves in the bottom canopy were avoided.
 
 \
-Prior to gas exchange measurements photosynthetic photon flux density (PPFD) was recorded as a point measurement at the individual leaf level and a spatially averaged measurement at the canopy position for each selected leaf.
-A hand-held quantum sensor (LI-COR, Lincoln, NE, USA) was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either full sun or full shade.
+Prior to gas exchange measurements photosynthetic photon flux density (PPFD) was recorded both as a point measurement at the individual leaf level and a spatially averaged measurement at the canopy position for each selected leaf.
+A hand-held quantum sensor (LI-COR, Lincoln, NE, USA) was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either sun or shade.
 A ceptometer (AccuPAR LP-80, Decagon Devices, Pullman, USA) was then used to measure a spatially averaged PPFD at the canopy height of each chosen leaf type.
@@
 \
-Synchronized gas exchange and C isotope discrimination measurements were made similarly as described in Tazoe et al.
+Concurrent gas exchange and C isotope discrimination measurements were conducted based on methods described in Tazoe et al.
 [-@tazoe2011using] and Evans & von Caemmerer [-@evans2013temperature].
 Leaf level gas exchange was measured with a standard (2 x 3 cm) leaf cuvette using a portable gas exchange system (LI-6400XT, LI-COR, Lincoln, NE, USA).
-This system was coupled with a tunable diode laser (TDL; TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of online C isotope discrimination.
+This system was coupled with a tunable diode laser (TDL; TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of ^13^C isotope discrimination.
 The CO~2~ in the leaf cuvette was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mol s^-1^.
@@
 PPFD in the cuvette was set to match the individual light environment of each leaf type (explained above).
-Periods of high irradiance were simulated for shade leaves by increasing the leaf cuvette PPFD (LI-COR red/blue light source) to match the light environment of the full sun leaf in the same tree.
+Sustained periods of high irradiance (sunflecks) were simulated for shade leaves by increasing the leaf cuvette PPFD (LI-COR red/blue light source) to match the light environment of the sun leaf in the same tree.
 The maximum sunfleck response of shade leaves was then recorded once CO~2~ and water vapour fluxes re-stabilized in the leaf cuvette (ca.
@@
 Reference, sample and two calibration gases were run on alternating 80 s loops (20 s each), one for each AT and ET leaf at a matched canopy position, for a total of 12 min.
-This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference line and samples line gas streams for calculations.
+This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference line and sample line gas streams for calculations.
 The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDL on each measurement cycle.
@@
 Full descriptions of the carbon isotope discrimination equations, with ternary corrections, are presented in Supporting Information Methods S1.
-The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck where compared as a function of C~i~:C~a~.
-Once g~m~ was calculated the chloroplast CO~2~ partial pressure (C~c~) and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were estimated from the relationship between g~m~ to the leaf photosynthesis rate (*A~n~*) by:
+The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck were compared as a function of C~i~:C~a~.
+Once g~m~ was calculated, the chloroplast CO~2~ partial pressure (C~c~) and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were estimated from the relationship between g~m~ and leaf photosynthesis rate (*A~n~*) by:
 
@@
 ##Biochemical parameters of photosynthesis
-Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each WTC prior to the initiation of the drought treatment.
+Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each WTC, sampling in February 2014 when all tress were well-watered in each temperature treatment (*n*=6).
 Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
@@
 ##Leaf chemistry and hydraulic parameters
-Following gas exchange measurements each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, dried and weighed.
+Following gas exchange measurements each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, oven-dried and weighed.
 These leaves were then milled and analyzed for leaf N content and $\delta$^13^C.
 Leaf samples were analysed on a Delta V Advantage coupled to a Flash HT and Conflo IV (Thermo Fisher Scientific, Bremen, Germany) in dual-reactor setup.
-Samples were flash combusted at 1000&deg;C and converted to CO~2~ and N~2~ and then subjected to stable isotope ratio mass spectrometry.
+Samples were flash combusted at 1000 &deg;C and converted to CO~2~ and N~2~ and then subjected to stable isotope ratio mass spectrometry.
 Leaf N is reported on an area basis (N~a~, g m^-2^) and isotopic signatures of dry matter are reported relative to standard Vienna Pee Dee Belemnite.
@@
 \
-Predawn ($\Psi_{pd}$) and midday ($\Psi_l$) leaf water potentials (MPa) were measured for sun and shade leaves during each gas exchange campaign using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA).
-The leaf closest to the leaf used for gas exchange was sampled for measurement of $\Psi_{pd}$.
-Predawn leaf water potential was measured before sunrise on the same day as gas exchange measurements.
-Leaves used for gas exchange were immediately sampled for $\Psi_l$ once measurements were completed.
+Prior to the gas exchange measurements, predawn ($\Psi_{pd}$) leaf water potentials (MPa) were measured for a separate set of sun and shade leaves using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA).
+The leaf closest to the leaf used for gas exchange was sampled for measurement of $\Psi_{pd}$ before sunrise on the same day as gas exchange measurements.
+Following the gas exchange measurements, leaves used for gas exchange were immediately sampled for midday ($\Psi_l$) leaf water potentials.
 All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
-Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange in the leaf cuvette were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) with the equation:
+Leaf water potentials and transpiration (E, mmol m^-2^ s^-1^) from gas exchange in the leaf cuvette were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) with the equation:
 
@@
 Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A~n~* divided by E.
-The g~1~ parameter, describing the plants water-use strategy, was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near].
-Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use, whereas large values imply a lower C construction cost and decreased water-use efficiency [@lin2015optimal].
+The g~1~ parameter, describing plant water-use strategy, was estimated from ITE vs.
+VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near].
+Small values of g~1~ indicate that transpiration is comparatively costly in C terms and reflecting conservative water use (increaed water-use efficiency), whereas large values imply a lower C cost and decreased water-use efficiency [@lin2015optimal].
 
 ##Data analysis
-Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R [@RDevelopmentCoreTeam2011] with WTC as a random effect.
+Differences in respones of dependent variables to either the warming treatment or leaf type were analysed by mixed-effects models in R [@RDevelopmentCoreTeam2011] with WTC as a random effect.
 Explained variance (R^2^) of mixed models were computed as in [@nakagawa2013general], in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
@@
 ##Leaf resource distribution
-Across six measurement campaigns over the 7 month period PPFD was  reduced on average by >75% in the shade (Figure `r parfig`).
+Across six measurement campaigns over the 7 month period, PPFD was reduced on average by >75% in the shade (Figure `r parfig`).
 Leaf-specific hydraulic conductance (K~l~) was similar across sun and shade leaves (Table `r wtc3tab1`).
@@
 \
-Mean *A~n~* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table `r wtc3tab2`).
+Mean *A~n~* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment and temperatures (Table `r wtc3tab2`).
 Additionally, leaf N~a~ was positively related to *A~n~* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
@@
 No effect of the warming treatment was detected on *A~n~* of sun leaves measured at high light or shade leaves at either low or high light.
-Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
+Photosynthesis within leaf types and warming treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
 
@@
 Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was therefore significantly lower than sun leaves (P < 0.001).
-Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment, but no warming effect was detected in shade leaves measured at low or high light (Table `r wtc3tab2`).
-The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
-For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal ITE model with a g~1~ value for each leaf type and treatment (Figure `r waterusefig`a).
+Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment compared to ambient, but no warming effect was detected in shade leaves measured at low or high light (Table `r wtc3tab2`).
+The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves measured at low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
+For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to the fitted response curves from the optimal ITE model with a specified g~1~ value for each leaf type and treatment (Figure `r waterusefig`a).
 Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).
@@
 No effects of the warming treatment on leaf $\delta$^13^C were detected.
-Leaf $\delta$^13^C and N~a~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C (higher water-use efficiency) and higher N investment in sun leaves.
+Leaf $\delta$^13^C and N~a~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C (higher water-use efficiency) and higher N investment in sun leaves than shade leaves.
 
 ##Leaf carbon isotope discrimination and mesophyll conductance
-The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
+The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for both leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
 Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
@@
 \
-Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
+Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment and temperatures (P < 0.001).
 Following the increase in PPFD from low to high light on the same leaf, g~m~ values of shade leaves increased on average 55% after approximately 25 min (Figure `r gmshadefig`).
@@
 ##Variation in intercellular and chloroplastic CO~2~ concentrations
-Higher rates of g~s~ in shade leaves under low and high light led to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
+Higher g~s~ in shade leaves under low and high light led to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
 The chloroplast CO~2~ partial pressure was comparable between shade leaves when measured at both low and high light conditions (Figure `r ciccfig`c).
@@
 The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and *A~n~* [@caemmerer2014temperature].
-This drawdown was similar between sun and shade leaves measured at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
+This drawdown was similar between sun and shade leaves measured at their local light and temperature environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
 This was the result of the proportional relationship between g~m~ and *A~n~* across all leaves.
@@
 Here we show that *A~n~* in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
-Although shade leaves in lower light environments exhibited relatively high g~s~ and C~i~, it was concomitant increases in g~m~ under periods of high light availability that allowed for this up-regulation of *A~n~*.
+Although shade leaves in lower light environments exhibited relatively high g~s~ and C~i~, it was concomitant increases in g~m~ under periods of high light availability that enabled up-regulation of *A~n~*.
 Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
-These findings offer new insights into how aspects of leaf physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake in future studies.
-Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology.
+These findings offer new insights into how aspects of leaf gas exchange physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake and water use in future studies.
+Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology, owing to temperature acclimation [@aspinwall2016convergent].
 
@@
 The allocation of N~a~ constrains *A~n~* and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
-Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy [@mooney1979environmental].
+Decreasing light availability should decrease the investment of N into photosynthetic enzymes within a canopy [@mooney1979environmental].
 As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical].
@@
 Leaf mass per area, however, was not different between sun and shade leaves.
-This could be due to leaf formation under comparative light conditions or possible unmeasured differences in total non-structural carbohydrates contents between leaf types.
+This could be due to leaf formation under comparable light conditions or possible unmeasured differences in total non-structural carbohydrates contents between leaf types that masked differences in leaf thickness or density.
 
@@
 In this study, variation in leaf N distribution and *A~n~* rates were not associated with subsequent changes in K~l~ between sun and shade leaves.
-Thus, no direct relationship between water supply and N distribution or *A~n~* within the canopy were detected.
+Thus, no direct relationship between water supply capacity and N distribution or *A~n~* within the canopy were detected.
 
 \
-Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased ITE in shade leaves throughout the experiment.
-From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A~n~* in sun leaves was characterized by low rates of g~s~ and low C~i~.
-Consistently lower leaf $\delta$^13^C in shade leaves also suggests that observed higher C~i~ and C~c~ in shade leaves was likely prevalent long term.
-Relative to the differences in *A~n~* between leaf types, higher rates of g~s~ in shade leaves appear to exhibit inefficient water use.
-As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower ITE compared to sun leaves.
+Unexpected higher g~s~ in shade leaves compared to sun leaves led to decreased ITE in shade leaves throughout the experiment.
+From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A~n~* in sun leaves was characterized by low g~s~ and low C~i~.
+Consistently lower leaf $\delta$^13^C in shade leaves also suggests that observed higher C~i~ and C~c~ in shade leaves was likely prevalent long term (Figure `r deltafig`).
+Relative to the differences in *A~n~* between leaf types, higher g~s~ in shade leaves appears to result in inefficient water use.
+As whole canopy C gain integrates the efficiency of all leaves, this raises the question of why shade leaves maintained a lower ITE compared to sun leaves.
 
@@
 It is possible that reducing stomatal response time, by sustaining higher g~s~, is a strategy to take advantage of high light quickly in shade leaves [@tausz2005dynamic].
-Evidence from this study supports this hypothesis, as shade leaves increased *A~n~* equivalent or even outperforming sun leaves under identical light intensity.
+Evidence from this study supports this hypothesis, as shade leaves increased *A~n~* to values equal or even outperforming sun leaves when subjected to a brief period of identical high-light intensity.
 Transpiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck induced thermal load [@schymanski2013stomatal].
 This is because rapid increases in leaf temperature with sunflecks have been shown to inhibit C gain [@leakey2003high].
-However, this response likely occurs at very high air temperatures and may not explain the observed g~s~ in shade leaves across the large natural range of temperatures included in this study.
+However, this response likely occurs at very high air temperatures and may not explain the observed g~s~ in shade leaves across the large natural seasonal range of temperatures included in this study.
 How prevalent each of these strategies are within tree canopies is still unknown, as empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants, often in deep shade [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution].
+though leaves within a forest canopy exhibited a similar uncoupling of *A~n~* and g~s~ across a light gradient [@tjoelker1995interaction].
 Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy [see @de1997simple].
@@
 This provides a potential explanation for the observed increases in g~m~, but the impacts of aquaporins on g~m~ are yet to be tested in leaves of tree species.
-Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
-Here we provide empirical data showing g~m~ not only varies within a canopy but the up-regulation of g~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.
+Our findings support growing evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
+Here we provide empirical data showing g~m~ not only varies within a canopy, but that the up-regulation of g~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.
 
@@
 If the valuation of sunflecks as a C resource is large enough, then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan or across the entire canopy when considering both sun and shade leaf types.
-For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to tolerate decreases in ITE.
+For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to accommodate the cost of decreases in ITE.
 However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.
@@
 ##Conclusions
-Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for up-scaling leaf level measurements to the canopy.
+Here we show that dynamic physiological responses of shade leaves to altered light environments have important implications for up-scaling leaf level measurements to the canopy.
 Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
@@
 **Table 1**.
-*Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments.
-Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
+*Eucalyptus tereticornis* leaf morphological and physiological traits of sun and shade leaves under ambient and elevated temperature treatments.
+Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across six measurement campaigns.
 Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
@@
 *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
-Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).
+Each value reflects the mean (&pm; 1 standard error) for each treatment across six gas exchange campaigns.
 Units for *A~n~* and E are $\mu$mol m^-2^ s^-1^, for g~s~ are mol m^-2^ s^-1^, for g~m~ are mol m^-2^ s^-1^ bar^-1^ and for VPD is kPa.
@@
 **Figure 2**.
-(a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.
+(a) ACi curves for sun and shade leaves grown under elevated (ET) and ambient (AT) temperature treatments.
 ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
@@
 **Figure 3**.
-The relationship between *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
+The relationship between *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments across six measurement campaigns.
 Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
@@
 **Figure 6**.
-The change in measured g~m~ for individual shade leaves following an increase in PPFD to match the light environment of full sun leaves in the same tree.
+The change in measured g~m~ for individual shade leaves following an increase in PPFD to match the light environment of full sun leaves in the same tree measured across six measurement dates.
 Measurements of g~m~ were recorded once CO~2~ and water vapour fluxes were stable in the leaf cuvette, which took approximately 25 minutes after light intensity was increased.
@@
 Response of *A~n~* (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
-Shaded symbols represents each monthly measurement campaign.
+Shaded symbols represent each monthly measurement campaign.
 Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
@@
 [-@evans1986carbon].
-In this experiment, a modified form of this equation presented in Evans & von Caemmerer [-@evans2013temperature] with ternary effect corrections by Farquhar & Cernusak [-@farquhar2012ternary] was used such that:
+In this experiment, a modified form of the equation presented in Evans & von Caemmerer [-@evans2013temperature] with ternary effect corrections by Farquhar & Cernusak [-@farquhar2012ternary] was used such that:
 

### 2016-06-19 gm_newphyt/manuscript_gmes.Rmd
use pacman for package loading

@@
 ```{r install packages, echo=FALSE, message=FALSE, warning=FALSE,results="hide"}
-#install needed packages if not already installed
-.packages = c("lubridate", "doBy", "plyr","plotrix", "wesanderson", "plantecophys", "scales","reshape", "pixiedust")
-
-# Install CRAN packages (if not already installed)
-.inst <- .packages %in% installed.packages()
-if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst], repos = "http://cran.at.r-project.org")
-
-# Load packages into session
-lapply(.packages, library, character.only=TRUE)
+r <- require(pacman)
+if(!r)install.packages("pacman")
+pacman::p_load(lubridate, doBy, plyr,plotrix, wesanderson, plantecophys, scales,reshape, pixiedust)
 ```

### 2016-06-19 gm_newphyt/manuscript_gmes.Rmd
fixed An, gs, gm formatting; more edits to introduction

@@
 \
-^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, 2751 Australia
+^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith 2751 NSW, Australia
 \
-^2^ ARC Centre of Excellence for Translational Photosynthesis, Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, 2601 Australia
+^2^ ARC Centre of Excellence for Translational Photosynthesis, Plant Science Division, Research School of Biology, The Australian National University, Canberra 2601 ACT, Australia
 
@@
 However, biochemical and diffusional constraints on gas exchange in sun and shade leaves in response to light remain poorly quantified, but critical for predicting carbon and water exchange in canopies.
-* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in the field in climate controlled whole tree chambers.
-* Compared to sun leaves, shade leaves had lower *A~n~*, g~m~, leaf nitrogen and photosynthetic biochemical parameters (Vc~max~ and J~max~), but g~s~ was similar.
-When light intensity was temporarily increased for shade leaves to match that of sun leaves, both g~s~ and g~m~ increased, leading to increases in *A~n~* greater than sun leaves.
+* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in the field in climate controlled whole tree chambers.
+* Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic biochemical parameters (Vc~max~ and J~max~), but *g*~s~ was similar.
+When light intensity was temporarily increased for shade leaves to match that of sun leaves, both *g*~s~ and *g*~m~ increased, leading to increases in *A*~n~ greater than sun leaves.
 * Here we show that dynamic physiological responses of shade leaves to altered light environments have important implications for up-scaling leaf level measurements and predicting whole canopy carbon gain.
-Despite exhibiting reduced photosynthetic biochemical capacity, the rapid up-regulation of g~m~ with increased light enables shade leaves to respond quickly to sunflecks, possibly representing a new mechanism underpinning leaf gas exchange responses to light.
+Despite exhibiting reduced photosynthetic biochemical capacity, the rapid up-regulation of *g*~m~ with increased light enables shade leaves to respond quickly to sunflecks, possibly representing a new mechanism underpinning leaf gas exchange responses to light.
 
@@
 Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index (ratio of leaf area to ground area), creating a steep light gradient from the canopy top to bottom [@monsi2005factor].
-Consequently, structural and functional properties of leaves within canopies are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance with canopy depth strongly affects rates of leaf photosynthesis (*A~n~*) [@evans1995carbon].
-To estimate whole canopy C gain it is thus necessary to account for the non-linear response of *A~n~* to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
+Consequently, structural and functional properties of leaves within canopies are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance with canopy depth strongly affects rates of leaf photosynthesis (*A*~n~) [@evans1995carbon].
+To estimate whole canopy C gain it is thus necessary to account for the non-linear response of *A*~n~ to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
 
 \
-The distribution of resources required for *A~n~*, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients.
-As *A~n~* has a saturating response with light and maximum rates depend, in part, on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co; @buckley2013does].
+The distribution of resources required for *A*~n~, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients.
+As *A*~n~ has a saturating response with light and maximum rates depend, in part, on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co; @buckley2013does].
 Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance of the growth environment [@evans2001photosynthetic].
 Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
-Higher rates of *A~n~* and stomatal conductance (g~s~) in sun leaves can only be sustained if the leaf specific hydraulic conductance (K~l~) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal].
-Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A~n~* if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co]; thus K~l~ should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
+Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A*~n~ if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co]; thus leaf-specific  hydraulic conductance (*K*~l~) should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects;@hubbard2001stomatal].
 
 \
-Rates of photosynthesis in C~3~ plants are limited by the [CO~2~] available for fixation by Rubisco within the chloroplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmosphere to the site of carboxylation [@warren2008stand].
-This drawdown consists of two major resistance pathways to CO~2~ diffusion, namely, the CO~2~ diffusion from the atmosphere through stomata into intercellular air spaces (stomatal conductance, g~s~) and the pathway from the intercellular air spaces into the chloroplast of the mesophyll cells (mesophyll conductance, g~m~).
-Based on optimality theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize *A~n~* [@peltoniemi2012co].
+The photosynthetic rate in C~3~ plants is limited by the [CO~2~] available for fixation by Rubisco within the chloroplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmosphere to the site of carboxylation [@warren2008stand].
+There are two major resistance pathways to CO~2~ diffusion: the CO~2~ diffusion from the atmosphere through stomata into intercellular air spaces (stomatal conductance, *g*~s~) and the pathway from the intercellular air spaces into the chloroplast of the mesophyll cells (mesophyll conductance, *g*~m~).
+Based on optimality theory, regulation of *g*~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize *A*~n~ [@peltoniemi2012co].
 This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
-Mesophyll conductance can also impose limitations on *A~n~* that can be as large as g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A~n~* [@niinemets2007photosynthesis], if g~m~ constrains CO~2~ supply to the chloroplast.
-Part of the variation in photosynthetic capacity between sun and shade leaves may arise from differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
-It is likely that leaf anatomical costs associated with minimizing the length and tortuosity of the g~m~ diffusion pathway are necessary to maintain the benefits of a high g~m~ [@hassiotou2009influence].
-Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack of empirical data currently hinders our ability to interpreting their coupled responses to *A~n~* across sun and shade leaves.
+Mesophyll conductance can also impose limitations on *A*~n~ that can be as large as *g*~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A*~n~ [@niinemets2007photosynthesis], if *g*~m~ constrains CO~2~ supply to the chloroplast.
+Part of the variation in photosynthetic capacity between sun and shade leaves may arise from differences in *g*~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
+It is likely that leaf anatomical costs associated with minimizing the length and tortuosity of the *g*~m~ diffusion pathway are necessary to maintain the benefits of a high *g*~m~ [@hassiotou2009influence].
+Nonetheless, like *g*~s~, *g*~m~ may also vary dynamically in response to environmental drivers [@flexas2008mesophyll].
+Thus, stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack of empirical data currently hinders our ability to interpreting their coupled responses to *A*~n~ across sun and shade leaves.
 
 \
-Additionally, accounting for short term light fluctuations within a canopy, via sunflecks, makes assessing shade leaf behaviour difficult.
-How shade leaves utilize sunflecks for short term C gain depends on the combined response time of g~s~ and g~m~ and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks;@tausz2005dynamic].
+How shade leaves utilize sunflecks for short term carbon gain depends on the combined response time of *g*~s~ and *g*~m~ and the underlying photosynthetic machinery acclimated to a low light environment [@pearcy1990sunflecks;@tausz2005dynamic].
 The utilization of sunflecks in shade leaves is first limited by delayed responses of stomata opening, which may take minutes, effectively limiting the maximum assimilation rate that can be achieved [@pearcy1990sunflecks; @vicomanz2011; @way2012sunflecks].
 Mesophyll conductance has been shown to respond to environmental factors (e.g.
-CO~2~, temperature or vapor pressure deficit) at timescales of minutes, possibly faster than g~s~ [@flexas2008mesophyll and references therein], yet the short term response to light availability is unclear.
-For example, g~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid].
-Anatomical parameters which regulate g~m~ with changing growth irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative; @terashima2001sun; @hanba2002effect] are also unlikely to adjust during transient fluctuations in light.
+CO~2~, temperature or vapor pressure deficit) at timescales of minutes, possibly faster than *g*~s~ [@flexas2008mesophyll and references therein], yet the short term response to light availability is unclear.
+For example, *g*~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid].
+Anatomical parameters which regulate *g*~m~ with changing growth irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative; @terashima2001sun; @hanba2002effect] are also unlikely to adjust during transient fluctuations in light.
 The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance of the growth environment and as a potential response to transitory light availability.
@@
 This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis].
-How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for up-scaling *A~n~* for sun and shade leaves.
-Additionally, light-saturated rates of *A~n~* are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
+How these stresses affect the diffusion of CO~2~, through either *g*~s~ or *g*~m~, will have implications for up-scaling *A*~n~ for sun and shade leaves.
+Additionally, light-saturated rates of *A*~n~ are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
 How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients.
@@
 \
-In this study we use *Eucalyptus tereticornis* trees, planted outdoors in naturally sunlit climate controlled whole tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to test whether the distribution of N and water supply and leaf physiological behaviour result in  higher photosynthetic capacity in sun leaves compared to shade leaves.
-We further aimed to quantify the contraints on An in sun and shade leaves arising from biochemical photosynthetic capacity and the CO~2~ diffusion pathway via stomatal and mesophyll conductance.
-As leaves which developed in the shade were expected to have lower biochemical photosynthetic capacity and correspondingly lower diffusive conductance than sun leaves, we predicted that sunfleck simulations would increase *A~n~* and g~s~ to values lower than that of sun leaves.
-We further predicted that climate warming would decrease g~s~ and leaf C gain in sun leaves more so thatn shade leaves during summer months, as increased evaporation demand from higher temperatures and irradiance lead to stomatal closure.
+In this study we use *Eucalyptus tereticornis* trees, planted in naturally sunlit climate controlled whole-tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to test whether the distribution of N and water supply capacity via leaf-specific hydraulic conductance and leaf physiological behaviour result in  higher photosynthetic capacity in sun leaves compared to shade leaves.
+We further aimed to quantify the contraints on *A*~n~ in sun and shade leaves arising from photosynthetic capacity and the CO~2~ diffusion pathway via stomatal and mesophyll conductance.
+As leaves which developed in the shade were expected to have lower photosynthetic capacity and correspondingly lower diffusive conductance (both *g*~s~ and *g*~m~) than sun leaves, we predicted that sunfleck simulations would increase *A*~n~ and *g*~s~ to values lower than that attained by sun leaves in similar conditions.
+We further predicted that climate warming would decrease *g*~s~ and leaf C gain in sun leaves more so than shade leaves during summer months, as increased evaporation demand from higher temperatures and irradiance lead to stomatal closure.
 
@@
 Due to a limited range of data for the drought treatment only well-watered trees are reported, which reduces the sample size of WTC (*n*=6 to *n*=3) for the final 3 months of the experiment.
-This limited amount of data is attributed to small sample sizes as well as the difficulty in generating a sufficient CO~2~ drawdown inside the leaf cuvette, due to drought conditions, needed to accurately measure g~m~.
+This limited amount of data is attributed to small sample sizes as well as the difficulty in generating a sufficient CO~2~ drawdown inside the leaf cuvette, due to drought conditions, needed to accurately measure *g*~m~.
 
@@
 The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDL on each measurement cycle.
-Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the 12 min interval.
+Photosynthesis, *g*~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the 12 min interval.
 \
@@
 
-where $\Delta$~o~ is the observed discrimination and $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, g~m~, respiration and photorespiration, respectively.
+where $\Delta$~o~ is the observed discrimination and $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, *g*~m~, respiration and photorespiration, respectively.
 For this study, the CO~2~ compensation point ($\Gamma_{*}$) and respiration during the day (R~d~) parameters originally derived for tobacco plants [@von1994kinetics] were replaced with parameters derived for *Eucalyptus globulus* from Crous et al.
-[-@crous2012light] when calculating g~m~.
+[-@crous2012light] when calculating *g*~m~.
 Full descriptions of the carbon isotope discrimination equations, with ternary corrections, are presented in Supporting Information Methods S1.
 The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck were compared as a function of C~i~:C~a~.
-Once g~m~ was calculated, the chloroplast CO~2~ partial pressure (C~c~) and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were estimated from the relationship between g~m~ and leaf photosynthesis rate (*A~n~*) by:
+Once *g*~m~ was calculated, the chloroplast CO~2~ partial pressure (C~c~) and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were estimated from the relationship between *g*~m~ and leaf photosynthesis rate (*A*~n~) by:
 
@@
 All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
-Leaf water potentials and transpiration (E, mmol m^-2^ s^-1^) from gas exchange in the leaf cuvette were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) with the equation:
+Leaf water potentials and transpiration (E, mmol m^-2^ s^-1^) from gas exchange in the leaf cuvette were then used to calculate leaf-specific hydraulic conductance (*K*~l~, mmol m^-2^ s^-1^ MPa) with the equation:
 
@@
 
-Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A~n~* divided by E.
+Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A*~n~ divided by E.
 The g~1~ parameter, describing plant water-use strategy, was estimated from ITE vs.
-VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near].
+VPD response curves by fitting a rearranged optimal *g*~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near].
 Small values of g~1~ indicate that transpiration is comparatively costly in C terms and reflecting conservative water use (increaed water-use efficiency), whereas large values imply a lower C cost and decreased water-use efficiency [@lin2015optimal].
@@
 Across six measurement campaigns over the 7 month period, PPFD was reduced on average by >75% in the shade (Figure `r parfig`).
-Leaf-specific hydraulic conductance (K~l~) was similar across sun and shade leaves (Table `r wtc3tab1`).
+Leaf-specific hydraulic conductance (*K*~l~) was similar across sun and shade leaves (Table `r wtc3tab1`).
 This was because neither leaf water potentials ($\Psi_{pd}$ and $\Psi_l$, Table `r wtc3tab1`) nor transpiration rates (E, Table `r wtc3tab2`) differed between leaf types.
@@
 Leaf mass per area (LMA) was not different between leaf types (Table `r wtc3tab1`).
-No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~a~ or LMA either within or across leaf types (P > 0.05).
+No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, *K*~l~, E, N~a~ or LMA either within or across leaf types (P > 0.05).
 
@@
 \
-Mean *A~n~* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment and temperatures (Table `r wtc3tab2`).
-Additionally, leaf N~a~ was positively related to *A~n~* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
-Following an increase in light intensity to match high-light conditions, *A~n~* of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table `r wtc3tab2`).
-No effect of the warming treatment was detected on *A~n~* of sun leaves measured at high light or shade leaves at either low or high light.
+Mean *A*~n~ was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment and temperatures (Table `r wtc3tab2`).
+Additionally, leaf N~a~ was positively related to *A*~n~ across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
+Following an increase in light intensity to match high-light conditions, *A*~n~ of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table `r wtc3tab2`).
+No effect of the warming treatment was detected on *A*~n~ of sun leaves measured at high light or shade leaves at either low or high light.
 Photosynthesis within leaf types and warming treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
@@
 ##Stomatal conductance and leaf water-use efficiency
-On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table `r wtc3tab2`).
-Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and *A~n~* were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
-Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
-No effect of the warming treatment was detected on g~s~ within or across leaf types.
+On average, *g*~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table `r wtc3tab2`).
+Photosynthesis was positively correlated with *g*~s~ in all leaves measured under high light conditions, however, *g*~s~ and *A*~n~ were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
+Following increased PPFD, *g*~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
+No effect of the warming treatment was detected on *g*~s~ within or across leaf types.
 Stomatal conductance within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get2fig`b).
@@
 The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for both leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
-Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
-The remainder $\Delta_o$ consists of the contributions of g~s~, respiration and photorespiration to discrimination.
+Carbon isotope discrimination associated with *g*~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
+The remainder $\Delta_o$ consists of the contributions of *g*~s~, respiration and photorespiration to discrimination.
 
 \
-Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment and temperatures (P < 0.001).
-Following the increase in PPFD from low to high light on the same leaf, g~m~ values of shade leaves increased on average 55% after approximately 25 min (Figure `r gmshadefig`).
-These measured values of g~m~ for shade leaves at high light were also equivalent or greater than those of sun leaves (Table `r wtc3tab2`).
-Proportional increases in g~m~ were matched by proportional increases in *A~n~* from low to high light in shade leaves (Figure `r gebarfig`b,c).
-Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186).
-The large increases in g~m~ in shade leaves under high light likely resulted in the highest rates of *A~n~* (Figure `r agmgsfig`b).
-No differences in g~m~ were detected with the warming treatment within leaf types.
+Mean *g*~m~ was higher in sun compared to shade leaves (+27%) under their local light environment and temperatures (P < 0.001).
+Following the increase in PPFD from low to high light on the same leaf, *g*~m~ values of shade leaves increased on average 55% after approximately 25 min (Figure `r gmshadefig`).
+These measured values of *g*~m~ for shade leaves at high light were also equivalent or greater than those of sun leaves (Table `r wtc3tab2`).
+Proportional increases in *g*~m~ were matched by proportional increases in *A*~n~ from low to high light in shade leaves (Figure `r gebarfig`b,c).
+Photosynthesis scaled positively with increases in *g*~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186).
+The large increases in *g*~m~ in shade leaves under high light likely resulted in the highest rates of *A*~n~ (Figure `r agmgsfig`b).
+No differences in *g*~m~ were detected with the warming treatment within leaf types.
 Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively).
@@
 ##Variation in intercellular and chloroplastic CO~2~ concentrations
-Higher g~s~ in shade leaves under low and high light led to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
+Higher *g*~s~ in shade leaves under low and high light led to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
 The chloroplast CO~2~ partial pressure was comparable between shade leaves when measured at both low and high light conditions (Figure `r ciccfig`c).
 In sun leaves C~c~ was significantly lower than shade leaves, consistent with a lower C~i~.
-The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and *A~n~* [@caemmerer2014temperature].
+The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between *g*~m~ and *A*~n~ [@caemmerer2014temperature].
 This drawdown was similar between sun and shade leaves measured at their local light and temperature environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
-This was the result of the proportional relationship between g~m~ and *A~n~* across all leaves.
+This was the result of the proportional relationship between *g*~m~ and *A*~n~ across all leaves.
 The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
@@
 #Discussion
-Here we show that *A~n~* in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
-Although shade leaves in lower light environments exhibited relatively high g~s~ and C~i~, it was concomitant increases in g~m~ under periods of high light availability that enabled up-regulation of *A~n~*.
+Here we show that *A*~n~ in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
+Although shade leaves in lower light environments exhibited relatively high *g*~s~ and C~i~, it was concomitant increases in *g*~m~ under periods of high light availability that enabled up-regulation of *A*~n~.
 Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
@@
 ##Resource distribution and photosynthetic capacity
-The allocation of N~a~ constrains *A~n~* and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
+The allocation of N~a~ constrains *A*~n~ and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
 Decreasing light availability should decrease the investment of N into photosynthetic enzymes within a canopy [@mooney1979environmental].
 As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical].
-Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and *A~n~* were all reduced in shade leaves.
+Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and *A*~n~ were all reduced in shade leaves.
 Leaf mass per area, however, was not different between sun and shade leaves.
@@
 Using a two-leaf model, Peltoniemi et al.
-[-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if K~l~ is also optimally distributed.
-In this study, variation in leaf N distribution and *A~n~* rates were not associated with subsequent changes in K~l~ between sun and shade leaves.
-Thus, no direct relationship between water supply capacity and N distribution or *A~n~* within the canopy were detected.
+[-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if *K*~l~ is also optimally distributed.
+In this study, variation in leaf N distribution and *A*~n~ rates were not associated with subsequent changes in *K*~l~ between sun and shade leaves.
+Thus, no direct relationship between water supply capacity and N distribution or *A*~n~ within the canopy were detected.
 
 \
-Unexpected higher g~s~ in shade leaves compared to sun leaves led to decreased ITE in shade leaves throughout the experiment.
-From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A~n~* in sun leaves was characterized by low g~s~ and low C~i~.
+Unexpected higher *g*~s~ in shade leaves compared to sun leaves led to decreased ITE in shade leaves throughout the experiment.
+From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A*~n~ in sun leaves was characterized by low *g*~s~ and low C~i~.
 Consistently lower leaf $\delta$^13^C in shade leaves also suggests that observed higher C~i~ and C~c~ in shade leaves was likely prevalent long term (Figure `r deltafig`).
-Relative to the differences in *A~n~* between leaf types, higher g~s~ in shade leaves appears to result in inefficient water use.
+Relative to the differences in *A*~n~ between leaf types, higher *g*~s~ in shade leaves appears to result in inefficient water use.
 As whole canopy C gain integrates the efficiency of all leaves, this raises the question of why shade leaves maintained a lower ITE compared to sun leaves.
@@
 ##Physiological behaviour of sun and shade leaves
-The pattern of inefficient water use in shade leaves is important as we hypothesized that g~s~ and *A~n~* would be proportional across sun and shade leaves.
-In sun leaves, *A~n~* and g~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
-However, lower rates of *A~n~* in shade leaves were not coupled with decreases in g~s~, leading to the observed decreases in ITE.
+The pattern of inefficient water use in shade leaves is important as we hypothesized that *g*~s~ and *A*~n~ would be proportional across sun and shade leaves.
+In sun leaves, *A*~n~ and *g*~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
+However, lower rates of *A*~n~ in shade leaves were not coupled with decreases in *g*~s~, leading to the observed decreases in ITE.
 This is significant as optimal stomatal regulation to balance C gain with water loss has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on sun leaves [e.g.
 @prentice2014balancing; @lin2015optimal].
-As a result, the often used economic framework of balancing costs of using water versus N allocation to predict *A~n~* [@wright2003least] may break down when considering all leaves within a tree canopy.
+As a result, the often used economic framework of balancing costs of using water versus N allocation to predict *A*~n~ [@wright2003least] may break down when considering all leaves within a tree canopy.
 
 \
-It is possible that reducing stomatal response time, by sustaining higher g~s~, is a strategy to take advantage of high light quickly in shade leaves [@tausz2005dynamic].
-Evidence from this study supports this hypothesis, as shade leaves increased *A~n~* to values equal or even outperforming sun leaves when subjected to a brief period of identical high-light intensity.
+It is possible that reducing stomatal response time, by sustaining higher *g*~s~, is a strategy to take advantage of high light quickly in shade leaves [@tausz2005dynamic].
+Evidence from this study supports this hypothesis, as shade leaves increased *A*~n~ to values equal or even outperforming sun leaves when subjected to a brief period of identical high-light intensity.
 Transpiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck induced thermal load [@schymanski2013stomatal].
 This is because rapid increases in leaf temperature with sunflecks have been shown to inhibit C gain [@leakey2003high].
-However, this response likely occurs at very high air temperatures and may not explain the observed g~s~ in shade leaves across the large natural seasonal range of temperatures included in this study.
+However, this response likely occurs at very high air temperatures and may not explain the observed *g*~s~ in shade leaves across the large natural seasonal range of temperatures included in this study.
 How prevalent each of these strategies are within tree canopies is still unknown, as empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants, often in deep shade [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution].
-though leaves within a forest canopy exhibited a similar uncoupling of *A~n~* and g~s~ across a light gradient [@tjoelker1995interaction].
+though leaves within a forest canopy exhibited a similar uncoupling of *A*~n~ and *g*~s~ across a light gradient [@tjoelker1995interaction].
 Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy [see @de1997simple].
@@
 \
-We found that *A~n~* and g~m~ scaled positively across leaf types and, unexpectedly, increased proportionately after leaf stability (~25 min) when light intensity was increased in shade leaves.
-Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
-This provides a potential explanation for the observed increases in g~m~, but the impacts of aquaporins on g~m~ are yet to be tested in leaves of tree species.
-Our findings support growing evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
-Here we provide empirical data showing g~m~ not only varies within a canopy, but that the up-regulation of g~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.
+We found that *A*~n~ and *g*~m~ scaled positively across leaf types and, unexpectedly, increased proportionately after leaf stability (~25 min) when light intensity was increased in shade leaves.
+Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of *g*~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
+This provides a potential explanation for the observed increases in *g*~m~, but the impacts of aquaporins on *g*~m~ are yet to be tested in leaves of tree species.
+Our findings support growing evidence that *g*~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
+Here we provide empirical data showing *g*~m~ not only varies within a canopy, but that the up-regulation of *g*~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.
 
@@
 If shade leaves "lie in wait" for sunflecks, then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution.
-This is because the role of g~s~ in regulating photosynthetic induction impacts the capacity of a leaf to utilize sunflecks [@way2012sunflecks].
+This is because the role of *g*~s~ in regulating photosynthetic induction impacts the capacity of a leaf to utilize sunflecks [@way2012sunflecks].
 If the valuation of sunflecks as a C resource is large enough, then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan or across the entire canopy when considering both sun and shade leaf types.
@@
 These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have important implications for process-based models that predict canopy C gain from rates of leaf photosynthesis.
-Furthermore, the dynamic nature of g~m~ cannot be simply parameterized in tree growth models and possibly should be excluded until it can be represented properly.
-Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of g~m~ to opportunistically increase CO~2~ drawdown.
+Furthermore, the dynamic nature of *g*~m~ cannot be simply parameterized in tree growth models and possibly should be excluded until it can be represented properly.
+Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of *g*~m~ to opportunistically increase CO~2~ drawdown.
 To improve our ability to predict whole canopy C gain, future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.
@@
 *Eucalyptus tereticornis* leaf morphological and physiological traits of sun and shade leaves under ambient and elevated temperature treatments.
-Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across six measurement campaigns.
+Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and *K*~l~ values represent treatment mean (&pm; 1 standard error) across six measurement campaigns.
 Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
-Units of LMA and leaf N~a~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, $\Psi$ is MPa and $\delta$^13^C is &permil;.
+Units of LMA and leaf N~a~ are g m^2^, *K*~l~ is mmol m^-2^ s^-1^ MPa, $\Psi$ is MPa and $\delta$^13^C is &permil;.
 Different letters represent significant differences between leaf type and temperature treatments.
@@
 Each value reflects the mean (&pm; 1 standard error) for each treatment across six gas exchange campaigns.
-Units for *A~n~* and E are $\mu$mol m^-2^ s^-1^, for g~s~ are mol m^-2^ s^-1^, for g~m~ are mol m^-2^ s^-1^ bar^-1^ and for VPD is kPa.
+Units for *A*~n~ and E are $\mu$mol m^-2^ s^-1^, for *g*~s~ are mol m^-2^ s^-1^, for *g*~m~ are mol m^-2^ s^-1^ bar^-1^ and for VPD is kPa.
 Different letters represent significant differences between leaf type, light environment and temperature treatments.
@@
 dust(getab2)%>%
-  sprinkle_colnames("Leaf", "Light","Temperature", "*A~n~*", "g~s~", "g~m~", "ITE","E","VPD","C~i~", "C~c~") %>%
+  sprinkle_colnames("Leaf", "Light","Temperature", "*A*~n~", "*g*~s~", "*g*~m~", "ITE","E","VPD","C~i~", "C~c~") %>%
   sprinkle(rows= 1, part="head",bold=TRUE)%>%
@@
 (b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light.
-(c) The relationship between *A~n~* and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
+(c) The relationship between *A*~n~ and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
 For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
@@
 **Figure 3**.
-The relationship between *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments across six measurement campaigns.
+The relationship between *A*~n~ to *g*~s~ (a) and *g*~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments across six measurement campaigns.
 Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
@@
 \
-**Figure 4**. Box plots of measured g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
+**Figure 4**. Box plots of measured *g*~s~ (a), *g*~m~ (b) and *A*~n~ (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
 \
@@
 **Figure 6**.
-The change in measured g~m~ for individual shade leaves following an increase in PPFD to match the light environment of full sun leaves in the same tree measured across six measurement dates.
-Measurements of g~m~ were recorded once CO~2~ and water vapour fluxes were stable in the leaf cuvette, which took approximately 25 minutes after light intensity was increased.
+The change in measured *g*~m~ for individual shade leaves following an increase in PPFD to match the light environment of full sun leaves in the same tree measured across six measurement dates.
+Measurements of *g*~m~ were recorded once CO~2~ and water vapour fluxes were stable in the leaf cuvette, which took approximately 25 minutes after light intensity was increased.
 The dashed line is the 1:1 relationship.
@@
 **Figure S3**.
-Response of *A~n~* (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
+Response of *A*~n~ (a), *g*~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
 Shaded symbols represent each monthly measurement campaign.
@@
 All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
-Weak negative relationships with g~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (R^2^ = 0.16 and 0.08, respectively).
+Weak negative relationships with *g*~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (R^2^ = 0.16 and 0.08, respectively).
 \
@@
 **Figure S4**.
-Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
+Response of VPD (a), *g*~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
 Shaded symbols represents each monthly measurement campaign.
@@
 
-###Description of the calculation of g~m~ from carbon isotope discrimination during C3 photosynthesis
+###Description of the calculation of *g*~m~ from carbon isotope discrimination during C3 photosynthesis
 \
-Using online C isotope discrimination measurements, the difference between the observed discrimination and what is predicted for light saturated gas exchange is proportional to g~m~ [@griffiths2013mesophyll].
+Using online C isotope discrimination measurements, the difference between the observed discrimination and what is predicted for light saturated gas exchange is proportional to *g*~m~ [@griffiths2013mesophyll].
 First, leaf C isotope discrimination was calculated by comparing the isotopic composition of the reference gas entering the leaf cuvette ($\delta$^13^C~e~) with the sample gas ($\delta$^13^C~o~) such that:
@@
 
-where $\Delta$~o~ is the observed discrimination and  $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, g~m~, respiration and photorespiration, respectively.
+where $\Delta$~o~ is the observed discrimination and  $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, *g*~m~, respiration and photorespiration, respectively.
 The equations for each are as follows:
@@
 \
-The g~m~ can then be calculated as:
+The *g*~m~ can then be calculated as:
 
@@
 
-Here, g~m~ is expressed in units of mol m^-2^ s^-1^ bar^-1^ to be consistent with Evans & von Caemmerer [-@evans2013temperature].
-As the pressure term is nearly equal to unity, suitable comparisons between g~m~ and g~s~ can be made.
+Here, *g*~m~ is expressed in units of mol m^-2^ s^-1^ bar^-1^ to be consistent with Evans & von Caemmerer [-@evans2013temperature].
+As the pressure term is nearly equal to unity, suitable comparisons between *g*~m~ and *g*~s~ can be made.
 

### 2016-06-20 gm_newphyt/manuscript_gmes.Rmd
More details on curve fitting; number of minor edits

@@
 seedlings, chosen from a single local Cumberland plain cohort, were planted in March 2013 into 12 whole-tree chambers (WTC) at the Hawkesbury Forest Experiment site near Richmond, New South Wales, Australia.
+The top soils at this site, used in the chambers, are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg^-1^ total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity.
+A root exclusion barrier extended from chamber walls to the hard layer (ca.
+1 m) and roots were allowed to grow freely below the barrier.
 Each chamber was 9 m tall, which accommodated growth of trees for 15 months.
@@
 [-@barton2010whole] and methods for this experiment in [@drake2016does].
+Coppiced *E.
+saligna* trees are present in the area between the whole-tree chambers, and these coppiced trees grew at rates similar to the trees in the chambers, thus providing ample shade.
 Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of ambient +3&deg;C (ET, Figure `r airvarfig`).
@@
 Trees were watered weekly with 70 L from March 2013 to September 2013.
-From October 2013 to the end of the experiment trees were watered every 15 days with the mean monthly rainfall amount for Richmond, NSW.
+From October 2013 to the end of the experiment **when was the end of the experiment?** trees were watered every 15 days with the mean monthly rainfall amount for Richmond, NSW.
 In February 2014 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering.
-Due to a limited range of data for the drought treatment only well-watered trees are reported, which reduces the sample size of WTC (*n*=6 to *n*=3) for the final 3 months of the experiment.
-This limited amount of data is attributed to small sample sizes as well as the difficulty in generating a sufficient CO~2~ drawdown inside the leaf cuvette, due to drought conditions, needed to accurately measure *g*~m~.
-
-\
-The top soils at this site, used in the chambers, are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg^-1^ total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity.
-A root exclusion barrier extended from chamber walls to the hard layer (ca.
-1 m) and roots were allowed to grow freely below the barrier.
+Because of the difficulty of measuring *g*~m~ in the droughted trees (the necessary CO~2~ drawdown was not achieved), we here report only data from well-watered trees, which reduces the sample size from *n*=6 to *n*=3 for the final 3 months of the experiment.
 
@@
 Leaf gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.
-At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm, height of 348&pm;15.1 cm and an leaf area of 3.9&pm;0.1 m^2^.
+At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm **SD, SE or CI?**, height of 348&pm;15.1 cm and an leaf area of 3.9&pm;0.1 m^2^.
 For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an leaf area of 6.2&pm;0.2 m^2^.
@@
 
-##Leaf gas exchange, couple with concurrent measurements of carbon isotope discrimination to estimate mesophyll conductance
+##Leaf gas exchange, coupled with concurrent measurements of carbon isotope discrimination to estimate mesophyll conductance
 Leaf gas exchange measurements were performed six times, beginning in October 2013 and monthly from December 2013 to April 2014.
@@
 
-##Biochemical parameters of photosynthesis
-Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each WTC, sampling in February 2014 when all tress were well-watered in each temperature treatment (*n*=6).
-Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011] using default parameters.
+##Photosynthetic capacity from A-C~i~ curves
+Photosynthetic CO~2~ response (A-C~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each WTC, sampling in February 2014 when all tress were well-watered in each temperature treatment (*n*=6).
+Each A-C~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+From these curves the photosynthetic parameters, *J*~max~ and *V*~cmax~, were quantified using the photosynthesis model of [@farquhar1980biochemical].
+We used the 'fitaci' function in the 'plantecophys' package in R, see [-@duursma2015] for detailed description of the fitting methods.
+This package uses non-linear regression to estimate *J*~max~, *V*~cmax~ and *R*~dark~ simultaneously.
+Because there was considerable practical difficulty in fitting A-C~c~ curves, we report standard C~i~-based *J*~max~ and *V*~cmax~, which effectively include a mesophyll conductance component.
+When fitting the A-Ci curves, we calculated $\Gamma_{*}$ with the same methods as for *g*~m~, described above (the estimated value at 25&deg;C was 38.9 $\mu$mol mol^-1^), and the combined Michaelis-Menten coefficient (*K*~m~) was estimated from [-@medlyn2002temperature] (713.4 $\mu$mol mol^-1^ at 25&deg;C).
+The temperature response parameters for *J*~max~, *V*~cmax~ and *K*~m~ are not relevant since *T*~leaf~ was kept constant at 25&deg;C.
 
@@
 ##Data analysis
-Differences in respones of dependent variables to either the warming treatment or leaf type were analysed by mixed-effects models in R [@RDevelopmentCoreTeam2011] with WTC as a random effect.
+Differences in respones of dependent variables to either the warming treatment or leaf type were analysed with linear mixed-effects models with WTC as a random effect.
 Explained variance (R^2^) of mixed models were computed as in [@nakagawa2013general], in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
 Confidence intervals (95 %) of mixed effect linear models were generated using bootstrapping methods with 999 simulations, using the bootMer function in the 'lme4' package [@bates2013lme4].
-For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package, using WTC as a random effect.
+For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package [@wood2006], using WTC as a random effect.
 All tests of statistical significance were conducted at an $\alpha$ of 0.05.
+All analyses were performed with R 3.3.0 [@RDevelopmentCoreTeam2011].
 
@@
 ##Photosynthetic capacity and leaf photosynthesis rates 
-The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from AC~i~ curves (Figure `r acifig`a).
+The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from A-C~i~ curves (Figure `r acifig`a).
 Within leaf types, no effect of the warming treatment was detected on either parameter.
@@
 
+### Acknowledgements
+We thank Craig Barton and Burhan Amiji for maintaining the Whole Tree Chambers experiment and for their outstanding technical assistance.
+We thank Sune Linder and the Swedish University for Agricultural Science for providing the whole-tree chambers.
+
+### Author Contributions
+CC contributed to the design of the research, data analysis, collection, interpretation and writing the manuscript.
+MT contributed to the design of the research, interpretation  and writing the manuscript.
+SC contributed to the design of the research, performance of the research and interpretation.
+RD contributed to the design of the research, data analysis, interpretation and writing the manuscript.
+
 #Tables
@@
 **Table 1**.
-*Eucalyptus tereticornis* leaf morphological and physiological traits of sun and shade leaves under ambient and elevated temperature treatments.
-Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and *K*~l~ values represent treatment mean (&pm; 1 standard error) across six measurement campaigns.
-Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light.
-Units of LMA and leaf N~a~ are g m^2^, *K*~l~ is mmol m^-2^ s^-1^ MPa, $\Psi$ is MPa and $\delta$^13^C is &permil;.
+Leaf morphological and physiological traits of sun and shade leaves under ambient and elevated temperature treatments for *Eucalyptus tereticornis*.
+Leaf mass per area (LMA, g m^2^), leaf nitrogen per unit area (N~a~, gN m^-2^), $\delta$^13^C (&permil;), pre-dawn leaf water potential ($\Psi$~pd~, MPa), midday leaf water potential ($\Psi$~l~, MPa) and leaf-specific hydraulic condctance (*K*~l~, mmol m^-2^ s^-1^ MPa^-1^) are given as treatment means (&pm; 1 standard error) across six measurement campaigns.
+Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from A-C~i~ curves measured in each chamber at saturating light.
 Different letters represent significant differences between leaf type and temperature treatments.
@@
 Boxplots of (a) intercellular CO~2~ concentration (C~i~), (b) CO~2~ concentration in the chloroplasts (C~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (C~i~-C~c~).
-
-### Acknowledgements
-We thank Craig Barton and Burhan Amiji for maintaining the Whole Tree Chambers experiment and for their outstanding technical assistance.
-We thank Sune Linder and the Swedish University for Agricultural Science for providing the whole-tree chambers.
-
-### Author Contributions
-CC contributed to the design of the research, data analysis, collection, interpretation and writing the manuscript.
-MT contributed to the design of the research, interpretation  and writing the manuscript.
-SC contributed to the design of the research, performance of the research and interpretation.
-RD contributed to the design of the research, data analysis, interpretation and writing the manuscript.
 

### 2016-06-21 gm_newphyt/manuscript_gmes.Rmd
Edits up to Discussion

@@
 * To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in the field in climate controlled whole tree chambers.
-* Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic biochemical parameters (Vc~max~ and J~max~), but *g*~s~ was similar.
+* Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic biochemical parameters (*V*~cmax~ and *J*~max~), but *g*~s~ was similar.
 When light intensity was temporarily increased for shade leaves to match that of sun leaves, both *g*~s~ and *g*~m~ increased, leading to increases in *A*~n~ greater than sun leaves.
@@
 Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
-Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A*~n~ if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co]; thus leaf-specific  hydraulic conductance (*K*~l~) should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects;@hubbard2001stomatal].
+Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A*~n~ if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co]; thus leaf-specific  hydraulic conductance (*K*~L~) should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects;@hubbard2001stomatal].
 
@@
 How these stresses affect the diffusion of CO~2~, through either *g*~s~ or *g*~m~, will have implications for up-scaling *A*~n~ for sun and shade leaves.
-Additionally, light-saturated rates of *A*~n~ are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
+Additionally, light-saturated rates of *A*~n~ are limited by the maximum rate of Rubisco carboxylation (*V*~cmax~) or the maximum rate of photosynthetic electron transport (*J*~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
 How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients.
@@
 Photosynthesis, *g*~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the 12 min interval.
+
 \
@@
 Full descriptions of the carbon isotope discrimination equations, with ternary corrections, are presented in Supporting Information Methods S1.
-The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck were compared as a function of C~i~:C~a~.
+The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck were compared as a function of C~i~/C~a~.
 Once *g*~m~ was calculated, the chloroplast CO~2~ partial pressure (C~c~) and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were estimated from the relationship between *g*~m~ and leaf photosynthesis rate (*A*~n~) by:
@@
 ##Leaf chemistry and hydraulic parameters
-Following gas exchange measurements each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, oven-dried and weighed.
+Following gas exchange measurements, each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, oven-dried and weighed.
 These leaves were then milled and analyzed for leaf N content and $\delta$^13^C.
@@
 Samples were flash combusted at 1000 &deg;C and converted to CO~2~ and N~2~ and then subjected to stable isotope ratio mass spectrometry.
-Leaf N is reported on an area basis (N~a~, g m^-2^) and isotopic signatures of dry matter are reported relative to standard Vienna Pee Dee Belemnite.
+Leaf N is reported on an area basis (*N*~a~, g m^-2^) and isotopic signatures of dry matter are reported relative to standard Vienna Pee Dee Belemnite.
 
@@
 The leaf closest to the leaf used for gas exchange was sampled for measurement of $\Psi_{pd}$ before sunrise on the same day as gas exchange measurements.
-Following the gas exchange measurements, leaves used for gas exchange were immediately sampled for midday ($\Psi_l$) leaf water potentials.
+Following the gas exchange measurements, leaves used for gas exchange were immediately sampled for midday ($\Psi_L$) leaf water potentials.
 All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
-Leaf water potentials and transpiration (E, mmol m^-2^ s^-1^) from gas exchange in the leaf cuvette were then used to calculate leaf-specific hydraulic conductance (*K*~l~, mmol m^-2^ s^-1^ MPa) with the equation:
+Leaf water potentials and transpiration (*E*~L~, mmol m^-2^ s^-1^) from gas exchange in the leaf cuvette were then used to calculate leaf-specific hydraulic conductance (*K*~L~, mmol m^-2^ s^-1^ MPa^-1^) with the equation:
 
@@
 (3)
-$$K_l = \frac{E}{\Psi_{pd}~-\Psi_l}$$
+$$K_L = \frac{E_L}{\Psi_{pd}~-\Psi_L}$$
 
-Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A*~n~ divided by E.
-The g~1~ parameter, describing plant water-use strategy, was estimated from ITE vs.
+Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A*~n~ divided by *E*~L~.
+The *g*~1~ parameter, describing plant water-use strategy, was estimated from ITE vs.
 VPD response curves by fitting a rearranged optimal *g*~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near].
-Small values of g~1~ indicate that transpiration is comparatively costly in C terms and reflecting conservative water use (increaed water-use efficiency), whereas large values imply a lower C cost and decreased water-use efficiency [@lin2015optimal].
+Small values of *g*~1~ indicate that transpiration is comparatively costly in C terms and reflecting conservative water use (increaed water-use efficiency), whereas large values imply a lower C cost and decreased water-use efficiency [@lin2015optimal].
 
 ##Data analysis
-Differences in respones of dependent variables to either the warming treatment or leaf type were analysed with linear mixed-effects models with WTC as a random effect.
-Explained variance (R^2^) of mixed models were computed as in [@nakagawa2013general], in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
+Differences in respones of dependent variables to either the warming treatment or leaf type were analysed with linear mixed-effects models with WTC as a random effect, and leaf type (sun, shade or shade in the sun) and warming treatment as fixed effects **(and the interaction??)**.
+Explained variance (R^2^) of mixed models were computed as in Nakagawa & Schielzeth [-@nakagawa2013general], in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
 Confidence intervals (95 %) of mixed effect linear models were generated using bootstrapping methods with 999 simulations, using the bootMer function in the 'lme4' package [@bates2013lme4].
@@
 Across six measurement campaigns over the 7 month period, PPFD was reduced on average by >75% in the shade (Figure `r parfig`).
-Leaf-specific hydraulic conductance (*K*~l~) was similar across sun and shade leaves (Table `r wtc3tab1`).
-This was because neither leaf water potentials ($\Psi_{pd}$ and $\Psi_l$, Table `r wtc3tab1`) nor transpiration rates (E, Table `r wtc3tab2`) differed between leaf types.
-Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table `r wtc3tab1`).
+Leaf-specific hydraulic conductance (*K*~L~) was similar across sun and shade leaves (Table `r wtc3tab1`).
+This was because neither leaf water potentials ($\Psi_{pd}$ and $\Psi_L$, Table `r wtc3tab1`) nor transpiration rates (E, Table `r wtc3tab2`) differed between leaf types.
+Leaf *N*~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table `r wtc3tab1`).
 Leaf mass per area (LMA) was not different between leaf types (Table `r wtc3tab1`).
-No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, *K*~l~, E, N~a~ or LMA either within or across leaf types (P > 0.05).
+No effect of the warming treatment **was interaction tested??** was detected with  PPFD, $\Psi_{pd}$, $\Psi_L$, *K*~L~, *E*~L~, *N*~a~ or LMA either within or across leaf types (P > 0.05).
 
 ##Photosynthetic capacity and leaf photosynthesis rates 
-The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from A-C~i~ curves (Figure `r acifig`a).
+The photosynthetic parameters *J*~max~ and *V*~cmax~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from A-C~i~ curves (Figure `r acifig`a).
 Within leaf types, no effect of the warming treatment was detected on either parameter.
-Among the sampled leaves, Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
+Among the sampled leaves, *V*~cmax~ was positively related to leaf *N*~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
 
@@
 Mean *A*~n~ was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment and temperatures (Table `r wtc3tab2`).
-Additionally, leaf N~a~ was positively related to *A*~n~ across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
-Following an increase in light intensity to match high-light conditions, *A*~n~ of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table `r wtc3tab2`).
+Additionally, *A*~n~ was positively related to *N*~a~ across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
+Following an increase in light intensity to match high-light conditions ('sunfleck simulations'), *A*~n~ of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table `r wtc3tab2`).
 No effect of the warming treatment was detected on *A*~n~ of sun leaves measured at high light or shade leaves at either low or high light.
@@
 Photosynthesis was positively correlated with *g*~s~ in all leaves measured under high light conditions, however, *g*~s~ and *A*~n~ were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
-Following increased PPFD, *g*~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
+In high light, *g*~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
 No effect of the warming treatment was detected on *g*~s~ within or across leaf types.
@@
 \
-Measured under ambient light and temperature, leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001).
-Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was therefore significantly lower than sun leaves (P < 0.001).
-Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment compared to ambient, but no warming effect was detected in shade leaves measured at low or high light (Table `r wtc3tab2`).
-The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves measured at low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
-For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to the fitted response curves from the optimal ITE model with a specified g~1~ value for each leaf type and treatment (Figure `r waterusefig`a).
+Measured under ambient light and temperature, leaf instantaneous transpiration efficiency (ITE) was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001).
+Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was therefore still significantly lower than sun leaves (P < 0.001).
+ITE in sun leaves was reduced in the warming treatment compared to ambient, but no warming effect was detected in shade leaves measured at low or high light (Table `r wtc3tab2`).
+The mean estimated *g*~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves measured at low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
+For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to the fitted response curves from the optimal ITE model with a specified *g*~1~ value for each leaf type and treatment (Figure `r waterusefig`a).
 Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).
@@
 \
-Bulk-leaf $\delta$^13^C, as an index of integrated water-use efficiency [@marshall2007sources], was significantly lower in shade leaves compared to sun leaves by ca.
+Bulk-leaf $\delta$^13^C was significantly lower in shade leaves compared to sun leaves by ca.
 2&permil; (Table `r wtc3tab1`).
 No effects of the warming treatment on leaf $\delta$^13^C were detected.
-Leaf $\delta$^13^C and N~a~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C (higher water-use efficiency) and higher N investment in sun leaves than shade leaves.
+Leaf $\delta$^13^C and *N*~a~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C and higher N investment in sun leaves than shade leaves.
 
 ##Leaf carbon isotope discrimination and mesophyll conductance
-The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for both leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
+The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~/C~a~ for both leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
 Carbon isotope discrimination associated with *g*~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
-The remainder $\Delta_o$ consists of the contributions of *g*~s~, respiration and photorespiration to discrimination.
+The remainder of the variation in $\Delta_o$ consists of the contributions of *g*~s~, respiration and photorespiration to discrimination.
 
 \
-Mean *g*~m~ was higher in sun compared to shade leaves (+27%) under their local light environment and temperatures (P < 0.001).
+Mean *g*~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
 Following the increase in PPFD from low to high light on the same leaf, *g*~m~ values of shade leaves increased on average 55% after approximately 25 min (Figure `r gmshadefig`).
@@
 ##Resource distribution and photosynthetic capacity
-The allocation of N~a~ constrains *A*~n~ and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
+The allocation of *N*~a~ constrains *A*~n~ and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
 Decreasing light availability should decrease the investment of N into photosynthetic enzymes within a canopy [@mooney1979environmental].
-As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical].
-Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and *A*~n~ were all reduced in shade leaves.
+As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters *V*~cmax~ and *J*~max~ [@farquhar1980biochemical].
+Our data agree with these conventional conclusions as the distribution of *N*~a~, both measures of photosynthetic capacity and *A*~n~ were all reduced in shade leaves.
 Leaf mass per area, however, was not different between sun and shade leaves.
 This could be due to leaf formation under comparable light conditions or possible unmeasured differences in total non-structural carbohydrates contents between leaf types that masked differences in leaf thickness or density.
+
+`http://www.sciencedirect.com/science/article/pii/S0378112709003661`
 
@@
 Using a two-leaf model, Peltoniemi et al.
-[-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if *K*~l~ is also optimally distributed.
-In this study, variation in leaf N distribution and *A*~n~ rates were not associated with subsequent changes in *K*~l~ between sun and shade leaves.
+[-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if *K*~L~ is also optimally distributed.
+In this study, variation in leaf N distribution and *A*~n~ rates were not associated with subsequent changes in *K*~L~ between sun and shade leaves.
 Thus, no direct relationship between water supply capacity and N distribution or *A*~n~ within the canopy were detected.
@@
 From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A*~n~ in sun leaves was characterized by low *g*~s~ and low C~i~.
-Consistently lower leaf $\delta$^13^C in shade leaves also suggests that observed higher C~i~ and C~c~ in shade leaves was likely prevalent long term (Figure `r deltafig`).
+Consistently lower leaf $\delta$^13^C in shade leaves also suggests that observed higher C~i~ and C~c~ in shade leaves was likely prevalent long term (Figure `r deltafig`) [@marshall2007sources], despite the observation that *g*~m~ accounted for nearly 70% of the variation in $\Delta$~o~.
 Relative to the differences in *A*~n~ between leaf types, higher *g*~s~ in shade leaves appears to result in inefficient water use.
-As whole canopy C gain integrates the efficiency of all leaves, this raises the question of why shade leaves maintained a lower ITE compared to sun leaves.
+As whole canopy C gain integrates the efficiency of all leaves, this raises the question of why shade leaves maintain a lower water-use efficiency compared to sun leaves.
 
@@
 Leaf morphological and physiological traits of sun and shade leaves under ambient and elevated temperature treatments for *Eucalyptus tereticornis*.
-Leaf mass per area (LMA, g m^2^), leaf nitrogen per unit area (N~a~, gN m^-2^), $\delta$^13^C (&permil;), pre-dawn leaf water potential ($\Psi$~pd~, MPa), midday leaf water potential ($\Psi$~l~, MPa) and leaf-specific hydraulic condctance (*K*~l~, mmol m^-2^ s^-1^ MPa^-1^) are given as treatment means (&pm; 1 standard error) across six measurement campaigns.
-Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from A-C~i~ curves measured in each chamber at saturating light.
+Leaf mass per area (LMA, g m^2^), leaf nitrogen per unit area (*N*~a~, gN m^-2^), $\delta$^13^C (&permil;), pre-dawn leaf water potential ($\Psi$~pd~, MPa), midday leaf water potential ($\Psi$~l~, MPa) and leaf-specific hydraulic condctance (*K*~L~, mmol m^-2^ s^-1^ MPa^-1^) are given as treatment means (&pm; 1 standard error) across six measurement campaigns.
+Values of *V*~cmax~ and *J*~max~ are treatment mean (&pm; 1 standard error) from A-C~i~ curves measured in each chamber at saturating light.
 Different letters represent significant differences between leaf type and temperature treatments.
@@
 dust(resourcetab2)%>%
-  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K",paste0("$\\Psi$","~pre~"),paste0("$\\Psi$","~l~"), paste0("$\\delta$","^13^C")) %>%
+  sprinkle_colnames("Leaf", "Temperature", "LMA", "*N*~a~", "*V*~cmax~", "*J*~max~","K",paste0("$\\Psi$","~pre~"),paste0("$\\Psi$","~l~"), paste0("$\\delta$","^13^C")) %>%
   sprinkle(rows= 1, part="head",bold=TRUE)%>%
@@
 ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-(b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light.
-(c) The relationship between *A*~n~ and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
+(b) The relationship between *V*~cmax~ and mean leaf *N*~a~ for each chamber, including sun leaves and shade leaves at low light.
+(c) The relationship between *A*~n~ and leaf *N*~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
 For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
@@
 (a) The relationship between instantaneous transpiration efficiency (ITE) and VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
-(b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light.
-For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment.
+(b) The relationship between leaf $\delta$^13^C and leaf *N*~a~ for sun leaves at high light and shade leaves at low light.
+For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a *g*~1~ value for each leaf type and treatment.
 For (b) the dashed line represents the significant linear model fit across all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.

### 2016-06-23 gm_newphyt/manuscript_gmes.Rmd
More edits discussion and figure captions; split supp info into standalone doc; preloading stuff into separate script

@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
-# Two useful packages for printing numbers, accessing content of objects.
-library(broom)
-library(reporttools)
-
 # Set Flag to include figures in text or not.
-includeFigs <- TRUE
+includeFigs <- FALSE
 ```
 
-```{r install packages, echo=FALSE, message=FALSE, warning=FALSE,results="hide"}
-r <- require(pacman)
-if(!r)install.packages("pacman")
-pacman::p_load(lubridate, doBy, plyr,plotrix, wesanderson, plantecophys, scales,reshape, pixiedust)
-```
+```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
 
-```{r global data/sourcing , include=FALSE}
-source("functions and packages/functions.R")
-source("master_scripts/plot_objects.R")
+source("gm_newphyt/manuscript_preload.R")
 
-##read table ready dataframes
-resourcetab <- read.csv("master_scripts/resource_table.csv", stringsAsFactors = FALSE)
-getab <- read.csv("master_scripts/ge_table.csv", stringsAsFactors = FALSE)
-pval <- c(0.781, 0.001, 0.028, 0.002, 0.973, 0.3486, 0.6385, 0.001)
-pval2 <- c(0.001,	0.001,	0.001,	0.001,	0.001,	0.005,	0.001,	0.001)
+# Make figure PDFs.
+source("gm_newphyt/make_figures.R")
+
 ```
@@
 Trees were watered weekly with 70 L from March 2013 to September 2013.
-From October 2013 to the end of the experiment **when was the end of the experiment?** trees were watered every 15 days with the mean monthly rainfall amount for Richmond, NSW.
+From October 2013 to the end of the experiment trees were watered every 15 days with the mean monthly rainfall amount for Richmond, NSW.
 In February 2014 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering.
@@
 The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDL on each measurement cycle.
-Photosynthesis, *g*~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the 12 min interval.
+Photosynthesis, *g*~s~, transpiration, VPD and intercellular [CO~2~] (*C*~i~) were auto-logged every 15 s for each gas exchange system over the 12 min interval.
 
@@
 
-where $\Delta$~o~ is the observed discrimination and $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, *g*~m~, respiration and photorespiration, respectively.
+where $\Delta$~o~ is the observed discrimination and $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if *C*~i~ = *C*~c~, *g*~m~, respiration and photorespiration, respectively.
 For this study, the CO~2~ compensation point ($\Gamma_{*}$) and respiration during the day (R~d~) parameters originally derived for tobacco plants [@von1994kinetics] were replaced with parameters derived for *Eucalyptus globulus* from Crous et al.
@@
 Full descriptions of the carbon isotope discrimination equations, with ternary corrections, are presented in Supporting Information Methods S1.
-The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck were compared as a function of C~i~/C~a~.
-Once *g*~m~ was calculated, the chloroplast CO~2~ partial pressure (C~c~) and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were estimated from the relationship between *g*~m~ and leaf photosynthesis rate (*A*~n~) by:
+The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck were compared as a function of *C*~i~/C~a~.
+Once *g*~m~ was calculated, the chloroplast CO~2~ partial pressure (*C*~c~) and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were estimated from the relationship between *g*~m~ and leaf photosynthesis rate (*A*~n~) by:
 
@@
 
-##Photosynthetic capacity from A-C~i~ curves
-Photosynthetic CO~2~ response (A-C~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each WTC, sampling in February 2014 when all tress were well-watered in each temperature treatment (*n*=6).
-Each A-C~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+##Photosynthetic capacity from A-*C*~i~ curves
+Photosynthetic CO~2~ response (A-*C*~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each WTC, sampling in February 2014 when all tress were well-watered in each temperature treatment (*n*=6).
+Each A-*C*~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
 From these curves the photosynthetic parameters, *J*~max~ and *V*~cmax~, were quantified using the photosynthesis model of [@farquhar1980biochemical].
@@
 This package uses non-linear regression to estimate *J*~max~, *V*~cmax~ and *R*~dark~ simultaneously.
-Because there was considerable practical difficulty in fitting A-C~c~ curves, we report standard C~i~-based *J*~max~ and *V*~cmax~, which effectively include a mesophyll conductance component.
+Because there was considerable practical difficulty in fitting A-*C*~c~ curves, we report standard *C*~i~-based *J*~max~ and *V*~cmax~, which effectively include a mesophyll conductance component.
 When fitting the A-Ci curves, we calculated $\Gamma_{*}$ with the same methods as for *g*~m~, described above (the estimated value at 25&deg;C was 38.9 $\mu$mol mol^-1^), and the combined Michaelis-Menten coefficient (*K*~m~) was estimated from [-@medlyn2002temperature] (713.4 $\mu$mol mol^-1^ at 25&deg;C).
@@
 ##Photosynthetic capacity and leaf photosynthesis rates 
-The photosynthetic parameters *J*~max~ and *V*~cmax~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from A-C~i~ curves (Figure `r acifig`a).
+The photosynthetic parameters *J*~max~ and *V*~cmax~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from A-*C*~i~ curves (Figure `r acifig`a).
 Within leaf types, no effect of the warming treatment was detected on either parameter.
@@
 ##Leaf carbon isotope discrimination and mesophyll conductance
-The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~/C~a~ for both leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
+The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with *C*~i~/C~a~ for both leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
 Carbon isotope discrimination associated with *g*~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
@@
 ##Variation in intercellular and chloroplastic CO~2~ concentrations
-Higher *g*~s~ in shade leaves under low and high light led to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
+Higher *g*~s~ in shade leaves under low and high light led to significant increases in *C*~i~ compared to sun leaves (Figure `r ciccfig`a).
 The chloroplast CO~2~ partial pressure was comparable between shade leaves when measured at both low and high light conditions (Figure `r ciccfig`c).
-In sun leaves C~c~ was significantly lower than shade leaves, consistent with a lower C~i~.
-The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between *g*~m~ and *A*~n~ [@caemmerer2014temperature].
+In sun leaves *C*~c~ was significantly lower than shade leaves, consistent with a lower *C*~i~.
+The drawdown of CO~2~ from intercellular spaces to the chloroplast, *C*~i~-*C*~c~, measures the coordination between *g*~m~ and *A*~n~ [@caemmerer2014temperature].
 This drawdown was similar between sun and shade leaves measured at their local light and temperature environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
 This was the result of the proportional relationship between *g*~m~ and *A*~n~ across all leaves.
-The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
+The CO~2~ drawdown from C~a~ to *C*~i~ and *C*~i~ to *C*~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
 
@@
 Here we show that *A*~n~ in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
-Although shade leaves in lower light environments exhibited relatively high *g*~s~ and C~i~, it was concomitant increases in *g*~m~ under periods of high light availability that enabled up-regulation of *A*~n~.
+Although shade leaves in lower light environments exhibited relatively high *g*~s~ and *C*~i~, it was concomitant increases in *g*~m~ under periods of high light availability that enabled up-regulation of *A*~n~.
 Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
@@
 Our data agree with these conventional conclusions as the distribution of *N*~a~, both measures of photosynthetic capacity and *A*~n~ were all reduced in shade leaves.
-Leaf mass per area, however, was not different between sun and shade leaves.
-This could be due to leaf formation under comparable light conditions or possible unmeasured differences in total non-structural carbohydrates contents between leaf types that masked differences in leaf thickness or density.
-
-`http://www.sciencedirect.com/science/article/pii/S0378112709003661`
+Leaf mass per area (LMA), however, was not different between sun and shade leaves, consistent with small differences reported in a clonal *Eucalyptus* plantation [@nouvellon2010].
+Similar LMA in sun and shade leaves could be the result of leaf formation under comparable light conditions or possible unmeasured differences in total non-structural carbohydrates contents between leaf types that masked differences in leaf thickness or density.
 
@@
 
+##Physiological behaviour of sun and shade leaves
 \
 Unexpected higher *g*~s~ in shade leaves compared to sun leaves led to decreased ITE in shade leaves throughout the experiment.
-From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A*~n~ in sun leaves was characterized by low *g*~s~ and low C~i~.
-Consistently lower leaf $\delta$^13^C in shade leaves also suggests that observed higher C~i~ and C~c~ in shade leaves was likely prevalent long term (Figure `r deltafig`) [@marshall2007sources], despite the observation that *g*~m~ accounted for nearly 70% of the variation in $\Delta$~o~.
-Relative to the differences in *A*~n~ between leaf types, higher *g*~s~ in shade leaves appears to result in inefficient water use.
+Consistently lower leaf $\delta$^13^C in shade leaves also suggests that observed higher *C*~i~ and *C*~c~ in shade leaves was prevalent long term (Figure `r deltafig`) [@marshall2007sources], despite the observation that *g*~m~ accounted for nearly 70% of the variation in $\Delta$~o~.
+Higher *g*~s~ in shade leaves relative to the differences in *A*~n~, appears to result in inefficient water use as more water is used per unit photosynthesis.
 As whole canopy C gain integrates the efficiency of all leaves, this raises the question of why shade leaves maintain a lower water-use efficiency compared to sun leaves.
 
-##Physiological behaviour of sun and shade leaves
-The pattern of inefficient water use in shade leaves is important as we hypothesized that *g*~s~ and *A*~n~ would be proportional across sun and shade leaves.
+\
+The observed pattern of inefficient water use in shade leaves leads to the rejection of the hypothesis that *g*~s~ and *A*~n~ are proportional across sun and shade leaves.
 In sun leaves, *A*~n~ and *g*~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
@@
 \
-It is possible that reducing stomatal response time, by sustaining higher *g*~s~, is a strategy to take advantage of high light quickly in shade leaves [@tausz2005dynamic].
-Evidence from this study supports this hypothesis, as shade leaves increased *A*~n~ to values equal or even outperforming sun leaves when subjected to a brief period of identical high-light intensity.
-Transpiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck induced thermal load [@schymanski2013stomatal].
-This is because rapid increases in leaf temperature with sunflecks have been shown to inhibit C gain [@leakey2003high].
-However, this response likely occurs at very high air temperatures and may not explain the observed *g*~s~ in shade leaves across the large natural seasonal range of temperatures included in this study.
-How prevalent each of these strategies are within tree canopies is still unknown, as empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants, often in deep shade [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution].
-though leaves within a forest canopy exhibited a similar uncoupling of *A*~n~ and *g*~s~ across a light gradient [@tjoelker1995interaction].
+It is possible that higher *g*~s~ in shade leaves is a strategy to take advantage of high light quickly in shade leaves, avoiding the delay with stomatal opening [@tausz2005dynamic].
+Our results support this hypothesis, as shade leaves increased *A*~n~ to values equal or even outperforming sun leaves when subjected to a brief period of identical high-light intensity.
+Transpiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck-induced rapid increases in leaf temperature [@schymanski2013stomatal], which have been shown to inhibit C gain [@leakey2003high].
+However, this response likely occurs at very high air temperatures and may not explain the observed *g*~s~ in shade leaves across the large natural seasonal range of temperatures included in our study.
+How prevalent each of these strategies are within tree canopies is still unknown, as empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants, often in deep shade [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution], though leaves within a forest canopy exhibited a similar uncoupling of *A*~n~ and *g*~s~ across a light gradient [@tjoelker1995interaction].
 Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy [see @de1997simple].
@@
 \
-We found that *A*~n~ and *g*~m~ scaled positively across leaf types and, unexpectedly, increased proportionately after leaf stability (~25 min) when light intensity was increased in shade leaves.
-Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of *g*~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
-This provides a potential explanation for the observed increases in *g*~m~, but the impacts of aquaporins on *g*~m~ are yet to be tested in leaves of tree species.
-Our findings support growing evidence that *g*~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
+We found that *A*~n~ and *g*~m~ scaled positively across leaf types (Fig.
+3b) and, unexpectedly, increased in shade leaves when light intensity was increased to values similar in sun leaves.
+Possible mechanisms for the rapid (~25 min) response of *g*~m~ to light intensity include regulation of aquaporins, which can facilitate increases in the CO~2~ permeability of the cell membranes [@hanba2004overexpression; @heinen2009role; @li2014plant] or movement of chloroplasts to facilitate CO~2~ diffusion [@tholen2008].
+Our findings support growing evidence that *g*~m~ is highly variable and can respond to environmental variables [@flexas2007rapid; @flexas2008mesophyll].
 Here we provide empirical data showing *g*~m~ not only varies within a canopy, but that the up-regulation of *g*~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.
@@
 If the valuation of sunflecks as a C resource is large enough, then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan or across the entire canopy when considering both sun and shade leaf types.
-For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to accommodate the cost of decreases in ITE.
+For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to accommodate the cost of increased water use in the shade.
 However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.
@@
 Here we show that dynamic physiological responses of shade leaves to altered light environments have important implications for up-scaling leaf level measurements to the canopy.
-Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
+Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves that determines C gain.
 These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have important implications for process-based models that predict canopy C gain from rates of leaf photosynthesis.
-Furthermore, the dynamic nature of *g*~m~ cannot be simply parameterized in tree growth models and possibly should be excluded until it can be represented properly.
-Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of *g*~m~ to opportunistically increase CO~2~ drawdown.
-To improve our ability to predict whole canopy C gain, future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.
+Furthermore, the dynamic nature of *g*~m~ cannot be simply parameterized in vegetation models, and possibly should be excluded until it can be represented properly.
+Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of *g*~m~ to respond to environmental drivers.
 
 ### Acknowledgements
-We thank Craig Barton and Burhan Amiji for maintaining the Whole Tree Chambers experiment and for their outstanding technical assistance.
+We thank Craig Barton and Burhan Amiji for maintaining the whole-tree chambers experiment and for their outstanding technical assistance.
 We thank Sune Linder and the Swedish University for Agricultural Science for providing the whole-tree chambers.
@@
 Leaf mass per area (LMA, g m^2^), leaf nitrogen per unit area (*N*~a~, gN m^-2^), $\delta$^13^C (&permil;), pre-dawn leaf water potential ($\Psi$~pd~, MPa), midday leaf water potential ($\Psi$~l~, MPa) and leaf-specific hydraulic condctance (*K*~L~, mmol m^-2^ s^-1^ MPa^-1^) are given as treatment means (&pm; 1 standard error) across six measurement campaigns.
-Values of *V*~cmax~ and *J*~max~ are treatment mean (&pm; 1 standard error) from A-C~i~ curves measured in each chamber at saturating light.
-Different letters represent significant differences between leaf type and temperature treatments.
+Values of *V*~cmax~ and *J*~max~ are treatment mean (&pm; 1 standard error) from A-*C*~i~ curves measured in each chamber at saturating light.
+Different letters represent significant differences between leaf type and temperature treatments (based on a Tukey test).
 The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.
@@
 **Table 2**.
-*Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
-Each value reflects the mean (&pm; 1 standard error) for each treatment across six gas exchange campaigns.
-Units for *A*~n~ and E are $\mu$mol m^-2^ s^-1^, for *g*~s~ are mol m^-2^ s^-1^, for *g*~m~ are mol m^-2^ s^-1^ bar^-1^ and for VPD is kPa.
-Different letters represent significant differences between leaf type, light environment and temperature treatments.
+Leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments for *Eucalyptus tereticornis*.
+Leaf net photosynthesis rate (*A*~n~, $\mu$mol m^-2^ s^-1^), leaf transpiration rate (*E*~L~, mmol m^-2^ s^-1^), stomatal conductance (*g*~s~, mol m^-2^ s^-1^), *g*~m~ (mol m^-2^ s^-1^ bar^-1^), intercellular CO~2~ concentration (*C*~i~, $\mu$mol mol^-1^), chloroplastic CO~2~ concentration (*C*~c~, $\mu$mol mol^-1^) and leaf-to-air vapour pressure deficit (VPD, kPa) are given as the mean (&pm; 1 standard error) for each treatment across six gas exchange campaigns.
+Different letters represent significant differences between leaf type, light environment and temperature treatments (based on a Tukey test).
 The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.
@@
 dust(getab2)%>%
-  sprinkle_colnames("Leaf", "Light","Temperature", "*A*~n~", "*g*~s~", "*g*~m~", "ITE","E","VPD","C~i~", "C~c~") %>%
+  sprinkle_colnames("Leaf", "Light","Temperature", "*A*~n~", "*g*~s~", "*g*~m~", "ITE","*E*~L~","VPD","*C*~i~", "*C*~c~") %>%
   sprinkle(rows= 1, part="head",bold=TRUE)%>%
@@
 
-```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png",message=FALSE, dpi=600, dev.args=list(type="cairo")}
+```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png",message=FALSE, dpi=600, dev.args=list(type="cairo"), eval=includeFigs}
 source("master_scripts/par2.R")
@@
 **Figure 1**.
-Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014.
-Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf.
-Each date represents the starting date for each measurement campaign.
-Points represent the mean (&pm; 1 standard error) daily maximum air temperature during each campaign period.
+Environmental conditions during the six measurement campaigns.
+Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014 (as photosynthetic photon flux density, PPFD).
+Means &pm; 1 standard error represent PPFD measured with a ceptometer at the height of each selected leaf in the canopy.
+Points connected by a line represent the mean (&pm; 1 standard error) daily maximum air temperature during each campaign period.
+Each date represents the starting date for each measurement campaign; each campaign was completed in **how many days???*.
 
 \
-```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600,message=FALSE, dev.args=list(type="cairo")}
+```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600,message=FALSE, dev.args=list(type="cairo"), eval=includeFigs}
  source("master_scripts/aci_vcmaxnitro.R")
@@
 **Figure 2**.
-(a) ACi curves for sun and shade leaves grown under elevated (ET) and ambient (AT) temperature treatments.
-ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-(b) The relationship between *V*~cmax~ and mean leaf *N*~a~ for each chamber, including sun leaves and shade leaves at low light.
-(c) The relationship between *A*~n~ and leaf *N*~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
+Photosynthetic capacity, photosynthetis rate and leaf nitrogen for sun and shade leaves.
+(a) A-*C*~i~ curves for sun and shade leaves grown under elevated (ET) and ambient (AT) temperature treatments.
+A-*C*~i~ curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+(b) The relationship between *V*~cmax~ and mean leaf N per area (*N*~a~) for each chamber, including sun leaves and shade leaves at low light.
+(c) The relationship between leaf net photosynthesis rate (*A*~n~) and leaf *N*~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
 For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
 \
-```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo"), eval=includeFigs}
 source("master_scripts/Agmgs.R")
@@
 **Figure 3**.
-The relationship between *A*~n~ to *g*~s~ (a) and *g*~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments across six measurement campaigns.
-Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
+Relationships between leaf net photosynthesis rate (*A*~n~), stomatal conductance (*g*~s~, panel (a)) and mesophyll conductance (*g*~m~, panel (b)) for sun leaves and shade leaves in low or high light.
+Sun leaves were measured at high light, shade leaves were measured at both low and high light.
+Data are cobined across elevated (ET, triangles) and ambient (ET, circles) temperature treatments, and six measurement campaigns.
+Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b), fitted across all data but accounting for random effects (see Methods).
 Grey areas are 95% confidence intervals for the mean.
 \
-```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600,message=FALSE, dev.args=list(type="cairo")}
+```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600,message=FALSE, dev.args=list(type="cairo"), eval=includeFigs}
 source("master_scripts/physiology_boxplots.R")
@@
 \
-**Figure 4**. Box plots of measured *g*~s~ (a), *g*~m~ (b) and *A*~n~ (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
+**Figure 4**.
+Box plots of stomatal conductance (*g*~s~) (a), mesophyll conductance (*g*~m~) (b) and leaf net photosynthesis rate (*A*~n~) (c) of sun leaves and shade leaves at low and high light pooled across six measurement dates and both temperature treatments.
 \
-```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=600}
+```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=600, eval=includeFigs}
 source("master_scripts/wateruse.R")
@@
 **Figure 5**.
-(a) The relationship between instantaneous transpiration efficiency (ITE) and VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
-(b) The relationship between leaf $\delta$^13^C and leaf *N*~a~ for sun leaves at high light and shade leaves at low light.
-For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a *g*~1~ value for each leaf type and treatment.
-For (b) the dashed line represents the significant linear model fit across all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
+Two measures of water-use efficiency related to vapour pressure deficit and leaf nitrogen per area (*N*~a~).
+(a) Instantaneous transpiration efficiency (ITE) declines with VPD, and is lower for shade leaves (either when measured in high or low light).
+VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a *g*~1~ value for each leaf type and temperature treatment (AT, ambient temperature; ET, elevated temperature).
+(b) The relationship between leaf $\delta$^13^C (higher values indicate higher water-use efficiency) and leaf *N*~a~ for sun leaves at high light and shade leaves at low light.
+The dashed line represents the significant linear model fit across all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
 \
-```{r gmshade, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600,message=FALSE, dev.args=list(type="cairo")}
+```{r gmshade, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600,message=FALSE, dev.args=list(type="cairo"), eval=includeFigs}
 source("master_scripts/gmes_shadeleaves_figure.R")
@@
 **Figure 6**.
-The change in measured *g*~m~ for individual shade leaves following an increase in PPFD to match the light environment of full sun leaves in the same tree measured across six measurement dates.
-Measurements of *g*~m~ were recorded once CO~2~ and water vapour fluxes were stable in the leaf cuvette, which took approximately 25 minutes after light intensity was increased.
+Mesophyll conductance (*g*~m~) measured on shade leaves in low light, and in high light (on the same leaves).
+After measurement in the shade, the photosynthetic photon flux density (PPFD) was increased to match the light environment of full sun leaves in the same tree.
+All six measurement dates are included, and symbols reflect the temperature treatment (AT, ambient temperature; ET, elevated temperature).
+Measurements of *g*~m~ were recorded after CO~2~ and water vapour fluxes were stable in the leaf cuvette, which took approximately 25 minutes after light intensity was increased.
 The dashed line is the 1:1 relationship.
 \
-```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
+```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo"), eval=includeFigs}
 source("master_scripts/cicc_boxplots.R")
@@
 **Figure 7**.
-Boxplots of (a) intercellular CO~2~ concentration (C~i~), (b) CO~2~ concentration in the chloroplasts (C~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (C~i~-C~c~).
-
-#Supporting Information
-
-##Figures
-
-```{r airvars, fig.width=7, fig.height=8.8, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/air_vars.R")
-```
-\
-**Figure S1**. Daily maximum and minimum temperature (a), daily maximum VPD (b) and total daily PPFD (c) for each chamber across the experiment duration.
-\
-```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/delta_cica.R")
-```  
-\
-**Figure S2**.
-Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis ($\Delta_o$) and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
-The solid line represents the theoretical line for C3 plants from Evans et al.
-[-@evans1986carbon].
-\
-```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/gasex_temp3.R")
-```  
-\
-**Figure S3**.
-Response of *A*~n~ (a), *g*~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
-Shaded symbols represent each monthly measurement campaign.
-Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
-All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
-Weak negative relationships with *g*~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (R^2^ = 0.16 and 0.08, respectively).
-\
-```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/gasex_temp4.R")
-```  
-\
-**Figure S4**.
-Response of VPD (a), *g*~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
-Shaded symbols represents each monthly measurement campaign.
-Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
-All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
-Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (R^2^ = 0.73, 0.58 and 0.72, respectively).
-
-##Methods S1
-
-###Description of the calculation of *g*~m~ from carbon isotope discrimination during C3 photosynthesis
-\
-Using online C isotope discrimination measurements, the difference between the observed discrimination and what is predicted for light saturated gas exchange is proportional to *g*~m~ [@griffiths2013mesophyll].
-First, leaf C isotope discrimination was calculated by comparing the isotopic composition of the reference gas entering the leaf cuvette ($\delta$^13^C~e~) with the sample gas ($\delta$^13^C~o~) such that:
-
-\
-(1)
-$$\delta = \left(\frac{R_s}{R_{stnd}} -1\right)1000$$
-
-where R~s~ is the isotopic ratio of the sample and R~stnd~ is the isotopic ratio of the standard Vienna Pee Dee Belemnite (VPDP).
-Next, the observed discrimination ($\Delta$~o~) is calculated from Evans et al.
-[-@evans1986carbon]:
-
-\
-(2)
-$$\Delta_{o} = \frac{\xi(\delta^13C_o-\delta^13C_e)}{1000+\delta^13C_e-\xi(\delta^13C_o-\delta^13C_e)}$$
-
-where:
-
-\
-(3)
-$$\xi = \frac{C_{e}}{C_{e}-C_{o}}$$
-
-and $\xi$ is the ratio of the CO~2~ entering the well mixed leaf cuvette to the CO~2~ draw down in the gas stream by the leaf.
-
-\
-Second, C isotope discrimination during C~3~ photosynthesis ($\Delta$) is the resultant discrimination from CO~2~ diffusion from the atmosphere to the site of carboxylation, consisting of a series of fractionation steps described in Evans et al.
-[-@evans1986carbon].
-In this experiment, a modified form of the equation presented in Evans & von Caemmerer [-@evans2013temperature] with ternary effect corrections by Farquhar & Cernusak [-@farquhar2012ternary] was used such that:
-
-\
-(4)
-$$\Delta_o = \Delta_i - \Delta_{gm} - \Delta_e - \Delta_f$$
-
-where $\Delta$~o~ is the observed discrimination and  $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, *g*~m~, respiration and photorespiration, respectively.
-The equations for each are as follows:
-
-\
-(5)
-$$
-\begin{aligned}
-\Delta_i = &\frac{1}{1-t}a'+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}
-\end{aligned}
-$$
-
-\
-(6)
-$$
-\begin{aligned}
-\Delta_{gm} = &\frac{1+t}{1-t}\left(b-a_i-\frac{eR_d}{A+R_d}\right)\frac{A}{g_mC_a}
-\end{aligned}
-$$
-
-\
-(7)
-$$
-\begin{aligned}
-\Delta_e = &\frac{1+t}{1-t}\left(\frac{eR_d}{(A+R_d)C_a}(C_i-\Gamma_*)\right)
-\end{aligned}
-$$
-
-\
-(8)
-$$
-\begin{aligned}
-\Delta_f = & \frac{1+t}{1-t}\left(f\frac{\Gamma_*}{C_a}\right)\\
-\end{aligned}
-$$
-
-where the different fractionation factors include; diffusion through water (a~i~, 1.8&permil;), Rubisco carboxylation (b, 29&permil;), the photorespiratory fractionation (f, 16.2&permil;) and the combined fractionation through the boundary layer and the stomata (a').
-a' is defined by:
-
-\
-(9)
-$$a' = \frac{a_b(C_a-C_i)+a(C_s-C_i)}{C_a-C_i}$$
-
-where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ is the fractionation from boundary layer diffusion (2.9&permil;) and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon].
-C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~).
-In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulus* from Crous et al.
-[-@crous2012light].
-The ternary effect corrections (t) are described by:
-
-\
-(10)
-$$t = \frac{(1+a')E}{2g^t_{ac}}$$
-
-where E denotes the transpiration rate and $\text{g}^t_{ac}$ is the total conductance to CO~2~ diffusion to both the boundary layer and stomatal conductance.
-
-\
-The *g*~m~ can then be calculated as:
-
-\
-(11)
-$$
-\begin{aligned}
-g_m = \left(\frac{1+t}{1-t}\left(b-a_i-\frac{eR_d}{A+R_d}\right)\frac{A}{C_a}\right) / \left(\Delta_i-\Delta_o-\Delta_e-\Delta_f\right)
-\end{aligned}
-$$
-
-Here, *g*~m~ is expressed in units of mol m^-2^ s^-1^ bar^-1^ to be consistent with Evans & von Caemmerer [-@evans2013temperature].
-As the pressure term is nearly equal to unity, suitable comparisons between *g*~m~ and *g*~s~ can be made.
+Boxplots of (a) intercellular CO~2~ concentration (*C*~i~), (b) CO~2~ concentration in the chloroplasts (*C*~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (*C*~i~-*C*~c~).
+All data across six measurement dates and both temperature treatments are included.
 

### 2016-06-23 gm_newphyt/manuscript_gmes_SuppInfo.Rmd
More edits discussion and figure captions; split supp info into standalone doc; preloading stuff into separate script

--- 2016-05-30 manuscript/manuscript_wtc3.Rmd
@@
 ---
-title: |
-  | Chapter 3 
-  | Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks
+title: Supporting information
 author: ""
@@
   word_document:
-    reference_docx: manuscriptstyle_thesis.docx
-  pdf_document: default
-  html_document:
-    number_sections: yes
-csl: tree-physiology.csl
+    reference_docx: manuscriptstyle.docx
+csl: new-phytologist.csl
 bibliography: references.bib
@@
 
-<!-- \ -->
-<!-- Courtney E. Campany^1^, Belinda Medlyn^1^, Mark Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^. -->
+Article title: Coupled response of stomatal and mesophyll conductance to light enhances photosynthesis of shade leaves under sunflecks.
 
-<!-- \ -->
-<!-- ^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia -->
-
-<!-- \ -->
-<!-- ^2^ Plant Science Division, Research School of Biology, The Australian National University, Canberra, ACT, Australia -->
-
-<!-- \ -->
-<!-- *Corresponding author*: -->
-<!-- Courtney Campany -->
-<!-- E: courtneycampany@gmail.com -->
+Authors: Courtney E. Campany, Mark G. Tjoelker, Susanne von Caemmerer, Remko A. Duursma.
 
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
-# Two useful packages for printing numbers, accessing content of objects.
-library(broom)
-library(reporttools)
-
 # Set Flag to include figures in text or not.
 includeFigs <- TRUE
-
-# This document can be compiled with the 'Knit Word' button in Rstudio, or:
-# library(rmarkdown)
-# render("manuscript.Rmd", "word_document", "manuscript.docx")
 ```
 
-```{r global data/sourcing , include=FALSE}
-source("functions and packages/functions.R")
-source("functions and packages/packages.R")
-source("master_scripts/plot_objects.R")
-library(pixiedust)
-
-##read table ready dataframes
-resourcetab <- read.csv("master_scripts/resource_table.csv", stringsAsFactors = FALSE)
-getab <- read.csv("master_scripts/ge_table.csv", stringsAsFactors = FALSE)
-pval <- c(0.781, 0.001, 0.028, 0.002, 0.973, 0.3486, 0.6385, 0.001)
-pval2 <- c(0.001,	0.001,	0.001,	0.001,	0.001,	0.005,	0.001,	0.001)
-# getab <- read.csv("master_scripts/ge_table.csv")
-
-#treatments <- read.csv("raw data/chamber_trt.csv")
+```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
+source("gm_newphyt/manuscript_preload.R")
 ```
 
-```{r, echo=FALSE}
+## Figures
 
-parfig <- "3.1"
-acifig <- "3.2"
-agmgsfig <- "3.3"
-gebarfig <- "3.4"
-waterusefig <- "3.5"
-deltafig <- "3.6"
-ciccfig <- "3.7"
+```{r airvars, fig.width=7, fig.height=8.8, echo=FALSE,warning=FALSE,dev="png", dpi=600,  eval=includeFigs, dev.args=list(type="cairo")}
+source("master_scripts/air_vars.R")
+```
+\
+**Figure S1**. Daily maximum and minimum temperature (a), daily maximum VPD (b) and total daily PPFD (c) for each chamber across the experiment duration.
+\
+```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, eval=includeFigs, dev.args=list(type="cairo")}
+source("master_scripts/delta_cica.R")
+```  
+\
+**Figure S2**.
+Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis ($\Delta_o$) and measured *C*~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
+The solid line represents the theoretical line for C3 plants from Evans et al.
+[-@evans1986carbon].
+\
+```{r gasex_temp3, fig.width=7, fig.height=9, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo"), eval=includeFigs}
+source("master_scripts/gasex_temp3.R")
+```  
+\
+**Figure S3**.
+Response of *A*~n~ (a), *g*~m~ (b) and *C*~i~-*C*~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
+Shaded symbols represent each monthly measurement campaign.
+Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
+All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
+Weak negative relationships with *g*~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (R^2^ = 0.16 and 0.08, respectively).
+\
+```{r gasex_temp4, fig.width=7, fig.height=9, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo"), eval=includeFigs}
+source("master_scripts/gasex_temp4.R")
+```  
+\
+**Figure S4**.
+Response of VPD (a), *g*~s~ (b) and C~a~-*C*~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
+Shaded symbols represents each monthly measurement campaign.
+Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
+All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
+Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (R^2^ = 0.73, 0.58 and 0.72, respectively).
 
-airvarfig <- "3.S1"
-accfig <- "3.S2"
-get1fig <- "3.S3"
-get2fig <- "3.S4"
+##Methods S1
 
-wtc3tab1 <- "3.1"
-wtc3tab2 <- "3.2"
-
-```
-
-#3.1  Abstract
-Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves.
-A lack of empirical data relating photosynthesis to leaf physiological behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
-To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
-Compared to sun leaves, shade leaves had lower *A~n~*, g~m~, Vc~max~, J~max~ and leaf nitrogen but g~s~ was similar.
-When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in *A~n~* greater than sun leaves at the same photosynthetic photon flux density.
-Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain.
-We argue that the rapid response of g~m~ with light enables shade leaves to respond quickly to sunflecks.
-Evidence that g~m~ not only varies within a canopy but can be up-regulated over short time intervals possibly represents a new mechanism underpinning leaf gas exchange responses to light.
-
-<!-- ## Key words -->
-<!-- photosynthesis, stomatal conductance, mesophyll conductance,  shade, leaf optimal behavior -->
-
-#3.2  Introduction
-Light availability is one of the most important environmental drivers of leaf carbon (C) uptake in trees.
-Predicting C uptake of forests usually involves upscaling leaf level measurements to assess whole canopy function.
-Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling based on solar irradiance problematic.
-Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor].
-Consequently, structural and functional properties of leaves are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of leaf photosynthesis (*A~n~*) [@evans1995carbon].
-To estimate whole canopy C gain it is thus necessary to account for the non-linear response of *A~n~* to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
-
+###Description of the calculation of *g*~m~ from carbon isotope discrimination during C3 photosynthesis
 \
-The distribution of resources required for *A~n~*, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients.
-As *A~n~* has a saturating response with light and maximum rates depend in part on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co].
-Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance [@evans2001photosynthetic].
-Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
-Higher rates of *A~n~* and stomatal conductance (g~s~) can only be sustained if the leaf specific hydraulic conductance (K~l~) is also large enough to avoid low leaf water potentials [@hubbard2001stomatal].
-Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A~n~* if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co], thus K~l~ should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects].
-
-\
-Rates of photosynthesis in C~3~ plants are limited by the [CO~2~] available for fixation by Rubisco within the chloroplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmosphere to the site of carboxylation [@warren2008stand].
-This drawdown consists of multiple resistance pathways to CO~2~ diffusion which include CO~2~ diffusion from the atmosphere through stomata (stomatal conductance, g~s~) and then from these intercellular air spaces into the chloroplast (mesophyll conductance, g~m~).
-Based on optimal theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize *A~n~* [@peltoniemi2012co].
-This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
-Mesophyll conductance can also impose limitations on *A~n~* as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A~n~* [@niinemets2007photosynthesis] if g~m~ constrains CO~2~ supply to the chloroplast.
-Part of the variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
-It is likely that leaf anatomical costs associated with minimizing the length, diameter and tortuosity of the g~m~ diffusion pathway are necessary to maintain a high g~m~ [@hassiotou2009influence].
-Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack of empirical data currently hinders our ability to interpreting their coupled responses to *A~n~* across sun and shade leaves.
-
-\
-Additionally, accounting for short term light fluctuations within a canopy, via sunflecks, makes assessing shade leaf behaviour difficult.
-How shade leaves utilize sunflecks for short term C gain depends on the combined response time of g~s~ and g~m~ and the underlying photosynthetic biochemistry acclimated to a low light environment [@pearcy1990sunflecks;@tausz2005dynamic].
-The utilization of sunflecks is first limited by delayed responses of stomata opening, which may take minutes, effectively limiting the maximum assimilation rate that can be achieved [@pearcy1990sunflecks; @vicomanz2011; @way2012sunflecks].
-Mesophyll conductance has been shown to respond to environmental factors (e.g.
-CO~2~, temperature or vapor pressure deficit) at timescales of minutes, possibly faster than g~s~ [@flexas2008mesophyll and references therein], yet short term response to light availability are unclear.
-For example, g~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid].
-Anatomical parameters which regulate g~m~ with changing irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative] are also unlikely to adjust during short light fluctuations.
-The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance and as a potential response to transitory light availability.
-
-\
-Climate warming may also affect the physiological behaviour of leaves within a canopy.
-This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis].
-How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for upscaling *A~n~* for sun and shade leaves.
-Additionally, light saturated rates of *A~n~* are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
-How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients.
-The impacts of warming on plant physiological processes are vast, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.
-
-\
-In this study we use *Eucalyptus tereticornis* Sm.
-trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to empirically evaluate the distribution of nitrogen and water supply and leaf physiological behaviour of sun and shade leaves.
-Our hypotheses are as follows:
-
-\
-1.	If whole tree canopies are optimized for C gain, then leaf N, hydraulic conductance and biochemical photosynthetic capacity are predicted to be higher in sun leaves compared to shade leaves.
-
-\
-2.	Stomatal conductance is proportional to *A~n~* across sun and shade leaves under similar leaf VPD and g~m~ scales positively with photosynthetic capacity.
-
-\
-3. As shade leaves were expected to develop lower biochemical photosynthetic capacity, increases in *A~n~* following sunfleck simulations were not expected to reach rates of full sun leaves.
-
-\
-4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporation demand from higher temperatures lead to stomatal closure.
-
-as higher temperatures
-
-#3.3  Materials and Methods
-
-##3.3.1 Whole tree chamber experimental design
-Twelve *Eucalyptus tereticornis* seedlings, chosen from a single local Cumberland plain cohort, were planted in March 2013 into 12 whole-tree chambers (WTC) at the Hawkesbury Forest Experiment site near Richmond, NSW, Australia.
-Each chamber has a height of 9 m and seedlings were grown for 15 months.
-A detailed description of the WTC operation and design is available in [-@barton2010whole].
-Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of +3&deg;C (ET, Figure `r airvarfig`).
-Mixing fans in each chamber maintained air movement to minimize leaf boundary layer, which allows leaf temperatures in each WTC close to those of outside control trees [@barton2010whole].
-Overall, minimal WTC effects on tree performance were previously detected across a range of structural characteristics and physiological parameters compared to outside trees [@medhurst2006whole].
-
-\
-Trees grew quickly and developed large canopies, with height growth reaching the top of the WTCs over the experiment duration.
-Trees were watered weekly with 70 L from March 2013 to November 2013.
-From October 2013 to the end of the experiment trees were watered every 15 days with the mean monthly rainfall amount for Richmond, NSW.
-In February 2013 half of the chambers (3 each of AT and ET) were subjected to a drought treatment by withholding watering.
-Due to a limited range of data for the drought treatment only well-watered trees are reported, which reduces the sample size of WTC (n=6 to n=3) for the final 3 months of the experiment.
-This limited amount of data is attributed to small sample sizes as well as a difficulty in generating a sufficient CO~2~ drawdown inside the leaf cuvette, due to drought conditions, needed to accurately measure g~m~.
-
-\
-Before seedlings were planted into each chamber they were maintained under well-watered conditions in 25 L pots and kept inside each chamber.
-This allowed for seedlings to gain sufficient size before planting while also allowing them to acclimate to chamber temperature treatments.
-Seedlings were planted into each chamber after mean seedling height reached ca.
-100 cm.
-The top soils at this site, used in both pots and chambers, are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg-1 total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity.
-A root exclusion barrier extended from chamber walls to the hard layer (ca.
-1 m) and roots were allowed to grow freely below the barrier.
-Leaf gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.
-At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm, height of 348&pm;15.1 cm and an leaf area of 3.9&pm;0.1 m^2^.
-For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an leaf area of 6.2&pm;0.2 m^2^.
-Leaf area was calculated based on complete leaf counts and mean leaf size from a subsample.
-
-##3.3.2 Leaf gas exchange, online carbon isotope discrimination and mesophyll conductance
-Leaf gas exchange measurements were performed six times, beginning in October 2013 and monthly from December 2013 to April 2014.
-Measurements were taken on a representative sun and shade leaf for each tree during each measurement campaign.
-The newest fully expanded leaf from the branch apex was chosen for gas exchange measurements and sun leaves were measured in the upper third of the canopy.
-Here, shade leaves are defined as inner-canopy leaves developing on secondary branches in a low light environment.
-Shade leaves were always measured in the lower canopy, but leaves were sampled on subsequent higher branches across measurement campaigns to minimize confounding effects of leaf age.
-As shade leaves most likely developed slower this assured that older leaves in the bottom canopy were avoided.
-
-\
-Prior to gas exchange measurements photosynthetic photon flux density (PPFD) was recorded as a point measurement at the individual leaf level and a spatially averaged measurement at the canopy position for each selected leaf.
-A hand-held quantum sensor (LI-COR, Lincoln, NE, USA) was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either full sun or full shade.
-A ceptometer (AccuPAR LP-80, Decagon Devices, Pullman, USA) was then used to measure a spatially averaged PPFD at the canopy height of each chosen leaf type.
-Each ceptometer reading integrated an array of 80 sensors over a total length of 84 cm.
-Five ceptometer readings were recorded at different locations within the canopy, but at the same height and close to each selected leaf.
-The mean of these readings was assumed to represent the local light environment of sun and shade leaves for each tree.
-All measurements of PPFD and gas exchange were performed on sunny days between 10:00-14:30 h.
-
-\
-Leaf level gas exchange was measured with a standard (2 x 3 cm) leaf cuvette using a portable gas exchange system (LI-6400XT, LI-COR, Lincoln, NE, USA).
-This system was coupled with a tunable diode laser (TDL; TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of online C isotope discrimination.
-The CO~2~ in the leaf cuvette was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mol s^-1^.
-We did not explicitly control relative humidity within the leaf cuvette.
-Two identical gas exchanges systems were run simultaneously, one in each of an AT or ET WTC.
-Leaf temperatures were controlled at the current AT or ET WTC air temperature.
-PPFD in the cuvette was set to match the individual light environment of each leaf type (explained above).
-Periods of high irradiance were simulated for shade leaves by increasing the leaf cuvette PPFD (LI-COR red/blue light source) to match the light environment of the full sun leaf in the same tree.
-The maximum sunfleck response of shade leaves was then recorded once CO~2~ and water vapor fluxes re-stabilized in the leaf cuvette (ca.
-25 min).
-
-\
-Once CO~2~ and water vapor flux values were stable for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet of the LI-6400.
-The gas streams were dried by passing through napion gas dryers in the respective gas lines, and then ^12^CO~2~ and ^13^CO~2~ concentrations were measured for each gas stream by the TDL.
-Reference, sample and two calibration gases were run on alternating 80 s loops (20 s each), one for each AT and ET leaf at a matched canopy position, for a total of 12 min.
-This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference line and samples line gas streams for calculations.
-The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDL on each measurement cycle.
-Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] (C~i~) were auto-logged every 15 s for each gas exchange system over the same 12 min interval.
-
-\
-Using online C isotope discrimination measurements, the difference between the observed discrimination and what is predicted for light saturated gas exchange is proportional to g~m~ [@griffiths2013mesophyll].
+Using online C isotope discrimination measurements, the difference between the observed discrimination and what is predicted for light saturated gas exchange is proportional to *g*~m~ [@griffiths2013mesophyll].
 First, leaf C isotope discrimination was calculated by comparing the isotopic composition of the reference gas entering the leaf cuvette ($\delta$^13^C~e~) with the sample gas ($\delta$^13^C~o~) such that:
@@
 
-where $\Delta$~o~ is the observed discrimination and  $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, g~m~, respiration and photorespiration, respectively.
+where $\Delta$~o~ is the observed discrimination and  $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if *C*~i~ = *C*~c~, *g*~m~, respiration and photorespiration, respectively.
 The equations for each are as follows:
@@
 where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ is the fractionation from boundary layer diffusion (2.9&permil;) and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon].
-C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~).
+C~a~ and *C*~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~).
 In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulus* from Crous et al.
@@
 
-\
-The CO~2~ diffusion from the intercellular airspace to the chloroplast, g~m~, is given by its relationship to the leaf photosynthesis rate (*A~n~*) by:
-
-\
-(12)
-$$g_m = \frac{A_n}{C_i-C_c}$$
-
-where C~c~ is the chloroplast CO~2~ partial pressure.
-Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were then estimated using Equation 12.
-Examples of this approach to measure gas exchange and C isotope discrimination are presented in Evans and Von Caemmerer [-@evans2013temperature].
-The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck where then compared as a function of C~i~:C~a~.
-
-##3.3.3 Biochemical parameters of photosynthesis
-Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each WTC prior to the initiation of the drought treatment.
-Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011] using default parameters.
-
-##3.3.4 Leaf chemistry and hydraulic parameters
-Following gas exchange measurements each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, dried and weighed.
-These leaves were then milled and analyzed for leaf N content and $\delta$^13^C.
-Leaf samples were analysed on a Delta V Advantage coupled to a Flash HT and Conflo IV (Thermo Fisher Scientific, Bremen, Germany) in dual-reactor setup.
-Samples were flash combusted at 1000&deg;C and converted to CO~2~ and N~2~ and then subjected to stable isotope ratio mass spectrometry.
-Leaf N is reported on an area basis (N~a~, g m^-2^) and isotopic signatures of dry leaf matter are reported relative to VPDP.
-
-\
-Predawn ($\Psi_{pd}$) and midday ($\Psi_l$) leaf water potentials (MPa) were measured for sun and shade leaves during each gas exchange campaign using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA).
-The leaf closest to the leaf used for gas exchange was sampled for measurement of $\Psi_{pd}$.
-Predawn leaf water potential was measured before sunrise on the same day as gas exchange measurements.
-Leaves used for gas exchange were immediately sampled for $\Psi_l$ once measurements were completed.
-All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
-The degree to which leaf cuvette conditions, especially for shade leaves, modified $\Psi_l$ is unknown.
-Leaf water potential and transpiration (E, mmol^-2^ m^-2^ s^-1^) from gas exchange in the leaf cuvette were then used to calculate leaf-specific hydraulic conductance (K~l~, mmol m^-2^ s^-1^ MPa) with the equation:
-
-\
-(13)
-$$K_l = \frac{E}{\Psi_{pd}~-\Psi_l}$$
-
-Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A~n~* divided by E.
-The g~1~ parameter, describing the plants water-use strategy, was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near].
-Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use, whereas large values imply a lower C construction cost and decreased water-use efficiency [@lin2015optimal].
-
-##3.3.5 Data analysis
-Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R [@RDevelopmentCoreTeam2011] with WTC as a random effect.
-Explained variance (R^2^) of mixed models were computed as in [@nakagawa2013general], in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
-Confidence intervals (95 %) of mixed effect linear models were generated using bootstrapping methods with 999 simulations, using the bootMer function in the 'lme4' package [@bates2013lme4].
-For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package, using WTC as a random effect.
-All tests of statistical significance were conducted at an $\alpha$ of 0.05.
-
-#3.4  Results
-
-##3.4.1 Leaf resource distribution
-Across six measurement campaigns over the 7 month period PPFD was  reduced on average by >75% in the shade (Figure `r parfig`).
-Leaf-specific hydraulic conductance (K~l~) was similar across sun and shade leaves (Table `r wtc3tab1`).
-This was because $\Psi_{pd}$, $\Psi_l$ (Table `r wtc3tab1`) and E (Table `r wtc3tab2`) did not differ between leaf types.
-Leaf N~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table `r wtc3tab1`).
-Leaf mass per area was not different between leaf types.
-No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K~l~, E, N~area~ or LMA either within or across leaf types (P > 0.05).
-\
-```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/par2.R")
-```  
-\
-**Figure 3.1**.
-Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014.
-Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf.
-Each date represents the starting date for each measurement campaign.
-Points represent the mean (&pm; 1 standard error) daily maximum air temperature during each campaign period.
-
-\
-**Table 3.1**.
-*Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments.
-Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6).
-Values of Vc~max~ and J~max~ are treatment means (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light iwth units $\mu$mol m^-2^ s^-1^.
-Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, $\Psi$ is MPa and $\delta$^13^C is &permil;.
-Different letters represent significant differences between leaf type and temperature treatments.
-The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.
-
-```{r, echo=FALSE}
-resourcetab2 <- rbind(resourcetab, c("P value","", as.character(pval)))
-greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
-
-dust(resourcetab2)%>%
-  sprinkle_colnames("Leaf", "Temperature", "LMA", "N~a~", "Vc~max~", "J~max~","K",paste0("$\\Psi$","~pre~"),paste0("$\\Psi$","~l~"), paste0("$\\delta$","^13^c")) %>%
-  sprinkle(rows= 1, part="head",bold=TRUE)%>%
-  sprinkle_print_method("markdown")
-```
-
-\
-**Table 3.2**.
-*Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.
-Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6).
-Units for *A~n~* and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD is kPa.
-Different letters represent significant differences between leaf type, light environment and temperature treatments.
-The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.
-
-```{r, echo=FALSE}
-getab2 <- rbind(getab, c("P value","","", as.character(pval2)))
-greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
-
-dust(getab2)%>%
-  sprinkle_colnames("Leaf", "Light","Temperature", "*A~n~*", "g~s~", "g~m~", "ITE","E","VPD","C~i~", "C~c~") %>%
-  sprinkle(rows= 1, part="head",bold=TRUE)%>%
-  sprinkle_print_method("markdown")
-```
-
-##3.4.2 Photosynthetic capacity and leaf photosynthesis rates 
-The photosynthetic parameters J~max~ and Vc~max~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from AC~i~ curves (Figure `r acifig`a).
-Within leaf types, no effect of the warming treatment was detected on either parameter.
-Among the sampled leaves, Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
-\
-```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/aci_vcmaxnitro.R")
-```  
-\
-**Figure 3.2**.
-(a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.
-ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-(b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light.
-(c) The relationship between *A~n~* and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
-For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
-
-\
-Mean *A~n~* was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment (Table `r wtc3tab2`).
-Additionally, leaf N~a~ was positively related to *A~n~* across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
-Following an increase in light intensity to match high-light conditions, *A~n~* of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table `r wtc3tab2`).
-No effect of the warming treatment was detected on *A~n~* of sun leaves measured at high light or shade leaves at either low or high light.
-Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
-
-##3.4.3 Stomatal conductance and leaf water-use efficiency
-On average, g~s~ was 18% higher in shade compared to sun leaves under their local light environment (Table `r wtc3tab2`).
-Photosynthesis was positively correlated with g~s~ in all leaves measured under high light conditions, however, g~s~ and *A~n~* were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
-Following increased PPFD, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
-No effect of the warming treatment was detected on g~s~ within or across leaf types.
-Stomatal conductance within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get2fig`b).
-\
-```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/Agmgs.R")
-```  
-\
-**Figure 3.3**.
-The response of *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments.
-Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b).
-Grey areas are 95% confidence intervals for the mean.
-\
-```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("scripts/physiology_barplots.R")
-```  
-\
-**Figure 3.4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
-
-\
-Measured under ambient light and temperature, leaf ITE was significantly greater in sun leaves than in shade leaves at low light (+21%, P = 0.001).
-Following an increase in  PPFD to high-light conditions, ITE of shade leaves did not differ from shade leaves at low light and was therefore significantly lower than sun leaves (P < 0.001).
-Instantaneous transpiration efficiency in sun leaves was reduced in the warming treatment, but no warming effect was detected in shade leaves measured at low or high light (Table `r wtc3tab2`).
-The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
-For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to response curves from the optimal ITE model with a g~1~ value for each leaf type and treatment (Figure `r waterusefig`a).
-Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).
-\
-```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=600}
-source("master_scripts/wateruse.R")
-```  
-\
-**Figure 3.5**.
-(a) Response of instantaneous transpiration efficiency (ITE) to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.
-(b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light.
-For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment.
-For (b) the dashed line represents the significant linear model fit across all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
-
-\
-Bulk-leaf $\delta$^13^C, as an index of integrated water-use efficiency [@marshall2007sources], was significantly lower in shade leaves compared to sun leaves by ca.
-2&permil; (Table `r wtc3tab1`).
-No effects of the warming treatment on leaf $\delta$^13^C were detected.
-Leaf $\delta$^13^C and N~a~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C (higher water-use efficiency) and higher N investment in sun leaves.
-
-##3.4.4 Leaf carbon isotope discrimination and mesophyll conductance
-The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
-Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
-The remainder $\Delta_o$ consists of the contributions of g~s~, respiration and photorespiration to discrimination.
-
-\
-```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/delta_cica.R")
-```  
-\
-**Figure 3.6**.
-Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis ($\Delta_o$) and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
-The solid line represents the theoretical line for C3 plants from Evans et al.
-[-@evans1986carbon].
-
-\
-Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
-Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table `r wtc3tab2`).
-Proportional increases in g~m~ were matched by proportional increases in *A~n~* from low to high light in shade leaves (Figure `r gebarfig`b,c).
-Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186).
-The large increases in g~m~ in shade leaves under high light likely resulted in the highest rates of *A~n~* (Figure `r agmgsfig`b).
-No differences in g~m~ were detected with the warming treatment within leaf types.
-Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively).
-We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~.
-Comparison of AC~c~ curves (Figure `r accfig`) and AC~i~ curves revealed similar differences between sun and shade leaves.
-
-##3.4.5 Variation in intercellular and chloroplastic CO~2~ concentrations
-Higher rates of g~s~ in shade leaves under low and high light led to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a).
-The chloroplast CO~2~ partial pressure was comparable between shade leaves when measured at both low and high light conditions (Figure `r ciccfig`c).
-In sun leaves C~c~ was significantly lower than shade leaves, consistent with a lower C~i~.
-The drawdown of CO~2~ from intercellular spaces to the chloroplast, C~i~-C~c~, measures the coordination between g~m~ and *A~n~* [@caemmerer2014temperature].
-This drawdown was similar between sun and shade leaves measured at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
-This was the result of the proportional relationship between g~m~ and *A~n~* across all leaves.
-The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
-\
-```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("scripts/cicc_bar.R")
-```  
-\
-**Figure 3.7**.
-The mean &pm; 1 standard error of (a) intercellular CO~2~ concentration (C~i~), (b) CO~2~ concentration in the chloroplasts (C~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (C~i~-C~c~).
-
-#3.5  Discussion
-Here we show that *A~n~* in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
-Although shade leaves in lower light environments exhibited relatively high g~s~ and C~i~, it was rapid increases in g~m~ under periods of high light availability that allowed for this up-regulation of *A~n~*.
-Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
-These findings offer new insights into how aspects of leaf physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake in future studies.
-Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology.
-
-##3.5.1 Resource distribution and photosynthetic capacity
-The allocation of N~a~ constrains *A~n~* and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
-Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy [@mooney1979environmental].
-As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical].
-Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and *A~n~* were all reduced in shade leaves.
-Leaf mass per area, however, was not different between sun and shade leaves.
-This could be due to leaf formation under comparative light conditions or possible unmeasured differences in total non-structural carbohydrates contents between leaf types.
-Alternatively, plasticity in leaf structure and anatomy of this species may have been necessary to achieve the observed dynamic physiological responses to light.
-As gradients in LMA within tree canopies are commonly observed, this finding warrants further study.
-
-\
-Photosynthesis is also limited by the ability to supply water to the upper canopy.
-Ultimately, the ability of tree hydraulic architecture to supply water to foliage across increasing pathlengths affects productivity and survival [@sellin2008effects].
-Using a two-leaf model, Peltoniemi et al.
-[-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if K~l~ is also optimally distributed.
-In this study, variation in leaf N distribution and *A~n~* rates were not associated with subsequent changes in K~l~ between sun and shade leaves.
-Thus, no direct relationship between water supply and N distribution or *A~n~* within the canopy were detected.
-
-\
-Unexpected higher rates of g~s~ in shade leaves compared to sun leaves led to decreased ITE in shade leaves throughout the experiment.
-Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was likely prevalent long term.
-From a canopy perspective this pattern in water-use efficiency initially appears to be detrimental to C gain as *A~n~* in sun leaves was characterized by low rates of g~s~ and low C~i~.
-Relative to the differences in *A~n~* between leaf types, higher rates of g~s~ in shade leaves appear to exhibit inefficient water use.
-As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower ITE compared to sun leaves.
-
-##3.5.2 Physiological behaviour of sun and shade leaves
-The pattern of inefficient water use in shade leaves is important as we hypothesized that g~s~ and *A~n~* would be proportional across sun and shade leaves.
-In sun leaves, *A~n~* and g~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
-However, lower rates of *A~n~* in shade leaves were not coupled with decreases in g~s~, leading to the observed decreases in ITE.
-This is significant as optimal stomatal regulation to balance C gain with water loss has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on sun leaves [e.g.
-@prentice2014balancing; @lin2015optimal].
-As a result, the often used economic framework of balancing costs of using water versus N allocation to predict *A~n~* [@wright2003least] may break down when considering all leaves within a tree canopy.
-
-\
-It is possible that reducing stomatal response time, by sustaining higher g~s~, is a strategy to take advantage of high light quickly in shade leaves [@tausz2005dynamic].
-Evidence from this study supports this hypothesis, as shade leaves increased *A~n~* equivalent or even outperforming sun leaves under identical light intensity.
-Transpiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck induced thermal load [@schymanski2013stomatal].
-This is because rapid increases in leaf temperature with sunflecks have been shown to inhibit C gain [@leakey2003high].
-However, this response likely occurs at very high air temperatures and may not explain the observed g~s~ in shade leaves across the large natural range of temperatures included in this study.
-How prevalent each of these strategies are within tree canopies is still unknown, as empirical studies assessing photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@pepin1997rates;@leakey2002relative] and understory plants, often in deep shade [@chazdon1991importance; @allen2000stomatal; @brantley2009contribution].
-Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy [see @de1997simple].
-
-\
-We found that *A~n~* and g~m~ scaled positively across leaf types and, surprisingly, increased rapidly (within minutes) and proportionately when light intensity was increased in shade leaves.
-Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant].
-This provides a potential explanation for the observed rapid increases in g~m~, but the impacts of aquaporins on g~m~ are yet to be tested in leaves of tree species.
-Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll].
-Here we provide empirical data showing g~m~ not only varies within a canopy but the up-regulation of g~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.
-
-\
-If shade leaves "lie in wait" for sunflecks, then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution.
-This is because the role of g~s~ in regulating photosynthetic induction impacts the capacity of a leaf to utilize sunflecks [@way2012sunflecks].
-If the valuation of sunflecks as a C resource is large enough, then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan or across the entire canopy when considering both sun and shade leaf types.
-For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to tolerate decreases in ITE.
-However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.
-Thus, models which predict leaf photosynthesis from N distribution within a canopy will be incomplete unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].
-
-##3.5.3 Conclusions
-Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy.
-Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.
-These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have important implications for process-based models that predict canopy C gain from rates of leaf photosynthesis.
-Furthermore, the dynamic nature of g~m~ cannot be simply parameterized in tree growth models and possibly should be excluded until it can be represented properly.
-Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown.
-To improve our ability to predict whole canopy C gain, future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.
-
-#3.6  Supporting Information
-
-```{r airvars, fig.width=7, fig.height=8.8, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/air_vars.R")
-```
-\
-**Figure 3.S1**. Daily maximum and minimum temperature (a), daily maximum VPD (b) and total daily PPFD (c) for each chamber across the experiment duration.
-\
-```{r acc, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("scripts/ACc_figure_points.R")
-```  
-\
-**Figure 3.S2**.
-Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.
-C~c~ values were predicted with g~m~ and curves represent chloroplastic photosynthetic parameters at 25°C and saturating light (1800 $\mu$mols m^-1^ s^-1^).
-\
-```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/gasex_temp3.R")
-```  
-\
-**Figure 3.S3**.
-Response of *A~n~* (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.
-Shaded symbols represents each monthly measurement campaign.
-Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
-All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
-Weak negative relationships with g~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (R^2^ = 0.16 and 0.08, respectively).
-\
-```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
-source("master_scripts/gasex_temp4.R")
-```  
-\
-**Figure 3.S4**.
-Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
-Shaded symbols represents each monthly measurement campaign.
-Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
-All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
-Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (R^2^ = 0.73, 0.58 and 0.72, respectively).
+Here, *g*~m~ is expressed in units of mol m^-2^ s^-1^ bar^-1^ to be consistent with Evans & von Caemmerer [-@evans2013temperature].
+As the pressure term is nearly equal to unity, suitable comparisons between *g*~m~ and *g*~s~ can be made.
 

### 2016-06-23 gm_newphyt/manuscript_gmes.Rmd
Minor edits and reference for multcomp

@@
 Leaf gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.
-At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm **SD, SE or CI?**, height of 348&pm;15.1 cm and an leaf area of 3.9&pm;0.1 m^2^.
+At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm (SE), height of 348&pm;15.1 cm and an leaf area of 3.9&pm;0.1 m^2^.
 For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an leaf area of 6.2&pm;0.2 m^2^.
@@
 ##Data analysis
-Differences in respones of dependent variables to either the warming treatment or leaf type were analysed with linear mixed-effects models with WTC as a random effect, and leaf type (sun, shade or shade in the sun) and warming treatment as fixed effects **(and the interaction??)**.
+Differences in respones of dependent variables to either the warming treatment or leaf type were analysed with linear mixed-effects models with WTC as a random effect, and leaf type (sun, shade or shade in the sun) and warming treatment, and the interaction as fixed effects.
+The interaction was never significant, so we only report results of the tests of the main effects.
 Explained variance (R^2^) of mixed models were computed as in Nakagawa & Schielzeth [-@nakagawa2013general], in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
@@
 All tests of statistical significance were conducted at an $\alpha$ of 0.05.
+For multiple post-hoc comparisons, we used Tukey pairwise comparison tests with the 'multcomp' package [@hothorn2008].
+In all results, we report the mean &pm; one standard error for the mean (SE).
 All analyses were performed with R 3.3.0 [@RDevelopmentCoreTeam2011].
@@
 Leaf mass per area (LMA) was not different between leaf types (Table `r wtc3tab1`).
-No effect of the warming treatment **was interaction tested??** was detected with  PPFD, $\Psi_{pd}$, $\Psi_L$, *K*~L~, *E*~L~, *N*~a~ or LMA either within or across leaf types (P > 0.05).
+No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_L$, *K*~L~, *E*~L~, *N*~a~ or LMA either within or across leaf types (P > 0.05).
 
@@
 Points connected by a line represent the mean (&pm; 1 standard error) daily maximum air temperature during each campaign period.
-Each date represents the starting date for each measurement campaign; each campaign was completed in **how many days???*.
+Each date represents the starting date for each measurement campaign; each campaign was completed in 3 days (**Check this- one or two took longer than this?**).
 

### 2016-06-27 gm_newphyt/manuscript_gmes.Rmd
Final round of edits everywhere

@@
 ---
-title: Coupled response of stomatal and mesophyll conductance to light enhances photosynthesis of shade leaves under sunflecks.
+title: Coupled response of stomatal and mesophyll conductance to light enhances photosynthesis of shade leaves under sunflecks
 author: ""
@@
 * To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in the field in climate controlled whole tree chambers.
-* Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic biochemical parameters (*V*~cmax~ and *J*~max~), but *g*~s~ was similar.
-When light intensity was temporarily increased for shade leaves to match that of sun leaves, both *g*~s~ and *g*~m~ increased, leading to increases in *A*~n~ greater than sun leaves.
-* Here we show that dynamic physiological responses of shade leaves to altered light environments have important implications for up-scaling leaf level measurements and predicting whole canopy carbon gain.
-Despite exhibiting reduced photosynthetic biochemical capacity, the rapid up-regulation of *g*~m~ with increased light enables shade leaves to respond quickly to sunflecks, possibly representing a new mechanism underpinning leaf gas exchange responses to light.
+* Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic capacity (*V*~cmax~ and *J*~max~), but *g*~s~ was similar.
+When light intensity was temporarily increased for shade leaves to match that of sun leaves, both *g*~s~ and *g*~m~ increased, and *A*~n~ increased to values greater than sun leaves.
+* Here we show that dynamic physiological responses of shade leaves to altered light environments have implications for up-scaling leaf level measurements and predicting whole canopy carbon gain.
+Despite exhibiting reduced photosynthetic capacity, the rapid up-regulation of *g*~m~ with increased light enables shade leaves to respond quickly to sunflecks.
 
@@
 This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal].
-Mesophyll conductance can also impose limitations on *A*~n~ that can be as large as *g*~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A*~n~ [@niinemets2007photosynthesis], if *g*~m~ constrains CO~2~ supply to the chloroplast.
+Mesophyll conductance can also impose limitations on *A*~n~ that can be as large as *g*~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A*~n~ [@niinemets2007photosynthesis].
 Part of the variation in photosynthetic capacity between sun and shade leaves may arise from differences in *g*~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
@@
 How shade leaves utilize sunflecks for short term carbon gain depends on the combined response time of *g*~s~ and *g*~m~ and the underlying photosynthetic machinery acclimated to a low light environment [@pearcy1990sunflecks;@tausz2005dynamic].
-The utilization of sunflecks in shade leaves is first limited by delayed responses of stomata opening, which may take minutes, effectively limiting the maximum assimilation rate that can be achieved [@pearcy1990sunflecks; @vicomanz2011; @way2012sunflecks].
+The utilization of sunflecks in shade leaves is initially limited by delayed responses of stomata opening, which may take minutes, limiting the assimilation rate that can be achieved [@pearcy1990sunflecks; @vicomanz2011; @way2012sunflecks].
 Mesophyll conductance has been shown to respond to environmental factors (e.g.
@@
 For example, *g*~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid].
-Anatomical parameters which regulate *g*~m~ with changing growth irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative; @terashima2001sun; @hanba2002effect] are also unlikely to adjust during transient fluctuations in light.
+Anatomical parameters which regulate *g*~m~ with changing growth irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative; @terashima2001sun; @hanba2002effect] are unlikely to adjust during transient fluctuations in light.
 The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance of the growth environment and as a potential response to transitory light availability.
@@
 Climate warming may also affect the physiological behaviour of leaves within a canopy.
-This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis].
+Leaves are exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis].
 How these stresses affect the diffusion of CO~2~, through either *g*~s~ or *g*~m~, will have implications for up-scaling *A*~n~ for sun and shade leaves.
@@
 How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients.
-The impacts of warming on plant physiological processes are vast, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.
+The impacts of warming on plant physiological processes are diverse, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.
 
 \
-In this study we use *Eucalyptus tereticornis* trees, planted in naturally sunlit climate controlled whole-tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to test whether the distribution of N and water supply capacity via leaf-specific hydraulic conductance and leaf physiological behaviour result in  higher photosynthetic capacity in sun leaves compared to shade leaves.
-We further aimed to quantify the contraints on *A*~n~ in sun and shade leaves arising from photosynthetic capacity and the CO~2~ diffusion pathway via stomatal and mesophyll conductance.
-As leaves which developed in the shade were expected to have lower photosynthetic capacity and correspondingly lower diffusive conductance (both *g*~s~ and *g*~m~) than sun leaves, we predicted that sunfleck simulations would increase *A*~n~ and *g*~s~ to values lower than that attained by sun leaves in similar conditions.
-We further predicted that climate warming would decrease *g*~s~ and leaf C gain in sun leaves more so than shade leaves during summer months, as increased evaporation demand from higher temperatures and irradiance lead to stomatal closure.
+In this study we use *Eucalyptus tereticornis* trees, planted in naturally sunlit climate controlled whole-tree chambers with ambient and elevated temperature (ambient +3&deg;C) treatments, to test whether the distribution of N,  water supply capacity and leaf physiological traits result in  higher photosynthetic capacity in sun leaves compared to shade leaves.
+We further aimed to quantify the constraints on *A*~n~ in sun and shade leaves arising from photosynthetic capacity and the CO~2~ diffusion pathway via stomatal and mesophyll conductance.
+As leaves which developed in the shade were expected to have lower photosynthetic capacity and correspondingly lower diffusive conductance (both *g*~s~ and *g*~m~) than sun leaves, we predicted that sunfleck simulations would increase *A*~n~ and *g*~s~ in shade leaves, but reach values lower than that attained by sun leaves in similar conditions.
+We further predicted that climate warming would decrease *g*~s~ and leaf C gain in sun leaves more so than shade leaves during summer months, as increased evaporative demand from higher temperatures and irradiance lead to stomatal closure.
 
@@
 Leaf gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.
-At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm (SE), height of 348&pm;15.1 cm and an leaf area of 3.9&pm;0.1 m^2^.
+At this point, trees under AT treatment had a mean stem diameter (measured at 65cm from the stem base) of 28.2&pm;1.1 mm (SE), height of 348&pm;15.1 cm and a total leaf area of 3.9&pm;0.1 m^2^.
 For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an leaf area of 6.2&pm;0.2 m^2^.
@@
 The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDL on each measurement cycle.
-Photosynthesis, *g*~s~, transpiration, VPD and intercellular [CO~2~] (*C*~i~) were auto-logged every 15 s for each gas exchange system over the 12 min interval.
+All gas exchange variables were auto-logged every 15 s for each gas exchange system over the 12 min interval.
 
@@
 ##Photosynthetic capacity from A-*C*~i~ curves
-Photosynthetic CO~2~ response (A-*C*~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each WTC, sampling in February 2014 when all tress were well-watered in each temperature treatment (*n*=6).
-Each A-*C*~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-From these curves the photosynthetic parameters, *J*~max~ and *V*~cmax~, were quantified using the photosynthesis model of [@farquhar1980biochemical].
-We used the 'fitaci' function in the 'plantecophys' package in R, see [-@duursma2015] for detailed description of the fitting methods.
-This package uses non-linear regression to estimate *J*~max~, *V*~cmax~ and *R*~dark~ simultaneously.
+Photosynthetic CO~2~ response (A-*C*~i~) curves were measured for one sun and shade leaf for each WTC in February 2014, when all trees were well-watered in each temperature treatment (*n*=6).
+Each A-*C*~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ with leaf temperature (*T*~leaf~) held constant at 25 &deg;C, and saturating PPFD (1800 $\mu$mols m^-1^ s^-1^).
+From these curves the photosynthetic parameters, *J*~max~ and *V*~cmax~, were quantified using the photosynthesis model of Farquhar et al.
+[-@farquhar1980biochemical].
+We used the 'fitaci' function in the 'plantecophys' package in R, see Duursma [-@duursma2015] for detailed description of the fitting methods.
+This method uses non-linear regression to estimate *J*~max~, *V*~cmax~ and *R*~dark~ simultaneously.
 Because there was considerable practical difficulty in fitting A-*C*~c~ curves, we report standard *C*~i~-based *J*~max~ and *V*~cmax~, which effectively include a mesophyll conductance component.
-When fitting the A-Ci curves, we calculated $\Gamma_{*}$ with the same methods as for *g*~m~, described above (the estimated value at 25&deg;C was 38.9 $\mu$mol mol^-1^), and the combined Michaelis-Menten coefficient (*K*~m~) was estimated from [-@medlyn2002temperature] (713.4 $\mu$mol mol^-1^ at 25&deg;C).
+When fitting the A-*C*~i~ curves, we calculated $\Gamma_{*}$ with the same methods as for *g*~m~, described above (the estimated value at 25&deg;C was 38.9 $\mu$mol mol^-1^), and the combined Michaelis-Menten coefficient (*K*~m~) was estimated from Medlyn et al.
+[-@medlyn2002temperature] (713.4 $\mu$mol mol^-1^ at 25&deg;C).
 The temperature response parameters for *J*~max~, *V*~cmax~ and *K*~m~ are not relevant since *T*~leaf~ was kept constant at 25&deg;C.
 
-##Leaf chemistry and hydraulic parameters
+##Leaf nitrogen and hydraulic conductance
 Following gas exchange measurements, each leaf was collected, measured for leaf water potential (explained below), scanned for leaf area, oven-dried and weighed.
@@
 VPD response curves by fitting a rearranged optimal *g*~s~ model for ITE [@medlyn2011reconciling] using non-linear regression [see @duursma2013near].
-Small values of *g*~1~ indicate that transpiration is comparatively costly in C terms and reflecting conservative water use (increaed water-use efficiency), whereas large values imply a lower C cost and decreased water-use efficiency [@lin2015optimal].
+Small values of *g*~1~ indicate that transpiration is comparatively costly in C terms and reflecting conservative water use (increased water-use efficiency), whereas large values imply a lower C cost and decreased water-use efficiency [@lin2015optimal].
 
 ##Data analysis
-Differences in respones of dependent variables to either the warming treatment or leaf type were analysed with linear mixed-effects models with WTC as a random effect, and leaf type (sun, shade or shade in the sun) and warming treatment, and the interaction as fixed effects.
+Differences in responses of dependent variables to either the warming treatment or leaf type were analysed with linear mixed-effects models with WTC as a random effect, and leaf type (sun, shade or shade in the sun) and warming treatment, and the interaction as fixed effects.
 The interaction was never significant, so we only report results of the tests of the main effects.
-Explained variance (R^2^) of mixed models were computed as in Nakagawa & Schielzeth [-@nakagawa2013general], in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
-Confidence intervals (95 %) of mixed effect linear models were generated using bootstrapping methods with 999 simulations, using the bootMer function in the 'lme4' package [@bates2013lme4].
+Explained variance (R^2^) of mixed models was computed as in Nakagawa & Schielzeth [-@nakagawa2013general], in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
+Confidence intervals (95 %) of linear mixed-effects models were generated using bootstrapping methods with 999 simulations, using the 'bootMer' function in the 'lme4' package [@bates2013lme4].
 For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package [@wood2006], using WTC as a random effect.
@@
 Leaf-specific hydraulic conductance (*K*~L~) was similar across sun and shade leaves (Table `r wtc3tab1`).
-This was because neither leaf water potentials ($\Psi_{pd}$ and $\Psi_L$, Table `r wtc3tab1`) nor transpiration rates (E, Table `r wtc3tab2`) differed between leaf types.
+This was because neither leaf water potentials ($\Psi_{pd}$ and $\Psi_L$, Table `r wtc3tab1`) nor transpiration rates (*E*~L~ , Table `r wtc3tab2`) differed between leaf types.
 Leaf *N*~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table `r wtc3tab1`).
 Leaf mass per area (LMA) was not different between leaf types (Table `r wtc3tab1`).
-No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_L$, *K*~L~, *E*~L~, *N*~a~ or LMA either within or across leaf types (P > 0.05).
+No effect of the warming treatment was detected on  PPFD, $\Psi_{pd}$, $\Psi_L$, *K*~L~, *E*~L~, *N*~a~ or LMA either within or across leaf types (P > 0.05).
 
@@
 ##Leaf carbon isotope discrimination and mesophyll conductance
-The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with *C*~i~/C~a~ for both leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
+The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with *C*~i~/*C*~a~ for both leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
 Carbon isotope discrimination associated with *g*~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
@@
 Proportional increases in *g*~m~ were matched by proportional increases in *A*~n~ from low to high light in shade leaves (Figure `r gebarfig`b,c).
-Photosynthesis scaled positively with increases in *g*~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186).
+Photosynthesis scaled positively with increases in *g*~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.019).
 The large increases in *g*~m~ in shade leaves under high light likely resulted in the highest rates of *A*~n~ (Figure `r agmgsfig`b).
 No differences in *g*~m~ were detected with the warming treatment within leaf types.
-Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively).
+Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 for sun leaves, P = 0.04 for shade leaves).
 
@@
 This drawdown was similar between sun and shade leaves measured at their local light and temperature environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c).
-This was the result of the proportional relationship between *g*~m~ and *A*~n~ across all leaves.
-The CO~2~ drawdown from C~a~ to *C*~i~ and *C*~i~ to *C*~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
+This result demonstrates the approximately proportional relationship between *g*~m~ and *A*~n~ across all leaves.
+The CO~2~ drawdown from *C*~a~ to *C*~i~ and *C*~i~ to *C*~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).
 
 #Discussion
-Here we show that *A*~n~ in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
-Although shade leaves in lower light environments exhibited relatively high *g*~s~ and *C*~i~, it was concomitant increases in *g*~m~ under periods of high light availability that enabled up-regulation of *A*~n~.
-Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
+Here we show that *A*~n~ in leaves within canopies of *Eucalyptus tereticornis* is limited by the local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased.
+Although shade leaves in lower light environments exhibited relatively high *g*~s~ and *C*~i~, the  increases in *g*~m~ under periods of high light availability enabled the up-regulation of *A*~n~.
+Although it is well-known that shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
 These findings offer new insights into how aspects of leaf gas exchange physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake and water use in future studies.
-Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology, owing to temperature acclimation [@aspinwall2016convergent].
+Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology, consistent with temperature acclimation demonstrated in this experiment [@aspinwall2016convergent].
 
 ##Resource distribution and photosynthetic capacity
-The allocation of *N*~a~ constrains *A*~n~ and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
+The allocation of *N*~a~ along canopy gradients constrains *A*~n~ and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.
 Decreasing light availability should decrease the investment of N into photosynthetic enzymes within a canopy [@mooney1979environmental].
 As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters *V*~cmax~ and *J*~max~ [@farquhar1980biochemical].
-Our data agree with these conventional conclusions as the distribution of *N*~a~, both measures of photosynthetic capacity and *A*~n~ were all reduced in shade leaves.
+Our data agree with these conventional conclusions as the distribution of *N*~a~, both measures of photosynthetic capacity, and *A*~n~ were all reduced in shade leaves.
 Leaf mass per area (LMA), however, was not different between sun and shade leaves, consistent with small differences reported in a clonal *Eucalyptus* plantation [@nouvellon2010].
@@
 \
-The observed pattern of inefficient water use in shade leaves leads to the rejection of the hypothesis that *g*~s~ and *A*~n~ are proportional across sun and shade leaves.
-In sun leaves, *A*~n~ and *g*~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory.
+In sun leaves, *A*~n~ and *g*~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory (Fig.
+5).
 However, lower rates of *A*~n~ in shade leaves were not coupled with decreases in *g*~s~, leading to the observed decreases in ITE.
@@
 Our results support this hypothesis, as shade leaves increased *A*~n~ to values equal or even outperforming sun leaves when subjected to a brief period of identical high-light intensity.
-Transpiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck-induced rapid increases in leaf temperature [@schymanski2013stomatal], which have been shown to inhibit C gain [@leakey2003high].
+Transpiration-induced cooling in shade leaves, by keeping stomata open, has also been suggested as an effective strategy to reduce sunfleck-induced rapid increase in leaf temperature [@schymanski2013stomatal], which has been shown to inhibit C gain [@leakey2003high].
 However, this response likely occurs at very high air temperatures and may not explain the observed *g*~s~ in shade leaves across the large natural seasonal range of temperatures included in our study.
@@
 This is because the role of *g*~s~ in regulating photosynthetic induction impacts the capacity of a leaf to utilize sunflecks [@way2012sunflecks].
-If the valuation of sunflecks as a C resource is large enough, then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan or across the entire canopy when considering both sun and shade leaf types.
+If the valuation of sunflecks as a C resource is large enough, then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan or across the entire canopy when considering both sun and shade leaf types [@vicomanz2011].
 For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to accommodate the cost of increased water use in the shade.
@@
 Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves that determines C gain.
-These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have important implications for process-based models that predict canopy C gain from rates of leaf photosynthesis.
+These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have implications for process-based models that predict canopy C gain from rates of leaf photosynthesis.
 Furthermore, the dynamic nature of *g*~m~ cannot be simply parameterized in vegetation models, and possibly should be excluded until it can be represented properly.
@@
 We thank Sune Linder and the Swedish University for Agricultural Science for providing the whole-tree chambers.
+The complete dataset can be downloaded from *link to figshare will be added*.
 
@@
 Leaf morphological and physiological traits of sun and shade leaves under ambient and elevated temperature treatments for *Eucalyptus tereticornis*.
-Leaf mass per area (LMA, g m^2^), leaf nitrogen per unit area (*N*~a~, gN m^-2^), $\delta$^13^C (&permil;), pre-dawn leaf water potential ($\Psi$~pd~, MPa), midday leaf water potential ($\Psi$~l~, MPa) and leaf-specific hydraulic condctance (*K*~L~, mmol m^-2^ s^-1^ MPa^-1^) are given as treatment means (&pm; 1 standard error) across six measurement campaigns.
+Leaf mass per area (LMA, g m^2^), leaf nitrogen per unit area (*N*~a~, gN m^-2^), $\delta$^13^C (&permil;), pre-dawn leaf water potential ($\Psi$~pd~, MPa), midday leaf water potential ($\Psi$~L~, MPa) and leaf-specific hydraulic condctance (*K*~L~, mmol m^-2^ s^-1^ MPa^-1^) are given as treatment means (&pm; 1 standard error) across six measurement campaigns.
 Values of *V*~cmax~ and *J*~max~ are treatment mean (&pm; 1 standard error) from A-*C*~i~ curves measured in each chamber at saturating light.
@@
 dust(resourcetab2)%>%
-  sprinkle_colnames("Leaf", "Temperature", "LMA", "*N*~a~", "*V*~cmax~", "*J*~max~","K",paste0("$\\Psi$","~pre~"),paste0("$\\Psi$","~l~"), paste0("$\\delta$","^13^C")) %>%
+  sprinkle_colnames("Leaf", "Temperature", "LMA", "*N*~a~", "*V*~cmax~", "*J*~max~","K",paste0("$\\Psi$","~pd~"),paste0("$\\Psi$","~L~"), paste0("$\\delta$","^13^C")) %>%
   sprinkle(rows= 1, part="head",bold=TRUE)%>%
@@
 Points connected by a line represent the mean (&pm; 1 standard error) daily maximum air temperature during each campaign period.
-Each date represents the starting date for each measurement campaign; each campaign was completed in 3 days (**Check this- one or two took longer than this?**).
+Each date represents the starting date for each measurement campaign; each campaign was completed in 3 days.
 
@@
 **Figure 2**.
-Photosynthetic capacity, photosynthetis rate and leaf nitrogen for sun and shade leaves.
+Photosynthetic capacity, photosynthesis rate and leaf nitrogen for sun and shade leaves.
 (a) A-*C*~i~ curves for sun and shade leaves grown under elevated (ET) and ambient (AT) temperature treatments.
@@
 Sun leaves were measured at high light, shade leaves were measured at both low and high light.
-Data are cobined across elevated (ET, triangles) and ambient (ET, circles) temperature treatments, and six measurement campaigns.
+Data are combined across elevated (ET, triangles) and ambient (ET, circles) temperature treatments, and six measurement campaigns.
 Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b), fitted across all data but accounting for random effects (see Methods).

### 2016-06-29 gm_newphyt/manuscript_gmes.Rmd
New colours and minor edits from MT and SvC

@@
 \
-Courtney E. Campany^1^, Mark G. Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^.
+Courtney E. Campany^1^, Mark G. Tjoelker^1^, Susanne von Caemmere*R*^2^, Remko A. Duursma^1^.
 
@@
 However, biochemical and diffusional constraints on gas exchange in sun and shade leaves in response to light remain poorly quantified, but critical for predicting carbon and water exchange in canopies.
-* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in the field in climate controlled whole tree chambers.
+* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with  concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in the field in climate controlled whole tree chambers.
 * Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic capacity (*V*~cmax~ and *J*~max~), but *g*~s~ was similar.
@@
 The CO~2~ in the leaf cuvette was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mol s^-1^.
-Two identical gas exchanges systems were run simultaneously, one in each of a randomly chosen WTC for each temperature treatment.
-Leaf temperatures were controlled at the current AT or ET WTC air temperature.
+Two identical gas exchange systems were run simultaneously, one in each of a randomly chosen WTC for each temperature treatment.
+Leaf temperatures were controlled at the current AT or ET WTC air temperature, and measurements were made at ambient air humidity.
 PPFD in the cuvette was set to match the individual light environment of each leaf type (explained above).
@@
 Reference, sample and two calibration gases were run on alternating 80 s loops (20 s each), one for each AT and ET leaf at a matched canopy position, for a total of 12 min.
-This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference line and sample line gas streams for calculations.
+This allowed for 4 or 5 measurements per leaf and data were averaged over the last 10 s of reference line and sample line gas streams for calculations.
 The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDL on each measurement cycle.
@@
 where $\Delta$~o~ is the observed discrimination and $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if *C*~i~ = *C*~c~, *g*~m~, respiration and photorespiration, respectively.
-For this study, the CO~2~ compensation point ($\Gamma_{*}$) and respiration during the day (R~d~) parameters originally derived for tobacco plants [@von1994kinetics] were replaced with parameters derived for *Eucalyptus globulus* from Crous et al.
+For this study, the CO~2~ compensation point ($\Gamma_{*}$) and respiration during the day (R~d~) parameters originally derived for tobacco plants [@von1994kinetics] were replaced with parameters derived for *Eucalyptus globulus* from Crous *et al*.
 [-@crous2012light] when calculating *g*~m~.
@@
 When fitting the A-*C*~i~ curves, we calculated $\Gamma_{*}$ with the same methods as for *g*~m~, described above (the estimated value at 25&deg;C was 38.9 $\mu$mol mol^-1^), and the combined Michaelis-Menten coefficient (*K*~m~) was estimated from Medlyn et al.
-[-@medlyn2002temperature] (713.4 $\mu$mol mol^-1^ at 25&deg;C).
+[-@medlyn2002temperature] (713.4 $\mu$mol mol^-1^ at 25&deg;C, using a *K*~c~ of 405 $\mu$mol mol^-1^, with an atmospheric pressure of 100 kPa).
 The temperature response parameters for *J*~max~, *V*~cmax~ and *K*~m~ are not relevant since *T*~leaf~ was kept constant at 25&deg;C.
@@
 The interaction was never significant, so we only report results of the tests of the main effects.
-Explained variance (R^2^) of mixed models was computed as in Nakagawa & Schielzeth [-@nakagawa2013general], in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
+Explained variance (*R*^2^) of mixed models was computed as in Nakagawa & Schielzeth [-@nakagawa2013general], in which the marginal *R*^2^ represents variance explained by fixed factors and the conditional *R*^2^ by both fixed and random factors.
 Confidence intervals (95 %) of linear mixed-effects models were generated using bootstrapping methods with 999 simulations, using the 'bootMer' function in the 'lme4' package [@bates2013lme4].
@@
 ##Photosynthetic capacity and leaf photosynthesis rates 
-The photosynthetic parameters *J*~max~ and *V*~cmax~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from A-*C*~i~ curves (Figure `r acifig`a).
+The photosynthetic parameters *J*~max~ and *V*~cmax~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from A-*C*~i~ curves measured at 25 &deg;C (Figure `r acifig`a).
 Within leaf types, no effect of the warming treatment was detected on either parameter.
@@
 Leaf mass per area (LMA, g m^2^), leaf nitrogen per unit area (*N*~a~, gN m^-2^), $\delta$^13^C (&permil;), pre-dawn leaf water potential ($\Psi$~pd~, MPa), midday leaf water potential ($\Psi$~L~, MPa) and leaf-specific hydraulic condctance (*K*~L~, mmol m^-2^ s^-1^ MPa^-1^) are given as treatment means (&pm; 1 standard error) across six measurement campaigns.
-Values of *V*~cmax~ and *J*~max~ are treatment mean (&pm; 1 standard error) from A-*C*~i~ curves measured in each chamber at saturating light.
+Values of *V*~cmax~ and *J*~max~ are treatment mean (&pm; 1 standard error) from A-*C*~i~ curves measured in each chamber at 25 &deg;C and saturating light.
 Different letters represent significant differences between leaf type and temperature treatments (based on a Tukey test).
@@
 (c) The relationship between leaf net photosynthesis rate (*A*~n~) and leaf *N*~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
-For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
+For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional *R*^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
 \
@@
 (b) The relationship between leaf $\delta$^13^C (higher values indicate higher water-use efficiency) and leaf *N*~a~ for sun leaves at high light and shade leaves at low light.
-The dashed line represents the significant linear model fit across all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
+The dashed line represents the significant linear model fit across all leaves with a marginal and conditional *R*^2^ of 0.41 and 0.45, respectively.
 \

### 2016-06-29 functions and packages/plot_objects_all.R

@@ -53,8 +53,8 @@ leglab2<- c(expression(paste(Sun, "-" ,"AT")), expression(paste(Sun, "-" ,"ET"))
             expression(paste(Shade,"-" ,"AT")), expression(paste(Shade, "-" ,"ET")))
 
 ###sun shade colors
-suncol <- alpha("forestgreen", alpha=.75)
-shacol <- alpha("yellow4", alpha=.75)
+suncol <- alpha(rgb(230,159,0,max=255),0.75)   #alpha("forestgreen", alpha=.75)
+shacol <- alpha(rgb(0,114,178,max=255), 0.75)      #alpha("yellow4", alpha=.75)
 leafcol <- c(suncol, shacol)
 leaflab2 <- c("Sun", "Shade")
 
@@ -62,7 +62,7 @@ colaci <-c(suncol, suncol, shacol, shacol)
 
 ###lights on shade cols
 leaflightlab <- c("shade-high", "shade-low", "sun")
-lightscol <- alpha("darkorange2", .75)
+lightscol <- alpha(rgb(230,159,0,max=255),0.75)     #alpha("darkorange2", .75)
 lightlab <- c(lightscol,shacol)
 lightleg <- c("High light", "Low light")
 

### 2016-06-29 master_scripts/plot_objects.R

@@ -83,22 +83,16 @@ acileg <-  c(expression(paste(Sun, "-" ,"AT")), expression(paste(Sun, "-" ,"ET")
 ltys <- c(1,2, 1,2)
 
 ###sun shade colors-------------------------------------------------------------------------------------------
-suncol2 <- "#005300"
 
 
-# shacol2 <- "#61726C"
-shacol2 <- "#9d2f16"
+suncol <- alpha("#ff7f00",0.75) #rgb(230,159,0,0.75*255,max=255)   #alpha("forestgreen", alpha=.75)
+shacol <- newshade <- newshacol <-  alpha("#377eb8", 0.75) 
+#rgb(0,114,178,0.75*255,max=255)      #alpha("yellow4", alpha=.75)
+lightscol <- alpha("#4daf4a", 0.75) #rgb(86,180,233,0.75*255,max=255)     #alpha("darkorange2", .75)
 
-
-lightscol2 <- "#E68A00"
-newshade <- ("#613E1C")
-
-suncol <- alpha(suncol2, alpha=.75)
-shacol <- alpha(shacol2 , alpha=.75)
-newshacol <- alpha(newshade, alpha=.75)
-
-lightscol <- alpha(lightscol2, .75)
-
+suncol2 <- alpha(suncol,1)
+shacol2 <- alpha(shacol,1)
+lightscol2 <- alpha(lightscol,1)
 
 shacol50 <- alpha(shacol, alpha=.5)
 suncol50 <- alpha(suncol, alpha=.5)

### 2016-07-06 gm_newphyt/manuscript_gmes.Rmd
minor edit on RH control
bad R2
text edit
last edits before manuscrupt submission

@@
 \
-Courtney E. Campany^1^, Mark G. Tjoelker^1^, Susanne von Caemmere*R*^2^, Remko A. Duursma^1^.
+Courtney E. Campany^1^, Mark G. Tjoelker^1^, Susanne von Caemmerer^2^, Remko A. Duursma^1^.
 
@@
 \
-Total word count (excluding summary, references and legends): 5563
+Total word count (excluding summary, references and legends): 5903
 \
@@
 \
-Introduction: 1186
+Introduction: 1242
 \
-Materials and Methods: 1863
+Materials and Methods: 2106
 \
-Results: 1098
+Results: 1114
 \
-Discussion: 1380
+Discussion: 1393
 \
-Acknowledgements: 36
+Acknowledgements: 48
 \
@@
 * Light gradients within tree canopies play a major role in the distribution of plant resources that define the photosynthetic capacity of sun and shade leaves.
-However, biochemical and diffusional constraints on gas exchange in sun and shade leaves in response to light remain poorly quantified, but critical for predicting carbon and water exchange in canopies.
-* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with  concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in the field in climate controlled whole tree chambers.
+However, biochemical and diffusional constraints on gas exchange in sun and shade leaves in response to light remain poorly quantified, but critical for predicting canopy carbon and water exchange.
+* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
 * Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic capacity (*V*~cmax~ and *J*~max~), but *g*~s~ was similar.
@@
 Leaf temperatures were controlled at the current AT or ET WTC air temperature, and measurements were made at ambient air humidity.
+Measurements were thus made across a range of air temperature and vapour pressure deficit (kPa) across the six campaigns.
 PPFD in the cuvette was set to match the individual light environment of each leaf type (explained above).
@@
 Each date represents the starting date for each measurement campaign; each campaign was completed in 3 days.
-
 \
@@
 ```
-
 \

### 2016-07-12 gm_newphyt/manuscript_gmes.Rmd
cleaning
need to remove anatomy and talk about total conductance behavior

@@
 CO~2~, temperature or vapor pressure deficit) at timescales of minutes, possibly faster than *g*~s~ [@flexas2008mesophyll and references therein], yet the short term response to light availability is unclear.
-For example, *g*~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid].
+For example, *g*~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid] and *Eucalyptus* seedlings [@douthe2011mesophyll; @douthe2012mesophyll].
+
+*missing data on tree canopies which have highly variable light environments* .combined responses....
+
 Anatomical parameters which regulate *g*~m~ with changing growth irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative; @terashima2001sun; @hanba2002effect] are unlikely to adjust during transient fluctuations in light.
+*is this sentence necessary?*
+
 The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance of the growth environment and as a potential response to transitory light availability.

### 2016-07-12 gm_pce/manuscript_gmes.Rmd
renamed manuscript folder, started making journal specific ms edits

--- 2016-07-12 gm_newphyt/manuscript_gmes.Rmd
@@
 \
-\
-Total word count (excluding summary, references and legends): 5903
-\
-Summary: 200
-\
-Introduction: 1242
-\
-Materials and Methods: 2106
-\
-Results: 1114
-\
-Discussion: 1393
-\
-Acknowledgements: 48
-\
-Number of figures: 7 (6 in color)
-\
-Number of tables: 2
-\
-Number of supporting information files: 4 figures, 1 methods
 
@@
 
-# Summary
-* Light gradients within tree canopies play a major role in the distribution of plant resources that define the photosynthetic capacity of sun and shade leaves.
+# Abstract
+Light gradients within tree canopies play a major role in the distribution of plant resources that define the photosynthetic capacity of sun and shade leaves.
 However, biochemical and diffusional constraints on gas exchange in sun and shade leaves in response to light remain poorly quantified, but critical for predicting canopy carbon and water exchange.
-* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
-* Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic capacity (*V*~cmax~ and *J*~max~), but *g*~s~ was similar.
+To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
+Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic capacity (*V*~cmax~ and *J*~max~), but *g*~s~ was similar.
 When light intensity was temporarily increased for shade leaves to match that of sun leaves, both *g*~s~ and *g*~m~ increased, and *A*~n~ increased to values greater than sun leaves.
-* Here we show that dynamic physiological responses of shade leaves to altered light environments have implications for up-scaling leaf level measurements and predicting whole canopy carbon gain.
+Here we show that dynamic physiological responses of shade leaves to altered light environments have implications for up-scaling leaf level measurements and predicting whole canopy carbon gain.
 Despite exhibiting reduced photosynthetic capacity, the rapid up-regulation of *g*~m~ with increased light enables shade leaves to respond quickly to sunflecks.
@@
 For example, *g*~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid] and *Eucalyptus* seedlings [@douthe2011mesophyll; @douthe2012mesophyll].
-
-*missing data on tree canopies which have highly variable light environments* .combined responses....
-
-Anatomical parameters which regulate *g*~m~ with changing growth irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative; @terashima2001sun; @hanba2002effect] are unlikely to adjust during transient fluctuations in light.
-*is this sentence necessary?*
-
+Anatomical parameters which regulate *g*~m~ with changing growth irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative; @terashima2001sun; @hanba2002effect] are also unlikely to adjust during transient fluctuations in light.
 The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance of the growth environment and as a potential response to transitory light availability.
@@
 These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have implications for process-based models that predict canopy C gain from rates of leaf photosynthesis.
-Furthermore, the dynamic nature of *g*~m~ cannot be simply parameterized in vegetation models, and possibly should be excluded until it can be represented properly.
+Furthermore, the dynamic nature of *g*~m~ cannot be simply parameterized in vegetation models, and further experimental evidence is needed so that it can be represented properly.
+possibly should be excluded until it can be represented properly.
 Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of *g*~m~ to respond to environmental drivers.
 
-### Acknowledgements
+# Acknowledgements
 We thank Craig Barton and Burhan Amiji for maintaining the whole-tree chambers experiment and for their outstanding technical assistance.
@@
 The complete dataset can be downloaded from *link to figshare will be added*.
-
-### Author Contributions
-CC contributed to the design of the research, data analysis, collection, interpretation and writing the manuscript.
-MT contributed to the design of the research, interpretation  and writing the manuscript.
-SC contributed to the design of the research, performance of the research and interpretation.
-RD contributed to the design of the research, data analysis, interpretation and writing the manuscript.
 

### 2016-07-12 gm_pce/manuscript_gmes_SuppInfo.Rmd
renamed manuscript folder, started making journal specific ms edits

--- 2016-06-23 gm_newphyt/manuscript_gmes_SuppInfo.Rmd
@@
 **Figure S2**.
-Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis ($\Delta_o$) and measured *C*~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
+Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis ($\Delta_o$) and measured *C*~i~/*C*~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
 The solid line represents the theoretical line for C3 plants from Evans et al.
@@
 All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
-Weak negative relationships with *g*~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (R^2^ = 0.16 and 0.08, respectively).
+Weak negative relationships with *g*~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (*R*^2^ = 0.16 and 0.08, respectively).
 \
@@
 **Figure S4**.
-Response of VPD (a), *g*~s~ (b) and C~a~-*C*~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
+Response of VPD (a), *g*~s~ (b) and *C*~a~-*C*~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
 Shaded symbols represents each monthly measurement campaign.
@@
 All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination.
-Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (R^2^ = 0.73, 0.58 and 0.72, respectively).
+Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (*R*^2^ = 0.73, 0.58 and 0.72, respectively).
 
@@
 where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ is the fractionation from boundary layer diffusion (2.9&permil;) and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon].
-C~a~ and *C*~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~).
+*C*~a~ and *C*~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~).
 In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulus* from Crous et al.

### 2016-07-13 gm_pce/manuscript_gmes.Rmd
making conversions to pce format

@@
     number_sections: yes
-csl: new-phytologist.csl
+csl: plant-cell-environment.csl
 bibliography: references.bib
@@
 
-source("gm_newphyt/manuscript_preload.R")
+source("gm_pce/manuscript_preload.R")
 
 # Make figure PDFs.
-source("gm_newphyt/make_figures.R")
+source("gm_pce/make_figures.R")
 

### 2016-07-13 gm_pce/manuscript_gmes.Rmd
pce edits

@@
   word_document:
-    reference_docx: manuscriptstyle.docx
+    reference_docx: manuscriptstyle_pce.docx
   pdf_document: default

### 2016-07-13 gm_pce/make_figures.R

@@ -13,7 +13,7 @@ source("functions and packages/functions.R")
 source("master_scripts/plot_objects.R")
 
 
-fn <- function(...)file.path("gm_newphyt",...)
+fn <- function(...)file.path("gm_pce",...)
 
 to_pdf(source("master_scripts/par2.R"), fn("Figure1.pdf"), width=7, height=7)
 to_pdf(source("master_scripts/aci_vcmaxnitro.R"), fn("Figure2.pdf"), width=8, height=10)

### 2016-07-13 gm_pce/manuscript_gmes.Rmd
mostly cite formatting

@@
 \
-*Tel: +61(0)432391114*
+*Tel: +61 2 4570 1421*
 \
@@
 \
+\
+
+**Running Title**
+\
+Mesophyll conductance response to light
 
@@
 However, biochemical and diffusional constraints on gas exchange in sun and shade leaves in response to light remain poorly quantified, but critical for predicting canopy carbon and water exchange.
-To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
+To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole-tree chambers.
 Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic capacity (*V*~cmax~ and *J*~max~), but *g*~s~ was similar.
@@
 
-##Key words
+##Keyword index
 *Eucalyptus tereticornis*, mesophyll conductance, photosynthesis, stomatal conductance, sunflecks
@@
 Mesophyll conductance has been shown to respond to environmental factors (e.g.
-CO~2~, temperature or vapor pressure deficit) at timescales of minutes, possibly faster than *g*~s~ [@flexas2008mesophyll and references therein], yet the short term response to light availability is unclear.
+CO~2~, temperature or vapour pressure deficit) at timescales of minutes, possibly faster than *g*~s~ [@flexas2008mesophyll and references therein], yet the short term response to light availability is unclear.
 For example, *g*~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid] and *Eucalyptus* seedlings [@douthe2011mesophyll; @douthe2012mesophyll].
@@
 
-##Whole tree chamber experimental design
+##Whole-tree chamber experimental design
 Twelve *Eucalyptus tereticornis* Sm.
@@
 Furthermore, the dynamic nature of *g*~m~ cannot be simply parameterized in vegetation models, and further experimental evidence is needed so that it can be represented properly.
-possibly should be excluded until it can be represented properly.
 Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of *g*~m~ to respond to environmental drivers.

### 2016-07-13 gm_pce/manuscript_gmes.Rmd
final edits, submitted to PCE

@@
 Leaf morphological and physiological traits of sun and shade leaves under ambient and elevated temperature treatments for *Eucalyptus tereticornis*.
-Leaf mass per area (LMA, g m^2^), leaf nitrogen per unit area (*N*~a~, gN m^-2^), $\delta$^13^C (&permil;), pre-dawn leaf water potential ($\Psi$~pd~, MPa), midday leaf water potential ($\Psi$~L~, MPa) and leaf-specific hydraulic condctance (*K*~L~, mmol m^-2^ s^-1^ MPa^-1^) are given as treatment means (&pm; 1 standard error) across six measurement campaigns.
+Leaf mass per area (LMA, g m^2^), leaf nitrogen per unit area (*N*~a~, gN m^-2^), $\delta$^13^C (&permil;), pre-dawn leaf water potential ($\Psi$~pd~, MPa), midday leaf water potential ($\Psi$~L~, MPa) and leaf-specific hydraulic conductance (*K*~L~, mmol m^-2^ s^-1^ MPa^-1^) are given as treatment means (&pm; 1 standard error) across six measurement campaigns.
 Values of *V*~cmax~ and *J*~max~ are treatment mean (&pm; 1 standard error) from A-*C*~i~ curves measured in each chamber at 25 &deg;C and saturating light.

### 2016-07-13 gm_pce/manuscript_gmes_SuppInfo.Rmd
final edits, submitted to PCE

@@
   word_document:
-    reference_docx: manuscriptstyle.docx
-csl: new-phytologist.csl
+    reference_docx: manuscriptstyle_pce.docx
+csl: plant-cell-environment.csl
 bibliography: references.bib
@@
 ```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
-source("gm_newphyt/manuscript_preload.R")
+source("gm_pce/manuscript_preload.R")
 ```

### 2016-08-23 master_scripts/Agmgs.R
changed colors
renamed thesis folder

@@ -1,6 +1,6 @@
-# source("functions and packages/functions.R")
-# source("master_scripts/plot_objects.R")
-# source("functions and packages/packages.R")
+source("functions and packages/functions.R")
+source("master_scripts/plot_objects.R")
+source("functions and packages/packages.R")
 
 #read in gm data set (no drought) and Cibar(discrimination)-------------------------------------------------------
 gmes <- read.csv("calculated_data/gmes_wellwatered.csv")

### 2016-08-23 master_scripts/aci_vcmaxnitro.R

@@ -1,6 +1,6 @@
-# source("functions and packages/packages.R")
-# source("functions and packages/functions.R")
-# source("master_scripts/plot_objects.R")
+source("functions and packages/packages.R")
+source("functions and packages/functions.R")
+source("master_scripts/plot_objects.R")
 
 treatments <- read.csv("raw data/temp_trt.csv") 
 

### 2016-08-23 master_scripts/par2.R

@@ -1,6 +1,6 @@
-# source("functions and packages/functions.R")
-# source("functions and packages/packages.R")
-# source("master_scripts/plot_objects.R")
+source("functions and packages/functions.R")
+source("functions and packages/packages.R")
+source("master_scripts/plot_objects.R")
 
 treatments <- read.csv("raw data/temp_trt.csv")
 

### 2016-08-26 gm_pce/review_response.Rmd
started reviewer response RMD started edits in manuscript

---
title: "Response to referees"
author: "Court Campany"
date: "26 August 2016"
output: word_document
---

```{r, results='asis', echo=FALSE, warning=FALSE}
library(knitr)
opts_knit$set(root.dir = '../')
```

*Editor*

*Referee #1*

*Referee #2*

- 1. Wingate et al. 2007:

\
- 2. The measurements were taken as a closed system with the coupled TDl-Licor 6400, independent from the whole-tree chamber conditions. Additionally, specifically choosen CO~2~ cyclinders used for the Licor 6440 were closer to ambient air 13C (~ 10&permil) instead of standard CO~2~ which are fossil fuel derived (~ 30&permil). Thus the WTC atmosphere or CO~2~ cylinder $\delta$^13^C should not interfer with steady state conditions of the leaf measuremnts.

__Specific Comments:__
- line 358:We corrected the text to match stats from Table 2

- line 392: We agree that thermal acclimation is not previosuly discussed before this point is made. However, the experimental design does include a +3C warming treatment, and the potential affects of warming on leaf physiology are mentioned in the introduction. As there were very few effects of climate warming observed in this aspect of the experiment, we feel it is important to address a possible explaination for the lack of observed warming effects using published findings from this experiment.

- Q: Nitrogen concentration in the leaves had a weak postive relationship with measured rates of mesophyll conductance if all leaves were considered. This weak relationship disappeared if sun and shade leaves were analyzed seperately. Since we can only speculate about the effect of aquaporins we do not include gm v. N in the manuscript.

- line 404-415:

- line 412-415: *REMKO*

- Q: On average mesophyll conductance and stomatal might be expected to be correlated when investigating interspecific comparisions, such as in Flexas et al. 2013 (Figure 1). The dynamic responses of leaf physiolgy to environmental conditions within a species/canopy are likely to alter this relationship. In this experiment, you can see a strong postive relationship when only sun leaves are considered (which is commonly what most experiments report), similar to the Flexas paper. However, due to the unexpected stomatal behavior in shade leaves this relationship was reversed. We have attached a quick figure below for the referee, but have not included it in the manuscript as it can be inferred from Table 1 and Figure 3 if necessary.

\
```{r gmgs, fig.align='center', fig.width=5, fig.height=5, echo=FALSE,warning=FALSE,dev="png",message=FALSE, dpi=600}
source("master_scripts/gmgs_referee2.R")
```

### 2016-08-26 gm_pce/review_response.Rmd
new leafK values expressed in manuscript table, not yet in results or text
add reviewer comments

@@
 *Editor*
+- 1.
+These leaves studied in this experiment juvenille leaves, as the *E.
+tereticornis* trees were grown for ~ 1 year.
+The fact that LMA was similar within the canopy may not be uncommon for this leaf type and has also been noted in Eucalpt plantations (Nouvellon et al.
+2010, cited in discussion).
+However, we have expanded the discussion to address fixed vs variable contols of leaf physiolgy specific to LMA with additional citiations (see reply to Referee # 2 specific comments).
+
+\
+- 2.
+We feel that colors are necessary in this document to visualize leaf types (+light), while also showing temperature treatments.
+It is near impossible to clearly distinguish all of these with symbols alone, and complicates visualization of confidence intervals on scatter plots.
+We have attempted to tone-done the color palette, while also distguishing colors properly for color-blind readers.
+
+\
+- 3/4.
+After careful thought we have decided to move the information regarding parameters derived ACi curves to the supplement.
+We only want to show that sun and shade leaves were fundamentaly different in their photosyntheitic capacity.
+As such, we have decided to use this data to generate Amax values at a common C~i~ between the leaf types.
+This very simply shows the point we aim to make.
+If the derivation and representation of J~max~/V~cmax~ values has distracted the editor from this point, it is likely to also do so to other readers who are plant physiolgists.
+These two parameters will be reported in the supplementary as additional evidence supporting significant difference in photosynthetic capacity between leaf types.
+
+*Remko*The fitaci package used here correctly estimates J~max~.
+
+We also acknowledge that AC~i~ response curves were intitally conceived with the notion that g~m~ was small enough to be ignored,     which we now know to not be true.
+We spent consideral time fitting AC~C~ curves that generated unrealistic V~cmax~ estimates of up to ~300 mumols m^-2^s^-1^.
+This will likely be a new issue for the plant physiolgy community as more g~m~ data are generated, and we hope are datasets will eventually help contribute to a solution.
+
+\
+-  5. depth of leaf ----  llyod paper appendix ---add to text as well
 
 *Referee #1*
+
+- 1. This experiment was harvested in May 2014, shortly after our measurements ceased.  Therefore, the suggestion to make diurnal measuremnts is no longer possible.
+
+\
+- 2.
+Leaves used for gas exchange were sampled and then measured for midday water potentails.
+This is explained in the first sentence of the subsection "Leaf Nitrogen and hydraulic conductance" in the methods.
+The method used here for measureing leaf hydraulic conductance is also one of the three methods described in Flexas et al.
+2013, which compared the coordination of leaf hydraulic conductance and mesophyll conductance.
+
+\
+- 3.
+Due to Referee #1 concern over cuvette measured transpiration, water potentials and subsequent leaf K calculations, we re-examined the calculation and analyses in the manuscript.
+Initially, leaf K was calculated with transpiration rates in shade leaves under low light conditions.
+As light was then increased in shade leaves and then transpiration re-measured, it is entirely possible that this change could have altered the water status of the leaf and thus subsequent water potential measurements (which we can not determine).
+As such we feel that the only leaf K which can properly reported is that of sun leaves vs shade leaves under high light.
+The data and results in the manuscript now reflect this change.
+This new data allows us to address aspects of Flexas et al.
+2013 (and others) mentioned above, but partially inhibits our ability to test optimization theories from Peltoniemi et al.
+2012.
 
@@
 __Specific Comments:__
+
+\
 - line 358:We corrected the text to match stats from Table 2
 
+\
 - line 392: We agree that thermal acclimation is not previosuly discussed before this point is made.
@@
 
+\
 - Q: Nitrogen concentration in the leaves had a weak postive relationship with measured rates of mesophyll conductance if all leaves were considered.
 This weak relationship disappeared if sun and shade leaves were analyzed seperately.
-Since we can only speculate about the effect of aquaporins we do not include gm v.
+Since we can only speculate about the effect of aquaporins we have not added gm v.
 N in the manuscript.
 
-- line 404-415:
+\
+- line 404-405:
 
+\
 - line 412-415: *REMKO*
 
+\
 - Q: On average mesophyll conductance and stomatal might be expected to be correlated when investigating interspecific comparisions, such as in Flexas et al.

### 2016-08-26 master_scripts/resource_table.R

@@ -36,7 +36,7 @@ canopy_agg <- summaryBy(lma+leafN_area+ c13 ~ leaf + temp, data=canopy_chem2, FU
 resoucetab <- merge(aci_agg, canopy_agg)
 
 #### leaf K and water potential
-waterdat <- read.csv("calculated_data/leafK_nodrought.csv")
+waterdat <- read.csv("calculated_data/leafK_nodrought_highlight.csv")
 water_agg <- summaryBy(pre_mp+mid_mp+leafK ~ leaf+temp, data=waterdat, FUN=c(mean,se))
 
 resoucetab <- merge(resoucetab, water_agg)

### 2016-08-26 stats/leafchemistry_stats.R

@@ -74,22 +74,31 @@ anova(lma_sha_temp)
 ####leaf K---------------------------------------------------------------------------------------------------------------
 
 leafK <- read.csv("calculated_data/leafK_nodrought.csv")
+#with highlight instead
+leafK2 <- read.csv("calculated_data/leafK_nodrought_highlight.csv")
 
 #examite data with boxplots, then remove any outliers
 boxplot(leafK~temp, data=leafK[leafK$leaf =="sun",])
 boxplot(leafK~temp, data=leafK[leafK$leaf =="shade",])
 
-leafK_clean <- leafK[leafK$leafK <=4.75,]
+boxplot(leafK~temp, data=leafK2[leafK2$leaf =="sun",])
+boxplot(leafK~temp, data=leafK2[leafK2$leaf =="shade",])
 
-leafK_leaf <- lme(leafK ~ leaf, random=~1|chamber, data=leafK_clean)
+leafK_clean <- leafK[leafK$leafK <=4.75,]
+leafK2_clean <- leafK2[leafK2$leafK <=4.75,]
+
+leafK_leaf <- lme(leafK ~ leaf, random=~1|chamber, data=leafK2_clean)
 summary(leafK_leaf)
 anova(leafK_leaf)
+visreg(leafK_leaf)
 
-leafK_sun_temp <- lme(leafK ~ temp ,random=~1|chamber, data=leafK_clean, subset=leaf=="sun")
+
+##no temperature effects
+leafK_sun_temp <- lme(leafK ~ temp ,random=~1|chamber, data=leafK2_clean, subset=leaf=="sun")
 summary(leafK_sun_temp)
 anova(leafK_sun_temp)
 
-leafK_sha_temp <- lme(leafK ~ temp ,random=~1|chamber, data=leafK_clean, subset=leaf=="shade")
+leafK_sha_temp <- lme(leafK ~ temp ,random=~1|chamber, data=leafK2_clean, subset=leaf=="shade")
 summary(leafK_sha_temp)
 anova(leafK_sha_temp)
 

### 2016-08-29 gm_pce/manuscript_gmes.Rmd
started reviewer response RMD started edits in manuscript
new amax figure and results
aci moved to supp with methods and table....no figure

@@
 parfig <- "1"
-acifig <- "2"
+amaxfig <- "2"
 agmgsfig <- "3"
@@
 get2fig <- "S4"
+acifig <- "S5"
 
@@
 Each A-*C*~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ with leaf temperature (*T*~leaf~) held constant at 25 &deg;C, and saturating PPFD (1800 $\mu$mols m^-1^ s^-1^).
-From these curves the photosynthetic parameters, *J*~max~ and *V*~cmax~, were quantified using the photosynthesis model of Farquhar et al.
-[-@farquhar1980biochemical].
-We used the 'fitaci' function in the 'plantecophys' package in R, see Duursma [-@duursma2015] for detailed description of the fitting methods.
-This method uses non-linear regression to estimate *J*~max~, *V*~cmax~ and *R*~dark~ simultaneously.
-Because there was considerable practical difficulty in fitting A-*C*~c~ curves, we report standard *C*~i~-based *J*~max~ and *V*~cmax~, which effectively include a mesophyll conductance component.
-When fitting the A-*C*~i~ curves, we calculated $\Gamma_{*}$ with the same methods as for *g*~m~, described above (the estimated value at 25&deg;C was 38.9 $\mu$mol mol^-1^), and the combined Michaelis-Menten coefficient (*K*~m~) was estimated from Medlyn et al.
-[-@medlyn2002temperature] (713.4 $\mu$mol mol^-1^ at 25&deg;C, using a *K*~c~ of 405 $\mu$mol mol^-1^, with an atmospheric pressure of 100 kPa).
-The temperature response parameters for *J*~max~, *V*~cmax~ and *K*~m~ are not relevant since *T*~leaf~ was kept constant at 25&deg;C.
+From these curves the photosynthetic capacity of each leaf (*A*~max~) was determined as the maximum rate of photosynthesis measured at the highest observed *C*~i~.
+The photosynthetic parameters, *J*~max~ and *V*~cmax~, were also generated from each A-*C*~i~ curve and their derivations are presented in Supporting Information Methods S2.
 
@@
 Across six measurement campaigns over the 7 month period, PPFD was reduced on average by >75% in the shade (Figure `r parfig`).
-Leaf-specific hydraulic conductance (*K*~L~) was similar across sun and shade leaves (Table `r wtc3tab1`).
-This was because neither leaf water potentials ($\Psi_{pd}$ and $\Psi_L$, Table `r wtc3tab1`) nor transpiration rates (*E*~L~ , Table `r wtc3tab2`) differed between leaf types.
 Leaf *N*~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table `r wtc3tab1`).
 Leaf mass per area (LMA) was not different between leaf types (Table `r wtc3tab1`).
+
+*Leaf-specific hydraulic conductance (*K*~L~) was similar across sun and shade leaves (Table `r wtc3tab1`).
+This was because neither leaf water potentials ($\Psi_{pd}$ and $\Psi_L$, Table `r wtc3tab1`) nor transpiration rates (*E*~L~ , Table `r wtc3tab2`) differed between leaf types.*
+
  No effect of the warming treatment was detected on  PPFD, $\Psi_{pd}$, $\Psi_L$, *K*~L~, *E*~L~, *N*~a~ or LMA either within or across leaf types (P > 0.05).
@@
 ##Photosynthetic capacity and leaf photosynthesis rates 
-The photosynthetic parameters *J*~max~ and *V*~cmax~ were higher in sun compared to shade leaves (Table `r wtc3tab1`), as estimated from A-*C*~i~ curves measured at 25 &deg;C (Figure `r acifig`a).
-Within leaf types, no effect of the warming treatment was detected on either parameter.
-Among the sampled leaves, *V*~cmax~ was positively related to leaf *N*~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).
+Mean photosynthetic capacity (*A*~max~) of sun leaves  was signifcantly higher (+27%, P = 0.001) compared to shade leaves, when measured at saturating light and 25 &deg;C.
+Within leaf types, no effect of the warming treatment was detected on *A*~max~.
+*A*~max~ was positively related to leaf *N*~a~ across leaf types and temperature treatments (P = 0.01, Figure `r amaxfig`).
+Similarily, the the photosynthetic parameters *J*~max~ and *V*~cmax~ were higher in sun compared to shade leaves (Supplementary Table 1).
 
@@
 Following the increase in PPFD from low to high light on the same leaf, *g*~m~ values of shade leaves increased on average 55% after approximately 25 min (Figure `r gmshadefig`).
-These measured values of *g*~m~ for shade leaves at high light were also equivalent or greater than those of sun leaves (Table `r wtc3tab2`).
+These measured values of *g*~m~ for shade leaves at high light were also indistinguishable than those of sun leaves (Table `r wtc3tab2`).
 Proportional increases in *g*~m~ were matched by proportional increases in *A*~n~ from low to high light in shade leaves (Figure `r gebarfig`b,c).
@@
 ```{r, echo=FALSE}
-resourcetab2 <- rbind(resourcetab, c("P value","", as.character(pval)))
+resourcetab2 <- rbind(resourcetab[, c(1:4,7:10)], c("P value","", as.character(pval)))
 greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
@@
 dust(resourcetab2)%>%
-  sprinkle_colnames("Leaf", "Temperature", "LMA", "*N*~a~", "*V*~cmax~", "*J*~max~","K",paste0("$\\Psi$","~pd~"),paste0("$\\Psi$","~L~"), paste0("$\\delta$","^13^C")) %>%
+  sprinkle_colnames("Leaf", "Temperature", "LMA", "*N*~a~", "K",paste0("$\\Psi$","~pd~"),paste0("$\\Psi$","~L~"),  paste0("$\\delta$","^13^C")) %>%
   sprinkle(rows= 1, part="head",bold=TRUE)%>%
@@
 
-```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png",message=FALSE, dpi=600, dev.args=list(type="cairo"), eval=includeFigs}
+```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png",message=FALSE, dpi=600, eval=includeFigs}
 source("master_scripts/par2.R")
@@
 \
-```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600,message=FALSE, dev.args=list(type="cairo"), eval=includeFigs}
+```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600,message=FALSE, eval=includeFigs}
  source("master_scripts/aci_vcmaxnitro.R")
@@
 **Figure 2**.
-Photosynthetic capacity, photosynthesis rate and leaf nitrogen for sun and shade leaves.
-(a) A-*C*~i~ curves for sun and shade leaves grown under elevated (ET) and ambient (AT) temperature treatments.
-A-*C*~i~ curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-(b) The relationship between *V*~cmax~ and mean leaf N per area (*N*~a~) for each chamber, including sun leaves and shade leaves at low light.
-(c) The relationship between leaf net photosynthesis rate (*A*~n~) and leaf *N*~a~ for sun and shade leaves measured under their ambient light and temperature conditions.
-For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional *R*^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
+The relationship between mean leaf N per area (*N*~a~) for each chamber and photosynthetic capacity (*A*~max) for sun and shade leaves grown under elevated (ET) and ambient (AT) temperature treatments.
+*A*~max represents the net photosynthetic rate at the highest C~i~ values taken from A-*C*~i~ curves measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+The dashed line represents the significant linear model fit for all leaves, with a marginal and conditional *R*^2^ of 0.59 and 0.62.
 \
-```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo"), eval=includeFigs}
+```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, eval=includeFigs}
 source("master_scripts/Agmgs.R")

### 2016-08-29 gm_pce/manuscript_gmes_SuppInfo.Rmd
new amax figure and results
aci moved to supp with methods and table....no figure

@@
 
+## Tables
+*jmax vcmax values*
+**Table S1**.
+Leaf morphological and physiological traits of sun and shade leaves under ambient and elevated temperature treatments for *Eucalyptus tereticornis*.
+Values of *V*~cmax~ and *J*~max~ are treatment mean (&pm; 1 standard error) from A-*C*~i~ curves measured in each chamber at 25 &deg;C and saturating light.
+Different letters represent significant differences between leaf type and temperature treatments (based on a Tukey test).
+The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.
+
+```{r, echo=FALSE}
+resourcetab2 <- rbind(resourcetab[,c(1:2, 5:6)], c("P value","", as.character(pval_aci)))
+
+dust(resourcetab2)%>%
+  sprinkle_colnames("Leaf", "Temperature", "*V*~cmax~", "*J*~max~") %>%
+  sprinkle(rows= 1, part="head",bold=TRUE)%>%
+  sprinkle_print_method("markdown")
+```
+
 ## Figures
 
-```{r airvars, fig.width=7, fig.height=8.8, echo=FALSE,warning=FALSE,dev="png", dpi=600,  eval=includeFigs, dev.args=list(type="cairo")}
+```{r airvars, fig.width=7, fig.height=8.8, echo=FALSE,warning=FALSE,dev="png", dpi=600,  eval=includeFigs}
 source("master_scripts/air_vars.R")
@@
 \
-```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, eval=includeFigs, dev.args=list(type="cairo")}
+```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, eval=includeFigs}
 source("master_scripts/delta_cica.R")
@@
 \
-```{r gasex_temp3, fig.width=7, fig.height=9, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo"), eval=includeFigs}
+```{r gasex_temp3, fig.width=7, fig.height=9, echo=FALSE,warning=FALSE,dev="png", dpi=600, eval=includeFigs}
 source("master_scripts/gasex_temp3.R")
@@
 \
-```{r gasex_temp4, fig.width=7, fig.height=9, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo"), eval=includeFigs}
+```{r gasex_temp4, fig.width=7, fig.height=9, echo=FALSE,warning=FALSE,dev="png", dpi=600, eval=includeFigs}
 source("master_scripts/gasex_temp4.R")
@@
 Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (*R*^2^ = 0.73, 0.58 and 0.72, respectively).
+
+<!-- \ -->
+<!-- ```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600,message=FALSE, eval=includeFigs} -->
+<!--  source("master_scripts/aci_vcmaxnitro.R") -->
+<!-- ``` -->
+<!-- \ -->
+<!-- **Figure S5**.
+Photosynthetic capacity, photosynthesis rate and leaf nitrogen for sun and shade leaves.
+(a) A-*C*~i~ curves for sun and shade leaves grown under elevated (ET) and ambient (AT) temperature treatments.
+A-*C*~i~ curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+(b) The relationship between *V*~cmax~ and mean leaf N per area (*N*~a~) for each chamber, including sun leaves and shade leaves at low light.
+For (b) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional *R*^2^ of 0.28 and 0.35.
+-->
 
@@
 
+##Methods S2
+
+###Photosynthetic parameters from A-*C*~i~ curves
+From each A-*C*~i~ curve the photosynthetic parameters, *J*~max~ and *V*~cmax~, were quantified using the photosynthesis model of Farquhar et al.
+[-@farquhar1980biochemical].
+We used the 'fitaci' function in the 'plantecophys' package in R, see Duursma [-@duursma2015] for detailed description of the fitting methods.
+This method uses non-linear regression to estimate *J*~max~, *V*~cmax~ and *R*~dark~ simultaneously.
+Because there was considerable practical difficulty in fitting A-*C*~c~ curves, we report standard *C*~i~-based *J*~max~ and *V*~cmax~, which effectively include a mesophyll conductance component.
+When fitting the A-*C*~i~ curves, we calculated $\Gamma_{*}$ with the same methods as for *g*~m~, described above (the estimated value at 25&deg;C was 38.9 $\mu$mol mol^-1^), and the combined Michaelis-Menten coefficient (*K*~m~) was estimated from Medlyn et al.
+[-@medlyn2002temperature] (713.4 $\mu$mol mol^-1^ at 25&deg;C, using a *K*~c~ of 405 $\mu$mol mol^-1^, with an atmospheric pressure of 100 kPa).
+The temperature response parameters for *J*~max~, *V*~cmax~ and *K*~m~ are not relevant since *T*~leaf~ was kept constant at 25&deg;C.
+
 # References {.unnumbered}

### 2016-08-29 master_scripts/aci_vcmaxnitro.R

@@ -123,30 +123,20 @@ plot(Photo~Ci ,data= acishade_clean, col=lightscol, ylim=c(0, 42), xlim=c(0,2000
   
   legend("topleft", c("Sun", "Shade-High Light", "AT", "ET"), pch=c(-1,-1, 16,17), 
          lty=c(1,1, 1,2), lwd=2,
-         col=c(suncol, lightscol2, "black", "black"),inset = 0.01, bty='n',cex=.7)
-  text(x=2040, y=42, "(a)", cex=.7)
+         col=c(suncol, lightscol2, "black", "black"),inset = 0.01, bty='n',cex=.8)
+  text(x=2020, y=42, "(a)", cex=1)
   
 ###panel 2: leafNarea ~ vcmax (bottom left)
-par(fig=c(0, .5, 0, .5),new=T,mar=c(4,4,1,1), cex=1.25, las=1, cex.axis=.8, cex.lab=1, mgp=c(2.5,1,0))
+par(fig=c(0, 1, 0, .5),new=T,mar=c(4,4,1,1), cex=1.25, las=1, cex.axis=.8, cex.lab=1, mgp=c(2.5,1,0))
 
-plot(Vcmax~leafN_area.mean, data=N_aci, col=as.factor(leaf),  ylim=c(40, 133),xlim=c(1,3.5),
+plot(Vcmax~leafN_area.mean, data=N_aci, col=as.factor(leaf),  ylim=c(40, 133),xlim=c(1.5,3.5),
        pch=c(16, 17)[pch=N_aci$temp],xlab=narealab, ylab=vclab)
   ablineclip(nitrovc_mod, x1=min(N_aci$leafN_area.mean), x2=max(N_aci$leafN_area.mean), lwd=2, lty=3)
   legend("topleft", leglab2, pch=c(16,17,16,17), col=trtcols,inset = 0.01, bty='n',cex=.7)
-  text(x=3.475, y=133, "(b)", cex=.7)
-  
-####panel bottom right 
-par(fig=c(.5, 1, 0, .5),new=T,mar=c(4,4,1,1), cex=1.25, las=1, cex.axis=.8, cex.lab=1, mgp=c(2.5,1,0))
-plot(Photo ~ leafN_area, data=Anitro[Anitro$leaflight=="sun-high",], col=suncol,  xlim=c(0,4), ylim=c(0,25),
[…]

### 2016-08-30 gm_pce/review_response.Rmd
fixed versus variable gm

@@
 tereticornis* trees were grown for ~ 1 year.
-The fact that LMA was similar within the canopy may not be uncommon for this leaf type and has also been noted in Eucalpt plantations (Nouvellon et al.
+The fact that LMA was similar within the canopy may be common in juvenille leaves and similar LMA within canopies has also been noted in Eucalpt plantations (Nouvellon et al.
 2010, cited in discussion).
@@
 We feel that colors are necessary in this document to visualize leaf types (+light), while also showing temperature treatments.
-It is near impossible to clearly distinguish all of these with symbols alone, and complicates visualization of confidence intervals on scatter plots.
-We have attempted to tone-done the color palette, while also distguishing colors properly for color-blind readers.
+It is very messy to distinguish all of these with symbols alone, and complicates visualization of confidence intervals on scatter plots.
+We have attempted to tone-done the color palette, while still distguishing colors properly for color-blind readers.
 
@@
 - 3/4.
-After careful thought we have decided to move the information regarding parameters derived ACi curves to the supplement.
+After careful thought we have decided to move the information regarding parameters derived AC~i~ curves to the supplementary information.
 We only want to show that sun and shade leaves were fundamentaly different in their photosyntheitic capacity.
-As such, we have decided to use this data to generate Amax values at a common C~i~ between the leaf types.
+As such, we have decided present Amax values, from these CO~2~ response curves, between the leaf types.
 This very simply shows the point we aim to make.
-If the derivation and representation of J~max~/V~cmax~ values has distracted the editor from this point, it is likely to also do so to other readers who are plant physiolgists.
-These two parameters will be reported in the supplementary as additional evidence supporting significant difference in photosynthetic capacity between leaf types.
+If the derivation and representation of J~max~/V~cmax~ values has distracted the editor from this point, it is likely to also distract readers who are plant physiolgists.
+J~max~ and V~cmax~ values are still described and reported in the supplementary as additional evidence to support differences in photosynthetic capacity between leaf types.
 
@@
 
-We also acknowledge that AC~i~ response curves were intitally conceived with the notion that g~m~ was small enough to be ignored,     which we now know to not be true.
-We spent consideral time fitting AC~C~ curves that generated unrealistic V~cmax~ estimates of up to ~300 mumols m^-2^s^-1^.
-This will likely be a new issue for the plant physiolgy community as more g~m~ data are generated, and we hope are datasets will eventually help contribute to a solution.
+We also acknowledge that AC~i~ response curves were intitally conceived with the notion that g~m~ was small enough to be ignored,     which we now know is untrue.
+We spent consideral time attempting to fit AC~C~ curves, which generated unrealistic V~cmax~ estimates of up to ~300 mumols m^-2^s^-1^.
+This represents a new issue for the plant physiolgy community as more g~m~ data are generated, and we hope our datasets will eventually help contribute to a solution.
 
@@
 
-- 1. Wingate et al. 2007:
+- 1. Wingate et al. 2007: accountd for by evan and cammereer 2012? (griffiths paper)
 
@@
 Thus the WTC atmosphere or CO~2~ cylinder $\delta$^13^C should not interfer with steady state conditions of the leaf measuremnts.
+
+- 3. fixed vs variable controls of g~m~ .....
 
@@
 \
-- line 404-405:
+- line 404-405: added Brooks et al. 1996 reference explaining patterns of morphology and physiology of when sun leaves become shaded.
 

### 2016-08-30 gm_pce/manuscript_gmes.Rmd
fixed versus variable gm

@@
 ##Photosynthetic capacity and leaf photosynthesis rates 
-Mean photosynthetic capacity (*A*~max~) of sun leaves  was signifcantly higher (+27%, P = 0.001) compared to shade leaves, when measured at saturating light and 25 &deg;C.
+Mean photosynthetic capacity (*A*~max~) of sun leaves  was significantly higher (+27%, P = 0.001) compared to shade leaves, when measured at saturating light and 25 &deg;C.
 Within leaf types, no effect of the warming treatment was detected on *A*~max~.
 *A*~max~ was positively related to leaf *N*~a~ across leaf types and temperature treatments (P = 0.01, Figure `r amaxfig`).
-Similarily, the the photosynthetic parameters *J*~max~ and *V*~cmax~ were higher in sun compared to shade leaves (Supplementary Table 1).
+Similarly, the the photosynthetic parameters *J*~max~ and *V*~cmax~ were higher in sun compared to shade leaves (Supplementary Table 1).
 
@@
 Similar LMA in sun and shade leaves could be the result of leaf formation under comparable light conditions or possible unmeasured differences in total non-structural carbohydrates contents between leaf types that masked differences in leaf thickness or density.
+For example, sun leaves of *Abies amabilis* foliage that became shaded exhibited physiology similar to shaded leaves, while retaining the morphology of sun leaves [@brooks1996effects].
 
@@
 \
-We found that *A*~n~ and *g*~m~ scaled positively across leaf types (Fig.
+**We found that *A*~n~ and *g*~m~ scaled positively across leaf types (Fig.
 3b) and, unexpectedly, increased in shade leaves when light intensity was increased to values similar in sun leaves.
@@
 Our findings support growing evidence that *g*~m~ is highly variable and can respond to environmental variables [@flexas2007rapid; @flexas2008mesophyll].
-Here we provide empirical data showing *g*~m~ not only varies within a canopy, but that the up-regulation of *g*~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.
+Here we provide empirical data showing *g*~m~ not only varies within a canopy, but that the up-regulation of *g*~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.**
+
+*fixed vs variable controls of g~m~*
+We found that *A*~n~ and *g*~m~ scaled positively across leaf types (Fig.
+3b) and, unexpectedly, increased in shade leaves when light intensity was increased to values similar in sun leaves.
+Comparable LMA within the canopy likely allowed for this up-regulation of *g*~m~ , which played a critical role in the photosynthetic response of shade leaves to sunflecks.
+Components of LMA, such as palisade thickness and cell surface exposed to intercellular air space, set an upper limit for *g*~m~, but considerable variation in *g*~m~ can exist for a given LMA [@flexas2008mesophyll; @milla2016leaf].
+Possible mechanisms for the rapid (~25 min) up-regulation of *g*~m~ to light intensity include regulation of aquaporins, which can facilitate increases in the CO~2~ permeability of the cell membranes [@hanba2004overexpression; @heinen2009role; @li2014plant] or movement of chloroplasts to facilitate CO~2~ diffusion [@tholen2008].
+It has also been proposed that *g*~m~ and *K*~L~ should be coordinated, as optimizing hydraulic supply allows both *g*~s~ and *g*~m~ to enhance *A*~n~ [@flexas2013leaf].
+Here, higher *K*~l~ in shade leaves under high light lends supports to the potential coordination *K*~l~ and *g*~m~ and also to the association of higher *A*~n~ with higher *K*~l~ when light becomes saturating reported by @franks2006higher.
 

### 2016-08-31 gm_pce/manuscript_gmes.Rmd
removed Aci added amax kl and amax in table
new stats

started new gmes

manuscript edits

@@
 
-##Photosynthetic capacity from A-*C*~i~ curves
-Photosynthetic CO~2~ response (A-*C*~i~) curves were measured for one sun and shade leaf for each WTC in February 2014, when all trees were well-watered in each temperature treatment (*n*=6).
-Each A-*C*~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ with leaf temperature (*T*~leaf~) held constant at 25 &deg;C, and saturating PPFD (1800 $\mu$mols m^-1^ s^-1^).
-From these curves the photosynthetic capacity of each leaf (*A*~max~) was determined as the maximum rate of photosynthesis measured at the highest observed *C*~i~.
-The photosynthetic parameters, *J*~max~ and *V*~cmax~, were also generated from each A-*C*~i~ curve and their derivations are presented in Supporting Information Methods S2.
+##Photosynthetic capacity
+Photosynthetic capacity (*A*~max~) of sun and shade leaves were measured at saturating light and [CO~2~] for one sun and shade leaf for each WTC.
+Measurements were completed in February 2014, when all trees were well-watered in each temperature treatment (*n*=6).
+*A*~max~ measurements were made at a PPFD of 1800 $\mu$mol m^-1^ s^-1^ and [CO~2~] of 1800 $\mu$l l^-1^, with leaf temperature (*T*~leaf~) held constant at 25 &deg;C .
 
@@
 
+\
+As light intensity was increased for shade leaves before they were detached, the *E*~L~ measured at high light was used to to calculate *K*~L~ for all shade leaves.
+
+\
 Leaf level instantaneous transpiration efficiency (ITE) was calculated as *A*~n~ divided by *E*~L~.
@@
 Leaf mass per area (LMA) was not different between leaf types (Table `r wtc3tab1`).
-
-*Leaf-specific hydraulic conductance (*K*~L~) was similar across sun and shade leaves (Table `r wtc3tab1`).
-This was because neither leaf water potentials ($\Psi_{pd}$ and $\Psi_L$, Table `r wtc3tab1`) nor transpiration rates (*E*~L~ , Table `r wtc3tab2`) differed between leaf types.*
-
+Leaf-specific hydraulic conductance (*K*~L~) was significantly higher in shade leaves (+58%) compared to sun leaves (Table `r wtc3tab1`).
+This was due to differences in transpiration rates (*E*~L~ , Table `r wtc3tab2`) between leaf types at high light, as neither leaf water potentials ($\Psi_{pd}$ and $\Psi_L$, Table `r wtc3tab1`) differed between leaf types.
 No effect of the warming treatment was detected on  PPFD, $\Psi_{pd}$, $\Psi_L$, *K*~L~, *E*~L~, *N*~a~ or LMA either within or across leaf types (P > 0.05).
@@
 ##Photosynthetic capacity and leaf photosynthesis rates 
-Mean photosynthetic capacity (*A*~max~) of sun leaves  was significantly higher (+27%, P = 0.001) compared to shade leaves, when measured at saturating light and 25 &deg;C.
+Mean photosynthetic capacity (*A*~max~) of sun leaves  was significantly higher (+27%) compared to shade leaves, when measured at 25 &deg;C (Table `r wtc3tab1`).
 Within leaf types, no effect of the warming treatment was detected on *A*~max~.
-*A*~max~ was positively related to leaf *N*~a~ across leaf types and temperature treatments (P = 0.01, Figure `r amaxfig`).
-Similarly, the the photosynthetic parameters *J*~max~ and *V*~cmax~ were higher in sun compared to shade leaves (Supplementary Table 1).
+*A*~max~ was positively related to mean leaf *N*~a~ across leaf types and temperature treatments (P = 0.01, Figure `r amaxfig`).
 
@@
 Mean *A*~n~ was significantly higher in sun compared to shade leaves (+23%), when measured at their local light environment and temperatures (Table `r wtc3tab2`).
-Additionally, *A*~n~ was positively related to *N*~a~ across gas exchange campaigns and leaf types measured under ambient light and temperature conditions (P < 0.001, Figure `r acifig`c).
 Following an increase in light intensity to match high-light conditions ('sunfleck simulations'), *A*~n~ of shade leaves increased to values significantly greater than sun leaves at high light (P < 0.001, Table `r wtc3tab2`).
@@
 ```{r, echo=FALSE}
-resourcetab2 <- rbind(resourcetab[, c(1:4,7:10)], c("P value","", as.character(pval)))
-greeks=c(alpha='\u03b1', Delta='\u0394', Sigma='\u03A3', mu='\u03bc',delta = '\u03b4')
+resourcetab2 <- rbind(resourcetab, c("P value","", as.character(pval)))
 
 dust(resourcetab2)%>%
-  sprinkle_colnames("Leaf", "Temperature", "LMA", "*N*~a~", "K",paste0("$\\Psi$","~pd~"),paste0("$\\Psi$","~L~"),  paste0("$\\delta$","^13^C")) %>%
+  sprinkle_colnames("Leaf", "Temperature", "LMA", "*A*~max~", "*N*~a~", "*K*~l~",paste0("$\\Psi$","~pd~"),paste0("$\\Psi$","~L~"),  paste0("$\\delta$","^13^C")) %>%
   sprinkle(rows= 1, part="head",bold=TRUE)%>%

### 2016-08-31 gm_pce/manuscript_gmes_SuppInfo.Rmd
removed Aci added amax kl and amax in table
new stats

started new gmes

manuscript edits

@@
 source("gm_pce/manuscript_preload.R")
-```
-
-## Tables
-*jmax vcmax values*
-**Table S1**.
-Leaf morphological and physiological traits of sun and shade leaves under ambient and elevated temperature treatments for *Eucalyptus tereticornis*.
-Values of *V*~cmax~ and *J*~max~ are treatment mean (&pm; 1 standard error) from A-*C*~i~ curves measured in each chamber at 25 &deg;C and saturating light.
-Different letters represent significant differences between leaf type and temperature treatments (based on a Tukey test).
-The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.
-
-```{r, echo=FALSE}
-resourcetab2 <- rbind(resourcetab[,c(1:2, 5:6)], c("P value","", as.character(pval_aci)))
-
-dust(resourcetab2)%>%
-  sprinkle_colnames("Leaf", "Temperature", "*V*~cmax~", "*J*~max~") %>%
-  sprinkle(rows= 1, part="head",bold=TRUE)%>%
-  sprinkle_print_method("markdown")
 ```
@@
 Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (*R*^2^ = 0.73, 0.58 and 0.72, respectively).
-
-<!-- \ -->
-<!-- ```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600,message=FALSE, eval=includeFigs} -->
-<!--  source("master_scripts/aci_vcmaxnitro.R") -->
-<!-- ``` -->
-<!-- \ -->
-<!-- **Figure S5**.
-Photosynthetic capacity, photosynthesis rate and leaf nitrogen for sun and shade leaves.
-(a) A-*C*~i~ curves for sun and shade leaves grown under elevated (ET) and ambient (AT) temperature treatments.
-A-*C*~i~ curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
-(b) The relationship between *V*~cmax~ and mean leaf N per area (*N*~a~) for each chamber, including sun leaves and shade leaves at low light.
-For (b) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional *R*^2^ of 0.28 and 0.35.
--->
 

### 2016-08-31 master_scripts/amax_nitro.R

@@ -104,18 +104,18 @@ legend("topleft", leglab2, pch=c(16,17,16,17), col=c(suncol,suncol, lightscol, l
 # amax_temp <- lme(Photo ~ temp ,random=~1|chamber, data=amax_dat)
 # summary(amax_temp)
 # anova(amax_temp)
-# 
+#
 # ##full model (sun_shade diff)
 # amax_leaf <- lme(Photo ~ tukeyid, random=~1|chamber, data=amax_dat)
 # summary(amax_leaf)
 # anova(amax_leaf)
 # library(visreg)
 # visreg(amax_leaf)
-# 
+#
 # tukey_amax<- glht(amax_leaf, linfct = mcp(tukeyid = "Tukey"))
 # amax_siglets<- cld(tukey_amax)
 # amax_siglets2 <- amax_siglets$mcletters$Letters
-
-#write.csv(lma_siglets2, "master_scripts/sigletters/slr_amax.csv", row.names=FALSE)
-#amax_agg <- summaryBy(Photo ~ leaf, data=amax_dat, FUN=c(mean, se))
-  
+# 
+# write.csv(amax_siglets2, "master_scripts/sigletters/slr_amax.csv", row.names=FALSE)
+# amax_agg <- summaryBy(Photo ~ leaf+temp, data=amax_dat, FUN=c(mean, se))
+# write.csv(amax_agg, "calculated_data/amax.csv", row.names = FALSE)

### 2016-08-31 master_scripts/newleafK.R

@@ -94,16 +94,7 @@ library(multcomp)
 library(nlme)
 
 
-##1. lowlight shade K
-leafK_norm$tukeyid <- as.factor(paste(leafK_norm$leaf, leafK_norm$temp, sep="-"))
-
-  ###temp trt?  NO
-  knorm_temp <- lme(leafK ~ temp ,random=~1|chamber, data=leafK_norm)
-  summary(knorm_temp)
-  anova(knorm_temp)
-
-
-##2. highlight shade K
+##1. highlight shade K
 leafK_fleck$tukeyid <- as.factor(paste(leafK_fleck$leaf, leafK_fleck$temp, sep="-"))
 
   ###temp trt?  NO
@@ -112,15 +103,18 @@ leafK_fleck$tukeyid <- as.factor(paste(leafK_fleck$leaf, leafK_fleck$temp, sep="
   anova(kfleck_temp)
 
 
-##3. Full model on high light leaves for manuscript
+##2. Full model on high light leaves for manuscript
 k_leaf <- lme(leafK~ tukeyid, random=~1|chamber, data=leafK_fleck)
-summary(k_leaf)
-anova(k_leaf)
-visreg(k_leaf)
+  summary(k_leaf)
+  anova(k_leaf)
+  visreg(k_leaf)
 
 tukey_k<- glht(k_leaf, linfct = mcp(tukeyid = "Tukey"))
 k_siglets<- cld(tukey_k)
 k_siglets2 <- k_siglets$mcletters$Letters
+#letters wrong order????rearrange
+k_siglets3 <- c("a","a", "b","b")
+names(k_siglets3) <- names(k_siglets2)
 
-write.csv(k_siglets2, "master_scripts/sigletters/slr_k.csv", row.names=FALSE)
+write.csv(k_siglets3, "master_scripts/sigletters/slr_k.csv", row.names=FALSE)
 

### 2016-09-01 gm_pce/manuscript_gmes.Rmd
new gmes values with wingate correction from evans 2013 paper.
can now remake figures by simply replacing wellwatered gmes dfr with #2

@@
 \
-**We found that *A*~n~ and *g*~m~ scaled positively across leaf types (Fig.
-3b) and, unexpectedly, increased in shade leaves when light intensity was increased to values similar in sun leaves.
-Possible mechanisms for the rapid (~25 min) response of *g*~m~ to light intensity include regulation of aquaporins, which can facilitate increases in the CO~2~ permeability of the cell membranes [@hanba2004overexpression; @heinen2009role; @li2014plant] or movement of chloroplasts to facilitate CO~2~ diffusion [@tholen2008].
-Our findings support growing evidence that *g*~m~ is highly variable and can respond to environmental variables [@flexas2007rapid; @flexas2008mesophyll].
-Here we provide empirical data showing *g*~m~ not only varies within a canopy, but that the up-regulation of *g*~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.**
-
-*fixed vs variable controls of g~m~*
 We found that *A*~n~ and *g*~m~ scaled positively across leaf types (Fig.
@@
 **Figure 2**.
-The relationship between mean leaf N per area (*N*~a~) for each chamber and photosynthetic capacity (*A*~max) for sun and shade leaves grown under elevated (ET) and ambient (AT) temperature treatments.
-*A*~max represents the net photosynthetic rate at the highest C~i~ values taken from A-*C*~i~ curves measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
+Photosynthetic capacity (*A*~max) was positively correlated with mean leaf N per area (*N*~a~) for sun and shade leaves grown under elevated (ET) and ambient (AT) temperature treatments.
+Leaf *N*~a~ is the mean of each leaf type sampled across gas exchange campaigns in each whole-tree chamber.
+*A*~max was measured once on all trees for each leaf type, in February 2014, at 25°C and at both saturating light and [CO~2~].
 The dashed line represents the significant linear model fit for all leaves, with a marginal and conditional *R*^2^ of 0.59 and 0.62.

### 2016-09-01 gm_pce/manuscript_gmes_SuppInfo.Rmd
new gmes values with wingate correction from evans 2013 paper.
can now remake figures by simply replacing wellwatered gmes dfr with #2

@@
 
-##Methods S2
-
-###Photosynthetic parameters from A-*C*~i~ curves
-From each A-*C*~i~ curve the photosynthetic parameters, *J*~max~ and *V*~cmax~, were quantified using the photosynthesis model of Farquhar et al.
-[-@farquhar1980biochemical].
-We used the 'fitaci' function in the 'plantecophys' package in R, see Duursma [-@duursma2015] for detailed description of the fitting methods.
-This method uses non-linear regression to estimate *J*~max~, *V*~cmax~ and *R*~dark~ simultaneously.
-Because there was considerable practical difficulty in fitting A-*C*~c~ curves, we report standard *C*~i~-based *J*~max~ and *V*~cmax~, which effectively include a mesophyll conductance component.
-When fitting the A-*C*~i~ curves, we calculated $\Gamma_{*}$ with the same methods as for *g*~m~, described above (the estimated value at 25&deg;C was 38.9 $\mu$mol mol^-1^), and the combined Michaelis-Menten coefficient (*K*~m~) was estimated from Medlyn et al.
-[-@medlyn2002temperature] (713.4 $\mu$mol mol^-1^ at 25&deg;C, using a *K*~c~ of 405 $\mu$mol mol^-1^, with an atmospheric pressure of 100 kPa).
-The temperature response parameters for *J*~max~, *V*~cmax~ and *K*~m~ are not relevant since *T*~leaf~ was kept constant at 25&deg;C.
-
 # References {.unnumbered}

### 2016-09-01 functions and packages/gmes_functions.R

@@ -240,7 +240,7 @@ gmcalc_func2 <- function(x, a=4.4, ab= 2.9, e=30, b=29, f=16.2,del_growth = -8 ,
   #fractionation where Ci=Cc in the absence ot respiratory fractionation
   x$delta_i <- (x$t2*x$a_prime)+(x$t2*((1+x$t)*b-x$a_prime)*x$CiCa)
   
-  #fractionation associated with the diffusion of CO2 from intercellular airsoace to chloroplast 
+  #fractionation associated with the diffusion of CO2 from intercellular airspace to chloroplast 
   x$delta_gm = x$t3*(b - x$a_prime - (e*x$Rd)/(x$Photo+x$Rd)) * (x$Photo/(x$gm * x$CO2R))
   
   #most of the fractionation associated with respiration
@@ -256,8 +256,8 @@ gmcalc_func2 <- function(x, a=4.4, ab= 2.9, e=30, b=29, f=16.2,del_growth = -8 ,
 ### new e
 gmcalc_func3 <- function(x, a=4.4, ab= 2.9, b=29, f=16.2,del_growth = -8 , delR=-5, 
                         k25r=0.728, k25g=38.89, Ea_r = 72.311, Ea_g = 20.437,Rgc=8.314472){
-  
-  e = del_growth - delR
+
+  e = delR - del_growth
   
   x$CiCa <- x$Ci/x$CO2R
   x$a_prime <- (ab*(x$CO2S-x$C2sfc)+a*(x$C2sfc-x$Ci))/(x$CO2S-x$Ci)
@@ -286,5 +286,34 @@ gmcalc_func3 <- function(x, a=4.4, ab= 2.9, b=29, f=16.2,del_growth = -8 , delR=
   
   x$gm <- x$t3 * (b - 1.8 - x$Rd * e / (x$Rd+x$Photo)) * x$Photo/x$CO2S/(x$DiminusDo - x$rd_term2 - x$f_term2)
   x$gm_bar <- x$gm*100/x$Press
+  
+  #different fractionation components as outputs--------------------------------------------------------------------
+  
+  #fractionation where Ci=Cc in the absence ot respiratory fractionation
+  x$delta_i <- (x$t2*x$a_prime)+(x$t2*((1+x$t)*b-x$a_prime)*x$CiCa)
+  
+  #fractionation associated with the diffusion of CO2 from intercellular airspace to chloroplast 
[…]

### 2016-09-01 gm_pce/review_response.Rmd
implemented new gm values, everywhere reviewer comments mostly addressed now

--- 2016-08-26 gm_pce/review_response.Rmd
@@
 ---
-title: "Response to referees"
-author: "Court Campany"
-date: "26 August 2016"
+title: 
+author: 
+date: 
 output: word_document
@@
 
-*Editor*
-
-*Referee #1*
-
-*Referee #2*
-
-- 1. Wingate et al. 2007:
+Keith A. Mott
+\
+Editor-in-Chief
+\
+Plant, Cell and Environment
 
 \
-- 2.
-The measurements were taken as a closed system with the coupled TDl-Licor 6400, independent from the whole-tree chamber conditions.
-Additionally, specifically choosen CO~2~ cyclinders used for the Licor 6440 were closer to ambient air 13C (~ 10&permil) instead of standard CO~2~ which are fossil fuel derived (~ 30&permil).
-Thus the WTC atmosphere or CO~2~ cylinder $\delta$^13^C should not interfer with steady state conditions of the leaf measuremnts.
+Dear Professor Mott,
 
-__Specific Comments:__
-- line 358:We corrected the text to match stats from Table 2
+We are grateful for the opportunity to resubmit our manuscript.
+We thank the reviewers for providing insightful comments to improve our manuscript.
+We have responded in detail to each comment, which we feel has improved the overall quality of the manuscript.
+Below, we have itemized our response to each coomment from the Editor and both reviewers.
 
+### **Editor**
+
+\
+- 1. *Where these mature/isobilateral leaves and does this affect the outcome?*
+\
+These leaves studied in this experiment were very liekly juvenille leaves, as the *E.
+tereticornis* trees were grown for ~ 1 year.
+Consequently, the fact that LMA was similar within the canopy is different from other reports as the editor notes.
+Similar LMA within canopies has been noted before in Eucalpt plantations (Nouvellon et al.
+2010), which is cited in the discussion.
+We have expanded the discussion to address the possible fixed contols of LMA on leaf physiology (see reply to Referee # 2 specific comments for full details).
+
+\
+- 2. *Stylistic concerns with colors* 
+\
+We feel that colors are necessary in this macnuscript to visualize leaf types (+light), while also differentiating temperature treatments.
+We attempted to use only symbols, which makes it neear impossible to distinguish all of the treatments.
+It also complicates visualization of confidence intervals on scatter plots.
+We have attempted to tone-done the color palette, while still distinguishing colors properly for color-blind readers.
+(*need to try colors again/drop orange??*)
+
+\
+- 3/4. *Concerns over parameters derived from AC~i~ curves* 
+\
+After careful thought we have decided to remove parameters derived from AC~i~ curves from the manuscript.
+Our main goal is to show that sun and shade leaves were fundamentaly different in their photosyntheitic capacity.
+As such, we have decided to instead present *A*~max~ values between the leaf types.
+This provides simple robust measurements of photosynthetic capacity, that reach our intended goal in a more straight forward manner.
+If the derivation and representation of *J*~max~/*V*~cmax~ values has distracted the editor from our main point, it is likely to also distract readers who are plant physiolgists.
+Figure 2.
+now shows the relationship between  Amax and leaf *N*~a~ and *A*~max~ values are reported in Table 1.
+
+\
+We also acknowledge that AC~i~ response curves were intitally conceived with the notion that g~m~ was small enough to be ignored,     which we now know is untrue.
+We spent consideral time attempting to fit AC~C~ curves, which often generated unrealistic V~cmax~ estimates.
+This represents a new issue for the plant physiolgy community as more g~m~ data are generated, and we hope our dataset will eventually help contribute to a solution.
+
+\
+- 5. *Does light penetration deeper in a shade leaf impact gm?* 
+\
+This question raises theoritical issues about how/if photosynthetic machinery are optimallty distributed along intraleaf light gradient.
+To our knowledge, these issues have yet to resolved empirically and are still a relevant question for most leaf-level experiments.
+For example, it is possible that Jmax/Vcmax could be partitioned within the leaf according to light gradients (see Llyod et al.
+1992), which could affect estimates of g~m~ at low light.
+Alternatively, greater light penetration in to the leaf but longer pathlengths for CO~2~ diffusion could interact to affect estimates of g~m~ at high light.
+Although we acknowledge that this question is relevant to this study and many others, we do not have the tools to address it further in the manuscript.
+
+### **Referee #1**
+
+\
+- 1. *Difficulty in conciliating optimal resource distribution without an enhancement of leaf hydraulic conductance*
+\
+Due to Referee #1 concern over cuvette measured transpiration, water potentials and subsequent leaf K calculations, we have re-examined the calculation and analyses in the manuscript.
+Initially, leaf K was calculated with transpiration rates in shade leaves under low light conditions.
+As shade leave were allowed to stablize under high light conditions before they were detached, the leaf water status reflected these conditions and not those of low light.
+We have corrected the calculation of *K*~l~ and made tje subsequent changes in the results and discussion of the manuscript.
+*K*~l~ is now signicantly higher in shade leaves, which makes sense in relation to the maintaing hydraullic supply to the observed stomatal behavior.
+We believe this actually makes the optimization of resources story much clearer throughout the manuscript, and we greatly appreciate the referee's comments.
+These corrected values now allow us to additionally address the possible coordination of *K*~l~ and *g*~m~ reported by Flexas et al.
+2013 in the discussion.
+
+\
+- 2. *Diurnal cycles of photosynthesis, stomatal conductance and light within the canopy* 
+\
+This experiment was harvested in May 2014, shortly after our measurements ceased. Therefore, the suggestion to make diurnal measuremnts is no longer possible.
+
+\
+- 3. *Concern over methods to measure leaf hydraulic conductance*
+\
+The method used here for measureing *K*~l~ is also one of the three methods described in Flexas et al.
+2013, which compared the coordination of leaf hydraulic conductance and mesophyll conductance.
+Leaves used for gas exchange were indeed sampled and then immediately measured for midday water potential.
+This is explained in the first sentence of the subsection "Leaf Nitrogen and hydraulic conductance" in the methods.
+
+### **Referee #2**
+
+\
+- 1. *Referee #2 correctly points out that the issues raised by Wingate et al. 2007 were not addressed*
+\
+We have recalculated the respiratory fractionation component 'e', according to Evans and von Caemmerer 2013, which addresses Wingate et al.
+2007.
+This correction increased values of *g*~m~ but does not alter observed patterns.
+
+\
+- 2. *Estimating turnover time of the pools*
+\
+We agree that this concept would be great for the plant physiological community.  However, we fill it would detract from main conclusions of this manuscript and we do not address this further.
+
+\
+- 3. *Was the chamber $\delta$^13^C measured?*
+\
+Measurements of chamber 13C throughout the day were measured for this experiment, but not at the same time as these measurements.
+A brief inspection of these data show that chamber $\delta$^13^C varied from around -9 to -12 &permil, likely depending on many factors.
+Although these values are different from the assumption that the $\delta$^13^C of the atmopshere is -8‰, we do not have a robust way to assign specific values to our measurements at each time point.
+Therefore we still assume that $\delta$^13^C of the atmopshere is -8‰ as most other studies in glass house conditions do also.
+Additionally, we used CO~2~ cyclinders for the Licor 6440 that were closer to much ambient air $\delta$^13^C instead of standard cylinders with  $\delta$^13^C > 30&permil, which also slightly improves calculations.
+
+\
+- 4. *Referee #1 Specific Comments:*
+
+\
+- line 358: We corrected the text to match stats from Table 2
+
+\
 - line 392: We agree that thermal acclimation is not previosuly discussed before this point is made.
@@
 
-- Q: Nitrogen concentration in the leaves had a weak postive relationship with measured rates of mesophyll conductance if all leaves were considered.
+\
+- Q: Nitrogen concentration in the leaves had a weak postive relationship with measured rates of *g*~m~ if all leaves were considered.
 This weak relationship disappeared if sun and shade leaves were analyzed seperately.
-Since we can only speculate about the effect of aquaporins we do not include gm v.
-N in the manuscript.
+Since we can only speculate about the effect of aquaporins we have not further addresssed *g*~m~ versus nitrogen in the manuscript.
 
-- line 404-415:
+\
+- line 404-405: added Brooks et al. 1996 citation explaining observed patterns of morphology and physiology when sun leaves become shaded.
 
+\
 - line 412-415: *REMKO*
 
+\
 - Q: On average mesophyll conductance and stomatal might be expected to be correlated when investigating interspecific comparisions, such as in Flexas et al.
 2013 (Figure 1).
-The dynamic responses of leaf physiolgy to environmental conditions within a species/canopy are likely to alter this relationship.
-In this experiment, you can see a strong postive relationship when only sun leaves are considered (which is commonly what most experiments report), similar to the Flexas paper.
-However, due to the unexpected stomatal behavior in shade leaves this relationship was reversed.
+The dynamic responses of leaf physiology to environmental conditions within a species/canopy are likely to alter this relationship.
+In this experiment, you can see a postive relationship when only sun leaves are considered, similar to the Flexas paper.
+Due to the unexpected stomatal behavior, however, this relationship was reversed in shade leaves.
 We have attached a quick figure below for the referee, but have not included it in the manuscript as it can be inferred from Table 1 and Figure 3 if necessary.

### 2016-09-01 master_scripts/Agmgs.R

@@ -3,7 +3,7 @@ source("master_scripts/plot_objects.R")
 source("functions and packages/packages.R")
 
 #read in gm data set (no drought) and Cibar(discrimination)-------------------------------------------------------
-gmes <- read.csv("calculated_data/gmes_wellwatered.csv")
+gmes <- read.csv("calculated_data/gmes_wellwatered2.csv")
 
 ###get average by id
 gm_agg <- summaryBy(Photo+ Cond + gm_bar ~ chamber+id+leaf +light+temp+leaflight+Month, data=gmes, FUN=mean, keep.names=TRUE)

### 2016-09-01 master_scripts/cicc_boxplots.R

@@ -5,7 +5,7 @@ source("master_scripts/plot_objects.R")
 source("functions and packages/packages.R")
 
 
-gasex <- read.csv("calculated_data/gmes_wellwatered.csv")
+gasex <- read.csv("calculated_data/gmes_wellwatered2.csv")
 
 ##calculate CC
 gasex$Cc<- with(gasex, Ci-Photo/gm)

### 2016-09-01 master_scripts/gasex_temp3.R

@@ -1,9 +1,9 @@
-# source("functions and packages/functions.R")
-# source("master_scripts/plot_objects.R")
-# source("functions and packages/packages.R")
+source("functions and packages/functions.R")
+source("master_scripts/plot_objects.R")
+source("functions and packages/packages.R")
 
 #read in gm data set (no drought) and Cibar(discrimination)-------------------------------------------------------
-gasex<- read.csv("calculated_data/gmes_wellwatered.csv")
+gasex<- read.csv("calculated_data/gmes_wellwatered2.csv")
   ##calculate CC
   gasex$Cc<- with(gasex, Ci-Photo/gm)
   ##calculate drawdown

### 2016-09-01 master_scripts/gmes_shadeleaves_figure.R

@@ -1,9 +1,9 @@
-# source("functions and packages/functions.R")
-# source("master_scripts/plot_objects.R")
-# source("functions and packages/packages.R")
+source("functions and packages/functions.R")
+source("master_scripts/plot_objects.R")
+source("functions and packages/packages.R")
 
 #read in gm data set (no drought) and Cibar(discrimination)-------------------------------------------------------
-gmes <- read.csv("calculated_data/gmes_wellwatered.csv")
+gmes <- read.csv("calculated_data/gmes_wellwatered2.csv")
 
 ###get average by id
 gm_agg <- summaryBy(Photo+ Cond + gm_bar ~ chamber+id+leaf +light+temp+leaflight+Month, data=gmes, FUN=mean, keep.names=TRUE)
@@ -11,7 +11,7 @@ gm_agg$leaf <- gsub("s", "S", gm_agg$leaf)
 gm_agg$light <- gsub("high", "Sun-light", gm_agg$light)
 gm_agg$light <- gsub("low", "Shade-light", gm_agg$light)
 
-###Im going to plot gmes of shade at different light levels, so seperate into two dfrs for ease (make sure they match correctly)
+###plot gmes of shade at different light levels, so seperate into two dfrs for ease (make sure they match correctly)
 
 shadat <- gm_agg[gm_agg$leaflight == "shade-low",]
 shadat <- droplevels(shadat)
@@ -27,6 +27,7 @@ names(fleckdat2)[4] <- "fleck_gm"
 
 #merge
 shadeleaf <- merge(shadat2, fleckdat2, by=c("chamber", "temp","Month"))
+# gm_increase <- (mean(shadeleaf$shade_gm)-mean(shadeleaf$fleck_gm))/mean(shadeleaf$shade_gm)
 
 ##plot
 # windows(7,7)

### 2016-09-01 master_scripts/gmgs_referee2.R

@@ -3,7 +3,7 @@ source("master_scripts/plot_objects.R")
 source("functions and packages/packages.R")
 
 ##read format phys data------------------------------------------------------------------------------------------------
-gasex <- read.csv("calculated_data/gmes_wellwatered.csv")
+gasex <- read.csv("calculated_data/gmes_wellwatered2.csv")
 gm_agg <- summaryBy(gm_bar+Cond+Photo ~ chamber+id+leaf +light+temp+leaflight+Month, 
                     data=gasex, FUN=mean, keep.names=TRUE)
 

### 2016-09-02 gm_pce/review_response.Rmd
colors removed from boxplots
wingate corrections made in response and in supp methods

@@
 We feel that colors are necessary in this macnuscript to visualize leaf types (+light), while also differentiating temperature treatments.
-We attempted to use only symbols, which makes it neear impossible to distinguish all of the treatments.
+We attempted to use only symbols, which makes it near impossible to distinguish all of the treatments.
 It also complicates visualization of confidence intervals on scatter plots.
-We have attempted to tone-done the color palette, while still distinguishing colors properly for color-blind readers.
+The choice of color palette is also used to distinguish colors properly for color-blind readers.
+We have dropped the use of colors for figures with boxplots, as they are not essential.
 (*need to try colors again/drop orange??*)
@@
 This provides simple robust measurements of photosynthetic capacity, that reach our intended goal in a more straight forward manner.
-If the derivation and representation of *J*~max~/*V*~cmax~ values has distracted the editor from our main point, it is likely to also distract readers who are plant physiolgists.
+If specific derivation or representations of *J*~max~/*V*~cmax~ values has distracted the editor from our main point, it is likely to also distract readers who are plant physiolgists.
 Figure 2.
-now shows the relationship between  Amax and leaf *N*~a~ and *A*~max~ values are reported in Table 1.
+now shows the relationship between Amax and leaf *N*~a~ and *A*~max~ values are also reported in Table 1.
 
 \
-We also acknowledge that AC~i~ response curves were intitally conceived with the notion that g~m~ was small enough to be ignored,     which we now know is untrue.
-We spent consideral time attempting to fit AC~C~ curves, which often generated unrealistic V~cmax~ estimates.
-This represents a new issue for the plant physiolgy community as more g~m~ data are generated, and we hope our dataset will eventually help contribute to a solution.
+We also acknowledge that AC~i~ response curves were intitally conceived with the notion that *g*~m~ was small enough to be ignored,     which we now know is untrue.
+This represents a new issue for the plant physiolgy community as more *g*~m~ data are continually  generated.
+We hope this dataset can eventually help be used to address this problem, but it is beyond the scope of this paper.
 
@@
 To our knowledge, these issues have yet to resolved empirically and are still a relevant question for most leaf-level experiments.
-For example, it is possible that Jmax/Vcmax could be partitioned within the leaf according to light gradients (see Llyod et al.
-1992), which could affect estimates of g~m~ at low light.
-Alternatively, greater light penetration in to the leaf but longer pathlengths for CO~2~ diffusion could interact to affect estimates of g~m~ at high light.
+For example, it is possible that *J*~max~/*V*~cmax~ could be partitioned within the leaf according to light gradients (see Llyod et al.
+1992), which could affect estimates of *g*~m~ at low light.
+Alternatively, greater light penetration in to the leaf but longer pathlengths for CO~2~ diffusion might interact to affect estimates of *g*~m~ at high light.
 Although we acknowledge that this question is relevant to this study and many others, we do not have the tools to address it further in the manuscript.
@@
 \
-Due to Referee #1 concern over cuvette measured transpiration, water potentials and subsequent leaf K calculations, we have re-examined the calculation and analyses in the manuscript.
-Initially, leaf K was calculated with transpiration rates in shade leaves under low light conditions.
+Due to Referee #1 concern over cuvette measured transpiration, water potentials and subsequent *K*~l~ calculations, we have re-examined the calculation and subsequent analyses in the manuscript.
+Initially, *K*~l~ was calculated with transpiration rates in shade leaves under low light conditions.
 As shade leave were allowed to stablize under high light conditions before they were detached, the leaf water status reflected these conditions and not those of low light.
-We have corrected the calculation of *K*~l~ and made tje subsequent changes in the results and discussion of the manuscript.
+We have corrected the calculation of *K*~l~ and made changes in the results and discussion sections.
 *K*~l~ is now signicantly higher in shade leaves, which makes sense in relation to the maintaing hydraullic supply to the observed stomatal behavior.
 We believe this actually makes the optimization of resources story much clearer throughout the manuscript, and we greatly appreciate the referee's comments.
-These corrected values now allow us to additionally address the possible coordination of *K*~l~ and *g*~m~ reported by Flexas et al.
-2013 in the discussion.
+These corrected values also allow us to address the possible coordination of *K*~l~ and *g*~m~ reported by Flexas et al.
+2013.
 
@@
 This correction increased values of *g*~m~ but does not alter observed patterns.
+The correct calculation of 'e' has been added to the supplementary methods.
+This correction also addresses comment #3 below.
 
@@
 \
-Measurements of chamber 13C throughout the day were measured for this experiment, but not at the same time as these measurements.
-A brief inspection of these data show that chamber $\delta$^13^C varied from around -9 to -12 &permil, likely depending on many factors.
-Although these values are different from the assumption that the $\delta$^13^C of the atmopshere is -8‰, we do not have a robust way to assign specific values to our measurements at each time point.
-Therefore we still assume that $\delta$^13^C of the atmopshere is -8‰ as most other studies in glass house conditions do also.
-Additionally, we used CO~2~ cyclinders for the Licor 6440 that were closer to much ambient air $\delta$^13^C instead of standard cylinders with  $\delta$^13^C > 30&permil, which also slightly improves calculations.
+Measurements of chamber $\delta$^13^C throughout the day were measured for this experiment during the same week as each gas exchange campaigns presented here.
+Since these measurements do not explicitly overlap, we do not have a robust way to assign specific values to each chamber at each time point.
+Instead, we took the mean chamber $\delta$^13^C for all chambers flux campaigns between 10am and 2 pm, to coincide with the timing of gas exchange campaigns.
+This gives a more realistic value than the steady state assumption of -8‰.
+This change is included in the corrections mentioned above.
+Additionally, we used CO~2~ cyclinders for the Licor 6400 that were much closer to ambient air $\delta$^13^C, instead of standard cylinders with  $\delta$^13^C > 30&permil, which slightly improves calculations.
 

### 2016-09-02 gm_pce/manuscript_gmes.Rmd
colors removed from boxplots
wingate corrections made in response and in supp methods

@@
 Mean *g*~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
-Following the increase in PPFD from low to high light on the same leaf, *g*~m~ values of shade leaves increased on average 55% after approximately 25 min (Figure `r gmshadefig`).
+Following the increase in PPFD from low to high light on the same leaf, *g*~m~ values of shade leaves increased on average 37% after approximately 25 min (Figure `r gmshadefig`).
 These measured values of *g*~m~ for shade leaves at high light were also indistinguishable than those of sun leaves (Table `r wtc3tab2`).

### 2016-09-02 gm_pce/manuscript_gmes_SuppInfo.Rmd
colors removed from boxplots
wingate corrections made in response and in supp methods

@@
 \
-```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, eval=includeFigs}
+```{r deltacica, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, eval=includeFigs, message=FALSE}
 source("master_scripts/delta_cica.R")
@@
 where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ is the fractionation from boundary layer diffusion (2.9&permil;) and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon].
+We calculate discrimination during day respiration (e) following Wingate et al.
+[-@wingate2007variations] as e = $\delta^13C_{tank}$ - $\delta^13C_{atmosphere}$.
+In this study, $\delta^13C_{tank}$ was -4 to -6&permil; and $\delta^13C_{atmosphere}$ was -12&permil;.
+$\delta^13C_{atmosphere}$ represents the mean whole-tree chamber $\delta^13C$ measured in a seperate experiment (data not shown), during similar diurnal time periods as the gas exchange campaigns presented here.
 *C*~a~ and *C*~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~).

### 2016-09-02 master_scripts/cicc_boxplots.R

@@ -31,19 +31,21 @@ cols <- c(shacol2, lightscol2,suncol2)
 # windows(12,6)
 par(mfrow=c(1,3),cex.axis=1.25, cex.lab = 1.5, mar=c(5,5,2,1))
 
-boxplot(Ci ~leaflight, gasex_agg, col=cols, ylab=expression(C[i]~~(paste(mu,bar,sep=""))), names=FALSE, ylim=c(0,325), outline=FALSE)
+#col=cols
+
+boxplot(Ci ~leaflight, gasex_agg, ylab=expression(C[i]~~(paste(mu,bar,sep=""))), names=FALSE, ylim=c(0,325), outline=FALSE)
   mtext("Shade \n Low Light", side=1,at=1, line=3.5)
   mtext("Shade \n High Light", side=1,at=2, line=3.5)
   mtext("Sun \n High Light", side=1,at=3, line=3.5)
   text(x=.55, y=325, "(a)", cex=1.5)
 
-boxplot(Cc ~leaflight, gasex_agg, col=cols, ylab=expression(C[c]~~(paste(mu,bar,sep=""))), names=FALSE,ylim=c(0,325),outline=FALSE)
+boxplot(Cc ~leaflight, gasex_agg,  ylab=expression(C[c]~~(paste(mu,bar,sep=""))), names=FALSE,ylim=c(0,325),outline=FALSE)
   mtext("Shade \n Low Light", side=1,at=1, line=3.5)
   mtext("Shade \n High Light", side=1,at=2, line=3.5)
   mtext("Sun \n High Light", side=1,at=3, line=3.5)
   text(x=.55, y=325, "(b)", cex=1.5)
 
-boxplot(drawdown~leaflight, gasex_agg, col=cols, ylab=drawdownlab3, xlab="", names=FALSE, ylim=c(0,325),outline=FALSE)
+boxplot(drawdown~leaflight, gasex_agg, ylab=drawdownlab3, xlab="", names=FALSE, ylim=c(0,325),outline=FALSE)
   mtext("Shade \n Low Light", side=1,at=1, line=3.5)
   mtext("Shade \n High Light", side=1,at=2, line=3.5)
   mtext("Sun \n High Light", side=1,at=3, line=3.5)

### 2016-09-02 gm_pce/review_response.Rmd
manuscript edits from reviewers now complete minus a few imputs from Remko
I did some heavy repo cleaning

@@
 We have responded in detail to each comment, which we feel has improved the overall quality of the manuscript.
-Below, we have itemized our response to each coomment from the Editor and both reviewers.
+Below, we have itemized our response to each comment from the Editor and both reviewers.
 
@@
 \
-These leaves studied in this experiment were very liekly juvenille leaves, as the *E.
+These leaves studied in this experiment were very likely juvenile leaves, as the *E.
 tereticornis* trees were grown for ~ 1 year.
@@
 2010), which is cited in the discussion.
-We have expanded the discussion to address the possible fixed contols of LMA on leaf physiology (see reply to Referee # 2 specific comments for full details).
+We have expanded the discussion to address the possible fixed controls of LMA on leaf physiology (see reply to Referee # 2 specific comments for full details).
 
@@
 \
-We feel that colors are necessary in this macnuscript to visualize leaf types (+light), while also differentiating temperature treatments.
+We feel that colors are necessary in this manuscript to visualize leaf types (+light), while also differentiating temperature treatments.
 We attempted to use only symbols, which makes it near impossible to distinguish all of the treatments.
@@
 The choice of color palette is also used to distinguish colors properly for color-blind readers.
-We have dropped the use of colors for figures with boxplots, as they are not essential.
+We have dropped the use of colors for figures with box plots, as they are not essential.
 (*need to try colors again/drop orange??*)
@@
 After careful thought we have decided to remove parameters derived from AC~i~ curves from the manuscript.
-Our main goal is to show that sun and shade leaves were fundamentaly different in their photosyntheitic capacity.
+Our main goal is to show that sun and shade leaves were fundamental different in their photosynthetic capacity.
 As such, we have decided to instead present *A*~max~ values between the leaf types.
 This provides simple robust measurements of photosynthetic capacity, that reach our intended goal in a more straight forward manner.
-If specific derivation or representations of *J*~max~/*V*~cmax~ values has distracted the editor from our main point, it is likely to also distract readers who are plant physiolgists.
+If specific derivation or representations of *J*~max~/*V*~cmax~ values has distracted the editor from our main point, it is likely to also distract readers who are plant physiologists.
 Figure 2.
-now shows the relationship between Amax and leaf *N*~a~ and *A*~max~ values are also reported in Table 1.
+now shows the relationship between *A*~max~ and leaf *N*~a~ and *A*~max~ values are also reported in Table 1.
 
 \
-We also acknowledge that AC~i~ response curves were intitally conceived with the notion that *g*~m~ was small enough to be ignored,     which we now know is untrue.
-This represents a new issue for the plant physiolgy community as more *g*~m~ data are continually  generated.
-We hope this dataset can eventually help be used to address this problem, but it is beyond the scope of this paper.
+We also acknowledge that AC~i~ response curves were initially conceived with the notion that *g*~m~ was small enough to be ignored,     which we now know is untrue.
+This represents a new issue for the plant physiology community as more *g*~m~ data are continually  generated.
+We hope this data set can eventually help be used to address this problem, but it is beyond the scope of this paper.
 
 \
-- 5. *Does light penetration deeper in a shade leaf impact gm?* 
+- 5. *Does light penetration deeper in a shade leaf impact mesophyll conductance?* 
 \
-This question raises theoritical issues about how/if photosynthetic machinery are optimallty distributed along intraleaf light gradient.
+This question raises theoretical issues about how/if photosynthetic machinery are optimally distributed along intraleaf light gradient.
 To our knowledge, these issues have yet to resolved empirically and are still a relevant question for most leaf-level experiments.
@@
 1992), which could affect estimates of *g*~m~ at low light.
-Alternatively, greater light penetration in to the leaf but longer pathlengths for CO~2~ diffusion might interact to affect estimates of *g*~m~ at high light.
+Alternatively, greater light penetration in to the leaf but longer path lengths for CO~2~ diffusion might interact to affect estimates of *g*~m~ at high light.
 Although we acknowledge that this question is relevant to this study and many others, we do not have the tools to address it further in the manuscript.
@@
 Initially, *K*~l~ was calculated with transpiration rates in shade leaves under low light conditions.
-As shade leave were allowed to stablize under high light conditions before they were detached, the leaf water status reflected these conditions and not those of low light.
+As shade leave were allowed to stabilize under high light conditions before they were detached, the leaf water status reflected these conditions and not those of low light.
 We have corrected the calculation of *K*~l~ and made changes in the results and discussion sections.
-*K*~l~ is now signicantly higher in shade leaves, which makes sense in relation to the maintaing hydraullic supply to the observed stomatal behavior.
+*K*~l~ is now significantly higher in shade leaves, which makes sense in relation to the maintain hydraulic supply to the observed stomatal behavior.
 We believe this actually makes the optimization of resources story much clearer throughout the manuscript, and we greatly appreciate the referee's comments.
@@
 \
-This experiment was harvested in May 2014, shortly after our measurements ceased. Therefore, the suggestion to make diurnal measuremnts is no longer possible.
+This experiment was harvested in May 2014, shortly after our measurements ceased. Therefore, the suggestion to make diurnal measurements is no longer possible.
 
@@
 \
-The method used here for measureing *K*~l~ is also one of the three methods described in Flexas et al.
+The method used here for measuring *K*~l~ is also one of the three methods described in Flexas et al.
 2013, which compared the coordination of leaf hydraulic conductance and mesophyll conductance.
@@
 \
-We have recalculated the respiratory fractionation component 'e', according to Evans and von Caemmerer 2013, which addresses Wingate et al.
+We have recalculated the respiratory fractional component 'e', according to Evans and von Caemmerer 2013, which addresses Wingate et al.
 2007.
@@
 This change is included in the corrections mentioned above.
-Additionally, we used CO~2~ cyclinders for the Licor 6400 that were much closer to ambient air $\delta$^13^C, instead of standard cylinders with  $\delta$^13^C > 30&permil, which slightly improves calculations.
+Additionally, we used CO~2~ cylinders for the Licor 6400 that were much closer to ambient air $\delta$^13^C, instead of standard cylinders with  $\delta$^13^C > 30‰, which slightly improves calculations.
 
@@
 \
-- line 392: We agree that thermal acclimation is not previosuly discussed before this point is made.
+- line 392: We agree that thermal acclimation is not previously discussed before this point is made.
 However, the experimental design does include a +3C warming treatment, and the potential affects of warming on leaf physiology are mentioned in the introduction.
-As there were very few effects of climate warming observed in this aspect of the experiment, we feel it is important to address a possible explaination for the lack of observed warming effects using published findings from this experiment.
+As there were very few effects of climate warming observed in this aspect of the experiment, we feel it is important to address a possible explanation for the lack of observed warming effects using published findings from this experiment.
 
 \
-- Q: Nitrogen concentration in the leaves had a weak postive relationship with measured rates of *g*~m~ if all leaves were considered.
-This weak relationship disappeared if sun and shade leaves were analyzed seperately.
-Since we can only speculate about the effect of aquaporins we have not further addresssed *g*~m~ versus nitrogen in the manuscript.
+- Q: Nitrogen concentration in the leaves had a weak postie relationship with measured rates of *g*~m~ if all leaves were considered.
+This weak relationship disappeared if sun and shade leaves were analyzed separately.
+Since we can only speculate about the effect of aquaporins we have not further addressed *g*~m~ versus nitrogen in the manuscript.
 
@@
 \
-- Q: On average mesophyll conductance and stomatal might be expected to be correlated when investigating interspecific comparisions, such as in Flexas et al.
+- Q: On average mesophyll conductance and stomatal might be expected to be correlated when investigating inter-specific comparisons, such as in Flexas et al.
 2013 (Figure 1).
 The dynamic responses of leaf physiology to environmental conditions within a species/canopy are likely to alter this relationship.
-In this experiment, you can see a postive relationship when only sun leaves are considered, similar to the Flexas paper.
+In this experiment, you can see a postie relationship when only sun leaves are considered, similar to the Flexas paper.
 Due to the unexpected stomatal behavior, however, this relationship was reversed in shade leaves.

### 2016-09-02 gm_pce/manuscript_gmes.Rmd
manuscript edits from reviewers now complete minus a few imputs from Remko
I did some heavy repo cleaning

@@
 The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with *C*~i~/*C*~a~ for both leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`).
-Carbon isotope discrimination associated with *g*~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments.
+Carbon isotope discrimination associated with *g*~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.3%) and varied little across measurement temperatures, leaf types, or warming treatments.
 The remainder of the variation in $\Delta_o$ consists of the contributions of *g*~s~, respiration and photorespiration to discrimination.
@@
 \
-Mean *g*~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).
+Mean *g*~m~ was higher in sun compared to shade leaves (+25%) under their local light environment (P < 0.001).
 Following the increase in PPFD from low to high light on the same leaf, *g*~m~ values of shade leaves increased on average 37% after approximately 25 min (Figure `r gmshadefig`).
@@
 No differences in *g*~m~ were detected with the warming treatment within leaf types.
-Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 for sun leaves, P = 0.04 for shade leaves).
+Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.002 for sun leaves, P = 0.044 for shade leaves).
 

### 2016-09-02 gm_pce/manuscript_gmes_SuppInfo.Rmd
manuscript edits from reviewers now complete minus a few imputs from Remko
I did some heavy repo cleaning

@@
 In this study, $\delta^13C_{tank}$ was -4 to -6&permil; and $\delta^13C_{atmosphere}$ was -12&permil;.
-$\delta^13C_{atmosphere}$ represents the mean whole-tree chamber $\delta^13C$ measured in a seperate experiment (data not shown), during similar diurnal time periods as the gas exchange campaigns presented here.
+$\delta^13C_{atmosphere}$ represents the mean whole-tree chamber $\delta^13C$ measured in a separate experiment (data not shown), during similar diurnal time periods as the gas exchange campaigns presented here.
 *C*~a~ and *C*~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~).

### 2016-09-02 master_scripts/data_tables.R

@@ -1,8 +1,6 @@
 source("functions and packages/functions.R")
 library(doBy)
 
-# gasexchange  <- read.csv("calculated_data/gmes_wtc.csv")
-#new wingate gm
 gasexchange  <- read.csv("calculated_data/gmes_wellwatered.csv")
 ##calculate CC
 gasexchange$Cc<- with(gasexchange, Ci-Photo/gm)
@@ -19,8 +17,6 @@ ge_overall <- summaryBy(Photo+Cond+Ci+Cc+Trmmol+gm_bar+VpdL ~ leaf +temp+leaflig
 ###split by drought and then rearrange
 ge_wet <- ge_overall[ge_overall$drydown == "control",]
 ge_wet <- droplevels(ge_wet)
-
-ge_dry <- ge_overall[ge_overall$drydown == "drought",]
 
 
 ###add ITE------------------------------------------------------------------------------------------------------------------- 

### 2016-09-02 master_scripts/gmgs_referee2.R

@@ -3,7 +3,7 @@ source("master_scripts/plot_objects.R")
 source("functions and packages/packages.R")
 
 ##read format phys data------------------------------------------------------------------------------------------------
-gasex <- read.csv("calculated_data/gmes_wellwatered2.csv")
+gasex <- read.csv("calculated_data/gmes_wellwatered.csv")
 gm_agg <- summaryBy(gm_bar+Cond+Photo ~ chamber+id+leaf +light+temp+leaflight+Month, 
                     data=gasex, FUN=mean, keep.names=TRUE)
 

### 2016-09-02 scripts/physiology_barplots.R

@@ -1,58 +1,0 @@
-##physiology bar plots of gs, gm and A overall means with sun, shade and lights on
-
-# source("functions and packages/functions.R")
-# source("master_scripts/plot_objects.R")
-# source("functions and packages/packages.R")
-
-gasex <- read.csv("calculated_data/gmes_wellwatered.csv")
-
-
-gasex_agg <- summaryBy(Photo+gm+Cond~ chamber+id+leaf +light+temp+leaflight, data=gasex, FUN=mean, keep.names=TRUE)
-  gasex_agg$leaflight <- gsub("shade-low", "Shade", gasex_agg$leaflight)
-  gasex_agg$leaflight <- gsub("sun-high", "Sun", gasex_agg$leaflight)
-  gasex_agg$leaflight <- gsub("shade-high", "Shade \n High Light", gasex_agg$leaflight)
-
-
-leaforder <- c("Shade", "Sun", "Shade \n High Light")  
-gasex_agg$leaflight <- factor(gasex_agg$leaflight, levels = leaforder)
-
-#lightson <- gasex_agg[gasex_agg$leaflight != "sun-high",]
-
-cols <- c(shacol2, suncol2, lightscol2)
-
-####PLOTTING-------------------------------------------------------------------------------------------------------------
-
-#1: stomatal conductance
-
-# windows(12,6)
-
-par(mfrow=c(1,3))
-
-bar(Cond, leaflight, gasex_agg, col=cols, half.errbar=FALSE, ylim=c(0, 0.3),names.arg=FALSE,xlab="",
-    mar=c(5,5,2,1), ylab=condlab, cex.axis=1.25, cex.lab = 1.5, cex.names=1.5,legend=F)
-abline(v=2.5, lty=5)
-text(x=.25, y=.2925, "(a)", cex=1.5)
-mtext("Shade \n Low Light", side=1,at=.725, line=3.5 )
-mtext("Sun \n High Light", side=1,at=1.9, line=3.5 )
-mtext("Shade \n High Light", side=1,at=3.15, line=3.5 )
-
-####mesophyll conductance
-bar(gm, leaflight, gasex_agg, col=cols, xlab="", half.errbar=FALSE, ylim=c(0, 0.225),names.arg=FALSE,
[…]

### 2016-09-02 scripts/read_data.R

@@ -1,3 +1,0 @@
-gm<- read.csv("raw data/gm_all.csv")
-leaf<- read.csv("raw data/leaf_data.csv")
-treatments <- read.csv("raw data/chamber_trt.csv")

### 2016-09-02 stats/aci_param_stats.R

@@ -1,51 +1,0 @@
-aci <- read.csv("calculated_data/aciparameters.csv")
-
-
-###linear models of gas exchange by treatments
-library(visreg)
-library(multcomp)
-library(nlme)
-
-####leaf Narea--------------------------------------------------------------------------------------------------
-#examite data with boxplots, then remove any outliers
-boxplot(Jmax~temp, data=aci[aci$leaf =="sun",])
-boxplot(Jmax~temp, data=aci[aci$leaf =="shade",])
-
-boxplot(Vcmax~temp, data=aci[aci$leaf =="sun",])
-boxplot(Vcmax~temp, data=aci[aci$leaf =="shade",])
-
-#Jmax different by leaf type but not warming
-J_sun_temp <- lme(Jmax ~ temp ,random=~1|chamber, data=aci, subset=leaf=="sun")
-summary(J_sun_temp)
-anova(J_sun_temp)
-visreg(J_sun_temp)
-
-J_sha_temp <- lme(Jmax ~ temp ,random=~1|chamber, data=aci, subset=leaf=="shade")
-summary(J_sha_temp)
-anova(J_sha_temp)
-visreg(J_sha_temp)
-
-J_leaf <- lme(Jmax ~ leaf, random=~1|chamber, data=aci)
-summary(J_leaf)
-anova(J_leaf)
-visreg(J_leaf)
-
-
-#Vcmax different by leaf type but not warming
-V_sun_temp <- lme(Vcmax ~ temp ,random=~1|chamber, data=aci, subset=leaf=="sun")
-summary(V_sun_temp)
-anova(V_sun_temp)
-visreg(V_sun_temp)
-
-V_sha_temp <- lme(Vcmax ~ temp ,random=~1|chamber, data=aci, subset=leaf=="shade")
-summary(V_sha_temp)
-anova(V_sha_temp)
-visreg(V_sha_temp)
-
-V_leaf <- lme(Vcmax ~ leaf, random=~1|chamber, data=aci)
-summary(V_leaf)
-anova(V_leaf)
-visreg(V_leaf)
-
-
-

### 2016-09-06 gm_pce/review_response.Rmd
Edits to reviewers response

@@
 \
-- 1. *Where these mature/isobilateral leaves and does this affect the outcome?*
+- *Editor: 1.
+Were these mature leaves and therefore isobilateral?
+Does this perhaps affect the outcome?
+The fact that LMA did not vary with shade is different from many other reports for dorsiventral leaves.*
+
 \
-These leaves studied in this experiment were very likely juvenile leaves, as the *E.
+*Response*: The leaves studied in this experiment were very likely juvenile leaves, as the *E.
 tereticornis* trees were grown for ~ 1 year.
 Consequently, the fact that LMA was similar within the canopy is different from other reports as the editor notes.
-Similar LMA within canopies has been noted before in Eucalpt plantations (Nouvellon et al.
+Similar LMA within canopies has been noted before in Eucalypt plantations (Nouvellon et al.
 2010), which is cited in the discussion.
@@
 \
-- 2. *Stylistic concerns with colors* 
+- *Editor: 2. (Stylistic) I found the colors to be distracting. Judicial use of color can help clarify data but unnecessary use of color can be distracting.*
+
 \
-We feel that colors are necessary in this manuscript to visualize leaf types (+light), while also differentiating temperature treatments.
+*Response*: We feel that the use of colour was necessary to visualize the three leaf types (sun, shade, shade+light), while also differentiating between the two temperature treatments.
 We attempted to use only symbols, which makes it near impossible to distinguish all of the treatments.
@@
 We have dropped the use of colors for figures with box plots, as they are not essential.
-(*need to try colors again/drop orange??*)
 
 \
-- 3/4. *Concerns over parameters derived from AC~i~ curves* 
-\
-After careful thought we have decided to remove parameters derived from AC~i~ curves from the manuscript.
-Our main goal is to show that sun and shade leaves were fundamental different in their photosynthetic capacity.
-As such, we have decided to instead present *A*~max~ values between the leaf types.
-This provides simple robust measurements of photosynthetic capacity, that reach our intended goal in a more straight forward manner.
-If specific derivation or representations of *J*~max~/*V*~cmax~ values has distracted the editor from our main point, it is likely to also distract readers who are plant physiologists.
-Figure 2.
-now shows the relationship between *A*~max~ and leaf *N*~a~ and *A*~max~ values are also reported in Table 1.
+*Editor: 3.
+The use of Jmax is consistent with the original use but not consistent with more modern use of the J that is found when light is extrapolated to infinity (similar to the concept of Vmax.
+Buckley and Diaz-Espejo recently published on this and given the use of Jmax in global models it might be good to limit the use of Jmax to the extrapolated number meaning that the J from an ACi curve analysis should be called something else (Jhigh?
+Jsat?) not sure specifically what to recommend, I avoid the issue by using just J but am not sure that is the best solution either.*
 
 \
-We also acknowledge that AC~i~ response curves were initially conceived with the notion that *g*~m~ was small enough to be ignored,     which we now know is untrue.
-This represents a new issue for the plant physiology community as more *g*~m~ data are continually  generated.
-We hope this data set can eventually help be used to address this problem, but it is beyond the scope of this paper.
+and
 
 \
-- 5. *Does light penetration deeper in a shade leaf impact mesophyll conductance?* 
+*4.
+I don't like expressing Vcmax on a "Ci" basis.
+This again changes the meaning of Vcmax from the extrapolated maximum rubisco rate (biochemically correct) to enzyme and biophysical diffusion components (no longer biochemically a Vmax and making two different definitions of Vcmax.
+Given that mesophyll conductance is a central part of this manuscript I don't understand why some reasonable constraints cannot be put on the A-Ci analysis so that parameters can be estimated based on Cc.
+I can accept that mesophyll conductance derived from ACi curves may not be sufficiently reliable, but the measurements made with the TDL should allow some reasonable Gm's to be applied to the ACi curves to avoid a Vcmax that is both biochemistry and biophysics.*
+
 \
-This question raises theoretical issues about how/if photosynthetic machinery are optimally distributed along intraleaf light gradient.
+*Response*: After careful thought we have decided to remove parameters derived from AC~i~ curves from the manuscript.
+Our main goal is to show that sun and shade leaves differed in their photosynthetic capacity, regardless of whether these differences are due to biochemistry or leaf structure (which we cannot satisfactorily address with the available data).
+.
+As such, we have decided to instead present photosynthetic rates at saturating CO~2~ concentrations and light (*A*~max~) for the different leaf types.
+This provides a simple metric of photosynthetic capacity, and shows that shade leaves have a lower photosynthetic capacity.
+We have adjusted the Methods (which is now shorter), as well as Figure 2 which now shows the relationship between *A*~max~ and leaf *N*~a~ and *A*~max~ values (these values are also reported in Table 1).
+
+\
+We acknowledge that AC~i~ response curves were initially conceived with the notion that *g*~m~ was large enough to be ignored, which we now know is untrue.
+We do think there are some practical concerns when fitting A-Cc curves, and we hope this data set can eventually help to address this issue, but it is beyond the scope of this paper.
+
+\
+*Editor: 5. Again because the leaves are isobilateral, to what degree is the increase in gm in high light a reflection of photosynthesis happening over a greater depth of leaf?*
+
+\
+*Response*: This question raises theoretical issues about how/if photosynthetic machinery are optimally distributed along intraleaf light gradient.
 To our knowledge, these issues have yet to resolved empirically and are still a relevant question for most leaf-level experiments.
-For example, it is possible that *J*~max~/*V*~cmax~ could be partitioned within the leaf according to light gradients (see Llyod et al.
-1992), which could affect estimates of *g*~m~ at low light.
+For example, it is possible that *J*~max~ and/or *V*~cmax~ could be partitioned within the leaf according to light gradients (see Lloyd et al.
+1992), which may affect estimates of *g*~m~ at low light.
 Alternatively, greater light penetration in to the leaf but longer path lengths for CO~2~ diffusion might interact to affect estimates of *g*~m~ at high light.
@@
 \
-- 1. *Difficulty in conciliating optimal resource distribution without an enhancement of leaf hydraulic conductance*
+*Reviewer: Campany et al have presented a work on spatial distribution of leaf gas exchange variables and parameters as a function of light regime and how they reflect an optimal distribution of resources in the canopy.
+The most remarkable result of this study is the occurrence of a higher gs in shaded leaves than in sunny ones.
+It is not clear to me the advantages of shaded leaves in having higher gs than ones at sun.
+I can see the point of enhancing An when sunflecks, but if the whole day is integrated it would result in a dramatic reduction of WUE, since shaded leaves with stomata open will transpire more or similarly to sunny ones (assuming that sunny leaves have higher temperature and therefore higher leaf-to-air VPD.
+Honestly, I cannot see any optimisation for canopy C in here, if this would be the case.
+As authors discussed, perhaps we should considered a dynamic light environment to understand the advantages of this strategy, but still I find difficult to conciliate with an optimal distribution of resources, in most cases very limiting.
+Data shows that there was an increase of gs, and hence, water use, without an enhancement of KL, and the relationship between gsmax and Vcmax, which usually found in canopies, but in this study is broken was broken.*
+
 \
-Due to Referee #1 concern over cuvette measured transpiration, water potentials and subsequent *K*~l~ calculations, we have re-examined the calculation and subsequent analyses in the manuscript.
+*Response*: Due to the concern over cuvette measured transpiration, water potentials and subsequent *K*~l~ calculations, we have re-examined the calculation and subsequent analyses in the manuscript.
 Initially, *K*~l~ was calculated with transpiration rates in shade leaves under low light conditions.
-As shade leave were allowed to stabilize under high light conditions before they were detached, the leaf water status reflected these conditions and not those of low light.
+As shade leave were allowed to stabilize under high light conditions before they were detached and measured for water potential, the leaf water status reflected these conditions and not those of low light.
 We have corrected the calculation of *K*~l~ and made changes in the results and discussion sections.
@@
 \
-- 2. *Diurnal cycles of photosynthesis, stomatal conductance and light within the canopy* 
-\
-This experiment was harvested in May 2014, shortly after our measurements ceased. Therefore, the suggestion to make diurnal measurements is no longer possible.
+*Reviewer:  Since the theory of considering a dynamic light environment is very appealing and novel, I would suggest to the authors to make diurnal cycles of light, gs and An at different locations in the canopy and check this hypothesis.
+They have all the equipment necessary to do so and they could integrate a whole day and see if it makes sense.*
 
 \
-- 3. *Concern over methods to measure leaf hydraulic conductance*
+*Response*: The trees in the whole-tree chambers were harvested in May 2014, shortly after our measurements were complete.
+Therefore, the suggestion to make diurnal measurements is no longer possible.
+We do agree with the reviewer that such detailed measurements would make for an interesting follow-up study.
+
 \
-The method used here for measuring *K*~l~ is also one of the three methods described in Flexas et al.
+*Reviewer: Another potential flaw I can see is the way KL was measured in this study.
+I would say it is not acceptable if we want to reach conclusions about the spatial distribution of K within the canopy.
+The assumed steady-state in the IRGAs chamber while measuring transpiration might not be the same one of the leaves in which leaf water potential was measured.
+Likely it is not, since leaves modify the VPD in the chamber respect to the one under natural conditions.
+Additionally, it seems that KL was not estimated from measurements of leaf water potential and E taken on the same leaf.
+At least, it is not specified.
+This can be subjected to errors in the estimation of KL.
+Finally, the gradient of water potential was considered between the soil (assuming leaf water potential as a good surrogate of matric potential) and the leaf.
+So authors were measuring the plant hydraulic conductance, from leaf to root.
+It is normal that it is similar in all leaves of the plant.*
+
+\
+*Response*: The main issue in this comment is whether or not leaf gas exchange and water potential were measured on the same leaves.
+As explained in the section "Leaf Nitrogen and hydraulic conductance", they were indeed measured on the same leaves.
+Leaves used for gas exchange were indeed sampled and then immediately measured for midday water potential.
+The method we used to estimate KL is standard, and presented as one of the most common methods in Flexas et al.
 2013, which compared the coordination of leaf hydraulic conductance and mesophyll conductance.
-Leaves used for gas exchange were indeed sampled and then immediately measured for midday water potential.
-This is explained in the first sentence of the subsection "Leaf Nitrogen and hydraulic conductance" in the methods.
+We now cite this work in the Methods section, as justification for the methods used.
 
@@
 \
-- 1. *Referee #2 correctly points out that the issues raised by Wingate et al. 2007 were not addressed*
-\
-We have recalculated the respiratory fractional component 'e', according to Evans and von Caemmerer 2013, which addresses Wingate et al.
-2007.
-This correction increased values of *g*~m~ but does not alter observed patterns.
-The correct calculation of 'e' has been added to the supplementary methods.
-This correction also addresses comment #3 below.
+*Reviewer: This manuscript describes the analysis of dynamic responses to changed light conditions in Eucalyptus trees growing at different temperatures in whole-tree chambers.
+The study focuses  on the differences between sun and shade leaves almost entirely; thermal acclimation was apparently so complete that it didn’t leave much to say about it.
+The surprising results were, first, that increasing the light intensity around a shade leaf to that of full sunlight caused it to ”over-compensate”, rising to photosynthetic rates that exceeded those of the sun-acclimated leaves.
+This must have caused the optimizers among this group of authors to scratch their heads.
+The second surprise was that the mesophyll conductance rose to levels indistinguishable from those of sun leaves when the shade leaves were put in full sun—and they did so within minutes.
+This result suggests that there is no structural component of mesophyll conductance in the study species.
+The similarity in LMA between sun and shade leaves tends to confirm this impression.
+The paper is well written, well analyzed, and well illustrated.*
 
 \
-- 2. *Estimating turnover time of the pools*
-\
-We agree that this concept would be great for the plant physiological community.  However, we fill it would detract from main conclusions of this manuscript and we do not address this further.
+*Response*: We thank the reviewer for the positive comments, and agree that the points raised were the most surprising results in the study.
 
 \
-- 3. *Was the chamber $\delta$^13^C measured?*
+*Reviewer: I have two significant concerns about it.
+First, the mesophyll conductance calculation apparently assumed that the respiratory and photorespiratory carbon pools were at isotopic steady-state.
+This assumption is most problematic in the leaves moved from shade to full sunlight, where the cc/ca likely fell substantially.
+It is not clear that the carbon pools supporting the respiratory and photorespiratory flux terms would have been able to shift to the new photosynthate in the brief acclimation period (ca.
+25 min) of these experiments.
+This issue was raised by Wingate et al.
+in a nice PCE ms.
+several years ago (WINGATE et al., 2007).
+Perhaps the modeling could be conducted with the assumption that old photosynthate makes up these pools and a sensitivity analysis conducted to determine whether it matters.
+Alternatively, perhaps the TDL data can be used to describe the shift in isotopic composition of the net flux; an asymptote might be used as evidence of a new steady-state having been reached.
+It would be great for the community if you were able to use a description of this shift to estimate the turnover time of the pools!*
+
 \
-Measurements of chamber $\delta$^13^C throughout the day were measured for this experiment during the same week as each gas exchange campaigns presented here.
+*Response*: The reviewer is correct that we did not account for 'old' carbon in respiratory pools.
+We have now recalculated the respiratory fractional component 'e' (according to Evans and von Caemmerer 2013), which addresses this issue (which was described by Wingate et al.
+2007).
+The correction increased values of *g*~m~ but does not alter observed patterns.
+The correct calculation of 'e' has been added to the supplementary methods.
+This correction also addresses the next comment.
+
+We agree with the reviewer that "Perhaps the modeling could be conducted with the assumption that old photosynthate makes up these pools and a sensitivity analysis conducted to determine whether it matters".
+However we think this idea is well outside the scope of the current study, but it is an issue that could be addressed in future work.
+
+\
+*Reviewer: Second, was the chamber δ13C measured?
+It could be quite different from the bulk atmosphere if the throughflow was low.
+It could also vary considerably during the day, further violating the steady-state assumption described above.
+The points above take nothing from the surprising net photosynthetic rate when the shade leaves were moved to sun.
+They only affect the gmes estimates.*
+
+\
+*Response*: Measurements of chamber $\delta$^13^C throughout the day were measured for this experiment during the same week as each gas exchange campaigns presented here.
 Since these measurements do not explicitly overlap, we do not have a robust way to assign specific values to each chamber at each time point.
@@
 \
-- 4. *Referee #1 Specific Comments:*
+*Reviewer: A third smaller point, which the authors allude to but do not emphasize, is that the ”shade” leaves may in fact have been produced in the sun when the trees were smaller.
+If so, it is unlikely that they could shift their morphology, but quite likely that they have shifted their biochemistry.
+This situation was described by Renee Brooks and colleagues (Brooks et al., 1996) many years ago in a paper that should be cited more—and should be cited here!
+More generally, it would seem that the discussion in the literature about the extent to which gmes is fixed by leaf morphology or flexible should be expanded here.
+I’m thinking of the papers that correlate gmes with LMA, for example.
+Flexas reviewed this literature several years ago and here’s a recent example of it (Milla-Moreno et al., 2016).
+The rapid adjustments described here, if they are real, clearly describe something different.*
 
 \
-- line 358: We corrected the text to match stats from Table 2
+*Response*: We agree with the reviewer's points, and have expanded (**check where**) the discussion on "the extent to which gmes is fixed by leaf morphology or flexible".
+We have also added the citation to Brooks et al.
+1996.
 
 \
-- line 392: We agree that thermal acclimation is not previously discussed before this point is made.
+*Reviewer: 358: the stats in table 2 do not support this statement. They say that the gm of shade leaves in light was indistinguishable from that of sun leaves. That’s good enough.*
+
+\
+*Response*: We corrected the text to match stats from Table 2
+
+\
+*Reviewer: 392: thermal acclimation is mentioned at this key point, but not much before this. I suggest being consistent and taking it out here too.*
+
+\
+*Response*: We agree that thermal acclimation is not previously discussed before this point is made.
 However, the experimental design does include a +3C warming treatment, and the potential affects of warming on leaf physiology are mentioned in the introduction.
@@
 \
-- Q: Nitrogen concentration in the leaves had a weak postie relationship with measured rates of *g*~m~ if all leaves were considered.
+*Reviewer: Q: Is nitrogen concentration related to gmes, or maximum gmes? If it’s aquaporins, one might expect so.*
+
+\
+*Response*: Nitrogen concentration in the leaves had a weak postie relationship with measured rates of *g*~m~ if all leaves were considered.
 This weak relationship disappeared if sun and shade leaves were analyzed separately.
@@
 \
-- line 404-405: added Brooks et al. 1996 citation explaining observed patterns of morphology and physiology when sun leaves become shaded.
+*Reviewer: 404 and 405: compare to Brooks’s study mentioned above and expand to address fixed vs. variable controls*
 
 \
-- line 412-415: *REMKO*
+*Response*: - line 404-405: added Brooks et al. 1996 citation explaining observed patterns of morphology and physiology when sun leaves become shaded.
 
 \
-- Q: On average mesophyll conductance and stomatal might be expected to be correlated when investigating inter-specific comparisons, such as in Flexas et al.
-2013 (Figure 1).
+*Reviewer: 412 to 415: if I recall correctly, optimization of WUE in the presence of similar hydraulics should lead to similar WUE throughout the canopy.
+Is this true?
+If so, it should be stated and the discrepancy made clearer.
+The point is made (I think), but it should be crystal clear that the optimization arguments are failing us here, which may mean that we’re missing something.
+Approximately this point is made on lines 431 to 434, but as presented it seems like a minor point.
+I think it is not so minor.*
+
+\
+*Response*: The work of Peltoniemi et al.
+(2012) demonstrated that similar WUE throughout the canopy is only achieved if the hydraulic conductance (*K*~l~ ) is proportional to the PPFD gradient (in the same way that leaf nitrogen should be proportional to PPFD).
+They also reasoned that if shade leaves have a higher hydraulic conductance (perhaps due to a shorter path length), the optimal response will be to have a higher Ci (and thus lower WUE) in shade leaves compared to sun leaves.
+Since we fixed a mistake in the calculation of *K*~l~, the results on WUE differences between sun and shade leaves are consistent with the theoretical predictions of Peltoniemi.
+We have expanded on this interesting point in the discussion (expanding on the section that previously included lines 431-434).
+(**haven't done this yet!**)
+
+\
+*Reviewer: Q.
+It has been argued that gmes should be correlated with gs.
+Not sure why that should be so, except that they are similar quantities, but it would be good to see what the correlation looks like here.*
+
+\
+*Response*: On average mesophyll conductance and stomatal conductance might be expected to be correlated when investigating inter-specific comparisons, such as in Flexas et al.
+2013.
 The dynamic responses of leaf physiology to environmental conditions within a species/canopy are likely to alter this relationship.
-In this experiment, you can see a postie relationship when only sun leaves are considered, similar to the Flexas paper.
+In this experiment, there was a positive relationship between the two variables when only sun leaves are considered, similar to the Flexas paper.
 Due to the unexpected stomatal behavior, however, this relationship was reversed in shade leaves.

### 2016-09-09 gm_pce/review_response.Rmd
More edits; some references added

@@
 \
-*Response*: Due to the concern over cuvette measured transpiration, water potentials and subsequent *K*~l~ calculations, we have re-examined the calculation and subsequent analyses in the manuscript.
-Initially, *K*~l~ was calculated with transpiration rates in shade leaves under low light conditions.
+*Response*: Due to the concern over cuvette measured transpiration, water potentials and subsequent *K*~L~ calculations, we have re-examined the calculation and subsequent analyses in the manuscript.
+Initially, *K*~L~ was calculated with transpiration rates in shade leaves under low light conditions.
 As shade leave were allowed to stabilize under high light conditions before they were detached and measured for water potential, the leaf water status reflected these conditions and not those of low light.
-We have corrected the calculation of *K*~l~ and made changes in the results and discussion sections.
-*K*~l~ is now significantly higher in shade leaves, which makes sense in relation to the maintain hydraulic supply to the observed stomatal behavior.
+We have corrected the calculation of *K*~L~ and made changes in the results and discussion sections.
+We have also clarified the calculation in the Methods section.
+
+After the recalculation, *K*~L~ is now significantly higher in shade leaves compared to sun leaves, which makes sense in relation to the maintain hydraulic supply to the observed stomatal behavior.
 We believe this actually makes the optimization of resources story much clearer throughout the manuscript, and we greatly appreciate the referee's comments.
-These corrected values also allow us to address the possible coordination of *K*~l~ and *g*~m~ reported by Flexas et al.
+These corrected values also allow us to address the possible coordination of *K*~L~ and *g*~m~ reported by Flexas et al.
 2013.

### 2016-09-09 gm_pce/manuscript_gmes.Rmd
More edits; some references added

@@
 To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole-tree chambers.
-Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic capacity (*V*~cmax~ and *J*~max~), but *g*~s~ was similar.
+Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic capacity (*A*~max~, i.e.
+*A*~n~ at saturating [CO~2~]), but *g*~s~ was similar.
 When light intensity was temporarily increased for shade leaves to match that of sun leaves, both *g*~s~ and *g*~m~ increased, and *A*~n~ increased to values greater than sun leaves.
@@
 Anatomical parameters which regulate *g*~m~ with changing growth irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative; @terashima2001sun; @hanba2002effect] are also unlikely to adjust during transient fluctuations in light.
-The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance of the growth environment and as a potential response to transitory light availability.
+The physiological behaviour of shade leaves to maximize C gain must be assessed as both a degree of acclimation to local irradiance and as a potential response to transitory light availability.
 
@@
 How these stresses affect the diffusion of CO~2~, through either *g*~s~ or *g*~m~, will have implications for up-scaling *A*~n~ for sun and shade leaves.
-Additionally, light-saturated rates of *A*~n~ are limited by the maximum rate of Rubisco carboxylation (*V*~cmax~) or the maximum rate of photosynthetic electron transport (*J*~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature].
-How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients.
+How these parameters, in combination with photosynthetic capacity, are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients.
 The impacts of warming on plant physiological processes are diverse, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.
@@
 We further aimed to quantify the constraints on *A*~n~ in sun and shade leaves arising from photosynthetic capacity and the CO~2~ diffusion pathway via stomatal and mesophyll conductance.
-As leaves which developed in the shade were expected to have lower photosynthetic capacity and correspondingly lower diffusive conductance (both *g*~s~ and *g*~m~) than sun leaves, we predicted that sunfleck simulations would increase *A*~n~ and *g*~s~ in shade leaves, but reach values lower than that attained by sun leaves in similar conditions.
+As leaves which were acclimated to shade conditions were expected to have lower photosynthetic capacity and correspondingly lower diffusive conductance (both *g*~s~ and *g*~m~) than sun leaves, we predicted that sunfleck simulations would increase *A*~n~ and *g*~s~ in shade leaves, but reach values lower than that attained by sun leaves in similar conditions.
 We further predicted that climate warming would decrease *g*~s~ and leaf C gain in sun leaves more so than shade leaves during summer months, as increased evaporative demand from higher temperatures and irradiance lead to stomatal closure.
@@
 These findings offer new insights into how aspects of leaf gas exchange physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake and water use in future studies.
-Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology, consistent with temperature acclimation demonstrated in this experiment [@aspinwall2016convergent].
+Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects of a +3 &deg;C warming treatment were detected on leaf physiology, consistent with temperature acclimation demonstrated in this experiment [@aspinwall2016convergent].
 
@@
 As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters *V*~cmax~ and *J*~max~ [@farquhar1980biochemical].
-Our data agree with these conventional conclusions as the distribution of *N*~a~, both measures of photosynthetic capacity, and *A*~n~ were all reduced in shade leaves.
-Leaf mass per area (LMA), however, was not different between sun and shade leaves, consistent with small differences reported in a clonal *Eucalyptus* plantation [@nouvellon2010].
+Our data agree with these conventional conclusions as the distribution of *N*~a~, *A*~max~, and *A*~n~ were all reduced in shade leaves.
+Leaf mass per area (LMA), however, was not different between sun and shade leaves, consistent with small differences reported in a clonal *Eucalyptus* plantation [@nouvellon2010], but in contrast to data for many other species [@niinemets2015].
 Similar LMA in sun and shade leaves could be the result of leaf formation under comparable light conditions or possible unmeasured differences in total non-structural carbohydrates contents between leaf types that masked differences in leaf thickness or density.
@@
 Using a two-leaf model, Peltoniemi et al.
-[-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if *K*~L~ is also optimally distributed.
-In this study, variation in leaf N distribution and *A*~n~ rates were not associated with subsequent changes in *K*~L~ between sun and shade leaves.
-Thus, no direct relationship between water supply capacity and N distribution or *A*~n~ within the canopy were detected.
+[-@peltoniemi2012co] theorized that optimal N distribution will be proportional to light distribution only if *K*~L~ is also optimally distributed.
+We found higher *K*~L~ in shade leaves compared to sun leaves, countering the prediction by Peltoniemi et al.
+[@peltoniemi2012co], but consistent with higher *g*~s~ attained by shade leaves in high light conditions.
+This finding is consistent with the idea that *g*~s~ is regulated, at least in part, by *K*~L~ [@hubbard2001stomatal].
 
@@
 @prentice2014balancing; @lin2015optimal].
-As a result, the often used economic framework of balancing costs of using water versus N allocation to predict *A*~n~ [@wright2003least] may break down when considering all leaves within a tree canopy.
+Consistent with our results, Buckley et al.
+[-@2014] found that shade leaves exhibited higher *g*~s~ than expected based on their environment.
+In our study, higher *g*~s~ in shade leaves may in part be explained by higher *K*~L~, which effectively reduces the cost of water transport, allowing more profligate water use in shade leaves.
 
 \
-It is possible that higher *g*~s~ in shade leaves is a strategy to take advantage of high light quickly in shade leaves, avoiding the delay with stomatal opening [@tausz2005dynamic].
+It is also possible that higher *g*~s~ in shade leaves is a strategy to take advantage of high light quickly in shade leaves, avoiding the delay with stomatal opening [@tausz2005dynamic;@vicomanz2011].
 Our results support this hypothesis, as shade leaves increased *A*~n~ to values equal or even outperforming sun leaves when subjected to a brief period of identical high-light intensity.
@@
 **Figure 2**.
-Photosynthetic capacity (*A*~max) was positively correlated with mean leaf N per area (*N*~a~) for sun and shade leaves grown under elevated (ET) and ambient (AT) temperature treatments.
+Photosynthetic capacity (*A*~max~) was positively correlated with mean leaf N per area (*N*~a~) for sun and shade leaves grown under elevated (ET) and ambient (AT) temperature treatments.
 Leaf *N*~a~ is the mean of each leaf type sampled across gas exchange campaigns in each whole-tree chamber.
-*A*~max was measured once on all trees for each leaf type, in February 2014, at 25°C and at both saturating light and [CO~2~].
+*A*~max~ was measured once on all trees for each leaf type, in February 2014, at 25°C and at both saturating light and [CO~2~].
 The dashed line represents the significant linear model fit for all leaves, with a marginal and conditional *R*^2^ of 0.59 and 0.62.
@@
 Box plots of stomatal conductance (*g*~s~) (a), mesophyll conductance (*g*~m~) (b) and leaf net photosynthesis rate (*A*~n~) (c) of sun leaves and shade leaves at low and high light pooled across six measurement dates and both temperature treatments.
+For statistical comparisons, see Table 2.
 \
@@
 All data across six measurement dates and both temperature treatments are included.
+For statistical comparisons, see Table 2.
 

### 2016-09-10 master_scripts/cicc_boxplots.R
added colors back to boxplots, changed this in reviewer response, renamed paper folder

@@ -31,21 +31,19 @@ cols <- c(shacol2, lightscol2,suncol2)
 # windows(12,6)
 par(mfrow=c(1,3),cex.axis=1.25, cex.lab = 1.5, mar=c(5,5,2,1))
 
-#col=cols
-
-boxplot(Ci ~leaflight, gasex_agg, ylab=expression(C[i]~~(paste(mu,bar,sep=""))), names=FALSE, ylim=c(0,325), outline=FALSE)
+boxplot(Ci ~leaflight, gasex_agg, col=cols,ylab=expression(C[i]~~(paste(mu,bar,sep=""))), names=FALSE, ylim=c(0,325), outline=FALSE)
   mtext("Shade \n Low Light", side=1,at=1, line=3.5)
   mtext("Shade \n High Light", side=1,at=2, line=3.5)
   mtext("Sun \n High Light", side=1,at=3, line=3.5)
   text(x=.55, y=325, "(a)", cex=1.5)
 
-boxplot(Cc ~leaflight, gasex_agg,  ylab=expression(C[c]~~(paste(mu,bar,sep=""))), names=FALSE,ylim=c(0,325),outline=FALSE)
+boxplot(Cc ~leaflight, gasex_agg,  col=cols,ylab=expression(C[c]~~(paste(mu,bar,sep=""))), names=FALSE,ylim=c(0,325),outline=FALSE)
   mtext("Shade \n Low Light", side=1,at=1, line=3.5)
   mtext("Shade \n High Light", side=1,at=2, line=3.5)
   mtext("Sun \n High Light", side=1,at=3, line=3.5)
   text(x=.55, y=325, "(b)", cex=1.5)
 
-boxplot(drawdown~leaflight, gasex_agg, ylab=drawdownlab3, xlab="", names=FALSE, ylim=c(0,325),outline=FALSE)
+boxplot(drawdown~leaflight, gasex_agg, col=cols,ylab=drawdownlab3, xlab="", names=FALSE, ylim=c(0,325),outline=FALSE)
   mtext("Shade \n Low Light", side=1,at=1, line=3.5)
   mtext("Shade \n High Light", side=1,at=2, line=3.5)
   mtext("Sun \n High Light", side=1,at=3, line=3.5)

### 2016-09-10 PCE_paper/review_response.Rmd
added colors back to boxplots, changed this in reviewer response, renamed paper folder
final edits to reviewer response

--- 2016-09-09 gm_pce/review_response.Rmd
@@
 \
-*Response*: We feel that the use of colour was necessary to visualize the three leaf types (sun, shade, shade+light), while also differentiating between the two temperature treatments.
+*Response*: We feel that the use of color was necessary to visualize the three leaf types (sun, shade, shade+light), while also differentiating between the two temperature treatments.
 We attempted to use only symbols, which makes it near impossible to distinguish all of the treatments.
@@
 The choice of color palette is also used to distinguish colors properly for color-blind readers.
-We have dropped the use of colors for figures with box plots, as they are not essential.
 
@@
 \
-*Response*: We agree with the reviewer's points, and have expanded (**check where**) the discussion on "the extent to which gmes is fixed by leaf morphology or flexible".
+*Response*: We agree with the reviewer's points, and have expanded the discussion on "the extent to which gmes is fixed by leaf morphology or flexible" (4^th^ paragraph in subsection 'Physiological behaviour of sun and shade leaves').
 We have also added the citation to Brooks et al.
@@
 We have expanded on this interesting point in the discussion (expanding on the section that previously included lines 431-434).
-(**haven't done this yet!**)
 

### 2016-09-10 PCE_paper/manuscript_gmes.Rmd
final edits to reviewer response

--- 2016-09-09 gm_pce/manuscript_gmes.Rmd
@@
 It has also been proposed that *g*~m~ and *K*~L~ should be coordinated, as optimizing hydraulic supply allows both *g*~s~ and *g*~m~ to enhance *A*~n~ [@flexas2013leaf].
-Here, higher *K*~l~ in shade leaves under high light lends supports to the potential coordination *K*~l~ and *g*~m~ and also to the association of higher *A*~n~ with higher *K*~l~ when light becomes saturating reported by @franks2006higher.
+Here, higher *K*~l~ in shade leaves under high light lends support to the potential coordination *K*~l~ and *g*~m~ and also to the association of higher *A*~n~ with higher *K*~l~ when light becomes saturating reported by @franks2006higher.
 

### 2016-09-11 PCE_paper/review_response.Rmd
manuscript edits

@@
 ---
-title: 
-author: 
-date: 
-output: word_document
+output:
+  word_document:
+    reference_docx: manuscriptstyle_pce2.docx
+  pdf_document: default
+  html_document:
+    number_sections: yes
 ---
@@
 We have responded in detail to each comment, which we feel has improved the overall quality of the manuscript.
-Below, we have itemized our response to each comment from the Editor and both reviewers.
+Below, we have itemized our response to each comment from the Editor and both referees.
 
 ### **Editor**
-
 \
@@
 It also complicates visualization of confidence intervals on scatter plots.
-The choice of color palette is also used to distinguish colors properly for color-blind readers.
+The choice of color palette is also used to properly colors for color-blind readers.
 
@@
 Our main goal is to show that sun and shade leaves differed in their photosynthetic capacity, regardless of whether these differences are due to biochemistry or leaf structure (which we cannot satisfactorily address with the available data).
-.
 As such, we have decided to instead present photosynthetic rates at saturating CO~2~ concentrations and light (*A*~max~) for the different leaf types.
 This provides a simple metric of photosynthetic capacity, and shows that shade leaves have a lower photosynthetic capacity.
-We have adjusted the Methods (which is now shorter), as well as Figure 2 which now shows the relationship between *A*~max~ and leaf *N*~a~ and *A*~max~ values (these values are also reported in Table 1).
+We have adjusted the Methods (which is now shorter), as well as Figure 2 which now shows the relationship between *A*~max~ and leaf *N*~a~.
+*A*~max~ values are also reported in Table 1.
 
@@
 \
-*Response*: This question raises theoretical issues about how/if photosynthetic machinery are optimally distributed along intraleaf light gradient.
-To our knowledge, these issues have yet to resolved empirically and are still a relevant question for most leaf-level experiments.
+*Response*: This question raises theoretical issues about how/if photosynthetic machinery are optimally distributed along intraleaf light gradients.
+To our knowledge, these issues have yet to be resolved empirically and are still a relevant question for most leaf-level experiments.
 For example, it is possible that *J*~max~ and/or *V*~cmax~ could be partitioned within the leaf according to light gradients (see Lloyd et al.
 1992), which may affect estimates of *g*~m~ at low light.
-Alternatively, greater light penetration in to the leaf but longer path lengths for CO~2~ diffusion might interact to affect estimates of *g*~m~ at high light.
+Alternatively, greater light penetration into the leaf but longer path lengths for CO~2~ diffusion might interact to affect estimates of *g*~m~ at high light.
 Although we acknowledge that this question is relevant to this study and many others, we do not have the tools to address it further in the manuscript.
@@
 \
-*Response*: Due to the concern over cuvette measured transpiration, water potentials and subsequent *K*~L~ calculations, we have re-examined the calculation and subsequent analyses in the manuscript.
-Initially, *K*~L~ was calculated with transpiration rates in shade leaves under low light conditions.
+*Response*: Due to the concern over cuvette measured transpiration, water potentials and subsequent *K*~L~ calculations, we have re-examined the calculation and subsequent analyses of *K*~L~ in the manuscript.
+Initially, *K*~L~ was calculated with transpiration rates of shade leaves under low light conditions.
 As shade leave were allowed to stabilize under high light conditions before they were detached and measured for water potential, the leaf water status reflected these conditions and not those of low light.
@@
 
-After the recalculation, *K*~L~ is now significantly higher in shade leaves compared to sun leaves, which makes sense in relation to the maintain hydraulic supply to the observed stomatal behavior.
+\
+After the recalculation, *K*~L~ is now significantly higher in shade leaves compared to sun leaves, which makes sense to maintain hydraulic supply to the observed stomatal behavior.
 We believe this actually makes the optimization of resources story much clearer throughout the manuscript, and we greatly appreciate the referee's comments.
@@
 Leaves used for gas exchange were indeed sampled and then immediately measured for midday water potential.
-The method we used to estimate KL is standard, and presented as one of the most common methods in Flexas et al.
+The method we used to estimate KL is standard, and presented as a commonly used method in Flexas et al.
 2013, which compared the coordination of leaf hydraulic conductance and mesophyll conductance.
@@
 
+\
 We agree with the reviewer that "Perhaps the modeling could be conducted with the assumption that old photosynthate makes up these pools and a sensitivity analysis conducted to determine whether it matters".

### 2016-09-11 PCE_paper/manuscript_gmes.Rmd
manuscript edits

@@
 # Set Flag to include figures in text or not.
-includeFigs <- FALSE
+includeFigs <- TRUE
 ```
@@
 
-source("gm_pce/manuscript_preload.R")
+source("PCE_paper/manuscript_preload.R")
 
 # Make figure PDFs.
-source("gm_pce/make_figures.R")
+source("PCE_paper/make_figures.R")
 
@@
 When light intensity was temporarily increased for shade leaves to match that of sun leaves, both *g*~s~ and *g*~m~ increased, and *A*~n~ increased to values greater than sun leaves.
-Here we show that dynamic physiological responses of shade leaves to altered light environments have implications for up-scaling leaf level measurements and predicting whole canopy carbon gain.
+We show that dynamic physiological responses of shade leaves to altered light environments have implications for up-scaling leaf level measurements and predicting whole canopy carbon gain.
 Despite exhibiting reduced photosynthetic capacity, the rapid up-regulation of *g*~m~ with increased light enables shade leaves to respond quickly to sunflecks.
@@
 Consistent with our results, Buckley et al.
-[-@2014] found that shade leaves exhibited higher *g*~s~ than expected based on their environment.
+[-@buckley2014] found that shade leaves exhibited higher *g*~s~ than expected based on their environment.
 In our study, higher *g*~s~ in shade leaves may in part be explained by higher *K*~L~, which effectively reduces the cost of water transport, allowing more profligate water use in shade leaves.
@@
 \
-```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600,message=FALSE, eval=includeFigs}
- source("master_scripts/aci_vcmaxnitro.R")
+```{r amaxnitro, fig.width=6, fig.height=6, warning=FALSE,echo=FALSE,dev="png", dpi=600,message=FALSE, eval=includeFigs}
+ source("master_scripts/amax_nitro.R")
 ```

### 2016-09-11 master_scripts/par2.R

@@ -1,5 +1,5 @@
 source("functions and packages/functions.R")
-# source("functions and packages/packages.R")
+source("functions and packages/packages.R")
 source("master_scripts/plot_objects.R")
 
 treatments <- read.csv("raw data/temp_trt.csv")

### 2016-09-12 PCE_paper/review_response.Rmd
final reviwer edits, this version sent to coauthors

@@
 \
-After the recalculation, *K*~L~ is now significantly higher in shade leaves compared to sun leaves, which makes sense to maintain hydraulic supply to the observed stomatal behavior.
+After the recalculation, *K*~L~ is now significantly higher in shade leaves compared to sun leaves, which makes sense in order to maintain hydraulic supply to the observed stomatal behavior.
 We believe this actually makes the optimization of resources story much clearer throughout the manuscript, and we greatly appreciate the referee's comments.
@@
 Leaves used for gas exchange were indeed sampled and then immediately measured for midday water potential.
-The method we used to estimate KL is standard, and presented as a commonly used method in Flexas et al.
+The method we used to estimate *K*~L~ is standard, and presented as a commonly used method in Flexas et al.
 2013, which compared the coordination of leaf hydraulic conductance and mesophyll conductance.

### 2016-09-12 PCE_paper/manuscript_gmes.Rmd
final reviwer edits, this version sent to coauthors

@@
 Measurements were completed in February 2014, when all trees were well-watered in each temperature treatment (*n*=6).
-*A*~max~ measurements were made at a PPFD of 1800 $\mu$mol m^-1^ s^-1^ and [CO~2~] of 1800 $\mu$l l^-1^, with leaf temperature (*T*~leaf~) held constant at 25 &deg;C .
+*A*~max~ measurements were made at a PPFD of 1800 $\mu$mol m^-1^ s^-1^ and [CO~2~] of 1800 $\mu$l l^-1^, with leaf temperature held constant at 25 &deg;C .
 
@@
 All leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed.
-Leaf water potentials and transpiration (*E*~L~, mmol m^-2^ s^-1^) from gas exchange in the leaf cuvette were then used to calculate leaf-specific hydraulic conductance (*K*~L~, mmol m^-2^ s^-1^ MPa^-1^) with the equation:
+Leaf water potentials and transpiration (*E*~L~, mmol m^-2^ s^-1^) from gas exchange in the leaf cuvette were then used to calculate leaf-specific hydraulic conductance (*K*~L~, mmol m^-2^ s^-1^ MPa^-1^) from the equation [@whitehead1998regulation; @flexas2013leaf]:
 
@@
 We found higher *K*~L~ in shade leaves compared to sun leaves, countering the prediction by Peltoniemi et al.
-[@peltoniemi2012co], but consistent with higher *g*~s~ attained by shade leaves in high light conditions.
+[-@peltoniemi2012co], but consistent with higher *g*~s~ attained by shade leaves in high light conditions.
 This finding is consistent with the idea that *g*~s~ is regulated, at least in part, by *K*~L~ [@hubbard2001stomatal].
@@
 Leaf mass per area (LMA, g m^2^), leaf nitrogen per unit area (*N*~a~, gN m^-2^), $\delta$^13^C (&permil;), pre-dawn leaf water potential ($\Psi$~pd~, MPa), midday leaf water potential ($\Psi$~L~, MPa) and leaf-specific hydraulic conductance (*K*~L~, mmol m^-2^ s^-1^ MPa^-1^) are given as treatment means (&pm; 1 standard error) across six measurement campaigns.
-Values of *V*~cmax~ and *J*~max~ are treatment mean (&pm; 1 standard error) from A-*C*~i~ curves measured in each chamber at 25 &deg;C and saturating light.
+Values of *A*~max~ ($\mu$mol m^-2^ s^-1^) were measured once on all trees for each leaf type at 25°C and at both saturating light and [CO~2~].
 Different letters represent significant differences between leaf type and temperature treatments (based on a Tukey test).

### 2016-09-21 PCE_paper/review_response.Rmd
marks changes

@@
 We thank the reviewers for providing insightful comments to improve our manuscript.
-We have responded in detail to each comment, which we feel has improved the overall quality of the manuscript.
+We have responded in detail to each comment, which has improved the overall quality of the manuscript.
 Below, we have itemized our response to each comment from the Editor and both referees.
@@
 \
-*Response*: The leaves studied in this experiment were very likely juvenile leaves, as the *E.
-tereticornis* trees were grown for ~ 1 year.
-Consequently, the fact that LMA was similar within the canopy is different from other reports as the editor notes.
-Similar LMA within canopies has been noted before in Eucalypt plantations (Nouvellon et al.
+*Response*: Unlike some eucalyptus species (e.g.
+*E.
+globulus*), the juvenile and adult leaf phase morphotypes of *E.
+tereticornis* are similar in shape and size (disjunct, broad lanceolate to narrow lanceolate).
+
+The leaves of *E.
+tereticornis* trees studied in this experiment could have included both morphotypes, but were nonetheless all fully developed and mature.
+Importantly, the leaves were indeed isobilateral with stomata on both sides (20% lower stomatal density on upper vs.
+lower leaf surface).
+The fact that LMA did not differ between sun and shade leaves is different from other reports as the editor notes.
+Lack of differentiation in LMA profiles within canopies have been noted before in Eucalypt plantations (Nouvellon et al.
 2010), which is cited in the discussion.
-We have expanded the discussion to address the possible fixed controls of LMA on leaf physiology (see reply to Referee # 2 specific comments for full details).
+We have expanded the discussion to address the possible controls of a relatively fixed LMA on leaf physiology (see reply to Referee # 2 specific comments for full details).
 
@@
 *Response*: We feel that the use of color was necessary to visualize the three leaf types (sun, shade, shade+light), while also differentiating between the two temperature treatments.
-We attempted to use only symbols, which makes it near impossible to distinguish all of the treatments.
+We attempted to use only symbols, which makes it nearly impossible to readily distinguish all of the treatments.
 It also complicates visualization of confidence intervals on scatter plots.
-The choice of color palette is also used to properly colors for color-blind readers.
+We have darkened several of the original colors, but this choice of color palette is also used to ensure accessibility for color-blind readers.
 
@@
 \
-and
+*Response*: All references to J have been removed from the manuscript (see repsonse below).
 
 \
-*4.
+*Editor: 4.
 I don't like expressing Vcmax on a "Ci" basis.
@@
 *Response*: After careful thought we have decided to remove parameters derived from AC~i~ curves from the manuscript.
-Our main goal is to show that sun and shade leaves differed in their photosynthetic capacity, regardless of whether these differences are due to biochemistry or leaf structure (which we cannot satisfactorily address with the available data).
+Thus, references to J~max~ and V~cmax have been removed.
+The presentation and discussion of the diffusion pathways, including Ci and Cc~ and draw down remain (Figure 7).
+Our main goal is to show that sun and shade leaves differed in their photosynthetic capacity, regardless of whether these differences are due to biochemistry or leaf ultrastructure or both (which we cannot satisfactorily address with the available data).
 As such, we have decided to instead present photosynthetic rates at saturating CO~2~ concentrations and light (*A*~max~) for the different leaf types.
-This provides a simple metric of photosynthetic capacity, and shows that shade leaves have a lower photosynthetic capacity.
+This provides a simple metric of photosynthetic capacity that is well supported in the literature, and shows that shade leaves have a lower photosynthetic capacity, irrespective of the relative contribution of biochemical and biophysical contraints.
 We have adjusted the Methods (which is now shorter), as well as Figure 2 which now shows the relationship between *A*~max~ and leaf *N*~a~.
@@
 We acknowledge that AC~i~ response curves were initially conceived with the notion that *g*~m~ was large enough to be ignored, which we now know is untrue.
-We do think there are some practical concerns when fitting A-Cc curves, and we hope this data set can eventually help to address this issue, but it is beyond the scope of this paper.
+We beleive there are some practical concerns when fitting A-Cc curves, and we hope this data set can eventually help to address this issue, but it is beyond the scope of this paper.
 
@@
 Alternatively, greater light penetration into the leaf but longer path lengths for CO~2~ diffusion might interact to affect estimates of *g*~m~ at high light.
-Although we acknowledge that this question is relevant to this study and many others, we do not have the tools to address it further in the manuscript.
+Although we acknowledge that this question is relevant to this study and many others, we do not have the data or tools to address it further in the manuscript.
 
@@
 \
-*Reviewer: Campany et al have presented a work on spatial distribution of leaf gas exchange variables and parameters as a function of light regime and how they reflect an optimal distribution of resources in the canopy.
+*Reviewer: 1.
+Campany et al have presented a work on spatial distribution of leaf gas exchange variables and parameters as a function of light regime and how they reflect an optimal distribution of resources in the canopy.
 The most remarkable result of this study is the occurrence of a higher gs in shaded leaves than in sunny ones.
@@
 \
-*Response*: Due to the concern over cuvette measured transpiration, water potentials and subsequent *K*~L~ calculations, we have re-examined the calculation and subsequent analyses of *K*~L~ in the manuscript.
+*Response*: We agree that the finding of higher gs in shade than sun leaves, particularly under high light conditions, is remarkable and indeed has implications for reductions in integrated WUE as shown and discussed in the manuscript (Table 2, Figure 5).
+In light of the reviewer’s point regarding lack of enhancement of *K*~L~ in shade leaves, we have re-examined the calculation and subsequent analyses of *K*~L~ in the manuscript by investigating issues that can arise through use of cuvette-measured transpiration.
 Initially, *K*~L~ was calculated with transpiration rates of shade leaves under low light conditions.
 As shade leave were allowed to stabilize under high light conditions before they were detached and measured for water potential, the leaf water status reflected these conditions and not those of low light.
-We have corrected the calculation of *K*~L~ and made changes in the results and discussion sections.
+Thus, we have corrected the calculation of shade leaf *K*~L~ with transpiration rates under high light and have made changes in the results and discussion sections.
 We have also clarified the calculation in the Methods section.
@@
 After the recalculation, *K*~L~ is now significantly higher in shade leaves compared to sun leaves, which makes sense in order to maintain hydraulic supply to the observed stomatal behavior.
-We believe this actually makes the optimization of resources story much clearer throughout the manuscript, and we greatly appreciate the referee's comments.
+We believe this actually makes the test if the optimization of resources framework and subsequent discussion and interpretation much clearer throughout the manuscript, and we greatly appreciate the referee's comments.
 These corrected values also allow us to address the possible coordination of *K*~L~ and *g*~m~ reported by Flexas et al.
@@
 \
-*Reviewer:  Since the theory of considering a dynamic light environment is very appealing and novel, I would suggest to the authors to make diurnal cycles of light, gs and An at different locations in the canopy and check this hypothesis.
+*Reviewer: 2.
+Since the theory of considering a dynamic light environment is very appealing and novel, I would suggest to the authors to make diurnal cycles of light, gs and An at different locations in the canopy and check this hypothesis.
 They have all the equipment necessary to do so and they could integrate a whole day and see if it makes sense.*
@@
 \
-*Reviewer: Another potential flaw I can see is the way KL was measured in this study.
+*Reviewer: 3.
+Another potential flaw I can see is the way KL was measured in this study.
 I would say it is not acceptable if we want to reach conclusions about the spatial distribution of K within the canopy.
@@
 *Response*: The main issue in this comment is whether or not leaf gas exchange and water potential were measured on the same leaves.
-As explained in the section "Leaf Nitrogen and hydraulic conductance", they were indeed measured on the same leaves.
+As explained in the section "Leaf Nitrogen and hydraulic conductance", they were measured on the same leaves.
 Leaves used for gas exchange were indeed sampled and then immediately measured for midday water potential.
@@
 \
-*Reviewer: This manuscript describes the analysis of dynamic responses to changed light conditions in Eucalyptus trees growing at different temperatures in whole-tree chambers.
+*Reviewer: 1.
+This manuscript describes the analysis of dynamic responses to changed light conditions in Eucalyptus trees growing at different temperatures in whole-tree chambers.
 The study focuses  on the differences between sun and shade leaves almost entirely; thermal acclimation was apparently so complete that it didn’t leave much to say about it.
@@
 \
-*Reviewer: I have two significant concerns about it.
+*Reviewer: 2.
+I have two significant concerns about it.
 First, the mesophyll conductance calculation apparently assumed that the respiratory and photorespiratory carbon pools were at isotopic steady-state.
@@
 \
-*Reviewer: Second, was the chamber δ13C measured?
+*Reviewer: 3.
+Second, was the chamber δ13C measured?
 It could be quite different from the bulk atmosphere if the throughflow was low.
@@
 \
-*Response*: Measurements of chamber $\delta$^13^C throughout the day were measured for this experiment during the same week as each gas exchange campaigns presented here.
+*Response*: Measurements of chamber $\delta$^13^C throughout the day were measured for this experiment during the same week as each of the gas exchange campaigns presented here.
 Since these measurements do not explicitly overlap, we do not have a robust way to assign specific values to each chamber at each time point.
@@
 This change is included in the corrections mentioned above.
-Additionally, we used CO~2~ cylinders for the Licor 6400 that were much closer to ambient air $\delta$^13^C, instead of standard cylinders with  $\delta$^13^C > 30‰, which slightly improves calculations.
+Additionally, for our cuvette measurements we used CO~2~ cylinders for the Licor 6400 that were much closer to ambient air $\delta$^13^C, instead of standard cylinders with  $\delta$^13^C > 30‰, which slightly improves calculations.
 
 \
-*Reviewer: A third smaller point, which the authors allude to but do not emphasize, is that the ”shade” leaves may in fact have been produced in the sun when the trees were smaller.
+*Reviewer: 4.
+A third smaller point, which the authors allude to but do not emphasize, is that the ”shade” leaves may in fact have been produced in the sun when the trees were smaller.
 If so, it is unlikely that they could shift their morphology, but quite likely that they have shifted their biochemistry.
@@
 \
-*Reviewer: 358: the stats in table 2 do not support this statement. They say that the gm of shade leaves in light was indistinguishable from that of sun leaves. That’s good enough.*
+*Reviewer: 5. the stats in table 2 do not support this statement. They say that the gm of shade leaves in light was indistinguishable from that of sun leaves. That’s good enough.*
 
 \
-*Response*: We corrected the text to match stats from Table 2
+*Response*: We corrected the text to match stats from Table 2.
 
 \
-*Reviewer: 392: thermal acclimation is mentioned at this key point, but not much before this. I suggest being consistent and taking it out here too.*
+*Reviewer: 6. thermal acclimation is mentioned at this key point, but not much before this. I suggest being consistent and taking it out here too.*
 
@@
 As there were very few effects of climate warming observed in this aspect of the experiment, we feel it is important to address a possible explanation for the lack of observed warming effects using published findings from this experiment.
+Nonetheless, our findings and conclusions regarding sun and shade leaf contrasts are robust between trees grown in contrasting temperatures.
 
 \
-*Reviewer: Q: Is nitrogen concentration related to gmes, or maximum gmes? If it’s aquaporins, one might expect so.*
+*Reviewer: 7. Q: Is nitrogen concentration related to gmes, or maximum gmes? If it’s aquaporins, one might expect so.*
 
 \
-*Response*: Nitrogen concentration in the leaves had a weak postie relationship with measured rates of *g*~m~ if all leaves were considered.
+*Response*: Nitrogen concentration in the leaves had a weak positive relationship with measured *g*~m~ if all leaves were considered.
 This weak relationship disappeared if sun and shade leaves were analyzed separately.
@@
 \
-*Reviewer: 404 and 405: compare to Brooks’s study mentioned above and expand to address fixed vs. variable controls*
+*Reviewer: 8. line 404 and 405: compare to Brooks’s study mentioned above and expand to address fixed vs. variable controls*
 
 \
-*Response*: - line 404-405: added Brooks et al. 1996 citation explaining observed patterns of morphology and physiology when sun leaves become shaded.
+*Response*: - line 404-405: We added the Brooks et al. 1996 citation explaining observed patterns of morphology and physiology when sun leaves become shaded during tree growth and crown development.
 
 \
-*Reviewer: 412 to 415: if I recall correctly, optimization of WUE in the presence of similar hydraulics should lead to similar WUE throughout the canopy.
+*Reviewer: 9.
+line 412 to 415: if I recall correctly, optimization of WUE in the presence of similar hydraulics should lead to similar WUE throughout the canopy.
 Is this true?
@@
 They also reasoned that if shade leaves have a higher hydraulic conductance (perhaps due to a shorter path length), the optimal response will be to have a higher Ci (and thus lower WUE) in shade leaves compared to sun leaves.
-Since we fixed a mistake in the calculation of *K*~l~, the results on WUE differences between sun and shade leaves are consistent with the theoretical predictions of Peltoniemi.
+Since we fixed a mistake in the calculation of *K*~l~, the results on WUE differences between sun and shade leaves are indeed consistent with the theoretical predictions of Peltoniemi.
 We have expanded on this interesting point in the discussion (expanding on the section that previously included lines 431-434).
@@
 \
-*Reviewer: Q.
-It has been argued that gmes should be correlated with gs.
+*Reviewer: 10.
+Q: It has been argued that gmes should be correlated with gs.
 Not sure why that should be so, except that they are similar quantities, but it would be good to see what the correlation looks like here.*

### 2016-09-21 PCE_paper/review_response.Rmd
MT edits to reviewer response

@@
 tereticornis* are similar in shape and size (disjunct, broad lanceolate to narrow lanceolate).
-
 The leaves of *E.
@@
 \
-*Response*: All references to J have been removed from the manuscript (see repsonse below).
+*Response*: All references to J have been removed from the manuscript (see response below).
 
@@
 *Response*: After careful thought we have decided to remove parameters derived from AC~i~ curves from the manuscript.
-Thus, references to J~max~ and V~cmax have been removed.
-The presentation and discussion of the diffusion pathways, including Ci and Cc~ and draw down remain (Figure 7).
-Our main goal is to show that sun and shade leaves differed in their photosynthetic capacity, regardless of whether these differences are due to biochemistry or leaf ultrastructure or both (which we cannot satisfactorily address with the available data).
+Thus, references to J~max~ and V~cmax~ have been removed.
+The presentation and discussion of the diffusion pathways, including C~i~ and C~c~ and draw down remain (Figure 7).
+Our main goal is to show that sun and shade leaves differed in their photosynthetic capacity, regardless of whether these differences are due to biochemistry or leaf ultra-structure or both (which we cannot satisfactorily address with the available data).
 As such, we have decided to instead present photosynthetic rates at saturating CO~2~ concentrations and light (*A*~max~) for the different leaf types.
-This provides a simple metric of photosynthetic capacity that is well supported in the literature, and shows that shade leaves have a lower photosynthetic capacity, irrespective of the relative contribution of biochemical and biophysical contraints.
+This provides a simple metric of photosynthetic capacity that is well supported in the literature, and shows that shade leaves have a lower photosynthetic capacity, irrespective of the relative contribution of biochemical and biophysical constraints.
 We have adjusted the Methods (which is now shorter), as well as Figure 2 which now shows the relationship between *A*~max~ and leaf *N*~a~.
@@
 We acknowledge that AC~i~ response curves were initially conceived with the notion that *g*~m~ was large enough to be ignored, which we now know is untrue.
-We beleive there are some practical concerns when fitting A-Cc curves, and we hope this data set can eventually help to address this issue, but it is beyond the scope of this paper.
+We believe there are some practical concerns when fitting A-Cc curves, and we hope this data set can eventually help to address this issue, but it is beyond the scope of this paper.
 

### 2016-09-22 PCE_paper/manuscript_gmes.Rmd
everything ready except for last color palette check
fixed answer to point 9 reviewer 2, and also expanded text in ms on this point

@@
 # Set Flag to include figures in text or not.
-includeFigs <- TRUE
+includeFigs <- FALSE
 ```
@@
 @prentice2014balancing; @lin2015optimal].
+For the entire canopy to exhibit optimal regulation of stomatal conductance, that is, to maximize photosynthesis minus transpiration cost [@cowan1977stomatal], sun and shade leaves should have equal ITE - provided environmental drivers are similar [@buckley2014].
+In our study, higher *g*~s~ in shade leaves may in part be explained by higher *K*~L~, which effectively reduces the cost of water transport, allowing more profligate water use in shade leaves.
 Consistent with our results, Buckley et al.
-[-@buckley2014] found that shade leaves exhibited higher *g*~s~ than expected based on their environment.
-In our study, higher *g*~s~ in shade leaves may in part be explained by higher *K*~L~, which effectively reduces the cost of water transport, allowing more profligate water use in shade leaves.
+[-@buckley2014] found that shade leaves in a grapevine canopy exhibited higher *g*~s~ than expected based on their environment and comparison to sun leaves.
+Further work is needed to confirm the generality of this observation.
 

### 2016-09-22 PCE_paper/manuscript_gmes.Rmd
marks edits

--- 2016-09-12 PCE_paper/manuscript_gmes.Rmd
@@
 Light gradients within tree canopies play a major role in the distribution of plant resources that define the photosynthetic capacity of sun and shade leaves.
-However, biochemical and diffusional constraints on gas exchange in sun and shade leaves in response to light remain poorly quantified, but critical for predicting canopy carbon and water exchange.
+However, the biochemical and diffusional constraints on gas exchange in sun and shade leaves in response to light remain poorly quantified, but critical for predicting canopy carbon and water exchange.
 To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole-tree chambers.
@@
 Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index (ratio of leaf area to ground area), creating a steep light gradient from the canopy top to bottom [@monsi2005factor].
-Consequently, structural and functional properties of leaves within canopies are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance with canopy depth strongly affects rates of leaf photosynthesis (*A*~n~) [@evans1995carbon].
-To estimate whole canopy C gain it is thus necessary to account for the non-linear response of *A*~n~ to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
+Consequently, structural and functional properties of leaves within canopies are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance with canopy depth strongly affects rates of leaf net photosynthesis (*A*~n~) [@evans1995carbon].
+To calculate whole canopy C gain it is thus necessary to account for the non-linear response of *A*~n~ to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].
 
@@
 Changes in chlorophyll per unit N, chlorophyll a:b ratios, electron transport capacity per unit chlorophyll and ratios of electron transport capacity to Rubisco activity can also occur in response to changes in irradiance of the growth environment [@evans2001photosynthetic].
-Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems developed for high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
+Sun leaves frequently experience greater water limitations in the upper canopy, despite effective vascular systems that accommodate high radiation loads and transpiration [@sellin2008effects; @niinemets2012optimization].
 Optimal photosynthetic N investment in the upper canopy will be ineffective in enhancing *A*~n~ if water supply is insufficient [@niinemets2012optimization; @peltoniemi2012co]; thus leaf-specific  hydraulic conductance (*K*~L~) should also be higher in the upper canopy to supply sunlit leaves with sufficient water [@burgess2006hydraulic;@sellin2007effects ;@sellin2008effects;@hubbard2001stomatal].
@@
 Mesophyll conductance can also impose limitations on *A*~n~ that can be as large as *g*~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in *A*~n~ [@niinemets2007photosynthesis].
-Part of the variation in photosynthetic capacity between sun and shade leaves may arise from differences in *g*~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
+Differences in photosynthetic capacity between sun and shade leaves may arise, in part, from differences in *g*~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified.
 It is likely that leaf anatomical costs associated with minimizing the length and tortuosity of the *g*~m~ diffusion pathway are necessary to maintain the benefits of a high *g*~m~ [@hassiotou2009influence].
@@
 CO~2~, temperature or vapour pressure deficit) at timescales of minutes, possibly faster than *g*~s~ [@flexas2008mesophyll and references therein], yet the short term response to light availability is unclear.
-For example, *g*~m~ was found to be independent of light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid] and *Eucalyptus* seedlings [@douthe2011mesophyll; @douthe2012mesophyll].
+For example, *g*~m~ was found to be independent of incident light intensity in wheat leaves [@tazoe2009light] but was responsive to light in tobacco [@flexas2007rapid] and *Eucalyptus* seedlings [@douthe2011mesophyll; @douthe2012mesophyll].
 Anatomical parameters which regulate *g*~m~ with changing growth irradiance such as chloroplast surface area [@terashima2006irradiance] and mesophyll thickness [@boardman1977comparative; @terashima2001sun; @hanba2002effect] are also unlikely to adjust during transient fluctuations in light.
@@
 A detailed description of the WTC operation and design is available in Barton et al.
-[-@barton2010whole] and methods for this experiment in [@drake2016does].
+[-@barton2010whole] and methods for this experiment in Drake et al.
+[-@drake2016does].
 Coppiced *E.
-saligna* trees are present in the area between the whole-tree chambers, and these coppiced trees grew at rates similar to the trees in the chambers, thus providing ample shade.
+saligna* trees were present in the area between the whole-tree chambers, and these coppiced trees grew at rates similar to the trees in the chambers, thus providing ample shade.
 Six chambers were set to match outside ambient air temperatures (AT) while the remaining 6 experienced an elevated air temperature treatment of ambient +3&deg;C (ET, Figure `r airvarfig`).
@@
 For ET treatments, trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an leaf area of 6.2&pm;0.2 m^2^.
-Leaf area was calculated based on complete leaf counts and mean leaf size from a subsample [@drake2016does].
+Leaf area was calculated based on complete leaf counts and mean leaf size from a subsample from each tree [@drake2016does].
 
@@
 Shade leaves were always measured in the lower canopy, but leaves were sampled on subsequent higher branches across measurement campaigns to minimize confounding effects of leaf age.
-As shade leaves most likely developed more slowly this assured that older leaves in the bottom canopy were avoided.
+As shade leaves most likely developed more slowly, this assured that older leaves in the bottom canopy were avoided.
 
@@
 [-@tazoe2011using] and Evans & von Caemmerer [-@evans2013temperature].
-Leaf level gas exchange was measured with a standard (2 x 3 cm) leaf cuvette using a portable gas exchange system (LI-6400XT, LI-COR, Lincoln, NE, USA).
+Leaf level gas exchange was measured with a standard (2 x 3 cm) leaf cuvette and red/blue light source using a portable gas exchange system (LI-6400XT, LI-COR, Lincoln, NE, USA).
 This system was coupled with a tunable diode laser (TDL; TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of ^13^C isotope discrimination.
@@
 Leaf temperatures were controlled at the current AT or ET WTC air temperature, and measurements were made at ambient air humidity.
-Measurements were thus made across a range of air temperature and vapour pressure deficit (kPa) across the six campaigns.
+Measurements were thus made across a range of air temperatures and vapour pressure deficits (kPa) across the six campaigns.
 PPFD in the cuvette was set to match the individual light environment of each leaf type (explained above).
@@
 \
-As light intensity was increased for shade leaves before they were detached, the *E*~L~ measured at high light was used to to calculate *K*~L~ for all shade leaves.
+As light intensity was increased for shade leaves before they were detached for measurement of $\Psi_L$, the *E*~L~ measured at high light was used to calculate *K*~L~ for all shade leaves.
 
@@
 ##Leaf resource distribution
-Across six measurement campaigns over the 7 month period, PPFD was reduced on average by >75% in the shade (Figure `r parfig`).
+Across six measurement campaigns over the 7-month period, PPFD was reduced on average by >75% in the shade (Figure `r parfig`).
 Leaf *N*~a~ was approximately 20% higher in sun leaves compared to shade leaves (Table `r wtc3tab1`).
@@
 Leaf-specific hydraulic conductance (*K*~L~) was significantly higher in shade leaves (+58%) compared to sun leaves (Table `r wtc3tab1`).
-This was due to differences in transpiration rates (*E*~L~ , Table `r wtc3tab2`) between leaf types at high light, as neither leaf water potentials ($\Psi_{pd}$ and $\Psi_L$, Table `r wtc3tab1`) differed between leaf types.
+This was due to differences in transpiration rates (*E*~L~, Table `r wtc3tab2`) between leaf types at high light, as neither leaf water potentials ($\Psi_{pd}$ and $\Psi_L$, Table `r wtc3tab1`) differed between leaf types.
 No effect of the warming treatment was detected on  PPFD, $\Psi_{pd}$, $\Psi_L$, *K*~L~, *E*~L~, *N*~a~ or LMA either within or across leaf types (P > 0.05).
@@
 No effect of the warming treatment was detected on *A*~n~ of sun leaves measured at high light or shade leaves at either low or high light.
-Photosynthesis within leaf types and warming treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
+Rates of photosynthesis within leaf types and warming treatments were similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).
 
@@
 Photosynthesis was positively correlated with *g*~s~ in all leaves measured under high light conditions, however, *g*~s~ and *A*~n~ were not correlated in shade leaves under low light (Figure `r agmgsfig`a).
-In high light, *g*~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
+In high light, *g*~s~ of shade leaves increased compared to low light and was significantly greater than sun leaves, pooled across all measurement dates (Figure `r gebarfig`a).
 No effect of the warming treatment was detected on *g*~s~ within or across leaf types.
-Stomatal conductance within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get2fig`b).
+Stomatal conductances within leaf types and treatments were similar through time and across the range of leaf temperatures measured (Figure `r get2fig`b).
 
@@
 ITE in sun leaves was reduced in the warming treatment compared to ambient, but no warming effect was detected in shade leaves measured at low or high light (Table `r wtc3tab2`).
-The mean estimated *g*~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves measured at low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.
+The mean estimated *g*~1~ for sun leaves was 1.51&pm;0.11.
+Mean *g*~1~ was 2.59&pm;0.12 for shade leaves measured at low light and 2.74&pm;0.04 when measured at high light.
 For all leaf types and light treatments there was a strong response of ITE to VPD and individual data points broadly corresponded to the fitted response curves from the optimal ITE model with a specified *g*~1~ value for each leaf type and treatment (Figure `r waterusefig`a).
-Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).
+Within leaf types and light treatments, the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a).
 
@@
 Although it is well-known that shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.
-These findings offer new insights into how aspects of leaf gas exchange physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake and water use in future studies.
+These findings offer new insights into how aspects of leaf gas exchange physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake and water use.
 Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects of a +3 &deg;C warming treatment were detected on leaf physiology, consistent with temperature acclimation demonstrated in this experiment [@aspinwall2016convergent].
@@
 Leaf mass per area (LMA), however, was not different between sun and shade leaves, consistent with small differences reported in a clonal *Eucalyptus* plantation [@nouvellon2010], but in contrast to data for many other species [@niinemets2015].
-Similar LMA in sun and shade leaves could be the result of leaf formation under comparable light conditions or possible unmeasured differences in total non-structural carbohydrates contents between leaf types that masked differences in leaf thickness or density.
+Similar LMA in sun and shade leaves could be the result of initial leaf formation and expansion under comparable light conditions and subsequent shade arising during canopy development in this evergreen species.
+Unmeasured differences in total non-structural carbohydrates contents between leaf types may also have masked differences in leaf thickness or density.
 For example, sun leaves of *Abies amabilis* foliage that became shaded exhibited physiology similar to shaded leaves, while retaining the morphology of sun leaves [@brooks1996effects].
@@
 \
-Photosynthesis is also limited by the ability to supply water to the upper canopy.
+Photosynthesis is also limited by the supply of water to the upper canopy.
 Ultimately, the ability of tree hydraulic architecture to supply water to foliage across increasing pathlengths affects productivity and survival [@sellin2008effects].
@@
 [-@peltoniemi2012co] theorized that optimal N distribution will be proportional to light distribution only if *K*~L~ is also optimally distributed.
-We found higher *K*~L~ in shade leaves compared to sun leaves, countering the prediction by Peltoniemi et al.
+We found higher *K*~L~ in shade leaves compared to sun leaves, contrary to the prediction by Peltoniemi et al.
 [-@peltoniemi2012co], but consistent with higher *g*~s~ attained by shade leaves in high light conditions.
@@
 Consistently lower leaf $\delta$^13^C in shade leaves also suggests that observed higher *C*~i~ and *C*~c~ in shade leaves was prevalent long term (Figure `r deltafig`) [@marshall2007sources], despite the observation that *g*~m~ accounted for nearly 70% of the variation in $\Delta$~o~.
-Higher *g*~s~ in shade leaves relative to the differences in *A*~n~, appears to result in inefficient water use as more water is used per unit photosynthesis.
-As whole canopy C gain integrates the efficiency of all leaves, this raises the question of why shade leaves maintain a lower water-use efficiency compared to sun leaves.
+Higher *g*~s~ in shade leaves relative to the differences in *A*~n~, appears to result in comparatively inefficient water use as more water is used per unit photosynthesis.
+As whole canopy C gain integrates the efficiency of all leaves, this raises the question of why shade leaves exhibit a lower water-use efficiency compared to sun leaves.
 
@@
 \
-It is also possible that higher *g*~s~ in shade leaves is a strategy to take advantage of high light quickly in shade leaves, avoiding the delay with stomatal opening [@tausz2005dynamic;@vicomanz2011].
+It is also possible that higher *g*~s~ in shade leaves is a strategy to take advantage transiet periods of high light quickly in shade leaves, avoiding the delay associated with stomatal opening [@tausz2005dynamic;@vicomanz2011].
 Our results support this hypothesis, as shade leaves increased *A*~n~ to values equal or even outperforming sun leaves when subjected to a brief period of identical high-light intensity.
@@
 It has also been proposed that *g*~m~ and *K*~L~ should be coordinated, as optimizing hydraulic supply allows both *g*~s~ and *g*~m~ to enhance *A*~n~ [@flexas2013leaf].
-Here, higher *K*~l~ in shade leaves under high light lends support to the potential coordination *K*~l~ and *g*~m~ and also to the association of higher *A*~n~ with higher *K*~l~ when light becomes saturating reported by @franks2006higher.
+Here, higher *K*~l~ in shade leaves under high light lends support to the potential coordination *K*~l~ and *g*~m~ and also to the association of higher *A*~n~ with higher *K*~l~ when light becomes saturating, as reported by @franks2006higher.
 
@@
 Here we show that dynamic physiological responses of shade leaves to altered light environments have important implications for up-scaling leaf level measurements to the canopy.
-Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves that determines C gain.
+Although resource allocation constrains leaf photosynthetic capacity, it is the physiological behaviour of individual leaves that determines C gain.
 These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have implications for process-based models that predict canopy C gain from rates of leaf photosynthesis.

### 2016-09-22 PCE_paper/review_response.Rmd

@@
 \
-*Response*: The work of Peltoniemi et al.
-(2012) demonstrated that similar WUE throughout the canopy is only achieved if the hydraulic conductance (*K*~l~ ) is proportional to the PPFD gradient (in the same way that leaf nitrogen should be proportional to PPFD).
-They also reasoned that if shade leaves have a higher hydraulic conductance (perhaps due to a shorter path length), the optimal response will be to have a higher Ci (and thus lower WUE) in shade leaves compared to sun leaves.
-Since we fixed a mistake in the calculation of *K*~l~, the results on WUE differences between sun and shade leaves are indeed consistent with the theoretical predictions of Peltoniemi.
-We have expanded on this interesting point in the discussion (expanding on the section that previously included lines 431-434).
+*Response*: If the cost of water (sensu Cowan & Farquhar 1977) is constant throughout the canopy, then we would expect similar WUE in sun and shade leaves.
+Since we found that shade leaves have lower WUE, this could indicate a lower cost of water.
+This is consistent with the observation that hydraulic conductance was higher in shade leaves.
+We have expanded the section (previously lines 431-434) to read:
+
+"For the entire canopy to exhibit optimal regulation of stomatal conductance, that is, to maximize photosynthesis minus transpiration cost (Cowan and Farquhar 1977), sun and shade leaves should have equal ITE - provided environmental drivers are similar (Buckley et al.
+2014).
+In our study, higher *g*~s~ in shade leaves may in part be explained by higher *K*~L~, which effectively reduces the cost of water transport, allowing more profligate water use in shade leaves.
+Consistent with our results, Buckley et al.
+(2014) found that shade leaves in a grapevine canopy exhibited higher *g*~s~ than expected based on their environment and comparison to sun leaves.
+Further work is needed to confirm the generality of this observation."
 

### 2016-09-22 PCE_paper/manuscript_gmes.Rmd
colors finished

@@
 Each chamber was 9 m tall, which accommodated growth of trees for 15 months.
-A detailed description of the WTC operation and design is available in Barton et al.
-[-@barton2010whole] and methods for this experiment in Drake et al.
-[-@drake2016does].
+A detailed description of the WTC operation and design is available in @barton2010whole and methods for this experiment in @drake2016does.
 Coppiced *E.
@@
 \
-Concurrent gas exchange and C isotope discrimination measurements were conducted based on methods described in Tazoe et al.
-[-@tazoe2011using] and Evans & von Caemmerer [-@evans2013temperature].
+Concurrent gas exchange and C isotope discrimination measurements were conducted based on methods described in @tazoe2011using and @evans2013temperature.
 Leaf level gas exchange was measured with a standard (2 x 3 cm) leaf cuvette and red/blue light source using a portable gas exchange system (LI-6400XT, LI-COR, Lincoln, NE, USA).
@@
 \
-Mesophyll conductance was calculated from carbon isotope discrimination with equations and fractionation factors as presented in Evans & von Caemmerer [-@evans2013temperature], including the ternary corrections proposed by Farquhar & Cernusak [-@farquhar2012ternary], such that:
+Mesophyll conductance was calculated from carbon isotope discrimination with equations and fractionation factors as presented in @evans2013temperature, including the ternary corrections proposed by @farquhar2012ternary, such that:
 
@@
 where $\Delta$~o~ is the observed discrimination and $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if *C*~i~ = *C*~c~, *g*~m~, respiration and photorespiration, respectively.
-For this study, the CO~2~ compensation point ($\Gamma_{*}$) and respiration during the day (R~d~) parameters originally derived for tobacco plants [@von1994kinetics] were replaced with parameters derived for *Eucalyptus globulus* from Crous *et al*.
-[-@crous2012light] when calculating *g*~m~.
+*C*~i~ and  *C*~c~ are the intercellular and chloroplast CO~2~ partial pressures.
+For this study, the CO~2~ compensation point ($\Gamma_{*}$) and respiration during the day (R~d~) parameters originally derived for tobacco plants [@von1994kinetics] were replaced with parameters derived for *Eucalyptus globulus* from @crous2012light when calculating *g*~m~.
 Full descriptions of the carbon isotope discrimination equations, with ternary corrections, are presented in Supporting Information Methods S1.
 The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck were compared as a function of *C*~i~/C~a~.
-Once *g*~m~ was calculated, the chloroplast CO~2~ partial pressure (*C*~c~) and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were estimated from the relationship between *g*~m~ and leaf photosynthesis rate (*A*~n~) by:
+Once *g*~m~ was calculated, *C*~c~ and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were estimated from the relationship between *g*~m~ and leaf photosynthesis rate (*A*~n~) by:
 
@@
 The interaction was never significant, so we only report results of the tests of the main effects.
-Explained variance (*R*^2^) of mixed models was computed as in Nakagawa & Schielzeth [-@nakagawa2013general], in which the marginal *R*^2^ represents variance explained by fixed factors and the conditional *R*^2^ by both fixed and random factors.
+Explained variance (*R*^2^) of mixed models was computed as in @nakagawa2013general, in which the marginal *R*^2^ represents variance explained by fixed factors and the conditional *R*^2^ by both fixed and random factors.
 Confidence intervals (95 %) of linear mixed-effects models were generated using bootstrapping methods with 999 simulations, using the 'bootMer' function in the 'lme4' package [@bates2013lme4].
@@
 Ultimately, the ability of tree hydraulic architecture to supply water to foliage across increasing pathlengths affects productivity and survival [@sellin2008effects].
-Using a two-leaf model, Peltoniemi et al.
-[-@peltoniemi2012co] theorized that optimal N distribution will be proportional to light distribution only if *K*~L~ is also optimally distributed.
-We found higher *K*~L~ in shade leaves compared to sun leaves, contrary to the prediction by Peltoniemi et al.
-[-@peltoniemi2012co], but consistent with higher *g*~s~ attained by shade leaves in high light conditions.
+Using a two-leaf model, @peltoniemi2012co theorized that optimal N distribution will be proportional to light distribution only if *K*~L~ is also optimally distributed.
+We found higher *K*~L~ in shade leaves compared to sun leaves, contrary to the prediction by @peltoniemi2012co, but consistent with higher *g*~s~ attained by shade leaves in high light conditions.
 This finding is consistent with the idea that *g*~s~ is regulated, at least in part, by *K*~L~ [@hubbard2001stomatal].
@@
 In our study, higher *g*~s~ in shade leaves may in part be explained by higher *K*~L~, which effectively reduces the cost of water transport, allowing more profligate water use in shade leaves.
-Consistent with our results, Buckley et al.
-[-@buckley2014] found that shade leaves in a grapevine canopy exhibited higher *g*~s~ than expected based on their environment and comparison to sun leaves.
+Consistent with our results, @buckley2014 found that shade leaves in a grapevine canopy exhibited higher *g*~s~ than expected based on their environment and comparison to sun leaves.
 Further work is needed to confirm the generality of this observation.
@@
 Photosynthetic capacity (*A*~max~) was positively correlated with mean leaf N per area (*N*~a~) for sun and shade leaves grown under elevated (ET) and ambient (AT) temperature treatments.
-Leaf *N*~a~ is the mean of each leaf type sampled across gas exchange campaigns in each whole-tree chamber.
+Leaf *N*~a~ is the mean of each leaf type sampled across six gas exchange campaigns in each whole-tree chamber.
 *A*~max~ was measured once on all trees for each leaf type, in February 2014, at 25°C and at both saturating light and [CO~2~].

### 2016-09-22 master_scripts/cicc_boxplots.R

@@ -25,7 +25,7 @@ leaforder <- c("Shade", "Sunfleck", "Sun")
 gasex_agg$leaflight <- factor(gasex_agg$leaflight, levels = leaforder)
 
 
-cols <- c(shacol2, lightscol2,suncol2)
+cols <- c(shacol, lightscol,suncol)
 
 
 # windows(12,6)

### 2016-09-22 master_scripts/gmes_shadeleaves_figure.R

@@ -35,7 +35,8 @@ shadeleaf <- merge(shadat2, fleckdat2, by=c("chamber", "temp","Month"))
 
 par(mar=c(4,4,1,1), las=1, cex.axis=.8, cex.lab=1, mgp=c(2.5,1,0))
 plot(fleck_gm~shade_gm, data=shadeleaf, ylim=c(0,.3), xlim=c(0,.3), xlab=shagmlab, ylab=fleckgmlab, 
-     pch=c(16, 17)[shadeleaf$temp], cex=1.25)
+     pch=c(21, 24)[shadeleaf$temp], cex=1.25, lwd=1.25)
 abline(0,1, lty=2)
-legend("topleft", c("AT", "ET"), pch=c(16,17), inset = 0.01, bty='n', title="Shade Leaves", cex=1, pt.cex=1.25)
+legend("topleft", c("AT", "ET"), pch=c(21,24), inset = 0.01, bty='n', title="Shade Leaves", cex=1, 
+       pt.cex=1.25, pt.lwd=1.25)
 

### 2016-09-22 master_scripts/physiology_boxplots.R

@@ -23,7 +23,7 @@ gasex_agg$leaflight <- factor(gasex_agg$leaflight, levels = leaforder)
 
 #lightson <- gasex_agg[gasex_agg$leaflight != "sun-high",]
 
-cols <- c(shacol2, lightscol2,suncol2)
+cols <- c(shacol, lightscol, suncol)
 
 ####PLOTTING boxplots--------------------------------------------------------------------------------
 

### 2016-09-22 master_scripts/wateruse.R

@@ -76,7 +76,7 @@ plot(ite~VpdL, data=ite_sunsha, subset=leaflight=="sun-high",  col=suncol, xlab=
   
   text(x=0, y=15.5 ,"(a)", cex=1)
   
-  legend("topright", alllab, pch=c(16,16,16,16,17), col=c(suncol, shacol, lightscol2, "black", "black"),lty=c(-1,-1,-1,1,2),
+  legend("topright", alllab, pch=c(16,16,16,16,17), col=c(suncol, shacol, lightscol, "black", "black"),lty=c(-1,-1,-1,1,2),
          bty='n',cex=.8)
 
 ##panel2

### 2016-09-22 PCE_paper/manuscript_gmes.Rmd
Amax figure edits

@@
 Photosynthetic capacity (*A*~max~) was positively correlated with mean leaf N per area (*N*~a~) for sun and shade leaves grown under elevated (ET) and ambient (AT) temperature treatments.
-Leaf *N*~a~ is the mean of each leaf type sampled across six gas exchange campaigns in each whole-tree chamber.
-*A*~max~ was measured once on all trees for each leaf type, in February 2014, at 25°C and at both saturating light and [CO~2~].
+Leaf *N*~a~ is the mean of each leaf type, in well-watered treatments, sampled across six gas exchange campaigns in each whole-tree chamber.
+*A*~max~ was measured once on all trees for each leaf type, before the initiation of the drought treatment, at 25°C and at both saturating light and [CO~2~].
 The dashed line represents the significant linear model fit for all leaves, with a marginal and conditional *R*^2^ of 0.59 and 0.62.

### 2016-09-22 master_scripts/amax_nitro.R

@@ -69,7 +69,7 @@ canopy_chem <- merge(leaf_mass[leaf_mass$wp_type=="mid",c(1, 3:4, 7:8)], leaf_ch
   
   canopy_chem2 <- canopy_chem[canopy_chem$drydown=="control",]
   canopy_chem3 <- canopy_chem2[, c("chamber", "Month", "leaf", "leafN_area")]
-  
+
 ##Amax and Narea dataset  (with stats for results)----------------------------------------------------------------
 Nagg <- summaryBy(leafN_area ~ chamber + leaf, data=canopy_chem, FUN=mean, keep.names = TRUE)
 
@@ -90,9 +90,10 @@ nitroamax_mod <- lm(Photo~leafN_area, data=N_amax)
 ##Amax vs Narea
 #png(filename = "figpic/amax.png", width = 11, height = 8.5, units = "in", res= 400)
 
+#xlim=c(1.5,3.5),ylim=c(15,35),
 
 par(mar=c(4,4,1,1), las=1, cex.axis=.8, cex.lab=1, mgp=c(2.5,1,0))
-plot(Photo~leafN_area, data=N_amax, col=leafcol3[as.factor(leaf)],xlim=c(1.5,3.5),ylim=c(15,35),
+plot(Photo~leafN_area, data=N_amax, col=leafcol3[as.factor(leaf)],xlim=c(01.5,3.5),ylim=c(15,35),
        pch=c(16, 17)[pch=N_amax$temp],xlab=narealab, ylab=amaxlab, cex=1.25)
 
 ablineclip(nitroamax_mod, x1=min(N_amax$leafN_area), x2=max(N_amax$leafN_area), lwd=2, lty=3)

### 2016-09-22 PCE_paper/review_response.Rmd
colors finished
edits to supp material, review response and supp material are uploaded to pce

@@
 We have darkened several of the original colors, but this choice of color palette is also used to ensure accessibility for color-blind readers.
+We replaced the filled black circles in Figure 6 with open symbols.
 
@@
 \
-*Response*: If the cost of water (sensu Cowan & Farquhar 1977) is constant throughout the canopy, then we would expect similar WUE in sun and shade leaves.
+*Response*: If the cost of water is constant throughout the canopy, then we would expect similar WUE in sun and shade leaves (see Cowan & Farquhar 1977).
 Since we found that shade leaves have lower WUE, this could indicate a lower cost of water.

### 2016-09-22 PCE_paper/manuscript_gmes.Rmd
edits to supp material, review response and supp material are uploaded to pce

@@
 Similar LMA in sun and shade leaves could be the result of initial leaf formation and expansion under comparable light conditions and subsequent shade arising during canopy development in this evergreen species.
-Unmeasured differences in total non-structural carbohydrates contents between leaf types may also have masked differences in leaf thickness or density.
+Unmeasured differences in total non-structural carbohydrate contents between leaf types may also have masked differences in leaf thickness or density.
 For example, sun leaves of *Abies amabilis* foliage that became shaded exhibited physiology similar to shaded leaves, while retaining the morphology of sun leaves [@brooks1996effects].

### 2016-09-22 PCE_paper/manuscript_gmes_SuppInfo.Rmd
manuscript edits
edits to supp material, review response and supp material are uploaded to pce

--- 2016-09-02 gm_pce/manuscript_gmes_SuppInfo.Rmd
@@
 ```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
-source("gm_pce/manuscript_preload.R")
+source("PCE_paper/manuscript_preload.R")
 ```
@@
 Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis ($\Delta_o$) and measured *C*~i~/*C*~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.
-The solid line represents the theoretical line for C3 plants from Evans et al.
-[-@evans1986carbon].
+The solid line represents the theoretical line for C3 plants from @evans1986carbon.
 \
@@
 Response of VPD (a), *g*~s~ (b) and *C*~a~-*C*~i~ to leaf temperature for sun leaves and shade leaves at low and high light.
-Shaded symbols represents each monthly measurement campaign.
+Shaded symbols represent each monthly measurement campaign.
 Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns.
@@
 where R~s~ is the isotopic ratio of the sample and R~stnd~ is the isotopic ratio of the standard Vienna Pee Dee Belemnite (VPDP).
-Next, the observed discrimination ($\Delta$~o~) is calculated from Evans et al.
-[-@evans1986carbon]:
+Next, the observed discrimination ($\Delta$~o~) is calculated from @evans1986carbon:
 
@@
 \
-Second, C isotope discrimination during C~3~ photosynthesis ($\Delta$) is the resultant discrimination from CO~2~ diffusion from the atmosphere to the site of carboxylation, consisting of a series of fractionation steps described in Evans et al.
-[-@evans1986carbon].
-In this experiment, a modified form of the equation presented in Evans & von Caemmerer [-@evans2013temperature] with ternary effect corrections by Farquhar & Cernusak [-@farquhar2012ternary] was used such that:
+Second, C isotope discrimination during C~3~ photosynthesis ($\Delta$) is the resultant discrimination from CO~2~ diffusion from the atmosphere to the site of carboxylation, consisting of a series of fractionation steps described in @evans1986carbon.
+In this experiment, a modified form of the equation presented in @evans2013temperature with ternary effect corrections by @farquhar2012ternary was used such that:
 
@@
 where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ is the fractionation from boundary layer diffusion (2.9&permil;) and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon].
-We calculate discrimination during day respiration (e) following Wingate et al.
-[-@wingate2007variations] as e = $\delta^13C_{tank}$ - $\delta^13C_{atmosphere}$.
+We calculate discrimination during day respiration (e) following @wingate2007variations as e = $\delta^13C_{tank}$ - $\delta^13C_{atmosphere}$.
 In this study, $\delta^13C_{tank}$ was -4 to -6&permil; and $\delta^13C_{atmosphere}$ was -12&permil;.
@@
 *C*~a~ and *C*~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~).
-In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulus* from Crous et al.
-[-@crous2012light].
+In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulus* from @crous2012light.
 The ternary effect corrections (t) are described by:
@@
 
-Here, *g*~m~ is expressed in units of mol m^-2^ s^-1^ bar^-1^ to be consistent with Evans & von Caemmerer [-@evans2013temperature].
+Here, *g*~m~ is expressed in units of mol m^-2^ s^-1^ bar^-1^ to be consistent with @evans2013temperature.
 As the pressure term is nearly equal to unity, suitable comparisons between *g*~m~ and *g*~s~ can be made.

### 2016-09-27 PCE_paper/manuscript_gmes.Rmd
getting data ready for figshare, small edits to manuscript

@@
 To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole-tree chambers.
-Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic capacity (*A*~max~, i.e.
-*A*~n~ at saturating [CO~2~]), but *g*~s~ was similar.
+Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic capacity (*A*~max~), but *g*~s~ was similar.
 When light intensity was temporarily increased for shade leaves to match that of sun leaves, both *g*~s~ and *g*~m~ increased, and *A*~n~ increased to values greater than sun leaves.
@@
 Leaf mass per area (LMA) was not different between leaf types (Table `r wtc3tab1`).
-Leaf-specific hydraulic conductance (*K*~L~) was significantly higher in shade leaves (+58%) compared to sun leaves (Table `r wtc3tab1`).
+Leaf-specific hydraulic conductance (*K*~L~) was significantly higher in shade leaves (+44%) compared to sun leaves (Table `r wtc3tab1`).
 This was due to differences in transpiration rates (*E*~L~, Table `r wtc3tab2`) between leaf types at high light, as neither leaf water potentials ($\Psi_{pd}$ and $\Psi_L$, Table `r wtc3tab1`) differed between leaf types.

### 2016-09-27 master_scripts/amax_nitro.R

@@ -54,6 +54,8 @@ acishade <- merge(acishade, treatments)
 amax_dat <- rbind(shademax, sunmax)  
   amax_dat <- merge(amax_dat, treatments)
   
+write.csv(amax_dat, "calculated_data/amax_raw.csv", row.names=FALSE)
+  
 ###Narea data------------------------------------------------------------------------------------------------------
 leaf_chem <- read.csv("raw data/leaf_chem.csv")
 leaf_mass <- read.csv("raw data/leaf_data.csv")

### 2016-09-27 master_scripts/newleafK.R

@@ -33,27 +33,22 @@ licor_raw <- read.csv("raw data/gm_licor.csv") #this is base raw data
   transp <- add_Month(transp)
   transp <- wptrt_func(transp)
   
-  #means by campaign
-  transp_agg <- summaryBy(Trmmol+Cond ~ Month +chamber+leaf+light+drydown, data= transp, FUN=c(mean), keep.names=TRUE)
+#means by campaign
+transp_agg <- summaryBy(Trmmol+Cond ~ Month +chamber+leaf+light+drydown, data= transp, FUN=c(mean), keep.names=TRUE)
     transp_agg$leaflight <- paste(transp_agg$leaf, transp_agg$light, sep="-")
-###remove drought treatment from raw licor data
+
+    ###remove drought treatment from raw licor data
 E_dat <- transp_agg[transp_agg$drydown == "control",]
   E_dat <- droplevels(E_dat)
-  
-##prelim plot  
-  E_dat2 <- subset(E_dat,  chamber != "ch10" | Month != "Mar")
-  plot(E_dat2[E_dat2$leaflight=="shade-high","Trmmol"] ~ E_dat2[E_dat2$leaflight=="shade-low","Trmmol"], data=E_dat2,
-       xlab= "Trmmol Shade-Low", ylab = "Trmmol Shade-High", xlim=c(0,8), ylim=c(0,8))
-  abline(0,1)
 
-  ##seperate files by leaflight of shade leaves:
-E_norm <- subset(E_dat, transp_agg$leaflight !=  "shade-high")
+##seperate files by leaflight of shade leaves:
+E_norm <- subset(E_dat, E_dat$leaflight !=  "shade-high")
   E_norm$Month <- as.factor(E_norm$Month)
   E_norm$chamber<- as.factor(E_norm$chamber)
   E_norm$leaf<- as.factor(E_norm$leaf)
 
-E_fleck <- subset(E_dat, transp_agg$leaflight !=  "shade-low")
-  E_fleck$Month <- as.factor(E_fleck$Month)
+E_fleck <- subset(E_dat, E_dat$leaflight !=  "shade-low")
+E_fleck$Month <- as.factor(E_fleck$Month)
   E_fleck$chamber<- as.factor(E_fleck$chamber)
   E_fleck$leaf<- as.factor(E_fleck$leaf)
   
@@ -68,21 +63,8 @@ leafK_fleck <- merge(E_fleck ,WP)
[…]

### 2016-10-24 PCE_paper/manuscript_gmes.Rmd
fixed mistakes from proofs

@@
 \
-The distribution of resources required for *A*~n~, including leaf nitrogen (N) and supply of water, are also partially defined by canopy light gradients.
+The distribution of resources required for *A*~n~, including leaf nitrogen (N) and supply of water, is also partially defined by canopy light gradients.
 As *A*~n~ has a saturating response with light and maximum rates depend, in part, on N-rich photosynthetic machinery, it has been argued that leaf N should be proportional to PPFD along the canopy light gradient to maximize canopy C gain at a given total canopy N [@field1983allocating; @field1986photosynthesis; @peltoniemi2012co; @buckley2013does].
@@
 It has also been proposed that *g*~m~ and *K*~L~ should be coordinated, as optimizing hydraulic supply allows both *g*~s~ and *g*~m~ to enhance *A*~n~ [@flexas2013leaf].
-Here, higher *K*~l~ in shade leaves under high light lends support to the potential coordination *K*~l~ and *g*~m~ and also to the association of higher *A*~n~ with higher *K*~l~ when light becomes saturating, as reported by @franks2006higher.
+Here, higher *K*~L~ in shade leaves under high light lends support to the potential coordination *K*~L~ and *g*~m~ and also to the association of higher *A*~n~ with higher *K*~L~ when light becomes saturating, as reported by @franks2006higher.
 
@@
 If the valuation of sunflecks as a C resource is large enough, then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan or across the entire canopy when considering both sun and shade leaf types [@vicomanz2011].
-For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to accommodate the cost of increased water use in the shade.
+For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PPFD may be large enough to accommodate the cost of increased water use in the shade.
 However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies.

### 2016-10-24 master_scripts/plot_objects.R

@@ -52,8 +52,8 @@ cicalab <- "Ci/Ca"
 cicalab2 <- (expression(italic(C)[i]/italic(C)[a]))
 
 drawdownlab <- expression(italic(C)[i]-italic(C)[c]~~(mu*mol~mol^-1))
-drawdownlab2 <- expression(italic(C)[a]-italic(C)[i]~~(mu*mol~mol^-1))
-drawdownlab3 <- expression(italic(C)[i]-italic(C)[c]~~(mu*mol~mol^-1))
+drawdownlab2 <- expression(italic(C)[a]-italic(C)[i]~~(paste(mu,bar,sep="")))
+drawdownlab3 <- expression(italic(C)[i]-italic(C)[c]~~(paste(mu,bar,sep="")))
 
 relparlab <- expression(l[shade]:l[sun])
 relnitrolab <- expression(N[shade]:N[sun])
